### 2020-02-05 v-research_experiment-alpha.tex
experiment alpha - first commit

\documentclass{article}
\usepackage[gen]{eurosym}
\usepackage{tabularx}
\usepackage{epigraph}
\usepackage{textcomp}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{csquotes}% Recommended
\usepackage{rotating}
\usepackage[style=numeric]{biblatex}
% \bibliography{<mybibfile>}% ONLY selects .bib file; syntax for version <= 1.1b
\addbibresource{literature.bib}% Syntax for version >= 1.2

\usepackage{graphicx}
\graphicspath{{./00-images/}}

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{}
\chead{}
\rhead{}
\lfoot{}
\cfoot{\thepage}
\rfoot{Property of V-Research S.r.l.}
\renewcommand\headrulewidth{0pt}
\renewcommand\footrulewidth{0pt}

\usepackage[colorinlistoftodos]{todonotes}

\newcommand{\Paragraph}[1]{\smallskip\noindent{\bf #1.}}
\newcommand{\fixnote}[2]{\textbf{\color{red}{NOTE}}\footnote{{\bf #1:} #2}}
\newcommand{\fix}[2]{{\color{red} #2}}
% \renewcommand{\fixnote}[2]{}
% \renewcommand{\fix}[2]{}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{corollary}
\newtheorem{corollary}{Corollary}[section]
\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}[section]

\usepackage{authblk}
\date{}                     %% if you don't need date to appear
\setcounter{Maxaffil}{0}
\renewcommand\Affilfont{\itshape\small}

% for fancy table with rotated column titles
\usepackage{adjustbox}
\newcolumntype{R}[2]{%
    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
    l%
    <{\egroup}%
}
\newcommand*\rot{\multicolumn{1}{R{45}{0.01em}}}% no optional argument here, please!
\newcommand*\rota{\multicolumn{1}{R{90}{0.01em}}}% no optional argument here, please!
\def\hb{\hbox to 10.7 cm{}}

% circles
\usepackage{wasysym}
\newcommand{\Tdot}{$\CIRCLE$}
\newcommand{\Thdot}{$\LEFTcircle$}
\newcommand{\Twdot}{$\Circle$}

% formulas
\newcommand{\kframe}{K}
\newcommand{\possibleworlds}{G}
\newcommand{\modalrelation}{R}
\newcommand{\actualworld}{\omega*}
\newcommand{\world}{\omega}
\newcommand{\World}{\Omega}
\newcommand{\kmodel}{M}
\newcommand{\interpretation}{\sigma}
\newcommand{\knowledge}[1]{\mathbb{K}_{#1}}
\newcommand{\knows}[2]{K_{#1}#2}
\newcommand{\belief}[1]{\mathbb{B}_{#1}}
\newcommand{\believe}[2]{B_{#1}#2}

\makeindex

\begin{document}

\title{On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}}
\author[1]{Francesco Beltramini}
\author[1]{Marco Rocchetto}
\author[2]{Luca Vigan\`o}
\affil[1]{V-Research, Verona, Italy}
\affil[2]{King's College London, UK}

\maketitle

\begin{abstract}
	The objective of this research is to develop of a theory that defines
	(all and only) the possible insecurity and security configurations of
	any abstract system. The theory is structured upon other theories that 
	defines how a component of a system can be abstracted into an agent, defining how
	agents can be formalized (both syntactically and semantically) to
	describe an abstract system, such as a graph. Some of these theories
	(e.g. used for the semantic
	definition of the abstract system) are the epistemological definition
	of knowledge, the Belief-Desire-Intent and the Assertion-Belief-Fact 
	framework of reference, 
	mereology, and topological structure. We argue that a mereology is the
	most appropriate abstract underlying structure, do to its generality,
	for defining the expressiveness of the system abstraction.  Furthermore, a
	mereology allows us to define an ontology rather than a taxonomy.  We
	also correlate different abstractions of the system to the TRL and the
	engineering V-model.

	We implemented a formal theory (of axioms) of a mereotopology, and of
	the region connection calculus (RCC3 and RCC5) in a Python program that
	uses the Z3 SMT solver. The results show that a single component (i.e.
	agent) of an abstract system has a definite number of  different
	insecurity configurations (e.g. 53 using RCC5 over a topological
	structure) and only 1 secure (i.e. expected) configurations. The
	configurations are reported as models satisfying the abstract system
	semantics.

	We considered the philosophical definition of truth behind our
	approach, rejecting ``proof'' by induction from partial empirical evidences.
	Our theory can be applied to system engineering and we show a concrete
	application of our theory to the risk assessment of an ad-hoc system.
	Finally, we provide a number of ideas to support the engineering of
	secure systems (e.g. purely cyber or cyber-physical).
\end{abstract}
\newpage

\section{Introduction}\label{sec:intro}
\epigraph{Humanum est errare}{{\itshape Seneca the Elder}}
The European Commission states in\autocite{EU2019market} that: ``Cybersecurity
is one of the priority areas [\ldots] of the Commission initiative on ICT
Standards, which is part of the Digitising European
Industry\autocite{EU2019standard} strategy launched on 19 April 2016.  The aim
is to identify the essential ICT standards and present measures to accelerate
their development in support of digital innovations across the economy''. The
same document (i.e.\autocite{EU2019market}) states that ``The EU will invest up
to \euro450 million [\ldots], under its research and innovation programme
Horizon 2020''. The EU, in 2016 published a press release\autocite{EU2016press}
in which they present a strategy to invest \euro1.8 \emph{billion} to
``increase measures to address cyber threats''. The EU is not the only investor
in cybersecurity, most of the developed countries and several companies are
investing enormous amount of money towards various aspects of cybersecurity
(e.g. The US vulnerability databases\autocite{NIST2020NVD} maintained by the
National Institute of Standards and Technologies, i.e. NIST, of the US
Department of Commerce).

The cybersecurity industry is growing fast, e.g. as reported
in\autocite{Nasdaq2018market}. For example, in\autocite{Forbes2017market},
published by the Forbes, is stated that \euro5.3 billion of funding where
poured by venture capitalist into cybersecurity companies in 2018. The Forbes,
in the same article, also highlights another peculiar (as seemly contradictory)
trend: ``[\ldots] during the same time period, the number of cybersecurity
breaches increased exponentially''. The data reported by the NIST 
through the official CPE (Common Platform Enumeration) Dictionary Statistics 
on the NVD websites in \autocite{NIST2020CPEstatistics}, show that in 2016 the
number of reported vulnerabilities reported where around $~6000$ while in 2019
the number of vulnerabilities was above $16000$.  The scientific community also
reports similar findings.  In fact, in\autocite{Herley2009so}, Cormac Herley
(Microsoft Research) shows how basic cybersecurity principles (such as the
confidentiality benefit over the clear text for passwords typed into forms,
e.g. for logins in websites) are not fully understood or shared between the
cybersecurity research community\autocite{Nielsen2009stop}.  The lack of
understanding of basic security principle, the inverse proportionality between
investments in cybersecurity and the number of reported vulnerabilities year
after year, can be linked to the lack of a foundational theory on
cybersecurity, as already highlighted by Cormac Herley
in\autocite{Herley2016unfalsifiability}.

In this article, we give the first scientific theory (to the best of our knowledge)
on security.

\Paragraph{Structure} In Section~\ref{sec:problem} we define and formalize the problem statement. 
In Section~\ref{sec:theory} we outline our security theory, and in Section~\ref{sec:prototest}
we describe the implementation of the theory and some empirical tests of the theory.
Finally, in Section~\ref{sec:related} we conclude the paper with an overview of the related work.

\section{Problem Statement}\label{sec:problem}

\begin{figure}[t]
	\centering
	\includegraphics[width=.6\textwidth]{protocol-example.pdf}
	\caption{Abstraction of an ad-hoc esemplificative protocol execution}
	\label{fig:protocol-example}
\end{figure}

In\autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
``an asymmetry in computer security'', which he defines as follows: ``Things
can be declared insecure by observation, but not the reverse. There is no
observation that allows us to declare an arbitrary system or technique
secure''. Herley then uses this argument to show that ``claims that any measure
is necessary for security are empirically unfalsifiable''. Given that, any
theory which is not falsifiable by an empirical experiment is well
known\footnote{``A theory which is not refutable by any conceivable event is
nonscientific. Irrefutability is not a virtue of a theory (as people often
think) but a vice.'' -- Karl Popper, Conjectures and
Refutations\autocite{popper1962conjectures}} to be nonscientific (i.e.
unfalsifiability is a fallacy of a theory), Herley concludes that there is no
scientific theory on cybersecurity; which means that cybersecurity lays in the
realm of pseudo-sciences\autocite{Herley2016usenixvideo}.  Herley, e.g.
in\autocite{Herley2017justifying}, discusses the implications of a
nonscientific approach to cybersecurity, and highlights the tremendous impact on
all the scientific research and engineering of systems; leading often to
terrorism and wars, and wasting of resources in useless protections or
overspending.  While the criticism is investigated
in\autocite{Herley2016unfalsifiability}, no solution is provided.  On the
contrary, the goal of this work \emph{is} to lay the foundations of a
scientific cybersecurity theory. Furthermore, in Section~\ref{sec:sicurezza},
we consider the problem raised by Herley not confined to ``computer security''
but to any abstract system (so that our theory may hold for any sound
implementation such as networks, mechanical, cyber, or cyber-physical system,
or even a single computer or a single device such as an hard-drive).  There is
also an apparent inconsistency in\autocite{Herley2016unfalsifiability} that we
seek to clarify before following (as we agree) the scientific path draw by
Herley: cybersecurity is defined as an abstract property in many formal
approaches to the investigation of the security of systems, and the security of the design of a
formally verified protocol is indeed falsifiable.  For example, in the protocol
verification community, security is often defined as a formalization of the
high-level properties confidentiality, integrity, and availability. The problem
in such approaches is not the definition of what cybersecurity is, but the use
of theories (such as the Dolev-Yao attacker model\footnote{For the sake of
simplicity, the Dolev-Yao attacker can be considered as an abstraction of an
active attacker who controls the network but cannot break
cryptography.}\autocite{Dolev1983security})
that only applies to specific instances (often called scenarios) and
abstraction of the protocol. This, in turn, creates a false sense of security
since requires assumptions on the abstraction of the system of which security
is verified. As an example, for the formal security verification of the system
in Figure~\ref{fig:protocol-example}, a formalized scenario needs to be defined
by a modeler who chooses (among others): (i) a scope of the formalization (e.g.
excluding the server that distributes the public key is often done when
verifying the security of authentication protocols), (ii) the number of
sessions (even tough some approaches do reason on an infinite number of
sessions such as\autocite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
the peers (e.g.  in the ASLan++ language\autocite{Oheimb2010aslan++}), and (iv)
the abstraction of the cryptographic primitives (e.g.  ProVerif vs
CryptoVerif\autocite{Blanchet2017symbolic}).  Some of the choices will
completely change the results of the formal verification of the system. For
example, under the perfect cryptography assumption\footnote{As defined
in\autocite{Rocchetto2016cpdy}: ``In the so called perfect cryptography
assumption, the security encryption scheme is suppose to be perfect, without
any exploitable flaw, and so the only way for the attacker to decrypt a message
is by using the proper key. That assumption is widely accepted in the security
protocol community, and most of the formal reasoning tools for the analysis of
security protocols abstract away the mathematical and implementation details of
the encryption
scheme\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
and assuming that no violation to any security property is done after message
I); in Figure~\ref{fig:protocol-example}, the freedom of choosing the scope
determines that the flaws related to the dishonest impersonation of the Server
may or may not be considered in the verification process.  This choice has
tremendous impact on the focus and findings of the verification of the security
of the protocol.  While this may seem to turn upon minutiae and foreseeable,
this highlights the false sense of security that may derive from a
non-scientific theory of system security\footnote{``To the superficial
observer, the analysis of these forms seems to turn upon minutiae. It does in
fact deal with minutiae, but they are of the same order as those dealt with in
microscopic anatomy.'' -- Karl Marx, Capital Volume 1, 1867}.

\subsection{Sicurezza: Safety and Security}\label{sec:sicurezza}
In most of the natural languages, and in Italian too, the concepts of safety
and security are not syntactically differentiated and both terms (safety and
security) are expressed by the same word, e.g. sicurezza in Italian.  A
semantic distinction between safety and security is correlated to a
belief\footnote{A belief has to be intended as a proposition which is supposed
to be true by the majority of humans in our society without a scientific
underlying theory but based on partial empirical evidences or inductive proofs.} that
safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
natural environment (e.g. natural events such as wearing of hardware
components) while security deals with \emph{incidents} posed by mankind (e.g.
attackers and bugs).  The fundamental difference between nature and mankind (and,
in turn, between safety and cybersecurity) is believed to be on the different
intents\footnote{``The belief–desire–intention software model (BDI) is a
software model developed for programming intelligent
agents.''\autocite{wiki-bdi}. In the BDI model, the intents represents the
deliberative state of an agent which determines the choice of that agent on
what to do.} (accidents are unfortunate while incidents are not) of the causes
that generates the threat; namely, nature is believed not to have malicious
intents (but unfortunate causes-effects) while threats generated by mankind are
believed to be malicious\footnote{Of course, logical flaws or bugs may be
introduced by other means (e.g. ignorance) without explicit malicious intents,
but the exploitation of those flaws is considered (for now, and detailed
afterwards in the article) malicious, and then we consider any vulnerability to
be malicious (without loss of generality) even if due to the lack of skills.}.
An overview on the aforementioned aspects of safety and security is depicted in
Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
the terms that structure our current understanding of safety and security. 
\begin{itemize}
	\item \emph{Mankind} ``refers collectively to humans''
\autocite{wiki-mankind}, while the concept of \emph{Nature} is
		related ``to the intrinsic characteristics that plants,
		animals, and other features of the world develop of their own
		accord'' (e.g. the physical universe)\autocite{wiki-nature}. 
		\begin{itemize}
			\item So far, we have used several terms to refer to an
				\emph{attacker}, i.e. threat agent or threat source,
				considering those terms to be semantically
				equivalent.  This ``shallowness'' raise form the
				necessity of properly citing the different sources, but,
				in the reminder of this paper, we consider the
				Causality principle to be the \emph{threat
				source}, Nature or Mankind to be the
				\emph{threat agents} and an \emph{attacker} as
				a specific malicious threat agent which materialize a
				threat.
		\end{itemize}
	\item \emph{Vulnerability}\footnote{The term vulnerability is not
		present in the Encyclopedia of Cryptography and Security, while
		it is used in 12 entries (such as in the definition of
		``penetration testing''\autocite{caddy2005pentest})
		highlighting how commonly this word is used without a proper
		supporting semantics}, as defined in\autocite{cnssi20104009}
		(and adopted in\autocite{nist2013800-53}), is ``weakness in an
		information system, system security procedures, internal
		controls, or implementation that could be exploited by a threat
		source''. On the one hand, the definition is broad to enclose
		as much causes (that generates a vulnerability) as possible; on
		the other hand, it derives from empirical evidences (which
		should be considered beliefs\footnote{``For this view, that
		\emph{That Which Is Not} exists, can never predominate. You
		must debar your thought from this way of search, nor let
		ordinary experience in its variety force you along this way,
		(namely, that of allowing) the eye, sightless as it is, and the
		ear, full of sound, and the tongue, to rule; but (you must)
		judge by means of the Reason (Logos) the much-contested proof
		which is expounded by me.'' -- Parmenides of Elea, On Nature
		(circa 500 B.C.), fragments B7.1–8.2
		\autocite{Hakim2016philosophy}} since they are partial results in nature) 
		while a vulnerability should
		be defined in a way that is empirically falsifiable. This means
		that the term vulnerability should have a complete and sound
		definition, so that no other causes (e.g.  other sources) but
		the ones in the definition are responsible for a vulnerability.
		Furthermore, the term ``threat sources'' used in the definition
		in\autocite{cnssi20104009} may be identified with both Nature
		and Mankind, not differentiating between safety and security.
		In Definition~\ref{def:vulnerability}, we provide a formal
		theory of vulnerability (so that the scientific community can
		identify tests for the completeness and soundness of the
		definition itself).
\end{itemize}

\begin{figure}[t]
	\centering
	\includegraphics[width=\textwidth]{safety-security.pdf}
	\caption{Overview security and safety keywords}
	\label{fig:safety-security}
\end{figure}

Most of the safety-preserving principles in the field of engineering of
safety-critical cyber-physical systems (such as elevators and aircraft), upon
which safety requirements (e.g. in standards such as the IEC 61508 or 61511\autocite{IEC201761511}) are defined, have
been defined following empirical tests and measurements. While reasoning by
induction based on the empirical observation should be avoided, since it may
easily lead to false beliefs instead of scientific theories, this approach is
often justified by the supposed impossibility of defining a theory that
correctly predicts failures which, in turn, pose hazards to a system. 
To the best of our knowledge, and supported by\autocite{Herley2016unfalsifiability}, the
correlation between predictability of environment and believed unpredictability
of attackers (i.e. a malicious
environment) has not been correlated to a theory on cybersecurity. 
Therefore, inductive research efforts in
predicting malicious effects are accepted (and published) in scientific
conferences (e.g.~\autocite{Rocchetto2014CSRF}). A failure of a wire due to
environment (e.g. due to humidity, dust, heat \&c) is defined from empirical
evidences and processes have been standardized to test qualities of hardware components
%convinced by the induction principle. What is already a
%weak argument in general, 
This process completely breaks down when a malicious environment (i.e. an attacker)
is considered instead of the (supposedly honest and predictable)
natural environment. Therefore, the same approach that is in use for safety,
seems not to be applicable to test security.

Going back to Figure~\ref{fig:safety-security}, a vulnerability does not
necessarily become a threat for the system, unless exploited ``through a
channel that allows the violation of the security policy
[\ldots]''\autocite{cnssi20104009} (e.g. a software or procedure) that takes
advantage of the vulnerability causing an \emph{attack} to the system, which may
result in several correlated incidents and threats.  The process of
exploitation of a defect as a vulnerability is reported in
Figure~\ref{fig:safety-security} such that the difference between exploit and failure,
and attack and accident is to be found just in the maliciousness of the intents
that causes this process (i.e. excluding the intent, the terms are just syntactic transformation from a vulnerability to defect, from
accident to incident). In the following, we conclude the informal definition of
the terms that we used in this section and in Figure~\ref{fig:safety-security}.

\begin{itemize}
	\item \emph{Causality} refers to the causality principle; defined
		in\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
		connection of phenomena through which one thing (the cause)
		under certain conditions gives rise to, causes something else
		(the effect). The essence of causality is the generation and
		determination of one phenomenon by another. In this respect
		causality differs from various other kinds of connection, for
		example, the simple temporal sequence of phenomena, of the
		regularities of accompanying processes''.
	\item An \emph{Exploit}\footnote{We note that the term exploit is only
		used as a verb in\autocite{ISO2009information}} is ``An exploit
		(from the English verb to exploit, meaning to use something to
		one’s own advantage) is a piece of software, a chunk of data,
		or a sequence of commands that takes advantage of a bug or
		vulnerability to cause unintended or unanticipated behavior to
		occur on computer software, hardware, or something electronic
		(usually computerized).''\autocite{wiki-exploit}.
	\item An \emph{Attack}, as defined by the International Standard
		ISO/IEC 27000 is an ``attempt to destroy, expose, alter,
		disable, steal or gain unauthorized access to or make
		unauthorized use of an asset''; where an \emph{Asset} is
		``anything that has value to the organization''. We note that for
		the purpose of this article, we do not want to focus on a specific
		organization or business to define asset but, in general, on any 
		abstract organization (e.g. a company or a society).
		We do not consider ethical hackers as attacking a system. 
		In fact, we consider the term \emph{hack} as
		non-malicious (see Hacker\autocite{Stallman2002hacker}).
	\item A \emph{Threat}, as defined in\autocite{cnssi20104009}, is ``Any
		circumstance or event with the potential to adversely impact
		organizational operations (including mission, functions, image,
		or reputation), organizational assets, individuals, other
		organizations, or the Nation through an information system via
		unauthorized access, destruction, disclosure, modification of
		information, and/or denial of service''.
	\item \emph{Defect}, ``anything that renders the product not reasonably
		safe''\autocite{Robinson2019writing} (i.e. a characteristic of
		an object which hinders its proper usability).
	\item \emph{Failure}, as defined in\autocite{Merriam2020failure} as ``a state of
		inability to perform a normal function''. The term is
		structured and detailed in
		\autocite{cnssi20104009,iet2017glossary} but relying on an
		abstract notion of failure without a specific definition.
	\item \emph{Hazard}, ``a potential source of
		harm''\autocite{iet2017glossary}.
\end{itemize}

\printbibliography

\end{document}

### 2020-02-05 v-research_experiment-alpha.tex
first draft second part (glossary formalization)

@@
 \documentclass{article}
+\usepackage{mathtools}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{enumitem}
+\usepackage{multicol}
+\usepackage[toc,page]{appendix}
 \usepackage[gen]{eurosym}
@@
 % \bibliography{<mybibfile>}% ONLY selects .bib file; syntax for version <= 1.1b
-\addbibresource{literature.bib}% Syntax for version >= 1.2
+\addbibresource{../../97-literature/literature.bib}% Syntax for version >= 1.2
 
@@
 \newcommand{\Paragraph}[1]{\smallskip\noindent{\bf #1.}}
-\newcommand{\fixnote}[2]{\textbf{\color{red}{NOTE}}\footnote{{\bf #1:} #2}}
-\newcommand{\fix}[2]{{\color{red} #2}}
+\newcommand{\fixnote}[2]{\textbf{\color{red}{FIX}}\footnote{{\bf #1:} #2}}
+\newcommand{\fix}[2]{{\color{red} {\bf tofix:} #2}}
 % \renewcommand{\fixnote}[2]{}
@@
 \newtheorem{lemma}{Lemma}[section]
+\theoremstyle{theorem}
+\newtheorem{theorem}{Theorem}[section]
+\theoremstyle{theorem}
+\newtheorem{example}{Example}
 
@@
 % formulas
-\newcommand{\kframe}{K}
+\newcommand{\kframe}{\mathbf{K}}
+\newcommand{\bframe}{\mathbf{B_K}}
 \newcommand{\possibleworlds}{G}
@@
 \newcommand{\believe}[2]{B_{#1}#2}
+\newcommand{\region}{\chi}
+\newcommand{\Region}{\mathrm{X}}
+\newcommand{\agentuniverse}{\mathcal{A}}
+\newcommand{\agent}{a}
+\newcommand{\follang}{\mathcal{L}}
+\newcommand{\weakness}{w}
+\newcommand{\Weakness}{W}
+\newcommand{\vulnerability}{v}
+\newcommand{\Vulnerability}{V}
+%rcc5
+\newcommand{\rcc}{rcc}
+\newcommand{\connects}[2]{C(#1,#2)}
+\newcommand{\disconnected}[2]{\neg C(#1,#2)}
+\newcommand{\partof}[2]{P(#1,#2)}
+\newcommand{\overlaps}[2]{O(#1,#2)}
+\newcommand{\eq}[2]{EQ(#1,#2)}
+\newcommand{\pp}[2]{PP(#1,#2)}
+\newcommand{\po}[2]{PO(#1,#2)}
+\newcommand{\ppi}[2]{PPi(#1,#2)}
+\newcommand{\dr}[2]{DR(#1,#2)}
 
@@
 	We implemented a formal theory (of axioms) of a mereotopology, and of
-	the region connection calculus (RCC3 and RCC5) in a Python program that
+	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
 	uses the Z3 SMT solver. The results show that a single component (i.e.
@@
 
+\subsection{Glossary -- A Formalization}\label{sec:glossary}
+We now define a formalizations of the concepts described in
+Section~\ref{sec:problem} and depicted in Figure~\ref{fig:safety-security}. We
+base our formalization on first order logic (FOL) with the standard
+truth-value semantics. The choice of this logic is required by the 
+semantics of the concepts formalized afterwards in this section.
+
+We consider a first-order language $\follang$ over a
+signature $\Sigma_P$ where $P,P',\ldots,P^n$ represent terms, 
+$\varphi,\varphi',\ldots,\varphi^n$ represent formulas.
+The syntax is defined as follows.
+\begin{displaymath}
+	\varphi := P~|~\neg\varphi~|~\varphi\wedge\varphi~|~\varphi\vee\varphi~|~\varphi\Rightarrow\varphi'~|~\forall P.~\varphi~|~\exists P.~\varphi
+\end{displaymath}
+where $\wedge$, $\neg$, $\vee$, and $\Rightarrow$ are connectives representing
+conjunction, negation, disjunction, and (material) implication respectively; while
+$\forall$ and $\exists$ represent the standard universal and existential
+quantifiers (resp.). Finally, the symbol ``$.$'' is just syntactic sugar. 
+We consider $\interpretation\subset\Phi\times\{\top,\bot\}$ as the interpretation
+function, where $\Phi$ is any collection of sentences in $\follang$ 
+and $\top$ and $\bot$ represent the concepts of ``Tautology/True''
+and ``Contradiction/False'' respectively.
+
+Following Figure~\ref{fig:safety-security}, we start by formalizing the outermost term: Causality.
+
+\subsubsection{Causality Principle}\label{sec:causality}
+We formalize the Causality Principle starting from a K modal
+logic\autocite{Garson2018modal} (i.e. without restrictions on the causality
+relation between worlds). The standard definition of K modal logic is given in
+Definition~\ref{def:modallogic} in terms of an interpretation function (which
+we named $\sigma$ with a slight abuse of notation) defined in
+Definition~\ref{def:modalinterpretation}. 
+\begin{definition}{\bf K Modal Logic --}\label{def:modallogic}
+A K-frame is a frame $ \kframe=\langle \possibleworlds,\modalrelation \rangle $
+	in the K modal logic where $ \modalrelation $ is the binary relation
+	(i.e. a set of ordered pairs) between possible worlds
+	$\modalrelation\subseteq\possibleworlds\times\possibleworlds$, where
+	$\possibleworlds$ represents the possible worlds, and
+	$\possibleworlds\neq\emptyset$. An actual world
+	$\actualworld\in\possibleworlds$ is assumed. For any proposition $P$,
+	an interpretation function $\interpretation(\world,P)$ returns the
+	truth value of $P$; e.g. $\interpretation(\world,P)=\top$ means that
+	$P$ holds in $\world$. A model is defined as the tuple
+	$\kmodel=\langle\possibleworlds,\modalrelation,\interpretation\rangle$.
+\end{definition}
+
+\begin{definition}{\bf Causality as K Modal Interpretation Function --}\label{def:modalinterpretation}
+	Causality is (recursively) defined as the modal interpretation function $\interpretation$, as follows. 
+	\begin{enumerate}[noitemsep]
+		\item[$(\interpretation0)$] if $\interpretation(\world,P)=\top$ then $\world\models P$
+		\item[$(\interpretation1)$] $\world\models\neg P$ iff $\world\not\models P$
+		\item[$(\interpretation2)$] $\world\models P \wedge Q$ iff $\world\models P$ and $\world\models Q$
+		\item[$(\interpretation3)$] $\world\models\Box P$ iff for any world $\world'\in\possibleworlds$ if $\world \modalrelation\world'$ then $\world'\models P$
+		\item[$(\interpretation4)$] $\world\models\Diamond P$ iff there exists a set of worlds $\World'\subset\possibleworlds$ such that for any $\world'\in\World'$, if $\world\modalrelation\world'$ then $\world'\models P$
+		\item[$(\interpretation5)$] $\models P$ iff $\actualworld\models P$
+	\end{enumerate}
+	where truth is defined as necessary with $\Box$ and possible with $\Diamond$.
+\end{definition}
+
+%\begin{definition}{\bf Causality --}\label{def:causality}
+%	The Causality principle is defined over a K Modal Logic (i.e. without
+%	restrictions on the causality relation between worlds) as in
+%	Definition~\ref{def:modallogic}, with the standard interpretation
+%	function defined in~\ref{def:modalinterpretation}
+%\end{definition}
+
+The causality principle has been defined in its
+generic form. In fact, the accessibility relation $\modalrelation$ is free from
+any axiomatic restriction (e.g. it's not reflexive nor anti-reflexive).  
+We will focus in Section~\ref{sec:process} (for our tests) on 
+the application of our theory to CPS system engineering.
+More detailed case studies (i.e. \fix{mr}{a CPS as a smart power-grid}), 
+will be defined in Section~\ref{sec:prototest}, where
+the accessibility relation defines in details how the system itself evolves, due to causal relation 
+(i.e. by restricting the causality principle only to those cause-effect that defines
+the system). Therefore, the definition of $\modalrelation$
+will be specialized in a more strict way based on the application domains and case study.
+
+\fix{mr}{We note that we have, so far, defined the
+Causality principle without considering it a \emph{threat source} since 
+we lack the concept of intent and maliciousness. We introduce those concepts
+in Section~\ref{sec:mankind-nature} and then formally define a threat agent
+in the same section.}
+
+\subsubsection{Agents: Mankind and Nature}\label{sec:mankind-nature}
+\begin{table}[t]
+\centering
+\setlength{\tabcolsep}{3.5pt}
+\renewcommand{\arraystretch}{1}
+\scriptsize
+\caption{RCC3, RCC5, and RCC8 relations between Regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
+\begin{tabular}{ccclll} 
+\rota{\textbf{RCC3}}&\rota{\textbf{RCC5}}&\rota{\textbf{RCC8}}&\textbf{Name} & \textbf{Notation} & \textbf{Definition} \\
+\hline
+&&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& $\mathit{X}\subseteq \mathit{Y}$ \\
+&&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& $\mathit{X}\not\subseteq \mathit{Y}$\\
+&&&Part of				& $\mathit{P}(\mathit{X},\mathit{Y})$		& $\forall \mathit{Z} ~\mathit{C}(\mathit{Z},\mathit{X}) \rightarrow \mathit{C}(\mathit{Z},\mathit{Y})$\\
+&&&Overlaps			& $\mathit{O}(\mathit{X},\mathit{Y})$		& $\exists \mathit{Z} ~\mathit{P}(\mathit{Z},\mathit{X})\wedge \mathit{P}(\mathit{Z},\mathit{Y})$\\
+\Tdot&&&  \textbf{Overlaps Not Equal} 	& $\mathit{ONE}(\mathit{X},\mathit{Y})$		& $\mathit{O}(\mathit{X},\mathit{Y}) \land \neg \mathit{EQ}(\mathit{X},\mathit{Y})$ \\
+\Tdot&\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
+\Tdot&\Tdot&\Tdot& \textbf{DiscRete from} 		& $\mathit{DR}(\mathit{X},\mathit{Y})$		& $\neg \mathit{O}(\mathit{X},\mathit{Y})$\\
+&\Tdot&\Tdot&\textbf{Partial-Overlap}	& $\mathit{PO}(\mathit{X},\mathit{Y})$ 		& $\mathit{O}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{X},\mathit{Y}) \wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\
+&\Tdot&&\textbf{Proper-Part-of} 	& $\mathit{PP}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\ 
+	&\Tdot&&\textbf{Proper-Part-of-\textit{\textbf{i}}nverse} & $\mathit{PPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{Y},\mathit{X}) \wedge \neg \mathit{P}(\mathit{X},\mathit{Y})$\\
+&&\Tdot&\textbf{Externally Connected} 	& $\mathit{EC}(\mathit{X},\mathit{Y})$ 		& $\mathit{C}(\mathit{X},\mathit{Y}) \wedge \neg\mathit{O}(\mathit{X},\mathit{Y})$\\ 
+&&\Tdot&\textbf{Tangential PP} 	& $\mathit{TPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
+&&\Tdot&\textbf{Tangential PPi} 	& $\mathit{TPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{TPP}(\mathit{Y},\mathit{X})$\\ 
+&&\Tdot&\textbf{Non-Tangential PP} 	& $\mathit{NTPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\neg\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
+&&\Tdot&\textbf{Non-Tangential PPi} 	& $\mathit{NTPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{NTPP}(\mathit{Y},\mathit{X})$\\ 
+\end{tabular}
+\end{table}
+
+Mankind and Nature are defined in Section~\ref{sec:problem} as two abstract
+agents, both as collections (i.e. an abstract type that does not imply a
+specific implementation) of their sub-agents (i.e. humans for Mankind and
+plants, animals, \&c.  for Nature). Similarly to\autocite{Santaca2016abf}, 
+we define Mankind and Nature, and any
+other agent in the reminder of this article, as a meronomy (an hierarchy of
+Part-Whole relations) based on a standard definition of mereology, i.e. based
+on the definition of Parthood relation between \emph{Parts}.  
+However, we will need to
+consider different types of Part; so, we extend the mereology to a
+mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
+to increase the number of different types considered and to generalize the relations between Parts (as
+in Table~\ref{tab:rcc358}).
+For the sake of readability, we use the term \emph{Region} both to refer to a
+mereological Part and to a topological Region.
+The choice of mereotopology is also correlated to the objective of defining a formal
+ontology, which we use to define the (formal) semantics of the terms (Parts) in
+Section~\ref{sec:problem}, and of the concepts of safety and security
+(whole). We aim at creating a meronomy instead of the taxonomies such as the one provided
+in\autocite{NIST2020NVD,MITRE2020CVE} or instead of the dogmatic 
+CVSS\autocite{Mell2007CVSS} scoring system.
+
+A mereotopology, as defined e.g. in\autocite{Rachavelpula2017mereotopology},
+is an ordered mathematical structure where the basic relation between Regions is the
+reflexive and symmetric\fixnote{mr}{I guess it must be monotonic as defined in\autocite{Rachavelpula2017mereotopology} but I don't find it consistently in other papers.} 
+\emph{Parthood} relation $\subseteq$. 
+\begin{definition}{\bf Parthood --}\label{def:parthood}
+Given any pair of mereotopological Regions $X$ and $Y$,
+	\begin{enumerate}[noitemsep]
+		\item Reflexivity: $\forall X.~ (X\subseteq X)$
+		\item Symmetry: $\forall X, Y.~ (X\subseteq Y \Rightarrow Y\subseteq X)$
+	\end{enumerate}
+\end{definition}
+
+As later defined in Definition~\ref{def:agent}, 
+the Parthood relation orders a universe of agents $\agentuniverse$ by defining
+the so called \emph{Connects with} (see in Table~\ref{tab:rcc358}) relation 
+between Regions. We want this universe $\agentuniverse$ to be 
+expressible in FOL. In this way, we
+can reason both on the constituent of security (i.e. its terms and agents
+defining a system where security needs to be considered), and on 
+the evolution of those constituent w.r.t. cause-effects relations according
+to the modal structure of causality we defined in Section~\ref{sec:causality}.
+This will allow us (in Section~\ref{sec:process} and Section~\ref{sec:prototest})
+a better\fixnote{mr}{Can we prove this by construction?} 
+positioning w.r.t. risk assessment technologies
+(which most often reason on the constituent of a system design), and
+protocol verification tools (which requires some formalization of a cause-effect relation, e.g.
+the Linear Temporal Logic).
+
+As we argued in Section~\ref{sec:problem}, we must correlate the definition of
+agent (i.e. Mankind and Nature) in Definition~\ref{def:mankind-nature} to the
+mathematical structure of the logic that defines them.  We express Mankind and
+Nature as formulas over the theory of mereology and then in terms of
+mereotopological Regions, extending the interpretation function
+$\interpretation$, to include a formal theory of mereotopology.  We use the
+Region Connection Calculus (RCC), as defined
+in~\cite{bennettLogics,improvingRCC}, to provide an axiomatization of the
+spatial concepts and relations in first-order logic to correlate the algebraic
+structure to mereology. In its broader definition, the RCC theory is composed
+by eight axioms, and is known as RCC8, but here we restrict to RCC5 (without
+loss of generality) by not considering tangential connections between spatial
+Regions. We discuss the choice of RCC5 in more detail in
+Appendix~\ref{app:rcc}. In Table~\ref{tab:rcc}, we
+summarize the axioms of the Region Connection Calculus (see, e.g., \autocite{Grutter2008rcc}).
+
+\begin{definition}{\bf RCC axiomatization --}
+	For any $X,Y$ pair of Regions in a mereotopology:
+	\begin{itemize}[noitemsep]
+	\item[$(\interpretation6)$] $\interpretation(X\subseteq Y)$ iff [$\interpretation(X\subseteq X)=\top$, and $\interpretation(X\subseteq Y)=\bot$ or $\interpretation(Y\subseteq X)=\top$]\fixnote{mr}{would it be better/clearer or just correct to write $\sigma(X\subseteq Y)$ iff $\sigma(X\subseteq X\wedge [X\subseteq Y \vee Y\subseteq X])=\top$}
+	\item[$(\interpretation7)$] $\world\models\connects{X}{Y}$ iff $\world\models X\subseteq Y$ 
+	\item[$(\interpretation8)$] $\world\models\disconnected{X}{Y}$ iff  $\world\not\models X\subseteq Y$
+	\item[$(\interpretation9)$] $\world\models\partof{X}{Y}$ iff $\world\models\forall Z.~\connects{Z}{X}\Rightarrow\connects{Z}{Y}$
+	\item[$(\interpretation10)$] $\world\models\overlaps{X}{Y}$ iff $\world\models\exists Z.~\partof{Z}{X}\wedge\partof{Z}{Y}$
+	\item[$(\interpretation11)$] $\world\models\eq{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\partof{Y}{X}$
+	\item[$(\interpretation12)$] $\world\models\dr{X}{Y}$ iff $\world\models\neg\overlaps{X}{Y}$
+	\item[$(\interpretation13)$] $\world\models\po{X}{Y}$ iff $\world\models\overlaps{X}{Y}\wedge\neg\partof{X}{Y}\wedge\neg\partof{Y}{X}$
+	\item[$(\interpretation14)$] $\world\models\pp{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\neg\partof{Y}{X}$
+	\item[$(\interpretation15)$] $\world\models\ppi{X}{Y}$ iff $\world\models\partof{Y}{X}\wedge\neg\partof{X}{Y}$
+\end{itemize}
+where $Z$ is a mereotopological Region.
+\end{definition}
+
+\begin{definition}{\bf Agent: Mankind or Nature --}\label{def:agent}
+	An agent $\agent\in\agentuniverse$ is a tuple
+	$\langle\rcc(\region',\region''),\ldots,\rcc(\region^{n-1},\region^n)\rangle$
+	of RCC relations $\rcc$ over mereotopological Regions
+	${\region',\ldots,\region^n}\subseteq\Region$. 
+\end{definition}
+
+Currently, we do not distinguish between Mankind and Nature (since we still
+lack of the definition of ``malicious intent'', which is defined in
+Section~\ref{sec:theory}) and we have defined them as two generic
+agents.\fixnote{mr}{check if it's a repetition and if links to sections are
+consistent}
+
+As depicted in Figure~\ref{fig:safety-security}, Causality, Mankind, and Nature
+have a dashed border representing their correlation in terms of cause-effect
+and then in terms of formal structure which defines them: Modal Logic.
+Vulnerability, Defect, Incident, Accident, and Threat, similarly, 
+are correlated (depicted as a solid border) in terms of underlying 
+formal structure: Mereotopology.
+
+\subsubsection{Regions: Vulnerability and Defect (and Weakness)}\label{sec:vulnerabilitydefect}
+As defined in Section~\ref{sec:problem}, a Vulnerability or a Defect is a
+\emph{Weakness}.  As an example, a categorization of weaknesses is given
+in\autocite{MITRE2020CWEresearch} with 808 weaknesses categorized as ``Research
+Concepts'', distributed as follows:
+\begin{itemize}[noitemsep]
+	\item Incorrect Calculation - ($682$)
+	\item Incorrect Access of Indexable Resource (``Range Error'') - ($118$)
+	\item Use of Insufficiently Random Values - ($330$)
+	\item Improper Interaction Between Multiple Correctly-Behaving Entities - ($435$)
+	\item Improper Control of a Resource Through its Lifetime - ($664$)
+	\item Insufficient Control Flow Management - ($691$)
+	\item Protection Mechanism Failure - ($693$)
+	\item Incorrect Comparison - ($697$)
+	\item Improper Check or Handling of Exceptional Conditions - ($703$)
+	\item Improper Enforcement of Message or Data Structure - ($707$)
+	\item Improper Adherence to Coding Standards - ($710$)
+\end{itemize}
+
+The definition given by the MITRE in\autocite{MITRE2020CWEweakness} of
+weakness is: ``Software weaknesses are errors that can lead to software
+vulnerabilities. A software vulnerability, such as those enumerated on the
+Common Vulnerabilities and Exposures (CVE) List, is a mistake in software that
+can be directly used by a hacker to gain access to a system or network''.
+The definition is circular if we interpret the word ``error'' and
+``mistake'' with the same semantics: a weakness is an error that leads to a
+vulnerability and a vulnerability is a mistake which, in turn, is a weakness.
+The only difference (between weakness and vulnerability) seems to be
+that one can consider weakness as a ground term and state that a
+vulnerability is caused by a weakness, i.e. 
+$\World,\Weakness\models\Diamond\Vulnerability$ where
+$\Weakness,\Vulnerability$ are Regions of Weaknesses and Vulnerabilities (resp.);
+accepting the hierarchy in the CWE\autocite{CWE} as a ground truth. 
+Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability) 
+or the CVE reported in the NVD as a ground truth.
+
+\begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
+	When a Region $\region\subseteq\agent$ of an agent
+	$\agent\in\agentuniverse$ is identified as Weakness $\Weakness$ (i.e.
+	representing a weakness introduced by the agent into any phase of
+	production, e.g.  of the secure process development life-cycle, of any
+	system or subsystem) iff there exists $\world\in\possibleworlds$ such
+	that $\World,\Weakness\models\Diamond\exists\Weakness',\Vulnerability'.~
+	\po{\Weakness'}{\Vulnerability'}$, where $\Weakness'\subseteq\Weakness$, and
+	$\Vulnerability'\subseteq\Vulnerability$ is a Region of an agent $\agent'\in\agentuniverse$.
+\end{definition}
+
+\begin{example}{CWE-116: Improper Encoding or Escaping of Output\autocite{CWE-116} --}
+	\begin{itemize} 
+		\item Description: The software prepares a structured message
+		for communication with another component, but encoding or
+		escaping of the data is either missing or done incorrectly. As
+		a result, the intended structure of the message is not
+		preserved. 
+		\item Example: This code displays an email
+		address that was submitted as part of a form. Example language JSP.
+		\begin{verbatim} 
+		<% String email = request.getParameter("email"); %> 
+		...
+		Email Address: <%= email %>
+		\end{verbatim}
+		The value read from the form parameter is
+		reflected back to the client browser without
+		having been encoded prior to output, allowing
+		various XSS attacks (CWE-79).
+		\item Observed Examples
+			\begin{itemize}
+			\item CVE-2008-4636\autocite{CVE-2008-4636}: OS command injection in backup
+				software using shell metacharacters in a
+				filename; correct behavior would
+				require that this filename could not be
+				changed.
+			\item CVE-2008-0769\autocite{CVE-2008-0769}: Web application does not set the
+				charset when sending a page to a browser,
+				allowing for XSS exploitation when a
+				browser chooses an unexpected encoding.
+			\item CVE-2008-0005\autocite{CVE-2008-0005}: Program does not set the charset
+				when sending a page to a browser, allowing for
+				XSS exploitation when a browser chooses
+				an unexpected encoding.
+			\item CVE-2008-5573\autocite{CVE-2008-5573}: SQL injection via password
+				parameter; a strong password might contain \&
+			\item CVE-2008-3773\autocite{CVE-2008-3773}: Cross-site scripting in chat
+				application via a message subject, which
+				normally might contain \& and other
+				XSS-related characters.
+			\item CVE-2008-0757\autocite{CVE-2008-0757}: Cross-site scripting in chat
+				application via a message, which normally might
+				be allowed to contain arbitrary
+				content.
+			\end{itemize}
+\end{itemize}
+
+\begin{figure}[t]
+	\centering
+	\begin{subfigure}[b]{.5\textwidth}
+		\centering
+		\includegraphics[width=.7\textwidth]{rcc5_CWE-116.pdf}
+		\caption{Representation with RCC5}
+		\label{fig:rcc5_CWE-116}
+	\end{subfigure}\begin{subfigure}[b]{.5\textwidth}
+		\centering
+		\includegraphics[width=.7\textwidth]{rcc8_CWE-116.pdf}
+		\caption{Representation with RCC8}
+		\label{fig:rcc8_CWE-116}
+	\end{subfigure}
+	\begin{subfigure}[b]{.5\textwidth}
+		\centering
+		\includegraphics[width=.7\textwidth]{rcc8_CWE-116.pdf}
+		\caption{Representation in K-Modal Logic}
+		\label{fig:KML_CWE-116}
+	\end{subfigure}
+	\caption{Relations between CWE-116 and correlated CVEs}
+\end{figure}
+
+\end{example}
+
+\begin{definition}{\bf Region: Vulnerability and Defect --}\label{def:vulnerability-defect}
+A region $\region\subseteq\agent$ of an agent $\agent\in\agentuniverse$ is
+	called \emph{Vulnerability} if the agent $\agent$ is referred to as Mankind, 
+	\emph{Defect} if the agent is referred to as Nature.
+\end{definition}
+
+In our informal definition depicted in Figure~\ref{fig:safety-security}, a
+Vulnerability is seen as a process through which an agent (i.e. Mankind)
+generates an Incident, i.e.
+Agent$\rightarrow$Vulnerability$\rightarrow$Exploit$\rightarrow$Attack$\rightarrow$Incident.
+We describe processes, in our current formalization, in terms of the K-Modal
+Structure in Definition~\ref{def:modalinterpretation} (i.e. in terms of
+causality principle).
+
+\begin{definition}{\bf Vulnerability-Incident Process --}\label{def:vulnerability-incident}
+	Given two 
+\end{definition}
+
+The whole-part relation \fix{mr}{(i.e.  HAS-A\footnote{HAS-A is just an hint for the
+mechanization of this theory.  We'll come back to this in
+Section~\ref{sec:implementation}})} is given, in the following, over the K Modal
+logic described beforehand. To do so, we first need to formally define the
+concept of \emph{agent} in its abstract form. We build it following the same
+structure of\autocite{Santaca2016abf} but over the terms used in
+Section~\ref{sec:problem}, i.e. knowledge, and belief as epistemic concepts. We'll then 
+introduce the concept of \emph{intent} to define a \emph{threat agent}.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.8\textwidth]{knowledge-belief.pdf}
+	\caption{Informal representation of Knowledge and Belief}
+	\label{fig:knowledge-belief}
+\end{figure}
+
+The difference between Knowledge and Belief is depicted in
+Figure~\ref{fig:knowledge-belief} (see \autocite{wiki-knowledgebelief}).  However,
+according to\autocite{Gettier2012knowledge},
+Knowledge\footnote{``\emph{Theaetetus}: [\ldots] He said that knowledge was
+true opinion accompanied by reason, but that unreasoning true opinion was
+outside of the sphere of knowledge; and matters of which there is not a
+rational explanation are unknowable -- yes, that is what he called them -- and
+those of which there is are knowable. [\ldots] \emph{Socrates}: [\ldots] the
+primary elements of which we and all else are composed admit of no rational
+explanation; for each alone by itself can only be named, and no qualification
+can be added, neither that it is nor that it is not, for that would at once be
+adding to it existence or non-existence, whereas we must add nothing to it, if
+we are to speak of that itself alone.  [\ldots]'' Plato -- Theaetetus 201
+\autocite{Plato1914Plato}} as an epistemological concept is difficult (and
+sometimes believed impossible\autocite{citation}) to formally define. In this
+work, we are not interested in 
+how knowledge can be precisely formalized from an epistemic standpoint. 
+We assume that a semantic of a correct definition of epistemic knowledge exists,
+for example the one given in\autocite{Hintikka1962knwoledge} by Hintikka, and we then define 
+knowledge in terms of the Kripke structure defined in Definition~\ref{def:causality}.
+
+\begin{definition}{\bf Knowledge --}\label{def:knowledge}
+Given an abstract collection of Agents $Ag$, and the modal operator
+	$\knows{a}{}$ (where $a\in Ag$), Knowledge is defined as a collection
+	of predicates known by an agent $\knowledge{a}=\bigcup_\Phi \knows{a}{\varphi}$ 
+	(where $\Phi$ is the collection of all the propositions known by $a$).
+	Given a proposition $P$, we extend the semantics of the Causality structure with:
+	\begin{enumerate}[noitemsep]
+		\item[$(\interpretation6)$] $\world\models\knows{a}{P}$ iff
+			$\world'\models P$ for all $\world'$ such that
+			$\world\modalrelation\world'$.
+	\end{enumerate}
+\end{definition}
+
+\begin{definition}{\bf Belief --}\label{def:belief}
+	\fix{mr}{we should help the reader in understanding the upgrades of the logic.
+We used propositional with modal operators with a K Kripke semantics and we are now introducing Belief as in Doxastic logic?
+and we have epistemological operators as knowledge}
+	Given an abstract collection of Agents $Ag$, and the modal operator
+	$\believe{a}{}$ (where $a\in Ag$), Belief is defined as a collection
+	of predicates believed by an agent $\belief{a}=\bigcup_\Phi \believe{a}{\varphi}$
+	(where $\Phi$ is the collection of all the propositions believed by $a$).
+	Given a proposition $P$, we extend the semantics of the Causality structure with:
+	\begin{enumerate}[noitemsep]
+		\item[$(\interpretation7)$] $\world\models\believe{a}{P}$ iff
+			$\world\models\neg\knows{a}{\neg P}$ (i.e. the agent $a$ considers $P$ possible) 
+			and $\world'\models P$ for all
+			$\world'$ such that $\world\modalrelation\world'$.
+	\end{enumerate}
+\end{definition}
+
+\begin{definition}{\bf Mankind/Nature --}\label{def:mankind-nature}
+Mankind/Nature is represented as an agent.  
+	%countable logical signature $\Sigma_M$/$\Sigma_N$ over an
+	%abstract algebraic structure.  We also consider a countable signature
+	%of variables, disjoint from $\Sigma_M$/$\Sigma_N$, and write
+	%$\Sigma_{M_n}$/$\Sigma_{N_n}$ for the symbols of $\Sigma_M$/$\Sigma_N$
+	%with arity $n$. Thus, $\Sigma_0$ is a collection of constants, which we
+	%assume to have disjoint sub-collections that we refer to as
+	%\emph{sub-agents} of Mankind or Nature (i.e. humans, plants, \&c).
+	%The variables are untyped and can be instantiated with arbitrary types,
+	%yielding an untyped model.
+\end{definition}
+
+\subsubsection{Mereotopological Representation of Safety and Security}\label{sec:formalGlossary}
+The objectives of this section are:
+\begin{enumerate}[noitemsep]
+	\item formalize the abstract algebraic structure (as a mereotopology)
+		on top of which we have formalized the terms in
+		Figure~\ref{fig:safety-security} so far, i.e. Mankind, Nature,
+		Truths, Beliefs, and Knowledge.
+	\item correlate mereological structure to Causality and then Kripke
+		structure by defining the relation $R$ over the Kripke
+		structure as the mereotopological basic relation connects with,
+		i.e. reflexive and symmetric
+	\item we map signature to mereotopological Regions to express the
+		logical formulation of terms (e.g. Mankind, Nature, Truths,
+		Beliefs, Knowledge, Vulnerabilities) over the mereotopological
+		structure
+	\item we formalize the relations between terms using the RCC(5)
+\end{enumerate}
+
+\fix{mr}{change propositions in propositional language\url{http://logic.stanford.edu/intrologic/lectures/lecture_02.pdf}}
+
+Given that the concepts of Truths, Beliefs, and Knowledge are positioned over
+an underlying structure of Propositions, the first concept to formalize is that
+of Propositions. We do not restrict Propositions to a specific structure but,
+as for Mankind and Nature, we refer to them as a collection.  We express the
+generality of a collection as a (propositional) signature over an abstract
+algebraic structure. In our formalization, we use mereotopology as algebraic
+structure (due to its generality, as discussed in Section~\ref{sec:problem})
+and the RCC to formalize concepts over a mereotopology.
+
+We can now define the possible mereotopological relations between Belief and
+Knowledge (informally depicted in Figure~\ref{fig:knowledge-belief}) and then
+define the peculiarity of the believed difference between Mankind and Nature 
+(i.e. maliciousness as intent). 
+We first start from the formalization of Figure~\ref{fig:knowledge-belief}, as
+follows.
+
+\begin{definition}{\bf Information}\label{def:information}
+\end{definition}
+\begin{definition}{\bf Desire}\label{def:desire}
+\end{definition}
+\begin{definition}{\bf Intent}\label{def:intent}
+\end{definition}
+
+\subsection{Vulnerability and Defect}\label{sec:vulnerability-defect}
+In order to formalize the semantics of the concepts (informally) described in
+Section~\ref{sec:problem},
+
+\begin{definition}{\bf Vulnerability}
+\end{definition}
+
+\subsection{Cybersecurity Engineering Process}\label{sec:process}
+%\epigraph{La differenza tra idee e realt\`a \`e la differenza tra
+%filosofia e ingegneria. Il lavoro per trasformare l'una nell'altra \`e ricerca
+%scientifica
+\epigraph{The difference between ideas and reality is the
+difference between philosophy and engineering. The work to transform one into
+the other is scientific research}{{\itshape V-Research}}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\textwidth]{vmodel.pdf}
+	\caption{Cybersecurity Engineering Life-cycle}
+	\label{fig:knowledge-belief}
+\end{figure}
+
+\subsubsection{Relevant Standards}\label{sec:standards}
+\begin{enumerate}[noitemsep]
+	\item DO-326A
+\end{enumerate}
+
+\section{Honesty and Maliciousness -- A Cybersecurity Theory}\label{sec:theory}
+\begin{enumerate}[noitemsep]
+	\item cybersecurity as a requirement
+	\item vulnerability\label{def:vulnerability}
+\end{enumerate}
+
+\section{Prototype and Empirical Tests}\label{sec:prototest}
+\subsection{Prototype Implementation}\label{sec:implementation}
+\subsection{Empirical Tests}\label{sec:tests}
+
+\section{Conclusion and Related Work}\label{sec:related}
+
 \printbibliography
 
+\begin{appendices}
+	\section{RCC3, RCC5, or RCC8}\label{app:rcc}
+	\section{Wrong idea 1}
+In order to apply consistently the mereotopological structure not just to
+reason the agents but also on their evolution, we extend the properties of the
+Parthood relation (reflexivity and symmetry) to the modal relation
+$\modalrelation$ of the K Modal logic, which in turns defines the causality
+principle. This updates the K Modal logic to a B (Brouwer) Modal Logic\autocite{Garson2018modal}.
+
+\begin{definition}{\bf B Modal Logic --}\label{thm:bk-parthood}
+	The B Modal Logic (defined, with a slight abuse of notation as
+	$\bframe=\langle\possibleworlds,\modalrelation\rangle$) is an extension
+	of the K modal logic with the following axioms
+	\begin{enumerate}[noitemsep]
+		\item[$(\interpretation8)$] $\world\models P\Rightarrow Q$ iff
+			$\world\not\models P$ or $\world\models Q$ (logical implication)~\footnote{In
+			Definition~\ref{def:modallogic} we didn't specify the
+			semantics in the case of the connective $\Rightarrow$
+			(i.e. logical implication), so we state it here. Given
+			that $P\Rightarrow Q$ is equisatisfiable to
+			$\neg(P\wedge \neg Q)$, for any proposition $P$ and
+			$Q$, $\world\models P\Rightarrow Q$ is equisatisfiable
+			to $\world\models \neg(P\wedge \neg Q)$. If
+			$\world\models \neg(P\wedge \neg Q)$ then
+			$\world\not\models P\wedge\neg Q$ and then
+			$\world\models \neg P$ or $\world\models Q$.}
+		\item[$(\interpretation9)$] $\Box P\Rightarrow P$ (which ``claims that whatever is necessary is the case''\autocite{Garson2018modal}), and
+		\item[$(\interpretation10)$] $P\Rightarrow\Box\Diamond P$
+			(which says that if $P$ is the case, then $P$ is
+			necessarily possible\autocite{Garson2018modal})
+	\end{enumerate}
+\end{definition}
+
+\begin{theorem}{\bf Causality, Parthood, and B Modal Logic}\label{thm:bk-parthood}
+	\begin{enumerate}[noitemsep]
+		\item (Reflexivity) $\bframe\models\Box P\Rightarrow P$ ($\interpretation8$) iff $\modalrelation$ is reflexive
+		\item (Symmetry) $\bframe\models P\Rightarrow\Box\Diamond$ ($\interpretation9$) iff $\modalrelation$ is symmetric
+		\item (Consistency) $\world\modalrelation\world'$ iff $\connects{\region}{\region'}$
+	\end{enumerate}
+	whenever $\world$ and $\world'$, and $\region$ and $\region'$ are logically equivalent ($\equiv$).
+\end{theorem}
+\begin{proof}
+	The proof is exhaustive over all possible cases.
+	\begin{enumerate}[noitemsep]
+		\item(Reflexivity) proven in\autocite{}
+		\item(Symmetry) proven in\autocite{}
+		\item(Equivalence) \fix{mr}{todo} If
+			$\world\modalrelation\world'$ and
+			$\world\equiv\region$, then if
+			it doesn't hold that $\connects{\region,\region'}$ it
+			follows that there is no reflexive and transitive
+			relation that correlates $\region$ to $\region'$. Given 
+			that Regions are defined as signatures
+	\end{enumerate}
+\end{proof}
+
+\end{appendices}
+
 \end{document}

### 2020-02-05 compile.sh
just copied the horrible compile script from branch master

@@ -1,0 +1,5 @@
+#!/bin/bash
+pdflatex v-research_experiment-alpha.tex
+biber v-research_experiment-alpha
+pdflatex v-research_experiment-alpha.tex
+pdflatex v-research_experiment-alpha.tex

### 2020-02-05 prototypes/agent_categorization/rcc5_connectsWith_infiniteDomain.py
prototypes for agent categorization based on ABF theory

@@ -1,0 +1,124 @@
+from z3 import *
+
+Region = DeclareSort('Region')
+
+X = Const('X', Region)
+Y = Const('Y', Region)
+Z = Const('Z', Region)
+
+C  = Function('C', Region, Region, BoolSort())
+P  = Function('P', Region, Region, BoolSort())
+O  = Function('O', Region, Region, BoolSort())
+EQ = Function('EQ', Region, Region, BoolSort())
+DR = Function('DR', Region, Region, BoolSort())
+PO = Function('PO', Region, Region, BoolSort())
+PP = Function('PP', Region, Region, BoolSort())
+PPi= Function('PPi', Region, Region, BoolSort())
+
+################################
+# CONNECTS WITH
+# define here the parthood relation C (primitive binary inclusion relation called connects with)  
+# which is reflexive and antisymmetric 
+################################
+
+reflexivity=ForAll(X, C(X, X))
+symmetry=ForAll([X,Y], Implies(C(X, Y), C(Y,X)) )
+#extensionality= ForAll([X,Y], Implies(ForAll(Z, C(Z, X) == C(Z, Y)), X == Y))
+
+################################
+# PART OF
+# P(X,Y) : forall Z C(Z,X) => C(Z,Y)
+################################
+
+part_of=ForAll([X,Y], P(X,Y) == ForAll(Z, Implies(C(Z,X), C(Z,Y))))
+
+################################
+# OVERLAPS 
+# O(X,Y) : exists Z P(Z, X) /\ P(Z, Y) 
+################################
+
+overlaps=ForAll([X,Y], O(X,Y) == Exists(Z, And(P(Z,X), P(Z,Y))))
+
+################################
+# EQUAL TO 
+# E(X,Y) : P(X, Y ) /\ P(Y, X ) 
+################################
+
+equal_to=ForAll([X,Y], EQ(X,Y) == And(P(X,Y), P(Y,X)))
+
+################################
+# DISCRETE FROM
+# DR(X,Y) : not O(X,Y)
+################################
+
+discrete_from=ForAll([X,Y], DR(X,Y) == Not(O(X,Y)))
+
+################################
+# PARTIAL OVERLAP
+# PO(X,Y) : O(X, Y) /\ (not P(X, Y)) /\ (not P(Y, X))
+################################
+
+partial_overlap=ForAll([X,Y], PO(X,Y) == And(O(X,Y), Not(P(X,Y)), Not(P(Y,X))))
+
+################################
[…]

### 2020-02-06 compile.sh
improve compile script

@@ -1,5 +1,23 @@
 #!/bin/bash
-pdflatex v-research_experiment-alpha.tex
-biber v-research_experiment-alpha
-pdflatex v-research_experiment-alpha.tex
-pdflatex v-research_experiment-alpha.tex
+latexOutputDir="latex-output-dir"
+latexInputFile="v-research_experiment-alpha.tex"
+pdfOutputDir="${latexOutputDir}/${latexInputFile%.*}.pdf"
+
+if [[ ! -d $latexOutputDir ]]; then 
+	mkdir $latexOutputDir
+fi
+
+if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
+then 
+	if biber --nodieonerror --input-directory=$latexOutputDir --output-directory=$latexOutputDir ${latexInputFile%.*}
+	then
+		if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
+		then 
+			pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile 
+			if [[ -f $pdfOutputDir ]]
+			then
+				mv $pdfOutputDir `dirname $0`
+			fi
+		fi
+	fi
+fi

### 2020-02-07 v-research_experiment-alpha.tex
minor changes
minor changes (weakness definition)

@@
 % \bibliography{<mybibfile>}% ONLY selects .bib file; syntax for version <= 1.1b
-\addbibresource{../../97-literature/literature.bib}% Syntax for version >= 1.2
+\addbibresource{literature.bib}% Syntax for version >= 1.2
 
 \usepackage{graphicx}
-\graphicspath{{./00-images/}}
+\graphicspath{{./images/}}
 
@@
 \begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
-	When a Region $\region\subseteq\agent$ of an agent
-	$\agent\in\agentuniverse$ is identified as Weakness $\Weakness$ (i.e.
+	A Region $\region\subseteq\agent$ of an agent
+	$\agent\in\agentuniverse$ is defined as Weakness $\Weakness$ (i.e.
 	representing a weakness introduced by the agent into any phase of

### 2020-02-07 v-research_experiment-alpha.tex
conclude first draft sec 2.2.3 (vuln & defect)

@@
 vulnerability is caused by a weakness, i.e. 
-$\World,\Weakness\models\Diamond\Vulnerability$ where
+$\World,\Weakness\models\Diamond\Vulnerability\wedge\Weakness$ where
 $\Weakness,\Vulnerability$ are Regions of Weaknesses and Vulnerabilities (resp.);
@@
 	A Region $\region\subseteq\agent$ of an agent
-	$\agent\in\agentuniverse$ is defined as Weakness $\Weakness$ (i.e.
+	$\agent\in\agentuniverse$, is defined as Weakness $\Weakness$ (i.e.
 	representing a weakness introduced by the agent into any phase of
@@
 	system or subsystem) iff there exists $\world\in\possibleworlds$ such
-	that $\World,\Weakness\models\Diamond\exists\Weakness',\Vulnerability'.~
-	\po{\Weakness'}{\Vulnerability'}$, where $\Weakness'\subseteq\Weakness$, and
+	that $\world,\Weakness\models\Diamond\exists\Weakness',\Vulnerability'.~
+	\rcc(\Weakness',\Vulnerability')\wedge\neg\dr{\Weakness'}{\Vulnerability'}$, 
+	where $\Weakness'\subseteq\Weakness$, and
 	$\Vulnerability'\subseteq\Vulnerability$ is a Region of an agent $\agent'\in\agentuniverse$.
@@
 	\end{subfigure}
-	\begin{subfigure}[b]{.5\textwidth}
+	\begin{subfigure}[b]{\textwidth}
 		\centering
-		\includegraphics[width=.7\textwidth]{rcc8_CWE-116.pdf}
+		\includegraphics[width=\textwidth]{KML_CWE-116.pdf}
 		\caption{Representation in K-Modal Logic}
@@
 
+\subsubsection{Process: Incidents and Accidents}\label{sec:vulnerabilitydefect}
 In our informal definition depicted in Figure~\ref{fig:safety-security}, a
@@
 \end{figure}
+
+\subsubsection{Region: Threat and Asset}\label{sec:vulnerabilitydefect}
 

### 2020-02-18 compile.sh
add -short -s option to compile without biber

@@ -9,14 +9,17 @@ fi
 
 if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
 then 
-	if biber --nodieonerror --input-directory=$latexOutputDir --output-directory=$latexOutputDir ${latexInputFile%.*}
-	then
-		if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
-		then 
-			pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile 
-			if [[ -f $pdfOutputDir ]]
-			then
-				mv $pdfOutputDir `dirname $0`
+	if [ $1 != "-short" ] && [ $1 != "-s" ]
+	then 
+		if [ biber --nodieonerror --input-directory=$latexOutputDir --output-directory=$latexOutputDir ${latexInputFile%.*} ] 
+		then
+			if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
+			then 
+				pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile 
+				if [[ -f $pdfOutputDir ]]
+				then
+					mv $pdfOutputDir `dirname $0`
+				fi
 			fi
 		fi
 	fi

### 2020-02-18 compile.sh
bug fixing

@@ -9,9 +9,9 @@ fi
 
 if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
 then 
-	if [ $1 != "-short" ] && [ $1 != "-s" ]
+	if [ "$1" != "-short" ] && [ "$1" != "-s" ] 
 	then 
-		if [ biber --nodieonerror --input-directory=$latexOutputDir --output-directory=$latexOutputDir ${latexInputFile%.*} ] 
+		if biber --nodieonerror --input-directory=$latexOutputDir --output-directory=$latexOutputDir ${latexInputFile%.*}  
 		then
 			if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
 			then 

### 2020-02-18 compile.sh
bugfix

@@ -9,18 +9,19 @@ fi
 
 if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
 then 
-	if [ "$1" != "-short" ] && [ "$1" != "-s" ] 
+	if [ "$1" != "-short" ] && [ "$1" != "-ss" ] 
 	then 
 		if biber --nodieonerror --input-directory=$latexOutputDir --output-directory=$latexOutputDir ${latexInputFile%.*}  
 		then
 			if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
 			then 
 				pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile 
-				if [[ -f $pdfOutputDir ]]
-				then
-					mv $pdfOutputDir `dirname $0`
-				fi
 			fi
 		fi
+	fi
+
+	if [[ -f $pdfOutputDir ]]
+	then
+		mv $pdfOutputDir `dirname $0`
 	fi
 fi

### 2020-02-18 compile.sh
minor changes

@@ -9,7 +9,7 @@ fi
 
 if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
 then 
-	if [ "$1" != "-short" ] && [ "$1" != "-ss" ] 
+	if [ "$1" != "-short" ] && [ "$1" != "-s" ] 
 	then 
 		if biber --nodieonerror --input-directory=$latexOutputDir --output-directory=$latexOutputDir ${latexInputFile%.*}  
 		then

### 2020-02-18 v-research_experiment-alpha.tex
complete example on CWE-116 and review

@@
 We consider a first-order language $\follang$ over a
-signature $\Sigma_P$ where $P,P',\ldots,P^n$ represent terms, 
+signature $\Sigma_P$ where $P,P',\ldots,P^n$ represent terms, and
 $\varphi,\varphi',\ldots,\varphi^n$ represent formulas.
@@
 the application of our theory to CPS system engineering.
-More detailed case studies (i.e. \fix{mr}{a CPS as a smart power-grid}), 
+More detailed case studies (i.e. \fix{mr}{a CPS as a smart power-grid}) 
 will be defined in Section~\ref{sec:prototest}, where
@@
 
-\fix{mr}{We note that we have, so far, defined the
+We note that we have defined the
 Causality principle without considering it a \emph{threat source} since 
-we lack the concept of intent and maliciousness. We introduce those concepts
-in Section~\ref{sec:mankind-nature} and then formally define a threat agent
-in the same section.}
+we lack the concept of intent and maliciousness. Similarly, in the 
+next sections we won't discriminate between Nature and Mankind until we introduce 
+the concepts of maliciousness in Section~\ref{sec:theory} 
+and then formally define a threat agent
+in the same section.
 
@@
 on the definition of Parthood relation between \emph{Parts}.  
-However, we will need to
+However, shall 
 consider different types of Part; so, we extend the mereology to a
@@
 (whole). We aim at creating a meronomy instead of the taxonomies such as the one provided
-in\autocite{NIST2020NVD,MITRE2020CVE} or instead of the dogmatic 
+in\autocite{NIST2020NVD,MITRE2020CVE} or instead of the poorly justified 
 CVSS\autocite{Mell2007CVSS} scoring system.
@@
 protocol verification tools (which requires some formalization of a cause-effect relation, e.g.
-the Linear Temporal Logic).
+in Linear Temporal Logic).
 
@@
 structure to mereology. In its broader definition, the RCC theory is composed
-by eight axioms, and is known as RCC8, but here we restrict to RCC5 (without
+by eight axioms, and is known as RCC8. In the text, for brevity, 
+we will often focus only on RCC5 (without
 loss of generality) by not considering tangential connections between spatial
@@
 Section~\ref{sec:theory}) and we have defined them as two generic
-agents.\fixnote{mr}{check if it's a repetition and if links to sections are
-consistent}
+agents.
 
@@
 $\Weakness,\Vulnerability$ are Regions of Weaknesses and Vulnerabilities (resp.);
-accepting the hierarchy in the CWE\autocite{CWE} as a ground truth. 
+accepting the hierarchy in the CWE\autocite{CWE} as ground truth. 
 Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability) 
@@
 		\includegraphics[width=.7\textwidth]{rcc5_CWE-116.pdf}
-		\caption{Representation with RCC5}
+		\caption{Representation with RCC3}
 		\label{fig:rcc5_CWE-116}
@@
 \end{figure}
-
+Definition~\ref{def:weakness} states that CWE-116 is a weakness iff there exists a world in
+K Modal Logic, representing the system in which this weakness exists, 
+such that the natural evolution of this system (i.e. formalized by the causality principle) 
+make it possible to reach another state of the system (i.e. another, accessible world)
+where there exist a vulnerability that can be implemented from CWE-116 (and this relation 
+is not DR). The CWE website proposes the connection between CWE-116 and, for example, 
+CVE-2008-5573; a vulnerability of the login sub-system 
+of the Poll Pro v2.0\autocite{pollpro} system.
+The formal relation between the two is given as a link (i.e. URI) between the CWE and
+the CVE, the description of the relation is not defined but it is supposed to be
+inferred from the descriptions of the CWE-CVE. 
+We can formally represent this link as the ONE connection in RCC3, depicted in 
+Figure~\ref{fig:rcc5_CWE-116}, or EC connection in RCC8 (Figure~\ref{fig:rcc8_CWE-116}).
+
+\fixnote{mr}{shouldn't we consider the agent who introduces the weakness as dishonest? don't we say this in part 1?}
 \end{example}
+
+It is interesting to note that the CWE-CVE relation expresses the correlation
+between weaknesses and vulnerabilities in the most simple form, such that we
+can formalize all the relation as ONE in RCC3 or EC in RCC8.  To express a more
+complex relation between the two we shall analyze the definition of CWE-116.
+This is related to the Weakness-Vulnerability-Incident process (i.e. to the
+details on the implementation/realization in Figure~\ref{fig:safety-security})
+that we analyze in the next section, Section~\ref{sec:incidentaccident}.
 
@@
 
-\subsubsection{Process: Incidents and Accidents}\label{sec:vulnerabilitydefect}
+\subsubsection{Process: Incidents and Accidents}\label{sec:incidentaccident}
 In our informal definition depicted in Figure~\ref{fig:safety-security}, a
@@
 We describe processes, in our current formalization, in terms of the K-Modal
-Structure in Definition~\ref{def:modalinterpretation} (i.e. in terms of
+Structure as in Definition~\ref{def:modalinterpretation} (i.e. in terms of
 causality principle).
@@
 
-\subsubsection{Region: Threat and Asset}\label{sec:vulnerabilitydefect}
-
+\begin{example}
+given in natural language in the description of the 
+CWE-116 which we can analyze as follows:
+\begin{itemize}
+	\item ``the software prepares a structured message''
+	\item [the message is intended] ``for communication with another component''
+	\item ``but''
+	\item ``encoding or escaping of the data is either missing or done incorrectly''
+	\item ``[\ldots] result, the intended structure of the message is not preserved''
+\end{itemize}
+\end{example}
+
+\subsubsection{Region: Threat and Asset}\label{sec:threatasset}
 The difference between Knowledge and Belief is depicted in

### 2020-02-19 v-research_experiment-alpha.tex
extend with exploitation flow

@@
 \subsubsection{Process: Incidents and Accidents}\label{sec:incidentaccident}
-In our informal definition depicted in Figure~\ref{fig:safety-security}, a
-Vulnerability is seen as a process through which an agent (i.e. Mankind)
+In our informal definition depicted in Figure~\ref{fig:safety-security}, an
+Incident is generated through a process caused by a Vulnerability (which,
+in turn, is caused by a Weakness). Symmetrically, Nature has a process from Defect
+to Accidents. We start by analyzing the process generated by Mankind, which is
+divided into the following phases (explained in the remainder of this sub-Section):
+%Agent$\rightarrow$Vulnerability$\rightarrow$Exploit$\rightarrow$Attack$\rightarrow$Incident.
+\begin{enumerate}[noitemsep]
+	\item An agent (Mankind or Nature) defines a system, e.g.
+by following an engineering process for creating a new system, or for redesigning or improving an existing system (e.g.
+a component or a collection of components)
+	\item The dogmatic definition of a system contains a Weakness (i.e.
+the agent, Mankind or Nature, introduces an unintended behavior), for example as an un-intended behavior or as an architectural fallacy.
+The Weakness can be applied to the abstract definition of the system.
+	\item The Weakness can be defined as a malicious process for a system, called Vulnerability (or Defect for Nature)
+	\item The process expressed by the Vulnerability can be made operational for a specific implementation of the system (i.e. implemented by Mankind or realized by Nature) 
+\end{enumerate}
+Those additional details are depicted in Figure~\ref{fig:safety-security_2}, an update
+of Figure~\ref{fig:safety-security}.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\textwidth]{safety-security_2.pdf}
+	\caption{Overview security and safety keywords}
+	\label{fig:safety-security_2}
+\end{figure}
+
+\Paragraph{Exploitation Flow} An \emph{Exploit} is an \emph{implementation} of
+a Vulnerability, where a Vulnerability is one of the many ways to transform a
+Weakness into a malicious process that can be applied\footnote{Meaning that the
+malicious process, i.e.  any structured procedure such as a protocol, can be
+somehow made operational, e.g. implemented} to one or many systems. So, the act
+of an agent (Mankind) of ``exploiting a Vulnerability'' is the process of
+making the Vulnerability operational, by implementing the Vulnerability w.r.t.
+a specific target system.  Whenever the agent is Nature, the implementation is
+more broadly considered a realization\footnote{``The state of being
+realized''\autocite{Merriam2020realization}}
+
+The Common Attack Pattern Enumeration and
+Classification\autocite{MITRE2020CAPEC} (CAPEC) online service provides a
+``community resource'' which aims at ``identifying and understanding attacks''.
+The MITRE writes on the CAPEC homepage that ``Understanding how the adversary
+operates is essential to effective cyber security. CAPEC helps by providing a
+comprehensive dictionary of known patterns of attack employed by adversaries to
+exploit known weaknesses in cyber-enabled capabilities. It can be used by
+analysts, developers, testers, and educators to advance community understanding
+and enhance defenses.''. We can extrapolate two major concepts: 
+\begin{enumerate}
+	\item while understanding the exploitation process is the aim of the
+		CWE and CVE initiatives, understanding the hacking process is
+		the aim of the CAPEC initiative.
+	\item CAPEC provides a \emph{dictionary} of \emph{known} attack patterns.
+		As stated largely in Section~\ref{sec:}, this dictionary contains
+		empirical evidences and should not be used to induce a cybersecurity 
+		theory; but it can be used to validate a theory.
+\end{enumerate}
+
+The process continues as follows.
+\begin{enumerate}[noitemsep]
+\setcounter{enumi}{4}
+	\item an agent (Mankind\footnote{For the sake of readability we focus only on Mankind as an agent}) the exploitation flow can be used to hack a system.
+We use the term hack (as hacking) to stress that nothing prevents this process from being honest, however, for the sake of simplicity, we focus on Attacks which are malicious by definition.
+	\item An Attack is applied to a system with malicious intent by an agent
+	\item The application of the attack results in an Incident which impacts, i.e. poses a Threat to, an Asset
+\end{enumerate}
+
+\Paragraph{Hacking Flow}
+
+\begin{example}
+given in natural language in the description of the 
+CWE-116 which we can analyze as follows:
+\begin{itemize}
+	\item ``the software prepares a structured message''
+	\item the message is intended ``for communication with another component''
+	\item ``but''
+	\item ``encoding or escaping of the data is either missing or done incorrectly''
+	\item ``[\ldots] result, the intended structure of the message is not preserved''
+\end{itemize}
+\end{example}
+
+In this Vulnerability is seen as a process through which an agent (i.e. Mankind)
 generates an Incident, i.e.
-Agent$\rightarrow$Vulnerability$\rightarrow$Exploit$\rightarrow$Attack$\rightarrow$Incident.
+
 We describe processes, in our current formalization, in terms of the K-Modal
@@
 \end{figure}
-
-\begin{example}
-given in natural language in the description of the 
-CWE-116 which we can analyze as follows:
-\begin{itemize}
-	\item ``the software prepares a structured message''
-	\item [the message is intended] ``for communication with another component''
-	\item ``but''
-	\item ``encoding or escaping of the data is either missing or done incorrectly''
-	\item ``[\ldots] result, the intended structure of the message is not preserved''
-\end{itemize}
-\end{example}
 

### 2020-02-19 v-research_experiment-alpha.tex
refine description of exploitation and hacking processes

@@
 in\autocite{Herley2016unfalsifiability}.
+
+Another way of looking at the same problem is by analyzing the different
+definitions of security with respect to the scientific method of enquiry
+applied to get to the definition itself. In the following we categorize the
+related work into three categories, as defined by Sextus Empiricus in
+\autocite{Empiricus1990Pyrrhonism}.  To the best of our knowledge, there is no
+evidence that this categorization isn't complete.  ``The natural result of any
+investigation is that the investigators either discover the object of search or
+deny that it is discoverable and confess it to be in-apprehensible or persist
+in their search. [\ldots] This is probably
+why''\autocite{Empiricus1990Pyrrhonism}: 
+\begin{itemize}
+	\item The \emph{dogmatists} ``have claimed to have discovered the truth'' on what cybersecurity is
+		\begin{itemize}
+			\item Wikipedia defines cybersecurity in
+				\autocite{wiki-cybersecurity} as the protection
+				of computer systems and networks from the theft
+				of or damage to their hardware, software, or
+				electronic data, as well as from the disruption
+				or misdirection of the services they provide
+		\end{itemize}
+	\item The \emph{academics} ``have asserted that it cannot be apprehended''
+		\begin{itemize}
+			\item Eugene H. Spafford, Professor at Purdue
+				University, defines cybersecurity as follow.
+				``The only truly secure system is one that is
+				powered off, cast in a block of concrete and
+				sealed in a lead-lined room with armed guards —
+				and even then I have my doubts.''
+				\autocite{Spafford2019Quotes}
+		\end{itemize}
+	\item The \emph{skeptics} ``go on inquiring''
+		\begin{itemize}
+			\item Cormac Herley reaches the conclusion that
+				cybersecurity has no definition. ``There is an
+				inherent asymmetry in computer security: things
+				can be declared insecure by observation, but
+				not the reverse. There is no test that allows
+				us to declare an arbitrary system or technique
+				secure. This implies that claims of necessary
+				conditions for security are unfalsifiable. ''
+				\autocite{Herley2016unfalsifiability}
+		\end{itemize}
+\end{itemize}
 
@@
 to Accidents. We start by analyzing the process generated by Mankind, which is
-divided into the following phases (explained in the remainder of this sub-Section):
-%Agent$\rightarrow$Vulnerability$\rightarrow$Exploit$\rightarrow$Attack$\rightarrow$Incident.
-\begin{enumerate}[noitemsep]
+divided into the following phases (as shown in Figure~\ref{fig:cwe-nvd-capec}):
+\begin{enumerate}
 	\item An agent (Mankind or Nature) defines a system, e.g.
@@
 	\includegraphics[width=\textwidth]{safety-security_2.pdf}
-	\caption{Overview security and safety keywords}
+	\caption{Updated overview security and safety keywords, where Process
+	has to be intended as ``Abstract Exploitation Process''}
 	\label{fig:safety-security_2}
@@
 
-\Paragraph{Exploitation Flow} An \emph{Exploit} is an \emph{implementation} of
-a Vulnerability, where a Vulnerability is one of the many ways to transform a
-Weakness into a malicious process that can be applied\footnote{Meaning that the
+\begin{definition}{\bf Exploitation Process --} An \emph{Exploit} is an \emph{implementation} of
+a Vulnerability, where a Vulnerability describes how to transform a
+Weakness into an abstract malicious process, which can be applied\footnote{Meaning that the
 malicious process, i.e.  any structured procedure such as a protocol, can be
@@
 more broadly considered a realization\footnote{``The state of being
-realized''\autocite{Merriam2020realization}}
+realized''\autocite{Merriam2020realization}}. 
+\end{definition}
 
@@
 and enhance defenses.''. We can extrapolate two major concepts: 
-\begin{enumerate}
-	\item while understanding the exploitation process is the aim of the
+\begin{itemize}
+	\item While understanding the exploitation process is the aim of the
 		CWE and CVE initiatives, understanding the hacking process is
@@
 	\item CAPEC provides a \emph{dictionary} of \emph{known} attack patterns.
-		As stated largely in Section~\ref{sec:}, this dictionary contains
+		As stated in details in Section~\ref{sec:problem}, this dictionary contains
 		empirical evidences and should not be used to induce a cybersecurity 
 		theory; but it can be used to validate a theory.
+\end{itemize}
+
+\begin{example}
+The CWE-116 doesn't just relate the Weakness to a number of Vulnerability in
+	the CVE but also relates the Weakness to a number ($4$ for the CWE-116)
+	of abstract attack path in the CAPEC:
+	\begin{itemize}
+		\item CAPEC-104\autocite{CAPEC-104} Cross Zone Scripting
+		\item CAPEC-73\autocite{CAPEC-73} User-Controlled Filename
+		\item CAPEC-81\autocite{CAPEC-81} Web Logs Tampering
+		\item CAPEC-85\autocite{CAPEC-85} AJAX Fingerprinting
+	\end{itemize}
+
+	Each CAPEC entry has a reference to one or many CWE, including CWE-116.
+\end{example}
+
+After the exploitation process, as depicted in Figure~\ref{fig:cwe-nvd-capec},
+another process starts and transforms a Vulnerability in an Incident by
+Attacking the system with a malicious Hack, as follows.
+\begin{enumerate}
+\setcounter{enumi}{4}
+	\item An agent (Mankind\footnote{For the sake of readability we focus
+		only on Mankind as an agent}) the exploitation process can be
+		used to hack a system. We use the term hack (as hacking) to
+		stress that nothing prevents this process from being honest,
+		however, for the sake of simplicity, we focus on Attacks which
+		are malicious by definition.
+	\item An Attack is applied to a system with malicious intent by an
+		agent
+	\item The application of the attack results in an Incident which
+		impacts, i.e. poses a Threat to, an Asset
 \end{enumerate}
 
-The process continues as follows.
-\begin{enumerate}[noitemsep]
-\setcounter{enumi}{4}
-	\item an agent (Mankind\footnote{For the sake of readability we focus only on Mankind as an agent}) the exploitation flow can be used to hack a system.
-We use the term hack (as hacking) to stress that nothing prevents this process from being honest, however, for the sake of simplicity, we focus on Attacks which are malicious by definition.
-	\item An Attack is applied to a system with malicious intent by an agent
-	\item The application of the attack results in an Incident which impacts, i.e. poses a Threat to, an Asset
+\begin{definition}{\bf Hacking Process}\label{def:hackingprocess}
+An \emph{Attack} is the act of making operational an \emph{Exploitation Process} 
+	causing an Incident. \fix{mr}{Threat? Asset?}
+\end{definition}
+
+\begin{example}	CAPEC-81\autocite{CAPEC-81} Web Logs Tampering --
+	\begin{itemize}
+		\item Description: An attacker is able to cause a victim to
+			load content into their web-browser that bypasses
+			security zone controls and gain access to increased
+			privileges to execute scripting code or other web
+			objects such as unsigned ActiveX controls or applets.
+			This is a privilege elevation attack targeted at
+			zone-based web-browser security. In a zone-based model,
+			pages belong to one of a set of zones corresponding to
+			the level of privilege assigned to that page. Pages in
+			an untrusted zone would have a lesser level of access
+			to the system and/or be restricted in the types of
+			executable content it was allowed to invoke. In a
+			cross-zone scripting attack, a page that should be
+			assigned to a less privileged zone is granted the
+			privileges of a more trusted zone. This can be
+			accomplished by exploiting bugs in the browser,
+			exploiting incorrect configuration in the zone
+			controls, through a cross-site scripting attack that
+			causes the attackers' content to be treated as coming
+			from a more trusted page, or by leveraging some piece
+			of system functionality that is accessible from both
+			the trusted and less trusted zone. This attack differs
+			from "Restful Privilege Escalation" in that the latter
+			correlates to the inadequate securing of RESTful access
+			methods (such as HTTP DELETE) on the server, while
+			cross-zone scripting attacks the concept of security
+			zones as implemented by a browser. 
+		\item Execution Flow: 
+			\begin{enumerate}
+				\item Explore: Find systems susceptible to the
+					attack: Find systems that contain
+					functionality that is accessed from
+					both the internet zone and the local
+					zone. There needs to be a way to supply
+					input to that functionality from the
+					internet zone and that original input
+					needs to be used later on a page from a
+					local zone. 
+				\item Experiment: Find the insertion point for
+					the payload: The attacker first needs
+					to find some system functionality or
+					possibly another weakness in the system
+					(e.g. susceptibility to cross site
+					scripting) that would provide the
+					attacker with a mechanism to deliver
+					the payload (i.e. the code to be
+					executed) to the user. The location
+					from which this code is executed in the
+					user's browser needs to be within the
+					local machine zone.
+				\item Exploit: Craft and inject the payload:
+					Develop the payload to be executed in
+					the higher privileged zone in the
+					user's browser. Inject the payload and
+					attempt to lure the victim (if
+					possible) into executing the
+					functionality which unleashes the
+					payload.
 			\end{enumerate}
-
-\Paragraph{Hacking Flow}
-
-\begin{example}
-given in natural language in the description of the 
-CWE-116 which we can analyze as follows:
+		\item Prerequisite: The target must be using a zone-aware browser. 
+		\item Consequences: 
 			\begin{itemize}
-	\item ``the software prepares a structured message''
-	\item the message is intended ``for communication with another component''
-	\item ``but''
-	\item ``encoding or escaping of the data is either missing or done incorrectly''
-	\item ``[\ldots] result, the intended structure of the message is not preserved''
-\end{itemize}
+				\item Integrity: Modify Data
+				\item Confidentiality: Read Data
+				\item Confidentiality, Access Control, Authorization: Gain Privileges
+				\item Confidentiality, Integrity, Availability: Execute Unauthorized Commands
+			\end{itemize}
+	\end{itemize}
+
 \end{example}
 
-In this Vulnerability is seen as a process through which an agent (i.e. Mankind)
-generates an Incident, i.e.
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\textwidth]{cwe-nvd-capec.pdf}
+	\caption{Cybersecurity online dictionaries and connections}
+	\label{fig:cwe-nvd-capec}
+\end{figure}
+
+The formalization of the Exploitation and Hacking processes is the
+formalization of two foundational cybersecurity concepts which lay the basis
+for the definition (and formalization) of the cybersecurity theory in the next
+Section (Section~\ref{sec:theory}). However, the formal definition of those
+processes shall (i) correlate the formal definition of agents (i.e. the
+Definition~\ref{def:agent}) with the Causality principle (i.e. the
+formalization of the Kripke structure in Definition~\ref{def:causality}), and,
+most importantly, (ii) formally define a system as a structure of agents. For
+the sake of clarity, we then postpone the formalization of the two processes to
+the next Section.
 
@@
 \begin{definition}{\bf Intent}\label{def:intent}
-\end{definition}
-
-\subsection{Vulnerability and Defect}\label{sec:vulnerability-defect}
-In order to formalize the semantics of the concepts (informally) described in
-Section~\ref{sec:problem},
-
-\begin{definition}{\bf Vulnerability}
 \end{definition}

### 2020-02-20 compile.sh
change filename to main.tex

@@ -1,6 +1,6 @@
 #!/bin/bash
 latexOutputDir="latex-output-dir"
-latexInputFile="v-research_experiment-alpha.tex"
+latexInputFile="main.tex"
 pdfOutputDir="${latexOutputDir}/${latexInputFile%.*}.pdf"
 
 if [[ ! -d $latexOutputDir ]]; then 

### 2020-02-20 v-research_experiment-alpha.tex
make git happy

@@
 An \emph{Attack} is the act of making operational an \emph{Exploitation Process} 
-	causing an Incident. \fix{mr}{Threat? Asset?}
+	causing an Incident. 
 \end{definition}
@@
 Definition~\ref{def:agent}) with the Causality principle (i.e. the
-formalization of the Kripke structure in Definition~\ref{def:causality}), and,
+formalization of the Kripke structure in Definition~\ref{def:modallogic}), and,
 most importantly, (ii) formally define a system as a structure of agents. For
@@
 the next Section.
+
+\section{Honesty and Maliciousness -- A Cybersecurity Theory}\label{sec:theory}
+\begin{enumerate}[noitemsep]
+	\item cybersecurity as a requirement
+	\item vulnerability\label{def:vulnerability}
+\end{enumerate}
 
@@
 	\item DO-326A
-\end{enumerate}
-
-\section{Honesty and Maliciousness -- A Cybersecurity Theory}\label{sec:theory}
-\begin{enumerate}[noitemsep]
-	\item cybersecurity as a requirement
-	\item vulnerability\label{def:vulnerability}
 \end{enumerate}

### 2020-02-20 abstract.tex
restructure

--- 2020-02-19 v-research_experiment-alpha.tex
@@
-\documentclass{article}
-\usepackage{mathtools}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{enumitem}
-\usepackage{multicol}
-\usepackage[toc,page]{appendix}
-\usepackage[gen]{eurosym}
-\usepackage{tabularx}
-\usepackage{epigraph}
-\usepackage{textcomp}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{csquotes}% Recommended
-\usepackage{rotating}
-\usepackage[style=numeric]{biblatex}
-% \bibliography{<mybibfile>}% ONLY selects .bib file; syntax for version <= 1.1b
-\addbibresource{literature.bib}% Syntax for version >= 1.2
-
-\usepackage{graphicx}
-\graphicspath{{./images/}}
-
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\lhead{}
-\chead{}
-\rhead{}
-\lfoot{}
-\cfoot{\thepage}
-\rfoot{Property of V-Research S.r.l.}
-\renewcommand\headrulewidth{0pt}
-\renewcommand\footrulewidth{0pt}
-
-\usepackage[colorinlistoftodos]{todonotes}
-
-\newcommand{\Paragraph}[1]{\smallskip\noindent{\bf #1.}}
-\newcommand{\fixnote}[2]{\textbf{\color{red}{FIX}}\footnote{{\bf #1:} #2}}
-\newcommand{\fix}[2]{{\color{red} {\bf tofix:} #2}}
-% \renewcommand{\fixnote}[2]{}
-% \renewcommand{\fix}[2]{}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\theoremstyle{corollary}
-\newtheorem{corollary}{Corollary}[section]
-\theoremstyle{lemma}
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{theorem}
-\newtheorem{theorem}{Theorem}[section]
-\theoremstyle{theorem}
-\newtheorem{example}{Example}
-
-\usepackage{authblk}
-\date{}                     %% if you don't need date to appear
-\setcounter{Maxaffil}{0}
-\renewcommand\Affilfont{\itshape\small}
-
-% for fancy table with rotated column titles
-\usepackage{adjustbox}
-\newcolumntype{R}[2]{%
-    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
-    l%
-    <{\egroup}%
-}
-\newcommand*\rot{\multicolumn{1}{R{45}{0.01em}}}% no optional argument here, please!
-\newcommand*\rota{\multicolumn{1}{R{90}{0.01em}}}% no optional argument here, please!
-\def\hb{\hbox to 10.7 cm{}}
-
-% circles
-\usepackage{wasysym}
-\newcommand{\Tdot}{$\CIRCLE$}
-\newcommand{\Thdot}{$\LEFTcircle$}
-\newcommand{\Twdot}{$\Circle$}
-
-% formulas
-\newcommand{\kframe}{\mathbf{K}}
-\newcommand{\bframe}{\mathbf{B_K}}
-\newcommand{\possibleworlds}{G}
-\newcommand{\modalrelation}{R}
-\newcommand{\actualworld}{\omega*}
-\newcommand{\world}{\omega}
-\newcommand{\World}{\Omega}
-\newcommand{\kmodel}{M}
-\newcommand{\interpretation}{\sigma}
-\newcommand{\knowledge}[1]{\mathbb{K}_{#1}}
-\newcommand{\knows}[2]{K_{#1}#2}
-\newcommand{\belief}[1]{\mathbb{B}_{#1}}
-\newcommand{\believe}[2]{B_{#1}#2}
-\newcommand{\region}{\chi}
-\newcommand{\Region}{\mathrm{X}}
-\newcommand{\agentuniverse}{\mathcal{A}}
-\newcommand{\agent}{a}
-\newcommand{\follang}{\mathcal{L}}
-\newcommand{\weakness}{w}
-\newcommand{\Weakness}{W}
-\newcommand{\vulnerability}{v}
-\newcommand{\Vulnerability}{V}
-%rcc5
-\newcommand{\rcc}{rcc}
-\newcommand{\connects}[2]{C(#1,#2)}
-\newcommand{\disconnected}[2]{\neg C(#1,#2)}
-\newcommand{\partof}[2]{P(#1,#2)}
-\newcommand{\overlaps}[2]{O(#1,#2)}
-\newcommand{\eq}[2]{EQ(#1,#2)}
-\newcommand{\pp}[2]{PP(#1,#2)}
-\newcommand{\po}[2]{PO(#1,#2)}
-\newcommand{\ppi}[2]{PPi(#1,#2)}
-\newcommand{\dr}[2]{DR(#1,#2)}
-
-\makeindex
-
-\begin{document}
-
-\title{On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}}
-\author[1]{Francesco Beltramini}
-\author[1]{Marco Rocchetto}
-\author[2]{Luca Vigan\`o}
-\affil[1]{V-Research, Verona, Italy}
-\affil[2]{King's College London, UK}
-
-\maketitle
-
-\begin{abstract}
 The objective of this research is to develop of a theory that defines
@@
 secure systems (e.g. purely cyber or cyber-physical).
-\end{abstract}
-\newpage
-
-\section{Introduction}\label{sec:intro}
-\epigraph{Humanum est errare}{{\itshape Seneca the Elder}}
-The European Commission states in\autocite{EU2019market} that: ``Cybersecurity
-is one of the priority areas [\ldots] of the Commission initiative on ICT
-Standards, which is part of the Digitising European
-Industry\autocite{EU2019standard} strategy launched on 19 April 2016.  The aim
-is to identify the essential ICT standards and present measures to accelerate
-their development in support of digital innovations across the economy''. The
-same document (i.e.\autocite{EU2019market}) states that ``The EU will invest up
-to \euro450 million [\ldots], under its research and innovation programme
-Horizon 2020''. The EU, in 2016 published a press release\autocite{EU2016press}
-in which they present a strategy to invest \euro1.8 \emph{billion} to
-``increase measures to address cyber threats''. The EU is not the only investor
-in cybersecurity, most of the developed countries and several companies are
-investing enormous amount of money towards various aspects of cybersecurity
-(e.g. The US vulnerability databases\autocite{NIST2020NVD} maintained by the
-National Institute of Standards and Technologies, i.e. NIST, of the US
-Department of Commerce).
-
-The cybersecurity industry is growing fast, e.g. as reported
-in\autocite{Nasdaq2018market}. For example, in\autocite{Forbes2017market},
-published by the Forbes, is stated that \euro5.3 billion of funding where
-poured by venture capitalist into cybersecurity companies in 2018. The Forbes,
-in the same article, also highlights another peculiar (as seemly contradictory)
-trend: ``[\ldots] during the same time period, the number of cybersecurity
-breaches increased exponentially''. The data reported by the NIST 
-through the official CPE (Common Platform Enumeration) Dictionary Statistics 
-on the NVD websites in \autocite{NIST2020CPEstatistics}, show that in 2016 the
-number of reported vulnerabilities reported where around $~6000$ while in 2019
-the number of vulnerabilities was above $16000$.  The scientific community also
-reports similar findings.  In fact, in\autocite{Herley2009so}, Cormac Herley
-(Microsoft Research) shows how basic cybersecurity principles (such as the
-confidentiality benefit over the clear text for passwords typed into forms,
-e.g. for logins in websites) are not fully understood or shared between the
-cybersecurity research community\autocite{Nielsen2009stop}.  The lack of
-understanding of basic security principle, the inverse proportionality between
-investments in cybersecurity and the number of reported vulnerabilities year
-after year, can be linked to the lack of a foundational theory on
-cybersecurity, as already highlighted by Cormac Herley
-in\autocite{Herley2016unfalsifiability}.
-
-In this article, we give the first scientific theory (to the best of our knowledge)
-on security.
-
-\Paragraph{Structure} In Section~\ref{sec:problem} we define and formalize the problem statement. 
-In Section~\ref{sec:theory} we outline our security theory, and in Section~\ref{sec:prototest}
-we describe the implementation of the theory and some empirical tests of the theory.
-Finally, in Section~\ref{sec:related} we conclude the paper with an overview of the related work.
-
-\section{Problem Statement}\label{sec:problem}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.6\textwidth]{protocol-example.pdf}
-	\caption{Abstraction of an ad-hoc esemplificative protocol execution}
-	\label{fig:protocol-example}
-\end{figure}
-
-In\autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
-``an asymmetry in computer security'', which he defines as follows: ``Things
-can be declared insecure by observation, but not the reverse. There is no
-observation that allows us to declare an arbitrary system or technique
-secure''. Herley then uses this argument to show that ``claims that any measure
-is necessary for security are empirically unfalsifiable''. Given that, any
-theory which is not falsifiable by an empirical experiment is well
-known\footnote{``A theory which is not refutable by any conceivable event is
-nonscientific. Irrefutability is not a virtue of a theory (as people often
-think) but a vice.'' -- Karl Popper, Conjectures and
-Refutations\autocite{popper1962conjectures}} to be nonscientific (i.e.
-unfalsifiability is a fallacy of a theory), Herley concludes that there is no
-scientific theory on cybersecurity; which means that cybersecurity lays in the
-realm of pseudo-sciences\autocite{Herley2016usenixvideo}.  Herley, e.g.
-in\autocite{Herley2017justifying}, discusses the implications of a
-nonscientific approach to cybersecurity, and highlights the tremendous impact on
-all the scientific research and engineering of systems; leading often to
-terrorism and wars, and wasting of resources in useless protections or
-overspending.  While the criticism is investigated
-in\autocite{Herley2016unfalsifiability}, no solution is provided.  On the
-contrary, the goal of this work \emph{is} to lay the foundations of a
-scientific cybersecurity theory. Furthermore, in Section~\ref{sec:sicurezza},
-we consider the problem raised by Herley not confined to ``computer security''
-but to any abstract system (so that our theory may hold for any sound
-implementation such as networks, mechanical, cyber, or cyber-physical system,
-or even a single computer or a single device such as an hard-drive).  There is
-also an apparent inconsistency in\autocite{Herley2016unfalsifiability} that we
-seek to clarify before following (as we agree) the scientific path draw by
-Herley: cybersecurity is defined as an abstract property in many formal
-approaches to the investigation of the security of systems, and the security of the design of a
-formally verified protocol is indeed falsifiable.  For example, in the protocol
-verification community, security is often defined as a formalization of the
-high-level properties confidentiality, integrity, and availability. The problem
-in such approaches is not the definition of what cybersecurity is, but the use
-of theories (such as the Dolev-Yao attacker model\footnote{For the sake of
-simplicity, the Dolev-Yao attacker can be considered as an abstraction of an
-active attacker who controls the network but cannot break
-cryptography.}\autocite{Dolev1983security})
-that only applies to specific instances (often called scenarios) and
-abstraction of the protocol. This, in turn, creates a false sense of security
-since requires assumptions on the abstraction of the system of which security
-is verified. As an example, for the formal security verification of the system
-in Figure~\ref{fig:protocol-example}, a formalized scenario needs to be defined
-by a modeler who chooses (among others): (i) a scope of the formalization (e.g.
-excluding the server that distributes the public key is often done when
-verifying the security of authentication protocols), (ii) the number of
-sessions (even tough some approaches do reason on an infinite number of
-sessions such as\autocite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
-the peers (e.g.  in the ASLan++ language\autocite{Oheimb2010aslan++}), and (iv)
-the abstraction of the cryptographic primitives (e.g.  ProVerif vs
-CryptoVerif\autocite{Blanchet2017symbolic}).  Some of the choices will
-completely change the results of the formal verification of the system. For
-example, under the perfect cryptography assumption\footnote{As defined
-in\autocite{Rocchetto2016cpdy}: ``In the so called perfect cryptography
-assumption, the security encryption scheme is suppose to be perfect, without
-any exploitable flaw, and so the only way for the attacker to decrypt a message
-is by using the proper key. That assumption is widely accepted in the security
-protocol community, and most of the formal reasoning tools for the analysis of
-security protocols abstract away the mathematical and implementation details of
-the encryption
-scheme\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
-and assuming that no violation to any security property is done after message
-I); in Figure~\ref{fig:protocol-example}, the freedom of choosing the scope
-determines that the flaws related to the dishonest impersonation of the Server
-may or may not be considered in the verification process.  This choice has
-tremendous impact on the focus and findings of the verification of the security
-of the protocol.  While this may seem to turn upon minutiae and foreseeable,
-this highlights the false sense of security that may derive from a
-non-scientific theory of system security\footnote{``To the superficial
-observer, the analysis of these forms seems to turn upon minutiae. It does in
-fact deal with minutiae, but they are of the same order as those dealt with in
-microscopic anatomy.'' -- Karl Marx, Capital Volume 1, 1867}.
-
-\subsection{Sicurezza: Safety and Security}\label{sec:sicurezza}
-In most of the natural languages, and in Italian too, the concepts of safety
-and security are not syntactically differentiated and both terms (safety and
-security) are expressed by the same word, e.g. sicurezza in Italian.  A
-semantic distinction between safety and security is correlated to a
-belief\footnote{A belief has to be intended as a proposition which is supposed
-to be true by the majority of humans in our society without a scientific
-underlying theory but based on partial empirical evidences or inductive proofs.} that
-safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
-natural environment (e.g. natural events such as wearing of hardware
-components) while security deals with \emph{incidents} posed by mankind (e.g.
-attackers and bugs).  The fundamental difference between nature and mankind (and,
-in turn, between safety and cybersecurity) is believed to be on the different
-intents\footnote{``The belief–desire–intention software model (BDI) is a
-software model developed for programming intelligent
-agents.''\autocite{wiki-bdi}. In the BDI model, the intents represents the
-deliberative state of an agent which determines the choice of that agent on
-what to do.} (accidents are unfortunate while incidents are not) of the causes
-that generates the threat; namely, nature is believed not to have malicious
-intents (but unfortunate causes-effects) while threats generated by mankind are
-believed to be malicious\footnote{Of course, logical flaws or bugs may be
-introduced by other means (e.g. ignorance) without explicit malicious intents,
-but the exploitation of those flaws is considered (for now, and detailed
-afterwards in the article) malicious, and then we consider any vulnerability to
-be malicious (without loss of generality) even if due to the lack of skills.}.
-An overview on the aforementioned aspects of safety and security is depicted in
-Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
-the terms that structure our current understanding of safety and security. 
-\begin{itemize}
-	\item \emph{Mankind} ``refers collectively to humans''
-\autocite{wiki-mankind}, while the concept of \emph{Nature} is
-		related ``to the intrinsic characteristics that plants,
-		animals, and other features of the world develop of their own
-		accord'' (e.g. the physical universe)\autocite{wiki-nature}. 
-		\begin{itemize}
-			\item So far, we have used several terms to refer to an
-				\emph{attacker}, i.e. threat agent or threat source,
-				considering those terms to be semantically
-				equivalent.  This ``shallowness'' raise form the
-				necessity of properly citing the different sources, but,
-				in the reminder of this paper, we consider the
-				Causality principle to be the \emph{threat
-				source}, Nature or Mankind to be the
-				\emph{threat agents} and an \emph{attacker} as
-				a specific malicious threat agent which materialize a
-				threat.
-		\end{itemize}
-	\item \emph{Vulnerability}\footnote{The term vulnerability is not
-		present in the Encyclopedia of Cryptography and Security, while
-		it is used in 12 entries (such as in the definition of
-		``penetration testing''\autocite{caddy2005pentest})
-		highlighting how commonly this word is used without a proper
-		supporting semantics}, as defined in\autocite{cnssi20104009}
-		(and adopted in\autocite{nist2013800-53}), is ``weakness in an
-		information system, system security procedures, internal
-		controls, or implementation that could be exploited by a threat
-		source''. On the one hand, the definition is broad to enclose
-		as much causes (that generates a vulnerability) as possible; on
-		the other hand, it derives from empirical evidences (which
-		should be considered beliefs\footnote{``For this view, that
-		\emph{That Which Is Not} exists, can never predominate. You
-		must debar your thought from this way of search, nor let
-		ordinary experience in its variety force you along this way,
-		(namely, that of allowing) the eye, sightless as it is, and the
-		ear, full of sound, and the tongue, to rule; but (you must)
-		judge by means of the Reason (Logos) the much-contested proof
-		which is expounded by me.'' -- Parmenides of Elea, On Nature
-		(circa 500 B.C.), fragments B7.1–8.2
-		\autocite{Hakim2016philosophy}} since they are partial results in nature) 
-		while a vulnerability should
-		be defined in a way that is empirically falsifiable. This means
-		that the term vulnerability should have a complete and sound
-		definition, so that no other causes (e.g.  other sources) but
-		the ones in the definition are responsible for a vulnerability.
-		Furthermore, the term ``threat sources'' used in the definition
-		in\autocite{cnssi20104009} may be identified with both Nature
-		and Mankind, not differentiating between safety and security.
-		In Definition~\ref{def:vulnerability}, we provide a formal
-		theory of vulnerability (so that the scientific community can
-		identify tests for the completeness and soundness of the
-		definition itself).
-\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{safety-security.pdf}
-	\caption{Overview security and safety keywords}
-	\label{fig:safety-security}
-\end{figure}
-
-Most of the safety-preserving principles in the field of engineering of
-safety-critical cyber-physical systems (such as elevators and aircraft), upon
-which safety requirements (e.g. in standards such as the IEC 61508 or 61511\autocite{IEC201761511}) are defined, have
-been defined following empirical tests and measurements. While reasoning by
-induction based on the empirical observation should be avoided, since it may
-easily lead to false beliefs instead of scientific theories, this approach is
-often justified by the supposed impossibility of defining a theory that
-correctly predicts failures which, in turn, pose hazards to a system. 
-To the best of our knowledge, and supported by\autocite{Herley2016unfalsifiability}, the
-correlation between predictability of environment and believed unpredictability
-of attackers (i.e. a malicious
-environment) has not been correlated to a theory on cybersecurity. 
-Therefore, inductive research efforts in
-predicting malicious effects are accepted (and published) in scientific
-conferences (e.g.~\autocite{Rocchetto2014CSRF}). A failure of a wire due to
-environment (e.g. due to humidity, dust, heat \&c) is defined from empirical
-evidences and processes have been standardized to test qualities of hardware components
-%convinced by the induction principle. What is already a
-%weak argument in general, 
-This process completely breaks down when a malicious environment (i.e. an attacker)
-is considered instead of the (supposedly honest and predictable)
-natural environment. Therefore, the same approach that is in use for safety,
-seems not to be applicable to test security.
-
-Going back to Figure~\ref{fig:safety-security}, a vulnerability does not
-necessarily become a threat for the system, unless exploited ``through a
-channel that allows the violation of the security policy
-[\ldots]''\autocite{cnssi20104009} (e.g. a software or procedure) that takes
-advantage of the vulnerability causing an \emph{attack} to the system, which may
-result in several correlated incidents and threats.  The process of
-exploitation of a defect as a vulnerability is reported in
-Figure~\ref{fig:safety-security} such that the difference between exploit and failure,
-and attack and accident is to be found just in the maliciousness of the intents
-that causes this process (i.e. excluding the intent, the terms are just syntactic transformation from a vulnerability to defect, from
-accident to incident). In the following, we conclude the informal definition of
-the terms that we used in this section and in Figure~\ref{fig:safety-security}.
-
-\begin{itemize}
-	\item \emph{Causality} refers to the causality principle; defined
-		in\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
-		connection of phenomena through which one thing (the cause)
-		under certain conditions gives rise to, causes something else
-		(the effect). The essence of causality is the generation and
-		determination of one phenomenon by another. In this respect
-		causality differs from various other kinds of connection, for
-		example, the simple temporal sequence of phenomena, of the
-		regularities of accompanying processes''.
-	\item An \emph{Exploit}\footnote{We note that the term exploit is only
-		used as a verb in\autocite{ISO2009information}} is ``An exploit
-		(from the English verb to exploit, meaning to use something to
-		one’s own advantage) is a piece of software, a chunk of data,
-		or a sequence of commands that takes advantage of a bug or
-		vulnerability to cause unintended or unanticipated behavior to
-		occur on computer software, hardware, or something electronic
-		(usually computerized).''\autocite{wiki-exploit}.
-	\item An \emph{Attack}, as defined by the International Standard
-		ISO/IEC 27000 is an ``attempt to destroy, expose, alter,
-		disable, steal or gain unauthorized access to or make
-		unauthorized use of an asset''; where an \emph{Asset} is
-		``anything that has value to the organization''. We note that for
-		the purpose of this article, we do not want to focus on a specific
-		organization or business to define asset but, in general, on any 
-		abstract organization (e.g. a company or a society).
-		We do not consider ethical hackers as attacking a system. 
-		In fact, we consider the term \emph{hack} as
-		non-malicious (see Hacker\autocite{Stallman2002hacker}).
-	\item A \emph{Threat}, as defined in\autocite{cnssi20104009}, is ``Any
-		circumstance or event with the potential to adversely impact
-		organizational operations (including mission, functions, image,
-		or reputation), organizational assets, individuals, other
-		organizations, or the Nation through an information system via
-		unauthorized access, destruction, disclosure, modification of
-		information, and/or denial of service''.
-	\item \emph{Defect}, ``anything that renders the product not reasonably
-		safe''\autocite{Robinson2019writing} (i.e. a characteristic of
-		an object which hinders its proper usability).
-	\item \emph{Failure}, as defined in\autocite{Merriam2020failure} as ``a state of
-		inability to perform a normal function''. The term is
-		structured and detailed in
-		\autocite{cnssi20104009,iet2017glossary} but relying on an
-		abstract notion of failure without a specific definition.
-	\item \emph{Hazard}, ``a potential source of
-		harm''\autocite{iet2017glossary}.
-\end{itemize}
-
-\subsection{Glossary -- A Formalization}\label{sec:glossary}
-We now define a formalizations of the concepts described in
-Section~\ref{sec:problem} and depicted in Figure~\ref{fig:safety-security}. We
-base our formalization on first order logic (FOL) with the standard
-truth-value semantics. The choice of this logic is required by the 
-semantics of the concepts formalized afterwards in this section.
-
-We consider a first-order language $\follang$ over a
-signature $\Sigma_P$ where $P,P',\ldots,P^n$ represent terms, and
-$\varphi,\varphi',\ldots,\varphi^n$ represent formulas.
-The syntax is defined as follows.
-\begin{displaymath}
-	\varphi := P~|~\neg\varphi~|~\varphi\wedge\varphi~|~\varphi\vee\varphi~|~\varphi\Rightarrow\varphi'~|~\forall P.~\varphi~|~\exists P.~\varphi
-\end{displaymath}
-where $\wedge$, $\neg$, $\vee$, and $\Rightarrow$ are connectives representing
-conjunction, negation, disjunction, and (material) implication respectively; while
-$\forall$ and $\exists$ represent the standard universal and existential
-quantifiers (resp.). Finally, the symbol ``$.$'' is just syntactic sugar. 
-We consider $\interpretation\subset\Phi\times\{\top,\bot\}$ as the interpretation
-function, where $\Phi$ is any collection of sentences in $\follang$ 
-and $\top$ and $\bot$ represent the concepts of ``Tautology/True''
-and ``Contradiction/False'' respectively.
-
-Following Figure~\ref{fig:safety-security}, we start by formalizing the outermost term: Causality.
-
-\subsubsection{Causality Principle}\label{sec:causality}
-We formalize the Causality Principle starting from a K modal
-logic\autocite{Garson2018modal} (i.e. without restrictions on the causality
-relation between worlds). The standard definition of K modal logic is given in
-Definition~\ref{def:modallogic} in terms of an interpretation function (which
-we named $\sigma$ with a slight abuse of notation) defined in
-Definition~\ref{def:modalinterpretation}. 
-\begin{definition}{\bf K Modal Logic --}\label{def:modallogic}
-A K-frame is a frame $ \kframe=\langle \possibleworlds,\modalrelation \rangle $
-	in the K modal logic where $ \modalrelation $ is the binary relation
-	(i.e. a set of ordered pairs) between possible worlds
-	$\modalrelation\subseteq\possibleworlds\times\possibleworlds$, where
-	$\possibleworlds$ represents the possible worlds, and
-	$\possibleworlds\neq\emptyset$. An actual world
-	$\actualworld\in\possibleworlds$ is assumed. For any proposition $P$,
-	an interpretation function $\interpretation(\world,P)$ returns the
-	truth value of $P$; e.g. $\interpretation(\world,P)=\top$ means that
-	$P$ holds in $\world$. A model is defined as the tuple
-	$\kmodel=\langle\possibleworlds,\modalrelation,\interpretation\rangle$.
-\end{definition}
-
-\begin{definition}{\bf Causality as K Modal Interpretation Function --}\label{def:modalinterpretation}
-	Causality is (recursively) defined as the modal interpretation function $\interpretation$, as follows. 
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation0)$] if $\interpretation(\world,P)=\top$ then $\world\models P$
-		\item[$(\interpretation1)$] $\world\models\neg P$ iff $\world\not\models P$
-		\item[$(\interpretation2)$] $\world\models P \wedge Q$ iff $\world\models P$ and $\world\models Q$
-		\item[$(\interpretation3)$] $\world\models\Box P$ iff for any world $\world'\in\possibleworlds$ if $\world \modalrelation\world'$ then $\world'\models P$
-		\item[$(\interpretation4)$] $\world\models\Diamond P$ iff there exists a set of worlds $\World'\subset\possibleworlds$ such that for any $\world'\in\World'$, if $\world\modalrelation\world'$ then $\world'\models P$
-		\item[$(\interpretation5)$] $\models P$ iff $\actualworld\models P$
-	\end{enumerate}
-	where truth is defined as necessary with $\Box$ and possible with $\Diamond$.
-\end{definition}
-
-%\begin{definition}{\bf Causality --}\label{def:causality}
-%	The Causality principle is defined over a K Modal Logic (i.e. without
-%	restrictions on the causality relation between worlds) as in
-%	Definition~\ref{def:modallogic}, with the standard interpretation
-%	function defined in~\ref{def:modalinterpretation}
-%\end{definition}
-
-The causality principle has been defined in its
-generic form. In fact, the accessibility relation $\modalrelation$ is free from
-any axiomatic restriction (e.g. it's not reflexive nor anti-reflexive).  
-We will focus in Section~\ref{sec:process} (for our tests) on 
-the application of our theory to CPS system engineering.
-More detailed case studies (i.e. \fix{mr}{a CPS as a smart power-grid}) 
-will be defined in Section~\ref{sec:prototest}, where
-the accessibility relation defines in details how the system itself evolves, due to causal relation 
-(i.e. by restricting the causality principle only to those cause-effect that defines
-the system). Therefore, the definition of $\modalrelation$
-will be specialized in a more strict way based on the application domains and case study.
-
-We note that we have defined the
-Causality principle without considering it a \emph{threat source} since 
-we lack the concept of intent and maliciousness. Similarly, in the 
-next sections we won't discriminate between Nature and Mankind until we introduce 
-the concepts of maliciousness in Section~\ref{sec:theory} 
-and then formally define a threat agent
-in the same section.
-
-\subsubsection{Agents: Mankind and Nature}\label{sec:mankind-nature}
-\begin{table}[t]
-\centering
-\setlength{\tabcolsep}{3.5pt}
-\renewcommand{\arraystretch}{1}
-\scriptsize
-\caption{RCC3, RCC5, and RCC8 relations between Regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
-\begin{tabular}{ccclll} 
-\rota{\textbf{RCC3}}&\rota{\textbf{RCC5}}&\rota{\textbf{RCC8}}&\textbf{Name} & \textbf{Notation} & \textbf{Definition} \\
-\hline
-&&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& $\mathit{X}\subseteq \mathit{Y}$ \\
-&&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& $\mathit{X}\not\subseteq \mathit{Y}$\\
-&&&Part of				& $\mathit{P}(\mathit{X},\mathit{Y})$		& $\forall \mathit{Z} ~\mathit{C}(\mathit{Z},\mathit{X}) \rightarrow \mathit{C}(\mathit{Z},\mathit{Y})$\\
-&&&Overlaps			& $\mathit{O}(\mathit{X},\mathit{Y})$		& $\exists \mathit{Z} ~\mathit{P}(\mathit{Z},\mathit{X})\wedge \mathit{P}(\mathit{Z},\mathit{Y})$\\
-\Tdot&&&  \textbf{Overlaps Not Equal} 	& $\mathit{ONE}(\mathit{X},\mathit{Y})$		& $\mathit{O}(\mathit{X},\mathit{Y}) \land \neg \mathit{EQ}(\mathit{X},\mathit{Y})$ \\
-\Tdot&\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
-\Tdot&\Tdot&\Tdot& \textbf{DiscRete from} 		& $\mathit{DR}(\mathit{X},\mathit{Y})$		& $\neg \mathit{O}(\mathit{X},\mathit{Y})$\\
-&\Tdot&\Tdot&\textbf{Partial-Overlap}	& $\mathit{PO}(\mathit{X},\mathit{Y})$ 		& $\mathit{O}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{X},\mathit{Y}) \wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\
-&\Tdot&&\textbf{Proper-Part-of} 	& $\mathit{PP}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\ 
-	&\Tdot&&\textbf{Proper-Part-of-\textit{\textbf{i}}nverse} & $\mathit{PPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{Y},\mathit{X}) \wedge \neg \mathit{P}(\mathit{X},\mathit{Y})$\\
-&&\Tdot&\textbf{Externally Connected} 	& $\mathit{EC}(\mathit{X},\mathit{Y})$ 		& $\mathit{C}(\mathit{X},\mathit{Y}) \wedge \neg\mathit{O}(\mathit{X},\mathit{Y})$\\ 
-&&\Tdot&\textbf{Tangential PP} 	& $\mathit{TPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Tangential PPi} 	& $\mathit{TPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{TPP}(\mathit{Y},\mathit{X})$\\ 
-&&\Tdot&\textbf{Non-Tangential PP} 	& $\mathit{NTPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\neg\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Non-Tangential PPi} 	& $\mathit{NTPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{NTPP}(\mathit{Y},\mathit{X})$\\ 
-\end{tabular}
-\end{table}
-
-Mankind and Nature are defined in Section~\ref{sec:problem} as two abstract
-agents, both as collections (i.e. an abstract type that does not imply a
-specific implementation) of their sub-agents (i.e. humans for Mankind and
-plants, animals, \&c.  for Nature). Similarly to\autocite{Santaca2016abf}, 
-we define Mankind and Nature, and any
-other agent in the reminder of this article, as a meronomy (an hierarchy of
-Part-Whole relations) based on a standard definition of mereology, i.e. based
-on the definition of Parthood relation between \emph{Parts}.  
-However, shall 
-consider different types of Part; so, we extend the mereology to a
-mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
-to increase the number of different types considered and to generalize the relations between Parts (as
-in Table~\ref{tab:rcc358}).
-For the sake of readability, we use the term \emph{Region} both to refer to a
-mereological Part and to a topological Region.
-The choice of mereotopology is also correlated to the objective of defining a formal
-ontology, which we use to define the (formal) semantics of the terms (Parts) in
-Section~\ref{sec:problem}, and of the concepts of safety and security
-(whole). We aim at creating a meronomy instead of the taxonomies such as the one provided
-in\autocite{NIST2020NVD,MITRE2020CVE} or instead of the poorly justified 
-CVSS\autocite{Mell2007CVSS} scoring system.
-
-A mereotopology, as defined e.g. in\autocite{Rachavelpula2017mereotopology},
-is an ordered mathematical structure where the basic relation between Regions is the
-reflexive and symmetric\fixnote{mr}{I guess it must be monotonic as defined in\autocite{Rachavelpula2017mereotopology} but I don't find it consistently in other papers.} 
-\emph{Parthood} relation $\subseteq$. 
-\begin{definition}{\bf Parthood --}\label{def:parthood}
-Given any pair of mereotopological Regions $X$ and $Y$,
-	\begin{enumerate}[noitemsep]
-		\item Reflexivity: $\forall X.~ (X\subseteq X)$
-		\item Symmetry: $\forall X, Y.~ (X\subseteq Y \Rightarrow Y\subseteq X)$
-	\end{enumerate}
-\end{definition}
-
-As later defined in Definition~\ref{def:agent}, 
-the Parthood relation orders a universe of agents $\agentuniverse$ by defining
-the so called \emph{Connects with} (see in Table~\ref{tab:rcc358}) relation 
-between Regions. We want this universe $\agentuniverse$ to be 
-expressible in FOL. In this way, we
-can reason both on the constituent of security (i.e. its terms and agents
-defining a system where security needs to be considered), and on 
-the evolution of those constituent w.r.t. cause-effects relations according
-to the modal structure of causality we defined in Section~\ref{sec:causality}.
-This will allow us (in Section~\ref{sec:process} and Section~\ref{sec:prototest})
-a better\fixnote{mr}{Can we prove this by construction?} 
-positioning w.r.t. risk assessment technologies
-(which most often reason on the constituent of a system design), and
-protocol verification tools (which requires some formalization of a cause-effect relation, e.g.
-in Linear Temporal Logic).
-
-As we argued in Section~\ref{sec:problem}, we must correlate the definition of
-agent (i.e. Mankind and Nature) in Definition~\ref{def:mankind-nature} to the
-mathematical structure of the logic that defines them.  We express Mankind and
-Nature as formulas over the theory of mereology and then in terms of
-mereotopological Regions, extending the interpretation function
-$\interpretation$, to include a formal theory of mereotopology.  We use the
-Region Connection Calculus (RCC), as defined
-in~\cite{bennettLogics,improvingRCC}, to provide an axiomatization of the
-spatial concepts and relations in first-order logic to correlate the algebraic
-structure to mereology. In its broader definition, the RCC theory is composed
-by eight axioms, and is known as RCC8. In the text, for brevity, 
-we will often focus only on RCC5 (without
-loss of generality) by not considering tangential connections between spatial
-Regions. We discuss the choice of RCC5 in more detail in
-Appendix~\ref{app:rcc}. In Table~\ref{tab:rcc}, we
-summarize the axioms of the Region Connection Calculus (see, e.g., \autocite{Grutter2008rcc}).
-
-\begin{definition}{\bf RCC axiomatization --}
-	For any $X,Y$ pair of Regions in a mereotopology:
-	\begin{itemize}[noitemsep]
-	\item[$(\interpretation6)$] $\interpretation(X\subseteq Y)$ iff [$\interpretation(X\subseteq X)=\top$, and $\interpretation(X\subseteq Y)=\bot$ or $\interpretation(Y\subseteq X)=\top$]\fixnote{mr}{would it be better/clearer or just correct to write $\sigma(X\subseteq Y)$ iff $\sigma(X\subseteq X\wedge [X\subseteq Y \vee Y\subseteq X])=\top$}
-	\item[$(\interpretation7)$] $\world\models\connects{X}{Y}$ iff $\world\models X\subseteq Y$ 
-	\item[$(\interpretation8)$] $\world\models\disconnected{X}{Y}$ iff  $\world\not\models X\subseteq Y$
-	\item[$(\interpretation9)$] $\world\models\partof{X}{Y}$ iff $\world\models\forall Z.~\connects{Z}{X}\Rightarrow\connects{Z}{Y}$
-	\item[$(\interpretation10)$] $\world\models\overlaps{X}{Y}$ iff $\world\models\exists Z.~\partof{Z}{X}\wedge\partof{Z}{Y}$
-	\item[$(\interpretation11)$] $\world\models\eq{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\partof{Y}{X}$
-	\item[$(\interpretation12)$] $\world\models\dr{X}{Y}$ iff $\world\models\neg\overlaps{X}{Y}$
-	\item[$(\interpretation13)$] $\world\models\po{X}{Y}$ iff $\world\models\overlaps{X}{Y}\wedge\neg\partof{X}{Y}\wedge\neg\partof{Y}{X}$
-	\item[$(\interpretation14)$] $\world\models\pp{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\neg\partof{Y}{X}$
-	\item[$(\interpretation15)$] $\world\models\ppi{X}{Y}$ iff $\world\models\partof{Y}{X}\wedge\neg\partof{X}{Y}$
-\end{itemize}
-where $Z$ is a mereotopological Region.
-\end{definition}
-
-\begin{definition}{\bf Agent: Mankind or Nature --}\label{def:agent}
-	An agent $\agent\in\agentuniverse$ is a tuple
-	$\langle\rcc(\region',\region''),\ldots,\rcc(\region^{n-1},\region^n)\rangle$
-	of RCC relations $\rcc$ over mereotopological Regions
-	${\region',\ldots,\region^n}\subseteq\Region$. 
-\end{definition}
-
-Currently, we do not distinguish between Mankind and Nature (since we still
-lack of the definition of ``malicious intent'', which is defined in
-Section~\ref{sec:theory}) and we have defined them as two generic
-agents.
-
-As depicted in Figure~\ref{fig:safety-security}, Causality, Mankind, and Nature
-have a dashed border representing their correlation in terms of cause-effect
-and then in terms of formal structure which defines them: Modal Logic.
-Vulnerability, Defect, Incident, Accident, and Threat, similarly, 
-are correlated (depicted as a solid border) in terms of underlying 
-formal structure: Mereotopology.
-
-\subsubsection{Regions: Vulnerability and Defect (and Weakness)}\label{sec:vulnerabilitydefect}
-As defined in Section~\ref{sec:problem}, a Vulnerability or a Defect is a
-\emph{Weakness}.  As an example, a categorization of weaknesses is given
-in\autocite{MITRE2020CWEresearch} with 808 weaknesses categorized as ``Research
-Concepts'', distributed as follows:
-\begin{itemize}[noitemsep]
-	\item Incorrect Calculation - ($682$)
-	\item Incorrect Access of Indexable Resource (``Range Error'') - ($118$)
-	\item Use of Insufficiently Random Values - ($330$)
-	\item Improper Interaction Between Multiple Correctly-Behaving Entities - ($435$)
-	\item Improper Control of a Resource Through its Lifetime - ($664$)
-	\item Insufficient Control Flow Management - ($691$)
-	\item Protection Mechanism Failure - ($693$)
-	\item Incorrect Comparison - ($697$)
-	\item Improper Check or Handling of Exceptional Conditions - ($703$)
-	\item Improper Enforcement of Message or Data Structure - ($707$)
-	\item Improper Adherence to Coding Standards - ($710$)
-\end{itemize}
-
-The definition given by the MITRE in\autocite{MITRE2020CWEweakness} of
-weakness is: ``Software weaknesses are errors that can lead to software
-vulnerabilities. A software vulnerability, such as those enumerated on the
-Common Vulnerabilities and Exposures (CVE) List, is a mistake in software that
-can be directly used by a hacker to gain access to a system or network''.
-The definition is circular if we interpret the word ``error'' and
-``mistake'' with the same semantics: a weakness is an error that leads to a
-vulnerability and a vulnerability is a mistake which, in turn, is a weakness.
-The only difference (between weakness and vulnerability) seems to be
-that one can consider weakness as a ground term and state that a
-vulnerability is caused by a weakness, i.e. 
-$\World,\Weakness\models\Diamond\Vulnerability\wedge\Weakness$ where
-$\Weakness,\Vulnerability$ are Regions of Weaknesses and Vulnerabilities (resp.);
-accepting the hierarchy in the CWE\autocite{CWE} as ground truth. 
-Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability) 
-or the CVE reported in the NVD as a ground truth.
-
-\begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
-	A Region $\region\subseteq\agent$ of an agent
-	$\agent\in\agentuniverse$, is defined as Weakness $\Weakness$ (i.e.
-	representing a weakness introduced by the agent into any phase of
-	production, e.g.  of the secure process development life-cycle, of any
-	system or subsystem) iff there exists $\world\in\possibleworlds$ such
-	that $\world,\Weakness\models\Diamond\exists\Weakness',\Vulnerability'.~
-	\rcc(\Weakness',\Vulnerability')\wedge\neg\dr{\Weakness'}{\Vulnerability'}$, 
-	where $\Weakness'\subseteq\Weakness$, and
-	$\Vulnerability'\subseteq\Vulnerability$ is a Region of an agent $\agent'\in\agentuniverse$.
-\end{definition}
-
-\begin{example}{CWE-116: Improper Encoding or Escaping of Output\autocite{CWE-116} --}
-	\begin{itemize} 
-		\item Description: The software prepares a structured message
-		for communication with another component, but encoding or
-		escaping of the data is either missing or done incorrectly. As
-		a result, the intended structure of the message is not
-		preserved. 
-		\item Example: This code displays an email
-		address that was submitted as part of a form. Example language JSP.
-		\begin{verbatim} 
-		<% String email = request.getParameter("email"); %> 
-		...
-		Email Address: <%= email %>
-		\end{verbatim}
-		The value read from the form parameter is
-		reflected back to the client browser without
-		having been encoded prior to output, allowing
-		various XSS attacks (CWE-79).
-		\item Observed Examples
-			\begin{itemize}
-			\item CVE-2008-4636\autocite{CVE-2008-4636}: OS command injection in backup
-				software using shell metacharacters in a
-				filename; correct behavior would
-				require that this filename could not be
-				changed.
-			\item CVE-2008-0769\autocite{CVE-2008-0769}: Web application does not set the
-				charset when sending a page to a browser,
-				allowing for XSS exploitation when a
-				browser chooses an unexpected encoding.
-			\item CVE-2008-0005\autocite{CVE-2008-0005}: Program does not set the charset
-				when sending a page to a browser, allowing for
-				XSS exploitation when a browser chooses
-				an unexpected encoding.
-			\item CVE-2008-5573\autocite{CVE-2008-5573}: SQL injection via password
-				parameter; a strong password might contain \&
-			\item CVE-2008-3773\autocite{CVE-2008-3773}: Cross-site scripting in chat
-				application via a message subject, which
-				normally might contain \& and other
-				XSS-related characters.
-			\item CVE-2008-0757\autocite{CVE-2008-0757}: Cross-site scripting in chat
-				application via a message, which normally might
-				be allowed to contain arbitrary
-				content.
-			\end{itemize}
-\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\begin{subfigure}[b]{.5\textwidth}
-		\centering
-		\includegraphics[width=.7\textwidth]{rcc5_CWE-116.pdf}
-		\caption{Representation with RCC3}
-		\label{fig:rcc5_CWE-116}
-	\end{subfigure}\begin{subfigure}[b]{.5\textwidth}
-		\centering
-		\includegraphics[width=.7\textwidth]{rcc8_CWE-116.pdf}
-		\caption{Representation with RCC8}
-		\label{fig:rcc8_CWE-116}
-	\end{subfigure}
-	\begin{subfigure}[b]{\textwidth}
-		\centering
-		\includegraphics[width=\textwidth]{KML_CWE-116.pdf}
-		\caption{Representation in K-Modal Logic}
-		\label{fig:KML_CWE-116}
-	\end{subfigure}
-	\caption{Relations between CWE-116 and correlated CVEs}
-\end{figure}
-Definition~\ref{def:weakness} states that CWE-116 is a weakness iff there exists a world in
-K Modal Logic, representing the system in which this weakness exists, 
-such that the natural evolution of this system (i.e. formalized by the causality principle) 
-make it possible to reach another state of the system (i.e. another, accessible world)
-where there exist a vulnerability that can be implemented from CWE-116 (and this relation 
-is not DR). The CWE website proposes the connection between CWE-116 and, for example, 
-CVE-2008-5573; a vulnerability of the login sub-system 
-of the Poll Pro v2.0\autocite{pollpro} system.
-The formal relation between the two is given as a link (i.e. URI) between the CWE and
-the CVE, the description of the relation is not defined but it is supposed to be
-inferred from the descriptions of the CWE-CVE. 
-We can formally represent this link as the ONE connection in RCC3, depicted in 
-Figure~\ref{fig:rcc5_CWE-116}, or EC connection in RCC8 (Figure~\ref{fig:rcc8_CWE-116}).
-
-\fixnote{mr}{shouldn't we consider the agent who introduces the weakness as dishonest? don't we say this in part 1?}
-\end{example}
-
-It is interesting to note that the CWE-CVE relation expresses the correlation
-between weaknesses and vulnerabilities in the most simple form, such that we
-can formalize all the relation as ONE in RCC3 or EC in RCC8.  To express a more
-complex relation between the two we shall analyze the definition of CWE-116.
-This is related to the Weakness-Vulnerability-Incident process (i.e. to the
-details on the implementation/realization in Figure~\ref{fig:safety-security})
-that we analyze in the next section, Section~\ref{sec:incidentaccident}.
-
-\begin{definition}{\bf Region: Vulnerability and Defect --}\label{def:vulnerability-defect}
-A region $\region\subseteq\agent$ of an agent $\agent\in\agentuniverse$ is
-	called \emph{Vulnerability} if the agent $\agent$ is referred to as Mankind, 
-	\emph{Defect} if the agent is referred to as Nature.
-\end{definition}
-
-\subsubsection{Process: Incidents and Accidents}\label{sec:incidentaccident}
-In our informal definition depicted in Figure~\ref{fig:safety-security}, an
-Incident is generated through a process caused by a Vulnerability (which,
-in turn, is caused by a Weakness). Symmetrically, Nature has a process from Defect
-to Accidents. We start by analyzing the process generated by Mankind, which is
-divided into the following phases (explained in the remainder of this sub-Section):
-%Agent$\rightarrow$Vulnerability$\rightarrow$Exploit$\rightarrow$Attack$\rightarrow$Incident.
-\begin{enumerate}[noitemsep]
-	\item An agent (Mankind or Nature) defines a system, e.g.
-by following an engineering process for creating a new system, or for redesigning or improving an existing system (e.g.
-a component or a collection of components)
-	\item The dogmatic definition of a system contains a Weakness (i.e.
-the agent, Mankind or Nature, introduces an unintended behavior), for example as an un-intended behavior or as an architectural fallacy.
-The Weakness can be applied to the abstract definition of the system.
-	\item The Weakness can be defined as a malicious process for a system, called Vulnerability (or Defect for Nature)
-	\item The process expressed by the Vulnerability can be made operational for a specific implementation of the system (i.e. implemented by Mankind or realized by Nature) 
-\end{enumerate}
-Those additional details are depicted in Figure~\ref{fig:safety-security_2}, an update
-of Figure~\ref{fig:safety-security}.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{safety-security_2.pdf}
-	\caption{Overview security and safety keywords}
-	\label{fig:safety-security_2}
-\end{figure}
-
-\Paragraph{Exploitation Flow} An \emph{Exploit} is an \emph{implementation} of
-a Vulnerability, where a Vulnerability is one of the many ways to transform a
-Weakness into a malicious process that can be applied\footnote{Meaning that the
-malicious process, i.e.  any structured procedure such as a protocol, can be
-somehow made operational, e.g. implemented} to one or many systems. So, the act
-of an agent (Mankind) of ``exploiting a Vulnerability'' is the process of
-making the Vulnerability operational, by implementing the Vulnerability w.r.t.
-a specific target system.  Whenever the agent is Nature, the implementation is
-more broadly considered a realization\footnote{``The state of being
-realized''\autocite{Merriam2020realization}}
-
-The Common Attack Pattern Enumeration and
-Classification\autocite{MITRE2020CAPEC} (CAPEC) online service provides a
-``community resource'' which aims at ``identifying and understanding attacks''.
-The MITRE writes on the CAPEC homepage that ``Understanding how the adversary
-operates is essential to effective cyber security. CAPEC helps by providing a
-comprehensive dictionary of known patterns of attack employed by adversaries to
-exploit known weaknesses in cyber-enabled capabilities. It can be used by
-analysts, developers, testers, and educators to advance community understanding
-and enhance defenses.''. We can extrapolate two major concepts: 
-\begin{enumerate}
-	\item while understanding the exploitation process is the aim of the
-		CWE and CVE initiatives, understanding the hacking process is
-		the aim of the CAPEC initiative.
-	\item CAPEC provides a \emph{dictionary} of \emph{known} attack patterns.
-		As stated largely in Section~\ref{sec:}, this dictionary contains
-		empirical evidences and should not be used to induce a cybersecurity 
-		theory; but it can be used to validate a theory.
-\end{enumerate}
-
-The process continues as follows.
-\begin{enumerate}[noitemsep]
-\setcounter{enumi}{4}
-	\item an agent (Mankind\footnote{For the sake of readability we focus only on Mankind as an agent}) the exploitation flow can be used to hack a system.
-We use the term hack (as hacking) to stress that nothing prevents this process from being honest, however, for the sake of simplicity, we focus on Attacks which are malicious by definition.
-	\item An Attack is applied to a system with malicious intent by an agent
-	\item The application of the attack results in an Incident which impacts, i.e. poses a Threat to, an Asset
-\end{enumerate}
-
-\Paragraph{Hacking Flow}
-
-\begin{example}
-given in natural language in the description of the 
-CWE-116 which we can analyze as follows:
-\begin{itemize}
-	\item ``the software prepares a structured message''
-	\item the message is intended ``for communication with another component''
-	\item ``but''
-	\item ``encoding or escaping of the data is either missing or done incorrectly''
-	\item ``[\ldots] result, the intended structure of the message is not preserved''
-\end{itemize}
-\end{example}
-
-In this Vulnerability is seen as a process through which an agent (i.e. Mankind)
-generates an Incident, i.e.
-
-We describe processes, in our current formalization, in terms of the K-Modal
-Structure as in Definition~\ref{def:modalinterpretation} (i.e. in terms of
-causality principle).
-
-\begin{definition}{\bf Vulnerability-Incident Process --}\label{def:vulnerability-incident}
-	Given two 
-\end{definition}
-
-The whole-part relation \fix{mr}{(i.e.  HAS-A\footnote{HAS-A is just an hint for the
-mechanization of this theory.  We'll come back to this in
-Section~\ref{sec:implementation}})} is given, in the following, over the K Modal
-logic described beforehand. To do so, we first need to formally define the
-concept of \emph{agent} in its abstract form. We build it following the same
-structure of\autocite{Santaca2016abf} but over the terms used in
-Section~\ref{sec:problem}, i.e. knowledge, and belief as epistemic concepts. We'll then 
-introduce the concept of \emph{intent} to define a \emph{threat agent}.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.8\textwidth]{knowledge-belief.pdf}
-	\caption{Informal representation of Knowledge and Belief}
-	\label{fig:knowledge-belief}
-\end{figure}
-
-\subsubsection{Region: Threat and Asset}\label{sec:threatasset}
-The difference between Knowledge and Belief is depicted in
-Figure~\ref{fig:knowledge-belief} (see \autocite{wiki-knowledgebelief}).  However,
-according to\autocite{Gettier2012knowledge},
-Knowledge\footnote{``\emph{Theaetetus}: [\ldots] He said that knowledge was
-true opinion accompanied by reason, but that unreasoning true opinion was
-outside of the sphere of knowledge; and matters of which there is not a
-rational explanation are unknowable -- yes, that is what he called them -- and
-those of which there is are knowable. [\ldots] \emph{Socrates}: [\ldots] the
-primary elements of which we and all else are composed admit of no rational
-explanation; for each alone by itself can only be named, and no qualification
-can be added, neither that it is nor that it is not, for that would at once be
-adding to it existence or non-existence, whereas we must add nothing to it, if
-we are to speak of that itself alone.  [\ldots]'' Plato -- Theaetetus 201
-\autocite{Plato1914Plato}} as an epistemological concept is difficult (and
-sometimes believed impossible\autocite{citation}) to formally define. In this
-work, we are not interested in 
-how knowledge can be precisely formalized from an epistemic standpoint. 
-We assume that a semantic of a correct definition of epistemic knowledge exists,
-for example the one given in\autocite{Hintikka1962knwoledge} by Hintikka, and we then define 
-knowledge in terms of the Kripke structure defined in Definition~\ref{def:causality}.
-
-\begin{definition}{\bf Knowledge --}\label{def:knowledge}
-Given an abstract collection of Agents $Ag$, and the modal operator
-	$\knows{a}{}$ (where $a\in Ag$), Knowledge is defined as a collection
-	of predicates known by an agent $\knowledge{a}=\bigcup_\Phi \knows{a}{\varphi}$ 
-	(where $\Phi$ is the collection of all the propositions known by $a$).
-	Given a proposition $P$, we extend the semantics of the Causality structure with:
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation6)$] $\world\models\knows{a}{P}$ iff
-			$\world'\models P$ for all $\world'$ such that
-			$\world\modalrelation\world'$.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Belief --}\label{def:belief}
-	\fix{mr}{we should help the reader in understanding the upgrades of the logic.
-We used propositional with modal operators with a K Kripke semantics and we are now introducing Belief as in Doxastic logic?
-and we have epistemological operators as knowledge}
-	Given an abstract collection of Agents $Ag$, and the modal operator
-	$\believe{a}{}$ (where $a\in Ag$), Belief is defined as a collection
-	of predicates believed by an agent $\belief{a}=\bigcup_\Phi \believe{a}{\varphi}$
-	(where $\Phi$ is the collection of all the propositions believed by $a$).
-	Given a proposition $P$, we extend the semantics of the Causality structure with:
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation7)$] $\world\models\believe{a}{P}$ iff
-			$\world\models\neg\knows{a}{\neg P}$ (i.e. the agent $a$ considers $P$ possible) 
-			and $\world'\models P$ for all
-			$\world'$ such that $\world\modalrelation\world'$.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Mankind/Nature --}\label{def:mankind-nature}
-Mankind/Nature is represented as an agent.  
-	%countable logical signature $\Sigma_M$/$\Sigma_N$ over an
-	%abstract algebraic structure.  We also consider a countable signature
-	%of variables, disjoint from $\Sigma_M$/$\Sigma_N$, and write
-	%$\Sigma_{M_n}$/$\Sigma_{N_n}$ for the symbols of $\Sigma_M$/$\Sigma_N$
-	%with arity $n$. Thus, $\Sigma_0$ is a collection of constants, which we
-	%assume to have disjoint sub-collections that we refer to as
-	%\emph{sub-agents} of Mankind or Nature (i.e. humans, plants, \&c).
-	%The variables are untyped and can be instantiated with arbitrary types,
-	%yielding an untyped model.
-\end{definition}
-
-\subsubsection{Mereotopological Representation of Safety and Security}\label{sec:formalGlossary}
-The objectives of this section are:
-\begin{enumerate}[noitemsep]
-	\item formalize the abstract algebraic structure (as a mereotopology)
-		on top of which we have formalized the terms in
-		Figure~\ref{fig:safety-security} so far, i.e. Mankind, Nature,
-		Truths, Beliefs, and Knowledge.
-	\item correlate mereological structure to Causality and then Kripke
-		structure by defining the relation $R$ over the Kripke
-		structure as the mereotopological basic relation connects with,
-		i.e. reflexive and symmetric
-	\item we map signature to mereotopological Regions to express the
-		logical formulation of terms (e.g. Mankind, Nature, Truths,
-		Beliefs, Knowledge, Vulnerabilities) over the mereotopological
-		structure
-	\item we formalize the relations between terms using the RCC(5)
-\end{enumerate}
-
-\fix{mr}{change propositions in propositional language\url{http://logic.stanford.edu/intrologic/lectures/lecture_02.pdf}}
-
-Given that the concepts of Truths, Beliefs, and Knowledge are positioned over
-an underlying structure of Propositions, the first concept to formalize is that
-of Propositions. We do not restrict Propositions to a specific structure but,
-as for Mankind and Nature, we refer to them as a collection.  We express the
-generality of a collection as a (propositional) signature over an abstract
-algebraic structure. In our formalization, we use mereotopology as algebraic
-structure (due to its generality, as discussed in Section~\ref{sec:problem})
-and the RCC to formalize concepts over a mereotopology.
-
-We can now define the possible mereotopological relations between Belief and
-Knowledge (informally depicted in Figure~\ref{fig:knowledge-belief}) and then
-define the peculiarity of the believed difference between Mankind and Nature 
-(i.e. maliciousness as intent). 
-We first start from the formalization of Figure~\ref{fig:knowledge-belief}, as
-follows.
-
-\begin{definition}{\bf Information}\label{def:information}
-\end{definition}
-\begin{definition}{\bf Desire}\label{def:desire}
-\end{definition}
-\begin{definition}{\bf Intent}\label{def:intent}
-\end{definition}
-
-\subsection{Vulnerability and Defect}\label{sec:vulnerability-defect}
-In order to formalize the semantics of the concepts (informally) described in
-Section~\ref{sec:problem},
-
-\begin{definition}{\bf Vulnerability}
-\end{definition}
-
-\subsection{Cybersecurity Engineering Process}\label{sec:process}
-%\epigraph{La differenza tra idee e realt\`a \`e la differenza tra
-%filosofia e ingegneria. Il lavoro per trasformare l'una nell'altra \`e ricerca
-%scientifica
-\epigraph{The difference between ideas and reality is the
-difference between philosophy and engineering. The work to transform one into
-the other is scientific research}{{\itshape V-Research}}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{vmodel.pdf}
-	\caption{Cybersecurity Engineering Life-cycle}
-	\label{fig:knowledge-belief}
-\end{figure}
-
-\subsubsection{Relevant Standards}\label{sec:standards}
-\begin{enumerate}[noitemsep]
-	\item DO-326A
-\end{enumerate}
-
-\section{Honesty and Maliciousness -- A Cybersecurity Theory}\label{sec:theory}
-\begin{enumerate}[noitemsep]
-	\item cybersecurity as a requirement
-	\item vulnerability\label{def:vulnerability}
-\end{enumerate}
-
-\section{Prototype and Empirical Tests}\label{sec:prototest}
-\subsection{Prototype Implementation}\label{sec:implementation}
-\subsection{Empirical Tests}\label{sec:tests}
-
-\section{Conclusion and Related Work}\label{sec:related}
-
-\printbibliography
-
-\begin{appendices}
-	\section{RCC3, RCC5, or RCC8}\label{app:rcc}
-	\section{Wrong idea 1}
-In order to apply consistently the mereotopological structure not just to
-reason the agents but also on their evolution, we extend the properties of the
-Parthood relation (reflexivity and symmetry) to the modal relation
-$\modalrelation$ of the K Modal logic, which in turns defines the causality
-principle. This updates the K Modal logic to a B (Brouwer) Modal Logic\autocite{Garson2018modal}.
-
-\begin{definition}{\bf B Modal Logic --}\label{thm:bk-parthood}
-	The B Modal Logic (defined, with a slight abuse of notation as
-	$\bframe=\langle\possibleworlds,\modalrelation\rangle$) is an extension
-	of the K modal logic with the following axioms
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation8)$] $\world\models P\Rightarrow Q$ iff
-			$\world\not\models P$ or $\world\models Q$ (logical implication)~\footnote{In
-			Definition~\ref{def:modallogic} we didn't specify the
-			semantics in the case of the connective $\Rightarrow$
-			(i.e. logical implication), so we state it here. Given
-			that $P\Rightarrow Q$ is equisatisfiable to
-			$\neg(P\wedge \neg Q)$, for any proposition $P$ and
-			$Q$, $\world\models P\Rightarrow Q$ is equisatisfiable
-			to $\world\models \neg(P\wedge \neg Q)$. If
-			$\world\models \neg(P\wedge \neg Q)$ then
-			$\world\not\models P\wedge\neg Q$ and then
-			$\world\models \neg P$ or $\world\models Q$.}
-		\item[$(\interpretation9)$] $\Box P\Rightarrow P$ (which ``claims that whatever is necessary is the case''\autocite{Garson2018modal}), and
-		\item[$(\interpretation10)$] $P\Rightarrow\Box\Diamond P$
-			(which says that if $P$ is the case, then $P$ is
-			necessarily possible\autocite{Garson2018modal})
-	\end{enumerate}
-\end{definition}
-
-\begin{theorem}{\bf Causality, Parthood, and B Modal Logic}\label{thm:bk-parthood}
-	\begin{enumerate}[noitemsep]
-		\item (Reflexivity) $\bframe\models\Box P\Rightarrow P$ ($\interpretation8$) iff $\modalrelation$ is reflexive
-		\item (Symmetry) $\bframe\models P\Rightarrow\Box\Diamond$ ($\interpretation9$) iff $\modalrelation$ is symmetric
-		\item (Consistency) $\world\modalrelation\world'$ iff $\connects{\region}{\region'}$
-	\end{enumerate}
-	whenever $\world$ and $\world'$, and $\region$ and $\region'$ are logically equivalent ($\equiv$).
-\end{theorem}
-\begin{proof}
-	The proof is exhaustive over all possible cases.
-	\begin{enumerate}[noitemsep]
-		\item(Reflexivity) proven in\autocite{}
-		\item(Symmetry) proven in\autocite{}
-		\item(Equivalence) \fix{mr}{todo} If
-			$\world\modalrelation\world'$ and
-			$\world\equiv\region$, then if
-			it doesn't hold that $\connects{\region,\region'}$ it
-			follows that there is no reflexive and transitive
-			relation that correlates $\region$ to $\region'$. Given 
-			that Regions are defined as signatures
-	\end{enumerate}
-\end{proof}
-
-\end{appendices}
-
-\end{document}

### 2020-02-20 formal_glossary.tex
restructure

--- 2020-02-20 v-research_experiment-alpha.tex
@@
-\documentclass{article}
-\usepackage{mathtools}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{enumitem}
-\usepackage{multicol}
-\usepackage[toc,page]{appendix}
-\usepackage[gen]{eurosym}
-\usepackage{tabularx}
-\usepackage{epigraph}
-\usepackage{textcomp}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{csquotes}% Recommended
-\usepackage{rotating}
-\usepackage[style=numeric]{biblatex}
-% \bibliography{<mybibfile>}% ONLY selects .bib file; syntax for version <= 1.1b
-\addbibresource{literature.bib}% Syntax for version >= 1.2
-
-\usepackage{graphicx}
-\graphicspath{{./images/}}
-
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\lhead{}
-\chead{}
-\rhead{}
-\lfoot{}
-\cfoot{\thepage}
-\rfoot{Property of V-Research S.r.l.}
-\renewcommand\headrulewidth{0pt}
-\renewcommand\footrulewidth{0pt}
-
-\usepackage[colorinlistoftodos]{todonotes}
-
-\newcommand{\Paragraph}[1]{\smallskip\noindent{\bf #1.}}
-\newcommand{\fixnote}[2]{\textbf{\color{red}{FIX}}\footnote{{\bf #1:} #2}}
-\newcommand{\fix}[2]{{\color{red} {\bf tofix:} #2}}
-% \renewcommand{\fixnote}[2]{}
-% \renewcommand{\fix}[2]{}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\theoremstyle{corollary}
-\newtheorem{corollary}{Corollary}[section]
-\theoremstyle{lemma}
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{theorem}
-\newtheorem{theorem}{Theorem}[section]
-\theoremstyle{theorem}
-\newtheorem{example}{Example}
-
-\usepackage{authblk}
-\date{}                     %% if you don't need date to appear
-\setcounter{Maxaffil}{0}
-\renewcommand\Affilfont{\itshape\small}
-
-% for fancy table with rotated column titles
-\usepackage{adjustbox}
-\newcolumntype{R}[2]{%
-    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
-    l%
-    <{\egroup}%
-}
-\newcommand*\rot{\multicolumn{1}{R{45}{0.01em}}}% no optional argument here, please!
-\newcommand*\rota{\multicolumn{1}{R{90}{0.01em}}}% no optional argument here, please!
-\def\hb{\hbox to 10.7 cm{}}
-
-% circles
-\usepackage{wasysym}
-\newcommand{\Tdot}{$\CIRCLE$}
-\newcommand{\Thdot}{$\LEFTcircle$}
-\newcommand{\Twdot}{$\Circle$}
-
-% formulas
-\newcommand{\kframe}{\mathbf{K}}
-\newcommand{\bframe}{\mathbf{B_K}}
-\newcommand{\possibleworlds}{G}
-\newcommand{\modalrelation}{R}
-\newcommand{\actualworld}{\omega*}
-\newcommand{\world}{\omega}
-\newcommand{\World}{\Omega}
-\newcommand{\kmodel}{M}
-\newcommand{\interpretation}{\sigma}
-\newcommand{\knowledge}[1]{\mathbb{K}_{#1}}
-\newcommand{\knows}[2]{K_{#1}#2}
-\newcommand{\belief}[1]{\mathbb{B}_{#1}}
-\newcommand{\believe}[2]{B_{#1}#2}
-\newcommand{\region}{\chi}
-\newcommand{\Region}{\mathrm{X}}
-\newcommand{\agentuniverse}{\mathcal{A}}
-\newcommand{\agent}{a}
-\newcommand{\follang}{\mathcal{L}}
-\newcommand{\weakness}{w}
-\newcommand{\Weakness}{W}
-\newcommand{\vulnerability}{v}
-\newcommand{\Vulnerability}{V}
-%rcc5
-\newcommand{\rcc}{rcc}
-\newcommand{\connects}[2]{C(#1,#2)}
-\newcommand{\disconnected}[2]{\neg C(#1,#2)}
-\newcommand{\partof}[2]{P(#1,#2)}
-\newcommand{\overlaps}[2]{O(#1,#2)}
-\newcommand{\eq}[2]{EQ(#1,#2)}
-\newcommand{\pp}[2]{PP(#1,#2)}
-\newcommand{\po}[2]{PO(#1,#2)}
-\newcommand{\ppi}[2]{PPi(#1,#2)}
-\newcommand{\dr}[2]{DR(#1,#2)}
-
-\makeindex
-
-\begin{document}
-
-\title{On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}}
-\author[1]{Francesco Beltramini}
-\author[1]{Marco Rocchetto}
-\author[2]{Luca Vigan\`o}
-\affil[1]{V-Research, Verona, Italy}
-\affil[2]{King's College London, UK}
-
-\maketitle
-
-\begin{abstract}
-	The objective of this research is to develop of a theory that defines
-	(all and only) the possible insecurity and security configurations of
-	any abstract system. The theory is structured upon other theories that 
-	defines how a component of a system can be abstracted into an agent, defining how
-	agents can be formalized (both syntactically and semantically) to
-	describe an abstract system, such as a graph. Some of these theories
-	(e.g. used for the semantic
-	definition of the abstract system) are the epistemological definition
-	of knowledge, the Belief-Desire-Intent and the Assertion-Belief-Fact 
-	framework of reference, 
-	mereology, and topological structure. We argue that a mereology is the
-	most appropriate abstract underlying structure, do to its generality,
-	for defining the expressiveness of the system abstraction.  Furthermore, a
-	mereology allows us to define an ontology rather than a taxonomy.  We
-	also correlate different abstractions of the system to the TRL and the
-	engineering V-model.
-
-	We implemented a formal theory (of axioms) of a mereotopology, and of
-	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
-	uses the Z3 SMT solver. The results show that a single component (i.e.
-	agent) of an abstract system has a definite number of  different
-	insecurity configurations (e.g. 53 using RCC5 over a topological
-	structure) and only 1 secure (i.e. expected) configurations. The
-	configurations are reported as models satisfying the abstract system
-	semantics.
-
-	We considered the philosophical definition of truth behind our
-	approach, rejecting ``proof'' by induction from partial empirical evidences.
-	Our theory can be applied to system engineering and we show a concrete
-	application of our theory to the risk assessment of an ad-hoc system.
-	Finally, we provide a number of ideas to support the engineering of
-	secure systems (e.g. purely cyber or cyber-physical).
-\end{abstract}
-\newpage
-
-\section{Introduction}\label{sec:intro}
-\epigraph{Humanum est errare}{{\itshape Seneca the Elder}}
-The European Commission states in\autocite{EU2019market} that: ``Cybersecurity
-is one of the priority areas [\ldots] of the Commission initiative on ICT
-Standards, which is part of the Digitising European
-Industry\autocite{EU2019standard} strategy launched on 19 April 2016.  The aim
-is to identify the essential ICT standards and present measures to accelerate
-their development in support of digital innovations across the economy''. The
-same document (i.e.\autocite{EU2019market}) states that ``The EU will invest up
-to \euro450 million [\ldots], under its research and innovation programme
-Horizon 2020''. The EU, in 2016 published a press release\autocite{EU2016press}
-in which they present a strategy to invest \euro1.8 \emph{billion} to
-``increase measures to address cyber threats''. The EU is not the only investor
-in cybersecurity, most of the developed countries and several companies are
-investing enormous amount of money towards various aspects of cybersecurity
-(e.g. The US vulnerability databases\autocite{NIST2020NVD} maintained by the
-National Institute of Standards and Technologies, i.e. NIST, of the US
-Department of Commerce).
-
-The cybersecurity industry is growing fast, e.g. as reported
-in\autocite{Nasdaq2018market}. For example, in\autocite{Forbes2017market},
-published by the Forbes, is stated that \euro5.3 billion of funding where
-poured by venture capitalist into cybersecurity companies in 2018. The Forbes,
-in the same article, also highlights another peculiar (as seemly contradictory)
-trend: ``[\ldots] during the same time period, the number of cybersecurity
-breaches increased exponentially''. The data reported by the NIST 
-through the official CPE (Common Platform Enumeration) Dictionary Statistics 
-on the NVD websites in \autocite{NIST2020CPEstatistics}, show that in 2016 the
-number of reported vulnerabilities reported where around $~6000$ while in 2019
-the number of vulnerabilities was above $16000$.  The scientific community also
-reports similar findings.  In fact, in\autocite{Herley2009so}, Cormac Herley
-(Microsoft Research) shows how basic cybersecurity principles (such as the
-confidentiality benefit over the clear text for passwords typed into forms,
-e.g. for logins in websites) are not fully understood or shared between the
-cybersecurity research community\autocite{Nielsen2009stop}.  The lack of
-understanding of basic security principle, the inverse proportionality between
-investments in cybersecurity and the number of reported vulnerabilities year
-after year, can be linked to the lack of a foundational theory on
-cybersecurity, as already highlighted by Cormac Herley
-in\autocite{Herley2016unfalsifiability}.
-
-Another way of looking at the same problem is by analyzing the different
-definitions of security with respect to the scientific method of enquiry
-applied to get to the definition itself. In the following we categorize the
-related work into three categories, as defined by Sextus Empiricus in
-\autocite{Empiricus1990Pyrrhonism}.  To the best of our knowledge, there is no
-evidence that this categorization isn't complete.  ``The natural result of any
-investigation is that the investigators either discover the object of search or
-deny that it is discoverable and confess it to be in-apprehensible or persist
-in their search. [\ldots] This is probably
-why''\autocite{Empiricus1990Pyrrhonism}: 
-\begin{itemize}
-	\item The \emph{dogmatists} ``have claimed to have discovered the truth'' on what cybersecurity is
-		\begin{itemize}
-			\item Wikipedia defines cybersecurity in
-				\autocite{wiki-cybersecurity} as the protection
-				of computer systems and networks from the theft
-				of or damage to their hardware, software, or
-				electronic data, as well as from the disruption
-				or misdirection of the services they provide
-		\end{itemize}
-	\item The \emph{academics} ``have asserted that it cannot be apprehended''
-		\begin{itemize}
-			\item Eugene H. Spafford, Professor at Purdue
-				University, defines cybersecurity as follow.
-				``The only truly secure system is one that is
-				powered off, cast in a block of concrete and
-				sealed in a lead-lined room with armed guards —
-				and even then I have my doubts.''
-				\autocite{Spafford2019Quotes}
-		\end{itemize}
-	\item The \emph{skeptics} ``go on inquiring''
-		\begin{itemize}
-			\item Cormac Herley reaches the conclusion that
-				cybersecurity has no definition. ``There is an
-				inherent asymmetry in computer security: things
-				can be declared insecure by observation, but
-				not the reverse. There is no test that allows
-				us to declare an arbitrary system or technique
-				secure. This implies that claims of necessary
-				conditions for security are unfalsifiable. ''
-				\autocite{Herley2016unfalsifiability}
-		\end{itemize}
-\end{itemize}
-
-In this article, we give the first scientific theory (to the best of our knowledge)
-on security.
-
-\Paragraph{Structure} In Section~\ref{sec:problem} we define and formalize the problem statement. 
-In Section~\ref{sec:theory} we outline our security theory, and in Section~\ref{sec:prototest}
-we describe the implementation of the theory and some empirical tests of the theory.
-Finally, in Section~\ref{sec:related} we conclude the paper with an overview of the related work.
-
-\section{Problem Statement}\label{sec:problem}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.6\textwidth]{protocol-example.pdf}
-	\caption{Abstraction of an ad-hoc esemplificative protocol execution}
-	\label{fig:protocol-example}
-\end{figure}
-
-In\autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
-``an asymmetry in computer security'', which he defines as follows: ``Things
-can be declared insecure by observation, but not the reverse. There is no
-observation that allows us to declare an arbitrary system or technique
-secure''. Herley then uses this argument to show that ``claims that any measure
-is necessary for security are empirically unfalsifiable''. Given that, any
-theory which is not falsifiable by an empirical experiment is well
-known\footnote{``A theory which is not refutable by any conceivable event is
-nonscientific. Irrefutability is not a virtue of a theory (as people often
-think) but a vice.'' -- Karl Popper, Conjectures and
-Refutations\autocite{popper1962conjectures}} to be nonscientific (i.e.
-unfalsifiability is a fallacy of a theory), Herley concludes that there is no
-scientific theory on cybersecurity; which means that cybersecurity lays in the
-realm of pseudo-sciences\autocite{Herley2016usenixvideo}.  Herley, e.g.
-in\autocite{Herley2017justifying}, discusses the implications of a
-nonscientific approach to cybersecurity, and highlights the tremendous impact on
-all the scientific research and engineering of systems; leading often to
-terrorism and wars, and wasting of resources in useless protections or
-overspending.  While the criticism is investigated
-in\autocite{Herley2016unfalsifiability}, no solution is provided.  On the
-contrary, the goal of this work \emph{is} to lay the foundations of a
-scientific cybersecurity theory. Furthermore, in Section~\ref{sec:sicurezza},
-we consider the problem raised by Herley not confined to ``computer security''
-but to any abstract system (so that our theory may hold for any sound
-implementation such as networks, mechanical, cyber, or cyber-physical system,
-or even a single computer or a single device such as an hard-drive).  There is
-also an apparent inconsistency in\autocite{Herley2016unfalsifiability} that we
-seek to clarify before following (as we agree) the scientific path draw by
-Herley: cybersecurity is defined as an abstract property in many formal
-approaches to the investigation of the security of systems, and the security of the design of a
-formally verified protocol is indeed falsifiable.  For example, in the protocol
-verification community, security is often defined as a formalization of the
-high-level properties confidentiality, integrity, and availability. The problem
-in such approaches is not the definition of what cybersecurity is, but the use
-of theories (such as the Dolev-Yao attacker model\footnote{For the sake of
-simplicity, the Dolev-Yao attacker can be considered as an abstraction of an
-active attacker who controls the network but cannot break
-cryptography.}\autocite{Dolev1983security})
-that only applies to specific instances (often called scenarios) and
-abstraction of the protocol. This, in turn, creates a false sense of security
-since requires assumptions on the abstraction of the system of which security
-is verified. As an example, for the formal security verification of the system
-in Figure~\ref{fig:protocol-example}, a formalized scenario needs to be defined
-by a modeler who chooses (among others): (i) a scope of the formalization (e.g.
-excluding the server that distributes the public key is often done when
-verifying the security of authentication protocols), (ii) the number of
-sessions (even tough some approaches do reason on an infinite number of
-sessions such as\autocite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
-the peers (e.g.  in the ASLan++ language\autocite{Oheimb2010aslan++}), and (iv)
-the abstraction of the cryptographic primitives (e.g.  ProVerif vs
-CryptoVerif\autocite{Blanchet2017symbolic}).  Some of the choices will
-completely change the results of the formal verification of the system. For
-example, under the perfect cryptography assumption\footnote{As defined
-in\autocite{Rocchetto2016cpdy}: ``In the so called perfect cryptography
-assumption, the security encryption scheme is suppose to be perfect, without
-any exploitable flaw, and so the only way for the attacker to decrypt a message
-is by using the proper key. That assumption is widely accepted in the security
-protocol community, and most of the formal reasoning tools for the analysis of
-security protocols abstract away the mathematical and implementation details of
-the encryption
-scheme\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
-and assuming that no violation to any security property is done after message
-I); in Figure~\ref{fig:protocol-example}, the freedom of choosing the scope
-determines that the flaws related to the dishonest impersonation of the Server
-may or may not be considered in the verification process.  This choice has
-tremendous impact on the focus and findings of the verification of the security
-of the protocol.  While this may seem to turn upon minutiae and foreseeable,
-this highlights the false sense of security that may derive from a
-non-scientific theory of system security\footnote{``To the superficial
-observer, the analysis of these forms seems to turn upon minutiae. It does in
-fact deal with minutiae, but they are of the same order as those dealt with in
-microscopic anatomy.'' -- Karl Marx, Capital Volume 1, 1867}.
-
-\subsection{Sicurezza: Safety and Security}\label{sec:sicurezza}
-In most of the natural languages, and in Italian too, the concepts of safety
-and security are not syntactically differentiated and both terms (safety and
-security) are expressed by the same word, e.g. sicurezza in Italian.  A
-semantic distinction between safety and security is correlated to a
-belief\footnote{A belief has to be intended as a proposition which is supposed
-to be true by the majority of humans in our society without a scientific
-underlying theory but based on partial empirical evidences or inductive proofs.} that
-safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
-natural environment (e.g. natural events such as wearing of hardware
-components) while security deals with \emph{incidents} posed by mankind (e.g.
-attackers and bugs).  The fundamental difference between nature and mankind (and,
-in turn, between safety and cybersecurity) is believed to be on the different
-intents\footnote{``The belief–desire–intention software model (BDI) is a
-software model developed for programming intelligent
-agents.''\autocite{wiki-bdi}. In the BDI model, the intents represents the
-deliberative state of an agent which determines the choice of that agent on
-what to do.} (accidents are unfortunate while incidents are not) of the causes
-that generates the threat; namely, nature is believed not to have malicious
-intents (but unfortunate causes-effects) while threats generated by mankind are
-believed to be malicious\footnote{Of course, logical flaws or bugs may be
-introduced by other means (e.g. ignorance) without explicit malicious intents,
-but the exploitation of those flaws is considered (for now, and detailed
-afterwards in the article) malicious, and then we consider any vulnerability to
-be malicious (without loss of generality) even if due to the lack of skills.}.
-An overview on the aforementioned aspects of safety and security is depicted in
-Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
-the terms that structure our current understanding of safety and security. 
-\begin{itemize}
-	\item \emph{Mankind} ``refers collectively to humans''
-\autocite{wiki-mankind}, while the concept of \emph{Nature} is
-		related ``to the intrinsic characteristics that plants,
-		animals, and other features of the world develop of their own
-		accord'' (e.g. the physical universe)\autocite{wiki-nature}. 
-		\begin{itemize}
-			\item So far, we have used several terms to refer to an
-				\emph{attacker}, i.e. threat agent or threat source,
-				considering those terms to be semantically
-				equivalent.  This ``shallowness'' raise form the
-				necessity of properly citing the different sources, but,
-				in the reminder of this paper, we consider the
-				Causality principle to be the \emph{threat
-				source}, Nature or Mankind to be the
-				\emph{threat agents} and an \emph{attacker} as
-				a specific malicious threat agent which materialize a
-				threat.
-		\end{itemize}
-	\item \emph{Vulnerability}\footnote{The term vulnerability is not
-		present in the Encyclopedia of Cryptography and Security, while
-		it is used in 12 entries (such as in the definition of
-		``penetration testing''\autocite{caddy2005pentest})
-		highlighting how commonly this word is used without a proper
-		supporting semantics}, as defined in\autocite{cnssi20104009}
-		(and adopted in\autocite{nist2013800-53}), is ``weakness in an
-		information system, system security procedures, internal
-		controls, or implementation that could be exploited by a threat
-		source''. On the one hand, the definition is broad to enclose
-		as much causes (that generates a vulnerability) as possible; on
-		the other hand, it derives from empirical evidences (which
-		should be considered beliefs\footnote{``For this view, that
-		\emph{That Which Is Not} exists, can never predominate. You
-		must debar your thought from this way of search, nor let
-		ordinary experience in its variety force you along this way,
-		(namely, that of allowing) the eye, sightless as it is, and the
-		ear, full of sound, and the tongue, to rule; but (you must)
-		judge by means of the Reason (Logos) the much-contested proof
-		which is expounded by me.'' -- Parmenides of Elea, On Nature
-		(circa 500 B.C.), fragments B7.1–8.2
-		\autocite{Hakim2016philosophy}} since they are partial results in nature) 
-		while a vulnerability should
-		be defined in a way that is empirically falsifiable. This means
-		that the term vulnerability should have a complete and sound
-		definition, so that no other causes (e.g.  other sources) but
-		the ones in the definition are responsible for a vulnerability.
-		Furthermore, the term ``threat sources'' used in the definition
-		in\autocite{cnssi20104009} may be identified with both Nature
-		and Mankind, not differentiating between safety and security.
-		In Definition~\ref{def:vulnerability}, we provide a formal
-		theory of vulnerability (so that the scientific community can
-		identify tests for the completeness and soundness of the
-		definition itself).
-\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{safety-security.pdf}
-	\caption{Overview security and safety keywords}
-	\label{fig:safety-security}
-\end{figure}
-
-Most of the safety-preserving principles in the field of engineering of
-safety-critical cyber-physical systems (such as elevators and aircraft), upon
-which safety requirements (e.g. in standards such as the IEC 61508 or 61511\autocite{IEC201761511}) are defined, have
-been defined following empirical tests and measurements. While reasoning by
-induction based on the empirical observation should be avoided, since it may
-easily lead to false beliefs instead of scientific theories, this approach is
-often justified by the supposed impossibility of defining a theory that
-correctly predicts failures which, in turn, pose hazards to a system. 
-To the best of our knowledge, and supported by\autocite{Herley2016unfalsifiability}, the
-correlation between predictability of environment and believed unpredictability
-of attackers (i.e. a malicious
-environment) has not been correlated to a theory on cybersecurity. 
-Therefore, inductive research efforts in
-predicting malicious effects are accepted (and published) in scientific
-conferences (e.g.~\autocite{Rocchetto2014CSRF}). A failure of a wire due to
-environment (e.g. due to humidity, dust, heat \&c) is defined from empirical
-evidences and processes have been standardized to test qualities of hardware components
-%convinced by the induction principle. What is already a
-%weak argument in general, 
-This process completely breaks down when a malicious environment (i.e. an attacker)
-is considered instead of the (supposedly honest and predictable)
-natural environment. Therefore, the same approach that is in use for safety,
-seems not to be applicable to test security.
-
-Going back to Figure~\ref{fig:safety-security}, a vulnerability does not
-necessarily become a threat for the system, unless exploited ``through a
-channel that allows the violation of the security policy
-[\ldots]''\autocite{cnssi20104009} (e.g. a software or procedure) that takes
-advantage of the vulnerability causing an \emph{attack} to the system, which may
-result in several correlated incidents and threats.  The process of
-exploitation of a defect as a vulnerability is reported in
-Figure~\ref{fig:safety-security} such that the difference between exploit and failure,
-and attack and accident is to be found just in the maliciousness of the intents
-that causes this process (i.e. excluding the intent, the terms are just syntactic transformation from a vulnerability to defect, from
-accident to incident). In the following, we conclude the informal definition of
-the terms that we used in this section and in Figure~\ref{fig:safety-security}.
-
-\begin{itemize}
-	\item \emph{Causality} refers to the causality principle; defined
-		in\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
-		connection of phenomena through which one thing (the cause)
-		under certain conditions gives rise to, causes something else
-		(the effect). The essence of causality is the generation and
-		determination of one phenomenon by another. In this respect
-		causality differs from various other kinds of connection, for
-		example, the simple temporal sequence of phenomena, of the
-		regularities of accompanying processes''.
-	\item An \emph{Exploit}\footnote{We note that the term exploit is only
-		used as a verb in\autocite{ISO2009information}} is ``An exploit
-		(from the English verb to exploit, meaning to use something to
-		one’s own advantage) is a piece of software, a chunk of data,
-		or a sequence of commands that takes advantage of a bug or
-		vulnerability to cause unintended or unanticipated behavior to
-		occur on computer software, hardware, or something electronic
-		(usually computerized).''\autocite{wiki-exploit}.
-	\item An \emph{Attack}, as defined by the International Standard
-		ISO/IEC 27000 is an ``attempt to destroy, expose, alter,
-		disable, steal or gain unauthorized access to or make
-		unauthorized use of an asset''; where an \emph{Asset} is
-		``anything that has value to the organization''. We note that for
-		the purpose of this article, we do not want to focus on a specific
-		organization or business to define asset but, in general, on any 
-		abstract organization (e.g. a company or a society).
-		We do not consider ethical hackers as attacking a system. 
-		In fact, we consider the term \emph{hack} as
-		non-malicious (see Hacker\autocite{Stallman2002hacker}).
-	\item A \emph{Threat}, as defined in\autocite{cnssi20104009}, is ``Any
-		circumstance or event with the potential to adversely impact
-		organizational operations (including mission, functions, image,
-		or reputation), organizational assets, individuals, other
-		organizations, or the Nation through an information system via
-		unauthorized access, destruction, disclosure, modification of
-		information, and/or denial of service''.
-	\item \emph{Defect}, ``anything that renders the product not reasonably
-		safe''\autocite{Robinson2019writing} (i.e. a characteristic of
-		an object which hinders its proper usability).
-	\item \emph{Failure}, as defined in\autocite{Merriam2020failure} as ``a state of
-		inability to perform a normal function''. The term is
-		structured and detailed in
-		\autocite{cnssi20104009,iet2017glossary} but relying on an
-		abstract notion of failure without a specific definition.
-	\item \emph{Hazard}, ``a potential source of
-		harm''\autocite{iet2017glossary}.
-\end{itemize}
-
-\subsection{Glossary -- A Formalization}\label{sec:glossary}
 We now define a formalizations of the concepts described in
@@
 the next Section.
-
-\section{Honesty and Maliciousness -- A Cybersecurity Theory}\label{sec:theory}
-\begin{enumerate}[noitemsep]
-	\item cybersecurity as a requirement
-	\item vulnerability\label{def:vulnerability}
-\end{enumerate}
-
-We describe processes, in our current formalization, in terms of the K-Modal
-Structure as in Definition~\ref{def:modalinterpretation} (i.e. in terms of
-causality principle).
-
-\begin{definition}{\bf Vulnerability-Incident Process --}\label{def:vulnerability-incident}
-	Given two 
-\end{definition}
-
-The whole-part relation \fix{mr}{(i.e.  HAS-A\footnote{HAS-A is just an hint for the
-mechanization of this theory.  We'll come back to this in
-Section~\ref{sec:implementation}})} is given, in the following, over the K Modal
-logic described beforehand. To do so, we first need to formally define the
-concept of \emph{agent} in its abstract form. We build it following the same
-structure of\autocite{Santaca2016abf} but over the terms used in
-Section~\ref{sec:problem}, i.e. knowledge, and belief as epistemic concepts. We'll then 
-introduce the concept of \emph{intent} to define a \emph{threat agent}.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.8\textwidth]{knowledge-belief.pdf}
-	\caption{Informal representation of Knowledge and Belief}
-	\label{fig:knowledge-belief}
-\end{figure}
-
-\subsubsection{Region: Threat and Asset}\label{sec:threatasset}
-The difference between Knowledge and Belief is depicted in
-Figure~\ref{fig:knowledge-belief} (see \autocite{wiki-knowledgebelief}).  However,
-according to\autocite{Gettier2012knowledge},
-Knowledge\footnote{``\emph{Theaetetus}: [\ldots] He said that knowledge was
-true opinion accompanied by reason, but that unreasoning true opinion was
-outside of the sphere of knowledge; and matters of which there is not a
-rational explanation are unknowable -- yes, that is what he called them -- and
-those of which there is are knowable. [\ldots] \emph{Socrates}: [\ldots] the
-primary elements of which we and all else are composed admit of no rational
-explanation; for each alone by itself can only be named, and no qualification
-can be added, neither that it is nor that it is not, for that would at once be
-adding to it existence or non-existence, whereas we must add nothing to it, if
-we are to speak of that itself alone.  [\ldots]'' Plato -- Theaetetus 201
-\autocite{Plato1914Plato}} as an epistemological concept is difficult (and
-sometimes believed impossible\autocite{citation}) to formally define. In this
-work, we are not interested in 
-how knowledge can be precisely formalized from an epistemic standpoint. 
-We assume that a semantic of a correct definition of epistemic knowledge exists,
-for example the one given in\autocite{Hintikka1962knwoledge} by Hintikka, and we then define 
-knowledge in terms of the Kripke structure defined in Definition~\ref{def:causality}.
-
-\begin{definition}{\bf Knowledge --}\label{def:knowledge}
-Given an abstract collection of Agents $Ag$, and the modal operator
-	$\knows{a}{}$ (where $a\in Ag$), Knowledge is defined as a collection
-	of predicates known by an agent $\knowledge{a}=\bigcup_\Phi \knows{a}{\varphi}$ 
-	(where $\Phi$ is the collection of all the propositions known by $a$).
-	Given a proposition $P$, we extend the semantics of the Causality structure with:
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation6)$] $\world\models\knows{a}{P}$ iff
-			$\world'\models P$ for all $\world'$ such that
-			$\world\modalrelation\world'$.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Belief --}\label{def:belief}
-	\fix{mr}{we should help the reader in understanding the upgrades of the logic.
-We used propositional with modal operators with a K Kripke semantics and we are now introducing Belief as in Doxastic logic?
-and we have epistemological operators as knowledge}
-	Given an abstract collection of Agents $Ag$, and the modal operator
-	$\believe{a}{}$ (where $a\in Ag$), Belief is defined as a collection
-	of predicates believed by an agent $\belief{a}=\bigcup_\Phi \believe{a}{\varphi}$
-	(where $\Phi$ is the collection of all the propositions believed by $a$).
-	Given a proposition $P$, we extend the semantics of the Causality structure with:
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation7)$] $\world\models\believe{a}{P}$ iff
-			$\world\models\neg\knows{a}{\neg P}$ (i.e. the agent $a$ considers $P$ possible) 
-			and $\world'\models P$ for all
-			$\world'$ such that $\world\modalrelation\world'$.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Mankind/Nature --}\label{def:mankind-nature}
-Mankind/Nature is represented as an agent.  
-	%countable logical signature $\Sigma_M$/$\Sigma_N$ over an
-	%abstract algebraic structure.  We also consider a countable signature
-	%of variables, disjoint from $\Sigma_M$/$\Sigma_N$, and write
-	%$\Sigma_{M_n}$/$\Sigma_{N_n}$ for the symbols of $\Sigma_M$/$\Sigma_N$
-	%with arity $n$. Thus, $\Sigma_0$ is a collection of constants, which we
-	%assume to have disjoint sub-collections that we refer to as
-	%\emph{sub-agents} of Mankind or Nature (i.e. humans, plants, \&c).
-	%The variables are untyped and can be instantiated with arbitrary types,
-	%yielding an untyped model.
-\end{definition}
-
-\subsubsection{Mereotopological Representation of Safety and Security}\label{sec:formalGlossary}
-The objectives of this section are:
-\begin{enumerate}[noitemsep]
-	\item formalize the abstract algebraic structure (as a mereotopology)
-		on top of which we have formalized the terms in
-		Figure~\ref{fig:safety-security} so far, i.e. Mankind, Nature,
-		Truths, Beliefs, and Knowledge.
-	\item correlate mereological structure to Causality and then Kripke
-		structure by defining the relation $R$ over the Kripke
-		structure as the mereotopological basic relation connects with,
-		i.e. reflexive and symmetric
-	\item we map signature to mereotopological Regions to express the
-		logical formulation of terms (e.g. Mankind, Nature, Truths,
-		Beliefs, Knowledge, Vulnerabilities) over the mereotopological
-		structure
-	\item we formalize the relations between terms using the RCC(5)
-\end{enumerate}
-
-\fix{mr}{change propositions in propositional language\url{http://logic.stanford.edu/intrologic/lectures/lecture_02.pdf}}
-
-Given that the concepts of Truths, Beliefs, and Knowledge are positioned over
-an underlying structure of Propositions, the first concept to formalize is that
-of Propositions. We do not restrict Propositions to a specific structure but,
-as for Mankind and Nature, we refer to them as a collection.  We express the
-generality of a collection as a (propositional) signature over an abstract
-algebraic structure. In our formalization, we use mereotopology as algebraic
-structure (due to its generality, as discussed in Section~\ref{sec:problem})
-and the RCC to formalize concepts over a mereotopology.
-
-We can now define the possible mereotopological relations between Belief and
-Knowledge (informally depicted in Figure~\ref{fig:knowledge-belief}) and then
-define the peculiarity of the believed difference between Mankind and Nature 
-(i.e. maliciousness as intent). 
-We first start from the formalization of Figure~\ref{fig:knowledge-belief}, as
-follows.
-
-\begin{definition}{\bf Information}\label{def:information}
-\end{definition}
-\begin{definition}{\bf Desire}\label{def:desire}
-\end{definition}
-\begin{definition}{\bf Intent}\label{def:intent}
-\end{definition}
-
-\subsection{Cybersecurity Engineering Process}\label{sec:process}
-%\epigraph{La differenza tra idee e realt\`a \`e la differenza tra
-%filosofia e ingegneria. Il lavoro per trasformare l'una nell'altra \`e ricerca
-%scientifica
-\epigraph{The difference between ideas and reality is the
-difference between philosophy and engineering. The work to transform one into
-the other is scientific research}{{\itshape V-Research}}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{vmodel.pdf}
-	\caption{Cybersecurity Engineering Life-cycle}
-	\label{fig:knowledge-belief}
-\end{figure}
-
-\subsubsection{Relevant Standards}\label{sec:standards}
-\begin{enumerate}[noitemsep]
-	\item DO-326A
-\end{enumerate}
-
-\section{Prototype and Empirical Tests}\label{sec:prototest}
-\subsection{Prototype Implementation}\label{sec:implementation}
-\subsection{Empirical Tests}\label{sec:tests}
-
-\section{Conclusion and Related Work}\label{sec:related}
-
-\printbibliography
-
-\begin{appendices}
-	\section{RCC3, RCC5, or RCC8}\label{app:rcc}
-	\section{Wrong idea 1}
-In order to apply consistently the mereotopological structure not just to
-reason the agents but also on their evolution, we extend the properties of the
-Parthood relation (reflexivity and symmetry) to the modal relation
-$\modalrelation$ of the K Modal logic, which in turns defines the causality
-principle. This updates the K Modal logic to a B (Brouwer) Modal Logic\autocite{Garson2018modal}.
-
-\begin{definition}{\bf B Modal Logic --}\label{thm:bk-parthood}
-	The B Modal Logic (defined, with a slight abuse of notation as
-	$\bframe=\langle\possibleworlds,\modalrelation\rangle$) is an extension
-	of the K modal logic with the following axioms
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation8)$] $\world\models P\Rightarrow Q$ iff
-			$\world\not\models P$ or $\world\models Q$ (logical implication)~\footnote{In
-			Definition~\ref{def:modallogic} we didn't specify the
-			semantics in the case of the connective $\Rightarrow$
-			(i.e. logical implication), so we state it here. Given
-			that $P\Rightarrow Q$ is equisatisfiable to
-			$\neg(P\wedge \neg Q)$, for any proposition $P$ and
-			$Q$, $\world\models P\Rightarrow Q$ is equisatisfiable
-			to $\world\models \neg(P\wedge \neg Q)$. If
-			$\world\models \neg(P\wedge \neg Q)$ then
-			$\world\not\models P\wedge\neg Q$ and then
-			$\world\models \neg P$ or $\world\models Q$.}
-		\item[$(\interpretation9)$] $\Box P\Rightarrow P$ (which ``claims that whatever is necessary is the case''\autocite{Garson2018modal}), and
-		\item[$(\interpretation10)$] $P\Rightarrow\Box\Diamond P$
-			(which says that if $P$ is the case, then $P$ is
-			necessarily possible\autocite{Garson2018modal})
-	\end{enumerate}
-\end{definition}
-
-\begin{theorem}{\bf Causality, Parthood, and B Modal Logic}\label{thm:bk-parthood}
-	\begin{enumerate}[noitemsep]
-		\item (Reflexivity) $\bframe\models\Box P\Rightarrow P$ ($\interpretation8$) iff $\modalrelation$ is reflexive
-		\item (Symmetry) $\bframe\models P\Rightarrow\Box\Diamond$ ($\interpretation9$) iff $\modalrelation$ is symmetric
-		\item (Consistency) $\world\modalrelation\world'$ iff $\connects{\region}{\region'}$
-	\end{enumerate}
-	whenever $\world$ and $\world'$, and $\region$ and $\region'$ are logically equivalent ($\equiv$).
-\end{theorem}
-\begin{proof}
-	The proof is exhaustive over all possible cases.
-	\begin{enumerate}[noitemsep]
-		\item(Reflexivity) proven in\autocite{}
-		\item(Symmetry) proven in\autocite{}
-		\item(Equivalence) \fix{mr}{todo} If
-			$\world\modalrelation\world'$ and
-			$\world\equiv\region$, then if
-			it doesn't hold that $\connects{\region,\region'}$ it
-			follows that there is no reflexive and transitive
-			relation that correlates $\region$ to $\region'$. Given 
-			that Regions are defined as signatures
-	\end{enumerate}
-\end{proof}
-
-\end{appendices}
-
-\end{document}

### 2020-02-20 theory.tex
restructure

--- 2020-02-19 v-research_experiment-alpha.tex
@@
-\documentclass{article}
-\usepackage{mathtools}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{enumitem}
-\usepackage{multicol}
-\usepackage[toc,page]{appendix}
-\usepackage[gen]{eurosym}
-\usepackage{tabularx}
-\usepackage{epigraph}
-\usepackage{textcomp}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{csquotes}% Recommended
-\usepackage{rotating}
-\usepackage[style=numeric]{biblatex}
-% \bibliography{<mybibfile>}% ONLY selects .bib file; syntax for version <= 1.1b
-\addbibresource{literature.bib}% Syntax for version >= 1.2
-
-\usepackage{graphicx}
-\graphicspath{{./images/}}
-
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\lhead{}
-\chead{}
-\rhead{}
-\lfoot{}
-\cfoot{\thepage}
-\rfoot{Property of V-Research S.r.l.}
-\renewcommand\headrulewidth{0pt}
-\renewcommand\footrulewidth{0pt}
-
-\usepackage[colorinlistoftodos]{todonotes}
-
-\newcommand{\Paragraph}[1]{\smallskip\noindent{\bf #1.}}
-\newcommand{\fixnote}[2]{\textbf{\color{red}{FIX}}\footnote{{\bf #1:} #2}}
-\newcommand{\fix}[2]{{\color{red} {\bf tofix:} #2}}
-% \renewcommand{\fixnote}[2]{}
-% \renewcommand{\fix}[2]{}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\theoremstyle{corollary}
-\newtheorem{corollary}{Corollary}[section]
-\theoremstyle{lemma}
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{theorem}
-\newtheorem{theorem}{Theorem}[section]
-\theoremstyle{theorem}
-\newtheorem{example}{Example}
-
-\usepackage{authblk}
-\date{}                     %% if you don't need date to appear
-\setcounter{Maxaffil}{0}
-\renewcommand\Affilfont{\itshape\small}
-
-% for fancy table with rotated column titles
-\usepackage{adjustbox}
-\newcolumntype{R}[2]{%
-    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
-    l%
-    <{\egroup}%
-}
-\newcommand*\rot{\multicolumn{1}{R{45}{0.01em}}}% no optional argument here, please!
-\newcommand*\rota{\multicolumn{1}{R{90}{0.01em}}}% no optional argument here, please!
-\def\hb{\hbox to 10.7 cm{}}
-
-% circles
-\usepackage{wasysym}
-\newcommand{\Tdot}{$\CIRCLE$}
-\newcommand{\Thdot}{$\LEFTcircle$}
-\newcommand{\Twdot}{$\Circle$}
-
-% formulas
-\newcommand{\kframe}{\mathbf{K}}
-\newcommand{\bframe}{\mathbf{B_K}}
-\newcommand{\possibleworlds}{G}
-\newcommand{\modalrelation}{R}
-\newcommand{\actualworld}{\omega*}
-\newcommand{\world}{\omega}
-\newcommand{\World}{\Omega}
-\newcommand{\kmodel}{M}
-\newcommand{\interpretation}{\sigma}
-\newcommand{\knowledge}[1]{\mathbb{K}_{#1}}
-\newcommand{\knows}[2]{K_{#1}#2}
-\newcommand{\belief}[1]{\mathbb{B}_{#1}}
-\newcommand{\believe}[2]{B_{#1}#2}
-\newcommand{\region}{\chi}
-\newcommand{\Region}{\mathrm{X}}
-\newcommand{\agentuniverse}{\mathcal{A}}
-\newcommand{\agent}{a}
-\newcommand{\follang}{\mathcal{L}}
-\newcommand{\weakness}{w}
-\newcommand{\Weakness}{W}
-\newcommand{\vulnerability}{v}
-\newcommand{\Vulnerability}{V}
-%rcc5
-\newcommand{\rcc}{rcc}
-\newcommand{\connects}[2]{C(#1,#2)}
-\newcommand{\disconnected}[2]{\neg C(#1,#2)}
-\newcommand{\partof}[2]{P(#1,#2)}
-\newcommand{\overlaps}[2]{O(#1,#2)}
-\newcommand{\eq}[2]{EQ(#1,#2)}
-\newcommand{\pp}[2]{PP(#1,#2)}
-\newcommand{\po}[2]{PO(#1,#2)}
-\newcommand{\ppi}[2]{PPi(#1,#2)}
-\newcommand{\dr}[2]{DR(#1,#2)}
-
-\makeindex
-
-\begin{document}
-
-\title{On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}}
-\author[1]{Francesco Beltramini}
-\author[1]{Marco Rocchetto}
-\author[2]{Luca Vigan\`o}
-\affil[1]{V-Research, Verona, Italy}
-\affil[2]{King's College London, UK}
-
-\maketitle
-
-\begin{abstract}
-	The objective of this research is to develop of a theory that defines
-	(all and only) the possible insecurity and security configurations of
-	any abstract system. The theory is structured upon other theories that 
-	defines how a component of a system can be abstracted into an agent, defining how
-	agents can be formalized (both syntactically and semantically) to
-	describe an abstract system, such as a graph. Some of these theories
-	(e.g. used for the semantic
-	definition of the abstract system) are the epistemological definition
-	of knowledge, the Belief-Desire-Intent and the Assertion-Belief-Fact 
-	framework of reference, 
-	mereology, and topological structure. We argue that a mereology is the
-	most appropriate abstract underlying structure, do to its generality,
-	for defining the expressiveness of the system abstraction.  Furthermore, a
-	mereology allows us to define an ontology rather than a taxonomy.  We
-	also correlate different abstractions of the system to the TRL and the
-	engineering V-model.
-
-	We implemented a formal theory (of axioms) of a mereotopology, and of
-	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
-	uses the Z3 SMT solver. The results show that a single component (i.e.
-	agent) of an abstract system has a definite number of  different
-	insecurity configurations (e.g. 53 using RCC5 over a topological
-	structure) and only 1 secure (i.e. expected) configurations. The
-	configurations are reported as models satisfying the abstract system
-	semantics.
-
-	We considered the philosophical definition of truth behind our
-	approach, rejecting ``proof'' by induction from partial empirical evidences.
-	Our theory can be applied to system engineering and we show a concrete
-	application of our theory to the risk assessment of an ad-hoc system.
-	Finally, we provide a number of ideas to support the engineering of
-	secure systems (e.g. purely cyber or cyber-physical).
-\end{abstract}
-\newpage
-
-\section{Introduction}\label{sec:intro}
-\epigraph{Humanum est errare}{{\itshape Seneca the Elder}}
-The European Commission states in\autocite{EU2019market} that: ``Cybersecurity
-is one of the priority areas [\ldots] of the Commission initiative on ICT
-Standards, which is part of the Digitising European
-Industry\autocite{EU2019standard} strategy launched on 19 April 2016.  The aim
-is to identify the essential ICT standards and present measures to accelerate
-their development in support of digital innovations across the economy''. The
-same document (i.e.\autocite{EU2019market}) states that ``The EU will invest up
-to \euro450 million [\ldots], under its research and innovation programme
-Horizon 2020''. The EU, in 2016 published a press release\autocite{EU2016press}
-in which they present a strategy to invest \euro1.8 \emph{billion} to
-``increase measures to address cyber threats''. The EU is not the only investor
-in cybersecurity, most of the developed countries and several companies are
-investing enormous amount of money towards various aspects of cybersecurity
-(e.g. The US vulnerability databases\autocite{NIST2020NVD} maintained by the
-National Institute of Standards and Technologies, i.e. NIST, of the US
-Department of Commerce).
-
-The cybersecurity industry is growing fast, e.g. as reported
-in\autocite{Nasdaq2018market}. For example, in\autocite{Forbes2017market},
-published by the Forbes, is stated that \euro5.3 billion of funding where
-poured by venture capitalist into cybersecurity companies in 2018. The Forbes,
-in the same article, also highlights another peculiar (as seemly contradictory)
-trend: ``[\ldots] during the same time period, the number of cybersecurity
-breaches increased exponentially''. The data reported by the NIST 
-through the official CPE (Common Platform Enumeration) Dictionary Statistics 
-on the NVD websites in \autocite{NIST2020CPEstatistics}, show that in 2016 the
-number of reported vulnerabilities reported where around $~6000$ while in 2019
-the number of vulnerabilities was above $16000$.  The scientific community also
-reports similar findings.  In fact, in\autocite{Herley2009so}, Cormac Herley
-(Microsoft Research) shows how basic cybersecurity principles (such as the
-confidentiality benefit over the clear text for passwords typed into forms,
-e.g. for logins in websites) are not fully understood or shared between the
-cybersecurity research community\autocite{Nielsen2009stop}.  The lack of
-understanding of basic security principle, the inverse proportionality between
-investments in cybersecurity and the number of reported vulnerabilities year
-after year, can be linked to the lack of a foundational theory on
-cybersecurity, as already highlighted by Cormac Herley
-in\autocite{Herley2016unfalsifiability}.
-
-In this article, we give the first scientific theory (to the best of our knowledge)
-on security.
-
-\Paragraph{Structure} In Section~\ref{sec:problem} we define and formalize the problem statement. 
-In Section~\ref{sec:theory} we outline our security theory, and in Section~\ref{sec:prototest}
-we describe the implementation of the theory and some empirical tests of the theory.
-Finally, in Section~\ref{sec:related} we conclude the paper with an overview of the related work.
-
-\section{Problem Statement}\label{sec:problem}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.6\textwidth]{protocol-example.pdf}
-	\caption{Abstraction of an ad-hoc esemplificative protocol execution}
-	\label{fig:protocol-example}
-\end{figure}
-
-In\autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
-``an asymmetry in computer security'', which he defines as follows: ``Things
-can be declared insecure by observation, but not the reverse. There is no
-observation that allows us to declare an arbitrary system or technique
-secure''. Herley then uses this argument to show that ``claims that any measure
-is necessary for security are empirically unfalsifiable''. Given that, any
-theory which is not falsifiable by an empirical experiment is well
-known\footnote{``A theory which is not refutable by any conceivable event is
-nonscientific. Irrefutability is not a virtue of a theory (as people often
-think) but a vice.'' -- Karl Popper, Conjectures and
-Refutations\autocite{popper1962conjectures}} to be nonscientific (i.e.
-unfalsifiability is a fallacy of a theory), Herley concludes that there is no
-scientific theory on cybersecurity; which means that cybersecurity lays in the
-realm of pseudo-sciences\autocite{Herley2016usenixvideo}.  Herley, e.g.
-in\autocite{Herley2017justifying}, discusses the implications of a
-nonscientific approach to cybersecurity, and highlights the tremendous impact on
-all the scientific research and engineering of systems; leading often to
-terrorism and wars, and wasting of resources in useless protections or
-overspending.  While the criticism is investigated
-in\autocite{Herley2016unfalsifiability}, no solution is provided.  On the
-contrary, the goal of this work \emph{is} to lay the foundations of a
-scientific cybersecurity theory. Furthermore, in Section~\ref{sec:sicurezza},
-we consider the problem raised by Herley not confined to ``computer security''
-but to any abstract system (so that our theory may hold for any sound
-implementation such as networks, mechanical, cyber, or cyber-physical system,
-or even a single computer or a single device such as an hard-drive).  There is
-also an apparent inconsistency in\autocite{Herley2016unfalsifiability} that we
-seek to clarify before following (as we agree) the scientific path draw by
-Herley: cybersecurity is defined as an abstract property in many formal
-approaches to the investigation of the security of systems, and the security of the design of a
-formally verified protocol is indeed falsifiable.  For example, in the protocol
-verification community, security is often defined as a formalization of the
-high-level properties confidentiality, integrity, and availability. The problem
-in such approaches is not the definition of what cybersecurity is, but the use
-of theories (such as the Dolev-Yao attacker model\footnote{For the sake of
-simplicity, the Dolev-Yao attacker can be considered as an abstraction of an
-active attacker who controls the network but cannot break
-cryptography.}\autocite{Dolev1983security})
-that only applies to specific instances (often called scenarios) and
-abstraction of the protocol. This, in turn, creates a false sense of security
-since requires assumptions on the abstraction of the system of which security
-is verified. As an example, for the formal security verification of the system
-in Figure~\ref{fig:protocol-example}, a formalized scenario needs to be defined
-by a modeler who chooses (among others): (i) a scope of the formalization (e.g.
-excluding the server that distributes the public key is often done when
-verifying the security of authentication protocols), (ii) the number of
-sessions (even tough some approaches do reason on an infinite number of
-sessions such as\autocite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
-the peers (e.g.  in the ASLan++ language\autocite{Oheimb2010aslan++}), and (iv)
-the abstraction of the cryptographic primitives (e.g.  ProVerif vs
-CryptoVerif\autocite{Blanchet2017symbolic}).  Some of the choices will
-completely change the results of the formal verification of the system. For
-example, under the perfect cryptography assumption\footnote{As defined
-in\autocite{Rocchetto2016cpdy}: ``In the so called perfect cryptography
-assumption, the security encryption scheme is suppose to be perfect, without
-any exploitable flaw, and so the only way for the attacker to decrypt a message
-is by using the proper key. That assumption is widely accepted in the security
-protocol community, and most of the formal reasoning tools for the analysis of
-security protocols abstract away the mathematical and implementation details of
-the encryption
-scheme\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
-and assuming that no violation to any security property is done after message
-I); in Figure~\ref{fig:protocol-example}, the freedom of choosing the scope
-determines that the flaws related to the dishonest impersonation of the Server
-may or may not be considered in the verification process.  This choice has
-tremendous impact on the focus and findings of the verification of the security
-of the protocol.  While this may seem to turn upon minutiae and foreseeable,
-this highlights the false sense of security that may derive from a
-non-scientific theory of system security\footnote{``To the superficial
-observer, the analysis of these forms seems to turn upon minutiae. It does in
-fact deal with minutiae, but they are of the same order as those dealt with in
-microscopic anatomy.'' -- Karl Marx, Capital Volume 1, 1867}.
-
-\subsection{Sicurezza: Safety and Security}\label{sec:sicurezza}
-In most of the natural languages, and in Italian too, the concepts of safety
-and security are not syntactically differentiated and both terms (safety and
-security) are expressed by the same word, e.g. sicurezza in Italian.  A
-semantic distinction between safety and security is correlated to a
-belief\footnote{A belief has to be intended as a proposition which is supposed
-to be true by the majority of humans in our society without a scientific
-underlying theory but based on partial empirical evidences or inductive proofs.} that
-safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
-natural environment (e.g. natural events such as wearing of hardware
-components) while security deals with \emph{incidents} posed by mankind (e.g.
-attackers and bugs).  The fundamental difference between nature and mankind (and,
-in turn, between safety and cybersecurity) is believed to be on the different
-intents\footnote{``The belief–desire–intention software model (BDI) is a
-software model developed for programming intelligent
-agents.''\autocite{wiki-bdi}. In the BDI model, the intents represents the
-deliberative state of an agent which determines the choice of that agent on
-what to do.} (accidents are unfortunate while incidents are not) of the causes
-that generates the threat; namely, nature is believed not to have malicious
-intents (but unfortunate causes-effects) while threats generated by mankind are
-believed to be malicious\footnote{Of course, logical flaws or bugs may be
-introduced by other means (e.g. ignorance) without explicit malicious intents,
-but the exploitation of those flaws is considered (for now, and detailed
-afterwards in the article) malicious, and then we consider any vulnerability to
-be malicious (without loss of generality) even if due to the lack of skills.}.
-An overview on the aforementioned aspects of safety and security is depicted in
-Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
-the terms that structure our current understanding of safety and security. 
-\begin{itemize}
-	\item \emph{Mankind} ``refers collectively to humans''
-\autocite{wiki-mankind}, while the concept of \emph{Nature} is
-		related ``to the intrinsic characteristics that plants,
-		animals, and other features of the world develop of their own
-		accord'' (e.g. the physical universe)\autocite{wiki-nature}. 
-		\begin{itemize}
-			\item So far, we have used several terms to refer to an
-				\emph{attacker}, i.e. threat agent or threat source,
-				considering those terms to be semantically
-				equivalent.  This ``shallowness'' raise form the
-				necessity of properly citing the different sources, but,
-				in the reminder of this paper, we consider the
-				Causality principle to be the \emph{threat
-				source}, Nature or Mankind to be the
-				\emph{threat agents} and an \emph{attacker} as
-				a specific malicious threat agent which materialize a
-				threat.
-		\end{itemize}
-	\item \emph{Vulnerability}\footnote{The term vulnerability is not
-		present in the Encyclopedia of Cryptography and Security, while
-		it is used in 12 entries (such as in the definition of
-		``penetration testing''\autocite{caddy2005pentest})
-		highlighting how commonly this word is used without a proper
-		supporting semantics}, as defined in\autocite{cnssi20104009}
-		(and adopted in\autocite{nist2013800-53}), is ``weakness in an
-		information system, system security procedures, internal
-		controls, or implementation that could be exploited by a threat
-		source''. On the one hand, the definition is broad to enclose
-		as much causes (that generates a vulnerability) as possible; on
-		the other hand, it derives from empirical evidences (which
-		should be considered beliefs\footnote{``For this view, that
-		\emph{That Which Is Not} exists, can never predominate. You
-		must debar your thought from this way of search, nor let
-		ordinary experience in its variety force you along this way,
-		(namely, that of allowing) the eye, sightless as it is, and the
-		ear, full of sound, and the tongue, to rule; but (you must)
-		judge by means of the Reason (Logos) the much-contested proof
-		which is expounded by me.'' -- Parmenides of Elea, On Nature
-		(circa 500 B.C.), fragments B7.1–8.2
-		\autocite{Hakim2016philosophy}} since they are partial results in nature) 
-		while a vulnerability should
-		be defined in a way that is empirically falsifiable. This means
-		that the term vulnerability should have a complete and sound
-		definition, so that no other causes (e.g.  other sources) but
-		the ones in the definition are responsible for a vulnerability.
-		Furthermore, the term ``threat sources'' used in the definition
-		in\autocite{cnssi20104009} may be identified with both Nature
-		and Mankind, not differentiating between safety and security.
-		In Definition~\ref{def:vulnerability}, we provide a formal
-		theory of vulnerability (so that the scientific community can
-		identify tests for the completeness and soundness of the
-		definition itself).
-\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{safety-security.pdf}
-	\caption{Overview security and safety keywords}
-	\label{fig:safety-security}
-\end{figure}
-
-Most of the safety-preserving principles in the field of engineering of
-safety-critical cyber-physical systems (such as elevators and aircraft), upon
-which safety requirements (e.g. in standards such as the IEC 61508 or 61511\autocite{IEC201761511}) are defined, have
-been defined following empirical tests and measurements. While reasoning by
-induction based on the empirical observation should be avoided, since it may
-easily lead to false beliefs instead of scientific theories, this approach is
-often justified by the supposed impossibility of defining a theory that
-correctly predicts failures which, in turn, pose hazards to a system. 
-To the best of our knowledge, and supported by\autocite{Herley2016unfalsifiability}, the
-correlation between predictability of environment and believed unpredictability
-of attackers (i.e. a malicious
-environment) has not been correlated to a theory on cybersecurity. 
-Therefore, inductive research efforts in
-predicting malicious effects are accepted (and published) in scientific
-conferences (e.g.~\autocite{Rocchetto2014CSRF}). A failure of a wire due to
-environment (e.g. due to humidity, dust, heat \&c) is defined from empirical
-evidences and processes have been standardized to test qualities of hardware components
-%convinced by the induction principle. What is already a
-%weak argument in general, 
-This process completely breaks down when a malicious environment (i.e. an attacker)
-is considered instead of the (supposedly honest and predictable)
-natural environment. Therefore, the same approach that is in use for safety,
-seems not to be applicable to test security.
-
-Going back to Figure~\ref{fig:safety-security}, a vulnerability does not
-necessarily become a threat for the system, unless exploited ``through a
-channel that allows the violation of the security policy
-[\ldots]''\autocite{cnssi20104009} (e.g. a software or procedure) that takes
-advantage of the vulnerability causing an \emph{attack} to the system, which may
-result in several correlated incidents and threats.  The process of
-exploitation of a defect as a vulnerability is reported in
-Figure~\ref{fig:safety-security} such that the difference between exploit and failure,
-and attack and accident is to be found just in the maliciousness of the intents
-that causes this process (i.e. excluding the intent, the terms are just syntactic transformation from a vulnerability to defect, from
-accident to incident). In the following, we conclude the informal definition of
-the terms that we used in this section and in Figure~\ref{fig:safety-security}.
-
-\begin{itemize}
-	\item \emph{Causality} refers to the causality principle; defined
-		in\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
-		connection of phenomena through which one thing (the cause)
-		under certain conditions gives rise to, causes something else
-		(the effect). The essence of causality is the generation and
-		determination of one phenomenon by another. In this respect
-		causality differs from various other kinds of connection, for
-		example, the simple temporal sequence of phenomena, of the
-		regularities of accompanying processes''.
-	\item An \emph{Exploit}\footnote{We note that the term exploit is only
-		used as a verb in\autocite{ISO2009information}} is ``An exploit
-		(from the English verb to exploit, meaning to use something to
-		one’s own advantage) is a piece of software, a chunk of data,
-		or a sequence of commands that takes advantage of a bug or
-		vulnerability to cause unintended or unanticipated behavior to
-		occur on computer software, hardware, or something electronic
-		(usually computerized).''\autocite{wiki-exploit}.
-	\item An \emph{Attack}, as defined by the International Standard
-		ISO/IEC 27000 is an ``attempt to destroy, expose, alter,
-		disable, steal or gain unauthorized access to or make
-		unauthorized use of an asset''; where an \emph{Asset} is
-		``anything that has value to the organization''. We note that for
-		the purpose of this article, we do not want to focus on a specific
-		organization or business to define asset but, in general, on any 
-		abstract organization (e.g. a company or a society).
-		We do not consider ethical hackers as attacking a system. 
-		In fact, we consider the term \emph{hack} as
-		non-malicious (see Hacker\autocite{Stallman2002hacker}).
-	\item A \emph{Threat}, as defined in\autocite{cnssi20104009}, is ``Any
-		circumstance or event with the potential to adversely impact
-		organizational operations (including mission, functions, image,
-		or reputation), organizational assets, individuals, other
-		organizations, or the Nation through an information system via
-		unauthorized access, destruction, disclosure, modification of
-		information, and/or denial of service''.
-	\item \emph{Defect}, ``anything that renders the product not reasonably
-		safe''\autocite{Robinson2019writing} (i.e. a characteristic of
-		an object which hinders its proper usability).
-	\item \emph{Failure}, as defined in\autocite{Merriam2020failure} as ``a state of
-		inability to perform a normal function''. The term is
-		structured and detailed in
-		\autocite{cnssi20104009,iet2017glossary} but relying on an
-		abstract notion of failure without a specific definition.
-	\item \emph{Hazard}, ``a potential source of
-		harm''\autocite{iet2017glossary}.
-\end{itemize}
-
-\subsection{Glossary -- A Formalization}\label{sec:glossary}
-We now define a formalizations of the concepts described in
-Section~\ref{sec:problem} and depicted in Figure~\ref{fig:safety-security}. We
-base our formalization on first order logic (FOL) with the standard
-truth-value semantics. The choice of this logic is required by the 
-semantics of the concepts formalized afterwards in this section.
-
-We consider a first-order language $\follang$ over a
-signature $\Sigma_P$ where $P,P',\ldots,P^n$ represent terms, and
-$\varphi,\varphi',\ldots,\varphi^n$ represent formulas.
-The syntax is defined as follows.
-\begin{displaymath}
-	\varphi := P~|~\neg\varphi~|~\varphi\wedge\varphi~|~\varphi\vee\varphi~|~\varphi\Rightarrow\varphi'~|~\forall P.~\varphi~|~\exists P.~\varphi
-\end{displaymath}
-where $\wedge$, $\neg$, $\vee$, and $\Rightarrow$ are connectives representing
-conjunction, negation, disjunction, and (material) implication respectively; while
-$\forall$ and $\exists$ represent the standard universal and existential
-quantifiers (resp.). Finally, the symbol ``$.$'' is just syntactic sugar. 
-We consider $\interpretation\subset\Phi\times\{\top,\bot\}$ as the interpretation
-function, where $\Phi$ is any collection of sentences in $\follang$ 
-and $\top$ and $\bot$ represent the concepts of ``Tautology/True''
-and ``Contradiction/False'' respectively.
-
-Following Figure~\ref{fig:safety-security}, we start by formalizing the outermost term: Causality.
-
-\subsubsection{Causality Principle}\label{sec:causality}
-We formalize the Causality Principle starting from a K modal
-logic\autocite{Garson2018modal} (i.e. without restrictions on the causality
-relation between worlds). The standard definition of K modal logic is given in
-Definition~\ref{def:modallogic} in terms of an interpretation function (which
-we named $\sigma$ with a slight abuse of notation) defined in
-Definition~\ref{def:modalinterpretation}. 
-\begin{definition}{\bf K Modal Logic --}\label{def:modallogic}
-A K-frame is a frame $ \kframe=\langle \possibleworlds,\modalrelation \rangle $
-	in the K modal logic where $ \modalrelation $ is the binary relation
-	(i.e. a set of ordered pairs) between possible worlds
-	$\modalrelation\subseteq\possibleworlds\times\possibleworlds$, where
-	$\possibleworlds$ represents the possible worlds, and
-	$\possibleworlds\neq\emptyset$. An actual world
-	$\actualworld\in\possibleworlds$ is assumed. For any proposition $P$,
-	an interpretation function $\interpretation(\world,P)$ returns the
-	truth value of $P$; e.g. $\interpretation(\world,P)=\top$ means that
-	$P$ holds in $\world$. A model is defined as the tuple
-	$\kmodel=\langle\possibleworlds,\modalrelation,\interpretation\rangle$.
-\end{definition}
-
-\begin{definition}{\bf Causality as K Modal Interpretation Function --}\label{def:modalinterpretation}
-	Causality is (recursively) defined as the modal interpretation function $\interpretation$, as follows. 
 \begin{enumerate}[noitemsep]
-		\item[$(\interpretation0)$] if $\interpretation(\world,P)=\top$ then $\world\models P$
-		\item[$(\interpretation1)$] $\world\models\neg P$ iff $\world\not\models P$
-		\item[$(\interpretation2)$] $\world\models P \wedge Q$ iff $\world\models P$ and $\world\models Q$
-		\item[$(\interpretation3)$] $\world\models\Box P$ iff for any world $\world'\in\possibleworlds$ if $\world \modalrelation\world'$ then $\world'\models P$
-		\item[$(\interpretation4)$] $\world\models\Diamond P$ iff there exists a set of worlds $\World'\subset\possibleworlds$ such that for any $\world'\in\World'$, if $\world\modalrelation\world'$ then $\world'\models P$
-		\item[$(\interpretation5)$] $\models P$ iff $\actualworld\models P$
+	\item cybersecurity as a requirement
+	\item vulnerability\label{def:vulnerability}
 \end{enumerate}
-	where truth is defined as necessary with $\Box$ and possible with $\Diamond$.
-\end{definition}
-
-%\begin{definition}{\bf Causality --}\label{def:causality}
-%	The Causality principle is defined over a K Modal Logic (i.e. without
-%	restrictions on the causality relation between worlds) as in
-%	Definition~\ref{def:modallogic}, with the standard interpretation
-%	function defined in~\ref{def:modalinterpretation}
-%\end{definition}
-
-The causality principle has been defined in its
-generic form. In fact, the accessibility relation $\modalrelation$ is free from
-any axiomatic restriction (e.g. it's not reflexive nor anti-reflexive).  
-We will focus in Section~\ref{sec:process} (for our tests) on 
-the application of our theory to CPS system engineering.
-More detailed case studies (i.e. \fix{mr}{a CPS as a smart power-grid}) 
-will be defined in Section~\ref{sec:prototest}, where
-the accessibility relation defines in details how the system itself evolves, due to causal relation 
-(i.e. by restricting the causality principle only to those cause-effect that defines
-the system). Therefore, the definition of $\modalrelation$
-will be specialized in a more strict way based on the application domains and case study.
-
-We note that we have defined the
-Causality principle without considering it a \emph{threat source} since 
-we lack the concept of intent and maliciousness. Similarly, in the 
-next sections we won't discriminate between Nature and Mankind until we introduce 
-the concepts of maliciousness in Section~\ref{sec:theory} 
-and then formally define a threat agent
-in the same section.
-
-\subsubsection{Agents: Mankind and Nature}\label{sec:mankind-nature}
-\begin{table}[t]
-\centering
-\setlength{\tabcolsep}{3.5pt}
-\renewcommand{\arraystretch}{1}
-\scriptsize
-\caption{RCC3, RCC5, and RCC8 relations between Regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
-\begin{tabular}{ccclll} 
-\rota{\textbf{RCC3}}&\rota{\textbf{RCC5}}&\rota{\textbf{RCC8}}&\textbf{Name} & \textbf{Notation} & \textbf{Definition} \\
-\hline
-&&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& $\mathit{X}\subseteq \mathit{Y}$ \\
-&&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& $\mathit{X}\not\subseteq \mathit{Y}$\\
-&&&Part of				& $\mathit{P}(\mathit{X},\mathit{Y})$		& $\forall \mathit{Z} ~\mathit{C}(\mathit{Z},\mathit{X}) \rightarrow \mathit{C}(\mathit{Z},\mathit{Y})$\\
-&&&Overlaps			& $\mathit{O}(\mathit{X},\mathit{Y})$		& $\exists \mathit{Z} ~\mathit{P}(\mathit{Z},\mathit{X})\wedge \mathit{P}(\mathit{Z},\mathit{Y})$\\
-\Tdot&&&  \textbf{Overlaps Not Equal} 	& $\mathit{ONE}(\mathit{X},\mathit{Y})$		& $\mathit{O}(\mathit{X},\mathit{Y}) \land \neg \mathit{EQ}(\mathit{X},\mathit{Y})$ \\
-\Tdot&\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
-\Tdot&\Tdot&\Tdot& \textbf{DiscRete from} 		& $\mathit{DR}(\mathit{X},\mathit{Y})$		& $\neg \mathit{O}(\mathit{X},\mathit{Y})$\\
-&\Tdot&\Tdot&\textbf{Partial-Overlap}	& $\mathit{PO}(\mathit{X},\mathit{Y})$ 		& $\mathit{O}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{X},\mathit{Y}) \wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\
-&\Tdot&&\textbf{Proper-Part-of} 	& $\mathit{PP}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\ 
-	&\Tdot&&\textbf{Proper-Part-of-\textit{\textbf{i}}nverse} & $\mathit{PPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{Y},\mathit{X}) \wedge \neg \mathit{P}(\mathit{X},\mathit{Y})$\\
-&&\Tdot&\textbf{Externally Connected} 	& $\mathit{EC}(\mathit{X},\mathit{Y})$ 		& $\mathit{C}(\mathit{X},\mathit{Y}) \wedge \neg\mathit{O}(\mathit{X},\mathit{Y})$\\ 
-&&\Tdot&\textbf{Tangential PP} 	& $\mathit{TPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Tangential PPi} 	& $\mathit{TPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{TPP}(\mathit{Y},\mathit{X})$\\ 
-&&\Tdot&\textbf{Non-Tangential PP} 	& $\mathit{NTPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\neg\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Non-Tangential PPi} 	& $\mathit{NTPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{NTPP}(\mathit{Y},\mathit{X})$\\ 
-\end{tabular}
-\end{table}
-
-Mankind and Nature are defined in Section~\ref{sec:problem} as two abstract
-agents, both as collections (i.e. an abstract type that does not imply a
-specific implementation) of their sub-agents (i.e. humans for Mankind and
-plants, animals, \&c.  for Nature). Similarly to\autocite{Santaca2016abf}, 
-we define Mankind and Nature, and any
-other agent in the reminder of this article, as a meronomy (an hierarchy of
-Part-Whole relations) based on a standard definition of mereology, i.e. based
-on the definition of Parthood relation between \emph{Parts}.  
-However, shall 
-consider different types of Part; so, we extend the mereology to a
-mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
-to increase the number of different types considered and to generalize the relations between Parts (as
-in Table~\ref{tab:rcc358}).
-For the sake of readability, we use the term \emph{Region} both to refer to a
-mereological Part and to a topological Region.
-The choice of mereotopology is also correlated to the objective of defining a formal
-ontology, which we use to define the (formal) semantics of the terms (Parts) in
-Section~\ref{sec:problem}, and of the concepts of safety and security
-(whole). We aim at creating a meronomy instead of the taxonomies such as the one provided
-in\autocite{NIST2020NVD,MITRE2020CVE} or instead of the poorly justified 
-CVSS\autocite{Mell2007CVSS} scoring system.
-
-A mereotopology, as defined e.g. in\autocite{Rachavelpula2017mereotopology},
-is an ordered mathematical structure where the basic relation between Regions is the
-reflexive and symmetric\fixnote{mr}{I guess it must be monotonic as defined in\autocite{Rachavelpula2017mereotopology} but I don't find it consistently in other papers.} 
-\emph{Parthood} relation $\subseteq$. 
-\begin{definition}{\bf Parthood --}\label{def:parthood}
-Given any pair of mereotopological Regions $X$ and $Y$,
-	\begin{enumerate}[noitemsep]
-		\item Reflexivity: $\forall X.~ (X\subseteq X)$
-		\item Symmetry: $\forall X, Y.~ (X\subseteq Y \Rightarrow Y\subseteq X)$
-	\end{enumerate}
-\end{definition}
-
-As later defined in Definition~\ref{def:agent}, 
-the Parthood relation orders a universe of agents $\agentuniverse$ by defining
-the so called \emph{Connects with} (see in Table~\ref{tab:rcc358}) relation 
-between Regions. We want this universe $\agentuniverse$ to be 
-expressible in FOL. In this way, we
-can reason both on the constituent of security (i.e. its terms and agents
-defining a system where security needs to be considered), and on 
-the evolution of those constituent w.r.t. cause-effects relations according
-to the modal structure of causality we defined in Section~\ref{sec:causality}.
-This will allow us (in Section~\ref{sec:process} and Section~\ref{sec:prototest})
-a better\fixnote{mr}{Can we prove this by construction?} 
-positioning w.r.t. risk assessment technologies
-(which most often reason on the constituent of a system design), and
-protocol verification tools (which requires some formalization of a cause-effect relation, e.g.
-in Linear Temporal Logic).
-
-As we argued in Section~\ref{sec:problem}, we must correlate the definition of
-agent (i.e. Mankind and Nature) in Definition~\ref{def:mankind-nature} to the
-mathematical structure of the logic that defines them.  We express Mankind and
-Nature as formulas over the theory of mereology and then in terms of
-mereotopological Regions, extending the interpretation function
-$\interpretation$, to include a formal theory of mereotopology.  We use the
-Region Connection Calculus (RCC), as defined
-in~\cite{bennettLogics,improvingRCC}, to provide an axiomatization of the
-spatial concepts and relations in first-order logic to correlate the algebraic
-structure to mereology. In its broader definition, the RCC theory is composed
-by eight axioms, and is known as RCC8. In the text, for brevity, 
-we will often focus only on RCC5 (without
-loss of generality) by not considering tangential connections between spatial
-Regions. We discuss the choice of RCC5 in more detail in
-Appendix~\ref{app:rcc}. In Table~\ref{tab:rcc}, we
-summarize the axioms of the Region Connection Calculus (see, e.g., \autocite{Grutter2008rcc}).
-
-\begin{definition}{\bf RCC axiomatization --}
-	For any $X,Y$ pair of Regions in a mereotopology:
-	\begin{itemize}[noitemsep]
-	\item[$(\interpretation6)$] $\interpretation(X\subseteq Y)$ iff [$\interpretation(X\subseteq X)=\top$, and $\interpretation(X\subseteq Y)=\bot$ or $\interpretation(Y\subseteq X)=\top$]\fixnote{mr}{would it be better/clearer or just correct to write $\sigma(X\subseteq Y)$ iff $\sigma(X\subseteq X\wedge [X\subseteq Y \vee Y\subseteq X])=\top$}
-	\item[$(\interpretation7)$] $\world\models\connects{X}{Y}$ iff $\world\models X\subseteq Y$ 
-	\item[$(\interpretation8)$] $\world\models\disconnected{X}{Y}$ iff  $\world\not\models X\subseteq Y$
-	\item[$(\interpretation9)$] $\world\models\partof{X}{Y}$ iff $\world\models\forall Z.~\connects{Z}{X}\Rightarrow\connects{Z}{Y}$
-	\item[$(\interpretation10)$] $\world\models\overlaps{X}{Y}$ iff $\world\models\exists Z.~\partof{Z}{X}\wedge\partof{Z}{Y}$
-	\item[$(\interpretation11)$] $\world\models\eq{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\partof{Y}{X}$
-	\item[$(\interpretation12)$] $\world\models\dr{X}{Y}$ iff $\world\models\neg\overlaps{X}{Y}$
-	\item[$(\interpretation13)$] $\world\models\po{X}{Y}$ iff $\world\models\overlaps{X}{Y}\wedge\neg\partof{X}{Y}\wedge\neg\partof{Y}{X}$
-	\item[$(\interpretation14)$] $\world\models\pp{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\neg\partof{Y}{X}$
-	\item[$(\interpretation15)$] $\world\models\ppi{X}{Y}$ iff $\world\models\partof{Y}{X}\wedge\neg\partof{X}{Y}$
-\end{itemize}
-where $Z$ is a mereotopological Region.
-\end{definition}
-
-\begin{definition}{\bf Agent: Mankind or Nature --}\label{def:agent}
-	An agent $\agent\in\agentuniverse$ is a tuple
-	$\langle\rcc(\region',\region''),\ldots,\rcc(\region^{n-1},\region^n)\rangle$
-	of RCC relations $\rcc$ over mereotopological Regions
-	${\region',\ldots,\region^n}\subseteq\Region$. 
-\end{definition}
-
-Currently, we do not distinguish between Mankind and Nature (since we still
-lack of the definition of ``malicious intent'', which is defined in
-Section~\ref{sec:theory}) and we have defined them as two generic
-agents.
-
-As depicted in Figure~\ref{fig:safety-security}, Causality, Mankind, and Nature
-have a dashed border representing their correlation in terms of cause-effect
-and then in terms of formal structure which defines them: Modal Logic.
-Vulnerability, Defect, Incident, Accident, and Threat, similarly, 
-are correlated (depicted as a solid border) in terms of underlying 
-formal structure: Mereotopology.
-
-\subsubsection{Regions: Vulnerability and Defect (and Weakness)}\label{sec:vulnerabilitydefect}
-As defined in Section~\ref{sec:problem}, a Vulnerability or a Defect is a
-\emph{Weakness}.  As an example, a categorization of weaknesses is given
-in\autocite{MITRE2020CWEresearch} with 808 weaknesses categorized as ``Research
-Concepts'', distributed as follows:
-\begin{itemize}[noitemsep]
-	\item Incorrect Calculation - ($682$)
-	\item Incorrect Access of Indexable Resource (``Range Error'') - ($118$)
-	\item Use of Insufficiently Random Values - ($330$)
-	\item Improper Interaction Between Multiple Correctly-Behaving Entities - ($435$)
-	\item Improper Control of a Resource Through its Lifetime - ($664$)
-	\item Insufficient Control Flow Management - ($691$)
-	\item Protection Mechanism Failure - ($693$)
-	\item Incorrect Comparison - ($697$)
-	\item Improper Check or Handling of Exceptional Conditions - ($703$)
-	\item Improper Enforcement of Message or Data Structure - ($707$)
-	\item Improper Adherence to Coding Standards - ($710$)
-\end{itemize}
-
-The definition given by the MITRE in\autocite{MITRE2020CWEweakness} of
-weakness is: ``Software weaknesses are errors that can lead to software
-vulnerabilities. A software vulnerability, such as those enumerated on the
-Common Vulnerabilities and Exposures (CVE) List, is a mistake in software that
-can be directly used by a hacker to gain access to a system or network''.
-The definition is circular if we interpret the word ``error'' and
-``mistake'' with the same semantics: a weakness is an error that leads to a
-vulnerability and a vulnerability is a mistake which, in turn, is a weakness.
-The only difference (between weakness and vulnerability) seems to be
-that one can consider weakness as a ground term and state that a
-vulnerability is caused by a weakness, i.e. 
-$\World,\Weakness\models\Diamond\Vulnerability\wedge\Weakness$ where
-$\Weakness,\Vulnerability$ are Regions of Weaknesses and Vulnerabilities (resp.);
-accepting the hierarchy in the CWE\autocite{CWE} as ground truth. 
-Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability) 
-or the CVE reported in the NVD as a ground truth.
-
-\begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
-	A Region $\region\subseteq\agent$ of an agent
-	$\agent\in\agentuniverse$, is defined as Weakness $\Weakness$ (i.e.
-	representing a weakness introduced by the agent into any phase of
-	production, e.g.  of the secure process development life-cycle, of any
-	system or subsystem) iff there exists $\world\in\possibleworlds$ such
-	that $\world,\Weakness\models\Diamond\exists\Weakness',\Vulnerability'.~
-	\rcc(\Weakness',\Vulnerability')\wedge\neg\dr{\Weakness'}{\Vulnerability'}$, 
-	where $\Weakness'\subseteq\Weakness$, and
-	$\Vulnerability'\subseteq\Vulnerability$ is a Region of an agent $\agent'\in\agentuniverse$.
-\end{definition}
-
-\begin{example}{CWE-116: Improper Encoding or Escaping of Output\autocite{CWE-116} --}
-	\begin{itemize} 
-		\item Description: The software prepares a structured message
-		for communication with another component, but encoding or
-		escaping of the data is either missing or done incorrectly. As
-		a result, the intended structure of the message is not
-		preserved. 
-		\item Example: This code displays an email
-		address that was submitted as part of a form. Example language JSP.
-		\begin{verbatim} 
-		<% String email = request.getParameter("email"); %> 
-		...
-		Email Address: <%= email %>
-		\end{verbatim}
-		The value read from the form parameter is
-		reflected back to the client browser without
-		having been encoded prior to output, allowing
-		various XSS attacks (CWE-79).
-		\item Observed Examples
-			\begin{itemize}
-			\item CVE-2008-4636\autocite{CVE-2008-4636}: OS command injection in backup
-				software using shell metacharacters in a
-				filename; correct behavior would
-				require that this filename could not be
-				changed.
-			\item CVE-2008-0769\autocite{CVE-2008-0769}: Web application does not set the
-				charset when sending a page to a browser,
-				allowing for XSS exploitation when a
-				browser chooses an unexpected encoding.
-			\item CVE-2008-0005\autocite{CVE-2008-0005}: Program does not set the charset
-				when sending a page to a browser, allowing for
-				XSS exploitation when a browser chooses
-				an unexpected encoding.
-			\item CVE-2008-5573\autocite{CVE-2008-5573}: SQL injection via password
-				parameter; a strong password might contain \&
-			\item CVE-2008-3773\autocite{CVE-2008-3773}: Cross-site scripting in chat
-				application via a message subject, which
-				normally might contain \& and other
-				XSS-related characters.
-			\item CVE-2008-0757\autocite{CVE-2008-0757}: Cross-site scripting in chat
-				application via a message, which normally might
-				be allowed to contain arbitrary
-				content.
-			\end{itemize}
-\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\begin{subfigure}[b]{.5\textwidth}
-		\centering
-		\includegraphics[width=.7\textwidth]{rcc5_CWE-116.pdf}
-		\caption{Representation with RCC3}
-		\label{fig:rcc5_CWE-116}
-	\end{subfigure}\begin{subfigure}[b]{.5\textwidth}
-		\centering
-		\includegraphics[width=.7\textwidth]{rcc8_CWE-116.pdf}
-		\caption{Representation with RCC8}
-		\label{fig:rcc8_CWE-116}
-	\end{subfigure}
-	\begin{subfigure}[b]{\textwidth}
-		\centering
-		\includegraphics[width=\textwidth]{KML_CWE-116.pdf}
-		\caption{Representation in K-Modal Logic}
-		\label{fig:KML_CWE-116}
-	\end{subfigure}
-	\caption{Relations between CWE-116 and correlated CVEs}
-\end{figure}
-Definition~\ref{def:weakness} states that CWE-116 is a weakness iff there exists a world in
-K Modal Logic, representing the system in which this weakness exists, 
-such that the natural evolution of this system (i.e. formalized by the causality principle) 
-make it possible to reach another state of the system (i.e. another, accessible world)
-where there exist a vulnerability that can be implemented from CWE-116 (and this relation 
-is not DR). The CWE website proposes the connection between CWE-116 and, for example, 
-CVE-2008-5573; a vulnerability of the login sub-system 
-of the Poll Pro v2.0\autocite{pollpro} system.
-The formal relation between the two is given as a link (i.e. URI) between the CWE and
-the CVE, the description of the relation is not defined but it is supposed to be
-inferred from the descriptions of the CWE-CVE. 
-We can formally represent this link as the ONE connection in RCC3, depicted in 
-Figure~\ref{fig:rcc5_CWE-116}, or EC connection in RCC8 (Figure~\ref{fig:rcc8_CWE-116}).
-
-\fixnote{mr}{shouldn't we consider the agent who introduces the weakness as dishonest? don't we say this in part 1?}
-\end{example}
-
-It is interesting to note that the CWE-CVE relation expresses the correlation
-between weaknesses and vulnerabilities in the most simple form, such that we
-can formalize all the relation as ONE in RCC3 or EC in RCC8.  To express a more
-complex relation between the two we shall analyze the definition of CWE-116.
-This is related to the Weakness-Vulnerability-Incident process (i.e. to the
-details on the implementation/realization in Figure~\ref{fig:safety-security})
-that we analyze in the next section, Section~\ref{sec:incidentaccident}.
-
-\begin{definition}{\bf Region: Vulnerability and Defect --}\label{def:vulnerability-defect}
-A region $\region\subseteq\agent$ of an agent $\agent\in\agentuniverse$ is
-	called \emph{Vulnerability} if the agent $\agent$ is referred to as Mankind, 
-	\emph{Defect} if the agent is referred to as Nature.
-\end{definition}
-
-\subsubsection{Process: Incidents and Accidents}\label{sec:incidentaccident}
-In our informal definition depicted in Figure~\ref{fig:safety-security}, an
-Incident is generated through a process caused by a Vulnerability (which,
-in turn, is caused by a Weakness). Symmetrically, Nature has a process from Defect
-to Accidents. We start by analyzing the process generated by Mankind, which is
-divided into the following phases (explained in the remainder of this sub-Section):
-%Agent$\rightarrow$Vulnerability$\rightarrow$Exploit$\rightarrow$Attack$\rightarrow$Incident.
-\begin{enumerate}[noitemsep]
-	\item An agent (Mankind or Nature) defines a system, e.g.
-by following an engineering process for creating a new system, or for redesigning or improving an existing system (e.g.
-a component or a collection of components)
-	\item The dogmatic definition of a system contains a Weakness (i.e.
-the agent, Mankind or Nature, introduces an unintended behavior), for example as an un-intended behavior or as an architectural fallacy.
-The Weakness can be applied to the abstract definition of the system.
-	\item The Weakness can be defined as a malicious process for a system, called Vulnerability (or Defect for Nature)
-	\item The process expressed by the Vulnerability can be made operational for a specific implementation of the system (i.e. implemented by Mankind or realized by Nature) 
-\end{enumerate}
-Those additional details are depicted in Figure~\ref{fig:safety-security_2}, an update
-of Figure~\ref{fig:safety-security}.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{safety-security_2.pdf}
-	\caption{Overview security and safety keywords}
-	\label{fig:safety-security_2}
-\end{figure}
-
-\Paragraph{Exploitation Flow} An \emph{Exploit} is an \emph{implementation} of
-a Vulnerability, where a Vulnerability is one of the many ways to transform a
-Weakness into a malicious process that can be applied\footnote{Meaning that the
-malicious process, i.e.  any structured procedure such as a protocol, can be
-somehow made operational, e.g. implemented} to one or many systems. So, the act
-of an agent (Mankind) of ``exploiting a Vulnerability'' is the process of
-making the Vulnerability operational, by implementing the Vulnerability w.r.t.
-a specific target system.  Whenever the agent is Nature, the implementation is
-more broadly considered a realization\footnote{``The state of being
-realized''\autocite{Merriam2020realization}}
-
-The Common Attack Pattern Enumeration and
-Classification\autocite{MITRE2020CAPEC} (CAPEC) online service provides a
-``community resource'' which aims at ``identifying and understanding attacks''.
-The MITRE writes on the CAPEC homepage that ``Understanding how the adversary
-operates is essential to effective cyber security. CAPEC helps by providing a
-comprehensive dictionary of known patterns of attack employed by adversaries to
-exploit known weaknesses in cyber-enabled capabilities. It can be used by
-analysts, developers, testers, and educators to advance community understanding
-and enhance defenses.''. We can extrapolate two major concepts: 
-\begin{enumerate}
-	\item while understanding the exploitation process is the aim of the
-		CWE and CVE initiatives, understanding the hacking process is
-		the aim of the CAPEC initiative.
-	\item CAPEC provides a \emph{dictionary} of \emph{known} attack patterns.
-		As stated largely in Section~\ref{sec:}, this dictionary contains
-		empirical evidences and should not be used to induce a cybersecurity 
-		theory; but it can be used to validate a theory.
-\end{enumerate}
-
-The process continues as follows.
-\begin{enumerate}[noitemsep]
-\setcounter{enumi}{4}
-	\item an agent (Mankind\footnote{For the sake of readability we focus only on Mankind as an agent}) the exploitation flow can be used to hack a system.
-We use the term hack (as hacking) to stress that nothing prevents this process from being honest, however, for the sake of simplicity, we focus on Attacks which are malicious by definition.
-	\item An Attack is applied to a system with malicious intent by an agent
-	\item The application of the attack results in an Incident which impacts, i.e. poses a Threat to, an Asset
-\end{enumerate}
-
-\Paragraph{Hacking Flow}
-
-\begin{example}
-given in natural language in the description of the 
-CWE-116 which we can analyze as follows:
-\begin{itemize}
-	\item ``the software prepares a structured message''
-	\item the message is intended ``for communication with another component''
-	\item ``but''
-	\item ``encoding or escaping of the data is either missing or done incorrectly''
-	\item ``[\ldots] result, the intended structure of the message is not preserved''
-\end{itemize}
-\end{example}
-
-In this Vulnerability is seen as a process through which an agent (i.e. Mankind)
-generates an Incident, i.e.
 
@@
 \end{definition}
-
-\subsection{Vulnerability and Defect}\label{sec:vulnerability-defect}
-In order to formalize the semantics of the concepts (informally) described in
-Section~\ref{sec:problem},
-
-\begin{definition}{\bf Vulnerability}
-\end{definition}
-
-\subsection{Cybersecurity Engineering Process}\label{sec:process}
-%\epigraph{La differenza tra idee e realt\`a \`e la differenza tra
-%filosofia e ingegneria. Il lavoro per trasformare l'una nell'altra \`e ricerca
-%scientifica
-\epigraph{The difference between ideas and reality is the
-difference between philosophy and engineering. The work to transform one into
-the other is scientific research}{{\itshape V-Research}}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{vmodel.pdf}
-	\caption{Cybersecurity Engineering Life-cycle}
-	\label{fig:knowledge-belief}
-\end{figure}
-
-\subsubsection{Relevant Standards}\label{sec:standards}
-\begin{enumerate}[noitemsep]
-	\item DO-326A
-\end{enumerate}
-
-\section{Honesty and Maliciousness -- A Cybersecurity Theory}\label{sec:theory}
-\begin{enumerate}[noitemsep]
-	\item cybersecurity as a requirement
-	\item vulnerability\label{def:vulnerability}
-\end{enumerate}
-
-\section{Prototype and Empirical Tests}\label{sec:prototest}
-\subsection{Prototype Implementation}\label{sec:implementation}
-\subsection{Empirical Tests}\label{sec:tests}
-
-\section{Conclusion and Related Work}\label{sec:related}
-
-\printbibliography
-
-\begin{appendices}
-	\section{RCC3, RCC5, or RCC8}\label{app:rcc}
-	\section{Wrong idea 1}
-In order to apply consistently the mereotopological structure not just to
-reason the agents but also on their evolution, we extend the properties of the
-Parthood relation (reflexivity and symmetry) to the modal relation
-$\modalrelation$ of the K Modal logic, which in turns defines the causality
-principle. This updates the K Modal logic to a B (Brouwer) Modal Logic\autocite{Garson2018modal}.
-
-\begin{definition}{\bf B Modal Logic --}\label{thm:bk-parthood}
-	The B Modal Logic (defined, with a slight abuse of notation as
-	$\bframe=\langle\possibleworlds,\modalrelation\rangle$) is an extension
-	of the K modal logic with the following axioms
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation8)$] $\world\models P\Rightarrow Q$ iff
-			$\world\not\models P$ or $\world\models Q$ (logical implication)~\footnote{In
-			Definition~\ref{def:modallogic} we didn't specify the
-			semantics in the case of the connective $\Rightarrow$
-			(i.e. logical implication), so we state it here. Given
-			that $P\Rightarrow Q$ is equisatisfiable to
-			$\neg(P\wedge \neg Q)$, for any proposition $P$ and
-			$Q$, $\world\models P\Rightarrow Q$ is equisatisfiable
-			to $\world\models \neg(P\wedge \neg Q)$. If
-			$\world\models \neg(P\wedge \neg Q)$ then
-			$\world\not\models P\wedge\neg Q$ and then
-			$\world\models \neg P$ or $\world\models Q$.}
-		\item[$(\interpretation9)$] $\Box P\Rightarrow P$ (which ``claims that whatever is necessary is the case''\autocite{Garson2018modal}), and
-		\item[$(\interpretation10)$] $P\Rightarrow\Box\Diamond P$
-			(which says that if $P$ is the case, then $P$ is
-			necessarily possible\autocite{Garson2018modal})
-	\end{enumerate}
-\end{definition}
-
-\begin{theorem}{\bf Causality, Parthood, and B Modal Logic}\label{thm:bk-parthood}
-	\begin{enumerate}[noitemsep]
-		\item (Reflexivity) $\bframe\models\Box P\Rightarrow P$ ($\interpretation8$) iff $\modalrelation$ is reflexive
-		\item (Symmetry) $\bframe\models P\Rightarrow\Box\Diamond$ ($\interpretation9$) iff $\modalrelation$ is symmetric
-		\item (Consistency) $\world\modalrelation\world'$ iff $\connects{\region}{\region'}$
-	\end{enumerate}
-	whenever $\world$ and $\world'$, and $\region$ and $\region'$ are logically equivalent ($\equiv$).
-\end{theorem}
-\begin{proof}
-	The proof is exhaustive over all possible cases.
-	\begin{enumerate}[noitemsep]
-		\item(Reflexivity) proven in\autocite{}
-		\item(Symmetry) proven in\autocite{}
-		\item(Equivalence) \fix{mr}{todo} If
-			$\world\modalrelation\world'$ and
-			$\world\equiv\region$, then if
-			it doesn't hold that $\connects{\region,\region'}$ it
-			follows that there is no reflexive and transitive
-			relation that correlates $\region$ to $\region'$. Given 
-			that Regions are defined as signatures
-	\end{enumerate}
-\end{proof}
-
-\end{appendices}
-
-\end{document}

### 2020-02-20 introduction.tex

--- 2020-02-19 v-research_experiment-alpha.tex
@@
-\documentclass{article}
-\usepackage{mathtools}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{enumitem}
-\usepackage{multicol}
-\usepackage[toc,page]{appendix}
-\usepackage[gen]{eurosym}
-\usepackage{tabularx}
-\usepackage{epigraph}
-\usepackage{textcomp}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{csquotes}% Recommended
-\usepackage{rotating}
-\usepackage[style=numeric]{biblatex}
-% \bibliography{<mybibfile>}% ONLY selects .bib file; syntax for version <= 1.1b
-\addbibresource{literature.bib}% Syntax for version >= 1.2
-
-\usepackage{graphicx}
-\graphicspath{{./images/}}
-
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\lhead{}
-\chead{}
-\rhead{}
-\lfoot{}
-\cfoot{\thepage}
-\rfoot{Property of V-Research S.r.l.}
-\renewcommand\headrulewidth{0pt}
-\renewcommand\footrulewidth{0pt}
-
-\usepackage[colorinlistoftodos]{todonotes}
-
-\newcommand{\Paragraph}[1]{\smallskip\noindent{\bf #1.}}
-\newcommand{\fixnote}[2]{\textbf{\color{red}{FIX}}\footnote{{\bf #1:} #2}}
-\newcommand{\fix}[2]{{\color{red} {\bf tofix:} #2}}
-% \renewcommand{\fixnote}[2]{}
-% \renewcommand{\fix}[2]{}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\theoremstyle{corollary}
-\newtheorem{corollary}{Corollary}[section]
-\theoremstyle{lemma}
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{theorem}
-\newtheorem{theorem}{Theorem}[section]
-\theoremstyle{theorem}
-\newtheorem{example}{Example}
-
-\usepackage{authblk}
-\date{}                     %% if you don't need date to appear
-\setcounter{Maxaffil}{0}
-\renewcommand\Affilfont{\itshape\small}
-
-% for fancy table with rotated column titles
-\usepackage{adjustbox}
-\newcolumntype{R}[2]{%
-    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
-    l%
-    <{\egroup}%
-}
-\newcommand*\rot{\multicolumn{1}{R{45}{0.01em}}}% no optional argument here, please!
-\newcommand*\rota{\multicolumn{1}{R{90}{0.01em}}}% no optional argument here, please!
-\def\hb{\hbox to 10.7 cm{}}
-
-% circles
-\usepackage{wasysym}
-\newcommand{\Tdot}{$\CIRCLE$}
-\newcommand{\Thdot}{$\LEFTcircle$}
-\newcommand{\Twdot}{$\Circle$}
-
-% formulas
-\newcommand{\kframe}{\mathbf{K}}
-\newcommand{\bframe}{\mathbf{B_K}}
-\newcommand{\possibleworlds}{G}
-\newcommand{\modalrelation}{R}
-\newcommand{\actualworld}{\omega*}
-\newcommand{\world}{\omega}
-\newcommand{\World}{\Omega}
-\newcommand{\kmodel}{M}
-\newcommand{\interpretation}{\sigma}
-\newcommand{\knowledge}[1]{\mathbb{K}_{#1}}
-\newcommand{\knows}[2]{K_{#1}#2}
-\newcommand{\belief}[1]{\mathbb{B}_{#1}}
-\newcommand{\believe}[2]{B_{#1}#2}
-\newcommand{\region}{\chi}
-\newcommand{\Region}{\mathrm{X}}
-\newcommand{\agentuniverse}{\mathcal{A}}
-\newcommand{\agent}{a}
-\newcommand{\follang}{\mathcal{L}}
-\newcommand{\weakness}{w}
-\newcommand{\Weakness}{W}
-\newcommand{\vulnerability}{v}
-\newcommand{\Vulnerability}{V}
-%rcc5
-\newcommand{\rcc}{rcc}
-\newcommand{\connects}[2]{C(#1,#2)}
-\newcommand{\disconnected}[2]{\neg C(#1,#2)}
-\newcommand{\partof}[2]{P(#1,#2)}
-\newcommand{\overlaps}[2]{O(#1,#2)}
-\newcommand{\eq}[2]{EQ(#1,#2)}
-\newcommand{\pp}[2]{PP(#1,#2)}
-\newcommand{\po}[2]{PO(#1,#2)}
-\newcommand{\ppi}[2]{PPi(#1,#2)}
-\newcommand{\dr}[2]{DR(#1,#2)}
-
-\makeindex
-
-\begin{document}
-
-\title{On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}}
-\author[1]{Francesco Beltramini}
-\author[1]{Marco Rocchetto}
-\author[2]{Luca Vigan\`o}
-\affil[1]{V-Research, Verona, Italy}
-\affil[2]{King's College London, UK}
-
-\maketitle
-
-\begin{abstract}
-	The objective of this research is to develop of a theory that defines
-	(all and only) the possible insecurity and security configurations of
-	any abstract system. The theory is structured upon other theories that 
-	defines how a component of a system can be abstracted into an agent, defining how
-	agents can be formalized (both syntactically and semantically) to
-	describe an abstract system, such as a graph. Some of these theories
-	(e.g. used for the semantic
-	definition of the abstract system) are the epistemological definition
-	of knowledge, the Belief-Desire-Intent and the Assertion-Belief-Fact 
-	framework of reference, 
-	mereology, and topological structure. We argue that a mereology is the
-	most appropriate abstract underlying structure, do to its generality,
-	for defining the expressiveness of the system abstraction.  Furthermore, a
-	mereology allows us to define an ontology rather than a taxonomy.  We
-	also correlate different abstractions of the system to the TRL and the
-	engineering V-model.
-
-	We implemented a formal theory (of axioms) of a mereotopology, and of
-	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
-	uses the Z3 SMT solver. The results show that a single component (i.e.
-	agent) of an abstract system has a definite number of  different
-	insecurity configurations (e.g. 53 using RCC5 over a topological
-	structure) and only 1 secure (i.e. expected) configurations. The
-	configurations are reported as models satisfying the abstract system
-	semantics.
-
-	We considered the philosophical definition of truth behind our
-	approach, rejecting ``proof'' by induction from partial empirical evidences.
-	Our theory can be applied to system engineering and we show a concrete
-	application of our theory to the risk assessment of an ad-hoc system.
-	Finally, we provide a number of ideas to support the engineering of
-	secure systems (e.g. purely cyber or cyber-physical).
-\end{abstract}
-\newpage
-
-\section{Introduction}\label{sec:intro}
 \epigraph{Humanum est errare}{{\itshape Seneca the Elder}}
@@
 
-In this article, we give the first scientific theory (to the best of our knowledge)
-on security.
-
-\Paragraph{Structure} In Section~\ref{sec:problem} we define and formalize the problem statement. 
-In Section~\ref{sec:theory} we outline our security theory, and in Section~\ref{sec:prototest}
-we describe the implementation of the theory and some empirical tests of the theory.
-Finally, in Section~\ref{sec:related} we conclude the paper with an overview of the related work.
-
-\section{Problem Statement}\label{sec:problem}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.6\textwidth]{protocol-example.pdf}
-	\caption{Abstraction of an ad-hoc esemplificative protocol execution}
-	\label{fig:protocol-example}
-\end{figure}
-
-In\autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
-``an asymmetry in computer security'', which he defines as follows: ``Things
-can be declared insecure by observation, but not the reverse. There is no
-observation that allows us to declare an arbitrary system or technique
-secure''. Herley then uses this argument to show that ``claims that any measure
-is necessary for security are empirically unfalsifiable''. Given that, any
-theory which is not falsifiable by an empirical experiment is well
-known\footnote{``A theory which is not refutable by any conceivable event is
-nonscientific. Irrefutability is not a virtue of a theory (as people often
-think) but a vice.'' -- Karl Popper, Conjectures and
-Refutations\autocite{popper1962conjectures}} to be nonscientific (i.e.
-unfalsifiability is a fallacy of a theory), Herley concludes that there is no
-scientific theory on cybersecurity; which means that cybersecurity lays in the
-realm of pseudo-sciences\autocite{Herley2016usenixvideo}.  Herley, e.g.
-in\autocite{Herley2017justifying}, discusses the implications of a
-nonscientific approach to cybersecurity, and highlights the tremendous impact on
-all the scientific research and engineering of systems; leading often to
-terrorism and wars, and wasting of resources in useless protections or
-overspending.  While the criticism is investigated
-in\autocite{Herley2016unfalsifiability}, no solution is provided.  On the
-contrary, the goal of this work \emph{is} to lay the foundations of a
-scientific cybersecurity theory. Furthermore, in Section~\ref{sec:sicurezza},
-we consider the problem raised by Herley not confined to ``computer security''
-but to any abstract system (so that our theory may hold for any sound
-implementation such as networks, mechanical, cyber, or cyber-physical system,
-or even a single computer or a single device such as an hard-drive).  There is
-also an apparent inconsistency in\autocite{Herley2016unfalsifiability} that we
-seek to clarify before following (as we agree) the scientific path draw by
-Herley: cybersecurity is defined as an abstract property in many formal
-approaches to the investigation of the security of systems, and the security of the design of a
-formally verified protocol is indeed falsifiable.  For example, in the protocol
-verification community, security is often defined as a formalization of the
-high-level properties confidentiality, integrity, and availability. The problem
-in such approaches is not the definition of what cybersecurity is, but the use
-of theories (such as the Dolev-Yao attacker model\footnote{For the sake of
-simplicity, the Dolev-Yao attacker can be considered as an abstraction of an
-active attacker who controls the network but cannot break
-cryptography.}\autocite{Dolev1983security})
-that only applies to specific instances (often called scenarios) and
-abstraction of the protocol. This, in turn, creates a false sense of security
-since requires assumptions on the abstraction of the system of which security
-is verified. As an example, for the formal security verification of the system
-in Figure~\ref{fig:protocol-example}, a formalized scenario needs to be defined
-by a modeler who chooses (among others): (i) a scope of the formalization (e.g.
-excluding the server that distributes the public key is often done when
-verifying the security of authentication protocols), (ii) the number of
-sessions (even tough some approaches do reason on an infinite number of
-sessions such as\autocite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
-the peers (e.g.  in the ASLan++ language\autocite{Oheimb2010aslan++}), and (iv)
-the abstraction of the cryptographic primitives (e.g.  ProVerif vs
-CryptoVerif\autocite{Blanchet2017symbolic}).  Some of the choices will
-completely change the results of the formal verification of the system. For
-example, under the perfect cryptography assumption\footnote{As defined
-in\autocite{Rocchetto2016cpdy}: ``In the so called perfect cryptography
-assumption, the security encryption scheme is suppose to be perfect, without
-any exploitable flaw, and so the only way for the attacker to decrypt a message
-is by using the proper key. That assumption is widely accepted in the security
-protocol community, and most of the formal reasoning tools for the analysis of
-security protocols abstract away the mathematical and implementation details of
-the encryption
-scheme\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
-and assuming that no violation to any security property is done after message
-I); in Figure~\ref{fig:protocol-example}, the freedom of choosing the scope
-determines that the flaws related to the dishonest impersonation of the Server
-may or may not be considered in the verification process.  This choice has
-tremendous impact on the focus and findings of the verification of the security
-of the protocol.  While this may seem to turn upon minutiae and foreseeable,
-this highlights the false sense of security that may derive from a
-non-scientific theory of system security\footnote{``To the superficial
-observer, the analysis of these forms seems to turn upon minutiae. It does in
-fact deal with minutiae, but they are of the same order as those dealt with in
-microscopic anatomy.'' -- Karl Marx, Capital Volume 1, 1867}.
-
-\subsection{Sicurezza: Safety and Security}\label{sec:sicurezza}
-In most of the natural languages, and in Italian too, the concepts of safety
-and security are not syntactically differentiated and both terms (safety and
-security) are expressed by the same word, e.g. sicurezza in Italian.  A
-semantic distinction between safety and security is correlated to a
-belief\footnote{A belief has to be intended as a proposition which is supposed
-to be true by the majority of humans in our society without a scientific
-underlying theory but based on partial empirical evidences or inductive proofs.} that
-safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
-natural environment (e.g. natural events such as wearing of hardware
-components) while security deals with \emph{incidents} posed by mankind (e.g.
-attackers and bugs).  The fundamental difference between nature and mankind (and,
-in turn, between safety and cybersecurity) is believed to be on the different
-intents\footnote{``The belief–desire–intention software model (BDI) is a
-software model developed for programming intelligent
-agents.''\autocite{wiki-bdi}. In the BDI model, the intents represents the
-deliberative state of an agent which determines the choice of that agent on
-what to do.} (accidents are unfortunate while incidents are not) of the causes
-that generates the threat; namely, nature is believed not to have malicious
-intents (but unfortunate causes-effects) while threats generated by mankind are
-believed to be malicious\footnote{Of course, logical flaws or bugs may be
-introduced by other means (e.g. ignorance) without explicit malicious intents,
-but the exploitation of those flaws is considered (for now, and detailed
-afterwards in the article) malicious, and then we consider any vulnerability to
-be malicious (without loss of generality) even if due to the lack of skills.}.
-An overview on the aforementioned aspects of safety and security is depicted in
-Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
-the terms that structure our current understanding of safety and security. 
+Another way of looking at the same problem is by analyzing the different
+definitions of security with respect to the scientific method of enquiry
+applied to get to the definition itself. In the following we categorize the
+related work into three categories, as defined by Sextus Empiricus in
+\autocite{Empiricus1990Pyrrhonism}.  To the best of our knowledge, there is no
+evidence that this categorization isn't complete.  ``The natural result of any
+investigation is that the investigators either discover the object of search or
+deny that it is discoverable and confess it to be in-apprehensible or persist
+in their search. [\ldots] This is probably
+why''\autocite{Empiricus1990Pyrrhonism}: 
 \begin{itemize}
-	\item \emph{Mankind} ``refers collectively to humans''
-\autocite{wiki-mankind}, while the concept of \emph{Nature} is
-		related ``to the intrinsic characteristics that plants,
-		animals, and other features of the world develop of their own
-		accord'' (e.g. the physical universe)\autocite{wiki-nature}. 
+	\item The \emph{dogmatists} ``have claimed to have discovered the truth'' on what cybersecurity is
 		\begin{itemize}
-			\item So far, we have used several terms to refer to an
-				\emph{attacker}, i.e. threat agent or threat source,
-				considering those terms to be semantically
-				equivalent.  This ``shallowness'' raise form the
-				necessity of properly citing the different sources, but,
-				in the reminder of this paper, we consider the
-				Causality principle to be the \emph{threat
-				source}, Nature or Mankind to be the
-				\emph{threat agents} and an \emph{attacker} as
-				a specific malicious threat agent which materialize a
-				threat.
+			\item Wikipedia defines cybersecurity in
+				\autocite{wiki-cybersecurity} as the protection
+				of computer systems and networks from the theft
+				of or damage to their hardware, software, or
+				electronic data, as well as from the disruption
+				or misdirection of the services they provide
 		\end{itemize}
-	\item \emph{Vulnerability}\footnote{The term vulnerability is not
-		present in the Encyclopedia of Cryptography and Security, while
-		it is used in 12 entries (such as in the definition of
-		``penetration testing''\autocite{caddy2005pentest})
-		highlighting how commonly this word is used without a proper
-		supporting semantics}, as defined in\autocite{cnssi20104009}
-		(and adopted in\autocite{nist2013800-53}), is ``weakness in an
-		information system, system security procedures, internal
-		controls, or implementation that could be exploited by a threat
-		source''. On the one hand, the definition is broad to enclose
-		as much causes (that generates a vulnerability) as possible; on
-		the other hand, it derives from empirical evidences (which
-		should be considered beliefs\footnote{``For this view, that
-		\emph{That Which Is Not} exists, can never predominate. You
-		must debar your thought from this way of search, nor let
-		ordinary experience in its variety force you along this way,
-		(namely, that of allowing) the eye, sightless as it is, and the
-		ear, full of sound, and the tongue, to rule; but (you must)
-		judge by means of the Reason (Logos) the much-contested proof
-		which is expounded by me.'' -- Parmenides of Elea, On Nature
-		(circa 500 B.C.), fragments B7.1–8.2
-		\autocite{Hakim2016philosophy}} since they are partial results in nature) 
-		while a vulnerability should
-		be defined in a way that is empirically falsifiable. This means
-		that the term vulnerability should have a complete and sound
-		definition, so that no other causes (e.g.  other sources) but
-		the ones in the definition are responsible for a vulnerability.
-		Furthermore, the term ``threat sources'' used in the definition
-		in\autocite{cnssi20104009} may be identified with both Nature
-		and Mankind, not differentiating between safety and security.
-		In Definition~\ref{def:vulnerability}, we provide a formal
-		theory of vulnerability (so that the scientific community can
-		identify tests for the completeness and soundness of the
-		definition itself).
+	\item The \emph{academics} ``have asserted that it cannot be apprehended''
+		\begin{itemize}
+			\item Eugene H. Spafford, Professor at Purdue
+				University, defines cybersecurity as follow.
+				``The only truly secure system is one that is
+				powered off, cast in a block of concrete and
+				sealed in a lead-lined room with armed guards —
+				and even then I have my doubts.''
+				\autocite{Spafford2019Quotes}
 		\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{safety-security.pdf}
-	\caption{Overview security and safety keywords}
-	\label{fig:safety-security}
-\end{figure}
-
-Most of the safety-preserving principles in the field of engineering of
-safety-critical cyber-physical systems (such as elevators and aircraft), upon
-which safety requirements (e.g. in standards such as the IEC 61508 or 61511\autocite{IEC201761511}) are defined, have
-been defined following empirical tests and measurements. While reasoning by
-induction based on the empirical observation should be avoided, since it may
-easily lead to false beliefs instead of scientific theories, this approach is
-often justified by the supposed impossibility of defining a theory that
-correctly predicts failures which, in turn, pose hazards to a system. 
-To the best of our knowledge, and supported by\autocite{Herley2016unfalsifiability}, the
-correlation between predictability of environment and believed unpredictability
-of attackers (i.e. a malicious
-environment) has not been correlated to a theory on cybersecurity. 
-Therefore, inductive research efforts in
-predicting malicious effects are accepted (and published) in scientific
-conferences (e.g.~\autocite{Rocchetto2014CSRF}). A failure of a wire due to
-environment (e.g. due to humidity, dust, heat \&c) is defined from empirical
-evidences and processes have been standardized to test qualities of hardware components
-%convinced by the induction principle. What is already a
-%weak argument in general, 
-This process completely breaks down when a malicious environment (i.e. an attacker)
-is considered instead of the (supposedly honest and predictable)
-natural environment. Therefore, the same approach that is in use for safety,
-seems not to be applicable to test security.
-
-Going back to Figure~\ref{fig:safety-security}, a vulnerability does not
-necessarily become a threat for the system, unless exploited ``through a
-channel that allows the violation of the security policy
-[\ldots]''\autocite{cnssi20104009} (e.g. a software or procedure) that takes
-advantage of the vulnerability causing an \emph{attack} to the system, which may
-result in several correlated incidents and threats.  The process of
-exploitation of a defect as a vulnerability is reported in
-Figure~\ref{fig:safety-security} such that the difference between exploit and failure,
-and attack and accident is to be found just in the maliciousness of the intents
-that causes this process (i.e. excluding the intent, the terms are just syntactic transformation from a vulnerability to defect, from
-accident to incident). In the following, we conclude the informal definition of
-the terms that we used in this section and in Figure~\ref{fig:safety-security}.
-
+	\item The \emph{skeptics} ``go on inquiring''
 		\begin{itemize}
-	\item \emph{Causality} refers to the causality principle; defined
-		in\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
-		connection of phenomena through which one thing (the cause)
-		under certain conditions gives rise to, causes something else
-		(the effect). The essence of causality is the generation and
-		determination of one phenomenon by another. In this respect
-		causality differs from various other kinds of connection, for
-		example, the simple temporal sequence of phenomena, of the
-		regularities of accompanying processes''.
-	\item An \emph{Exploit}\footnote{We note that the term exploit is only
-		used as a verb in\autocite{ISO2009information}} is ``An exploit
-		(from the English verb to exploit, meaning to use something to
-		one’s own advantage) is a piece of software, a chunk of data,
-		or a sequence of commands that takes advantage of a bug or
-		vulnerability to cause unintended or unanticipated behavior to
-		occur on computer software, hardware, or something electronic
-		(usually computerized).''\autocite{wiki-exploit}.
-	\item An \emph{Attack}, as defined by the International Standard
-		ISO/IEC 27000 is an ``attempt to destroy, expose, alter,
-		disable, steal or gain unauthorized access to or make
-		unauthorized use of an asset''; where an \emph{Asset} is
-		``anything that has value to the organization''. We note that for
-		the purpose of this article, we do not want to focus on a specific
-		organization or business to define asset but, in general, on any 
-		abstract organization (e.g. a company or a society).
-		We do not consider ethical hackers as attacking a system. 
-		In fact, we consider the term \emph{hack} as
-		non-malicious (see Hacker\autocite{Stallman2002hacker}).
-	\item A \emph{Threat}, as defined in\autocite{cnssi20104009}, is ``Any
-		circumstance or event with the potential to adversely impact
-		organizational operations (including mission, functions, image,
-		or reputation), organizational assets, individuals, other
-		organizations, or the Nation through an information system via
-		unauthorized access, destruction, disclosure, modification of
-		information, and/or denial of service''.
-	\item \emph{Defect}, ``anything that renders the product not reasonably
-		safe''\autocite{Robinson2019writing} (i.e. a characteristic of
-		an object which hinders its proper usability).
-	\item \emph{Failure}, as defined in\autocite{Merriam2020failure} as ``a state of
-		inability to perform a normal function''. The term is
-		structured and detailed in
-		\autocite{cnssi20104009,iet2017glossary} but relying on an
-		abstract notion of failure without a specific definition.
-	\item \emph{Hazard}, ``a potential source of
-		harm''\autocite{iet2017glossary}.
-\end{itemize}
-
-\subsection{Glossary -- A Formalization}\label{sec:glossary}
-We now define a formalizations of the concepts described in
-Section~\ref{sec:problem} and depicted in Figure~\ref{fig:safety-security}. We
-base our formalization on first order logic (FOL) with the standard
-truth-value semantics. The choice of this logic is required by the 
-semantics of the concepts formalized afterwards in this section.
-
-We consider a first-order language $\follang$ over a
-signature $\Sigma_P$ where $P,P',\ldots,P^n$ represent terms, and
-$\varphi,\varphi',\ldots,\varphi^n$ represent formulas.
-The syntax is defined as follows.
-\begin{displaymath}
-	\varphi := P~|~\neg\varphi~|~\varphi\wedge\varphi~|~\varphi\vee\varphi~|~\varphi\Rightarrow\varphi'~|~\forall P.~\varphi~|~\exists P.~\varphi
-\end{displaymath}
-where $\wedge$, $\neg$, $\vee$, and $\Rightarrow$ are connectives representing
-conjunction, negation, disjunction, and (material) implication respectively; while
-$\forall$ and $\exists$ represent the standard universal and existential
-quantifiers (resp.). Finally, the symbol ``$.$'' is just syntactic sugar. 
-We consider $\interpretation\subset\Phi\times\{\top,\bot\}$ as the interpretation
-function, where $\Phi$ is any collection of sentences in $\follang$ 
-and $\top$ and $\bot$ represent the concepts of ``Tautology/True''
-and ``Contradiction/False'' respectively.
-
-Following Figure~\ref{fig:safety-security}, we start by formalizing the outermost term: Causality.
-
-\subsubsection{Causality Principle}\label{sec:causality}
-We formalize the Causality Principle starting from a K modal
-logic\autocite{Garson2018modal} (i.e. without restrictions on the causality
-relation between worlds). The standard definition of K modal logic is given in
-Definition~\ref{def:modallogic} in terms of an interpretation function (which
-we named $\sigma$ with a slight abuse of notation) defined in
-Definition~\ref{def:modalinterpretation}. 
-\begin{definition}{\bf K Modal Logic --}\label{def:modallogic}
-A K-frame is a frame $ \kframe=\langle \possibleworlds,\modalrelation \rangle $
-	in the K modal logic where $ \modalrelation $ is the binary relation
-	(i.e. a set of ordered pairs) between possible worlds
-	$\modalrelation\subseteq\possibleworlds\times\possibleworlds$, where
-	$\possibleworlds$ represents the possible worlds, and
-	$\possibleworlds\neq\emptyset$. An actual world
-	$\actualworld\in\possibleworlds$ is assumed. For any proposition $P$,
-	an interpretation function $\interpretation(\world,P)$ returns the
-	truth value of $P$; e.g. $\interpretation(\world,P)=\top$ means that
-	$P$ holds in $\world$. A model is defined as the tuple
-	$\kmodel=\langle\possibleworlds,\modalrelation,\interpretation\rangle$.
-\end{definition}
-
-\begin{definition}{\bf Causality as K Modal Interpretation Function --}\label{def:modalinterpretation}
-	Causality is (recursively) defined as the modal interpretation function $\interpretation$, as follows. 
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation0)$] if $\interpretation(\world,P)=\top$ then $\world\models P$
-		\item[$(\interpretation1)$] $\world\models\neg P$ iff $\world\not\models P$
-		\item[$(\interpretation2)$] $\world\models P \wedge Q$ iff $\world\models P$ and $\world\models Q$
-		\item[$(\interpretation3)$] $\world\models\Box P$ iff for any world $\world'\in\possibleworlds$ if $\world \modalrelation\world'$ then $\world'\models P$
-		\item[$(\interpretation4)$] $\world\models\Diamond P$ iff there exists a set of worlds $\World'\subset\possibleworlds$ such that for any $\world'\in\World'$, if $\world\modalrelation\world'$ then $\world'\models P$
-		\item[$(\interpretation5)$] $\models P$ iff $\actualworld\models P$
-	\end{enumerate}
-	where truth is defined as necessary with $\Box$ and possible with $\Diamond$.
-\end{definition}
-
-%\begin{definition}{\bf Causality --}\label{def:causality}
-%	The Causality principle is defined over a K Modal Logic (i.e. without
-%	restrictions on the causality relation between worlds) as in
-%	Definition~\ref{def:modallogic}, with the standard interpretation
-%	function defined in~\ref{def:modalinterpretation}
-%\end{definition}
-
-The causality principle has been defined in its
-generic form. In fact, the accessibility relation $\modalrelation$ is free from
-any axiomatic restriction (e.g. it's not reflexive nor anti-reflexive).  
-We will focus in Section~\ref{sec:process} (for our tests) on 
-the application of our theory to CPS system engineering.
-More detailed case studies (i.e. \fix{mr}{a CPS as a smart power-grid}) 
-will be defined in Section~\ref{sec:prototest}, where
-the accessibility relation defines in details how the system itself evolves, due to causal relation 
-(i.e. by restricting the causality principle only to those cause-effect that defines
-the system). Therefore, the definition of $\modalrelation$
-will be specialized in a more strict way based on the application domains and case study.
-
-We note that we have defined the
-Causality principle without considering it a \emph{threat source} since 
-we lack the concept of intent and maliciousness. Similarly, in the 
-next sections we won't discriminate between Nature and Mankind until we introduce 
-the concepts of maliciousness in Section~\ref{sec:theory} 
-and then formally define a threat agent
-in the same section.
-
-\subsubsection{Agents: Mankind and Nature}\label{sec:mankind-nature}
-\begin{table}[t]
-\centering
-\setlength{\tabcolsep}{3.5pt}
-\renewcommand{\arraystretch}{1}
-\scriptsize
-\caption{RCC3, RCC5, and RCC8 relations between Regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
-\begin{tabular}{ccclll} 
-\rota{\textbf{RCC3}}&\rota{\textbf{RCC5}}&\rota{\textbf{RCC8}}&\textbf{Name} & \textbf{Notation} & \textbf{Definition} \\
-\hline
-&&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& $\mathit{X}\subseteq \mathit{Y}$ \\
-&&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& $\mathit{X}\not\subseteq \mathit{Y}$\\
-&&&Part of				& $\mathit{P}(\mathit{X},\mathit{Y})$		& $\forall \mathit{Z} ~\mathit{C}(\mathit{Z},\mathit{X}) \rightarrow \mathit{C}(\mathit{Z},\mathit{Y})$\\
-&&&Overlaps			& $\mathit{O}(\mathit{X},\mathit{Y})$		& $\exists \mathit{Z} ~\mathit{P}(\mathit{Z},\mathit{X})\wedge \mathit{P}(\mathit{Z},\mathit{Y})$\\
-\Tdot&&&  \textbf{Overlaps Not Equal} 	& $\mathit{ONE}(\mathit{X},\mathit{Y})$		& $\mathit{O}(\mathit{X},\mathit{Y}) \land \neg \mathit{EQ}(\mathit{X},\mathit{Y})$ \\
-\Tdot&\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
-\Tdot&\Tdot&\Tdot& \textbf{DiscRete from} 		& $\mathit{DR}(\mathit{X},\mathit{Y})$		& $\neg \mathit{O}(\mathit{X},\mathit{Y})$\\
-&\Tdot&\Tdot&\textbf{Partial-Overlap}	& $\mathit{PO}(\mathit{X},\mathit{Y})$ 		& $\mathit{O}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{X},\mathit{Y}) \wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\
-&\Tdot&&\textbf{Proper-Part-of} 	& $\mathit{PP}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\ 
-	&\Tdot&&\textbf{Proper-Part-of-\textit{\textbf{i}}nverse} & $\mathit{PPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{Y},\mathit{X}) \wedge \neg \mathit{P}(\mathit{X},\mathit{Y})$\\
-&&\Tdot&\textbf{Externally Connected} 	& $\mathit{EC}(\mathit{X},\mathit{Y})$ 		& $\mathit{C}(\mathit{X},\mathit{Y}) \wedge \neg\mathit{O}(\mathit{X},\mathit{Y})$\\ 
-&&\Tdot&\textbf{Tangential PP} 	& $\mathit{TPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Tangential PPi} 	& $\mathit{TPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{TPP}(\mathit{Y},\mathit{X})$\\ 
-&&\Tdot&\textbf{Non-Tangential PP} 	& $\mathit{NTPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\neg\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Non-Tangential PPi} 	& $\mathit{NTPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{NTPP}(\mathit{Y},\mathit{X})$\\ 
-\end{tabular}
-\end{table}
-
-Mankind and Nature are defined in Section~\ref{sec:problem} as two abstract
-agents, both as collections (i.e. an abstract type that does not imply a
-specific implementation) of their sub-agents (i.e. humans for Mankind and
-plants, animals, \&c.  for Nature). Similarly to\autocite{Santaca2016abf}, 
-we define Mankind and Nature, and any
-other agent in the reminder of this article, as a meronomy (an hierarchy of
-Part-Whole relations) based on a standard definition of mereology, i.e. based
-on the definition of Parthood relation between \emph{Parts}.  
-However, shall 
-consider different types of Part; so, we extend the mereology to a
-mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
-to increase the number of different types considered and to generalize the relations between Parts (as
-in Table~\ref{tab:rcc358}).
-For the sake of readability, we use the term \emph{Region} both to refer to a
-mereological Part and to a topological Region.
-The choice of mereotopology is also correlated to the objective of defining a formal
-ontology, which we use to define the (formal) semantics of the terms (Parts) in
-Section~\ref{sec:problem}, and of the concepts of safety and security
-(whole). We aim at creating a meronomy instead of the taxonomies such as the one provided
-in\autocite{NIST2020NVD,MITRE2020CVE} or instead of the poorly justified 
-CVSS\autocite{Mell2007CVSS} scoring system.
-
-A mereotopology, as defined e.g. in\autocite{Rachavelpula2017mereotopology},
-is an ordered mathematical structure where the basic relation between Regions is the
-reflexive and symmetric\fixnote{mr}{I guess it must be monotonic as defined in\autocite{Rachavelpula2017mereotopology} but I don't find it consistently in other papers.} 
-\emph{Parthood} relation $\subseteq$. 
-\begin{definition}{\bf Parthood --}\label{def:parthood}
-Given any pair of mereotopological Regions $X$ and $Y$,
-	\begin{enumerate}[noitemsep]
-		\item Reflexivity: $\forall X.~ (X\subseteq X)$
-		\item Symmetry: $\forall X, Y.~ (X\subseteq Y \Rightarrow Y\subseteq X)$
-	\end{enumerate}
-\end{definition}
-
-As later defined in Definition~\ref{def:agent}, 
-the Parthood relation orders a universe of agents $\agentuniverse$ by defining
-the so called \emph{Connects with} (see in Table~\ref{tab:rcc358}) relation 
-between Regions. We want this universe $\agentuniverse$ to be 
-expressible in FOL. In this way, we
-can reason both on the constituent of security (i.e. its terms and agents
-defining a system where security needs to be considered), and on 
-the evolution of those constituent w.r.t. cause-effects relations according
-to the modal structure of causality we defined in Section~\ref{sec:causality}.
-This will allow us (in Section~\ref{sec:process} and Section~\ref{sec:prototest})
-a better\fixnote{mr}{Can we prove this by construction?} 
-positioning w.r.t. risk assessment technologies
-(which most often reason on the constituent of a system design), and
-protocol verification tools (which requires some formalization of a cause-effect relation, e.g.
-in Linear Temporal Logic).
-
-As we argued in Section~\ref{sec:problem}, we must correlate the definition of
-agent (i.e. Mankind and Nature) in Definition~\ref{def:mankind-nature} to the
-mathematical structure of the logic that defines them.  We express Mankind and
-Nature as formulas over the theory of mereology and then in terms of
-mereotopological Regions, extending the interpretation function
-$\interpretation$, to include a formal theory of mereotopology.  We use the
-Region Connection Calculus (RCC), as defined
-in~\cite{bennettLogics,improvingRCC}, to provide an axiomatization of the
-spatial concepts and relations in first-order logic to correlate the algebraic
-structure to mereology. In its broader definition, the RCC theory is composed
-by eight axioms, and is known as RCC8. In the text, for brevity, 
-we will often focus only on RCC5 (without
-loss of generality) by not considering tangential connections between spatial
-Regions. We discuss the choice of RCC5 in more detail in
-Appendix~\ref{app:rcc}. In Table~\ref{tab:rcc}, we
-summarize the axioms of the Region Connection Calculus (see, e.g., \autocite{Grutter2008rcc}).
-
-\begin{definition}{\bf RCC axiomatization --}
-	For any $X,Y$ pair of Regions in a mereotopology:
-	\begin{itemize}[noitemsep]
-	\item[$(\interpretation6)$] $\interpretation(X\subseteq Y)$ iff [$\interpretation(X\subseteq X)=\top$, and $\interpretation(X\subseteq Y)=\bot$ or $\interpretation(Y\subseteq X)=\top$]\fixnote{mr}{would it be better/clearer or just correct to write $\sigma(X\subseteq Y)$ iff $\sigma(X\subseteq X\wedge [X\subseteq Y \vee Y\subseteq X])=\top$}
-	\item[$(\interpretation7)$] $\world\models\connects{X}{Y}$ iff $\world\models X\subseteq Y$ 
-	\item[$(\interpretation8)$] $\world\models\disconnected{X}{Y}$ iff  $\world\not\models X\subseteq Y$
-	\item[$(\interpretation9)$] $\world\models\partof{X}{Y}$ iff $\world\models\forall Z.~\connects{Z}{X}\Rightarrow\connects{Z}{Y}$
-	\item[$(\interpretation10)$] $\world\models\overlaps{X}{Y}$ iff $\world\models\exists Z.~\partof{Z}{X}\wedge\partof{Z}{Y}$
-	\item[$(\interpretation11)$] $\world\models\eq{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\partof{Y}{X}$
-	\item[$(\interpretation12)$] $\world\models\dr{X}{Y}$ iff $\world\models\neg\overlaps{X}{Y}$
-	\item[$(\interpretation13)$] $\world\models\po{X}{Y}$ iff $\world\models\overlaps{X}{Y}\wedge\neg\partof{X}{Y}\wedge\neg\partof{Y}{X}$
-	\item[$(\interpretation14)$] $\world\models\pp{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\neg\partof{Y}{X}$
-	\item[$(\interpretation15)$] $\world\models\ppi{X}{Y}$ iff $\world\models\partof{Y}{X}\wedge\neg\partof{X}{Y}$
-\end{itemize}
-where $Z$ is a mereotopological Region.
-\end{definition}
-
-\begin{definition}{\bf Agent: Mankind or Nature --}\label{def:agent}
-	An agent $\agent\in\agentuniverse$ is a tuple
-	$\langle\rcc(\region',\region''),\ldots,\rcc(\region^{n-1},\region^n)\rangle$
-	of RCC relations $\rcc$ over mereotopological Regions
-	${\region',\ldots,\region^n}\subseteq\Region$. 
-\end{definition}
-
-Currently, we do not distinguish between Mankind and Nature (since we still
-lack of the definition of ``malicious intent'', which is defined in
-Section~\ref{sec:theory}) and we have defined them as two generic
-agents.
-
-As depicted in Figure~\ref{fig:safety-security}, Causality, Mankind, and Nature
-have a dashed border representing their correlation in terms of cause-effect
-and then in terms of formal structure which defines them: Modal Logic.
-Vulnerability, Defect, Incident, Accident, and Threat, similarly, 
-are correlated (depicted as a solid border) in terms of underlying 
-formal structure: Mereotopology.
-
-\subsubsection{Regions: Vulnerability and Defect (and Weakness)}\label{sec:vulnerabilitydefect}
-As defined in Section~\ref{sec:problem}, a Vulnerability or a Defect is a
-\emph{Weakness}.  As an example, a categorization of weaknesses is given
-in\autocite{MITRE2020CWEresearch} with 808 weaknesses categorized as ``Research
-Concepts'', distributed as follows:
-\begin{itemize}[noitemsep]
-	\item Incorrect Calculation - ($682$)
-	\item Incorrect Access of Indexable Resource (``Range Error'') - ($118$)
-	\item Use of Insufficiently Random Values - ($330$)
-	\item Improper Interaction Between Multiple Correctly-Behaving Entities - ($435$)
-	\item Improper Control of a Resource Through its Lifetime - ($664$)
-	\item Insufficient Control Flow Management - ($691$)
-	\item Protection Mechanism Failure - ($693$)
-	\item Incorrect Comparison - ($697$)
-	\item Improper Check or Handling of Exceptional Conditions - ($703$)
-	\item Improper Enforcement of Message or Data Structure - ($707$)
-	\item Improper Adherence to Coding Standards - ($710$)
-\end{itemize}
-
-The definition given by the MITRE in\autocite{MITRE2020CWEweakness} of
-weakness is: ``Software weaknesses are errors that can lead to software
-vulnerabilities. A software vulnerability, such as those enumerated on the
-Common Vulnerabilities and Exposures (CVE) List, is a mistake in software that
-can be directly used by a hacker to gain access to a system or network''.
-The definition is circular if we interpret the word ``error'' and
-``mistake'' with the same semantics: a weakness is an error that leads to a
-vulnerability and a vulnerability is a mistake which, in turn, is a weakness.
-The only difference (between weakness and vulnerability) seems to be
-that one can consider weakness as a ground term and state that a
-vulnerability is caused by a weakness, i.e. 
-$\World,\Weakness\models\Diamond\Vulnerability\wedge\Weakness$ where
-$\Weakness,\Vulnerability$ are Regions of Weaknesses and Vulnerabilities (resp.);
-accepting the hierarchy in the CWE\autocite{CWE} as ground truth. 
-Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability) 
-or the CVE reported in the NVD as a ground truth.
-
-\begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
-	A Region $\region\subseteq\agent$ of an agent
-	$\agent\in\agentuniverse$, is defined as Weakness $\Weakness$ (i.e.
-	representing a weakness introduced by the agent into any phase of
-	production, e.g.  of the secure process development life-cycle, of any
-	system or subsystem) iff there exists $\world\in\possibleworlds$ such
-	that $\world,\Weakness\models\Diamond\exists\Weakness',\Vulnerability'.~
-	\rcc(\Weakness',\Vulnerability')\wedge\neg\dr{\Weakness'}{\Vulnerability'}$, 
-	where $\Weakness'\subseteq\Weakness$, and
-	$\Vulnerability'\subseteq\Vulnerability$ is a Region of an agent $\agent'\in\agentuniverse$.
-\end{definition}
-
-\begin{example}{CWE-116: Improper Encoding or Escaping of Output\autocite{CWE-116} --}
-	\begin{itemize} 
-		\item Description: The software prepares a structured message
-		for communication with another component, but encoding or
-		escaping of the data is either missing or done incorrectly. As
-		a result, the intended structure of the message is not
-		preserved. 
-		\item Example: This code displays an email
-		address that was submitted as part of a form. Example language JSP.
-		\begin{verbatim} 
-		<% String email = request.getParameter("email"); %> 
-		...
-		Email Address: <%= email %>
-		\end{verbatim}
-		The value read from the form parameter is
-		reflected back to the client browser without
-		having been encoded prior to output, allowing
-		various XSS attacks (CWE-79).
-		\item Observed Examples
-			\begin{itemize}
-			\item CVE-2008-4636\autocite{CVE-2008-4636}: OS command injection in backup
-				software using shell metacharacters in a
-				filename; correct behavior would
-				require that this filename could not be
-				changed.
-			\item CVE-2008-0769\autocite{CVE-2008-0769}: Web application does not set the
-				charset when sending a page to a browser,
-				allowing for XSS exploitation when a
-				browser chooses an unexpected encoding.
-			\item CVE-2008-0005\autocite{CVE-2008-0005}: Program does not set the charset
-				when sending a page to a browser, allowing for
-				XSS exploitation when a browser chooses
-				an unexpected encoding.
-			\item CVE-2008-5573\autocite{CVE-2008-5573}: SQL injection via password
-				parameter; a strong password might contain \&
-			\item CVE-2008-3773\autocite{CVE-2008-3773}: Cross-site scripting in chat
-				application via a message subject, which
-				normally might contain \& and other
-				XSS-related characters.
-			\item CVE-2008-0757\autocite{CVE-2008-0757}: Cross-site scripting in chat
-				application via a message, which normally might
-				be allowed to contain arbitrary
-				content.
+			\item Cormac Herley reaches the conclusion that
+				cybersecurity has no definition. ``There is an
+				inherent asymmetry in computer security: things
+				can be declared insecure by observation, but
+				not the reverse. There is no test that allows
+				us to declare an arbitrary system or technique
+				secure. This implies that claims of necessary
+				conditions for security are unfalsifiable. ''
+				\autocite{Herley2016unfalsifiability}
 		\end{itemize}
@@
 
-\begin{figure}[t]
-	\centering
-	\begin{subfigure}[b]{.5\textwidth}
-		\centering
-		\includegraphics[width=.7\textwidth]{rcc5_CWE-116.pdf}
-		\caption{Representation with RCC3}
-		\label{fig:rcc5_CWE-116}
-	\end{subfigure}\begin{subfigure}[b]{.5\textwidth}
-		\centering
-		\includegraphics[width=.7\textwidth]{rcc8_CWE-116.pdf}
-		\caption{Representation with RCC8}
-		\label{fig:rcc8_CWE-116}
-	\end{subfigure}
-	\begin{subfigure}[b]{\textwidth}
-		\centering
-		\includegraphics[width=\textwidth]{KML_CWE-116.pdf}
-		\caption{Representation in K-Modal Logic}
-		\label{fig:KML_CWE-116}
-	\end{subfigure}
-	\caption{Relations between CWE-116 and correlated CVEs}
-\end{figure}
-Definition~\ref{def:weakness} states that CWE-116 is a weakness iff there exists a world in
-K Modal Logic, representing the system in which this weakness exists, 
-such that the natural evolution of this system (i.e. formalized by the causality principle) 
-make it possible to reach another state of the system (i.e. another, accessible world)
-where there exist a vulnerability that can be implemented from CWE-116 (and this relation 
-is not DR). The CWE website proposes the connection between CWE-116 and, for example, 
-CVE-2008-5573; a vulnerability of the login sub-system 
-of the Poll Pro v2.0\autocite{pollpro} system.
-The formal relation between the two is given as a link (i.e. URI) between the CWE and
-the CVE, the description of the relation is not defined but it is supposed to be
-inferred from the descriptions of the CWE-CVE. 
-We can formally represent this link as the ONE connection in RCC3, depicted in 
-Figure~\ref{fig:rcc5_CWE-116}, or EC connection in RCC8 (Figure~\ref{fig:rcc8_CWE-116}).
+In this article, we give the first scientific theory (to the best of our
+knowledge) on security.
 
-\fixnote{mr}{shouldn't we consider the agent who introduces the weakness as dishonest? don't we say this in part 1?}
-\end{example}
-
-It is interesting to note that the CWE-CVE relation expresses the correlation
-between weaknesses and vulnerabilities in the most simple form, such that we
-can formalize all the relation as ONE in RCC3 or EC in RCC8.  To express a more
-complex relation between the two we shall analyze the definition of CWE-116.
-This is related to the Weakness-Vulnerability-Incident process (i.e. to the
-details on the implementation/realization in Figure~\ref{fig:safety-security})
-that we analyze in the next section, Section~\ref{sec:incidentaccident}.
-
-\begin{definition}{\bf Region: Vulnerability and Defect --}\label{def:vulnerability-defect}
-A region $\region\subseteq\agent$ of an agent $\agent\in\agentuniverse$ is
-	called \emph{Vulnerability} if the agent $\agent$ is referred to as Mankind, 
-	\emph{Defect} if the agent is referred to as Nature.
-\end{definition}
-
-\subsubsection{Process: Incidents and Accidents}\label{sec:incidentaccident}
-In our informal definition depicted in Figure~\ref{fig:safety-security}, an
-Incident is generated through a process caused by a Vulnerability (which,
-in turn, is caused by a Weakness). Symmetrically, Nature has a process from Defect
-to Accidents. We start by analyzing the process generated by Mankind, which is
-divided into the following phases (explained in the remainder of this sub-Section):
-%Agent$\rightarrow$Vulnerability$\rightarrow$Exploit$\rightarrow$Attack$\rightarrow$Incident.
-\begin{enumerate}[noitemsep]
-	\item An agent (Mankind or Nature) defines a system, e.g.
-by following an engineering process for creating a new system, or for redesigning or improving an existing system (e.g.
-a component or a collection of components)
-	\item The dogmatic definition of a system contains a Weakness (i.e.
-the agent, Mankind or Nature, introduces an unintended behavior), for example as an un-intended behavior or as an architectural fallacy.
-The Weakness can be applied to the abstract definition of the system.
-	\item The Weakness can be defined as a malicious process for a system, called Vulnerability (or Defect for Nature)
-	\item The process expressed by the Vulnerability can be made operational for a specific implementation of the system (i.e. implemented by Mankind or realized by Nature) 
-\end{enumerate}
-Those additional details are depicted in Figure~\ref{fig:safety-security_2}, an update
-of Figure~\ref{fig:safety-security}.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{safety-security_2.pdf}
-	\caption{Overview security and safety keywords}
-	\label{fig:safety-security_2}
-\end{figure}
-
-\Paragraph{Exploitation Flow} An \emph{Exploit} is an \emph{implementation} of
-a Vulnerability, where a Vulnerability is one of the many ways to transform a
-Weakness into a malicious process that can be applied\footnote{Meaning that the
-malicious process, i.e.  any structured procedure such as a protocol, can be
-somehow made operational, e.g. implemented} to one or many systems. So, the act
-of an agent (Mankind) of ``exploiting a Vulnerability'' is the process of
-making the Vulnerability operational, by implementing the Vulnerability w.r.t.
-a specific target system.  Whenever the agent is Nature, the implementation is
-more broadly considered a realization\footnote{``The state of being
-realized''\autocite{Merriam2020realization}}
-
-The Common Attack Pattern Enumeration and
-Classification\autocite{MITRE2020CAPEC} (CAPEC) online service provides a
-``community resource'' which aims at ``identifying and understanding attacks''.
-The MITRE writes on the CAPEC homepage that ``Understanding how the adversary
-operates is essential to effective cyber security. CAPEC helps by providing a
-comprehensive dictionary of known patterns of attack employed by adversaries to
-exploit known weaknesses in cyber-enabled capabilities. It can be used by
-analysts, developers, testers, and educators to advance community understanding
-and enhance defenses.''. We can extrapolate two major concepts: 
-\begin{enumerate}
-	\item while understanding the exploitation process is the aim of the
-		CWE and CVE initiatives, understanding the hacking process is
-		the aim of the CAPEC initiative.
-	\item CAPEC provides a \emph{dictionary} of \emph{known} attack patterns.
-		As stated largely in Section~\ref{sec:}, this dictionary contains
-		empirical evidences and should not be used to induce a cybersecurity 
-		theory; but it can be used to validate a theory.
-\end{enumerate}
-
-The process continues as follows.
-\begin{enumerate}[noitemsep]
-\setcounter{enumi}{4}
-	\item an agent (Mankind\footnote{For the sake of readability we focus only on Mankind as an agent}) the exploitation flow can be used to hack a system.
-We use the term hack (as hacking) to stress that nothing prevents this process from being honest, however, for the sake of simplicity, we focus on Attacks which are malicious by definition.
-	\item An Attack is applied to a system with malicious intent by an agent
-	\item The application of the attack results in an Incident which impacts, i.e. poses a Threat to, an Asset
-\end{enumerate}
-
-\Paragraph{Hacking Flow}
-
-\begin{example}
-given in natural language in the description of the 
-CWE-116 which we can analyze as follows:
-\begin{itemize}
-	\item ``the software prepares a structured message''
-	\item the message is intended ``for communication with another component''
-	\item ``but''
-	\item ``encoding or escaping of the data is either missing or done incorrectly''
-	\item ``[\ldots] result, the intended structure of the message is not preserved''
-\end{itemize}
-\end{example}
-
-In this Vulnerability is seen as a process through which an agent (i.e. Mankind)
-generates an Incident, i.e.
-
-We describe processes, in our current formalization, in terms of the K-Modal
-Structure as in Definition~\ref{def:modalinterpretation} (i.e. in terms of
-causality principle).
-
-\begin{definition}{\bf Vulnerability-Incident Process --}\label{def:vulnerability-incident}
-	Given two 
-\end{definition}
-
-The whole-part relation \fix{mr}{(i.e.  HAS-A\footnote{HAS-A is just an hint for the
-mechanization of this theory.  We'll come back to this in
-Section~\ref{sec:implementation}})} is given, in the following, over the K Modal
-logic described beforehand. To do so, we first need to formally define the
-concept of \emph{agent} in its abstract form. We build it following the same
-structure of\autocite{Santaca2016abf} but over the terms used in
-Section~\ref{sec:problem}, i.e. knowledge, and belief as epistemic concepts. We'll then 
-introduce the concept of \emph{intent} to define a \emph{threat agent}.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.8\textwidth]{knowledge-belief.pdf}
-	\caption{Informal representation of Knowledge and Belief}
-	\label{fig:knowledge-belief}
-\end{figure}
-
-\subsubsection{Region: Threat and Asset}\label{sec:threatasset}
-The difference between Knowledge and Belief is depicted in
-Figure~\ref{fig:knowledge-belief} (see \autocite{wiki-knowledgebelief}).  However,
-according to\autocite{Gettier2012knowledge},
-Knowledge\footnote{``\emph{Theaetetus}: [\ldots] He said that knowledge was
-true opinion accompanied by reason, but that unreasoning true opinion was
-outside of the sphere of knowledge; and matters of which there is not a
-rational explanation are unknowable -- yes, that is what he called them -- and
-those of which there is are knowable. [\ldots] \emph{Socrates}: [\ldots] the
-primary elements of which we and all else are composed admit of no rational
-explanation; for each alone by itself can only be named, and no qualification
-can be added, neither that it is nor that it is not, for that would at once be
-adding to it existence or non-existence, whereas we must add nothing to it, if
-we are to speak of that itself alone.  [\ldots]'' Plato -- Theaetetus 201
-\autocite{Plato1914Plato}} as an epistemological concept is difficult (and
-sometimes believed impossible\autocite{citation}) to formally define. In this
-work, we are not interested in 
-how knowledge can be precisely formalized from an epistemic standpoint. 
-We assume that a semantic of a correct definition of epistemic knowledge exists,
-for example the one given in\autocite{Hintikka1962knwoledge} by Hintikka, and we then define 
-knowledge in terms of the Kripke structure defined in Definition~\ref{def:causality}.
-
-\begin{definition}{\bf Knowledge --}\label{def:knowledge}
-Given an abstract collection of Agents $Ag$, and the modal operator
-	$\knows{a}{}$ (where $a\in Ag$), Knowledge is defined as a collection
-	of predicates known by an agent $\knowledge{a}=\bigcup_\Phi \knows{a}{\varphi}$ 
-	(where $\Phi$ is the collection of all the propositions known by $a$).
-	Given a proposition $P$, we extend the semantics of the Causality structure with:
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation6)$] $\world\models\knows{a}{P}$ iff
-			$\world'\models P$ for all $\world'$ such that
-			$\world\modalrelation\world'$.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Belief --}\label{def:belief}
-	\fix{mr}{we should help the reader in understanding the upgrades of the logic.
-We used propositional with modal operators with a K Kripke semantics and we are now introducing Belief as in Doxastic logic?
-and we have epistemological operators as knowledge}
-	Given an abstract collection of Agents $Ag$, and the modal operator
-	$\believe{a}{}$ (where $a\in Ag$), Belief is defined as a collection
-	of predicates believed by an agent $\belief{a}=\bigcup_\Phi \believe{a}{\varphi}$
-	(where $\Phi$ is the collection of all the propositions believed by $a$).
-	Given a proposition $P$, we extend the semantics of the Causality structure with:
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation7)$] $\world\models\believe{a}{P}$ iff
-			$\world\models\neg\knows{a}{\neg P}$ (i.e. the agent $a$ considers $P$ possible) 
-			and $\world'\models P$ for all
-			$\world'$ such that $\world\modalrelation\world'$.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Mankind/Nature --}\label{def:mankind-nature}
-Mankind/Nature is represented as an agent.  
-	%countable logical signature $\Sigma_M$/$\Sigma_N$ over an
-	%abstract algebraic structure.  We also consider a countable signature
-	%of variables, disjoint from $\Sigma_M$/$\Sigma_N$, and write
-	%$\Sigma_{M_n}$/$\Sigma_{N_n}$ for the symbols of $\Sigma_M$/$\Sigma_N$
-	%with arity $n$. Thus, $\Sigma_0$ is a collection of constants, which we
-	%assume to have disjoint sub-collections that we refer to as
-	%\emph{sub-agents} of Mankind or Nature (i.e. humans, plants, \&c).
-	%The variables are untyped and can be instantiated with arbitrary types,
-	%yielding an untyped model.
-\end{definition}
-
-\subsubsection{Mereotopological Representation of Safety and Security}\label{sec:formalGlossary}
-The objectives of this section are:
-\begin{enumerate}[noitemsep]
-	\item formalize the abstract algebraic structure (as a mereotopology)
-		on top of which we have formalized the terms in
-		Figure~\ref{fig:safety-security} so far, i.e. Mankind, Nature,
-		Truths, Beliefs, and Knowledge.
-	\item correlate mereological structure to Causality and then Kripke
-		structure by defining the relation $R$ over the Kripke
-		structure as the mereotopological basic relation connects with,
-		i.e. reflexive and symmetric
-	\item we map signature to mereotopological Regions to express the
-		logical formulation of terms (e.g. Mankind, Nature, Truths,
-		Beliefs, Knowledge, Vulnerabilities) over the mereotopological
-		structure
-	\item we formalize the relations between terms using the RCC(5)
-\end{enumerate}
-
-\fix{mr}{change propositions in propositional language\url{http://logic.stanford.edu/intrologic/lectures/lecture_02.pdf}}
-
-Given that the concepts of Truths, Beliefs, and Knowledge are positioned over
-an underlying structure of Propositions, the first concept to formalize is that
-of Propositions. We do not restrict Propositions to a specific structure but,
-as for Mankind and Nature, we refer to them as a collection.  We express the
-generality of a collection as a (propositional) signature over an abstract
-algebraic structure. In our formalization, we use mereotopology as algebraic
-structure (due to its generality, as discussed in Section~\ref{sec:problem})
-and the RCC to formalize concepts over a mereotopology.
-
-We can now define the possible mereotopological relations between Belief and
-Knowledge (informally depicted in Figure~\ref{fig:knowledge-belief}) and then
-define the peculiarity of the believed difference between Mankind and Nature 
-(i.e. maliciousness as intent). 
-We first start from the formalization of Figure~\ref{fig:knowledge-belief}, as
-follows.
-
-\begin{definition}{\bf Information}\label{def:information}
-\end{definition}
-\begin{definition}{\bf Desire}\label{def:desire}
-\end{definition}
-\begin{definition}{\bf Intent}\label{def:intent}
-\end{definition}
-
-\subsection{Vulnerability and Defect}\label{sec:vulnerability-defect}
-In order to formalize the semantics of the concepts (informally) described in
-Section~\ref{sec:problem},
-
-\begin{definition}{\bf Vulnerability}
-\end{definition}
-
-\subsection{Cybersecurity Engineering Process}\label{sec:process}
-%\epigraph{La differenza tra idee e realt\`a \`e la differenza tra
-%filosofia e ingegneria. Il lavoro per trasformare l'una nell'altra \`e ricerca
-%scientifica
-\epigraph{The difference between ideas and reality is the
-difference between philosophy and engineering. The work to transform one into
-the other is scientific research}{{\itshape V-Research}}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{vmodel.pdf}
-	\caption{Cybersecurity Engineering Life-cycle}
-	\label{fig:knowledge-belief}
-\end{figure}
-
-\subsubsection{Relevant Standards}\label{sec:standards}
-\begin{enumerate}[noitemsep]
-	\item DO-326A
-\end{enumerate}
-
-\section{Honesty and Maliciousness -- A Cybersecurity Theory}\label{sec:theory}
-\begin{enumerate}[noitemsep]
-	\item cybersecurity as a requirement
-	\item vulnerability\label{def:vulnerability}
-\end{enumerate}
-
-\section{Prototype and Empirical Tests}\label{sec:prototest}
-\subsection{Prototype Implementation}\label{sec:implementation}
-\subsection{Empirical Tests}\label{sec:tests}
-
-\section{Conclusion and Related Work}\label{sec:related}
-
-\printbibliography
-
-\begin{appendices}
-	\section{RCC3, RCC5, or RCC8}\label{app:rcc}
-	\section{Wrong idea 1}
-In order to apply consistently the mereotopological structure not just to
-reason the agents but also on their evolution, we extend the properties of the
-Parthood relation (reflexivity and symmetry) to the modal relation
-$\modalrelation$ of the K Modal logic, which in turns defines the causality
-principle. This updates the K Modal logic to a B (Brouwer) Modal Logic\autocite{Garson2018modal}.
-
-\begin{definition}{\bf B Modal Logic --}\label{thm:bk-parthood}
-	The B Modal Logic (defined, with a slight abuse of notation as
-	$\bframe=\langle\possibleworlds,\modalrelation\rangle$) is an extension
-	of the K modal logic with the following axioms
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation8)$] $\world\models P\Rightarrow Q$ iff
-			$\world\not\models P$ or $\world\models Q$ (logical implication)~\footnote{In
-			Definition~\ref{def:modallogic} we didn't specify the
-			semantics in the case of the connective $\Rightarrow$
-			(i.e. logical implication), so we state it here. Given
-			that $P\Rightarrow Q$ is equisatisfiable to
-			$\neg(P\wedge \neg Q)$, for any proposition $P$ and
-			$Q$, $\world\models P\Rightarrow Q$ is equisatisfiable
-			to $\world\models \neg(P\wedge \neg Q)$. If
-			$\world\models \neg(P\wedge \neg Q)$ then
-			$\world\not\models P\wedge\neg Q$ and then
-			$\world\models \neg P$ or $\world\models Q$.}
-		\item[$(\interpretation9)$] $\Box P\Rightarrow P$ (which ``claims that whatever is necessary is the case''\autocite{Garson2018modal}), and
-		\item[$(\interpretation10)$] $P\Rightarrow\Box\Diamond P$
-			(which says that if $P$ is the case, then $P$ is
-			necessarily possible\autocite{Garson2018modal})
-	\end{enumerate}
-\end{definition}
-
-\begin{theorem}{\bf Causality, Parthood, and B Modal Logic}\label{thm:bk-parthood}
-	\begin{enumerate}[noitemsep]
-		\item (Reflexivity) $\bframe\models\Box P\Rightarrow P$ ($\interpretation8$) iff $\modalrelation$ is reflexive
-		\item (Symmetry) $\bframe\models P\Rightarrow\Box\Diamond$ ($\interpretation9$) iff $\modalrelation$ is symmetric
-		\item (Consistency) $\world\modalrelation\world'$ iff $\connects{\region}{\region'}$
-	\end{enumerate}
-	whenever $\world$ and $\world'$, and $\region$ and $\region'$ are logically equivalent ($\equiv$).
-\end{theorem}
-\begin{proof}
-	The proof is exhaustive over all possible cases.
-	\begin{enumerate}[noitemsep]
-		\item(Reflexivity) proven in\autocite{}
-		\item(Symmetry) proven in\autocite{}
-		\item(Equivalence) \fix{mr}{todo} If
-			$\world\modalrelation\world'$ and
-			$\world\equiv\region$, then if
-			it doesn't hold that $\connects{\region,\region'}$ it
-			follows that there is no reflexive and transitive
-			relation that correlates $\region$ to $\region'$. Given 
-			that Regions are defined as signatures
-	\end{enumerate}
-\end{proof}
-
-\end{appendices}
-
-\end{document}
+\Paragraph{Structure} In Section~\ref{sec:problem} we define and formalize the
+problem statement.  In Section~\ref{sec:theory} we outline our security theory,
+and in Section~\ref{sec:prototest} we describe the implementation of the theory
+and some empirical tests of the theory.  Finally, in Section~\ref{sec:related}
+we conclude the paper with an overview of the related work.

### 2020-02-20 problem_statement.tex

--- 2020-02-19 v-research_experiment-alpha.tex
@@
-\documentclass{article}
-\usepackage{mathtools}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{enumitem}
-\usepackage{multicol}
-\usepackage[toc,page]{appendix}
-\usepackage[gen]{eurosym}
-\usepackage{tabularx}
-\usepackage{epigraph}
-\usepackage{textcomp}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{csquotes}% Recommended
-\usepackage{rotating}
-\usepackage[style=numeric]{biblatex}
-% \bibliography{<mybibfile>}% ONLY selects .bib file; syntax for version <= 1.1b
-\addbibresource{literature.bib}% Syntax for version >= 1.2
-
-\usepackage{graphicx}
-\graphicspath{{./images/}}
-
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\lhead{}
-\chead{}
-\rhead{}
-\lfoot{}
-\cfoot{\thepage}
-\rfoot{Property of V-Research S.r.l.}
-\renewcommand\headrulewidth{0pt}
-\renewcommand\footrulewidth{0pt}
-
-\usepackage[colorinlistoftodos]{todonotes}
-
-\newcommand{\Paragraph}[1]{\smallskip\noindent{\bf #1.}}
-\newcommand{\fixnote}[2]{\textbf{\color{red}{FIX}}\footnote{{\bf #1:} #2}}
-\newcommand{\fix}[2]{{\color{red} {\bf tofix:} #2}}
-% \renewcommand{\fixnote}[2]{}
-% \renewcommand{\fix}[2]{}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\theoremstyle{corollary}
-\newtheorem{corollary}{Corollary}[section]
-\theoremstyle{lemma}
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{theorem}
-\newtheorem{theorem}{Theorem}[section]
-\theoremstyle{theorem}
-\newtheorem{example}{Example}
-
-\usepackage{authblk}
-\date{}                     %% if you don't need date to appear
-\setcounter{Maxaffil}{0}
-\renewcommand\Affilfont{\itshape\small}
-
-% for fancy table with rotated column titles
-\usepackage{adjustbox}
-\newcolumntype{R}[2]{%
-    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
-    l%
-    <{\egroup}%
-}
-\newcommand*\rot{\multicolumn{1}{R{45}{0.01em}}}% no optional argument here, please!
-\newcommand*\rota{\multicolumn{1}{R{90}{0.01em}}}% no optional argument here, please!
-\def\hb{\hbox to 10.7 cm{}}
-
-% circles
-\usepackage{wasysym}
-\newcommand{\Tdot}{$\CIRCLE$}
-\newcommand{\Thdot}{$\LEFTcircle$}
-\newcommand{\Twdot}{$\Circle$}
-
-% formulas
-\newcommand{\kframe}{\mathbf{K}}
-\newcommand{\bframe}{\mathbf{B_K}}
-\newcommand{\possibleworlds}{G}
-\newcommand{\modalrelation}{R}
-\newcommand{\actualworld}{\omega*}
-\newcommand{\world}{\omega}
-\newcommand{\World}{\Omega}
-\newcommand{\kmodel}{M}
-\newcommand{\interpretation}{\sigma}
-\newcommand{\knowledge}[1]{\mathbb{K}_{#1}}
-\newcommand{\knows}[2]{K_{#1}#2}
-\newcommand{\belief}[1]{\mathbb{B}_{#1}}
-\newcommand{\believe}[2]{B_{#1}#2}
-\newcommand{\region}{\chi}
-\newcommand{\Region}{\mathrm{X}}
-\newcommand{\agentuniverse}{\mathcal{A}}
-\newcommand{\agent}{a}
-\newcommand{\follang}{\mathcal{L}}
-\newcommand{\weakness}{w}
-\newcommand{\Weakness}{W}
-\newcommand{\vulnerability}{v}
-\newcommand{\Vulnerability}{V}
-%rcc5
-\newcommand{\rcc}{rcc}
-\newcommand{\connects}[2]{C(#1,#2)}
-\newcommand{\disconnected}[2]{\neg C(#1,#2)}
-\newcommand{\partof}[2]{P(#1,#2)}
-\newcommand{\overlaps}[2]{O(#1,#2)}
-\newcommand{\eq}[2]{EQ(#1,#2)}
-\newcommand{\pp}[2]{PP(#1,#2)}
-\newcommand{\po}[2]{PO(#1,#2)}
-\newcommand{\ppi}[2]{PPi(#1,#2)}
-\newcommand{\dr}[2]{DR(#1,#2)}
-
-\makeindex
-
-\begin{document}
-
-\title{On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}}
-\author[1]{Francesco Beltramini}
-\author[1]{Marco Rocchetto}
-\author[2]{Luca Vigan\`o}
-\affil[1]{V-Research, Verona, Italy}
-\affil[2]{King's College London, UK}
-
-\maketitle
-
-\begin{abstract}
-	The objective of this research is to develop of a theory that defines
-	(all and only) the possible insecurity and security configurations of
-	any abstract system. The theory is structured upon other theories that 
-	defines how a component of a system can be abstracted into an agent, defining how
-	agents can be formalized (both syntactically and semantically) to
-	describe an abstract system, such as a graph. Some of these theories
-	(e.g. used for the semantic
-	definition of the abstract system) are the epistemological definition
-	of knowledge, the Belief-Desire-Intent and the Assertion-Belief-Fact 
-	framework of reference, 
-	mereology, and topological structure. We argue that a mereology is the
-	most appropriate abstract underlying structure, do to its generality,
-	for defining the expressiveness of the system abstraction.  Furthermore, a
-	mereology allows us to define an ontology rather than a taxonomy.  We
-	also correlate different abstractions of the system to the TRL and the
-	engineering V-model.
-
-	We implemented a formal theory (of axioms) of a mereotopology, and of
-	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
-	uses the Z3 SMT solver. The results show that a single component (i.e.
-	agent) of an abstract system has a definite number of  different
-	insecurity configurations (e.g. 53 using RCC5 over a topological
-	structure) and only 1 secure (i.e. expected) configurations. The
-	configurations are reported as models satisfying the abstract system
-	semantics.
-
-	We considered the philosophical definition of truth behind our
-	approach, rejecting ``proof'' by induction from partial empirical evidences.
-	Our theory can be applied to system engineering and we show a concrete
-	application of our theory to the risk assessment of an ad-hoc system.
-	Finally, we provide a number of ideas to support the engineering of
-	secure systems (e.g. purely cyber or cyber-physical).
-\end{abstract}
-\newpage
-
-\section{Introduction}\label{sec:intro}
-\epigraph{Humanum est errare}{{\itshape Seneca the Elder}}
-The European Commission states in\autocite{EU2019market} that: ``Cybersecurity
-is one of the priority areas [\ldots] of the Commission initiative on ICT
-Standards, which is part of the Digitising European
-Industry\autocite{EU2019standard} strategy launched on 19 April 2016.  The aim
-is to identify the essential ICT standards and present measures to accelerate
-their development in support of digital innovations across the economy''. The
-same document (i.e.\autocite{EU2019market}) states that ``The EU will invest up
-to \euro450 million [\ldots], under its research and innovation programme
-Horizon 2020''. The EU, in 2016 published a press release\autocite{EU2016press}
-in which they present a strategy to invest \euro1.8 \emph{billion} to
-``increase measures to address cyber threats''. The EU is not the only investor
-in cybersecurity, most of the developed countries and several companies are
-investing enormous amount of money towards various aspects of cybersecurity
-(e.g. The US vulnerability databases\autocite{NIST2020NVD} maintained by the
-National Institute of Standards and Technologies, i.e. NIST, of the US
-Department of Commerce).
-
-The cybersecurity industry is growing fast, e.g. as reported
-in\autocite{Nasdaq2018market}. For example, in\autocite{Forbes2017market},
-published by the Forbes, is stated that \euro5.3 billion of funding where
-poured by venture capitalist into cybersecurity companies in 2018. The Forbes,
-in the same article, also highlights another peculiar (as seemly contradictory)
-trend: ``[\ldots] during the same time period, the number of cybersecurity
-breaches increased exponentially''. The data reported by the NIST 
-through the official CPE (Common Platform Enumeration) Dictionary Statistics 
-on the NVD websites in \autocite{NIST2020CPEstatistics}, show that in 2016 the
-number of reported vulnerabilities reported where around $~6000$ while in 2019
-the number of vulnerabilities was above $16000$.  The scientific community also
-reports similar findings.  In fact, in\autocite{Herley2009so}, Cormac Herley
-(Microsoft Research) shows how basic cybersecurity principles (such as the
-confidentiality benefit over the clear text for passwords typed into forms,
-e.g. for logins in websites) are not fully understood or shared between the
-cybersecurity research community\autocite{Nielsen2009stop}.  The lack of
-understanding of basic security principle, the inverse proportionality between
-investments in cybersecurity and the number of reported vulnerabilities year
-after year, can be linked to the lack of a foundational theory on
-cybersecurity, as already highlighted by Cormac Herley
-in\autocite{Herley2016unfalsifiability}.
-
-In this article, we give the first scientific theory (to the best of our knowledge)
-on security.
-
-\Paragraph{Structure} In Section~\ref{sec:problem} we define and formalize the problem statement. 
-In Section~\ref{sec:theory} we outline our security theory, and in Section~\ref{sec:prototest}
-we describe the implementation of the theory and some empirical tests of the theory.
-Finally, in Section~\ref{sec:related} we conclude the paper with an overview of the related work.
-
-\section{Problem Statement}\label{sec:problem}
-
 \begin{figure}[t]
@@
 \end{itemize}
-
-\subsection{Glossary -- A Formalization}\label{sec:glossary}
-We now define a formalizations of the concepts described in
-Section~\ref{sec:problem} and depicted in Figure~\ref{fig:safety-security}. We
-base our formalization on first order logic (FOL) with the standard
-truth-value semantics. The choice of this logic is required by the 
-semantics of the concepts formalized afterwards in this section.
-
-We consider a first-order language $\follang$ over a
-signature $\Sigma_P$ where $P,P',\ldots,P^n$ represent terms, and
-$\varphi,\varphi',\ldots,\varphi^n$ represent formulas.
-The syntax is defined as follows.
-\begin{displaymath}
-	\varphi := P~|~\neg\varphi~|~\varphi\wedge\varphi~|~\varphi\vee\varphi~|~\varphi\Rightarrow\varphi'~|~\forall P.~\varphi~|~\exists P.~\varphi
-\end{displaymath}
-where $\wedge$, $\neg$, $\vee$, and $\Rightarrow$ are connectives representing
-conjunction, negation, disjunction, and (material) implication respectively; while
-$\forall$ and $\exists$ represent the standard universal and existential
-quantifiers (resp.). Finally, the symbol ``$.$'' is just syntactic sugar. 
-We consider $\interpretation\subset\Phi\times\{\top,\bot\}$ as the interpretation
-function, where $\Phi$ is any collection of sentences in $\follang$ 
-and $\top$ and $\bot$ represent the concepts of ``Tautology/True''
-and ``Contradiction/False'' respectively.
-
-Following Figure~\ref{fig:safety-security}, we start by formalizing the outermost term: Causality.
-
-\subsubsection{Causality Principle}\label{sec:causality}
-We formalize the Causality Principle starting from a K modal
-logic\autocite{Garson2018modal} (i.e. without restrictions on the causality
-relation between worlds). The standard definition of K modal logic is given in
-Definition~\ref{def:modallogic} in terms of an interpretation function (which
-we named $\sigma$ with a slight abuse of notation) defined in
-Definition~\ref{def:modalinterpretation}. 
-\begin{definition}{\bf K Modal Logic --}\label{def:modallogic}
-A K-frame is a frame $ \kframe=\langle \possibleworlds,\modalrelation \rangle $
-	in the K modal logic where $ \modalrelation $ is the binary relation
-	(i.e. a set of ordered pairs) between possible worlds
-	$\modalrelation\subseteq\possibleworlds\times\possibleworlds$, where
-	$\possibleworlds$ represents the possible worlds, and
-	$\possibleworlds\neq\emptyset$. An actual world
-	$\actualworld\in\possibleworlds$ is assumed. For any proposition $P$,
-	an interpretation function $\interpretation(\world,P)$ returns the
-	truth value of $P$; e.g. $\interpretation(\world,P)=\top$ means that
-	$P$ holds in $\world$. A model is defined as the tuple
-	$\kmodel=\langle\possibleworlds,\modalrelation,\interpretation\rangle$.
-\end{definition}
-
-\begin{definition}{\bf Causality as K Modal Interpretation Function --}\label{def:modalinterpretation}
-	Causality is (recursively) defined as the modal interpretation function $\interpretation$, as follows. 
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation0)$] if $\interpretation(\world,P)=\top$ then $\world\models P$
-		\item[$(\interpretation1)$] $\world\models\neg P$ iff $\world\not\models P$
-		\item[$(\interpretation2)$] $\world\models P \wedge Q$ iff $\world\models P$ and $\world\models Q$
-		\item[$(\interpretation3)$] $\world\models\Box P$ iff for any world $\world'\in\possibleworlds$ if $\world \modalrelation\world'$ then $\world'\models P$
-		\item[$(\interpretation4)$] $\world\models\Diamond P$ iff there exists a set of worlds $\World'\subset\possibleworlds$ such that for any $\world'\in\World'$, if $\world\modalrelation\world'$ then $\world'\models P$
-		\item[$(\interpretation5)$] $\models P$ iff $\actualworld\models P$
-	\end{enumerate}
-	where truth is defined as necessary with $\Box$ and possible with $\Diamond$.
-\end{definition}
-
-%\begin{definition}{\bf Causality --}\label{def:causality}
-%	The Causality principle is defined over a K Modal Logic (i.e. without
-%	restrictions on the causality relation between worlds) as in
-%	Definition~\ref{def:modallogic}, with the standard interpretation
-%	function defined in~\ref{def:modalinterpretation}
-%\end{definition}
-
-The causality principle has been defined in its
-generic form. In fact, the accessibility relation $\modalrelation$ is free from
-any axiomatic restriction (e.g. it's not reflexive nor anti-reflexive).  
-We will focus in Section~\ref{sec:process} (for our tests) on 
-the application of our theory to CPS system engineering.
-More detailed case studies (i.e. \fix{mr}{a CPS as a smart power-grid}) 
-will be defined in Section~\ref{sec:prototest}, where
-the accessibility relation defines in details how the system itself evolves, due to causal relation 
-(i.e. by restricting the causality principle only to those cause-effect that defines
-the system). Therefore, the definition of $\modalrelation$
-will be specialized in a more strict way based on the application domains and case study.
-
-We note that we have defined the
-Causality principle without considering it a \emph{threat source} since 
-we lack the concept of intent and maliciousness. Similarly, in the 
-next sections we won't discriminate between Nature and Mankind until we introduce 
-the concepts of maliciousness in Section~\ref{sec:theory} 
-and then formally define a threat agent
-in the same section.
-
-\subsubsection{Agents: Mankind and Nature}\label{sec:mankind-nature}
-\begin{table}[t]
-\centering
-\setlength{\tabcolsep}{3.5pt}
-\renewcommand{\arraystretch}{1}
-\scriptsize
-\caption{RCC3, RCC5, and RCC8 relations between Regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
-\begin{tabular}{ccclll} 
-\rota{\textbf{RCC3}}&\rota{\textbf{RCC5}}&\rota{\textbf{RCC8}}&\textbf{Name} & \textbf{Notation} & \textbf{Definition} \\
-\hline
-&&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& $\mathit{X}\subseteq \mathit{Y}$ \\
-&&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& $\mathit{X}\not\subseteq \mathit{Y}$\\
-&&&Part of				& $\mathit{P}(\mathit{X},\mathit{Y})$		& $\forall \mathit{Z} ~\mathit{C}(\mathit{Z},\mathit{X}) \rightarrow \mathit{C}(\mathit{Z},\mathit{Y})$\\
-&&&Overlaps			& $\mathit{O}(\mathit{X},\mathit{Y})$		& $\exists \mathit{Z} ~\mathit{P}(\mathit{Z},\mathit{X})\wedge \mathit{P}(\mathit{Z},\mathit{Y})$\\
-\Tdot&&&  \textbf{Overlaps Not Equal} 	& $\mathit{ONE}(\mathit{X},\mathit{Y})$		& $\mathit{O}(\mathit{X},\mathit{Y}) \land \neg \mathit{EQ}(\mathit{X},\mathit{Y})$ \\
-\Tdot&\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
-\Tdot&\Tdot&\Tdot& \textbf{DiscRete from} 		& $\mathit{DR}(\mathit{X},\mathit{Y})$		& $\neg \mathit{O}(\mathit{X},\mathit{Y})$\\
-&\Tdot&\Tdot&\textbf{Partial-Overlap}	& $\mathit{PO}(\mathit{X},\mathit{Y})$ 		& $\mathit{O}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{X},\mathit{Y}) \wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\
-&\Tdot&&\textbf{Proper-Part-of} 	& $\mathit{PP}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\ 
-	&\Tdot&&\textbf{Proper-Part-of-\textit{\textbf{i}}nverse} & $\mathit{PPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{Y},\mathit{X}) \wedge \neg \mathit{P}(\mathit{X},\mathit{Y})$\\
-&&\Tdot&\textbf{Externally Connected} 	& $\mathit{EC}(\mathit{X},\mathit{Y})$ 		& $\mathit{C}(\mathit{X},\mathit{Y}) \wedge \neg\mathit{O}(\mathit{X},\mathit{Y})$\\ 
-&&\Tdot&\textbf{Tangential PP} 	& $\mathit{TPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Tangential PPi} 	& $\mathit{TPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{TPP}(\mathit{Y},\mathit{X})$\\ 
-&&\Tdot&\textbf{Non-Tangential PP} 	& $\mathit{NTPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\neg\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Non-Tangential PPi} 	& $\mathit{NTPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{NTPP}(\mathit{Y},\mathit{X})$\\ 
-\end{tabular}
-\end{table}
-
-Mankind and Nature are defined in Section~\ref{sec:problem} as two abstract
-agents, both as collections (i.e. an abstract type that does not imply a
-specific implementation) of their sub-agents (i.e. humans for Mankind and
-plants, animals, \&c.  for Nature). Similarly to\autocite{Santaca2016abf}, 
-we define Mankind and Nature, and any
-other agent in the reminder of this article, as a meronomy (an hierarchy of
-Part-Whole relations) based on a standard definition of mereology, i.e. based
-on the definition of Parthood relation between \emph{Parts}.  
-However, shall 
-consider different types of Part; so, we extend the mereology to a
-mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
-to increase the number of different types considered and to generalize the relations between Parts (as
-in Table~\ref{tab:rcc358}).
-For the sake of readability, we use the term \emph{Region} both to refer to a
-mereological Part and to a topological Region.
-The choice of mereotopology is also correlated to the objective of defining a formal
-ontology, which we use to define the (formal) semantics of the terms (Parts) in
-Section~\ref{sec:problem}, and of the concepts of safety and security
-(whole). We aim at creating a meronomy instead of the taxonomies such as the one provided
-in\autocite{NIST2020NVD,MITRE2020CVE} or instead of the poorly justified 
-CVSS\autocite{Mell2007CVSS} scoring system.
-
-A mereotopology, as defined e.g. in\autocite{Rachavelpula2017mereotopology},
-is an ordered mathematical structure where the basic relation between Regions is the
-reflexive and symmetric\fixnote{mr}{I guess it must be monotonic as defined in\autocite{Rachavelpula2017mereotopology} but I don't find it consistently in other papers.} 
-\emph{Parthood} relation $\subseteq$. 
-\begin{definition}{\bf Parthood --}\label{def:parthood}
-Given any pair of mereotopological Regions $X$ and $Y$,
-	\begin{enumerate}[noitemsep]
-		\item Reflexivity: $\forall X.~ (X\subseteq X)$
-		\item Symmetry: $\forall X, Y.~ (X\subseteq Y \Rightarrow Y\subseteq X)$
-	\end{enumerate}
-\end{definition}
-
-As later defined in Definition~\ref{def:agent}, 
-the Parthood relation orders a universe of agents $\agentuniverse$ by defining
-the so called \emph{Connects with} (see in Table~\ref{tab:rcc358}) relation 
-between Regions. We want this universe $\agentuniverse$ to be 
-expressible in FOL. In this way, we
-can reason both on the constituent of security (i.e. its terms and agents
-defining a system where security needs to be considered), and on 
-the evolution of those constituent w.r.t. cause-effects relations according
-to the modal structure of causality we defined in Section~\ref{sec:causality}.
-This will allow us (in Section~\ref{sec:process} and Section~\ref{sec:prototest})
-a better\fixnote{mr}{Can we prove this by construction?} 
-positioning w.r.t. risk assessment technologies
-(which most often reason on the constituent of a system design), and
-protocol verification tools (which requires some formalization of a cause-effect relation, e.g.
-in Linear Temporal Logic).
-
-As we argued in Section~\ref{sec:problem}, we must correlate the definition of
-agent (i.e. Mankind and Nature) in Definition~\ref{def:mankind-nature} to the
-mathematical structure of the logic that defines them.  We express Mankind and
-Nature as formulas over the theory of mereology and then in terms of
-mereotopological Regions, extending the interpretation function
-$\interpretation$, to include a formal theory of mereotopology.  We use the
-Region Connection Calculus (RCC), as defined
-in~\cite{bennettLogics,improvingRCC}, to provide an axiomatization of the
-spatial concepts and relations in first-order logic to correlate the algebraic
-structure to mereology. In its broader definition, the RCC theory is composed
-by eight axioms, and is known as RCC8. In the text, for brevity, 
-we will often focus only on RCC5 (without
-loss of generality) by not considering tangential connections between spatial
-Regions. We discuss the choice of RCC5 in more detail in
-Appendix~\ref{app:rcc}. In Table~\ref{tab:rcc}, we
-summarize the axioms of the Region Connection Calculus (see, e.g., \autocite{Grutter2008rcc}).
-
-\begin{definition}{\bf RCC axiomatization --}
-	For any $X,Y$ pair of Regions in a mereotopology:
-	\begin{itemize}[noitemsep]
-	\item[$(\interpretation6)$] $\interpretation(X\subseteq Y)$ iff [$\interpretation(X\subseteq X)=\top$, and $\interpretation(X\subseteq Y)=\bot$ or $\interpretation(Y\subseteq X)=\top$]\fixnote{mr}{would it be better/clearer or just correct to write $\sigma(X\subseteq Y)$ iff $\sigma(X\subseteq X\wedge [X\subseteq Y \vee Y\subseteq X])=\top$}
-	\item[$(\interpretation7)$] $\world\models\connects{X}{Y}$ iff $\world\models X\subseteq Y$ 
-	\item[$(\interpretation8)$] $\world\models\disconnected{X}{Y}$ iff  $\world\not\models X\subseteq Y$
-	\item[$(\interpretation9)$] $\world\models\partof{X}{Y}$ iff $\world\models\forall Z.~\connects{Z}{X}\Rightarrow\connects{Z}{Y}$
-	\item[$(\interpretation10)$] $\world\models\overlaps{X}{Y}$ iff $\world\models\exists Z.~\partof{Z}{X}\wedge\partof{Z}{Y}$
-	\item[$(\interpretation11)$] $\world\models\eq{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\partof{Y}{X}$
-	\item[$(\interpretation12)$] $\world\models\dr{X}{Y}$ iff $\world\models\neg\overlaps{X}{Y}$
-	\item[$(\interpretation13)$] $\world\models\po{X}{Y}$ iff $\world\models\overlaps{X}{Y}\wedge\neg\partof{X}{Y}\wedge\neg\partof{Y}{X}$
-	\item[$(\interpretation14)$] $\world\models\pp{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\neg\partof{Y}{X}$
-	\item[$(\interpretation15)$] $\world\models\ppi{X}{Y}$ iff $\world\models\partof{Y}{X}\wedge\neg\partof{X}{Y}$
-\end{itemize}
-where $Z$ is a mereotopological Region.
-\end{definition}
-
-\begin{definition}{\bf Agent: Mankind or Nature --}\label{def:agent}
-	An agent $\agent\in\agentuniverse$ is a tuple
-	$\langle\rcc(\region',\region''),\ldots,\rcc(\region^{n-1},\region^n)\rangle$
-	of RCC relations $\rcc$ over mereotopological Regions
-	${\region',\ldots,\region^n}\subseteq\Region$. 
-\end{definition}
-
-Currently, we do not distinguish between Mankind and Nature (since we still
-lack of the definition of ``malicious intent'', which is defined in
-Section~\ref{sec:theory}) and we have defined them as two generic
-agents.
-
-As depicted in Figure~\ref{fig:safety-security}, Causality, Mankind, and Nature
-have a dashed border representing their correlation in terms of cause-effect
-and then in terms of formal structure which defines them: Modal Logic.
-Vulnerability, Defect, Incident, Accident, and Threat, similarly, 
-are correlated (depicted as a solid border) in terms of underlying 
-formal structure: Mereotopology.
-
-\subsubsection{Regions: Vulnerability and Defect (and Weakness)}\label{sec:vulnerabilitydefect}
-As defined in Section~\ref{sec:problem}, a Vulnerability or a Defect is a
-\emph{Weakness}.  As an example, a categorization of weaknesses is given
-in\autocite{MITRE2020CWEresearch} with 808 weaknesses categorized as ``Research
-Concepts'', distributed as follows:
-\begin{itemize}[noitemsep]
-	\item Incorrect Calculation - ($682$)
-	\item Incorrect Access of Indexable Resource (``Range Error'') - ($118$)
-	\item Use of Insufficiently Random Values - ($330$)
-	\item Improper Interaction Between Multiple Correctly-Behaving Entities - ($435$)
-	\item Improper Control of a Resource Through its Lifetime - ($664$)
-	\item Insufficient Control Flow Management - ($691$)
-	\item Protection Mechanism Failure - ($693$)
-	\item Incorrect Comparison - ($697$)
-	\item Improper Check or Handling of Exceptional Conditions - ($703$)
-	\item Improper Enforcement of Message or Data Structure - ($707$)
-	\item Improper Adherence to Coding Standards - ($710$)
-\end{itemize}
-
-The definition given by the MITRE in\autocite{MITRE2020CWEweakness} of
-weakness is: ``Software weaknesses are errors that can lead to software
-vulnerabilities. A software vulnerability, such as those enumerated on the
-Common Vulnerabilities and Exposures (CVE) List, is a mistake in software that
-can be directly used by a hacker to gain access to a system or network''.
-The definition is circular if we interpret the word ``error'' and
-``mistake'' with the same semantics: a weakness is an error that leads to a
-vulnerability and a vulnerability is a mistake which, in turn, is a weakness.
-The only difference (between weakness and vulnerability) seems to be
-that one can consider weakness as a ground term and state that a
-vulnerability is caused by a weakness, i.e. 
-$\World,\Weakness\models\Diamond\Vulnerability\wedge\Weakness$ where
-$\Weakness,\Vulnerability$ are Regions of Weaknesses and Vulnerabilities (resp.);
-accepting the hierarchy in the CWE\autocite{CWE} as ground truth. 
-Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability) 
-or the CVE reported in the NVD as a ground truth.
-
-\begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
-	A Region $\region\subseteq\agent$ of an agent
-	$\agent\in\agentuniverse$, is defined as Weakness $\Weakness$ (i.e.
-	representing a weakness introduced by the agent into any phase of
-	production, e.g.  of the secure process development life-cycle, of any
-	system or subsystem) iff there exists $\world\in\possibleworlds$ such
-	that $\world,\Weakness\models\Diamond\exists\Weakness',\Vulnerability'.~
-	\rcc(\Weakness',\Vulnerability')\wedge\neg\dr{\Weakness'}{\Vulnerability'}$, 
-	where $\Weakness'\subseteq\Weakness$, and
-	$\Vulnerability'\subseteq\Vulnerability$ is a Region of an agent $\agent'\in\agentuniverse$.
-\end{definition}
-
-\begin{example}{CWE-116: Improper Encoding or Escaping of Output\autocite{CWE-116} --}
-	\begin{itemize} 
-		\item Description: The software prepares a structured message
-		for communication with another component, but encoding or
-		escaping of the data is either missing or done incorrectly. As
-		a result, the intended structure of the message is not
-		preserved. 
-		\item Example: This code displays an email
-		address that was submitted as part of a form. Example language JSP.
-		\begin{verbatim} 
-		<% String email = request.getParameter("email"); %> 
-		...
-		Email Address: <%= email %>
-		\end{verbatim}
-		The value read from the form parameter is
-		reflected back to the client browser without
-		having been encoded prior to output, allowing
-		various XSS attacks (CWE-79).
-		\item Observed Examples
-			\begin{itemize}
-			\item CVE-2008-4636\autocite{CVE-2008-4636}: OS command injection in backup
-				software using shell metacharacters in a
-				filename; correct behavior would
-				require that this filename could not be
-				changed.
-			\item CVE-2008-0769\autocite{CVE-2008-0769}: Web application does not set the
-				charset when sending a page to a browser,
-				allowing for XSS exploitation when a
-				browser chooses an unexpected encoding.
-			\item CVE-2008-0005\autocite{CVE-2008-0005}: Program does not set the charset
-				when sending a page to a browser, allowing for
-				XSS exploitation when a browser chooses
-				an unexpected encoding.
-			\item CVE-2008-5573\autocite{CVE-2008-5573}: SQL injection via password
-				parameter; a strong password might contain \&
-			\item CVE-2008-3773\autocite{CVE-2008-3773}: Cross-site scripting in chat
-				application via a message subject, which
-				normally might contain \& and other
-				XSS-related characters.
-			\item CVE-2008-0757\autocite{CVE-2008-0757}: Cross-site scripting in chat
-				application via a message, which normally might
-				be allowed to contain arbitrary
-				content.
-			\end{itemize}
-\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\begin{subfigure}[b]{.5\textwidth}
-		\centering
-		\includegraphics[width=.7\textwidth]{rcc5_CWE-116.pdf}
-		\caption{Representation with RCC3}
-		\label{fig:rcc5_CWE-116}
-	\end{subfigure}\begin{subfigure}[b]{.5\textwidth}
-		\centering
-		\includegraphics[width=.7\textwidth]{rcc8_CWE-116.pdf}
-		\caption{Representation with RCC8}
-		\label{fig:rcc8_CWE-116}
-	\end{subfigure}
-	\begin{subfigure}[b]{\textwidth}
-		\centering
-		\includegraphics[width=\textwidth]{KML_CWE-116.pdf}
-		\caption{Representation in K-Modal Logic}
-		\label{fig:KML_CWE-116}
-	\end{subfigure}
-	\caption{Relations between CWE-116 and correlated CVEs}
-\end{figure}
-Definition~\ref{def:weakness} states that CWE-116 is a weakness iff there exists a world in
-K Modal Logic, representing the system in which this weakness exists, 
-such that the natural evolution of this system (i.e. formalized by the causality principle) 
-make it possible to reach another state of the system (i.e. another, accessible world)
-where there exist a vulnerability that can be implemented from CWE-116 (and this relation 
-is not DR). The CWE website proposes the connection between CWE-116 and, for example, 
-CVE-2008-5573; a vulnerability of the login sub-system 
-of the Poll Pro v2.0\autocite{pollpro} system.
-The formal relation between the two is given as a link (i.e. URI) between the CWE and
-the CVE, the description of the relation is not defined but it is supposed to be
-inferred from the descriptions of the CWE-CVE. 
-We can formally represent this link as the ONE connection in RCC3, depicted in 
-Figure~\ref{fig:rcc5_CWE-116}, or EC connection in RCC8 (Figure~\ref{fig:rcc8_CWE-116}).
-
-\fixnote{mr}{shouldn't we consider the agent who introduces the weakness as dishonest? don't we say this in part 1?}
-\end{example}
-
-It is interesting to note that the CWE-CVE relation expresses the correlation
-between weaknesses and vulnerabilities in the most simple form, such that we
-can formalize all the relation as ONE in RCC3 or EC in RCC8.  To express a more
-complex relation between the two we shall analyze the definition of CWE-116.
-This is related to the Weakness-Vulnerability-Incident process (i.e. to the
-details on the implementation/realization in Figure~\ref{fig:safety-security})
-that we analyze in the next section, Section~\ref{sec:incidentaccident}.
-
-\begin{definition}{\bf Region: Vulnerability and Defect --}\label{def:vulnerability-defect}
-A region $\region\subseteq\agent$ of an agent $\agent\in\agentuniverse$ is
-	called \emph{Vulnerability} if the agent $\agent$ is referred to as Mankind, 
-	\emph{Defect} if the agent is referred to as Nature.
-\end{definition}
-
-\subsubsection{Process: Incidents and Accidents}\label{sec:incidentaccident}
-In our informal definition depicted in Figure~\ref{fig:safety-security}, an
-Incident is generated through a process caused by a Vulnerability (which,
-in turn, is caused by a Weakness). Symmetrically, Nature has a process from Defect
-to Accidents. We start by analyzing the process generated by Mankind, which is
-divided into the following phases (explained in the remainder of this sub-Section):
-%Agent$\rightarrow$Vulnerability$\rightarrow$Exploit$\rightarrow$Attack$\rightarrow$Incident.
-\begin{enumerate}[noitemsep]
-	\item An agent (Mankind or Nature) defines a system, e.g.
-by following an engineering process for creating a new system, or for redesigning or improving an existing system (e.g.
-a component or a collection of components)
-	\item The dogmatic definition of a system contains a Weakness (i.e.
-the agent, Mankind or Nature, introduces an unintended behavior), for example as an un-intended behavior or as an architectural fallacy.
-The Weakness can be applied to the abstract definition of the system.
-	\item The Weakness can be defined as a malicious process for a system, called Vulnerability (or Defect for Nature)
-	\item The process expressed by the Vulnerability can be made operational for a specific implementation of the system (i.e. implemented by Mankind or realized by Nature) 
-\end{enumerate}
-Those additional details are depicted in Figure~\ref{fig:safety-security_2}, an update
-of Figure~\ref{fig:safety-security}.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{safety-security_2.pdf}
-	\caption{Overview security and safety keywords}
-	\label{fig:safety-security_2}
-\end{figure}
-
-\Paragraph{Exploitation Flow} An \emph{Exploit} is an \emph{implementation} of
-a Vulnerability, where a Vulnerability is one of the many ways to transform a
-Weakness into a malicious process that can be applied\footnote{Meaning that the
-malicious process, i.e.  any structured procedure such as a protocol, can be
-somehow made operational, e.g. implemented} to one or many systems. So, the act
-of an agent (Mankind) of ``exploiting a Vulnerability'' is the process of
-making the Vulnerability operational, by implementing the Vulnerability w.r.t.
-a specific target system.  Whenever the agent is Nature, the implementation is
-more broadly considered a realization\footnote{``The state of being
-realized''\autocite{Merriam2020realization}}
-
-The Common Attack Pattern Enumeration and
-Classification\autocite{MITRE2020CAPEC} (CAPEC) online service provides a
-``community resource'' which aims at ``identifying and understanding attacks''.
-The MITRE writes on the CAPEC homepage that ``Understanding how the adversary
-operates is essential to effective cyber security. CAPEC helps by providing a
-comprehensive dictionary of known patterns of attack employed by adversaries to
-exploit known weaknesses in cyber-enabled capabilities. It can be used by
-analysts, developers, testers, and educators to advance community understanding
-and enhance defenses.''. We can extrapolate two major concepts: 
-\begin{enumerate}
-	\item while understanding the exploitation process is the aim of the
-		CWE and CVE initiatives, understanding the hacking process is
-		the aim of the CAPEC initiative.
-	\item CAPEC provides a \emph{dictionary} of \emph{known} attack patterns.
-		As stated largely in Section~\ref{sec:}, this dictionary contains
-		empirical evidences and should not be used to induce a cybersecurity 
-		theory; but it can be used to validate a theory.
-\end{enumerate}
-
-The process continues as follows.
-\begin{enumerate}[noitemsep]
-\setcounter{enumi}{4}
-	\item an agent (Mankind\footnote{For the sake of readability we focus only on Mankind as an agent}) the exploitation flow can be used to hack a system.
-We use the term hack (as hacking) to stress that nothing prevents this process from being honest, however, for the sake of simplicity, we focus on Attacks which are malicious by definition.
-	\item An Attack is applied to a system with malicious intent by an agent
-	\item The application of the attack results in an Incident which impacts, i.e. poses a Threat to, an Asset
-\end{enumerate}
-
-\Paragraph{Hacking Flow}
-
-\begin{example}
-given in natural language in the description of the 
-CWE-116 which we can analyze as follows:
-\begin{itemize}
-	\item ``the software prepares a structured message''
-	\item the message is intended ``for communication with another component''
-	\item ``but''
-	\item ``encoding or escaping of the data is either missing or done incorrectly''
-	\item ``[\ldots] result, the intended structure of the message is not preserved''
-\end{itemize}
-\end{example}
-
-In this Vulnerability is seen as a process through which an agent (i.e. Mankind)
-generates an Incident, i.e.
-
-We describe processes, in our current formalization, in terms of the K-Modal
-Structure as in Definition~\ref{def:modalinterpretation} (i.e. in terms of
-causality principle).
-
-\begin{definition}{\bf Vulnerability-Incident Process --}\label{def:vulnerability-incident}
-	Given two 
-\end{definition}
-
-The whole-part relation \fix{mr}{(i.e.  HAS-A\footnote{HAS-A is just an hint for the
-mechanization of this theory.  We'll come back to this in
-Section~\ref{sec:implementation}})} is given, in the following, over the K Modal
-logic described beforehand. To do so, we first need to formally define the
-concept of \emph{agent} in its abstract form. We build it following the same
-structure of\autocite{Santaca2016abf} but over the terms used in
-Section~\ref{sec:problem}, i.e. knowledge, and belief as epistemic concepts. We'll then 
-introduce the concept of \emph{intent} to define a \emph{threat agent}.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.8\textwidth]{knowledge-belief.pdf}
-	\caption{Informal representation of Knowledge and Belief}
-	\label{fig:knowledge-belief}
-\end{figure}
-
-\subsubsection{Region: Threat and Asset}\label{sec:threatasset}
-The difference between Knowledge and Belief is depicted in
-Figure~\ref{fig:knowledge-belief} (see \autocite{wiki-knowledgebelief}).  However,
-according to\autocite{Gettier2012knowledge},
-Knowledge\footnote{``\emph{Theaetetus}: [\ldots] He said that knowledge was
-true opinion accompanied by reason, but that unreasoning true opinion was
-outside of the sphere of knowledge; and matters of which there is not a
-rational explanation are unknowable -- yes, that is what he called them -- and
-those of which there is are knowable. [\ldots] \emph{Socrates}: [\ldots] the
-primary elements of which we and all else are composed admit of no rational
-explanation; for each alone by itself can only be named, and no qualification
-can be added, neither that it is nor that it is not, for that would at once be
-adding to it existence or non-existence, whereas we must add nothing to it, if
-we are to speak of that itself alone.  [\ldots]'' Plato -- Theaetetus 201
-\autocite{Plato1914Plato}} as an epistemological concept is difficult (and
-sometimes believed impossible\autocite{citation}) to formally define. In this
-work, we are not interested in 
-how knowledge can be precisely formalized from an epistemic standpoint. 
-We assume that a semantic of a correct definition of epistemic knowledge exists,
-for example the one given in\autocite{Hintikka1962knwoledge} by Hintikka, and we then define 
-knowledge in terms of the Kripke structure defined in Definition~\ref{def:causality}.
-
-\begin{definition}{\bf Knowledge --}\label{def:knowledge}
-Given an abstract collection of Agents $Ag$, and the modal operator
-	$\knows{a}{}$ (where $a\in Ag$), Knowledge is defined as a collection
-	of predicates known by an agent $\knowledge{a}=\bigcup_\Phi \knows{a}{\varphi}$ 
-	(where $\Phi$ is the collection of all the propositions known by $a$).
-	Given a proposition $P$, we extend the semantics of the Causality structure with:
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation6)$] $\world\models\knows{a}{P}$ iff
-			$\world'\models P$ for all $\world'$ such that
-			$\world\modalrelation\world'$.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Belief --}\label{def:belief}
-	\fix{mr}{we should help the reader in understanding the upgrades of the logic.
-We used propositional with modal operators with a K Kripke semantics and we are now introducing Belief as in Doxastic logic?
-and we have epistemological operators as knowledge}
-	Given an abstract collection of Agents $Ag$, and the modal operator
-	$\believe{a}{}$ (where $a\in Ag$), Belief is defined as a collection
-	of predicates believed by an agent $\belief{a}=\bigcup_\Phi \believe{a}{\varphi}$
-	(where $\Phi$ is the collection of all the propositions believed by $a$).
-	Given a proposition $P$, we extend the semantics of the Causality structure with:
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation7)$] $\world\models\believe{a}{P}$ iff
-			$\world\models\neg\knows{a}{\neg P}$ (i.e. the agent $a$ considers $P$ possible) 
-			and $\world'\models P$ for all
-			$\world'$ such that $\world\modalrelation\world'$.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Mankind/Nature --}\label{def:mankind-nature}
-Mankind/Nature is represented as an agent.  
-	%countable logical signature $\Sigma_M$/$\Sigma_N$ over an
-	%abstract algebraic structure.  We also consider a countable signature
-	%of variables, disjoint from $\Sigma_M$/$\Sigma_N$, and write
-	%$\Sigma_{M_n}$/$\Sigma_{N_n}$ for the symbols of $\Sigma_M$/$\Sigma_N$
-	%with arity $n$. Thus, $\Sigma_0$ is a collection of constants, which we
-	%assume to have disjoint sub-collections that we refer to as
-	%\emph{sub-agents} of Mankind or Nature (i.e. humans, plants, \&c).
-	%The variables are untyped and can be instantiated with arbitrary types,
-	%yielding an untyped model.
-\end{definition}
-
-\subsubsection{Mereotopological Representation of Safety and Security}\label{sec:formalGlossary}
-The objectives of this section are:
-\begin{enumerate}[noitemsep]
-	\item formalize the abstract algebraic structure (as a mereotopology)
-		on top of which we have formalized the terms in
-		Figure~\ref{fig:safety-security} so far, i.e. Mankind, Nature,
-		Truths, Beliefs, and Knowledge.
-	\item correlate mereological structure to Causality and then Kripke
-		structure by defining the relation $R$ over the Kripke
-		structure as the mereotopological basic relation connects with,
-		i.e. reflexive and symmetric
-	\item we map signature to mereotopological Regions to express the
-		logical formulation of terms (e.g. Mankind, Nature, Truths,
-		Beliefs, Knowledge, Vulnerabilities) over the mereotopological
-		structure
-	\item we formalize the relations between terms using the RCC(5)
-\end{enumerate}
-
-\fix{mr}{change propositions in propositional language\url{http://logic.stanford.edu/intrologic/lectures/lecture_02.pdf}}
-
-Given that the concepts of Truths, Beliefs, and Knowledge are positioned over
-an underlying structure of Propositions, the first concept to formalize is that
-of Propositions. We do not restrict Propositions to a specific structure but,
-as for Mankind and Nature, we refer to them as a collection.  We express the
-generality of a collection as a (propositional) signature over an abstract
-algebraic structure. In our formalization, we use mereotopology as algebraic
-structure (due to its generality, as discussed in Section~\ref{sec:problem})
-and the RCC to formalize concepts over a mereotopology.
-
-We can now define the possible mereotopological relations between Belief and
-Knowledge (informally depicted in Figure~\ref{fig:knowledge-belief}) and then
-define the peculiarity of the believed difference between Mankind and Nature 
-(i.e. maliciousness as intent). 
-We first start from the formalization of Figure~\ref{fig:knowledge-belief}, as
-follows.
-
-\begin{definition}{\bf Information}\label{def:information}
-\end{definition}
-\begin{definition}{\bf Desire}\label{def:desire}
-\end{definition}
-\begin{definition}{\bf Intent}\label{def:intent}
-\end{definition}
-
-\subsection{Vulnerability and Defect}\label{sec:vulnerability-defect}
-In order to formalize the semantics of the concepts (informally) described in
-Section~\ref{sec:problem},
-
-\begin{definition}{\bf Vulnerability}
-\end{definition}
-
-\subsection{Cybersecurity Engineering Process}\label{sec:process}
-%\epigraph{La differenza tra idee e realt\`a \`e la differenza tra
-%filosofia e ingegneria. Il lavoro per trasformare l'una nell'altra \`e ricerca
-%scientifica
-\epigraph{The difference between ideas and reality is the
-difference between philosophy and engineering. The work to transform one into
-the other is scientific research}{{\itshape V-Research}}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{vmodel.pdf}
-	\caption{Cybersecurity Engineering Life-cycle}
-	\label{fig:knowledge-belief}
-\end{figure}
-
-\subsubsection{Relevant Standards}\label{sec:standards}
-\begin{enumerate}[noitemsep]
-	\item DO-326A
-\end{enumerate}
-
-\section{Honesty and Maliciousness -- A Cybersecurity Theory}\label{sec:theory}
-\begin{enumerate}[noitemsep]
-	\item cybersecurity as a requirement
-	\item vulnerability\label{def:vulnerability}
-\end{enumerate}
-
-\section{Prototype and Empirical Tests}\label{sec:prototest}
-\subsection{Prototype Implementation}\label{sec:implementation}
-\subsection{Empirical Tests}\label{sec:tests}
-
-\section{Conclusion and Related Work}\label{sec:related}
-
-\printbibliography
-
-\begin{appendices}
-	\section{RCC3, RCC5, or RCC8}\label{app:rcc}
-	\section{Wrong idea 1}
-In order to apply consistently the mereotopological structure not just to
-reason the agents but also on their evolution, we extend the properties of the
-Parthood relation (reflexivity and symmetry) to the modal relation
-$\modalrelation$ of the K Modal logic, which in turns defines the causality
-principle. This updates the K Modal logic to a B (Brouwer) Modal Logic\autocite{Garson2018modal}.
-
-\begin{definition}{\bf B Modal Logic --}\label{thm:bk-parthood}
-	The B Modal Logic (defined, with a slight abuse of notation as
-	$\bframe=\langle\possibleworlds,\modalrelation\rangle$) is an extension
-	of the K modal logic with the following axioms
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation8)$] $\world\models P\Rightarrow Q$ iff
-			$\world\not\models P$ or $\world\models Q$ (logical implication)~\footnote{In
-			Definition~\ref{def:modallogic} we didn't specify the
-			semantics in the case of the connective $\Rightarrow$
-			(i.e. logical implication), so we state it here. Given
-			that $P\Rightarrow Q$ is equisatisfiable to
-			$\neg(P\wedge \neg Q)$, for any proposition $P$ and
-			$Q$, $\world\models P\Rightarrow Q$ is equisatisfiable
-			to $\world\models \neg(P\wedge \neg Q)$. If
-			$\world\models \neg(P\wedge \neg Q)$ then
-			$\world\not\models P\wedge\neg Q$ and then
-			$\world\models \neg P$ or $\world\models Q$.}
-		\item[$(\interpretation9)$] $\Box P\Rightarrow P$ (which ``claims that whatever is necessary is the case''\autocite{Garson2018modal}), and
-		\item[$(\interpretation10)$] $P\Rightarrow\Box\Diamond P$
-			(which says that if $P$ is the case, then $P$ is
-			necessarily possible\autocite{Garson2018modal})
-	\end{enumerate}
-\end{definition}
-
-\begin{theorem}{\bf Causality, Parthood, and B Modal Logic}\label{thm:bk-parthood}
-	\begin{enumerate}[noitemsep]
-		\item (Reflexivity) $\bframe\models\Box P\Rightarrow P$ ($\interpretation8$) iff $\modalrelation$ is reflexive
-		\item (Symmetry) $\bframe\models P\Rightarrow\Box\Diamond$ ($\interpretation9$) iff $\modalrelation$ is symmetric
-		\item (Consistency) $\world\modalrelation\world'$ iff $\connects{\region}{\region'}$
-	\end{enumerate}
-	whenever $\world$ and $\world'$, and $\region$ and $\region'$ are logically equivalent ($\equiv$).
-\end{theorem}
-\begin{proof}
-	The proof is exhaustive over all possible cases.
-	\begin{enumerate}[noitemsep]
-		\item(Reflexivity) proven in\autocite{}
-		\item(Symmetry) proven in\autocite{}
-		\item(Equivalence) \fix{mr}{todo} If
-			$\world\modalrelation\world'$ and
-			$\world\equiv\region$, then if
-			it doesn't hold that $\connects{\region,\region'}$ it
-			follows that there is no reflexive and transitive
-			relation that correlates $\region$ to $\region'$. Given 
-			that Regions are defined as signatures
-	\end{enumerate}
-\end{proof}
-
-\end{appendices}
-
-\end{document}

### 2020-02-20 theory.tex
add intro to theory

@@
-\begin{enumerate}[noitemsep]
-	\item cybersecurity as a requirement
-	\item vulnerability\label{def:vulnerability}
-\end{enumerate}
+In order to address the problem raised by Herley, 
+we shall define how to distinguish between a secure and an insecure system.
+While most of the literature (and this paper so far) have
+correlated the problem of in-security to the maliciousness of
+agents interacting with the system, we show that security
+doesn't seem to have a malicious nature but, rather, insecurity 
+raises from the lack of well-defined security requirements for the
+design process of a system. We argue that the high number
+of security vulnerability reported today are simply the realization
+of potential system configurations, which deviate from the unintended 
+behavior just because the intended behavior of the system is not precisely 
+defined in the specification of the design process.
 
-We describe processes, in our current formalization, in terms of the K-Modal
-Structure as in Definition~\ref{def:modalinterpretation} (i.e. in terms of
-causality principle).
+Given that the application of our theory is \emph{systems engineering} and that
+the two foundational processes (exploitation and hacking) are based on the
+concept of a system, we shall begin by defining what a system is, in its
+abstract form.
 
-\begin{definition}{\bf Vulnerability-Incident Process --}\label{def:vulnerability-incident}
-	Given two 
-\end{definition}
+\subsection{Abstract Systems}\label{sec:system}
+The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) 
+provides a definition of system as ``A combination of interacting elements organized to
+achieve one or more stated purposes.''\autocite{ISO201515288}.
+Therefore, a system can be considered as a single agent where it's interacting
+elements are the constituents of the agent itself. For the sake of simplicity
+we first define a system as an agent and then extend the definition
+to a ``combination of interacting'' agents.
 
-The whole-part relation \fix{mr}{(i.e.  HAS-A\footnote{HAS-A is just an hint for the
-mechanization of this theory.  We'll come back to this in
-Section~\ref{sec:implementation}})} is given, in the following, over the K Modal
-logic described beforehand. To do so, we first need to formally define the
-concept of \emph{agent} in its abstract form. We build it following the same
-structure of\autocite{Santaca2016abf} but over the terms used in
-Section~\ref{sec:problem}, i.e. knowledge, and belief as epistemic concepts. We'll then 
-introduce the concept of \emph{intent} to define a \emph{threat agent}.
+We use the same process of \autocite{Santaca2016abf} but over slightly different
+terms. The differences and commonalities with \autocite{Santaca2016abf} 
+are described in Appendix~\ref{app:abf}. There is no 
+agreement between the research communities (e.g. Multi-Agent-System, Epistemic Logic)
+on which are the constituent of an agent as a system. However,
+the same ideas revolved around for thousands of years. Some examples
+follows.
+\begin{itemize}
+	\item In \autocite{Georgeoff1995bdi}, the BDI (Belief, Desire, Intent)
+		paradigm is explained; where an agent can be defined over those
+		three constituents.
+	\item In \autocite{Hintikka1962knwoledge}, Hintikka describes the
+		difference between Knowledge and Belief (as epistemological
+		concepts), and the whole Doxastic logic defines in details how
+		Beliefs can be formalized.
+	\item In \autocite{Hintikka1993Information}, Hintikka describes the concept
+		of Information and the difference with Knowledge and Belief.
+	\item In \autocite{Empiricus1990Pyrrhonism}, the author states: ``The
+		logical criterion also may be used in three senses -- of the
+		agent, or the instrument, or the ``according to what''; the
+		agent, for instance, may be a man, the instrument either sense
+		perception or intelligence, and the ``according to what'' the
+		application of the impression ``according to'' which the man
+		proceeds to judge by means of one of the aforesaid
+		instruments.'' 
+	\item In \autocite{Santaca2016abf}, the authors defines an agent as a
+		tuple of Assertions, Beliefs, and Facts.
+\end{itemize}
 

### 2020-02-24 theory.tex
add image to represent the relation between knowledge and belief in theory.tex
definition of system in abf

@@
 of security vulnerability reported today are simply the realization
-of potential system configurations, which deviate from the unintended 
+of potential system configurations, which deviate from the nominal
 behavior just because the intended behavior of the system is not precisely 
@@
 
-Given that the application of our theory is \emph{systems engineering} and that
+Given that we apply our theory to \emph{systems engineering} and that
 the two foundational processes (exploitation and hacking) are based on the
@@
 \subsection{Abstract Systems}\label{sec:system}
-The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) 
-provides a definition of system as ``A combination of interacting elements organized to
-achieve one or more stated purposes.''\autocite{ISO201515288}.
-Therefore, a system can be considered as a single agent where it's interacting
-elements are the constituents of the agent itself. For the sake of simplicity
-we first define a system as an agent and then extend the definition
-to a ``combination of interacting'' agents.
+The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
+of system as ``A combination of interacting elements organized to achieve one
+or more stated purposes.''\autocite{ISO201515288}.  Therefore, a system can be
+considered as a single agent where its interacting elements are the
+constituents of the agent itself. For the sake of simplicity we first define a
+system as an agent and then extend the definition to a ``combination of
+interacting'' agents.  We use the same process of \autocite{Santaca2016abf} but
+over slightly different terms. The differences and commonalities with
+\autocite{Santaca2016abf} are described in Appendix~\ref{app:abf}.
 
-We use the same process of \autocite{Santaca2016abf} but over slightly different
-terms. The differences and commonalities with \autocite{Santaca2016abf} 
-are described in Appendix~\ref{app:abf}. There is no 
-agreement between the research communities (e.g. Multi-Agent-System, Epistemic Logic)
-on which are the constituent of an agent as a system. However,
-the same ideas revolved around for thousands of years. Some examples
-follows.
+There is no agreement between the research communities (e.g.
+Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent
+as a system. However, the same ideas revolved around for thousands of years.
+Some examples follows.
 \begin{itemize}
@@
 		three constituents.
-	\item In \autocite{Hintikka1962knwoledge}, Hintikka describes the
+	\item In \autocite{Hintikka1962knowledge}, Hintikka describes the
 		difference between Knowledge and Belief (as epistemological
@@
 
+If we consider \autocite{Empiricus1990Pyrrhonism}, the author applies a
+``logical criterion'' to an agent, for instance a man, for instance a device in
+a system (or a system) as in \autocite{Santaca2016abf}.  The instrument identified is sense
+perception, because a man is considered as example, but we can abstract the
+concept of instrument as any received information, where information is
+considered as defined in \autocite{Hintikka1993Information}; which can be
+summarized by the following axiom: ``Information is specified by specifying
+which alternatives concerning the reality it admits and which alternatives
+excludes''. This means, for our argument, that considering a propositional
+variable, which admits the two alternatives True/False, its information is
+defined as excluding either to be True or False and then, as we'll discuss
+later on, Believed to be the opposite. Similarly, when reasoning on a
+mereotopological region w.r.t. the RCC (e.g. RCC3 admits two regions to be
+equal, disjoint, or overlapping) and information on this region should exclude
+some of the alternatives. Finally, in \autocite{Empiricus1990Pyrrhonism}
+the author states that the logical criterium can be judged according to something.
+For our argument, as in \autocite{Santaca2016abf}, we define this concept
+as Knowledge (or Facts); where knowledge is defined, as in \autocite{Steup2020epistemology}, as
+a set of proposition known by an agent, such that: (i) knowledge requires belief,
+(ii) knowledge require truth, (iii) knowledge must be \emph{properly justified}.
+
+Before detailing what ``properly justified'' means, we discuss the mentioned
+relation between agent and belief and we introduce a more formal definition
+of knowledge and belief. For our argument, the only objective of
+Information is to interact with the behavior of a system or an agent. 
+Due to the definition of Information and its relation to the probabilistic
+correlation between information and truth/reality, we must consider
+information in relation with an agent's beliefs. Similarly, we consider
+beliefs to define the behavior of an agent or a system. On the other hand,
+Knowledge drives the nominal behavior of an agent or a system. This view
+is not considered in most (if not all) the approaches to protocol verification
+and the Dolev-Yao theory is usually applied to a representation/abstraction
+of a protocol that only considers knowledge and transfer of knowledge (i.e.
+if an agent sends a message, the recipient knows that message). However,
+the security of a system is tightly related to the difference between
+the nominal behavior and the implemented one.
+
+\begin{definition}{\bf System --}\label{def:system}
+	A system (or a sub-system) is an agent (see Definition~\ref{def:agent})
+	such that
+	$s=\langle\rcc(\knowledgeRegion,\beliefRegion),\rcc(\knowledgeRegion,\informationRegion),\rcc(\beliefRegion,\informationRegion)\rangle$,
+	where $\knowledgeRegion,\beliefRegion,\informationRegion$ are Regions of Knowledge, Beliefs, and Information respectively.
+\end{definition}
+A system is then defined as a superposition of multiple states where knowledge,
+beliefs, and information are related between each other (in a mereotopological
+space) through the RCC. A \emph{system instance} is a specific definition of
+the $\rcc$ relations between the three regions. Therefore, a set of 
+system instances defines a set of choices for those $\rcc$ relations.
+
+% \item  ``normally concern the state or the history of the entire universe but only
+% of some small part of
+% it''.  
+% \item  ``Information and probability are inversely related'', i.e. ``
+% the more alternatives a proposition admits of, the more probable and the less
+% informative it is, and vice versa''
+
 \begin{figure}[t]
@@
 
-\subsubsection{Region: Threat and Asset}\label{sec:threatasset}
 The difference between Knowledge and Belief is depicted in
@@
 we are to speak of that itself alone.  [\ldots]'' Plato -- Theaetetus 201
-\autocite{Plato1914Plato}} as an epistemological concept is difficult (and
-sometimes believed impossible\autocite{citation}) to formally define. In this
-work, we are not interested in 
-how knowledge can be precisely formalized from an epistemic standpoint. 
-We assume that a semantic of a correct definition of epistemic knowledge exists,
-for example the one given in\autocite{Hintikka1962knwoledge} by Hintikka, and we then define 
-knowledge in terms of the Kripke structure defined in Definition~\ref{def:causality}.
+\autocite{Plato1914Plato}} as an epistemological concept is difficult to formally define. 
+Similarly for the concept of information, Hintikka in \autocite{Hintikka1993Information}
+states that ``A purely logical definition of information is impossible''.
+In this work, however, we are not interested in how knowledge, information, or
+belief can be precisely formalized from an epistemic standpoint.  We assume
+that a semantic of a correct definition of epistemic knowledge exists, for
+example the one given in\autocite{Hintikka1962knowledge} by Hintikka, and we
+then define knowledge in terms of the Kripke structure defined in
+Definition~\ref{def:modallogic}; similarly for Belief.
 
@@
 	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation6)$] $\world\models\knows{a}{P}$ iff
+		\item[$(\interpretation16)$] $\world\models\knows{a}{P}$ iff
 			$\world'\models P$ for all $\world'$ such that
@@
 \begin{definition}{\bf Belief --}\label{def:belief}
-	\fix{mr}{we should help the reader in understanding the upgrades of the logic.
-We used propositional with modal operators with a K Kripke semantics and we are now introducing Belief as in Doxastic logic?
-and we have epistemological operators as knowledge}
 	Given an abstract collection of Agents $Ag$, and the modal operator
@@
 	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation7)$] $\world\models\believe{a}{P}$ iff
+		\item[$(\interpretation17)$] $\world\models\believe{a}{P}$ iff
 			$\world\models\neg\knows{a}{\neg P}$ (i.e. the agent $a$ considers $P$ possible) 
@@
 
-\begin{definition}{\bf Mankind/Nature --}\label{def:mankind-nature}
-Mankind/Nature is represented as an agent.  
-	%countable logical signature $\Sigma_M$/$\Sigma_N$ over an
-	%abstract algebraic structure.  We also consider a countable signature
-	%of variables, disjoint from $\Sigma_M$/$\Sigma_N$, and write
-	%$\Sigma_{M_n}$/$\Sigma_{N_n}$ for the symbols of $\Sigma_M$/$\Sigma_N$
-	%with arity $n$. Thus, $\Sigma_0$ is a collection of constants, which we
-	%assume to have disjoint sub-collections that we refer to as
-	%\emph{sub-agents} of Mankind or Nature (i.e. humans, plants, \&c).
-	%The variables are untyped and can be instantiated with arbitrary types,
-	%yielding an untyped model.
+\begin{definition}{\bf Information --}\label{def:information}
+	Given an abstract collection of Agents $Ag$, and the modal operator
+	$\informs{a}{}$ (where $a\in Ag$), Information is defined as a collection
+	of beliefs asserted by an agents $a$.
+	%and transferred from $a$ to $b$ ($a\rightarrow b$).
+	$\information{a}=\bigcup_\Phi \informs{a}{}{\varphi}$
+	(where $\Phi$ is the collection of all the propositions believed and asserted by $a$).
+	Given a proposition $P$, we extend the semantics of the Causality structure with:
+	\begin{enumerate}[noitemsep]
+		\item[$(\interpretation18)$] $\world\models\informs{a}{P}$ iff
+			$\world\models\neg\knows{a}{\neg P}$ (i.e. the agent $a$ considers $P$ possible) 
+			and $\world'\models P$ for all
+			$\world'$ such that $\world\modalrelation\world'$.
+	\end{enumerate}
 \end{definition}
 
-\subsubsection{Mereotopological Representation of Safety and Security}\label{sec:formalGlossary}
-The objectives of this section are:
-\begin{enumerate}[noitemsep]
-	\item formalize the abstract algebraic structure (as a mereotopology)
-		on top of which we have formalized the terms in
-		Figure~\ref{fig:safety-security} so far, i.e. Mankind, Nature,
-		Truths, Beliefs, and Knowledge.
-	\item correlate mereological structure to Causality and then Kripke
-		structure by defining the relation $R$ over the Kripke
-		structure as the mereotopological basic relation connects with,
-		i.e. reflexive and symmetric
-	\item we map signature to mereotopological Regions to express the
-		logical formulation of terms (e.g. Mankind, Nature, Truths,
-		Beliefs, Knowledge, Vulnerabilities) over the mereotopological
-		structure
-	\item we formalize the relations between terms using the RCC(5)
-\end{enumerate}
+When dealing with a specification of a system (i.e. the design if the system
+for a specific operation (as a generic operating/operational system),
+considering information may be an over-complication since we are usually not
+interested in the information in abstract but to the transfer of that
+information, which we call Assertions. Therefore, we consider information only
+if there is a sender and a recipient.  Similarly, the portion of Beliefs we
+consider is the one that builds the behavior of an agent and we consider a set
+of axiomatic Facts instead of Knowledge.
 
-\fix{mr}{change propositions in propositional language\url{http://logic.stanford.edu/intrologic/lectures/lecture_02.pdf}}
+\begin{definition}{\bf Assertion}\label{def:assertion}
+\end{definition}
+\begin{definition}{\bf Behavior}\label{def:behavior}
+\end{definition}
+\begin{definition}{\bf Fact}\label{def:fact}
+\end{definition}
 
-Given that the concepts of Truths, Beliefs, and Knowledge are positioned over
-an underlying structure of Propositions, the first concept to formalize is that
-of Propositions. We do not restrict Propositions to a specific structure but,
-as for Mankind and Nature, we refer to them as a collection.  We express the
-generality of a collection as a (propositional) signature over an abstract
-algebraic structure. In our formalization, we use mereotopology as algebraic
-structure (due to its generality, as discussed in Section~\ref{sec:problem})
-and the RCC to formalize concepts over a mereotopology.
+\begin{definition}{\bf Operational System --}\label{def:system}
+	An operational system (or a sub-system) is an agent (see Definition~\ref{def:agent})
+	such that
+	$s=\langle\rcc(\factRegion,\behaviorRegion),\rcc(\factRegion,\assertionRegion),\rcc(\behaviorRegion,\assertionRegion)\rangle$,
+	where $\factRegion,\behaviorRegion,\assertionRegion$, are regions of facts, behavior, and assertions respectively.
+\end{definition}
 
-We can now define the possible mereotopological relations between Belief and
-Knowledge (informally depicted in Figure~\ref{fig:knowledge-belief}) and then
-define the peculiarity of the believed difference between Mankind and Nature 
-(i.e. maliciousness as intent). 
-We first start from the formalization of Figure~\ref{fig:knowledge-belief}, as
-follows.
+As already presented in \autocite{Santaca2016abf} it follows that, defining
+a system (or an operational system) with 3 regions, there exist
+an upper-bound to the number of possible configuration of a system, defined by
+the possible relations between the different regions.
+For completeness we report in  stated in the next paragraph
+the calculation done in \autocite{Santaca2016abf}.
 
-\begin{definition}{\bf Information}\label{def:information}
-\end{definition}
-\begin{definition}{\bf Desire}\label{def:desire}
-\end{definition}
-\begin{definition}{\bf Intent}\label{def:intent}
-\end{definition}
+\Paragraph{Number of different configurations of a system}
+\fix{mr}{
+The general formula to calculate the number of different types of agents is
+$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
+between $n$ different sets, where $r^e$ is the number of permutation of $r$
+relations over $e$ elements with repetitions, with $e$ being the number of
+$k$-ary combinations of $n$ sets, $\binom{n}{k}$.
+%and the proof is given in the appendix (Section~\ref{app:proof}).
+In our case, $\binom{n}{k}=3$ since we consider $3$ sets
+($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
+considered in the RCC are binary.  Hence, using RCC5 (with five different
+spatial relations) over three sets, we can theoretically define up to 125
+different type of agents. However, only 54 of the 125 (as showed in
+\cite{improvingRCC} and derived by the composition table of RCC5 in
+Table~\ref{tab:5com}) combinations are topologically correct with respect to
+the definition of the relations of RCC5. Generalizing to all the RCCs, in
+Table~\ref{tab:relations} we calculate the number of different agents with
+respect to all the variations of RCC (i.e., with 3, 5 or 8 spatial relations).
+
+\begin{itemize}
+\item \emph{RCC3} --- theoretical: $3^3=27$,  correct: 15 
+\item \emph{RCC5} --- theoretical: $5^3=125$, correct: 54
+\item \emph{RCC8} --- theoretical: $8^3=512$, correct: 193
+\end{itemize}
+
+Hence, even if considering a different number of sets than the three
+$\assertionRegion$, $\behaviorRegion$ and $\factRegion$ exponentially affects
+the number of theoretical agents, the application of RCC downscales that number
+of a factor that ranges from 1.8 to 2.5. In addition, using RCC5 we consider
+3.6 times more (different) types of agents than RCC3, but using RCC8 would
+allow us to consider 3.5 times more different agents.}

### 2020-02-24 theory.tex
minor changes

@@
 the possible relations between the different regions.
-For completeness we report in  stated in the next paragraph
+For completeness we report in the next paragraph 
 the calculation done in \autocite{Santaca2016abf}.
@@
 \Paragraph{Number of different configurations of a system}
-\fix{mr}{
 The general formula to calculate the number of different types of agents is
@@
 $k$-ary combinations of $n$ sets, $\binom{n}{k}$.
-%and the proof is given in the appendix (Section~\ref{app:proof}).
 In our case, $\binom{n}{k}=3$ since we consider $3$ sets
-($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
+($\assertionRegion,\behaviorRegion,\factRegion$), and all the relations
 considered in the RCC are binary.  Hence, using RCC5 (with five different
@@
 different type of agents. However, only 54 of the 125 (as showed in
-\cite{improvingRCC} and derived by the composition table of RCC5 in
-Table~\ref{tab:5com}) combinations are topologically correct with respect to
-the definition of the relations of RCC5. Generalizing to all the RCCs, in
-Table~\ref{tab:relations} we calculate the number of different agents with
-respect to all the variations of RCC (i.e., with 3, 5 or 8 spatial relations).
+\cite{improvingRCC}) combinations are topologically correct with respect to
+the definition of the relations of RCC5. Generalizing to all the RCCs:
 
-\begin{itemize}
+\begin{itemize}%[nosep]
 \item \emph{RCC3} --- theoretical: $3^3=27$,  correct: 15 
@@
 3.6 times more (different) types of agents than RCC3, but using RCC8 would
-allow us to consider 3.5 times more different agents.}
+allow us to consider 3.5 times more different agents.

### 2020-02-26 formal_glossary.tex
move image
hacking process -> black-hat hacking process

@@
 
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\textwidth]{cwe-nvd-capec.pdf}
+	\caption{Cybersecurity online dictionaries and connections}
+	\label{fig:cwe-nvd-capec}
+\end{figure}
+
 The Common Attack Pattern Enumeration and
@@
 
-\begin{definition}{\bf Hacking Process}\label{def:hackingprocess}
+\begin{definition}{\bf Black-hat Hacking Process}\label{def:hackingprocess}
 An \emph{Attack} is the act of making operational an \emph{Exploitation Process} 
@@
 
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{cwe-nvd-capec.pdf}
-	\caption{Cybersecurity online dictionaries and connections}
-	\label{fig:cwe-nvd-capec}
-\end{figure}
-
-The formalization of the Exploitation and Hacking processes is the
+The formalization of the Exploitation and Black-hat Hacking processes is the
 formalization of two foundational cybersecurity concepts which lay the basis

### 2020-02-26 prototypes/agent_categorization/rcc5_finiteDomain_mereology.py
parallel computation - first attempt

@@ -1,6 +1,8 @@
 from z3 import *
 import scipy.special
 import itertools
+from multiprocessing import Process
+from multiprocessing import Queue 
 
 ################################
 # this method returns the rcc5 satisfiability table 
@@ -99,65 +101,65 @@ def rcc_five_sat_table(x,y,z,EQ,PP,PO,PPi,DR):
 
 ################################
 # since we unfold the quantifiers we have to calculate
-# the (minimal) number of subsets of A, B, and F that allows
+# the (minimal) number of subregions of A, B, and F that allows
 # Z3 to find a model that satisfies the agent configuration
 # 
-# For example, if we have A,B,F and no subsets of those 3 sets 
+# For example, if we have A,B,F and no subregions of those 3 regions 
 # there is no model that satisfies PO(A,B) /\ PO(B,F) /\ PO(A,F)
-# since PO(A,B) <-> PP(A,B) if A has no subsets 
+# since PO(A,B) <-> PP(A,B) if A has no subregions 
 #
 # requires scipy for binomial calculation: pip install scipy
 # otherwise download wheel from https://pypi.python.org/pypi/scipy
 # and install it with pip install <downloaded-file>
 ################################
-def calculate_sets(sets, Set):
-    set_of_subsets=[]
-    for k in range(1,len(sets)):
+def calculate_regions(regions, Region):
+    region_of_subregions=[]
+    for k in range(1,len(regions)):
         counter=0
-        for i in range(int(scipy.special.binom(len(sets), (k+1)))):
-            subset="subset_%s_%d"%((k+1), counter)
-            set_of_subsets.append(Const(subset, Set))
-            #print("%d - (%d %d)=%d"%(i,len(sets),(k+1),int(scipy.special.binom(len(sets), (k+1)))))
+        for i in range(int(scipy.special.binom(len(regions), (k+1)))):
+            subregion="subregion_%s_%d"%((k+1), counter)
+            region_of_subregions.append(Const(subregion, Region))
+            #print("%d - (%d %d)=%d"%(i,len(regions),(k+1),int(scipy.special.binom(len(regions), (k+1)))))
             counter+=1
         
-    return set_of_subsets
+    return region_of_subregions
 
-def topology(solver, sets_and_subsets, P):
[…]

### 2020-02-26 theory.tex

@@
 
-\subsection{Abstract Systems}\label{sec:system}
+\subsection{Abstract System State}\label{sec:systemstate}
 The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
@@
 \begin{itemize}
-	\item In \autocite{Georgeoff1995bdi}, the BDI (Belief, Desire, Intent)
-		paradigm is explained; where an agent can be defined over those
-		three constituents.
 	\item In \autocite{Hintikka1962knowledge}, Hintikka describes the
@@
 variable, which admits the two alternatives True/False, its information is
-defined as excluding either to be True or False and then, as we'll discuss
-later on, Believed to be the opposite. Similarly, when reasoning on a
-mereotopological region w.r.t. the RCC (e.g. RCC3 admits two regions to be
-equal, disjoint, or overlapping) and information on this region should exclude
-some of the alternatives. Finally, in \autocite{Empiricus1990Pyrrhonism}
-the author states that the logical criterium can be judged according to something.
+defined as Believed to be True/False and not Believed to be the opposite. 
+%Similarly, when reasoning on a
+%mereotopological region w.r.t. the RCC (e.g. RCC3 admits two regions to be
+%equal, disjoint, or overlapping) information on this region should exclude
+%some of the alternatives. 
+Finally, in \autocite{Empiricus1990Pyrrhonism}
+the author states that the logical criterium can be judged according to a frame of reference.
 For our argument, as in \autocite{Santaca2016abf}, we define this concept
@@
 of knowledge and belief. For our argument, the only objective of
-Information is to interact with the behavior of a system or an agent. 
-Due to the definition of Information and its relation to the probabilistic
-correlation between information and truth/reality, we must consider
+Information is to exchange beliefs between agents.
+Due to the definition of Information and then with its relation to the probabilistic
+correlation to truth/reality, we consider
 information in relation with an agent's beliefs. Similarly, we consider
-beliefs to define the behavior of an agent or a system. On the other hand,
+beliefs to define the actual behavior of an agent or a system. On the other hand,
 Knowledge drives the nominal behavior of an agent or a system. This view
@@
 
-\begin{definition}{\bf System --}\label{def:system}
-	A system (or a sub-system) is an agent (see Definition~\ref{def:agent})
+\begin{definition}{\bf System State Space --}\label{def:system}
+	A state of a system (or a sub-system) $s$ is identified with 
+	an instance of an agent state (see Definition~\ref{def:agent})
 	such that
@@
 \end{definition}
-A system is then defined as a superposition of multiple states where knowledge,
+A system state space is then defined as a superposition of multiple system states where knowledge,
 beliefs, and information are related between each other (in a mereotopological
-space) through the RCC. A \emph{system instance} is a specific definition of
-the $\rcc$ relations between the three regions. Therefore, a set of 
-system instances defines a set of choices for those $\rcc$ relations.
+space) through the RCC. A \emph{system state} (as an instance of the system state space)
+is a specific configuration of
+the $\rcc$ relations between the three regions. Therefore, a collection of 
+system states defines a collection of choices for those $\rcc$ relations.
 
@@
 Given an abstract collection of Agents $Ag$, and the modal operator
-	$\knows{a}{}$ (where $a\in Ag$), Knowledge is defined as a collection
+	$\knows{a}{}$ (where $a\in Ag$), Knowledge is defined as a region 
 	of predicates known by an agent $\knowledge{a}=\bigcup_\Phi \knows{a}{\varphi}$ 
@@
 	Given an abstract collection of Agents $Ag$, and the modal operator
-	$\believe{a}{}$ (where $a\in Ag$), Belief is defined as a collection
+	$\believe{a}{}$ (where $a\in Ag$), Belief is defined as a region 
 	of predicates believed by an agent $\belief{a}=\bigcup_\Phi \believe{a}{\varphi}$
@@
 	Given an abstract collection of Agents $Ag$, and the modal operator
-	$\informs{a}{}$ (where $a\in Ag$), Information is defined as a collection
+	$\informs{a}{}$ (where $a\in Ag$), Information is defined as a region 
 	of beliefs asserted by an agents $a$.
@@
 	$\information{a}=\bigcup_\Phi \informs{a}{}{\varphi}$
-	(where $\Phi$ is the collection of all the propositions believed and asserted by $a$).
+	(where $\Phi$ is the region of all the propositions believed and asserted by $a$).
 	Given a proposition $P$, we extend the semantics of the Causality structure with:
@@
 		\item[$(\interpretation18)$] $\world\models\informs{a}{P}$ iff
-			$\world\models\neg\knows{a}{\neg P}$ (i.e. the agent $a$ considers $P$ possible) 
-			and $\world'\models P$ for all
+			$\world\models\belief{a}{P}$ (i.e. the agent $a$ considers $P$ possible), 
+			$\world\models\neg\belief{a}{\neg P}$, and $\world'\models P$ for all
 			$\world'$ such that $\world\modalrelation\world'$.
@@
 if there is a sender and a recipient.  Similarly, the portion of Beliefs we
-consider is the one that builds the behavior of an agent and we consider a set
-of axiomatic Facts instead of Knowledge.
+consider is the one that builds the behavior of an agent (strategy rules
+\footnote{``The logical
+structure of information is one of the most basic and one of the most basic
+and one of the simplest thing in the wide and wonderful world of logical
+analysis. This point can be put in a deeper perspective. A distinction [\ldots] ought to
+be made [--] between two kinds of rules (or principles) in any strategic activity like
+knowledge seeking. On the one hand you have the rules that define the game, e.g. how
+chessmen are moved on a board. The can be called \emph{definitory} rules.
+They must be distinguished from rules [\ldots] that deal with what is better and what is worse
+in the game in question. Definitory rules do not say anything about this subject. Rules
+which do can be called \emph{strategic rules}'' -- Hintikka in \autocite{Hintikka1993Information}}) 
+and we consider a set
+of axiomatic Facts (definitory rules) instead of considering the more general 
+epistemic definition of Knowledge.
 
-\begin{definition}{\bf Assertion}\label{def:assertion}
-\end{definition}
-\begin{definition}{\bf Behavior}\label{def:behavior}
-\end{definition}
-\begin{definition}{\bf Fact}\label{def:fact}
+\begin{definition}{\bf Assertion -- }\label{def:assertion}
+	An assertion is an intended transfer of beliefs between
+	two agents $a$ and $b$ such that
+	\begin{enumerate}[noitemsep]
+		\item[$(\interpretation19)$] $\world\models\assert{a}{b}{P}$ iff
+			$\world\models\belief{a}{P}$, 
+			$\world'\models\rcc(\belief{a}{P},\belief{b}{P})$, 
+			$\world'\models\belief{a}{\belief{b}{P}}$
+			and $\world'\models P$ for all
+			$\world'$ such that $\world\modalrelation\world'$.
+	\end{enumerate}
 \end{definition}
 
-\begin{definition}{\bf Operational System --}\label{def:system}
+\begin{definition}{\bf Fact -- }\label{def:fact}
+Facts are a sub-region of Knowledge $\factRegion = \bigcup_\Phi \knows{a}{\varphi}$
+	that predicate in a factive way, e.g.
+	if $a$ knows that $P$ then $P$, and the region of facts is monotone (no revision).
+	\begin{enumerate}[noitemsep]
+		\item[$(\interpretation20)$] $\world\models\fact{a}{P}$ iff
+			$\world \models P$ and 
+		.
+	\end{enumerate}
+\end{definition}
+
+\begin{definition}{\bf Behavior -- }\label{def:behavior}
+	The behavior of an agent is determined by its beliefs. 
+	\begin{enumerate}[noitemsep]
+		\item[$(\interpretation21)$] $\world\models\assert{a}{b}{P}$ iff
+			$\world\models\belief{a}{P}$, 
+			$\world'\models\rcc(\belief{a}{P},\belief{b}{P})$, and $\world'\models P$ for all
+			$\world'$ such that $\world\modalrelation\world'$.
+	\end{enumerate}
+\end{definition}
+
+\begin{definition}{\bf Operational System State --}\label{def:system}
 	An operational system (or a sub-system) is an agent (see Definition~\ref{def:agent})
@@
 allow us to consider 3.5 times more different agents.
+
+\subsection{System Security Evolution}\label{sec:systemevoution}
+\fix{mr}{
+Facts should be correlated to the causality principle. Therefore a
+specification in opnet should be translated into vmt and and by bounded model
+checking with nuxmv identify a path towards an undesired state, if that fails
+use abf to mutate the first state into one of the possible according to abf and
+re-run nuxmv.
+
+Knowledge is the theory itself, an instance is the facts of a design of it. A
+testing process tests facts against knowledge as much as we test a cps design
+chiices against its specification.  Information carries the knowledge of the
+negation of the opposite of the information. Therefore an assertion carries the
+belief (as a result of the behavior execution) of not the opposite. But, as
+Hintikka States in "popper" our previous statement is true if the probability
+distribution of the information is representative wrt the causality structure
+(i.e. knowledge and facts, see email before). So, based on our theory, the
+assertions should have a number between 3 to 5 possible negated meaning, one
+per each rcc if reference and not just a negation of the information. Those
+negation representative of an insecure configuration on the communication of
+the system. In turn, those configurations should categorize all the CWE.  The
+fact that weaknesses are generated by mankind but not exploited makes malicious
+attacks confined to the hackers (and black-hat hackers). Fix: move black-hat
+hackers as subregion of mankind with malicious intent. Describe the intent as a
+desire of reaching a breach of a cybersecurity property over an asset (agent of
+the system), instead of searching for a specific attack. The tool would then
+translate the system from opnet to abf formula. The abf tool generates all the
+sat state of the system fix: add state to all def before this text. And we
+shall define the causality relation between the different states generated by
+the abf tool as chosen nondeterministically from the possible sat states by a
+model checker such as nuxmv.
+
+By calculating sat instead of validity with abf tool and then adding it to the
+causality structure of a system fact description, e.g. in nuxmv; we can
+calculate a path for reachability of an insecurity property violated on an
+agent which represents an asset. Whenever nuxmv fails, we can add a new state
+and re-run the process. Those guesses should be better managed by a trained
+v-ai since the inference by induction from the partial information of the
+system in which we represent the assertions as negation of a beliefs and
+without rcc do not carry enough information for a Reasoner that doesn't
+implemented rcc.
+}
+
+\subsection{Security Properties and Assets}\label{sec:properties}
+In \autocite{Georgeoff1995bdi}, the BDI (Belief, Desire, Intent)
+paradigm is explained; where an agent can be defined over those
+three constituents.

### 2020-02-26 engineering.tex
add section on engineering process

--- 2020-02-19 v-research_experiment-alpha.tex
@@
-\documentclass{article}
-\usepackage{mathtools}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{enumitem}
-\usepackage{multicol}
-\usepackage[toc,page]{appendix}
-\usepackage[gen]{eurosym}
-\usepackage{tabularx}
-\usepackage{epigraph}
-\usepackage{textcomp}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{csquotes}% Recommended
-\usepackage{rotating}
-\usepackage[style=numeric]{biblatex}
-% \bibliography{<mybibfile>}% ONLY selects .bib file; syntax for version <= 1.1b
-\addbibresource{literature.bib}% Syntax for version >= 1.2
-
-\usepackage{graphicx}
-\graphicspath{{./images/}}
-
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\lhead{}
-\chead{}
-\rhead{}
-\lfoot{}
-\cfoot{\thepage}
-\rfoot{Property of V-Research S.r.l.}
-\renewcommand\headrulewidth{0pt}
-\renewcommand\footrulewidth{0pt}
-
-\usepackage[colorinlistoftodos]{todonotes}
-
-\newcommand{\Paragraph}[1]{\smallskip\noindent{\bf #1.}}
-\newcommand{\fixnote}[2]{\textbf{\color{red}{FIX}}\footnote{{\bf #1:} #2}}
-\newcommand{\fix}[2]{{\color{red} {\bf tofix:} #2}}
-% \renewcommand{\fixnote}[2]{}
-% \renewcommand{\fix}[2]{}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\theoremstyle{corollary}
-\newtheorem{corollary}{Corollary}[section]
-\theoremstyle{lemma}
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{theorem}
-\newtheorem{theorem}{Theorem}[section]
-\theoremstyle{theorem}
-\newtheorem{example}{Example}
-
-\usepackage{authblk}
-\date{}                     %% if you don't need date to appear
-\setcounter{Maxaffil}{0}
-\renewcommand\Affilfont{\itshape\small}
-
-% for fancy table with rotated column titles
-\usepackage{adjustbox}
-\newcolumntype{R}[2]{%
-    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
-    l%
-    <{\egroup}%
-}
-\newcommand*\rot{\multicolumn{1}{R{45}{0.01em}}}% no optional argument here, please!
-\newcommand*\rota{\multicolumn{1}{R{90}{0.01em}}}% no optional argument here, please!
-\def\hb{\hbox to 10.7 cm{}}
-
-% circles
-\usepackage{wasysym}
-\newcommand{\Tdot}{$\CIRCLE$}
-\newcommand{\Thdot}{$\LEFTcircle$}
-\newcommand{\Twdot}{$\Circle$}
-
-% formulas
-\newcommand{\kframe}{\mathbf{K}}
-\newcommand{\bframe}{\mathbf{B_K}}
-\newcommand{\possibleworlds}{G}
-\newcommand{\modalrelation}{R}
-\newcommand{\actualworld}{\omega*}
-\newcommand{\world}{\omega}
-\newcommand{\World}{\Omega}
-\newcommand{\kmodel}{M}
-\newcommand{\interpretation}{\sigma}
-\newcommand{\knowledge}[1]{\mathbb{K}_{#1}}
-\newcommand{\knows}[2]{K_{#1}#2}
-\newcommand{\belief}[1]{\mathbb{B}_{#1}}
-\newcommand{\believe}[2]{B_{#1}#2}
-\newcommand{\region}{\chi}
-\newcommand{\Region}{\mathrm{X}}
-\newcommand{\agentuniverse}{\mathcal{A}}
-\newcommand{\agent}{a}
-\newcommand{\follang}{\mathcal{L}}
-\newcommand{\weakness}{w}
-\newcommand{\Weakness}{W}
-\newcommand{\vulnerability}{v}
-\newcommand{\Vulnerability}{V}
-%rcc5
-\newcommand{\rcc}{rcc}
-\newcommand{\connects}[2]{C(#1,#2)}
-\newcommand{\disconnected}[2]{\neg C(#1,#2)}
-\newcommand{\partof}[2]{P(#1,#2)}
-\newcommand{\overlaps}[2]{O(#1,#2)}
-\newcommand{\eq}[2]{EQ(#1,#2)}
-\newcommand{\pp}[2]{PP(#1,#2)}
-\newcommand{\po}[2]{PO(#1,#2)}
-\newcommand{\ppi}[2]{PPi(#1,#2)}
-\newcommand{\dr}[2]{DR(#1,#2)}
-
-\makeindex
-
-\begin{document}
-
-\title{On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}}
-\author[1]{Francesco Beltramini}
-\author[1]{Marco Rocchetto}
-\author[2]{Luca Vigan\`o}
-\affil[1]{V-Research, Verona, Italy}
-\affil[2]{King's College London, UK}
-
-\maketitle
-
-\begin{abstract}
-	The objective of this research is to develop of a theory that defines
-	(all and only) the possible insecurity and security configurations of
-	any abstract system. The theory is structured upon other theories that 
-	defines how a component of a system can be abstracted into an agent, defining how
-	agents can be formalized (both syntactically and semantically) to
-	describe an abstract system, such as a graph. Some of these theories
-	(e.g. used for the semantic
-	definition of the abstract system) are the epistemological definition
-	of knowledge, the Belief-Desire-Intent and the Assertion-Belief-Fact 
-	framework of reference, 
-	mereology, and topological structure. We argue that a mereology is the
-	most appropriate abstract underlying structure, do to its generality,
-	for defining the expressiveness of the system abstraction.  Furthermore, a
-	mereology allows us to define an ontology rather than a taxonomy.  We
-	also correlate different abstractions of the system to the TRL and the
-	engineering V-model.
-
-	We implemented a formal theory (of axioms) of a mereotopology, and of
-	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
-	uses the Z3 SMT solver. The results show that a single component (i.e.
-	agent) of an abstract system has a definite number of  different
-	insecurity configurations (e.g. 53 using RCC5 over a topological
-	structure) and only 1 secure (i.e. expected) configurations. The
-	configurations are reported as models satisfying the abstract system
-	semantics.
-
-	We considered the philosophical definition of truth behind our
-	approach, rejecting ``proof'' by induction from partial empirical evidences.
-	Our theory can be applied to system engineering and we show a concrete
-	application of our theory to the risk assessment of an ad-hoc system.
-	Finally, we provide a number of ideas to support the engineering of
-	secure systems (e.g. purely cyber or cyber-physical).
-\end{abstract}
-\newpage
-
-\section{Introduction}\label{sec:intro}
-\epigraph{Humanum est errare}{{\itshape Seneca the Elder}}
-The European Commission states in\autocite{EU2019market} that: ``Cybersecurity
-is one of the priority areas [\ldots] of the Commission initiative on ICT
-Standards, which is part of the Digitising European
-Industry\autocite{EU2019standard} strategy launched on 19 April 2016.  The aim
-is to identify the essential ICT standards and present measures to accelerate
-their development in support of digital innovations across the economy''. The
-same document (i.e.\autocite{EU2019market}) states that ``The EU will invest up
-to \euro450 million [\ldots], under its research and innovation programme
-Horizon 2020''. The EU, in 2016 published a press release\autocite{EU2016press}
-in which they present a strategy to invest \euro1.8 \emph{billion} to
-``increase measures to address cyber threats''. The EU is not the only investor
-in cybersecurity, most of the developed countries and several companies are
-investing enormous amount of money towards various aspects of cybersecurity
-(e.g. The US vulnerability databases\autocite{NIST2020NVD} maintained by the
-National Institute of Standards and Technologies, i.e. NIST, of the US
-Department of Commerce).
-
-The cybersecurity industry is growing fast, e.g. as reported
-in\autocite{Nasdaq2018market}. For example, in\autocite{Forbes2017market},
-published by the Forbes, is stated that \euro5.3 billion of funding where
-poured by venture capitalist into cybersecurity companies in 2018. The Forbes,
-in the same article, also highlights another peculiar (as seemly contradictory)
-trend: ``[\ldots] during the same time period, the number of cybersecurity
-breaches increased exponentially''. The data reported by the NIST 
-through the official CPE (Common Platform Enumeration) Dictionary Statistics 
-on the NVD websites in \autocite{NIST2020CPEstatistics}, show that in 2016 the
-number of reported vulnerabilities reported where around $~6000$ while in 2019
-the number of vulnerabilities was above $16000$.  The scientific community also
-reports similar findings.  In fact, in\autocite{Herley2009so}, Cormac Herley
-(Microsoft Research) shows how basic cybersecurity principles (such as the
-confidentiality benefit over the clear text for passwords typed into forms,
-e.g. for logins in websites) are not fully understood or shared between the
-cybersecurity research community\autocite{Nielsen2009stop}.  The lack of
-understanding of basic security principle, the inverse proportionality between
-investments in cybersecurity and the number of reported vulnerabilities year
-after year, can be linked to the lack of a foundational theory on
-cybersecurity, as already highlighted by Cormac Herley
-in\autocite{Herley2016unfalsifiability}.
-
-In this article, we give the first scientific theory (to the best of our knowledge)
-on security.
-
-\Paragraph{Structure} In Section~\ref{sec:problem} we define and formalize the problem statement. 
-In Section~\ref{sec:theory} we outline our security theory, and in Section~\ref{sec:prototest}
-we describe the implementation of the theory and some empirical tests of the theory.
-Finally, in Section~\ref{sec:related} we conclude the paper with an overview of the related work.
-
-\section{Problem Statement}\label{sec:problem}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.6\textwidth]{protocol-example.pdf}
-	\caption{Abstraction of an ad-hoc esemplificative protocol execution}
-	\label{fig:protocol-example}
-\end{figure}
-
-In\autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
-``an asymmetry in computer security'', which he defines as follows: ``Things
-can be declared insecure by observation, but not the reverse. There is no
-observation that allows us to declare an arbitrary system or technique
-secure''. Herley then uses this argument to show that ``claims that any measure
-is necessary for security are empirically unfalsifiable''. Given that, any
-theory which is not falsifiable by an empirical experiment is well
-known\footnote{``A theory which is not refutable by any conceivable event is
-nonscientific. Irrefutability is not a virtue of a theory (as people often
-think) but a vice.'' -- Karl Popper, Conjectures and
-Refutations\autocite{popper1962conjectures}} to be nonscientific (i.e.
-unfalsifiability is a fallacy of a theory), Herley concludes that there is no
-scientific theory on cybersecurity; which means that cybersecurity lays in the
-realm of pseudo-sciences\autocite{Herley2016usenixvideo}.  Herley, e.g.
-in\autocite{Herley2017justifying}, discusses the implications of a
-nonscientific approach to cybersecurity, and highlights the tremendous impact on
-all the scientific research and engineering of systems; leading often to
-terrorism and wars, and wasting of resources in useless protections or
-overspending.  While the criticism is investigated
-in\autocite{Herley2016unfalsifiability}, no solution is provided.  On the
-contrary, the goal of this work \emph{is} to lay the foundations of a
-scientific cybersecurity theory. Furthermore, in Section~\ref{sec:sicurezza},
-we consider the problem raised by Herley not confined to ``computer security''
-but to any abstract system (so that our theory may hold for any sound
-implementation such as networks, mechanical, cyber, or cyber-physical system,
-or even a single computer or a single device such as an hard-drive).  There is
-also an apparent inconsistency in\autocite{Herley2016unfalsifiability} that we
-seek to clarify before following (as we agree) the scientific path draw by
-Herley: cybersecurity is defined as an abstract property in many formal
-approaches to the investigation of the security of systems, and the security of the design of a
-formally verified protocol is indeed falsifiable.  For example, in the protocol
-verification community, security is often defined as a formalization of the
-high-level properties confidentiality, integrity, and availability. The problem
-in such approaches is not the definition of what cybersecurity is, but the use
-of theories (such as the Dolev-Yao attacker model\footnote{For the sake of
-simplicity, the Dolev-Yao attacker can be considered as an abstraction of an
-active attacker who controls the network but cannot break
-cryptography.}\autocite{Dolev1983security})
-that only applies to specific instances (often called scenarios) and
-abstraction of the protocol. This, in turn, creates a false sense of security
-since requires assumptions on the abstraction of the system of which security
-is verified. As an example, for the formal security verification of the system
-in Figure~\ref{fig:protocol-example}, a formalized scenario needs to be defined
-by a modeler who chooses (among others): (i) a scope of the formalization (e.g.
-excluding the server that distributes the public key is often done when
-verifying the security of authentication protocols), (ii) the number of
-sessions (even tough some approaches do reason on an infinite number of
-sessions such as\autocite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
-the peers (e.g.  in the ASLan++ language\autocite{Oheimb2010aslan++}), and (iv)
-the abstraction of the cryptographic primitives (e.g.  ProVerif vs
-CryptoVerif\autocite{Blanchet2017symbolic}).  Some of the choices will
-completely change the results of the formal verification of the system. For
-example, under the perfect cryptography assumption\footnote{As defined
-in\autocite{Rocchetto2016cpdy}: ``In the so called perfect cryptography
-assumption, the security encryption scheme is suppose to be perfect, without
-any exploitable flaw, and so the only way for the attacker to decrypt a message
-is by using the proper key. That assumption is widely accepted in the security
-protocol community, and most of the formal reasoning tools for the analysis of
-security protocols abstract away the mathematical and implementation details of
-the encryption
-scheme\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
-and assuming that no violation to any security property is done after message
-I); in Figure~\ref{fig:protocol-example}, the freedom of choosing the scope
-determines that the flaws related to the dishonest impersonation of the Server
-may or may not be considered in the verification process.  This choice has
-tremendous impact on the focus and findings of the verification of the security
-of the protocol.  While this may seem to turn upon minutiae and foreseeable,
-this highlights the false sense of security that may derive from a
-non-scientific theory of system security\footnote{``To the superficial
-observer, the analysis of these forms seems to turn upon minutiae. It does in
-fact deal with minutiae, but they are of the same order as those dealt with in
-microscopic anatomy.'' -- Karl Marx, Capital Volume 1, 1867}.
-
-\subsection{Sicurezza: Safety and Security}\label{sec:sicurezza}
-In most of the natural languages, and in Italian too, the concepts of safety
-and security are not syntactically differentiated and both terms (safety and
-security) are expressed by the same word, e.g. sicurezza in Italian.  A
-semantic distinction between safety and security is correlated to a
-belief\footnote{A belief has to be intended as a proposition which is supposed
-to be true by the majority of humans in our society without a scientific
-underlying theory but based on partial empirical evidences or inductive proofs.} that
-safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
-natural environment (e.g. natural events such as wearing of hardware
-components) while security deals with \emph{incidents} posed by mankind (e.g.
-attackers and bugs).  The fundamental difference between nature and mankind (and,
-in turn, between safety and cybersecurity) is believed to be on the different
-intents\footnote{``The belief–desire–intention software model (BDI) is a
-software model developed for programming intelligent
-agents.''\autocite{wiki-bdi}. In the BDI model, the intents represents the
-deliberative state of an agent which determines the choice of that agent on
-what to do.} (accidents are unfortunate while incidents are not) of the causes
-that generates the threat; namely, nature is believed not to have malicious
-intents (but unfortunate causes-effects) while threats generated by mankind are
-believed to be malicious\footnote{Of course, logical flaws or bugs may be
-introduced by other means (e.g. ignorance) without explicit malicious intents,
-but the exploitation of those flaws is considered (for now, and detailed
-afterwards in the article) malicious, and then we consider any vulnerability to
-be malicious (without loss of generality) even if due to the lack of skills.}.
-An overview on the aforementioned aspects of safety and security is depicted in
-Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
-the terms that structure our current understanding of safety and security. 
-\begin{itemize}
-	\item \emph{Mankind} ``refers collectively to humans''
-\autocite{wiki-mankind}, while the concept of \emph{Nature} is
-		related ``to the intrinsic characteristics that plants,
-		animals, and other features of the world develop of their own
-		accord'' (e.g. the physical universe)\autocite{wiki-nature}. 
-		\begin{itemize}
-			\item So far, we have used several terms to refer to an
-				\emph{attacker}, i.e. threat agent or threat source,
-				considering those terms to be semantically
-				equivalent.  This ``shallowness'' raise form the
-				necessity of properly citing the different sources, but,
-				in the reminder of this paper, we consider the
-				Causality principle to be the \emph{threat
-				source}, Nature or Mankind to be the
-				\emph{threat agents} and an \emph{attacker} as
-				a specific malicious threat agent which materialize a
-				threat.
-		\end{itemize}
-	\item \emph{Vulnerability}\footnote{The term vulnerability is not
-		present in the Encyclopedia of Cryptography and Security, while
-		it is used in 12 entries (such as in the definition of
-		``penetration testing''\autocite{caddy2005pentest})
-		highlighting how commonly this word is used without a proper
-		supporting semantics}, as defined in\autocite{cnssi20104009}
-		(and adopted in\autocite{nist2013800-53}), is ``weakness in an
-		information system, system security procedures, internal
-		controls, or implementation that could be exploited by a threat
-		source''. On the one hand, the definition is broad to enclose
-		as much causes (that generates a vulnerability) as possible; on
-		the other hand, it derives from empirical evidences (which
-		should be considered beliefs\footnote{``For this view, that
-		\emph{That Which Is Not} exists, can never predominate. You
-		must debar your thought from this way of search, nor let
-		ordinary experience in its variety force you along this way,
-		(namely, that of allowing) the eye, sightless as it is, and the
-		ear, full of sound, and the tongue, to rule; but (you must)
-		judge by means of the Reason (Logos) the much-contested proof
-		which is expounded by me.'' -- Parmenides of Elea, On Nature
-		(circa 500 B.C.), fragments B7.1–8.2
-		\autocite{Hakim2016philosophy}} since they are partial results in nature) 
-		while a vulnerability should
-		be defined in a way that is empirically falsifiable. This means
-		that the term vulnerability should have a complete and sound
-		definition, so that no other causes (e.g.  other sources) but
-		the ones in the definition are responsible for a vulnerability.
-		Furthermore, the term ``threat sources'' used in the definition
-		in\autocite{cnssi20104009} may be identified with both Nature
-		and Mankind, not differentiating between safety and security.
-		In Definition~\ref{def:vulnerability}, we provide a formal
-		theory of vulnerability (so that the scientific community can
-		identify tests for the completeness and soundness of the
-		definition itself).
-\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{safety-security.pdf}
-	\caption{Overview security and safety keywords}
-	\label{fig:safety-security}
-\end{figure}
-
-Most of the safety-preserving principles in the field of engineering of
-safety-critical cyber-physical systems (such as elevators and aircraft), upon
-which safety requirements (e.g. in standards such as the IEC 61508 or 61511\autocite{IEC201761511}) are defined, have
-been defined following empirical tests and measurements. While reasoning by
-induction based on the empirical observation should be avoided, since it may
-easily lead to false beliefs instead of scientific theories, this approach is
-often justified by the supposed impossibility of defining a theory that
-correctly predicts failures which, in turn, pose hazards to a system. 
-To the best of our knowledge, and supported by\autocite{Herley2016unfalsifiability}, the
-correlation between predictability of environment and believed unpredictability
-of attackers (i.e. a malicious
-environment) has not been correlated to a theory on cybersecurity. 
-Therefore, inductive research efforts in
-predicting malicious effects are accepted (and published) in scientific
-conferences (e.g.~\autocite{Rocchetto2014CSRF}). A failure of a wire due to
-environment (e.g. due to humidity, dust, heat \&c) is defined from empirical
-evidences and processes have been standardized to test qualities of hardware components
-%convinced by the induction principle. What is already a
-%weak argument in general, 
-This process completely breaks down when a malicious environment (i.e. an attacker)
-is considered instead of the (supposedly honest and predictable)
-natural environment. Therefore, the same approach that is in use for safety,
-seems not to be applicable to test security.
-
-Going back to Figure~\ref{fig:safety-security}, a vulnerability does not
-necessarily become a threat for the system, unless exploited ``through a
-channel that allows the violation of the security policy
-[\ldots]''\autocite{cnssi20104009} (e.g. a software or procedure) that takes
-advantage of the vulnerability causing an \emph{attack} to the system, which may
-result in several correlated incidents and threats.  The process of
-exploitation of a defect as a vulnerability is reported in
-Figure~\ref{fig:safety-security} such that the difference between exploit and failure,
-and attack and accident is to be found just in the maliciousness of the intents
-that causes this process (i.e. excluding the intent, the terms are just syntactic transformation from a vulnerability to defect, from
-accident to incident). In the following, we conclude the informal definition of
-the terms that we used in this section and in Figure~\ref{fig:safety-security}.
-
-\begin{itemize}
-	\item \emph{Causality} refers to the causality principle; defined
-		in\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
-		connection of phenomena through which one thing (the cause)
-		under certain conditions gives rise to, causes something else
-		(the effect). The essence of causality is the generation and
-		determination of one phenomenon by another. In this respect
-		causality differs from various other kinds of connection, for
-		example, the simple temporal sequence of phenomena, of the
-		regularities of accompanying processes''.
-	\item An \emph{Exploit}\footnote{We note that the term exploit is only
-		used as a verb in\autocite{ISO2009information}} is ``An exploit
-		(from the English verb to exploit, meaning to use something to
-		one’s own advantage) is a piece of software, a chunk of data,
-		or a sequence of commands that takes advantage of a bug or
-		vulnerability to cause unintended or unanticipated behavior to
-		occur on computer software, hardware, or something electronic
-		(usually computerized).''\autocite{wiki-exploit}.
-	\item An \emph{Attack}, as defined by the International Standard
-		ISO/IEC 27000 is an ``attempt to destroy, expose, alter,
-		disable, steal or gain unauthorized access to or make
-		unauthorized use of an asset''; where an \emph{Asset} is
-		``anything that has value to the organization''. We note that for
-		the purpose of this article, we do not want to focus on a specific
-		organization or business to define asset but, in general, on any 
-		abstract organization (e.g. a company or a society).
-		We do not consider ethical hackers as attacking a system. 
-		In fact, we consider the term \emph{hack} as
-		non-malicious (see Hacker\autocite{Stallman2002hacker}).
-	\item A \emph{Threat}, as defined in\autocite{cnssi20104009}, is ``Any
-		circumstance or event with the potential to adversely impact
-		organizational operations (including mission, functions, image,
-		or reputation), organizational assets, individuals, other
-		organizations, or the Nation through an information system via
-		unauthorized access, destruction, disclosure, modification of
-		information, and/or denial of service''.
-	\item \emph{Defect}, ``anything that renders the product not reasonably
-		safe''\autocite{Robinson2019writing} (i.e. a characteristic of
-		an object which hinders its proper usability).
-	\item \emph{Failure}, as defined in\autocite{Merriam2020failure} as ``a state of
-		inability to perform a normal function''. The term is
-		structured and detailed in
-		\autocite{cnssi20104009,iet2017glossary} but relying on an
-		abstract notion of failure without a specific definition.
-	\item \emph{Hazard}, ``a potential source of
-		harm''\autocite{iet2017glossary}.
-\end{itemize}
-
-\subsection{Glossary -- A Formalization}\label{sec:glossary}
-We now define a formalizations of the concepts described in
-Section~\ref{sec:problem} and depicted in Figure~\ref{fig:safety-security}. We
-base our formalization on first order logic (FOL) with the standard
-truth-value semantics. The choice of this logic is required by the 
-semantics of the concepts formalized afterwards in this section.
-
-We consider a first-order language $\follang$ over a
-signature $\Sigma_P$ where $P,P',\ldots,P^n$ represent terms, and
-$\varphi,\varphi',\ldots,\varphi^n$ represent formulas.
-The syntax is defined as follows.
-\begin{displaymath}
-	\varphi := P~|~\neg\varphi~|~\varphi\wedge\varphi~|~\varphi\vee\varphi~|~\varphi\Rightarrow\varphi'~|~\forall P.~\varphi~|~\exists P.~\varphi
-\end{displaymath}
-where $\wedge$, $\neg$, $\vee$, and $\Rightarrow$ are connectives representing
-conjunction, negation, disjunction, and (material) implication respectively; while
-$\forall$ and $\exists$ represent the standard universal and existential
-quantifiers (resp.). Finally, the symbol ``$.$'' is just syntactic sugar. 
-We consider $\interpretation\subset\Phi\times\{\top,\bot\}$ as the interpretation
-function, where $\Phi$ is any collection of sentences in $\follang$ 
-and $\top$ and $\bot$ represent the concepts of ``Tautology/True''
-and ``Contradiction/False'' respectively.
-
-Following Figure~\ref{fig:safety-security}, we start by formalizing the outermost term: Causality.
-
-\subsubsection{Causality Principle}\label{sec:causality}
-We formalize the Causality Principle starting from a K modal
-logic\autocite{Garson2018modal} (i.e. without restrictions on the causality
-relation between worlds). The standard definition of K modal logic is given in
-Definition~\ref{def:modallogic} in terms of an interpretation function (which
-we named $\sigma$ with a slight abuse of notation) defined in
-Definition~\ref{def:modalinterpretation}. 
-\begin{definition}{\bf K Modal Logic --}\label{def:modallogic}
-A K-frame is a frame $ \kframe=\langle \possibleworlds,\modalrelation \rangle $
-	in the K modal logic where $ \modalrelation $ is the binary relation
-	(i.e. a set of ordered pairs) between possible worlds
-	$\modalrelation\subseteq\possibleworlds\times\possibleworlds$, where
-	$\possibleworlds$ represents the possible worlds, and
-	$\possibleworlds\neq\emptyset$. An actual world
-	$\actualworld\in\possibleworlds$ is assumed. For any proposition $P$,
-	an interpretation function $\interpretation(\world,P)$ returns the
-	truth value of $P$; e.g. $\interpretation(\world,P)=\top$ means that
-	$P$ holds in $\world$. A model is defined as the tuple
-	$\kmodel=\langle\possibleworlds,\modalrelation,\interpretation\rangle$.
-\end{definition}
-
-\begin{definition}{\bf Causality as K Modal Interpretation Function --}\label{def:modalinterpretation}
-	Causality is (recursively) defined as the modal interpretation function $\interpretation$, as follows. 
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation0)$] if $\interpretation(\world,P)=\top$ then $\world\models P$
-		\item[$(\interpretation1)$] $\world\models\neg P$ iff $\world\not\models P$
-		\item[$(\interpretation2)$] $\world\models P \wedge Q$ iff $\world\models P$ and $\world\models Q$
-		\item[$(\interpretation3)$] $\world\models\Box P$ iff for any world $\world'\in\possibleworlds$ if $\world \modalrelation\world'$ then $\world'\models P$
-		\item[$(\interpretation4)$] $\world\models\Diamond P$ iff there exists a set of worlds $\World'\subset\possibleworlds$ such that for any $\world'\in\World'$, if $\world\modalrelation\world'$ then $\world'\models P$
-		\item[$(\interpretation5)$] $\models P$ iff $\actualworld\models P$
-	\end{enumerate}
-	where truth is defined as necessary with $\Box$ and possible with $\Diamond$.
-\end{definition}
-
-%\begin{definition}{\bf Causality --}\label{def:causality}
-%	The Causality principle is defined over a K Modal Logic (i.e. without
-%	restrictions on the causality relation between worlds) as in
-%	Definition~\ref{def:modallogic}, with the standard interpretation
-%	function defined in~\ref{def:modalinterpretation}
-%\end{definition}
-
-The causality principle has been defined in its
-generic form. In fact, the accessibility relation $\modalrelation$ is free from
-any axiomatic restriction (e.g. it's not reflexive nor anti-reflexive).  
-We will focus in Section~\ref{sec:process} (for our tests) on 
-the application of our theory to CPS system engineering.
-More detailed case studies (i.e. \fix{mr}{a CPS as a smart power-grid}) 
-will be defined in Section~\ref{sec:prototest}, where
-the accessibility relation defines in details how the system itself evolves, due to causal relation 
-(i.e. by restricting the causality principle only to those cause-effect that defines
-the system). Therefore, the definition of $\modalrelation$
-will be specialized in a more strict way based on the application domains and case study.
-
-We note that we have defined the
-Causality principle without considering it a \emph{threat source} since 
-we lack the concept of intent and maliciousness. Similarly, in the 
-next sections we won't discriminate between Nature and Mankind until we introduce 
-the concepts of maliciousness in Section~\ref{sec:theory} 
-and then formally define a threat agent
-in the same section.
-
-\subsubsection{Agents: Mankind and Nature}\label{sec:mankind-nature}
-\begin{table}[t]
-\centering
-\setlength{\tabcolsep}{3.5pt}
-\renewcommand{\arraystretch}{1}
-\scriptsize
-\caption{RCC3, RCC5, and RCC8 relations between Regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
-\begin{tabular}{ccclll} 
-\rota{\textbf{RCC3}}&\rota{\textbf{RCC5}}&\rota{\textbf{RCC8}}&\textbf{Name} & \textbf{Notation} & \textbf{Definition} \\
-\hline
-&&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& $\mathit{X}\subseteq \mathit{Y}$ \\
-&&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& $\mathit{X}\not\subseteq \mathit{Y}$\\
-&&&Part of				& $\mathit{P}(\mathit{X},\mathit{Y})$		& $\forall \mathit{Z} ~\mathit{C}(\mathit{Z},\mathit{X}) \rightarrow \mathit{C}(\mathit{Z},\mathit{Y})$\\
-&&&Overlaps			& $\mathit{O}(\mathit{X},\mathit{Y})$		& $\exists \mathit{Z} ~\mathit{P}(\mathit{Z},\mathit{X})\wedge \mathit{P}(\mathit{Z},\mathit{Y})$\\
-\Tdot&&&  \textbf{Overlaps Not Equal} 	& $\mathit{ONE}(\mathit{X},\mathit{Y})$		& $\mathit{O}(\mathit{X},\mathit{Y}) \land \neg \mathit{EQ}(\mathit{X},\mathit{Y})$ \\
-\Tdot&\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
-\Tdot&\Tdot&\Tdot& \textbf{DiscRete from} 		& $\mathit{DR}(\mathit{X},\mathit{Y})$		& $\neg \mathit{O}(\mathit{X},\mathit{Y})$\\
-&\Tdot&\Tdot&\textbf{Partial-Overlap}	& $\mathit{PO}(\mathit{X},\mathit{Y})$ 		& $\mathit{O}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{X},\mathit{Y}) \wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\
-&\Tdot&&\textbf{Proper-Part-of} 	& $\mathit{PP}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\ 
-	&\Tdot&&\textbf{Proper-Part-of-\textit{\textbf{i}}nverse} & $\mathit{PPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{Y},\mathit{X}) \wedge \neg \mathit{P}(\mathit{X},\mathit{Y})$\\
-&&\Tdot&\textbf{Externally Connected} 	& $\mathit{EC}(\mathit{X},\mathit{Y})$ 		& $\mathit{C}(\mathit{X},\mathit{Y}) \wedge \neg\mathit{O}(\mathit{X},\mathit{Y})$\\ 
-&&\Tdot&\textbf{Tangential PP} 	& $\mathit{TPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Tangential PPi} 	& $\mathit{TPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{TPP}(\mathit{Y},\mathit{X})$\\ 
-&&\Tdot&\textbf{Non-Tangential PP} 	& $\mathit{NTPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\neg\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Non-Tangential PPi} 	& $\mathit{NTPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{NTPP}(\mathit{Y},\mathit{X})$\\ 
-\end{tabular}
-\end{table}
-
-Mankind and Nature are defined in Section~\ref{sec:problem} as two abstract
-agents, both as collections (i.e. an abstract type that does not imply a
-specific implementation) of their sub-agents (i.e. humans for Mankind and
-plants, animals, \&c.  for Nature). Similarly to\autocite{Santaca2016abf}, 
-we define Mankind and Nature, and any
-other agent in the reminder of this article, as a meronomy (an hierarchy of
-Part-Whole relations) based on a standard definition of mereology, i.e. based
-on the definition of Parthood relation between \emph{Parts}.  
-However, shall 
-consider different types of Part; so, we extend the mereology to a
-mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
-to increase the number of different types considered and to generalize the relations between Parts (as
-in Table~\ref{tab:rcc358}).
-For the sake of readability, we use the term \emph{Region} both to refer to a
-mereological Part and to a topological Region.
-The choice of mereotopology is also correlated to the objective of defining a formal
-ontology, which we use to define the (formal) semantics of the terms (Parts) in
-Section~\ref{sec:problem}, and of the concepts of safety and security
-(whole). We aim at creating a meronomy instead of the taxonomies such as the one provided
-in\autocite{NIST2020NVD,MITRE2020CVE} or instead of the poorly justified 
-CVSS\autocite{Mell2007CVSS} scoring system.
-
-A mereotopology, as defined e.g. in\autocite{Rachavelpula2017mereotopology},
-is an ordered mathematical structure where the basic relation between Regions is the
-reflexive and symmetric\fixnote{mr}{I guess it must be monotonic as defined in\autocite{Rachavelpula2017mereotopology} but I don't find it consistently in other papers.} 
-\emph{Parthood} relation $\subseteq$. 
-\begin{definition}{\bf Parthood --}\label{def:parthood}
-Given any pair of mereotopological Regions $X$ and $Y$,
-	\begin{enumerate}[noitemsep]
-		\item Reflexivity: $\forall X.~ (X\subseteq X)$
-		\item Symmetry: $\forall X, Y.~ (X\subseteq Y \Rightarrow Y\subseteq X)$
-	\end{enumerate}
-\end{definition}
-
-As later defined in Definition~\ref{def:agent}, 
-the Parthood relation orders a universe of agents $\agentuniverse$ by defining
-the so called \emph{Connects with} (see in Table~\ref{tab:rcc358}) relation 
-between Regions. We want this universe $\agentuniverse$ to be 
-expressible in FOL. In this way, we
-can reason both on the constituent of security (i.e. its terms and agents
-defining a system where security needs to be considered), and on 
-the evolution of those constituent w.r.t. cause-effects relations according
-to the modal structure of causality we defined in Section~\ref{sec:causality}.
-This will allow us (in Section~\ref{sec:process} and Section~\ref{sec:prototest})
-a better\fixnote{mr}{Can we prove this by construction?} 
-positioning w.r.t. risk assessment technologies
-(which most often reason on the constituent of a system design), and
-protocol verification tools (which requires some formalization of a cause-effect relation, e.g.
-in Linear Temporal Logic).
-
-As we argued in Section~\ref{sec:problem}, we must correlate the definition of
-agent (i.e. Mankind and Nature) in Definition~\ref{def:mankind-nature} to the
-mathematical structure of the logic that defines them.  We express Mankind and
-Nature as formulas over the theory of mereology and then in terms of
-mereotopological Regions, extending the interpretation function
-$\interpretation$, to include a formal theory of mereotopology.  We use the
-Region Connection Calculus (RCC), as defined
-in~\cite{bennettLogics,improvingRCC}, to provide an axiomatization of the
-spatial concepts and relations in first-order logic to correlate the algebraic
-structure to mereology. In its broader definition, the RCC theory is composed
-by eight axioms, and is known as RCC8. In the text, for brevity, 
-we will often focus only on RCC5 (without
-loss of generality) by not considering tangential connections between spatial
-Regions. We discuss the choice of RCC5 in more detail in
-Appendix~\ref{app:rcc}. In Table~\ref{tab:rcc}, we
-summarize the axioms of the Region Connection Calculus (see, e.g., \autocite{Grutter2008rcc}).
-
-\begin{definition}{\bf RCC axiomatization --}
-	For any $X,Y$ pair of Regions in a mereotopology:
-	\begin{itemize}[noitemsep]
-	\item[$(\interpretation6)$] $\interpretation(X\subseteq Y)$ iff [$\interpretation(X\subseteq X)=\top$, and $\interpretation(X\subseteq Y)=\bot$ or $\interpretation(Y\subseteq X)=\top$]\fixnote{mr}{would it be better/clearer or just correct to write $\sigma(X\subseteq Y)$ iff $\sigma(X\subseteq X\wedge [X\subseteq Y \vee Y\subseteq X])=\top$}
-	\item[$(\interpretation7)$] $\world\models\connects{X}{Y}$ iff $\world\models X\subseteq Y$ 
-	\item[$(\interpretation8)$] $\world\models\disconnected{X}{Y}$ iff  $\world\not\models X\subseteq Y$
-	\item[$(\interpretation9)$] $\world\models\partof{X}{Y}$ iff $\world\models\forall Z.~\connects{Z}{X}\Rightarrow\connects{Z}{Y}$
-	\item[$(\interpretation10)$] $\world\models\overlaps{X}{Y}$ iff $\world\models\exists Z.~\partof{Z}{X}\wedge\partof{Z}{Y}$
-	\item[$(\interpretation11)$] $\world\models\eq{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\partof{Y}{X}$
-	\item[$(\interpretation12)$] $\world\models\dr{X}{Y}$ iff $\world\models\neg\overlaps{X}{Y}$
-	\item[$(\interpretation13)$] $\world\models\po{X}{Y}$ iff $\world\models\overlaps{X}{Y}\wedge\neg\partof{X}{Y}\wedge\neg\partof{Y}{X}$
-	\item[$(\interpretation14)$] $\world\models\pp{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\neg\partof{Y}{X}$
-	\item[$(\interpretation15)$] $\world\models\ppi{X}{Y}$ iff $\world\models\partof{Y}{X}\wedge\neg\partof{X}{Y}$
-\end{itemize}
-where $Z$ is a mereotopological Region.
-\end{definition}
-
-\begin{definition}{\bf Agent: Mankind or Nature --}\label{def:agent}
-	An agent $\agent\in\agentuniverse$ is a tuple
-	$\langle\rcc(\region',\region''),\ldots,\rcc(\region^{n-1},\region^n)\rangle$
-	of RCC relations $\rcc$ over mereotopological Regions
-	${\region',\ldots,\region^n}\subseteq\Region$. 
-\end{definition}
-
-Currently, we do not distinguish between Mankind and Nature (since we still
-lack of the definition of ``malicious intent'', which is defined in
-Section~\ref{sec:theory}) and we have defined them as two generic
-agents.
-
-As depicted in Figure~\ref{fig:safety-security}, Causality, Mankind, and Nature
-have a dashed border representing their correlation in terms of cause-effect
-and then in terms of formal structure which defines them: Modal Logic.
-Vulnerability, Defect, Incident, Accident, and Threat, similarly, 
-are correlated (depicted as a solid border) in terms of underlying 
-formal structure: Mereotopology.
-
-\subsubsection{Regions: Vulnerability and Defect (and Weakness)}\label{sec:vulnerabilitydefect}
-As defined in Section~\ref{sec:problem}, a Vulnerability or a Defect is a
-\emph{Weakness}.  As an example, a categorization of weaknesses is given
-in\autocite{MITRE2020CWEresearch} with 808 weaknesses categorized as ``Research
-Concepts'', distributed as follows:
-\begin{itemize}[noitemsep]
-	\item Incorrect Calculation - ($682$)
-	\item Incorrect Access of Indexable Resource (``Range Error'') - ($118$)
-	\item Use of Insufficiently Random Values - ($330$)
-	\item Improper Interaction Between Multiple Correctly-Behaving Entities - ($435$)
-	\item Improper Control of a Resource Through its Lifetime - ($664$)
-	\item Insufficient Control Flow Management - ($691$)
-	\item Protection Mechanism Failure - ($693$)
-	\item Incorrect Comparison - ($697$)
-	\item Improper Check or Handling of Exceptional Conditions - ($703$)
-	\item Improper Enforcement of Message or Data Structure - ($707$)
-	\item Improper Adherence to Coding Standards - ($710$)
-\end{itemize}
-
-The definition given by the MITRE in\autocite{MITRE2020CWEweakness} of
-weakness is: ``Software weaknesses are errors that can lead to software
-vulnerabilities. A software vulnerability, such as those enumerated on the
-Common Vulnerabilities and Exposures (CVE) List, is a mistake in software that
-can be directly used by a hacker to gain access to a system or network''.
-The definition is circular if we interpret the word ``error'' and
-``mistake'' with the same semantics: a weakness is an error that leads to a
-vulnerability and a vulnerability is a mistake which, in turn, is a weakness.
-The only difference (between weakness and vulnerability) seems to be
-that one can consider weakness as a ground term and state that a
-vulnerability is caused by a weakness, i.e. 
-$\World,\Weakness\models\Diamond\Vulnerability\wedge\Weakness$ where
-$\Weakness,\Vulnerability$ are Regions of Weaknesses and Vulnerabilities (resp.);
-accepting the hierarchy in the CWE\autocite{CWE} as ground truth. 
-Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability) 
-or the CVE reported in the NVD as a ground truth.
-
-\begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
-	A Region $\region\subseteq\agent$ of an agent
-	$\agent\in\agentuniverse$, is defined as Weakness $\Weakness$ (i.e.
-	representing a weakness introduced by the agent into any phase of
-	production, e.g.  of the secure process development life-cycle, of any
-	system or subsystem) iff there exists $\world\in\possibleworlds$ such
-	that $\world,\Weakness\models\Diamond\exists\Weakness',\Vulnerability'.~
-	\rcc(\Weakness',\Vulnerability')\wedge\neg\dr{\Weakness'}{\Vulnerability'}$, 
-	where $\Weakness'\subseteq\Weakness$, and
-	$\Vulnerability'\subseteq\Vulnerability$ is a Region of an agent $\agent'\in\agentuniverse$.
-\end{definition}
-
-\begin{example}{CWE-116: Improper Encoding or Escaping of Output\autocite{CWE-116} --}
-	\begin{itemize} 
-		\item Description: The software prepares a structured message
-		for communication with another component, but encoding or
-		escaping of the data is either missing or done incorrectly. As
-		a result, the intended structure of the message is not
-		preserved. 
-		\item Example: This code displays an email
-		address that was submitted as part of a form. Example language JSP.
-		\begin{verbatim} 
-		<% String email = request.getParameter("email"); %> 
-		...
-		Email Address: <%= email %>
-		\end{verbatim}
-		The value read from the form parameter is
-		reflected back to the client browser without
-		having been encoded prior to output, allowing
-		various XSS attacks (CWE-79).
-		\item Observed Examples
-			\begin{itemize}
-			\item CVE-2008-4636\autocite{CVE-2008-4636}: OS command injection in backup
-				software using shell metacharacters in a
-				filename; correct behavior would
-				require that this filename could not be
-				changed.
-			\item CVE-2008-0769\autocite{CVE-2008-0769}: Web application does not set the
-				charset when sending a page to a browser,
-				allowing for XSS exploitation when a
-				browser chooses an unexpected encoding.
-			\item CVE-2008-0005\autocite{CVE-2008-0005}: Program does not set the charset
-				when sending a page to a browser, allowing for
-				XSS exploitation when a browser chooses
-				an unexpected encoding.
-			\item CVE-2008-5573\autocite{CVE-2008-5573}: SQL injection via password
-				parameter; a strong password might contain \&
-			\item CVE-2008-3773\autocite{CVE-2008-3773}: Cross-site scripting in chat
-				application via a message subject, which
-				normally might contain \& and other
-				XSS-related characters.
-			\item CVE-2008-0757\autocite{CVE-2008-0757}: Cross-site scripting in chat
-				application via a message, which normally might
-				be allowed to contain arbitrary
-				content.
-			\end{itemize}
-\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\begin{subfigure}[b]{.5\textwidth}
-		\centering
-		\includegraphics[width=.7\textwidth]{rcc5_CWE-116.pdf}
-		\caption{Representation with RCC3}
-		\label{fig:rcc5_CWE-116}
-	\end{subfigure}\begin{subfigure}[b]{.5\textwidth}
-		\centering
-		\includegraphics[width=.7\textwidth]{rcc8_CWE-116.pdf}
-		\caption{Representation with RCC8}
-		\label{fig:rcc8_CWE-116}
-	\end{subfigure}
-	\begin{subfigure}[b]{\textwidth}
-		\centering
-		\includegraphics[width=\textwidth]{KML_CWE-116.pdf}
-		\caption{Representation in K-Modal Logic}
-		\label{fig:KML_CWE-116}
-	\end{subfigure}
-	\caption{Relations between CWE-116 and correlated CVEs}
-\end{figure}
-Definition~\ref{def:weakness} states that CWE-116 is a weakness iff there exists a world in
-K Modal Logic, representing the system in which this weakness exists, 
-such that the natural evolution of this system (i.e. formalized by the causality principle) 
-make it possible to reach another state of the system (i.e. another, accessible world)
-where there exist a vulnerability that can be implemented from CWE-116 (and this relation 
-is not DR). The CWE website proposes the connection between CWE-116 and, for example, 
-CVE-2008-5573; a vulnerability of the login sub-system 
-of the Poll Pro v2.0\autocite{pollpro} system.
-The formal relation between the two is given as a link (i.e. URI) between the CWE and
-the CVE, the description of the relation is not defined but it is supposed to be
-inferred from the descriptions of the CWE-CVE. 
-We can formally represent this link as the ONE connection in RCC3, depicted in 
-Figure~\ref{fig:rcc5_CWE-116}, or EC connection in RCC8 (Figure~\ref{fig:rcc8_CWE-116}).
-
-\fixnote{mr}{shouldn't we consider the agent who introduces the weakness as dishonest? don't we say this in part 1?}
-\end{example}
-
-It is interesting to note that the CWE-CVE relation expresses the correlation
-between weaknesses and vulnerabilities in the most simple form, such that we
-can formalize all the relation as ONE in RCC3 or EC in RCC8.  To express a more
-complex relation between the two we shall analyze the definition of CWE-116.
-This is related to the Weakness-Vulnerability-Incident process (i.e. to the
-details on the implementation/realization in Figure~\ref{fig:safety-security})
-that we analyze in the next section, Section~\ref{sec:incidentaccident}.
-
-\begin{definition}{\bf Region: Vulnerability and Defect --}\label{def:vulnerability-defect}
-A region $\region\subseteq\agent$ of an agent $\agent\in\agentuniverse$ is
-	called \emph{Vulnerability} if the agent $\agent$ is referred to as Mankind, 
-	\emph{Defect} if the agent is referred to as Nature.
-\end{definition}
-
-\subsubsection{Process: Incidents and Accidents}\label{sec:incidentaccident}
-In our informal definition depicted in Figure~\ref{fig:safety-security}, an
-Incident is generated through a process caused by a Vulnerability (which,
-in turn, is caused by a Weakness). Symmetrically, Nature has a process from Defect
-to Accidents. We start by analyzing the process generated by Mankind, which is
-divided into the following phases (explained in the remainder of this sub-Section):
-%Agent$\rightarrow$Vulnerability$\rightarrow$Exploit$\rightarrow$Attack$\rightarrow$Incident.
-\begin{enumerate}[noitemsep]
-	\item An agent (Mankind or Nature) defines a system, e.g.
-by following an engineering process for creating a new system, or for redesigning or improving an existing system (e.g.
-a component or a collection of components)
-	\item The dogmatic definition of a system contains a Weakness (i.e.
-the agent, Mankind or Nature, introduces an unintended behavior), for example as an un-intended behavior or as an architectural fallacy.
-The Weakness can be applied to the abstract definition of the system.
-	\item The Weakness can be defined as a malicious process for a system, called Vulnerability (or Defect for Nature)
-	\item The process expressed by the Vulnerability can be made operational for a specific implementation of the system (i.e. implemented by Mankind or realized by Nature) 
-\end{enumerate}
-Those additional details are depicted in Figure~\ref{fig:safety-security_2}, an update
-of Figure~\ref{fig:safety-security}.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{safety-security_2.pdf}
-	\caption{Overview security and safety keywords}
-	\label{fig:safety-security_2}
-\end{figure}
-
-\Paragraph{Exploitation Flow} An \emph{Exploit} is an \emph{implementation} of
-a Vulnerability, where a Vulnerability is one of the many ways to transform a
-Weakness into a malicious process that can be applied\footnote{Meaning that the
-malicious process, i.e.  any structured procedure such as a protocol, can be
-somehow made operational, e.g. implemented} to one or many systems. So, the act
-of an agent (Mankind) of ``exploiting a Vulnerability'' is the process of
-making the Vulnerability operational, by implementing the Vulnerability w.r.t.
-a specific target system.  Whenever the agent is Nature, the implementation is
-more broadly considered a realization\footnote{``The state of being
-realized''\autocite{Merriam2020realization}}
-
-The Common Attack Pattern Enumeration and
-Classification\autocite{MITRE2020CAPEC} (CAPEC) online service provides a
-``community resource'' which aims at ``identifying and understanding attacks''.
-The MITRE writes on the CAPEC homepage that ``Understanding how the adversary
-operates is essential to effective cyber security. CAPEC helps by providing a
-comprehensive dictionary of known patterns of attack employed by adversaries to
-exploit known weaknesses in cyber-enabled capabilities. It can be used by
-analysts, developers, testers, and educators to advance community understanding
-and enhance defenses.''. We can extrapolate two major concepts: 
-\begin{enumerate}
-	\item while understanding the exploitation process is the aim of the
-		CWE and CVE initiatives, understanding the hacking process is
-		the aim of the CAPEC initiative.
-	\item CAPEC provides a \emph{dictionary} of \emph{known} attack patterns.
-		As stated largely in Section~\ref{sec:}, this dictionary contains
-		empirical evidences and should not be used to induce a cybersecurity 
-		theory; but it can be used to validate a theory.
-\end{enumerate}
-
-The process continues as follows.
-\begin{enumerate}[noitemsep]
-\setcounter{enumi}{4}
-	\item an agent (Mankind\footnote{For the sake of readability we focus only on Mankind as an agent}) the exploitation flow can be used to hack a system.
-We use the term hack (as hacking) to stress that nothing prevents this process from being honest, however, for the sake of simplicity, we focus on Attacks which are malicious by definition.
-	\item An Attack is applied to a system with malicious intent by an agent
-	\item The application of the attack results in an Incident which impacts, i.e. poses a Threat to, an Asset
-\end{enumerate}
-
-\Paragraph{Hacking Flow}
-
-\begin{example}
-given in natural language in the description of the 
-CWE-116 which we can analyze as follows:
-\begin{itemize}
-	\item ``the software prepares a structured message''
-	\item the message is intended ``for communication with another component''
-	\item ``but''
-	\item ``encoding or escaping of the data is either missing or done incorrectly''
-	\item ``[\ldots] result, the intended structure of the message is not preserved''
-\end{itemize}
-\end{example}
-
-In this Vulnerability is seen as a process through which an agent (i.e. Mankind)
-generates an Incident, i.e.
-
-We describe processes, in our current formalization, in terms of the K-Modal
-Structure as in Definition~\ref{def:modalinterpretation} (i.e. in terms of
-causality principle).
-
-\begin{definition}{\bf Vulnerability-Incident Process --}\label{def:vulnerability-incident}
-	Given two 
-\end{definition}
-
-The whole-part relation \fix{mr}{(i.e.  HAS-A\footnote{HAS-A is just an hint for the
-mechanization of this theory.  We'll come back to this in
-Section~\ref{sec:implementation}})} is given, in the following, over the K Modal
-logic described beforehand. To do so, we first need to formally define the
-concept of \emph{agent} in its abstract form. We build it following the same
-structure of\autocite{Santaca2016abf} but over the terms used in
-Section~\ref{sec:problem}, i.e. knowledge, and belief as epistemic concepts. We'll then 
-introduce the concept of \emph{intent} to define a \emph{threat agent}.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.8\textwidth]{knowledge-belief.pdf}
-	\caption{Informal representation of Knowledge and Belief}
-	\label{fig:knowledge-belief}
-\end{figure}
-
-\subsubsection{Region: Threat and Asset}\label{sec:threatasset}
-The difference between Knowledge and Belief is depicted in
-Figure~\ref{fig:knowledge-belief} (see \autocite{wiki-knowledgebelief}).  However,
-according to\autocite{Gettier2012knowledge},
-Knowledge\footnote{``\emph{Theaetetus}: [\ldots] He said that knowledge was
-true opinion accompanied by reason, but that unreasoning true opinion was
-outside of the sphere of knowledge; and matters of which there is not a
-rational explanation are unknowable -- yes, that is what he called them -- and
-those of which there is are knowable. [\ldots] \emph{Socrates}: [\ldots] the
-primary elements of which we and all else are composed admit of no rational
-explanation; for each alone by itself can only be named, and no qualification
-can be added, neither that it is nor that it is not, for that would at once be
-adding to it existence or non-existence, whereas we must add nothing to it, if
-we are to speak of that itself alone.  [\ldots]'' Plato -- Theaetetus 201
-\autocite{Plato1914Plato}} as an epistemological concept is difficult (and
-sometimes believed impossible\autocite{citation}) to formally define. In this
-work, we are not interested in 
-how knowledge can be precisely formalized from an epistemic standpoint. 
-We assume that a semantic of a correct definition of epistemic knowledge exists,
-for example the one given in\autocite{Hintikka1962knwoledge} by Hintikka, and we then define 
-knowledge in terms of the Kripke structure defined in Definition~\ref{def:causality}.
-
-\begin{definition}{\bf Knowledge --}\label{def:knowledge}
-Given an abstract collection of Agents $Ag$, and the modal operator
-	$\knows{a}{}$ (where $a\in Ag$), Knowledge is defined as a collection
-	of predicates known by an agent $\knowledge{a}=\bigcup_\Phi \knows{a}{\varphi}$ 
-	(where $\Phi$ is the collection of all the propositions known by $a$).
-	Given a proposition $P$, we extend the semantics of the Causality structure with:
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation6)$] $\world\models\knows{a}{P}$ iff
-			$\world'\models P$ for all $\world'$ such that
-			$\world\modalrelation\world'$.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Belief --}\label{def:belief}
-	\fix{mr}{we should help the reader in understanding the upgrades of the logic.
-We used propositional with modal operators with a K Kripke semantics and we are now introducing Belief as in Doxastic logic?
-and we have epistemological operators as knowledge}
-	Given an abstract collection of Agents $Ag$, and the modal operator
-	$\believe{a}{}$ (where $a\in Ag$), Belief is defined as a collection
-	of predicates believed by an agent $\belief{a}=\bigcup_\Phi \believe{a}{\varphi}$
-	(where $\Phi$ is the collection of all the propositions believed by $a$).
-	Given a proposition $P$, we extend the semantics of the Causality structure with:
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation7)$] $\world\models\believe{a}{P}$ iff
-			$\world\models\neg\knows{a}{\neg P}$ (i.e. the agent $a$ considers $P$ possible) 
-			and $\world'\models P$ for all
-			$\world'$ such that $\world\modalrelation\world'$.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Mankind/Nature --}\label{def:mankind-nature}
-Mankind/Nature is represented as an agent.  
-	%countable logical signature $\Sigma_M$/$\Sigma_N$ over an
-	%abstract algebraic structure.  We also consider a countable signature
-	%of variables, disjoint from $\Sigma_M$/$\Sigma_N$, and write
-	%$\Sigma_{M_n}$/$\Sigma_{N_n}$ for the symbols of $\Sigma_M$/$\Sigma_N$
-	%with arity $n$. Thus, $\Sigma_0$ is a collection of constants, which we
-	%assume to have disjoint sub-collections that we refer to as
-	%\emph{sub-agents} of Mankind or Nature (i.e. humans, plants, \&c).
-	%The variables are untyped and can be instantiated with arbitrary types,
-	%yielding an untyped model.
-\end{definition}
-
-\subsubsection{Mereotopological Representation of Safety and Security}\label{sec:formalGlossary}
-The objectives of this section are:
-\begin{enumerate}[noitemsep]
-	\item formalize the abstract algebraic structure (as a mereotopology)
-		on top of which we have formalized the terms in
-		Figure~\ref{fig:safety-security} so far, i.e. Mankind, Nature,
-		Truths, Beliefs, and Knowledge.
-	\item correlate mereological structure to Causality and then Kripke
-		structure by defining the relation $R$ over the Kripke
-		structure as the mereotopological basic relation connects with,
-		i.e. reflexive and symmetric
-	\item we map signature to mereotopological Regions to express the
-		logical formulation of terms (e.g. Mankind, Nature, Truths,
-		Beliefs, Knowledge, Vulnerabilities) over the mereotopological
-		structure
-	\item we formalize the relations between terms using the RCC(5)
-\end{enumerate}
-
-\fix{mr}{change propositions in propositional language\url{http://logic.stanford.edu/intrologic/lectures/lecture_02.pdf}}
-
-Given that the concepts of Truths, Beliefs, and Knowledge are positioned over
-an underlying structure of Propositions, the first concept to formalize is that
-of Propositions. We do not restrict Propositions to a specific structure but,
-as for Mankind and Nature, we refer to them as a collection.  We express the
-generality of a collection as a (propositional) signature over an abstract
-algebraic structure. In our formalization, we use mereotopology as algebraic
-structure (due to its generality, as discussed in Section~\ref{sec:problem})
-and the RCC to formalize concepts over a mereotopology.
-
-We can now define the possible mereotopological relations between Belief and
-Knowledge (informally depicted in Figure~\ref{fig:knowledge-belief}) and then
-define the peculiarity of the believed difference between Mankind and Nature 
-(i.e. maliciousness as intent). 
-We first start from the formalization of Figure~\ref{fig:knowledge-belief}, as
-follows.
-
-\begin{definition}{\bf Information}\label{def:information}
-\end{definition}
-\begin{definition}{\bf Desire}\label{def:desire}
-\end{definition}
-\begin{definition}{\bf Intent}\label{def:intent}
-\end{definition}
-
-\subsection{Vulnerability and Defect}\label{sec:vulnerability-defect}
-In order to formalize the semantics of the concepts (informally) described in
-Section~\ref{sec:problem},
-
-\begin{definition}{\bf Vulnerability}
-\end{definition}
-
-\subsection{Cybersecurity Engineering Process}\label{sec:process}
-%\epigraph{La differenza tra idee e realt\`a \`e la differenza tra
-%filosofia e ingegneria. Il lavoro per trasformare l'una nell'altra \`e ricerca
-%scientifica
 \epigraph{The difference between ideas and reality is the
@@
 
-\section{Honesty and Maliciousness -- A Cybersecurity Theory}\label{sec:theory}
-\begin{enumerate}[noitemsep]
-	\item cybersecurity as a requirement
-	\item vulnerability\label{def:vulnerability}
+\begin{enumerate}
+\item specification -- definition of the desired design of a system. The
+specification is verified w.r.t. the  $\abf$ theory. Checking the
+controls of the whole cybersecurity life-cycle and the relation with
+CWE.
+\item design -- the mitigations identified in the specification stage
+are implemented into the design. The security of the assets w.r.t. the design are
+verified with a, so called, cybersecurity risk assessment.
 \end{enumerate}
-
-\section{Prototype and Empirical Tests}\label{sec:prototest}
-\subsection{Prototype Implementation}\label{sec:implementation}
-\subsection{Empirical Tests}\label{sec:tests}
-
-\section{Conclusion and Related Work}\label{sec:related}
-
-\printbibliography
-
-\begin{appendices}
-	\section{RCC3, RCC5, or RCC8}\label{app:rcc}
-	\section{Wrong idea 1}
-In order to apply consistently the mereotopological structure not just to
-reason the agents but also on their evolution, we extend the properties of the
-Parthood relation (reflexivity and symmetry) to the modal relation
-$\modalrelation$ of the K Modal logic, which in turns defines the causality
-principle. This updates the K Modal logic to a B (Brouwer) Modal Logic\autocite{Garson2018modal}.
-
-\begin{definition}{\bf B Modal Logic --}\label{thm:bk-parthood}
-	The B Modal Logic (defined, with a slight abuse of notation as
-	$\bframe=\langle\possibleworlds,\modalrelation\rangle$) is an extension
-	of the K modal logic with the following axioms
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation8)$] $\world\models P\Rightarrow Q$ iff
-			$\world\not\models P$ or $\world\models Q$ (logical implication)~\footnote{In
-			Definition~\ref{def:modallogic} we didn't specify the
-			semantics in the case of the connective $\Rightarrow$
-			(i.e. logical implication), so we state it here. Given
-			that $P\Rightarrow Q$ is equisatisfiable to
-			$\neg(P\wedge \neg Q)$, for any proposition $P$ and
-			$Q$, $\world\models P\Rightarrow Q$ is equisatisfiable
-			to $\world\models \neg(P\wedge \neg Q)$. If
-			$\world\models \neg(P\wedge \neg Q)$ then
-			$\world\not\models P\wedge\neg Q$ and then
-			$\world\models \neg P$ or $\world\models Q$.}
-		\item[$(\interpretation9)$] $\Box P\Rightarrow P$ (which ``claims that whatever is necessary is the case''\autocite{Garson2018modal}), and
-		\item[$(\interpretation10)$] $P\Rightarrow\Box\Diamond P$
-			(which says that if $P$ is the case, then $P$ is
-			necessarily possible\autocite{Garson2018modal})
-	\end{enumerate}
-\end{definition}
-
-\begin{theorem}{\bf Causality, Parthood, and B Modal Logic}\label{thm:bk-parthood}
-	\begin{enumerate}[noitemsep]
-		\item (Reflexivity) $\bframe\models\Box P\Rightarrow P$ ($\interpretation8$) iff $\modalrelation$ is reflexive
-		\item (Symmetry) $\bframe\models P\Rightarrow\Box\Diamond$ ($\interpretation9$) iff $\modalrelation$ is symmetric
-		\item (Consistency) $\world\modalrelation\world'$ iff $\connects{\region}{\region'}$
-	\end{enumerate}
-	whenever $\world$ and $\world'$, and $\region$ and $\region'$ are logically equivalent ($\equiv$).
-\end{theorem}
-\begin{proof}
-	The proof is exhaustive over all possible cases.
-	\begin{enumerate}[noitemsep]
-		\item(Reflexivity) proven in\autocite{}
-		\item(Symmetry) proven in\autocite{}
-		\item(Equivalence) \fix{mr}{todo} If
-			$\world\modalrelation\world'$ and
-			$\world\equiv\region$, then if
-			it doesn't hold that $\connects{\region,\region'}$ it
-			follows that there is no reflexive and transitive
-			relation that correlates $\region$ to $\region'$. Given 
-			that Regions are defined as signatures
-	\end{enumerate}
-\end{proof}
-
-\end{appendices}
-
-\end{document}

### 2020-02-27 prototypes/agent_categorization/2a2w.py
add benchmarks

@@ -1,0 +1,271 @@
+#!/usr/bin/python3
+#author: Marco Rocchetto @V-Research
+
+#requirements
+# 1) z3 python API (https://github.com/Z3Prover/z3/blob/master/README.md)
+
+from z3 import *
+import scipy.special
+import itertools
+from multiprocessing import Process
+from multiprocessing import Queue 
+
+################################
+# since we unfold the quantifiers we have to calculate
+# the (minimal) number of subsets of A, B, and F that allows
+# Z3 to find a model that satisfies the agent configuration
+# 
+# For example, if we have A,B,F and no subsets of those 3 sets 
+# there is no model that satisfies PO(A,B) /\ PO(B,F) /\ PO(A,F)
+# since PO(A,B) <-> PP(A,B) if A has no subsets 
+#
+# requires scipy for binomial calculation: pip install scipy
+# otherwise download wheel from https://pypi.python.org/pypi/scipy
+# and install it with pip install <downloaded-file>
+################################
+def calculate_sets(sets, Set):
+    set_of_subsets=[]
+    for k in range(1,len(sets)):
+        counter=0
+        for i in range(int(scipy.special.binom(len(sets), (k+1)))):
+            subset="subset_%s_%d"%((k+1), counter)
+            set_of_subsets.append(Const(subset, Set))
+            #print("%d - (%d %d)=%d"%(i,len(sets),(k+1),int(scipy.special.binom(len(sets), (k+1)))))
+            counter+=1
+        
+    return set_of_subsets
+
+def topology(solver, sets_and_subsets, P):
+    ################################
+    # PART OF
+    ################################
+    
+    for s in sets_and_subsets:
+        #solver.add(P(s,s))
+        solver.assert_and_track(P(s,s), str("reflexivity(%s)"%s))
+    
+    for s1 in sets_and_subsets:
+        for s2 in sets_and_subsets:
+            #solver.add(Implies(And(P(s1,s2),P(s2,s1)), s1==s2))
+            solver.assert_and_track(Implies(And(P(s1,s2),P(s2,s1)), s1==s2), str("asymmetry(%s,%s)"%(s1,s2)))
+    
+    for s1 in sets_and_subsets:
+        for s2 in sets_and_subsets:
+            for s3 in sets_and_subsets:
[…]

### 2020-02-27 prototypes/agent_categorization/2a2w_sequential.py

@@ -1,0 +1,252 @@
+#!/usr/bin/python3
+#author: Marco Rocchetto @V-Research
+
+#requirements
+# 1) z3 python API (https://github.com/Z3Prover/z3/blob/master/README.md)
+
+from z3 import *
+import scipy.special
+import itertools
+
+################################
+# since we unfold the quantifiers we have to calculate
+# the (minimal) number of subsets of A, B, and F that allows
+# Z3 to find a model that satisfies the agent configuration
+# 
+# For example, if we have A,B,F and no subsets of those 3 sets 
+# there is no model that satisfies PO(A,B) /\ PO(B,F) /\ PO(A,F)
+# since PO(A,B) <-> PP(A,B) if A has no subsets 
+#
+# requires scipy for binomial calculation: pip install scipy
+# otherwise download wheel from https://pypi.python.org/pypi/scipy
+# and install it with pip install <downloaded-file>
+################################
+def calculate_sets(sets, Set):
+    set_of_subsets=[]
+    for k in range(1,len(sets)):
+        counter=0
+        for i in range(int(scipy.special.binom(len(sets), (k+1)))):
+            subset="subset_%s_%d"%((k+1), counter)
+            set_of_subsets.append(Const(subset, Set))
+            #print("%d - (%d %d)=%d"%(i,len(sets),(k+1),int(scipy.special.binom(len(sets), (k+1)))))
+            counter+=1
+        
+    return set_of_subsets
+
+def topology(solver, sets_and_subsets, P):
+    ################################
+    # PART OF
+    ################################
+    
+    for s in sets_and_subsets:
+        #solver.add(P(s,s))
+        solver.assert_and_track(P(s,s), str("reflexivity(%s)"%s))
+    
+    for s1 in sets_and_subsets:
+        for s2 in sets_and_subsets:
+            #solver.add(Implies(And(P(s1,s2),P(s2,s1)), s1==s2))
+            solver.assert_and_track(Implies(And(P(s1,s2),P(s2,s1)), s1==s2), str("asymmetry(%s,%s)"%(s1,s2)))
+    
+    for s1 in sets_and_subsets:
+        for s2 in sets_and_subsets:
+            for s3 in sets_and_subsets:
+               #solver.add(Implies(And(P(s1,s2),P(s2,s3)), P(s1,s3)))
[…]

### 2020-03-02 theory.tex
refine assertion, behavior, fact theory

@@
 
-\begin{definition}{\bf System State Space --}\label{def:system}
-	A state of a system (or a sub-system) $s$ is identified with 
-	an instance of an agent state (see Definition~\ref{def:agent})
-	such that
+\begin{definition}{\bf System State and State-Space --}\label{def:system}
+	The state of a system (or a sub-system) $s$ is identified with an
+	instance of an agent state (see Definition~\ref{def:agent}) such that
 	$s=\langle\rcc(\knowledgeRegion,\beliefRegion),\rcc(\knowledgeRegion,\informationRegion),\rcc(\beliefRegion,\informationRegion)\rangle$,
-	where $\knowledgeRegion,\beliefRegion,\informationRegion$ are Regions of Knowledge, Beliefs, and Information respectively.
+	where $\knowledgeRegion,\beliefRegion,\informationRegion$ are Regions
+	of Knowledge, Beliefs, and Information respectively, and $\rcc$ is a
+	specific relation between the Regions it predicates on.  The
+	state-space of a system is then defined by the different $\rcc$
+	relations between the three pairs of Regions defining the system and
+	all the sub-systems.
 \end{definition}
-A system state space is then defined as a superposition of multiple system states where knowledge,
+A system state space can be seen as a superposition of multiple agent states where knowledge,
 beliefs, and information are related between each other (in a mereotopological
-space) through the RCC. A \emph{system state} (as an instance of the system state space)
+space) through the all the admissible RCC relations. 
+A \emph{system state} (as an instance of the system state space)
 is a specific configuration of
@@
 			$\world\models\belief{a}{P}$ (i.e. the agent $a$ considers $P$ possible), 
-			$\world\models\neg\belief{a}{\neg P}$, and $\world'\models P$ for all
-			$\world'$ such that $\world\modalrelation\world'$.
+			$\world\models\neg\belief{a}{\neg P}$%, 
+			%and $\world'\models P$ for all
+			%$\world'$ such that $\world\modalrelation\world'$.
 	\end{enumerate}
@@
 
+\subsection{Engineering Operational System States}\label{sec:engsystemstate}
 When dealing with a specification of a system (i.e. the design if the system
 for a specific operation (as a generic operating/operational system),
-considering information may be an over-complication since we are usually not
-interested in the information in abstract but to the transfer of that
-information, which we call Assertions. Therefore, we consider information only
-if there is a sender and a recipient.  Similarly, the portion of Beliefs we
-consider is the one that builds the behavior of an agent (strategy rules
-\footnote{``The logical
-structure of information is one of the most basic and one of the most basic
-and one of the simplest thing in the wide and wonderful world of logical
-analysis. This point can be put in a deeper perspective. A distinction [\ldots] ought to
-be made [--] between two kinds of rules (or principles) in any strategic activity like
-knowledge seeking. On the one hand you have the rules that define the game, e.g. how
-chessmen are moved on a board. The can be called \emph{definitory} rules.
-They must be distinguished from rules [\ldots] that deal with what is better and what is worse
-in the game in question. Definitory rules do not say anything about this subject. Rules
-which do can be called \emph{strategic rules}'' -- Hintikka in \autocite{Hintikka1993Information}}) 
-and we consider a set
-of axiomatic Facts (definitory rules) instead of considering the more general 
-epistemic definition of Knowledge.
+considering abstract and general concepts such as knowledge, belief, or
+information may result to be too abstract for the objectives of the overall engineering
+process (e.g. verification and validation).
+As an example, we are usually not
+interested in information in abstract but to the transfer of that
+information, which we call Assertions, made by a subset of agents. 
+Therefore, we now define how we consider an agent for the engineering
+of CPS.
+
+\begin{enumerate}
+	\item We consider Information only when the intention of exchanging 
+		that Information is from a sender to 
+		recipient is defined; and we call it \emph{Assertion}  
+	\item Similarly, the portion of Beliefs we consider for system
+		engineering is the one that builds (input) or describes
+		(output) the behavior of an agent (strategy rules
+		\footnote{``The logical structure of information is one of the
+		most basic and one of the most basic and one of the simplest
+		thing in the wide and wonderful world of logical analysis. This
+		point can be put in a deeper perspective. A distinction
+		[\ldots] ought to be made [--] between two kinds of rules (or
+		principles) in any strategic activity like knowledge seeking.
+		On the one hand you have the rules that define the game, e.g.
+		how chessmen are moved on a board. The can be called
+		\emph{definitory} rules.  They must be distinguished from rules
+		[\ldots] that deal with what is better and what is worse in the
+		game in question.  Definitory rules do not say anything about
+		this subject. Rules which do can be called \emph{strategic
+		rules}'' -- Hintikka in \autocite{Hintikka1993Information}}),
+		and
+	\item We consider a set of axiomatic \emph{Facts} (definitory rules) instead
+		of considering the more general epistemic definition of
+		Knowledge. Specifically, Facts describes:
+		\begin{itemize}
+			\item the functional architecture of each agents
+			\item the structural (HW/SW) architecture of each
+				subsystem of agents and agent within a
+				subsystem
+			\item the data flow as a transfer of Beliefs through
+				Assertions (i.e. the Beliefs flow)
+		\end{itemize}
+\end{enumerate}
+
+We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.9\textwidth]{system-agent.pdf}
+	\caption{Example of system and agent structure}
+	\label{fig:system-agent}
+\end{figure}
 
@@
 
+\begin{definition}{\bf Behavior -- }\label{def:behavior}
+	The behavior of an agent is defined as the transformation process (e.g. defined as a
+	protocol, or a functional architecture) that determined the
+	output-beliefs based on input-beliefs (where input-beliefs and
+	output-beliefs are beliefs taken as input or output respectively).
+	\begin{enumerate}[noitemsep]
+		\item[$(\interpretation21)$] $\world\models\bigcup_{s\in S}\believe{s}{P_{s}}$ and $\world\models\behave{a}{[\bigcup_{s\in S}\believe{s}{P_{s}}]}$ iff
+			$\world\models\belief{a}{P'}$,
+
+		where $a$ is an agent, $S$ is a set of agents such that $a\neg\in S$, and the calculation of $P'$ is deteministic.
+	\end{enumerate}
+\end{definition}
+
 \begin{definition}{\bf Fact -- }\label{def:fact}
@@
 		\item[$(\interpretation20)$] $\world\models\fact{a}{P}$ iff
-			$\world \models P$ and 
-		.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Behavior -- }\label{def:behavior}
-	The behavior of an agent is determined by its beliefs. 
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation21)$] $\world\models\assert{a}{b}{P}$ iff
-			$\world\models\belief{a}{P}$, 
-			$\world'\models\rcc(\belief{a}{P},\belief{b}{P})$, and $\world'\models P$ for all
-			$\world'$ such that $\world\modalrelation\world'$.
+			$\world \models \Box P$. 
 	\end{enumerate}
@@
 \begin{definition}{\bf Operational System State --}\label{def:system}
-	An operational system (or a sub-system) is an agent (see Definition~\ref{def:agent})
+	An operational system (or a sub-system) is represented as an agent (see Definition~\ref{def:agent}) and then defined by the resulting state space,
 	such that
 	$s=\langle\rcc(\factRegion,\behaviorRegion),\rcc(\factRegion,\assertionRegion),\rcc(\behaviorRegion,\assertionRegion)\rangle$,
-	where $\factRegion,\behaviorRegion,\assertionRegion$, are regions of facts, behavior, and assertions respectively.
+	where $\assertionRegion,\behaviorRegion,\factRegion$, are regions of assertions, behavior, and facts respectively.
 \end{definition}
@@
 As already presented in \autocite{Santaca2016abf} it follows that, defining
-a system (or an operational system) with 3 regions, there exist
+a system (or an operational system) with a fixed number of regions, there exist
 an upper-bound to the number of possible configuration of a system, defined by
 the possible relations between the different regions.
-For completeness we report in the next paragraph 
+For completeness, we report in the next paragraph 
 the calculation done in \autocite{Santaca2016abf}.

### 2020-03-02 engineering.tex
first complete draft of abf theory

@@
 \end{figure}
+
+\subsection{Elicitation of Security Requirements}\label{sec:properties}
+Infosec, or Information Security, aims at mitigating the security
+risks related to information. The standard de-facto is 
+the so called CIA triad (where CIA stands for Confidentiality, Integrity, Availability).
+
+Confidentiality: the information is accessible only to authorised individuals.
+A set of rules that limits access to information.  The information can be
+accessed by X if and only if X is authorized to access it = X has the right
+attributes (i.e. keys, password) to access it.
+
+Integrity: the information is trustworthy and accurate (i.e. uncorrupted and
+unaltered). Protects data from modification or deletion by unauthorized sources
+and the ability to undo any damage done.  (for data at rest) If the information
+X is written, A will always be read. If information X is modified to
+information X', information X can always be recovered.  (for data in transit)
+If the information X is sent, A will be received. If information X is modified
+to information X', information X can always be recovered.
+
+Availability: the information is accessible and usable when required. Guarantee
+of reliable access to the information by authorized people If the information X
+needs to be accessed, information X can be retrieved. If
+
+Authenticity: assurance that the information is from the source it claims to be
+from. Authenticity involves proof of identity. Authenticity is verified through
+authentication.  If information X arrives from A to B, A sent information X and
+B can verify it.  ??Authenticity implies integrity but Integrity doesn't imply
+authenticity??  ??Authenticity doesn't imply confidentiality but
+confidentiality does imply authenticity??
+
+Non-repudiation: the ability to ensure that a party to a contract or a
+communication cannot deny the authenticity of their signature on information
+(or the sending of) that they originated.  If information X arrives from A, A
+sent it and cannot deny it (partial overlap??).  If sender A sends information
+X, it cannot say he didn't do it.  Non-repudiation implies authenticity and
+integrity. Authenticity doesn't imply non-repudiation (replay attacks?).
+Integrity doesn't imply non-repudiation (replay attacks?).
 

### 2020-03-02 theory.tex
first complete draft of abf theory

@@
 
-\subsection{Engineering Operational System States}\label{sec:engsystemstate}
+\subsection{Operational System State}\label{sec:engsystemstate}
 When dealing with a specification of a system (i.e. the design if the system
@@
 Facts should be correlated to the causality principle. Therefore a
-specification in opnet should be translated into vmt and and by bounded model
+specification in opnet should be translated into vmt and by bounded model
 checking with nuxmv identify a path towards an undesired state, if that fails
@@
 testing process tests facts against knowledge as much as we test a cps design
-chiices against its specification.  Information carries the knowledge of the
+choices against its specification.  Information carries the knowledge of the
 negation of the opposite of the information. Therefore an assertion carries the
@@
 }
-
-\subsection{Security Properties and Assets}\label{sec:properties}
-In \autocite{Georgeoff1995bdi}, the BDI (Belief, Desire, Intent)
-paradigm is explained; where an agent can be defined over those
-three constituents.

### 2020-03-02 engineering.tex
add intro: sec eng process and sec properties

@@
 
+Infosec, or Information Security, aims at mitigating the security risks related
+to information. The standard de-facto is the so called CIA triad 
+(where CIA stands for Confidentiality, Integrity, Availability), which is often
+criticized for being too general or non-adequate (e.g. by adding authenticity which is 
+often used as a building block for confidentiality and integrity) 
+to be effectively applied to the engineering of systems. Due to this,
+many researchers an organization tried to improve the CIA triad. 
+The evolutions/extensions of the CIA triad has been documented, e.g.,
+in \autocite{Samonas2014cia} and is summarized in Table~\ref{tab:ciaevolution}.
+
+\begin{table}[h]
+\centering
+%\setlength{\tabcolsep}{3.5pt}
+%\renewcommand{\arraystretch}{1}
+\small
+\begin{tabular}{rll} 
+	{\bf Year} & {\bf Definition} & {\bf Legend}\\
+	\hline
+	1970s & infosec = CIA & Confidentiality, Integrity, Availability\\
+	1980s & infosec += (Au, nR) & Authenticity and non-Repudiation\\
+	1990s & infosec += CSpec & Correctness in Specification\\
+	2000s & infosec += RITE & Responsibility, Integrity of people, Trust, Ethicality
+\end{tabular}
+\caption{Chronological progression of the CIA triad~\label{tab:ciaevolution}}
+\end{table}
+
+An example is the effort made by the OECD (Organisation for Economic
+Co-operation and Development) in \autocite{OECD2013guidelines} to define
+security guidelines for information system (1992) based on new principles such
+as ``awareness, ethics, risk assessment'' and maintain those revising the
+document (e.g.  following the multistakeholder expert consultation in 2013).
+Other similar efforts focused on defining entirely new principles or extending
+the CIA triad, such as \autocite{NISTSP800-160}. What is of interest for our
+argument is that, to the best of our knowledge, all of the approaches that aim
+at improving the CIA triad motivated by empirical evidences. On the contrary,
+we now define (in Section~\ref{sec:properties}) the CIA triad for an agent
+defined in the $\abftheory$ theory and we test if (i) other properties are
+allowed in the $\abftheory$ theory, and (ii) if and how the CIA triad can be
+detailed in the $\abftheory$ theory.
+
 \subsection{Elicitation of Security Requirements}\label{sec:properties}
-Infosec, or Information Security, aims at mitigating the security
-risks related to information. The standard de-facto is 
-the so called CIA triad (where CIA stands for Confidentiality, Integrity, Availability).
+In \autocite{Anderson1972report,Samonas2014cia}, CIA are defined as follows.
+\begin{itemize}
+	\item Unauthorized information release: an unauthorized person is able
+		to read and take  advantage  of  information  stored  in  the
+		computer.  This  category of concern  sometimes  extends  to
+		``traffic  analysis,''  in  which  the intruder  only observes
+		the  patterns  of  information  use.  From  those patterns,
+		the  intruder can   infer   some   information   content.
+		This category   also   includes   the unauthorized use of a
+		proprietary program.  (Confidentiality or Secrecy) 
+	\item Unauthorized  information  modification:  an  unauthorized person
+		is  able  to make changes in stored information –a form  of
+		sabotage.  It should be noted that in the case of this kind of
+		violation, the intruder does not necessarily see the
+		information he has changed.  (Integrity)
+	\item Unauthorized  denial  of  use:  an  intruder  can  prevent  an
+		authorized  user  from referring to, or from modifying
+		information, even though the intruder may not be able to refer
+		to, neither modify the information themselves. (Availability)
+\end{itemize}
+
+We now want to generalize the CIA triad to agents and not just 
+information and map it to the $\abftheory$ theory.
+\begin{definition}{\bf Infosec-Confidentiality --}\label{def:infosec-c}
+	An Information $\varphi$ is considered to be confidential over a set of
+	agents $Ag_c$ if, per each agent $a\in Ag_c$, the content of the
+	information (i.e. the data it carries) can only be understood (i.e. the relation
+	between the information and its data can be known) by the
+	agents in $Ag_c$.
+
+	\begin{itemize}
+		\item[$(\interpretation21)$]
+			for all agents $a\in Ag$, $\world\models\information{a}{\varphi}$ and
+			$\world\models\rcc(\varphi,Conf) \wedge \neg\dr{Conf}{\varphi}$ iff
+			$\world\models\knows{x}{[\information{a'}{\varphi}\rightarrow\varphi]}$
+			for any agent $a'$ such that $a'\not\in Ag_c$
+	\end{itemize}
+\end{definition}
+
+\begin{definition}{\bf System-Confidentiality --}\label{def:confidentiality}
+	The property of confidentiality holds in a system when its information
+	preserves confidentiality.
+\end{definition}
+
+\begin{definition}{\bf Agent-Confidentiality --}\label{def:confidentiality}
+An agent preserves the property of confidentiality iff 
+\end{definition}
 

### 2020-03-03 engineering.tex
complete def infosec-confidentiality and example

@@
 	\item Unauthorized  information  modification:  an  unauthorized person
-		is  able  to make changes in stored information –a form  of
+		is  able  to make changes in stored information -- a form  of
 		sabotage.  It should be noted that in the case of this kind of
@@
 \begin{definition}{\bf Infosec-Confidentiality --}\label{def:infosec-c}
-	An Information $\varphi$ is considered to be confidential over a set of
+	An Information $\varphi$ is considered to be confidential to a set of 
 	agents $Ag_c$ if, per each agent $a\in Ag_c$, the content of the
@@
 	\begin{itemize}
-		\item[$(\interpretation21)$]
-			for all agents $a\in Ag$, $\world\models\information{a}{\varphi}$ and
-			$\world\models\rcc(\varphi,Conf) \wedge \neg\dr{Conf}{\varphi}$ iff
-			$\world\models\knows{x}{[\information{a'}{\varphi}\rightarrow\varphi]}$
-			for any agent $a'$ such that $a'\not\in Ag_c$
+		\item[$(\interpretation21)$] for all agents $a\in Ag$,
+			$\world\models\information{a}{\varphi}$ and
+			$\world\models\rcc(\varphi,Conf) \wedge
+			\neg\dr{Conf}{\varphi}$ iff for any world $\world'$
+			such that $\world \modalrelation\world'$,
+			$\world'\models\believe{a}{\varphi}$, and
+			$\world'\models\knows{a}{\exists f.~
+			f(\information{a}{\varphi})=\varphi}$, where 
+			$Conf$ is a Region and $f$ is an uninterpreted function.
 	\end{itemize}
+\end{definition}
+
+\begin{example}{Dummy security protocol --}
+	The last part (part III) of the protocol in our first example
+	(Figure~\ref{fig:protocol-example}) represents the actual key exchange
+	which, in general, needs to be confidential; otherwise the key would be
+	known by any other agent who can, e.g., read on the communication
+	channel.  Therefore, the last message, in Alice-and-Bob notation can be
+	rephrased as: \begin{displaymath} Alice\Rightarrow Bob: enc(pk(Bob),
+	symmetricKey) \end{displaymath} where $enc$ is the standard asymmetric
+	encryption operator that encrypts the message $symmetricKey$ with the
+	public key of Bob $pk(Bob)$ (see, e.g.,
+	\autocite{Rocchetto2017interpolation} for a detailed formalization).
+	The function $f$ in Definition~\ref{def:infosec-c}, represents any
+	function used to enforce confidentiality on information.  In this
+	example, $f$ is identified with $enc$ (i.e.
+	$enc(pvt(Bob),symmetricKey)=symmetricKey$), the information is
+	$\information{Alice}{[enc(pk(Bob),symmetricKey)]}$ and the data, i.e.
+	$\varphi$ in Definition~\ref{def:infosec-c}, is $\varphi=symmetricKey$.
+	In the perfect cryptography assumption, and assuming for simplicity
+	that the logic of the whole protocol preserves confidentiality, the
+	message is confidential. In fact, if $Ag={Alice, Bob}$ and $\eq{symmetricKey}{Conf}$:
+	\begin{align*}
+		\world^0\models\belief{Alice}{symmetricKey},\\
+		\world^1\models\information{Alice}{symmetricKey}, \text{and}\\
+		\world^2\models\belief{Bob}{symmetricKey} ~\text{if} \\
+		enc(pvt(Bob), \information{Alice}{symmetricKey}=symmetricKey)=symmetricKey\\
+		(i.e. f=enc ~\text{and}~ \information{Alice}{symmetricKey}=enc(pub(Bob),symmetricKey))\\
+	\end{align*}
+\end{example}
+
+\begin{definition}{\bf Agent-Confidentiality --}\label{def:confidentiality}
+An agent preserves the property of confidentiality iff 
 \end{definition}
@@
 	preserves confidentiality.
-\end{definition}
-
-\begin{definition}{\bf Agent-Confidentiality --}\label{def:confidentiality}
-An agent preserves the property of confidentiality iff 
 \end{definition}

### 2020-03-03 theory.tex
qualitative evaluation of agents (and isk matrix)

@@
 allow us to consider 3.5 times more different agents.
+The quantitative evaluation of an agent is depicted in Figure~\ref{fig:quantitative}.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.8\textwidth]{quantitative.pdf}
+	\caption{Quantitative representation of the cybersecurity risk for a single agent}
+	\label{fig:quantitative}
+\end{figure}
+
+\subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}
+While a quantitative analysis reveals how many possible configurations
+of an agent (i.e. a system) may exist in $\abf$ (i.e. $54/125$ in RCC5), a qualitative
+analysis of the different configurations allows us to understand 
+which configurations are allowed by the $\abf$ theory.
+
+\begin{table}[t]
+\centering
+\begin{adjustbox}{width=\textwidth}
+\begin{tabular}{r||c|c|c|c|c} 
+& $\mathit{DR}$($\mathit{X}$, $\mathit{Y}$) & $\mathit{PO}$($\mathit{X}$,
+	$\mathit{Y}$)& $\mathit{EQ}$($\mathit{X}$, $\mathit{Y}$) &
+	$\mathit{PPi}$($\mathit{X}$, $\mathit{Y}$) & PP($\mathit{X}$,
+	$\mathit{Y}$) \\
+\hline
+\hline
+ \multirow{3}{*}{$\mathit{DR}$($\mathit{Y}$, $\mathit{Z}$)} &
+	\cellcolor{abfred} &
+	\cellcolor{abfred}$\mathit{DR}$($\mathit{X}$, $\mathit{Z}$) &
+	 \cellcolor{abfred} &
+	\cellcolor{abfred} &
+	\cellcolor{abfred}\multirow{3}{*}{$\mathit{DR}$($\mathit{X}$, $\mathit{Z}$)} \\
+& \cellcolor{abfred}$\mathit{T}$($\mathit{X}$, $\mathit{Z}$)&
+	\cellcolor{abfred}$\mathit{PO}$($\mathit{X}$, $\mathit{Z}$) &
+	\cellcolor{abfred}$\mathit{DR}$($\mathit{X}$, $\mathit{Z}$)&
+	\cellcolor{abfred}$\mathit{PO}$($\mathit{X}$, $\mathit{Z}$) &
+	\cellcolor{abfred}$\mathit{DR}$($\mathit{X}$, $\mathit{Z}$)\\
+ & \cellcolor{abfred}& 
+	\cellcolor{abfred}$\mathit{PP}$($\mathit{X}$,$\mathit{Z}$) & 
+	\cellcolor{abfred} &
+	$\cellcolor{abfred}\mathit{PP}$($\mathit{X}$, $\mathit{Z}$)
+	&\cellcolor{abfred}  \\
+\hline
+ \multirow{3}{*}{$\mathit{PO}$($\mathit{Y}$, $\mathit{Z}$)} &
+	\cellcolor{abfred}$\mathit{DR}$($\mathit{X}$, $\mathit{Z}$) &
+	\cellcolor{abforange} & 
+	\cellcolor{abforange} &
+	\cellcolor{abforange}$\mathit{PO}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abfred}$\mathit{DR}$($\mathit{X}$, $\mathit{Z}$) \\
+ & \cellcolor{abfred}$\mathit{PO}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abforange} $\mathit{T}$($\mathit{X}$,$\mathit{Z}$) & 
+	\cellcolor{abforange} $\mathit{PO}$($\mathit{X}$, $\mathit{Z}$)& 
+	\cellcolor{abforange}$\mathit{PPi}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abfred}$\mathit{PO}$($\mathit{X}$, $\mathit{Z}$) \\
+ & \cellcolor{abfred}$\mathit{PP}$($\mathit{X}$, $\mathit{Z}$) &
+	\cellcolor{abforange} & 
+	\cellcolor{abforange}& 
+	\cellcolor{abforange}&
+	\cellcolor{abfred}$\mathit{PP}$($\mathit{X}$, $\mathit{Z}$) \\
+\hline
+	$\mathit{EQ}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abfred}$\mathit{DR}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abforange} $\mathit{PO}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abfgreen} $\mathit{EQ}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abforange} $\mathit{PPi}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abfred}$\mathit{PP}$($\mathit{X}$, $\mathit{Z}$) \\
+\hline
+ \multirow{4}{*}{$\mathit{PP}$($\mathit{Y}$, $\mathit{Z}$)} &
+	\cellcolor{abfred}$\mathit{DR}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abforange}&
+	\cellcolor{abforange}&
+	\cellcolor{abforange}$\mathit{PO}$($\mathit{X}$, $\mathit{Z}$) &
+	\cellcolor{abfred}\\
+ & \cellcolor{abfred}$\mathit{PO}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abforange}$\mathit{PO}$($\mathit{X}$,$\mathit{Z}$) & 
+	\cellcolor{abforange}$\mathit{PP}$($\mathit{X}$, $\mathit{Z}$)& 
+	\cellcolor{abforange}$\mathit{EQ}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abfred}$\mathit{PP}$($\mathit{X}$, $\mathit{Z}$)\\
+ & \cellcolor{abfred}$\mathit{PP}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abforange}$\mathit{PP}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abforange}& 
+	\cellcolor{abforange}$\mathit{PP}$($\mathit{X}$, $\mathit{Z}$) & 
+	\cellcolor{abfred}\\
+ & \cellcolor{abfred}&\cellcolor{abforange} & \cellcolor{abforange}& $\cellcolor{abforange}\mathit{PPi}$($\mathit{X}$, $\mathit{Z}$) & \cellcolor{abfred}\\
+\hline
+ \multirow{3}{*}{$\mathit{PPi}$($\mathit{Y}$, $\mathit{Z}$)} &
+ \cellcolor{abfred}\multirow{3}{*}{} &
+ \cellcolor{abfred}$\mathit{DR}$($\mathit{X}$, $\mathit{Z}$) &
+ \cellcolor{abfred} &
+ \cellcolor{abfred}&
+ \cellcolor{abfred}\\
+& \cellcolor{abfred}$\mathit{DR}$($\mathit{X}$, $\mathit{Z}$) &
+ \cellcolor{abfred}$\mathit{PO}$($\mathit{X}$, $\mathit{Z}$) & 
+ \cellcolor{abfred}$\mathit{PPi}$($\mathit{X}$, $\mathit{Z}$)& 
+ \cellcolor{abfred}$\mathit{PPi}$($\mathit{X}$, $\mathit{Z}$)& 
+ \cellcolor{abfred}$\mathit{T}$($\mathit{X}$, $\mathit{Z}$)\\
+&\cellcolor{abfred} & \cellcolor{abfred}$\mathit{PPi}$($\mathit{X}$, $\mathit{Z}$) & \cellcolor{abfred}& \cellcolor{abfred}&\cellcolor{abfred} 
+\end{tabular}
+\end{adjustbox}
+\caption{RCC5 composition table over 3 sets.
+The results show that there exist 54 possible relations and the coloring anticipates the ideal risk matrix (green the secure state with low risk, orange the medium risk states, and red the high risk states).
+$\mathit{T}(\mathit{X},\mathit{Z}) = \{\mathit{DR}(\mathit{X}, \mathit{Z}), \, \mathit{PO}(\mathit{X},\mathit{Z}),\, \mathit{EQ}(\mathit{X},\mathit{Z}),\, \mathit{PP}(\mathit{X},\mathit{Z}),\, \mathit{PPi}(\mathit{X},\mathit{Z})\}$\label{tab:5com}}
+\end{table}
 

### 2020-03-10 formal_glossary.tex
refinement of images and conclusion of formal_glossary

@@
 for the definition (and formalization) of the cybersecurity theory in the next
-Section (Section~\ref{sec:theory}). However, the formal definition of those
-processes shall (i) correlate the formal definition of agents (i.e. the
-Definition~\ref{def:agent}) with the Causality principle (i.e. the
-formalization of the Kripke structure in Definition~\ref{def:modallogic}), and,
-most importantly, (ii) formally define a system as a structure of agents. For
-the sake of clarity, we then postpone the formalization of the two processes to
-the next Section.
+Section (Section~\ref{sec:theory}). The formal definition of those
+processes shall:
+\begin{itemize}
+	\item in Section~\ref{sec:theory}, formally define a system as a structure of agents, and
+	\item in Section~\ref{sec:process}, correlate the formal definition of agents (i.e. the
+		Definition~\ref{def:agent}) with the Causality principle (i.e.
+		the formalization of the Kripke structure in
+		Definition~\ref{def:modallogic}). 
+\end{itemize}

### 2020-03-10 theory.tex
add draft eng process

--- 2020-03-02 theory.tex
@@
 agents interacting with the system, we show that security
-doesn't seem to have a malicious nature but, rather, insecurity 
+doesn't seem to stem from a malicious nature but, rather, insecurity 
 raises from the lack of well-defined security requirements for the
@@
 of potential system configurations, which deviate from the nominal
-behavior just because the intended behavior of the system is not precisely 
-defined in the specification of the design process.
+behavior just because the \emph{intended behavior} of the system is not precisely 
+defined in the specification at the very early stages of the engineering process. 
+A reference engineering process (so called, system security lifecycle of a product)
+is described, in details, in Section~\ref{sec:process}. However,
+for the sake of simplicity, the reader can assume for this section that 
+we are currently dealing with just the relation between 
+the specification of a system and its design.
 
@@
 abstract form.
+
+\subsection{A Reference Model for Cybersecurity Engineering}\label{sec:vmodel}
 
@@
 		\begin{itemize}
-			\item the functional architecture of each agents
-			\item the structural (HW/SW) architecture of each
+			\item The functional architecture of each agents
+			\item The structural (HW/SW) architecture of each
 				subsystem of agents and agent within a
 				subsystem
-			\item the data flow as a transfer of Beliefs through
-				Assertions (i.e. the Beliefs flow)
+			\item Other important aspects of the system that we
+				describe afterwards in this paper, such as
+				the assets and security properties 
 		\end{itemize}
 \end{enumerate}
+We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}.
 
-We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}.
+We note here (and describe with more details afterwards in this paper) that the
+data flow can be defined as a transfer of Beliefs through Assertions (i.e. the
+Beliefs flow).
 
@@
 	The behavior of an agent is defined as the transformation process (e.g. defined as a
-	protocol, or a functional architecture) that determined the
+	protocol, or a functional architecture) that determines the
 	output-beliefs based on input-beliefs (where input-beliefs and
@@
 	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation21)$] $\world\models\bigcup_{s\in S}\believe{s}{P_{s}}$ and $\world\models\behave{a}{[\bigcup_{s\in S}\believe{s}{P_{s}}]}$ iff
+		\item[$(\interpretation20)$] $\world\models\bigcup_{s\in S}\believe{s}{P_{s}}$ and $\world\models\behave{a}{[\bigcup_{s\in S}\believe{s}{P_{s}}]}$ iff
 			$\world\models\belief{a}{P'}$,
@@
 	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation20)$] $\world\models\fact{a}{P}$ iff
+		\item[$(\interpretation21)$] $\world\models\fact{a}{P}$ iff
 			$\world \models \Box P$. 
@@
 
+In the quantitative evaluation of a single agent, depicted in Figure~\ref{fig:quantitative},
+we argue that only 1 configuration represents the nominal (expected) behavior 
+of the agent while the other configurations are either impossible to 
+implement or diverge from the intended nominal behavior.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.9\textwidth]{quantitative.pdf}
+	\caption{Quantitative representation of the cybersecurity risk for a single agent}
+	\label{fig:quantitative}
+\end{figure}
+
+\subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}
+While a quantitative analysis reveals how many possible configurations of an
+agent (i.e. a system) exist w.r.t. the $\abf$ theory (i.e. $54/125$ in RCC5), a
+qualitative analysis of the different configurations describes
+the configurations allowed by the $\abf$ theory, and how those configurations can be
+categorized.  In Table~\ref{tab:5com} we provide the generic composition table
+of RCC5 over 3 regions instantiated over $\abf$, which shows the whole state
+space for a single agent. The color coding of the table represents the 
+risk function as a risk matrix as we detail afterwards in this Section.
+
+A detailed qualitative categorization of agents over a theory similar to 
+the $\abf$ theory presented in this article is described in \autocite{Santaca2016abf}
+but the differences in the formalization and of the application requires
+another in-depth analysis. We categorize agents by first analyzing 
+the relations between each pair of Regions defining an agent (i.e. $\abf$),
+and then we categorize the different agents as tuple of the three Regions.
+
+\begin{enumerate}
+	\item \Rcc{$\assertionRegion$}{$\behaviorRegion$}, in order to reason
+		on the relation between assertions and behavior we first need
+		to consider that, by definition, assertions are defined as
+		transfer of information between two agents, e.g., $a$ and $b$.
+		Therefore, as depicted in Figure~\ref{fig:system-agent}, an
+		agent has two main categories of assertions, input assertions.
+		Given an agent $a$ and a collection of asserted predicates
+		$\Phi$, the Input assertions are those received
+		from a by an agent acting as a sender $s$,
+		$\assert{s}{a}{\Phi}$; similarly, output assertions are sent
+		from $a$ to a receiver $r$, $\assert{a}{r}{\Phi}$. We shall consider
+		two pairs of regions: 
+		\begin{itemize}
+			\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\behaviorRegion$}, where 
+				the relation between inputs and behavior describes the completeness
+				and correctness of the execution of the functional architecture
+				w.r.t. input elicitation. 
+			\item \Rcc{$\assertionRegion_{a\rightarrow r}$}{$\behaviorRegion$}, where
+				the relation between behavior and outputs describes the 
+		\end{itemize}
+	\item \Rcc{$\behaviorRegion$}{$\factRegion$} (Competence)
+	\item \Rcc{$\assertionRegion$}{$\factRegion$} (Honesty)
+\end{enumerate}
+
+\Paragraph{Security Requirements} The following security requirement for a CPS specification can be summarized:
+\begin{enumerate}
+	\item Proper interaction between multiple correctly-behaving agents (in
+		contrast with the ``Improper Interaction Between Multiple
+		Correctly-Behaving Entities'' defined by the CWE in the
+		categorization of ``research concepts'' in
+		\autocite{MITRE2020CWEresearch}) is defined as $\eq{\assertionRegion}{\behaviorRegion}$
+	\item The equality relation $\eq{\assertionRegion_{s\rightarrow a}}{\behaviorRegion}$
+		describes the intended secure behavior as: the beliefs generated by the behavior of the functional architecture
+		shall be complete w.r.t. the specified inputs of the agent. Therefore, \emph{the assertions received by an agent or a system
+		shall be compliant with the expected inputs of the functional architecture}. For example, the inputs
+		of the user of a SW must be sanitized to exclude deviations w.r.t. the expected inputs of the functions
+		implemented in the SW. Another example is the type checking between allowed inputs and expected inputs.
+\end{enumerate}
+
+\begin{example}
+	\begin{itemize}
+		\item https://cwe.mitre.org/data/definitions/1000.html highlight diff between CWE ``improper interaction\ldots'' and our ''Proper interatcion\ldots'' and the internal categorization of the classes and base of the subitems of this CWE
+	\end{itemize}
+\end{example}
+
+\begin{definition}{\bf Security and Insecurity of a System or an Agent --}\label{def:security}
+\end{definition}
+
+\begin{definition}{\bf Risk --}
+\end{definition}
+\begin{definition}{\bf Risk Matrix --}
+\end{definition}
+
+\begin{table}[t]
+\centering
+\begin{adjustbox}{width=\textwidth}
+\begin{tabular}{r||c|c|c|c|c} 
+& \dr{$\assertionRegion$}{$\behaviorRegion$} & 
+	\po{$\assertionRegion$}{$\behaviorRegion$}& 
+	\pp{$\assertionRegion$}{$\behaviorRegion$} &
+	\ppi{$\assertionRegion$}{$\behaviorRegion$} & 
+	\eq{$\assertionRegion$}{$\behaviorRegion$} \\
+\hline
+\hline %dr
+ \multirow{3}{*}{\dr{$\behaviorRegion$}{$\factRegion$}} & 
+	\cellcolor{abfred} & %dr
+	\cellcolor{abf-rg-1}\dr{$\assertionRegion$}{$\factRegion$} & %po
+	\cellcolor{abf-rg-2}\multirow{3}{*}{\dr{$\assertionRegion$}{$\factRegion$}} & %pp
+	\cellcolor{abf-rg-3} \dr{$\assertionRegion$}{$\factRegion$}& % ppi
+	 \cellcolor{abf-rg-4} \\ % eq
+& \cellcolor{abfred}\all{$\assertionRegion$}{$\factRegion$}& %dr
+	\cellcolor{abf-rg-1}\po{$\assertionRegion$}{$\factRegion$} & %po
+	\cellcolor{abf-rg-2}\dr{$\assertionRegion$}{$\factRegion$} & %pp
+	\cellcolor{abf-rg-3}\po{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-4}\dr{$\assertionRegion$}{$\factRegion$}\\ % e q
+ & \cellcolor{abfred}& %dr
+	\cellcolor{abf-rg-1}\pp{$\assertionRegion$}{$\factRegion$} & %po
+	\cellcolor{abf-rg-2}  & %pp
+	\cellcolor{abf-rg-3}\pp{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-4} \\ %eq
+\hline %po
+ \multirow{3}{*}{\po{$\behaviorRegion$}{$\factRegion$}} &
+	\cellcolor{abf-rg-1}\dr{$\assertionRegion$}{$\factRegion$} & %dr
+	\cellcolor{abf-rg-2} & %po
+	\cellcolor{abf-rg-3}\dr{$\assertionRegion$}{$\factRegion$} & %pp
+	\cellcolor{abf-rg-4}\po{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-5} \\ %eq
+ & \cellcolor{abf-rg-1}\po{$\assertionRegion$}{$\factRegion$} & %dr
+	\cellcolor{abf-rg-2} \all{$\assertionRegion$}{$\factRegion$} & %po 
+	\cellcolor{abf-rg-3}\po{$\assertionRegion$}{$\factRegion$} & %pp
+	\cellcolor{abf-rg-4}\ppi{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-5} \po{$\assertionRegion$}{$\factRegion$}\\%eq
+ & \cellcolor{abf-rg-1}\pp{$\assertionRegion$}{$\factRegion$} & %dr
+	\cellcolor{abf-rg-2} &  %po
+	\cellcolor{abf-rg-3}\pp{$\assertionRegion$}{$\factRegion$} & %pp
+	\cellcolor{abf-rg-4}& %ppi
+	\cellcolor{abf-rg-5}\\ %eq
+\hline %pp
+ \multirow{4}{*}{\pp{$\behaviorRegion$}{$\factRegion$}} &
+	\cellcolor{abf-rg-2}\dr{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-3}&
+	\cellcolor{abf-rg-4} & %pp
+	\cellcolor{abf-rg-5}\po{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-6}\\ %eq
+ & \cellcolor{abf-rg-2}\po{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-3}\po{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-4}\pp{$\assertionRegion$}{$\factRegion$} & %pp
+	\cellcolor{abf-rg-5}\eq{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-6}\pp{$\assertionRegion$}{$\factRegion$} \\ %eq
+ & \cellcolor{abf-rg-2}\pp{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-3}\pp{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-4} & %pp
+	\cellcolor{abf-rg-5}\pp{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-6} \\ %eq
+ & \cellcolor{abf-rg-2}&
+ 	\cellcolor{abf-rg-3} & 
+	\cellcolor{abf-rg-4}& %pp
+	\cellcolor{abf-rg-5}\ppi{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-6} \\ %eq
+\hline %ppi
+ \multirow{3}{*}{\ppi{$\behaviorRegion$}{$\factRegion$}} &
+ 	\cellcolor{abf-rg-3}\multirow{3}{*}{} &
+ 	\cellcolor{abf-rg-4}\dr{$\assertionRegion$}{$\factRegion$} &
+ 	\cellcolor{abf-rg-5}& %pp
+ 	\cellcolor{abf-rg-6}& %ppi
+ 	\cellcolor{abf-rg-7} \\ %eq
+& \cellcolor{abf-rg-3}\dr{$\assertionRegion$}{$\factRegion$} &
+ 	\cellcolor{abf-rg-4}\po{$\assertionRegion$}{$\factRegion$} & 
+ 	\cellcolor{abf-rg-5}\all{$\assertionRegion$}{$\factRegion$}& %pp
+ 	\cellcolor{abf-rg-6}\ppi{$\assertionRegion$}{$\factRegion$}& %ppi
+ 	\cellcolor{abf-rg-7}\ppi{$\assertionRegion$}{$\factRegion$}\\ %eq
+& \cellcolor{abf-rg-3} & 
+	\cellcolor{abf-rg-4}\ppi{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-5} & %pp
+	\cellcolor{abf-rg-6}& %ppi
+	\cellcolor{abf-rg-7}\\ %eq
+\hline %eq
+	\eq{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-4}\dr{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-5} \po{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-6}\pp{$\assertionRegion$}{$\factRegion$} & %pp
+	\cellcolor{abf-rg-7} \ppi{$\assertionRegion$}{$\factRegion$} & %ppi 
+	\cellcolor{abfgreen} \eq{$\assertionRegion$}{$\factRegion$}  %eq
+\end{tabular}
+\end{adjustbox}
+\caption{RCC5 composition table over 3 regions. The results show that there
+exist 54 possible relations and the coloring anticipates the ideal risk matrix
+(green the secure state with low risk, red the high risk state, and a gradient
+of medium risk states). 
+\all{$\assertionRegion$}{$\factRegion$} = \{\dr{$\assertionRegion$}{$\factRegion$}, \po{$\assertionRegion$}{$\factRegion$}, \pp{$\assertionRegion$}{$\factRegion$}, \ppi{$\assertionRegion$}{$\factRegion$}, \eq{$\assertionRegion$}{$\factRegion$}\}
+\label{tab:5com}}
+\end{table}
+
 \subsection{System Security Evolution}\label{sec:systemevoution}

### 2020-03-17 theory.tex
partial work on abf formalization

@@
 of potential system configurations, which deviate from the nominal
-behavior just because the \emph{intended behavior} of the system is not precisely 
+behavior just because the \emph{intended behavior}\fixnote{mr}{maybe ``nominal'' instead of ``intended'' is better and behavior can be confused with the
+relation between assertions and beliefs which is not what we mean} of the system is not precisely 
 defined in the specification at the very early stages of the engineering process. 
@@
 as a system. However, the same ideas revolved around for thousands of years.
-Some examples follows.
+Some relevant examples for our objective are the following.
 \begin{itemize}
@@
 a system (or a system) as in \autocite{Santaca2016abf}.  The instrument identified is sense
-perception, because a man is considered as example, but we can abstract the
-concept of instrument as any received information, where information is
-considered as defined in \autocite{Hintikka1993Information}; which can be
+perception, because a man is considered as example, but we can abstract from the
+concept of instrument and consider any received information, where information is
+considered as in \autocite{Hintikka1993Information}; which can be
 summarized by the following axiom: ``Information is specified by specifying
@@
 the security of a system is tightly related to the difference between
-the nominal behavior and the implemented one.
+the nominal behavior and the actual (e.g. the implemented) one.
 
@@
 \end{definition}
-A system state space can be seen as a superposition of multiple agent states where knowledge,
+Therefore, a system state space can be seen as a superposition of multiple agent states where knowledge,
 beliefs, and information are related between each other (in a mereotopological
-space) through the all the admissible RCC relations. 
-A \emph{system state} (as an instance of the system state space)
+space) through all the admissible RCC relations. 
+\fixnote{mr}{are the next lines redundant?}A \emph{system state} (as an instance of the system state space)
 is a specific configuration of
@@
 belief can be precisely formalized from an epistemic standpoint.  We assume
-that a semantic of a correct definition of epistemic knowledge exists, for
+that a semantic of a correct (i.e. commonly believed to be true) definition of epistemic knowledge exists, for
 example the one given in\autocite{Hintikka1962knowledge} by Hintikka, and we
@@
 			\item The functional architecture of each agents
-			\item The structural (HW/SW) architecture of each
+			\item The physical/structural (HW/SW) architecture of each
 				subsystem of agents and agent within a
@@
 	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation19)$] $\world\models\assert{a}{b}{P}$ iff
+		\item[$(\interpretation19)$] $\world\models\rassert{a}{b}{P}$ iff
 			$\world\models\belief{a}{P}$, 
+			$\world\models\neg\belief{a}{\neg P}$, 
 			$\world'\models\rcc(\belief{a}{P},\belief{b}{P})$, 
-			$\world'\models\belief{a}{\belief{b}{P}}$
-			and $\world'\models P$ for all
-			$\world'$ such that $\world\modalrelation\world'$.
+			for all $\world'$ such that $\world\modalrelation\world'$.
 	\end{enumerate}
@@
 	protocol, or a functional architecture) that determines the
-	output-beliefs based on input-beliefs (where input-beliefs and
+	output-beliefs based on input-beliefs and vice versa (where input-beliefs and
 	output-beliefs are beliefs taken as input or output respectively).
 	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation20)$] $\world\models\bigcup_{s\in S}\believe{s}{P_{s}}$ and $\world\models\behave{a}{[\bigcup_{s\in S}\believe{s}{P_{s}}]}$ iff
+		\item[$(\interpretation20)$] $\world\models\bigcup_{\Phi}\believe{s}{\varphi}$ and $\world\models\behave{a}{[\bigcup_{s\in S}\believe{s}{P_{s}}]}$ iff
 			$\world\models\belief{a}{P'}$,
@@
 	$s=\langle\rcc(\factRegion,\behaviorRegion),\rcc(\factRegion,\assertionRegion),\rcc(\behaviorRegion,\assertionRegion)\rangle$,
-	where $\assertionRegion,\behaviorRegion,\factRegion$, are regions of assertions, behavior, and facts respectively.
+	where $\assertionRegion,\behaviorRegion,\factRegion$, are regions of assertions, behavior (i.e. the beliefs generated by the behavior), and facts respectively.
 \end{definition}
@@
 
+As described beforehand in this section, the collection of Facts shall define
+the functional and the physical architecture of an agent, but Assets and
+security properties are defined as Facts. Given that understanding
+why an Asset is considered as such by the user doesn't necessarily allow
+us to better qualify an Asset, we consider the fact of ``being an asset'' as
+a Boolean property of an Agent.
+
+\begin{definition}{\bf Asset --}\label{def:asset}
+	\fix{mr}{asset def.}
+\end{definition}
+
+The definition of the security properties is given in Section~\ref{sec:properties}
+
 \subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}
@@
 
-A detailed qualitative categorization of agents over a theory similar to 
-the $\abf$ theory presented in this article is described in \autocite{Santaca2016abf}
-but the differences in the formalization and of the application requires
-another in-depth analysis. We categorize agents by first analyzing 
-the relations between each pair of Regions defining an agent (i.e. $\abf$),
-and then we categorize the different agents as tuple of the three Regions.
+A detailed qualitative categorization of agents over a theory similar to the
+$\abf$ theory presented in this article is described in
+\autocite{Santaca2016abf} but the differences in the formalization and of the
+application requires another in-depth analysis. As depicted in
+Figure~\ref{fig:soundness}, the relation between Facts, and Assertions and Behavior
+defines the soundness of the design (admissible configurations of Assertions,
+Behaviors and, in turn, Beliefs) w.r.t. the specification (what the
+specification mandates, such as by the nature of the physical/functional
+architectural models). Similarly, the relation between Assertions and Behavior defines the
+soundness of the functional architecture.
+We categorize agents by first analyzing the relations
+between each pair of Regions defining an agent (i.e. $\abf$), and then we
+categorize the different agents as tuple of the three Regions.
+For the sake of simplicity, soundness is opposed to non-soundness in the following, however,
+with the RCC we consider different ``degrees'' of soundness. For example,
+in RCC5, if we consider EQ between two Regions as representing soundness, DR over the same Regions
+represents non-soundness while PP, PO, PPi represents the different degrees of soundness.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.8\textwidth]{soundness.pdf}
+	\caption{Example relation between Facts, and Assertions and Beliefs}
+	\label{fig:soundness}
+\end{figure}
 
@@
 		Therefore, as depicted in Figure~\ref{fig:system-agent}, an
-		agent has two main categories of assertions, input assertions.
+		agent has two main categories of assertions, input and output assertions.
 		Given an agent $a$ and a collection of asserted predicates
 		$\Phi$, the Input assertions are those received
-		from a by an agent acting as a sender $s$,
-		$\assert{s}{a}{\Phi}$; similarly, output assertions are sent
-		from $a$ to a receiver $r$, $\assert{a}{r}{\Phi}$. We shall consider
+		by $a$ from an agent $s$ acting as a sender,
+		$\rassert{s}{a}{\Phi}$; similarly, output assertions are sent
+		from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
 		two pairs of regions: 
 		\begin{itemize}
-			\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\behaviorRegion$}, where 
-				the relation between inputs and behavior describes the completeness
-				and correctness of the execution of the functional architecture
-				w.r.t. input elicitation. 
-			\item \Rcc{$\assertionRegion_{a\rightarrow r}$}{$\behaviorRegion$}, where
-				the relation between behavior and outputs describes the 
+			\item \Rcc{$\assertionRegion_{s\rightarrow
+				a}$}{$\behaviorRegion$}, where the relation
+				between Input-assertions and behavior describes
+				the soundness of the execution of the
+				functional architecture w.r.t. input
+				elicitation. With more details, the ideal
+				specification of the functional architecture,
+				along with the expected inputs, defines the
+				functional behavior of an ideal system.  If all
+				the inputs (Assertions) are correctly handled
+				in the functional specification (behavior) the
+				specification is sound.
+
+			\item \Rcc{$\assertionRegion_{a\rightarrow
+				r}$}{$\behaviorRegion$}, where the relation
+				between behavior and outputs describes the
+				completeness of the behavior defined in the
+				specification w.r.t. the input elicitation.
+				With more details, if all the outputs
+				(assertions) of the functional architecture can
+				be produced, the functional architecture is
+				complete.
 		\end{itemize}

### 2020-03-24 theory.tex
first complete draft abf-theory

@@
 \subsection{A Reference Model for Cybersecurity Engineering}\label{sec:vmodel}
+In order to describe the $\abf$-theory we briefly introduce the 
+engineering concepts (detailed and extended in Section~\ref{sec:engineering})
+underlying the $\abf$-theory; namely, the first three stages of the, so called,
+engineering V-Model. Our definition is partial and we only use it for the 
+sake of simplicity and to better present the theory.
 
-\subsection{Abstract System State}\label{sec:systemstate}
+The development process of a new product (a CPS in our case) 
+should follow an engineering model. Our reference engineering process 
+(inspired by the waterfall V-Model\autocite{waterfallvmodel}), as depicted in
+Figure~\ref{fig:simplifiedvmodel}, is structured as the following ordered list
+of abstraction refinement steps.
+\begin{enumerate}
+	\item The process starts with the definition of a \emph{specification}
+		where the general requirements of the CPS are defined (e.g.
+		functional, physical, network, and design).  As an example, an
+		RFC of a communication or security protocol can be considered
+		as the result of this first phase (e.g. see
+		RFC-1\autocite{rfc1}.  In this first phase, we consider the
+		architectural aspects of a CPS such as the draft physical and
+		functional architectures.
+	\item The process then continues with the definition of a design where
+		the requirements of the previous phase holds; detailing the
+		logic of the physical (e.g. the physical-layer encoding of the
+		information) and functional architectures (e.g. the protocol
+		logic of communication over a number of potential sessions).
+	\item Finally, the design is implemented as hardware (HW) for the physical architecture
+		and software (SW) for the functional architecture.
+\end{enumerate}
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\textwidth]{vmodel.pdf}
+	\caption{Cybersecurity Engineering Life-cycle}
+	\label{fig:simplifiedvmodel}
+\end{figure}
+
+\subsection{Abstract System}\label{sec:systemstate}
 The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
@@
 
-\subsection{Operational System State}\label{sec:engsystemstate}
+\subsection{Operational System}\label{sec:engsystemstate}
 When dealing with a specification of a system (i.e. the design if the system
@@
 \begin{definition}{\bf Assertion -- }\label{def:assertion}
-	An assertion is an intended transfer of beliefs between
+	An assertion (\fix{mr}{subtype of Information}) of  is an intended transfer of beliefs between
 	two agents $a$ and $b$ such that
@@
 \begin{definition}{\bf Behavior -- }\label{def:behavior}
-	The behavior of an agent is defined as the transformation process (e.g. defined as a
+	The behavior (\fix{mr}{subtype of Beliefs}) of an agent is defined as the transformation process (e.g. defined as a
 	protocol, or a functional architecture) that determines the
@@
 \begin{definition}{\bf Fact -- }\label{def:fact}
-Facts are a sub-region of Knowledge $\factRegion = \bigcup_\Phi \knows{a}{\varphi}$
+	Facts are a \fix{mr}{sub-region/subtype} of Knowledge $\factRegion = \bigcup_\Phi \knows{a}{\varphi}$
 	that predicate in a factive way, e.g.
@@
 of the agent while the other configurations are either impossible to 
-implement or diverge from the intended nominal behavior.
+implement or diverge from the intended nominal behavior. We note 
+that, the numbers reported here do not consider the details of the
+engineering process and should be considered a limit of an abstract 
+representation of the system. In Section~\ref{sec:engineering} we
+detail this numeric evaluation to faitfully represent the
+number of insecurity configurations.
 
@@
 the functional and the physical architecture of an agent, but Assets and
-security properties are defined as Facts. Given that understanding
+security properties are defined as Facts, i.e. true. Given that understanding
 why an Asset is considered as such by the user doesn't necessarily allow
 us to better qualify an Asset, we consider the fact of ``being an asset'' as
-a Boolean property of an Agent.
+a predicate \fix{mr}{(or a true function)} of a component of a system.
 
 \begin{definition}{\bf Asset --}\label{def:asset}
-	\fix{mr}{asset def.}
+For any agent $a\in Ag$ in a collection of Agents $Ag$, an agent is considered an asset iff 
+	\begin{enumerate}[noitemsep]
+		\item[$(\interpretation22)$] $\world\models\asset{a}$ iff
+			$\world\models \Box\asset{a}$.
+	\end{enumerate}
 \end{definition}
@@
 space for a single agent. The color coding of the table represents the 
-risk function as a risk matrix as we detail afterwards in this Section.
+risk function as a risk matrix as we detail afterwards in this section.
 
 A detailed qualitative categorization of agents over a theory similar to the
-$\abf$ theory presented in this article is described in
-\autocite{Santaca2016abf} but the differences in the formalization and of the
+$\abf$ one, is presented in this article and described in
+\autocite{Santaca2016abf}. The differences in the formalization and of the
 application requires another in-depth analysis. As depicted in
@@
 specification mandates, such as by the nature of the physical/functional
-architectural models). Similarly, the relation between Assertions and Behavior defines the
-soundness of the functional architecture.
-We categorize agents by first analyzing the relations
+architectural models). We categorize agents by first analyzing the relations
 between each pair of Regions defining an agent (i.e. $\abf$), and then we
@@
 
-\begin{enumerate}
-	\item \Rcc{$\assertionRegion$}{$\behaviorRegion$}, in order to reason
-		on the relation between assertions and behavior we first need
-		to consider that, by definition, assertions are defined as
-		transfer of information between two agents, e.g., $a$ and $b$.
-		Therefore, as depicted in Figure~\ref{fig:system-agent}, an
-		agent has two main categories of assertions, input and output assertions.
-		Given an agent $a$ and a collection of asserted predicates
-		$\Phi$, the Input assertions are those received
-		by $a$ from an agent $s$ acting as a sender,
-		$\rassert{s}{a}{\Phi}$; similarly, output assertions are sent
-		from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
-		two pairs of regions: 
+\Paragraph{\Rcc{$\assertionRegion$}{$\behaviorRegion$} -- Collaboration}
+In order to reason on the relation between assertions and behavior we first
+need to consider that, by definition, assertions are defined as transfer of
+information between two agents, e.g., $a$ and $b$.  Therefore, as depicted in
+Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
+input and output assertions.  Given an agent $a$ and a collection of asserted
+predicates $\Phi$, the Input assertions are those received by $a$ from an agent
+$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
+are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
+two pairs of regions\footnote{``I am not asserting, as Lotze did, that a relation between $X$ and $Y$
+consists of a quality in $X$ and a quality in $Y$ -- a view which I regard as
+quite indefensible. -- I assert that a relation $Z$ between $X$ and $Y$ involves
+the existence in $X$ of the quality ``having the relation $Z$ to $Y$'' so that
+a difference of relations always involves a difference in quality, and a
+change of relations always involves a change of quality.'' --  Ellis J. McTaggart, The Unreality of Time\autocite{Mctaggart1908unreality}}: 
 \begin{itemize}
-			\item \Rcc{$\assertionRegion_{s\rightarrow
-				a}$}{$\behaviorRegion$}, where the relation
-				between Input-assertions and behavior describes
-				the soundness of the execution of the
-				functional architecture w.r.t. input
-				elicitation. With more details, the ideal
-				specification of the functional architecture,
-				along with the expected inputs, defines the
-				functional behavior of an ideal system.  If all
-				the inputs (Assertions) are correctly handled
-				in the functional specification (behavior) the
+	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\behaviorRegion$},
+		where the relation between Input-assertions and behavior
+		describes the soundness of the execution of the functional
+		architecture w.r.t. input elicitation. With more details, the
+		ideal specification of the functional architecture, along with
+		the expected inputs, defines the functional behavior of an
+		ideal system.  If all the inputs (Assertions) are correctly
+		handled in the functional specification (behavior) the
 		specification is sound.
 
-			\item \Rcc{$\assertionRegion_{a\rightarrow
-				r}$}{$\behaviorRegion$}, where the relation
-				between behavior and outputs describes the
-				completeness of the behavior defined in the
-				specification w.r.t. the input elicitation.
-				With more details, if all the outputs
-				(assertions) of the functional architecture can
-				be produced, the functional architecture is
-				complete.
+	\item \Rcc{$\assertionRegion_{a\rightarrow r}$}{$\behaviorRegion$},
+		where the relation between behavior and outputs describes the
+		completeness of the behavior defined in the specification
+		w.r.t. the input elicitation.  With more details, if all the
+		outputs (assertions) of the functional architecture can be
+		produced, the functional architecture is complete.
 \end{itemize}
-	\item \Rcc{$\behaviorRegion$}{$\factRegion$} (Competence)
-	\item \Rcc{$\assertionRegion$}{$\factRegion$} (Honesty)
-\end{enumerate}
 
-\Paragraph{Security Requirements} The following security requirement for a CPS specification can be summarized:
+\Paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and \Rcc{$\behaviorRegion$}{$\factRegion$} -- Honesty and Competence}
+While Assertions and Beliefs generated by the Behavior determines how a system should work, and the
+relation between the two defines the quality (i.e. the completeness w.r.t the specification) of the agent, the relation of
+Assertions and Behavior with Facts determines the quality with respect to the
+nominal (specified) system.  Given that Facts defines what is true in the
+system (e.g. the level of water in a tank may be considered as a Fact, while
+the reading of a sensor as a Belief), the relation of Assertions an Facts
+determines the degree of quality between the information circulating in a
+system (or within an agent) and reality.  Since the transfer of information through Assertions
+generates Beliefs, a dishonest agent may circulate false information,
+generating false Beliefs.  We note that it is often implied that the intention behind
+circulating false information discerns a dishonest and an incompetent agent,
+however, we consider honesty related to truth rather than related to what is
+true\footnote{``[\ldots] truth exists only in the good man, but the true in the bad
+man as well; for it is possible for the bad man to utter something true.''
+Sextus Empiricus, Outlines of Pyrrhonism, II-83\autocite{Empiricus1990Pyrrhonism}}.  This implies that the
+relation between Beliefs and Facts determines the competence (on the subjected
+defined by the Facts) of an agent (i.e. the more competent an agent is, the
+more likely a Belief of that agent is true).
+
+\subsubsection{$\abf$ Security Enumeration (SE)} The following security requirement for a CPS specification can be summarized:
 \begin{enumerate}
-	\item Proper interaction between multiple correctly-behaving agents (in
+	\item[SE-$1$] Proper interaction between multiple correctly-behaving agents (in
 		contrast with the ``Improper Interaction Between Multiple
-		Correctly-Behaving Entities'' defined by the CWE in the
-		categorization of ``research concepts'' in
-		\autocite{MITRE2020CWEresearch}) is defined as $\eq{\assertionRegion}{\behaviorRegion}$
-	\item The equality relation $\eq{\assertionRegion_{s\rightarrow a}}{\behaviorRegion}$
+		Correctly-Behaving Entities'' defined by the CWE--$435$ as one of the top ``view'' of the ``research concepts'' in
+		\autocite{MITRE2020CWEresearch}) is defined as $\eq{\assertionRegion_a}{\behaviorRegion_a}$ for an agent $a$
+	\begin{enumerate}
+		\item[SE-$1.1$] The equality relation $\eq{\assertionRegion_{s\rightarrow a}}{\behaviorRegion_a}$
 			describes the intended secure behavior as: the beliefs generated by the behavior of the functional architecture
@@
 			implemented in the SW. Another example is the type checking between allowed inputs and expected inputs.
+		\item[SE-$1.2$] Similarly, the equality relation $\eq{\assertionRegion_{a\rightarrow r}}{\behaviorRegion_a}$ defines that
+			the outputs of an agent $a$ shall be the outputs of the functional architecture.
+	\end{enumerate}
+	\item[SE-$2$]Sufficient Control Flow Management (in contrast with
+		the ``Insufficient Control Flow Management'' defined by MITRE
+		in the CWE--$691$ as one of the top ``view'' of the ``research
+		concepts'' in\autocite{MITRE2020CWEresearch}) is defined as
+		$\eq{\assertionRegion}{\factRegion}$.
+	\item[SE-$3$] Correct Calculation (in contrast with the ``Improper
+		Calculation'' defined by MITRE in the CWE--$682$ as one of the
+		top ``view'' of the ``research concepts''
+		in\autocite{MITRE2020CWEresearch}) is defined as
+		$\eq{\beliefRegion}{\factRegion}$.
 \end{enumerate}
 
-\begin{example}
-	\begin{itemize}
-		\item https://cwe.mitre.org/data/definitions/1000.html highlight diff between CWE ``improper interaction\ldots'' and our ''Proper interatcion\ldots'' and the internal categorization of the classes and base of the subitems of this CWE
-	\end{itemize}
-\end{example}
-
-\begin{definition}{\bf Security and Insecurity of a System or an Agent --}\label{def:security}
+We are now in the position to define what a secure system (with respect to the $\abf$-theory) is
+and, based on that definition, what the security risk is and how to quantify it in a risk matrix.
+\begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
+	A secure system is a system where SE-$1$, SE-$2$, and SE-$3$ holds for each 
+	agent composing the system.
 \end{definition}
-
+The ISO 31000 consider risk as the ``effect of uncertainty on objectives'' and the refers
+both of positive and negative consequences of uncertainty. Accordingly, we consider
+risk as follows. 
 \begin{definition}{\bf Risk --}
+The whole space of potential designs of a specification with respect to the
+$\abf$-theory.
 \end{definition}
+The definition of Risk leads to the risk matrix in Figure~\autoref{fig:quantitative}, defined as follows.
 \begin{definition}{\bf Risk Matrix --}
+	The risk matrix is a function between the three relations 
+	$s=\langle\rcc(\factRegion,\behaviorRegion),\rcc(\factRegion,\assertionRegion),\rcc(\behaviorRegion,\assertionRegion)\rangle$,
+	where the maximum risk is defined by the $DR$ relation between the three groups of Regions, and the 
+	minimum risk by the $EQ$ relation over the same Regions. In between the two extremes, the granularity
+	of possible intermediate configuration is defined by the calculus used (RCC5 in our case).
 \end{definition}
+We note that often the risk matrix is defined by the relation between likelihood and impact.
+We argue that the likelihood is related to the number of possible non secure (i.e. non maximum secure) 
+configurations of a system and, similarly, the impact.
 
@@
 belief (as a result of the behavior execution) of not the opposite. But, as
-Hintikka States in "popper" our previous statement is true if the probability
+Hintikka States in ``popper'' our previous statement is true if the probability
 distribution of the information is representative wrt the causality structure

### 2020-03-24 theory.tex
remove notes on system security evolution

@@
 \end{table}
-
-\subsection{System Security Evolution}\label{sec:systemevoution}
-\fix{mr}{
-Facts should be correlated to the causality principle. Therefore a
-specification in opnet should be translated into vmt and by bounded model
-checking with nuxmv identify a path towards an undesired state, if that fails
-use abf to mutate the first state into one of the possible according to abf and
-re-run nuxmv.
-
-Knowledge is the theory itself, an instance is the facts of a design of it. A
-testing process tests facts against knowledge as much as we test a cps design
-choices against its specification.  Information carries the knowledge of the
-negation of the opposite of the information. Therefore an assertion carries the
-belief (as a result of the behavior execution) of not the opposite. But, as
-Hintikka States in ``popper'' our previous statement is true if the probability
-distribution of the information is representative wrt the causality structure
-(i.e. knowledge and facts, see email before). So, based on our theory, the
-assertions should have a number between 3 to 5 possible negated meaning, one
-per each rcc if reference and not just a negation of the information. Those
-negation representative of an insecure configuration on the communication of
-the system. In turn, those configurations should categorize all the CWE.  The
-fact that weaknesses are generated by mankind but not exploited makes malicious
-attacks confined to the hackers (and black-hat hackers). Fix: move black-hat
-hackers as subregion of mankind with malicious intent. Describe the intent as a
-desire of reaching a breach of a cybersecurity property over an asset (agent of
-the system), instead of searching for a specific attack. The tool would then
-translate the system from opnet to abf formula. The abf tool generates all the
-sat state of the system fix: add state to all def before this text. And we
-shall define the causality relation between the different states generated by
-the abf tool as chosen nondeterministically from the possible sat states by a
-model checker such as nuxmv.
-
-By calculating sat instead of validity with abf tool and then adding it to the
-causality structure of a system fact description, e.g. in nuxmv; we can
-calculate a path for reachability of an insecurity property violated on an
-agent which represents an asset. Whenever nuxmv fails, we can add a new state
-and re-run the process. Those guesses should be better managed by a trained
-v-ai since the inference by induction from the partial information of the
-system in which we represent the assertions as negation of a beliefs and
-without rcc do not carry enough information for a Reasoner that doesn't
-implemented rcc.
-}

### 2020-03-24 abstract.tex
minor fix in abstract

@@
 mereology, and topological structure. We argue that a mereology is the
-most appropriate abstract underlying structure, do to its generality,
+most appropriate abstract underlying structure, due to its generality,
 for defining the expressiveness of the system abstraction.  Furthermore, a

### 2020-03-25 engineering.tex
qualitative evaluation of agents (and isk matrix)
add picture of a SPDL
first complete draft abf-theory
spdl description, use case and requirements draft

@@
-\epigraph{The difference between ideas and reality is the
+\epigraph{The difference between ideas and reality is like the
 difference between philosophy and engineering. The work to transform one into
 the other is scientific research}{{\itshape V-Research}}
+
+Several standards mandates a secure-by-design approach in which 
+cybersecurity shall be considered at the very early stages of the 
+design process. For example,
+\begin{enumerate}[noitemsep]
+	\item DO-326A -- ``Airworthiness Security Process Specification''
+		requires a cybersecurity risk assessment of the design and is
+		the ``are the only Acceptable Means of Compliance (AMC) by FAA
+		\& EASA for aviation cyber-security airworthiness certification,
+		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
+	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
+		Control System (ICS) Security''
+	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
+		for Cyber-Physical Vehicle Systems'' defines ``set of
+		high-level guiding principles for Cybersecurity as it relates
+		to cyber-physical vehicle systems'' and states that
+		``incorporate Cybersecurity into cyber-physical vehicle systems
+		from concept phase through production, operation, service, and
+		decommissioning''
+\end{enumerate}
+Companies who develop CPS (e.g. aerospace systems or elevator systems), must adhere
+to strict regulations and quality assurances processes. Introducing tools and procedures 
+in their processes requires a detailed and justified overview on how those
+tools and procedures can be used. In Figure~\ref{fig:spdl}, we provide an overview
+of a process that consider cybersecurity in the specification, design, and implementation
+stages. The overall process start with the specification of the architectural (both physical
+and functional) requirements from which the Weaknesses are automatically identified.
+Based on the Assets, architectures, and the Weaknesses the cybersecurity risk is
+calculated based on the $\abf$-theory. An iterative process between 
+the user (e.g. the engineer) and the \secramod module of the \abftool allows the user to move to the design phase
+when the risk level is considered acceptable. The design phase starts by automatically mapping
+the Weaknesses into Vulnerabilities which are, in turn, fed into the \designverifmod
+module along with the final specification. The Module returns a list of Vulnerabilities and
+potential Mitigations. The user then procees in implementing the CPS and the SW/HW choices
+can be tested by the \atgmod.
+
 \begin{figure}[t]
 	\centering
-	\includegraphics[width=\textwidth]{vmodel.pdf}
-	\caption{Cybersecurity Engineering Life-cycle}
-	\label{fig:knowledge-belief}
+	\includegraphics[width=\textwidth]{spdl.pdf}
+	\caption{Secure-by-design System Development Life-cycle}
+	\label{fig:spdl}
 \end{figure}
 
-Infosec, or Information Security, aims at mitigating the security risks related
-to information. The standard de-facto is the so called CIA triad 
-(where CIA stands for Confidentiality, Integrity, Availability), which is often
-criticized for being too general or non-adequate (e.g. by adding authenticity which is 
-often used as a building block for confidentiality and integrity) 
-to be effectively applied to the engineering of systems. Due to this,
-many researchers an organization tried to improve the CIA triad. 
-The evolutions/extensions of the CIA triad has been documented, e.g.,
-in \autocite{Samonas2014cia} and is summarized in Table~\ref{tab:ciaevolution}.
+In order to explain in details the SPDL, we use the following running example.
+\begin{example}
+	As depicted in Figure~\ref{fig:usecase}, the use case of the running example
+	(a small part of a subprocess of the SwAT testbed \autocite{}), shows
+	\begin{itemize}[noitemsep]
+		\item a tank that is filled with raw water coming from the inlet pipe 
+		\item a motorized valve (a valve with an actuator) that regulates that
+			\begin{itemize}
+				\item if opened, allows the raw water to flow into the tank
+				\item if closed, stops the raw water to flow into the tank
+			\end{itemize}
+		\item a sensor that reads the meters of raw water in the tank
+		\item a PLC controller, connected to the sensor and the actuator, that receives the readings from the sensor and regulates the quantity of water in the tank by communicating a change in the state of the actuator based on the following logi
+			\begin{itemize}
+				\item if the readings from the sensor reports a water level in the tank above (or equal) to 10m, the actuator shall close the valve
+				\item if th readings are below 10m the actuator shall open the valve.
+			\end{itemize}
+	\end{itemize}
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.55\textwidth]{usecase.pdf}
+	\caption{Use Case -- Running Example}
+	\label{fig:usecase}
+\end{figure}
+\end{example}
+
+\subsection{Specification -- Elicitation of Security Requirements}\label{sec:properties}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\textwidth]{spec.pdf}
+	\caption{Specification -- Running Example}
+	\label{fig:spec}
+\end{figure}
+The first step of the SPDL is the definition of the requirements and then, for
+the purpose of this paper of the (i) security requirements, (ii) architectural
+(physical and functional) requirements.  Our focus in mainly on security
+requirements but they are necessarily intertwined with the architectural
+requirements, as we will show afterwards in this section.  A complete study on
+how to properly define the architectural requirements is outside the scope of
+this paper so we propose a method that is convenient for the purpose of our
+assessment.
+
+\Paragraph{System Requirements}
+In our running example, there is a system (a CPS) composed by $5$ main
+subsystem which we consider atomic (i.e. not themself composed by subsystems)
+and then agents: the tank, the sensor, the controller, the actuator, and the
+motorized valve (valve from now on).  The whole specification (which we now
+describe) of the physical and functional architecture is depicted in
+Figure~\ref{fig:spec}.  In the use case, the communication flow is as follows:
+\begin{enumerate}
+	\item the sensor reads the level of water in the tank ($\text{tank}\rightarrow \text{sensor}$) \footnote{this communication depends on the technology in use for the readings.
+We assume a unidirectional communication without loss of generality}
+	\item the sensor communicates the readings to the controller ($\text{sensor}\rightarrow \text{controller}$)
+	\item the controller calculates the valve state and communicates it to the actuator ($\text{controller}\rightarrow \text{actuator}$)
+	\item the actuator translates (digital to analog) the states received and communicates\footnote{we do not distinguish, for the sake of simplicity between different types of channels and assume that communicating to the valve produces the expected change of physical state} it to the valve
+\end{enumerate}
+Each agent has an input or output port from which it communicates with other
+agents.  It is also composed by the functional blocks that defines its
+functional architecture and physical boundaries that defines its physical
+architecture, as follows.
+\begin{enumerate}[start=1, label={req\arabic*)}]
+	\item the tank has no functional architecture, it is a purely physical component which has the only purpose of containing water
+	\item the communication is unidirectional from the tank to the input port of the sensor
+	\item the sensor has an input-port that receives only incoming communication from the tank
+	\item the sensor has two functional blocks
+		\begin{enumerate}[label={req4.\arabic*)}]
+			\item read\_water\_level, accepts inputs only from the input-port and outputs a new reading every $t$ seconds
+			\item comm\_water\_level, accepts inputs only from the read\_water\_level and communicates the readings to the output-port
+		\end{enumerate}
+	\item the sensor has an output-port that only outputs the messages generated by the comm\_water\_level to the input-port of the controller
+	\item the controller has in input-port that only receives incoming communications from the sensor and sends the messages to the functional architecture
+	\item the functional architecture of the controller is composed by three functional blocks
+		\begin{enumerate}[label={req7.\arabic*)}]
+			\item get\_water\_level, accepts inputs only from the input-port and outputs the water level to the main control function
+			\item control, calculate the next state of the valve based on the inputs receive from the get\_water\_level
+			\item send\_state, receives the output of the control function and communicates it to the output-port
+		\end{enumerate}
+	\item the controller has an output-port that only outputs the messages generated by the send\_state to the input-port of the actuator 
+	\item the actuator has an input-port that receives only incoming communication from the controller 
+	\item the functional architecture of the actuator is composed by two functional blocks
+		\begin{enumerate}[label={req10.\arabic*)}]
+			\item read\_change\_state, accepts inputs only from the input-port and outputs the next state of the valve to the close\_open\_valve function
+			\item close\_open\_valve, converts the digital next state received from the read\_change\_state an analog signal and communicates it the output-port
+		\end{enumerate}
+	\item the valve accepts analog messages from the output port of the actuator and changes its state based on those messages
+\end{enumerate}
+
+While \emph{agents} and \emph{communication} between agents are described and
+defined in Section~\ref{sec:theory}, both the concepts of \emph{port},
+\emph{channel}, \emph{functional and physical architecture}, and
+\emph{functional block} have not yet been defined with respect to the $\abf$-theory.
+Informally, a channel identifies the place where the communication takes place.
+In this work we only consider mono-directional channels and communication but the 
+extension to bi-directional channel can be considered as the union of two
+unidirectional channels.
+
+\begin{definition}{\bf mono-directional channel --}\label{def:monochannel}
+	A mono-directional channel is defined by the message sent and received
+	over the channel. Given two agents $s$ (sender) and $r$ (receiver) a
+	mono-directional channel $s \rightarrow r$ defines the sending of
+	information between $s$ and $r$. 
+	%$\assertion{s}{r}$ is defined as a (improper) subset of the assertions of $s$, i.e.
+	%$\pp{\assertion{s}{r}}{\assertion{s}}
+	%PP(A λ S ,λ R , A λ S ) ∨ EQ(A λ S ,λ R , A λ S ) (since λ S may
+	%communicate with other agents).  The receipt of messages over a
+	%mono-directional channel is defined as the topolog- ical boundary B S
+	%,R (improper) subset of the beliefs of the receiver (i.e., the beliefs
+	%generated by the message exchange and by computations over the set of
+	%beliefs): PP(B S ,R , B R ) ∨ EQ(B S ,R , B R ).
+\end{definition}
+
+\begin{definition}{\bf \fix{mr}{informal Input or Output Port} --}\label{def:port} 
+	Ports are agents with the following predefined behavior: an input-port
+	changes the type of information from assertion to belief while an
+	output-port from belief to assertion; forwarding information from the
+	outside of an agent's boundary to the inside (input-port) or vice versa
+	(output-port).
+\end{definition}
+
+\Paragraph{Security Requirements} 
+While standards (such
+as IEC62443-1-3\fixnote{mr}{check the correct version}) defines requirements as
+``information in transit/at-rest should be encrypted'' we believe that, at
+specification level, security requirements can be formulated as subcategories of
+the well-known CIA triad (where CIA stands for Confidentiality, Integrity,
+Availability).
+Infosec, or Information Security, defines the security risks related to
+information with the de-facto standard CIA triad , which is often criticized
+\autocite{CIAcriticismCPS} for being too general or non-adequate (e.g. by
+adding authenticity which is often used as a building block for confidentiality
+and integrity) to be effectively applied to the engineering of systems. Due to
+this, many researchers an organization tried to improve the CIA triad.  The
+evolutions/extensions of the CIA triad has been documented, e.g., in
+\autocite{Samonas2014cia} and is summarized in Table~\ref{tab:ciaevolution}.
 
@@
 
-An example is the effort made by the OECD (Organisation for Economic
+An representative example is the effort made by the OECD (Organisation for Economic
 Co-operation and Development) in \autocite{OECD2013guidelines} to define
@@
 argument is that, to the best of our knowledge, all of the approaches that aim
-at improving the CIA triad motivated by empirical evidences. On the contrary,
+at improving the CIA triad are besed on motivations related to empirical evidences. On the contrary,
 we now define (in Section~\ref{sec:properties}) the CIA triad for an agent
@@
 
-\subsection{Elicitation of Security Requirements}\label{sec:properties}
 In \autocite{Anderson1972report,Samonas2014cia}, CIA are defined as follows.
@@
 
+\fix{mr}{
 Confidentiality: the information is accessible only to authorised individuals.
@@
 
-\subsubsection{Relevant Standards}\label{sec:standards}
-\begin{enumerate}[noitemsep]
-	\item DO-326A
-\end{enumerate}
-
 \begin{enumerate}
@@
 \end{enumerate}
+}

### 2020-03-25 engineering.tex
def channel and port

@@
 \emph{channel}, \emph{functional and physical architecture}, and
-\emph{functional block} have not yet been defined with respect to the $\abf$-theory.
-Informally, a channel identifies the place where the communication takes place.
-In this work we only consider mono-directional channels and communication but the 
-extension to bi-directional channel can be considered as the union of two
-unidirectional channels.
+\emph{functional block} have not yet been defined with respect to the
+$\abf$-theory.  Informally, a channel identifies the place where the
+communication takes place.  In this work we only consider mono-directional
+channels and communication but the extension to bi-directional channel can be
+considered as the union of two unidirectional channels. A mono-directional
+channel is defined by the assertions sent or received (over the channel).
 
 \begin{definition}{\bf mono-directional channel --}\label{def:monochannel}
-	A mono-directional channel is defined by the message sent and received
-	over the channel. Given two agents $s$ (sender) and $r$ (receiver) a
-	mono-directional channel $s \rightarrow r$ defines the sending of
-	information between $s$ and $r$. 
-	%$\assertion{s}{r}$ is defined as a (improper) subset of the assertions of $s$, i.e.
-	%$\pp{\assertion{s}{r}}{\assertion{s}}
-	%PP(A λ S ,λ R , A λ S ) ∨ EQ(A λ S ,λ R , A λ S ) (since λ S may
-	%communicate with other agents).  The receipt of messages over a
-	%mono-directional channel is defined as the topolog- ical boundary B S
-	%,R (improper) subset of the beliefs of the receiver (i.e., the beliefs
-	%generated by the message exchange and by computations over the set of
-	%beliefs): PP(B S ,R , B R ) ∨ EQ(B S ,R , B R ).
+	Given two agents $s$ (sender) and $r$ (receiver) a
+	mono-directional channel $s \rightarrow r$ is defined as the union of all the assertions 
+	$\rassert{s}{r}$ sent by $s$ to $r$ ($\assertionRegion_s$) and received by $r$ from $s$ ($\assertionRegion_r$). 
 \end{definition}
+The \emph{quality of a mono-directional} channel is defined as the 
+$\rcc$ relation between the assertions made by the sender and the ones received by the receiver, i.e. $\Rcc{s}{r}$.
 
-\begin{definition}{\bf \fix{mr}{informal Input or Output Port} --}\label{def:port} 
-	Ports are agents with the following predefined behavior: an input-port
-	changes the type of information from assertion to belief while an
-	output-port from belief to assertion; forwarding information from the
-	outside of an agent's boundary to the inside (input-port) or vice versa
-	(output-port).
+\begin{definition}{\bf Input or Output Port --}\label{def:port} 
+	Ports are agents with the following predefined behavior: an input-port changes
+	the type of information from assertion from a sender $s$ to an agent $a$,
+	$\rassert{s}{a}$ to belief of the agent $a$ ($\beliefRegion_a$) while an
+	output-port from belief to assertion; forwarding information from the outside
+	of an agent's boundary to the inside (input-port) or vice versa (output-port). 
 \end{definition}
+The \emph{quality of a port} is determined by the $\rcc$ relation between the 
+assertions received or sent and the belief, i.e. $\Rcc{\assertionRegion_s}{\beliefRegion_a}$ for the input-port
+or $\Rcc{\assertionRegion_r}{\beliefRegion_a}$.
+
+\begin{definition}{\bf Secure Input Port --}\label{def:secport}
+	A secure input port allows information as incoming assertions to flow
+	from a sender $s$ to the behavior of the recipient $a$ (agent).  If $p_I$ is a secure input port,
+	$\world\models\varphi$, and $\varphi\in\assertionRegion_s$ where $s$
+	has the only input port $p_I$, $\world'\models\beliefRegion_a$ and $\varphi\in\beliefRegion_a$.
+\end{definition}
+From this definition, it follows that there exist only the following types of insecure port in RCC5:
+\begin{itemize}
+	\item \emph{drop port} ($s\circ\rightarrow a$), where assertions reaches the port but do not pass the boundary of the agent (i.e. do not become belief of the agent)
+	\item \emph{insertion port} ($s\rightarrow\bullet a$), where some new information is believed by $a$ as incoming from the port but $s$ didn't send it
+	\item \emph{injection port} ($s\circ\rightarrow\bullet a$), where information coming from $s$ is substituted with new information which becomes believed by $a$
+	\item \emph{selective port} ($s\hookrightarrow a$), where some
+		information passes the port while other is either dropped
+		($s\circ\hookrightarrow a$), inserted
+		($s\hookrightarrow\bullet a$), or injected
+		($s\circ\hookrightarrow\bullet a$)
+\end{itemize}
+
+\begin{proof}
+
+\end{proof}
 

### 2020-03-26 engineering.tex
add paragraph name "port weaknesses"
port proof (first draft)

@@
 
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.9\textwidth]{IOports.pdf}
+	\caption{Input and Output Port of an agent}
+	\label{fig:ioports}
+\end{figure}
+
 While \emph{agents} and \emph{communication} between agents are described and
-defined in Section~\ref{sec:theory}, both the concepts of \emph{port},
+defined in Section~\ref{sec:theory}, both the concepts of \emph{port} (depicted in Figure~\ref{fig:ioports},
 \emph{channel}, \emph{functional and physical architecture}, and
@@
 \end{definition}
+
+\Paragraph{Port weaknesses}
 From this definition, it follows that there exist only the following types of insecure port in RCC5:
 \begin{itemize}
-	\item \emph{drop port} ($s\circ\rightarrow a$), where assertions reaches the port but do not pass the boundary of the agent (i.e. do not become belief of the agent)
-	\item \emph{insertion port} ($s\rightarrow\bullet a$), where some new information is believed by $a$ as incoming from the port but $s$ didn't send it
-	\item \emph{injection port} ($s\circ\rightarrow\bullet a$), where information coming from $s$ is substituted with new information which becomes believed by $a$
-	\item \emph{selective port} ($s\hookrightarrow a$), where some
+	\item \emph{drop port} ($s\dropport a$), where assertions reaches the port but do not pass the boundary of the agent (i.e. do not become belief of the agent)
+	\item \emph{insertion port} ($s\insertport a$), where some new information is believed by $a$ as incoming from the port but $s$ didn't send it
+	\item \emph{injection port} ($s\injectionport a$), where information coming from $s$ is substituted with new information which becomes believed by $a$
+	\item \emph{selective port}, where some
 		information passes the port while other is either dropped
-		($s\circ\hookrightarrow a$), inserted
-		($s\hookrightarrow\bullet a$), or injected
-		($s\circ\hookrightarrow\bullet a$)
+		($s\selectivedropport a$), inserted
+		($s\selectiveinsertport a$), or injected
+		($s\selectiveinjectionport a$)
 \end{itemize}
@@
 \begin{proof}
-
+An input port is, in the $\abf$-theory, defined secure as long as the relation
+	between the two regions of input assertions $\assertionRegion$ and
+	output beliefs $\beliefRegion$ are equal, i.e.
+	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
+	should result in a weakness (related to an insecurity flaw) of that
+	input port.  Using RCC5, there exist exactly other $4$ different type
+	of relations, one of which is the discrete-from (DR) relation, i.e.
+	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
+	the DR relations, they have no element (subregion) in common. Lets
+	define a function weight $|X|$ such as, for any region $X$, it
+	represents the smallest possible cardinality of a topological base for
+	$X$; where a base is a collection of regions in a topology such that
+	every open region can be written as union of elements of that base.  If
+	$\dr{\assertionRegion}{\beliefRegion}$ there exist only two mutually
+	exclusive options: either $|\assertionRegion|=|\beliefRegion|$ or
+	$|\assertionRegion|\neq|\beliefRegion|$. If, at the end of a
+	communication through the input-port
+	$|\assertionRegion|\neq|\beliefRegion|$, the two regions have a
+	different number of subregions and then there exist only two mutually
+	exclusive options: 
+	\begin{itemize}
+		\item either $|\assertionRegion|>|\beliefRegion|$, where one
+			ore more assertions didn't become Beliefs, which we
+			call \emph{drop port} since its behavior drops and
+			prevents incoming communications,
+		\item or $|\assertionRegion|<|\beliefRegion|$, where there are
+			one ore more Beliefs which have not been sent as
+			incoming communication, which we call \emph{insertion
+			port} since its behavior generates new Beliefs
+			unrelated to the incoming communications. 
+	\end{itemize}
+	If $|\assertionRegion|=|\beliefRegion|$, given that information is monotone by nature (due to its entropy)
+there exist only two mutually
+	exclusive options: 
+	\begin{itemize}
+		\item either information has been generated and transferred, increasing the cardinality of both
+			$\assertionRegion$ and $\beliefRegion$, i.e.
+			$|\assertionRegion_{t^0}|>|\assertionRegion_{t^1}|$ and
+			$|\beliefRegion_{t^0}|>|\beliefRegion_{t^1}|$ (where
+			$t^0$ and $t^1$ represent time at the beginning of
+			communication and right after respectively), which
+			means that Assertions has reached the port and new
+			Belief have been generated by the port but no
+			correlation the elements of the two regions exist, and
+			we call it \emph{injection port} since it has modified
+			the information carried by the assertions into new
+			corresponding beliefs.
+		\item or $|\assertionRegion_{t^0}|=|\assertionRegion_{t^1}|$
+			and $|\beliefRegion_{t^0}|=|\beliefRegion_{t^1}|$ but
+			this implies that no communication occurred, while we
+			assumed that a communication was taking place (absurd).
+	\end{itemize}
+	Similarly, 
+	\begin{itemize}
+		\item given that $\dr{\assertionRegion}{\beliefRegion}$ is
+			$\dr{\assertionRegion}{\beliefRegion}$ except for the
+			intersection in which case is
+			$\eq{\assertionRegion}{\beliefRegion}$ (see
+			\autocite{Santaca2016abf}),
+			we call it a selective injection port ($s\selectivedropport a$),
+		\item if $\pp{\assertionRegion}{\beliefRegion}$, we have that
+			some Belief are not correlated to incoming assertion
+			and we call it ($s\selectiveinsertport a$)
+		\item if $\ppi{\assertionRegion}{\beliefRegion}$ we have that
+			some incoming Assertions are not correlated to Belief 
+			and we call it ($s\selectiveinjectionport a$)
+	\end{itemize}
 \end{proof}

### 2020-03-27 engineering.tex
review proof for ports

--- 2020-03-25 engineering.tex
@@
 \begin{proof}
-
+An input port is, in the $\abf$-theory, defined secure as long as the relation
+	between the two regions of input assertions $\assertionRegion$ and
+	output beliefs $\beliefRegion$ are equal, i.e.
+	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
+	should result in a weakness (related to an insecurity flaw) of that
+	input port.  Using RCC5, there exist exactly other $4$ different type
+	of relations, one of which is the discrete-from (DR) relation, i.e.
+	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
+	the DR relations, they have no subregion in common. Lets
+	define a function weight $|X|$ such as, for any region $X$, it
+	represents the smallest possible cardinality of a (mereo)topological base for
+	$X$; where a base is a collection of regions in a (mereo)topology such that
+	every open region can be written as union of elements of that base.  If
+	a communication occurred and results in
+	$\dr{\assertionRegion}{\beliefRegion}$, there exist only two mutually
+	exclusive options: either $|\assertionRegion|=|\beliefRegion|$ or
+	$|\assertionRegion|\neq|\beliefRegion|$. 
+	\begin{itemize}
+		\item If $|\assertionRegion|\neq|\beliefRegion|$ (at the end of
+			a communication through the input-port), the two
+			regions have a different number of subregions, and,
+			then, there exist only two mutually exclusive options: 
+			\begin{itemize}
+				\item either $|\assertionRegion|>|\beliefRegion|$, where one
+					ore more Assertions didn't become Beliefs, which we
+					call \emph{drop port} ($s\dropport a$) since its behavior drops and
+					prevents incoming communications,
+				\item or $|\assertionRegion|<|\beliefRegion|$, where there are
+					one ore more Beliefs which have not been sent as
+					input-Assertions, which we call \emph{insertion
+					port} ($s\insertport a$) since its behavior generates new Beliefs
+					unrelated to the incoming communication. 
+			\end{itemize}
+		\item If $|\assertionRegion|=|\beliefRegion|$, supposing an increasing monotonic entropy on the information (Assertions) exchanged, 
+				there exist only two mutually exclusive options: 
+			\begin{itemize}
+				\item either information has been generated as Belief and transferred as Assertions, increasing the cardinality of both
+					$\assertionRegion$ and $\beliefRegion$, i.e.
+					$|\assertionRegion_{t^0}|>|\assertionRegion_{t^1}|$ and
+					$|\beliefRegion_{t^0}|>|\beliefRegion_{t^1}|$ (where
+					$t^0$ and $t^1$ represent time at the beginning of
+					communication and right after, respectively); which
+					means that Assertions have reached the port and new
+					Beliefs have been generated by the port, but no
+					correlation between the elements of the two regions exist. 
+					We call this \emph{injection port} ($s\injectionport a$) since it has modified
+					the information carried by the Assertions into new
+					corresponding Beliefs.
+				\item or $|\assertionRegion_{t^0}|=|\assertionRegion_{t^1}|$
+					and $|\beliefRegion_{t^0}|=|\beliefRegion_{t^1}|$ but
+					this implies that no communication occurred, while we
+					assumed that a communication was taking place (absurd).
+			\end{itemize}
+	\end{itemize}
+	Similarly, 
+	\begin{itemize}
+		\item given that $\po{\assertionRegion}{\beliefRegion}$ is equivalent to
+			$\dr{\assertionRegion}{\beliefRegion}$ except for the
+			overlapping part (see Appendix~\ref{app:selectiveinjectionport}) in which case is
+			$\eq{\assertionRegion}{\beliefRegion}$ (see
+			\autocite{Santaca2016abf}),
+			we call it a \emph{selective injection port} ($s\selectiveinjectionport a$); otherwise
+		\item $\pp{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Leftarrow\beliefRegion$ and
+			$\ppi{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Rightarrow\beliefRegion$ are subcases of
+			$\eq{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Leftrightarrow\beliefRegion$, and then
+		\begin{itemize}
+			\item if $\pp{\assertionRegion}{\beliefRegion}$, we have that
+				some Beliefs are not correlated to incoming Assertions
+				and we call it \emph{selective insertion port} ($s\selectiveinsertport a$),
+			\item if $\ppi{\assertionRegion}{\beliefRegion}$ we have that
+				some incoming Assertions are not correlated to Belief 
+				and we call it \emph{selective drop port} ($s\selectivedropport a$)
+		\end{itemize}
+	\end{itemize}
 \end{proof}

### 2020-03-27 engineering.tex
definition of channels

@@
 
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.9\textwidth]{channel.pdf}
+	\caption{Communication over a Mono-directional Channel}
+	\label{fig:channel}
+\end{figure}
+
 While \emph{agents} and \emph{communication} between agents are described and
-defined in Section~\ref{sec:theory}, both the concepts of \emph{port} (depicted in Figure~\ref{fig:ioports},
-\emph{channel}, \emph{functional and physical architecture}, and
+defined in Section~\ref{sec:theory}, the concepts of \emph{port} (depicted in
+Figure~\ref{fig:ioports}), \emph{channel} (depicted in
+Figure~\ref{fig:channel}), \emph{functional and physical architecture}, and
 \emph{functional block} have not yet been defined with respect to the
-$\abf$-theory.  Informally, a channel identifies the place where the
-communication takes place.  In this work we only consider mono-directional
-channels and communication but the extension to bi-directional channel can be
-considered as the union of two unidirectional channels. A mono-directional
-channel is defined by the assertions sent or received (over the channel).
+$\abf$-theory.  Informally, a channel\footnote{for now, we rely on the
+intuitive notion of channel as ``something'' that transfers information as
+Assertions between output and input port, i.e. monodirectional} identifies the
+place where the communication takes place.  In this work, we only consider
+mono-directional channels and communication but the extension to bi-directional
+channel can be considered as the union of two unidirectional channels. A
+mono-directional channel is defined by the assertions sent or received (over
+the channel).
 
-\begin{definition}{\bf mono-directional channel --}\label{def:monochannel}
-	Given two agents $s$ (sender) and $r$ (receiver) a
-	mono-directional channel $s \rightarrow r$ is defined as the union of all the assertions 
-	$\rassert{s}{r}$ sent by $s$ to $r$ ($\assertionRegion_s$) and received by $r$ from $s$ ($\assertionRegion_r$). 
-\end{definition}
-The \emph{quality of a mono-directional} channel is defined as the 
-$\rcc$ relation between the assertions made by the sender and the ones received by the receiver, i.e. $\Rcc{s}{r}$.
+\Paragraph{Input and Output Ports}
+Since the $\abf$-theory is a theory of agents, we consider ports as agents that
+allows the exchange of information between a channel and another
+agent.  However, a port must be considered as a special type of agent to avoid
+an \emph{infinite regress} in which a port needs a port to transfer information
+between the outside of the port to the inside of itself.
 
@@
 
-\Paragraph{Port weaknesses}
-From this definition, it follows that there exist only the following types of insecure port in RCC5:
-\begin{itemize}
+\Paragraph{Port Weaknesses}
+From this definition, it follows that there exist only the following \emph{six} types of weaknesses, generating six types of insecure port in RCC5:
+\begin{enumerate}[start=1, label={W\arabic*)}]
 	\item \emph{drop port} ($s\dropport a$), where assertions reaches the port but do not pass the boundary of the agent (i.e. do not become belief of the agent)
@@
 	\item \emph{injection port} ($s\injectionport a$), where information coming from $s$ is substituted with new information which becomes believed by $a$
-	\item \emph{selective port}, where some
-		information passes the port while other is either dropped
-		($s\selectivedropport a$), inserted
-		($s\selectiveinsertport a$), or injected
-		($s\selectiveinjectionport a$)
-\end{itemize}
+	\item \emph{selective port}, where some information passes the port and part is either:
+	\begin{enumerate}[start=1, label={W4.\arabic*)}]
+		\item dropped ($s\selectivedropport a$), 
+		\item inserted ($s\selectiveinsertport a$), or
+		\item injected ($s\selectiveinjectionport a$)
+\end{enumerate}
+\end{enumerate}
 
@@
 	\end{itemize}
+	In other words, $\dr{\assertionRegion}{\beliefRegion}\equiv\neg\overlaps{\assertionRegion}{\beliefRegion}\equiv\neg(\varphi\subseteq\assertionRegion\wedge\varphi\subseteq\beliefRegion)\equiv\varphi\not\subseteq\assertionRegion\vee\varphi\not\subseteq\beliefRegion$
 	Similarly, 
@@
 \end{proof}
+
+\Paragraph{Communication Channels}
+We start by considering the difference between a (communication)
+mono-directional channel (channel from now on) and an agent, as we did for the
+ports, since the $\abf$-theory is a theory of agents.  In fact, if a channel
+were considered an agent (channel-agent) then the question would be how an
+agent would transfer its Assertions to the channel-agent. If the channel
+between the agent and the channel-agent is again an agent, we would generate an
+\emph{infinite regress}. Therefore, we do allow channel-agents but we assume
+a finite depth (of detail) for a channel, where there exists a bottom-channel 
+which is not an agent. For now, we do not constrain a channel-agent in any way
+so there is no difference between a channel agent and agent. Therefore, for
+the sake of simplicity, we consider, in this text, channels to be bottom-channels,
+defined as agents with the pre-defined behavior of storing Assertions that are transferred
+over them (i.e. over the channel we are defining).
+
+\begin{definition}{\bf Mono-directional Channel --}\label{def:monochannel}
+	Given two agents $s$ (sender) and $r$ (receiver) a
+	mono-directional channel $s \rightarrow r$ is defined as the union of all the assertions 
+	$\rassert{s}{r}$ sent by $s$ to $r$ ($\assertionRegion_s$) and received by $r$ from $s$ ($\assertionRegion_r$) (for any given instance of time). 
+\end{definition}
+The \emph{quality of a mono-directional} channel is defined as the 
+$\rcc$ relation between the Assertions made by the sender and the ones received by the receiver, i.e. $\Rcc{\assertionRegion_s}{\beliefRegion_r}$.
+
+\Paragraph{Channel Weaknesses}
 
@@
 Integrity doesn't imply non-repudiation (replay attacks?). 
-
-\begin{enumerate}
-\item specification -- definition of the desired design of a system. The
-specification is verified w.r.t. the  $\abf$ theory. Checking the
-controls of the whole cybersecurity life-cycle and the relation with
-CWE.
-\item design -- the mitigations identified in the specification stage
-are implemented into the design. The security of the assets w.r.t. the design are
-verified with a, so called, cybersecurity risk assessment.
-\end{enumerate}
 }
+\subsection{Abstraction Level of a Specification in $\abf$-theory}
+As highlighted in \autocite{}, ``The level of abstraction is always a concern
+when it come sto the definition of a system model.'' (as we already mentioned in Section~\ref{sec:problem}).

### 2020-03-30 engineering.tex
first complete draft communication insecurity

@@
 intuitive notion of channel as ``something'' that transfers information as
-Assertions between output and input port, i.e. monodirectional} identifies the
+Assertions between output and input port, i.e. mono-directional} identifies the
 place where the communication takes place.  In this work, we only consider
@@
 \Paragraph{Port Weaknesses}
-From this definition, it follows that there exist only the following \emph{six} types of weaknesses, generating six types of insecure port in RCC5:
+From this definition, it follows that there exist only the following \emph{six} types of weaknesses, generating six types of insecure port in RCC5 (the notation is reported for input-ports, i.e.
+on the lef-hand side of the arrow):
 \begin{enumerate}[start=1, label={W\arabic*)}]
@@
 between the agent and the channel-agent is again an agent, we would generate an
-\emph{infinite regress}. Therefore, we do allow channel-agents but we assume
-a finite depth (of detail) for a channel, where there exists a bottom-channel 
+\emph{infinite regress}. Therefore, we do allow channel-agents but we assume a
+finite depth (of detail) for a channel, where there exists a bottom-channel
 which is not an agent. For now, we do not constrain a channel-agent in any way
-so there is no difference between a channel agent and agent. Therefore, for
-the sake of simplicity, we consider, in this text, channels to be bottom-channels,
-defined as agents with the pre-defined behavior of storing Assertions that are transferred
-over them (i.e. over the channel we are defining).
+so there is no difference between a channel agent and agent. Therefore, for the
+sake of simplicity, we consider, in this text, channels to be bottom-channels,
+defined as agents with the pre-defined behavior (i.e. defined in a dogmatic
+way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
+prevents us from introducing additional constraints to the channel as storing
+Assertions that are transferred over the channel, or filter out some
+input-assertions.}.
 
-\begin{definition}{\bf Mono-directional Channel --}\label{def:monochannel}
-	Given two agents $s$ (sender) and $r$ (receiver) a
-	mono-directional channel $s \rightarrow r$ is defined as the union of all the assertions 
-	$\rassert{s}{r}$ sent by $s$ to $r$ ($\assertionRegion_s$) and received by $r$ from $s$ ($\assertionRegion_r$) (for any given instance of time). 
+\begin{definition}{\bf Secure Mono-directional Channel (bottom-channel) --}\label{def:monochannel}
+	A mono-directional channel between the two agents ($s \rightarrow r$) is defined as 
+	an agent which behavior is (dogmatically) defined as: to forward any Assertion received from $s$ over an input-port, to  
+	the output-port where $r$ is listening to.
 \end{definition}
 The \emph{quality of a mono-directional} channel is defined as the 
-$\rcc$ relation between the Assertions made by the sender and the ones received by the receiver, i.e. $\Rcc{\assertionRegion_s}{\beliefRegion_r}$.
+$\rcc$ relation between the Assertions made by the sender and the ones received by the receiver, i.e. $\Rcc{\assertionRegion_s}{\assertionRegion_r}$.
 
 \Paragraph{Channel Weaknesses}
+Given that a mono-directional bottom-channel is assumed to be perfectly
+forwarding any Assertion from its input-port to its output-port, there is no
+insecure behavior but only the combination of the weaknesses of the input and
+output port; therefore there exists $(7^2)-2=47$ theoretical configurations
+($7^2$ because there are $6$ insecure types of port, plus $1$ secure type, on both input and oupput
+side; and we exclude the configuration with $2$ secure types as input
+and ouput, $-2$); where only 43 are possible.
+
+\begin{enumerate}[start=5, label={W\arabic*)}]
+	\item \emph{secure input port} and \emph{output drop port} ($s\sdcom r$), 
+	\item \emph{secure input port} and \emph{output insertion port} ($s\sicom r$), 
+	\item \emph{secure input port} and \emph{output injection port} ($s\sjcom r$), 
+	\item \emph{secure input port} and \emph{output selective drop port} ($s\ssdcom r$), 
+	\item \emph{secure input port} and \emph{output selective insertion port} ($s\ssicom r$), 
+	\item \emph{secure input port} and \emph{output selective injection port} ($s\ssjcom r$),
+
+	\item \emph{input drop port} and \emph{output drop port} ($s\ddcom r$)
+	\item \emph{input drop port} and \emph{output insertion port} ($s\dicom r$)
+	\item \emph{input drop port} and \emph{output secure port} ($s\dscom r$)
+
+	\item \emph{input insert port} and \emph{output drop port} ($s\idcom r$)
+	\item \emph{input insert port} and \emph{output insert port} ($s\iicom r$)
+	\item \emph{input insert port} and \emph{output injection port} ($s\ijcom r$)
+	\item \emph{input insert port} and \emph{output selective drop port} ($s\isdcom r$)
+	\item \emph{input insert port} and \emph{output selective insertion port} ($s\isicom r$)
+	\item \emph{input insert port} and \emph{output selective injection port} ($s\isjcom r$)
+	\item \emph{input insert port} and \emph{output secure port} ($s\iscom r$)
+
+	\item \emph{input injection port} and \emph{output drop port} ($s\jdcom r$)
+	\item \emph{input injection port} and \emph{output insertion port} ($s\jicom r$)
+	\item \emph{input injection port} and \emph{output injection port} ($s\jjcom r$)
+	\item \emph{input injection port} and \emph{output selective drop port} ($s\jsdcom r$)
+	\item \emph{input injection port} and \emph{output selective insertion port} ($s\jsicom r$)
+	\item \emph{input injection port} and \emph{output selective injection port} ($s\jsjcom r$)
+	\item \emph{input injection port} and \emph{output secure port} ($s\jscom r$)
+
+	\item \emph{input selective drop port} and \emph{output drop port} ($s\sddcom r$)
+	\item \emph{input selective drop port} and \emph{output insertion port} ($s\sdicom r$)
+	\item \emph{input selective drop port} and \emph{output injection port} ($s\sdjcom r$)
+	\item \emph{input selective drop port} and \emph{output selective drop port} ($s\sdsdcom r$)
+	\item \emph{input selective drop port} and \emph{output selective insertion port} ($s\sdsicom r$)
+	\item \emph{input selective drop port} and \emph{output selective injection port} ($s\sdsjcom r$)
+	\item \emph{input selective drop port} and \emph{output secure port} ($s\sdscom r$)
+
+	\item \emph{input selective insertion port} and \emph{output drop port} ($s\sidcom r$)
+	\item \emph{input selective insertion port} and \emph{output insertion port} ($s\siicom r$)
+	\item \emph{input selective insertion port} and \emph{output injection port} ($s\sijcom r$)
+	\item \emph{input selective insertion port} and \emph{output selective drop port} ($s\sisdcom r$)
+	\item \emph{input selective insertion port} and \emph{output selective insertion port} ($s\sisicom r$)
+	\item \emph{input selective insertion port} and \emph{output selective injection port} ($s\sisjcom r$)
+	\item \emph{input selective insertion port} and \emph{output secure port} ($s\siscom r$)
+
+	\item \emph{input selective injection port} and \emph{output drop port} ($s\sjdcom r$)
+	\item \emph{input selective injection port} and \emph{output insertion port} ($s\sjicom r$)
+	\item \emph{input selective injection port} and \emph{output injection port} ($s\sjjcom r$)
+	\item \emph{input selective injection port} and \emph{output selective drop port} ($s\sjsdcom r$)
+	\item \emph{input selective injection port} and \emph{output selective insertion port} ($s\sjsicom r$)
+	\item \emph{input selective injection port} and \emph{output selective injection port} ($s\sjsjcom r$)
+	\item \emph{input selective injection port} and \emph{output secure port}  ($s\sjcom r$)
+\end{enumerate}
+In Appendix~\ref{app:insecurecom}, we report the proof which is exhaustive on all
+possible combinations. The only
 
@@
 and integrity) to be effectively applied to the engineering of systems. Due to
-this, many researchers an organization tried to improve the CIA triad.  The
+this, many researchers and organization tried to improve the CIA triad.  The
 evolutions/extensions of the CIA triad has been documented, e.g., in
@@
 we now define (in Section~\ref{sec:properties}) the CIA triad for an agent
-defined in the $\abftheory$ theory and we test if (i) other properties are
-allowed in the $\abftheory$ theory, and (ii) if and how the CIA triad can be
-detailed in the $\abftheory$ theory.
+defined in the $\abftheory$-theory and we test if (i) other properties are
+allowed in the $\abftheory$-theory, and (ii) if and how the CIA triad can be
+detailed in the $\abftheory$-theory.
 
@@
 We now want to generalize the CIA triad to agents and not just 
-information and map it to the $\abftheory$ theory.
+information and map it to the $\abftheory$-theory. For the sake of simplicity,
+we revert the order and start from availability.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\textwidth]{client-server.pdf}
+	\caption{Client-Server paradigm in $\abf$-theory}
+	\label{fig:client-server}
+\end{figure}
+Availability, again, is the denial of use of a resource. In the $\abf$-theory, 
+in order to prevent a resource to be available, there
+exists only (i.e. inclusively) the following mutually exclusive cases
+that we illustrate using a client-server paradigm (depicted in Figure~\ref{fig:client-server}).
+\begin{enumerate}
+	\item\label{case:srv1} The resource on the server cannot be found, so that it is
+		possible for the client to contact the server and the server
+		sends a response back, but the output-port does not receive the
+		necessary information. This may be due to a software error such
+		that the functional architecture doesn't output a response, or
+		a problem in the flow of information (i.e. the functional
+		architecture doesn't receive the necessary information from
+		other resources), or an hardware failure \&c. This is formalized
+		as the relation between the beliefs produced by the functional
+		architecture and the ones received by the output port, i.e.
+		$\rcc(\behaviorRegion_s,\{\text{res}\})$ or 
+		$\rcc(\beliefRegion_s,\{\text{res}\})$ 
+	\item\label{case:srv2} The server hosting the resource is down, and then the server
+		cannot output responses (i.e. insecure output port due to drop-related issue) or receive
+		requests (insecure input port due to drop-related issue) or both.
+	\item The client cannot properly compute request (simmetrically to Case~\ref{case:srv1})
+	\item The client is down, symmetric to Case~\ref{case:srv2}
+	\item The channel is down due to a drop related issue in one of the two channels 
+\end{enumerate}
+
+\begin{definition}{\bf Infosec-Availability --}\label{def:infosec-c}
+\end{definition}
+
+\begin{definition}{\bf Agent-Availability --}\label{def:agent-a}
+\end{definition}
+
+\begin{definition}{\bf System-Availability --}\label{def:system-a}
+\end{definition}
+
 \begin{definition}{\bf Infosec-Confidentiality --}\label{def:infosec-c}

### 2020-03-30 insecureport.tex
first complete draft communication insecurity

\fix{mr}{this proof needs to be done}
\begin{proof}
In Appendix~\ref{app:commproof}, we report the proof which is exhaustive on all
possible combinations.  In the following we report only the excluded cases and
assume the other to be strightforwad (as a proof sketch).
\begin{enumerate}
	\item \emph{input secure port} and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\sdcom r$), 
			\item \emph{output insertion port} ($s\sicom r$), 
			\item \emph{output injection port} ($s\sjcom r$), 
			\item \emph{output selective drop port} ($s\ssdcom r$), 
			\item \emph{output selective insertion port} ($s\ssicom r$), 
			\item \emph{output selective injection port} ($s\ssjcom r$), 
		\end{enumerate}
	\item \emph{input drop port} (all the input-assertions are dropped) and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\ddcom r$), this
				configuration is possible but the input port
				doesn't forward any message therefore the
				output port will never have a message to drop
			\item \emph{output insertion port} ($s\dicom r$), where
				all the ouput-assertions have been inserted by
				the output port
			\item\label{drop-injection} \emph{output injection
				port}, this configuration is impossible since
				there is no Assertions from the input-port to
				inject (or it behaves as $s\dicom r$)
			\item\label{drop-seldrop} \emph{output selective drop
				port}, as in \ref{drop-injection} there is no
				Assertion from the input port therfore no
				selective port is possible in general. In this
				case this configuration is impossible or
				behaves as $s\ddcom r$
			\item \emph{output selective insertion port}, as in
				\ref{drop-seldrop} impossible, or behaves as
				$s\dicom r$
			\item \emph{output selective injection port}, as in
				\ref{drop-seldrop} impossible, or behaves as
				$s\dicom r$
			\item \emph{ouput secure port} ($s\dscom r$)
		\end{enumerate}

	\item \emph{input insertion port} (where all the assertion reaching the
		output-port have been artificially generated by the input-port) and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\idcom r$), this
				configuration is possible but the drop port
				drops every inserted message.
			\item \emph{output insertion port} ($s\iicom r$), new
				Assertions have all been inserted both at the left-hand
				side and righ-hand side of the communication.
			\item \emph{output injection port} ($s\ijcom r$), the
				new Assertions, inserted in the left-hand side,
				are then modified/injected on the right-hand side of the
				communication
			\item \emph{output selective drop port} ($s\isdcom r$),
				where some insertions of the input port are dropped 
			\item \emph{output selective insertion port} ($s\isicom
				r$), where some insertions of the input port
				are substituted with new insertion on the
				right-hand side of the communication
			\item \emph{output selective injection port} ($s\isjcom
				r$) where some insertions of the input port are
				modified/injected on the right-hand side of the
				communication 
			\item \emph{output secure port} ($s\iscom r$)
		\end{enumerate}

	\item \emph{input injection port} and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\jdcom r$)
			\item \emph{output insertion port} ($s\jicom r$)
			\item \emph{output injection port} ($s\jjcom r$)
			\item \emph{output selective drop port} ($s\jsdcom r$)
			\item \emph{output selective insertion port} ($s\jsicom r$)
			\item \emph{output selective injection port} ($s\jsjcom r$)
			\item \emph{output secure port} ($s\jscom r$)
		\end{enumerate}

	\item \emph{input selective drop port} and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\sddcom r$)
			\item \emph{output insertion port} ($s\sdicom r$)
			\item \emph{output injection port} ($s\sdjcom r$)
			\item \emph{output selective drop port} ($s\sdsdcom r$)
			\item \emph{output selective insertion port} ($s\sdsicom r$)
			\item \emph{output selective injection port} ($s\sdsjcom r$)
			\item \emph{output secure port} ($s\sdscom r$)
		\end{enumerate}

	\item \emph{input selective insertion port} and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\sidcom r$)
			\item \emph{output insertion port} ($s\siicom r$)
			\item \emph{output injection port} ($s\sijcom r$)
			\item \emph{output selective drop port} ($s\sisdcom r$)
			\item \emph{output selective insertion port} ($s\sisicom r$)
			\item \emph{output selective injection port} ($s\sisjcom r$)
			\item \emph{output secure port} ($s\siscom r$)
		\end{enumerate}

	\item \emph{input selective injection port} and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\sjdcom r$)
			\item \emph{output insertion port} ($s\sjicom r$)
			\item \emph{output injection port} ($s\sjjcom r$)
			\item \emph{output selective drop port} ($s\sjsdcom r$)
			\item \emph{output selective insertion port} ($s\sjsicom r$)
			\item \emph{output selective injection port} ($s\sjsjcom r$)
			\item \emph{output secure port}  ($s\sjcom r$)
		\end{enumerate}
\end{enumerate}
\end{proof}

### 2020-03-31 conclusion.tex
first full draft of the paper

In this paper, we proposed a foundational theory on security, arguing that
security-related issues are not related to the maliciousness of an agent but to
the vagueness of the security controls on the engineering processes.
The current state-of-the-art processes allows, given a specification,
an engineer to design the system in such a way that security issues arise due to the lack
of proper security risk assessment processes.
Those design, again, lack of security verification processes based on a solid 
security foundational theory and then permit the generation of insecure implementation.
The security verification and test-case generation will be the focus of our next steps.

We may conclude that the problem of security is a problem related to the 
many possible design and, in turn, implementation given a specification.
Philosophically, the problem is similar to the epistemological search for truth,
where the challenge is to relate Information and Belief to Human Knowledge.
Scientifically, generalizing the Human to an Agent, the problem is
to relate Assertions and Behaviors, to Facts. From an engineering standpoint,
by generalizing the concept of Agents as architectural subsystems, the
problem is to link Channels (and Ports) and Functional Architectures to 
Requirements.

### 2020-03-31 engineering.tex
first full draft of the paper

@@
 Based on the Assets, architectures, and the Weaknesses the cybersecurity risk is
-calculated based on the $\abf$-theory. An iterative process between 
+calculated based on the $\abftheory$-theory. An iterative process between 
 the user (e.g. the engineer) and the \secramod module of the \abftool allows the user to move to the design phase
@@
 \emph{functional block} have not yet been defined with respect to the
-$\abf$-theory.  Informally, a channel\footnote{for now, we rely on the
+$\abftheory$-theory.  Informally, a channel\footnote{for now, we rely on the
 intuitive notion of channel as ``something'' that transfers information as
@@
 \Paragraph{Input and Output Ports}
-Since the $\abf$-theory is a theory of agents, we consider ports as agents that
+Since the $\abftheory$-theory is a theory of agents, we consider ports as agents that
 allows the exchange of information between a channel and another
@@
 \begin{proof}
-An input port is, in the $\abf$-theory, defined secure as long as the relation
+An input port is, in the $\abftheory$-theory, defined secure as long as the relation
 	between the two regions of input assertions $\assertionRegion$ and
@@
 mono-directional channel (channel from now on) and an agent, as we did for the
-ports, since the $\abf$-theory is a theory of agents.  In fact, if a channel
+ports, since the $\abftheory$-theory is a theory of agents.  In fact, if a channel
 were considered an agent (channel-agent) then the question would be how an
@@
 In Appendix~\ref{app:insecurecom}, we report the proof which is exhaustive on all
-possible combinations. The only
+possible combinations.
+
+\Paragraph{Functional Block}
+A Functional Architecture, as already described beforehand, takes Information as input-Beliefs
+and transforms the Information into output-Beliefs. Those transformations 
+occurs within the Functional Architecture, where Functional Blocks transforms Beliefs into 
+other Beliefs. Similarly to channels, we could consider a Functional Block as a Functional Architecture
+occurring in an infinite regress. Therefore, we consider Functional Blocks as 
+correctly executing their abstract behavior, correctly transforming inputs into outputs (Beliefs).
+
+\begin{definition}{\bf Functional Block and Architecture --}\label{def:funblock}
+	A functional block of an agent $a$ 
+	takes a region of input-Beliefs and outputs a region of output-Beliefs. 
+	A functional architecture is an interconnected system of functional blocks.
+\end{definition}
+The quality of a functional block or architecture cannot be expressed as a relation 
+between input and output belief since there is a transformation process in the functional
+block which will be defined during the design and implementation phase, but not
+necessarily known in details in the specification phase. In order to
+define the quality of functional block we first need to introduce the engineering concept of
+Facts.
+
+\Paragraph{Facts}
+During the specification phase, for any agent, channel, port, functional block
+and architecture, there exist a Fact predicating over them.
+In other words, any requirement is defined as a Fact since they must be true in
+any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
+Figure~\ref{fig:soundness}, Facts are strategic rules that defines how
+the system shall behave (by specification), while reality may be shown to
+be insecure.
+
+The quality of a Functional Block is determined by the relation between the
+input-Beliefs formulas and the symmetric formulas in the Fact region
+(equivalently for output-Beliefs). Therefore, the quality of the functional
+architecture is not a purely architectural property but relies on the logic
+implemented in the functional blocks (which we abstract away at this stage of
+the engineering model).
+\begin{definition}{\bf Secure Functional Block and Architecture --}\label{def:secfun}
+	A functional block is secure if the input beliefs and output beliefs
+	are in $EQ$ relation with the corresponding Facts mandated by requirements.
+	A functional architecture composed exclusively by secure functional blocks
+	is considered secure.
+\end{definition}
+
+Therefore, there exists the following weaknesses in the functional
+architecture, where $\beliefRegion_I$ and $\beliefRegion_O$ are input and
+output Beliefs respectively, and $\factRegion_I$ and $\factRegion_O$ are
+the corresponding Facts (i.e. what really happens).
+\begin{enumerate}
+	\item $\dr{\beliefRegion_I}{\factRegion_I}$ and
+		$\dr{\beliefRegion_O}{\factRegion_O}$, the input beliefs are
+		misinterpreted and this result in the complete alteration of the
+		output beliefs
+	\item \fix{mr}{TBD}
+\end{enumerate}
 
@@
 Infosec, or Information Security, defines the security risks related to
-information with the de-facto standard CIA triad , which is often criticized
+information with the de-facto standard CIA triad, which is often criticized
 \autocite{CIAcriticismCPS} for being too general or non-adequate (e.g. by
@@
 	\includegraphics[width=\textwidth]{client-server.pdf}
-	\caption{Client-Server paradigm in $\abf$-theory}
+	\caption{Client-Server paradigm in $\abftheory$-theory}
 	\label{fig:client-server}
 \end{figure}
-Availability, again, is the denial of use of a resource. In the $\abf$-theory, 
+Availability, again, is the denial of use of a resource. In the $\abftheory$-theory, 
 in order to prevent a resource to be available, there
@@
 		requests (insecure input port due to drop-related issue) or both.
-	\item The client cannot properly compute request (simmetrically to Case~\ref{case:srv1})
+	\item\label{case:srv3} The client cannot properly compute request (simmetrically to Case~\ref{case:srv1})
 	\item The client is down, symmetric to Case~\ref{case:srv2}
@@
 \end{enumerate}
-
+%All the cases but Case~\ref{case:srv1} and Case~\ref{case:srv3}, tell us how to map the
+%property of availability over the weaknesses of ports and communications.  The
+%Cases~\ref{case:srv1} and~\ref{case:srv3}, are exceptions because the formulas $res$ and $req$ 
+%detail the specification with a singleton instead of a region of formulas (or, better, $|res|$ is equal to the one of the
+%base).
+%In other words, a channel, a port, an entity, or a functional block are
+%Fact (stated in a factorial way) and not Beliefs; what they produce or 
+%their non-specified internals are Beliefs.  In fact, we never assume an
+%entity not to be in the design implementing the specification.  Similarly, when a
+%functional block produces some information it produces a Belief, unless this Belief if detailed in the specification.
+%For example, the server produces the Belief $res$ but the Belief given that can only be in $EQ$
+%relation with the Fact $res$ (in contrast other Belief may not); therefore, $res$ is not really a belief but a Fact. 
+%$res$ as a Fact is something that
+%exists in the specification, while the Belief of $res$ in one of the possible
+%designs may not exist. 
+We shall, then, distinguish between two cases:
+\begin{definition}{\bf Agent-Availability --}\label{def:agent-a}
+when the property of availability is expressed over
+	an agent or a system, the list of possible designs violating the
+	availability property can be shown as the designs that allow drop ports
+	or drop channels. 
+\end{definition}
 \begin{definition}{\bf Infosec-Availability --}\label{def:infosec-c}
+when the property of availability is expressed
+	over a formula, the list of possible designs that violates the
+	availability property is reduced to the part of the system that
+	influences that specific variable.
 \end{definition}
 
-\begin{definition}{\bf Agent-Availability --}\label{def:agent-a}
+In a similar way we can define integrity since this requirement may 
+be expressed over an agent/system or a specific formula (i.e. element of a Region).
+
+\begin{definition}{\bf Agent-Integrity --}\label{def:agent-a}
+when the property of integrity is expressed over
+	an agent or a system, the list of possible designs violating the
+	integrity property can be shown as the designs that allow insert or inject ports,
+	or inert or inject channels. 
+\end{definition}
+\begin{definition}{\bf Infosec-Integrity --}\label{def:infosec-c}
+when the property of integrity is expressed
+	over a formula, the list of possible designs that violates the
+	integrity property is reduced to the part of the system that
+	influences that specific variable.
 \end{definition}
 
-\begin{definition}{\bf System-Availability --}\label{def:system-a}
-\end{definition}
+While availability and integrity are architectural property, meaning that they
+can be expressed over the behavior of a port or a channel (as constituent of
+the physical or functional architecture), confidentiality is not.
+Confidentiality states that an information shall not be understood and not that
+the information shall not be retrieved. So, if a port leaks an entire message,
+there is no confidentiality issue in general. On the other hand, if the
+information is not provided or altered there is an availability or integrity
+issue. A similar argument can be proposed for integrity sub-properties such as
+authenticity or non-repudiation. No wonder why it is often stated that
+availability and integrity are the first properties to be considered in CPS
+while confidentiality takes the first place when dealing with security
+protocols. In fact, integrity is often assumed by pattern-matching (e.g. in
+AVANTSSAR \autocite{avantssar}) and availability is hardly ever considered in
+protocol verification). The question is, can we express properties such as
+confidentiality, authentication, or non-repudiation as architectural properties
+in $\abftheory$-theory?
 
+\newpage
 \begin{definition}{\bf Infosec-Confidentiality --}\label{def:infosec-c}
@@
 }
-\subsection{Abstraction Level of a Specification in $\abf$-theory}
+\subsection{Abstraction Level of a Specification in $\abftheory$-theory}
 As highlighted in \autocite{}, ``The level of abstraction is always a concern

### 2020-03-31 theory.tex
spdl description, use case and requirements draft
first full draft of the paper

@@
 		and
-	\item We consider a set of axiomatic \emph{Facts} (definitory rules) instead
+	\item We consider a set of axiomatic \emph{Facts} (definitory rules\fixnote{mr}{facts are definitory rule, as they don't define how a real system is finally implemented (curently) but how it should, trugh a series of requirements which may not be respected, through insecurity, in the implemented system}) instead
 		of considering the more general epistemic definition of
@@
 determines the degree of quality between the information circulating in a
-system (or within an agent) and reality.  Since the transfer of information through Assertions
+system (or within an agent) and reality\fixnote{mr}{assertions are reality
+while facts are dogmatic assertions at specification level, i.e. requirements}.
+Since the transfer of information through Assertions
 generates Beliefs, a dishonest agent may circulate false information,

### 2020-04-01 engineering.tex
add confidentiality

@@
 
+\subsubsection{Availability and Integrity (Architecture Security)}
 \begin{figure}[t]
@@
 	or drop channels. 
-\end{definition}
+\end{definition}\fixnote{mr}{all security definitions defines when the property is
+not preserved instead of when it's preserved}
 \begin{definition}{\bf Infosec-Availability --}\label{def:infosec-c}
@@
 the information shall not be retrieved. So, if a port leaks an entire message,
-there is no confidentiality issue in general. On the other hand, if the
+there is no confidentiality issue in general, unless that message was not
+encrypted at all, in which case the confidentiality leakage is straightforward. 
+On the other hand, if the
 information is not provided or altered there is an availability or integrity
@@
 AVANTSSAR \autocite{avantssar}) and availability is hardly ever considered in
-protocol verification). The question is, can we express properties such as
+protocol verification. The question is, can we express properties such as
 confidentiality, authentication, or non-repudiation as architectural properties
-in $\abftheory$-theory?
+in $\abftheory$-theory? Confidentiality (and Availability) is based on the concept 
+of secrecy so we shall first start from describing what secrecy is and how it
+can be used to guarantee confidentiality in a system.
 
-\newpage
-\begin{definition}{\bf Infosec-Confidentiality --}\label{def:infosec-c}
-	An Information $\varphi$ is considered to be confidential to a set of 
-	agents $Ag_c$ if, per each agent $a\in Ag_c$, the content of the
-	information (i.e. the data it carries) can only be understood (i.e. the relation
-	between the information and its data can be known) by the
-	agents in $Ag_c$.
+\subsubsection{Secrecy, Authenticity, and Key Agreement (Cryptographic Security)}\label{sec:encryption}
+As Maurer puts it (\autocite{Maurer1999information}), there exist two types of
+cryptographic security (i.e. scientific theories aiming at guaranteeing
+security through the use of cryptographic functions): `` The  security  of a
+cryptographic system can rely either on the computational infeasibility of
+breaking it (computational security), or on the theoretical impossibility of
+breaking it, even using in infinite computing power (information-theoretic or
+unconditional security)''. We are not currently interested in detailing the
+difference between the two and we simplify it as follows. A
+computationally-secure cryptographic function (usually called encryption
+function) $E(\varphi)=e$ (where $\varphi$ is a plain-text and $e$ a
+cipher-text) can be considered perfectly secure if a malicious agent that
+obtains the result of the cryptographic function ($e$) has bounded
+computational power, while a unconditional secure cryptographic functions are
+considered perfectly secure. In this context, security means that is is
+impossible to calculate $\varphi$ from $e$, therefore impossible to compute
+$D(e)=\varphi$.  As one can evince from \autocite{Shannon1948information}, the
+commonality between the two types of cryptographic functions is that they both
+rely on the assumption about the probabilistic behavior of the universe. They
+do so, for the random (i.e. with high entropy
+\autocite{Shannon1948information}) generation of the so called \emph{key}, a
+piece of information, parametric to the cryptographic function, used to
+guarantee the security. Therefore, we can conclude that it is feasible for an
+agent who knows $k$ to decrypt the cipher-text obtaining the clear-text (i.e. $D(k, E(k,\varphi))=\varphi$)
+while it is impossible for any agent who has not the proper key. In this
+case, the information is considered to be \emph{secret or confidential}.
 
-	\begin{itemize}
-		\item[$(\interpretation21)$] for all agents $a\in Ag$,
-			$\world\models\information{a}{\varphi}$ and
-			$\world\models\rcc(\varphi,Conf) \wedge
-			\neg\dr{Conf}{\varphi}$ iff for any world $\world'$
-			such that $\world \modalrelation\world'$,
-			$\world'\models\believe{a}{\varphi}$, and
-			$\world'\models\knows{a}{\exists f.~
-			f(\information{a}{\varphi})=\varphi}$, where 
-			$Conf$ is a Region and $f$ is an uninterpreted function.
-	\end{itemize}
+\begin{definition}{Infosec-Secrecy or Confidentiality --}\label{def:confidentiality}
+	When an Assertion has been generated as an output of a functional or
+	physical agent, port, or functional block such that it is said (in a
+	factorial way) to encrypt the input, that Assertion is considered
+	confidential.
 \end{definition}
-
-\begin{example}{Dummy security protocol --}
-	The last part (part III) of the protocol in our first example
-	(Figure~\ref{fig:protocol-example}) represents the actual key exchange
-	which, in general, needs to be confidential; otherwise the key would be
-	known by any other agent who can, e.g., read on the communication
-	channel.  Therefore, the last message, in Alice-and-Bob notation can be
-	rephrased as: \begin{displaymath} Alice\Rightarrow Bob: enc(pk(Bob),
-	symmetricKey) \end{displaymath} where $enc$ is the standard asymmetric
-	encryption operator that encrypts the message $symmetricKey$ with the
-	public key of Bob $pk(Bob)$ (see, e.g.,
-	\autocite{Rocchetto2017interpolation} for a detailed formalization).
-	The function $f$ in Definition~\ref{def:infosec-c}, represents any
-	function used to enforce confidentiality on information.  In this
-	example, $f$ is identified with $enc$ (i.e.
-	$enc(pvt(Bob),symmetricKey)=symmetricKey$), the information is
-	$\information{Alice}{[enc(pk(Bob),symmetricKey)]}$ and the data, i.e.
-	$\varphi$ in Definition~\ref{def:infosec-c}, is $\varphi=symmetricKey$.
-	In the perfect cryptography assumption, and assuming for simplicity
-	that the logic of the whole protocol preserves confidentiality, the
-	message is confidential. In fact, if $Ag={Alice, Bob}$ and $\eq{symmetricKey}{Conf}$:
-	\begin{align*}
-		\world^0\models\belief{Alice}{symmetricKey},\\
-		\world^1\models\information{Alice}{symmetricKey}, \text{and}\\
-		\world^2\models\belief{Bob}{symmetricKey} ~\text{if} \\
-		enc(pvt(Bob), \information{Alice}{symmetricKey}=symmetricKey)=symmetricKey\\
-		(i.e. f=enc ~\text{and}~ \information{Alice}{symmetricKey}=enc(pub(Bob),symmetricKey))\\
-	\end{align*}
-\end{example}
-
-\begin{definition}{\bf Agent-Confidentiality --}\label{def:confidentiality}
-An agent preserves the property of confidentiality iff 
-\end{definition}
-
-\begin{definition}{\bf System-Confidentiality --}\label{def:confidentiality}
-	The property of confidentiality holds in a system when its information
-	preserves confidentiality.
-\end{definition}
+The quality of the confidentiality of information (Assertions or Beliefs depending on 
+the subsystem considered) are inversely proportional to the number of 
+integrity issues of the subsystem.
 
 \fix{mr}{
-Confidentiality: the information is accessible only to authorised individuals.
-A set of rules that limits access to information.  The information can be
-accessed by X if and only if X is authorized to access it = X has the right
-attributes (i.e. keys, password) to access it.
-
-Integrity: the information is trustworthy and accurate (i.e. uncorrupted and
-unaltered). Protects data from modification or deletion by unauthorized sources
-and the ability to undo any damage done.  (for data at rest) If the information
-X is written, A will always be read. If information X is modified to
-information X', information X can always be recovered.  (for data in transit)
-If the information X is sent, A will be received. If information X is modified
-to information X', information X can always be recovered.
-
-Availability: the information is accessible and usable when required. Guarantee
-of reliable access to the information by authorized people If the information X
-needs to be accessed, information X can be retrieved. If
-
 Authenticity: assurance that the information is from the source it claims to be

### 2020-04-02 engineering.tex
review introduction

@@
 	As depicted in Figure~\ref{fig:usecase}, the use case of the running example
-	(a small part of a subprocess of the SwAT testbed \autocite{}), shows
+	(a small part of a subprocess of the SwAT testbed \autocite{Mathur2016swat}), shows
 	\begin{itemize}[noitemsep]

### 2020-04-02 introduction.tex
review introduction

@@
 the number of vulnerabilities was above $16000$.  The scientific community also
-reports similar findings.  In fact, in\autocite{Herley2009so}, Cormac Herley
+reports seemingly unjustifiable findings.  In fact, in\autocite{Herley2009so}, Cormac Herley
 (Microsoft Research) shows how basic cybersecurity principles (such as the
@@
 
-In this article, we give the first scientific theory (to the best of our
-knowledge) on security.
-
-\Paragraph{Structure} In Section~\ref{sec:problem} we define and formalize the
-problem statement.  In Section~\ref{sec:theory} we outline our security theory,
-and in Section~\ref{sec:prototest} we describe the implementation of the theory
-and some empirical tests of the theory.  Finally, in Section~\ref{sec:related}
-we conclude the paper with an overview of the related work.
+The method applied in our line of research follows what the scientific method
+mandate. We start by detailing the problem statement by reporting the state of
+the art in Section~\ref{sec:problem}, we formulate (inductively)\footnote{``So,
+whenever they argue ``Every man is an animal and Socrates is a man; therefore
+Socrates is an animal,'' proposing to deduce from the universal proposition
+``every man is an animal'' the particular proposition ``Socrates therefore is
+an animal,'' which in fact goes (as we have mentioned) to enstablish by way of
+induction the universal proposition, the fall into the error of circular
+reasoning, since they are enstablishing the universal proposition inductively
+by means of each of the particulars and deducing the particular
+proposition from the universal syllogistically.'' Sextus Empiricus, Outlines of
+Pyrrhonism II-195 \autocite{Empiricus1990Pyrrhonism}} hypothesis based on our
+review (Section~\ref{sec:glossary}) which we use to propose a theory in
+Section~\ref{sec:theory} that apply to a abstract and standard (i.e. based on
+standards) secure process development lifecycle based on requirement
+engineering in Section~\ref{sec:engineering} (with an ad-hoc running example,
+based on the SWaT testbed \autocite{Mathur2016swat}).  We propose a tool for
+the risk assessment of a design based on our theory in Section~\ref{sec:tool}.
+Finally, in Section~\ref{sec:conclusion} we conclude the paper.

### 2020-04-02 problem_statement.tex
review problem statement

@@
 approaches to the investigation of the security of systems, and the security of the design of a
-formally verified protocol is indeed falsifiable.  For example, in the protocol
+formally verified protocol is indeed falsifiable (against the security properties verified).  For example, in the protocol
 verification community, security is often defined as a formalization of the
 high-level properties confidentiality, integrity, and availability. The problem
-in such approaches is not the definition of what cybersecurity is, but the use
-of theories (such as the Dolev-Yao attacker model\footnote{For the sake of
+in such approaches is not the definition of what cybersecurity is, but 
+generality of the results since they takle a specific step (and not the first) of the engineering of system.
+Therefore, the theories underlying the verification are based on 
+assumptions which non-evidently apply to a general security theory (since, as mentioned before,
+there is no sucha theory).
+As an example, the so called Dolev-Yao attacker model\footnote{For the sake of
 simplicity, the Dolev-Yao attacker can be considered as an abstraction of an
@@
 abstraction of the protocol. This, in turn, creates a false sense of security
-since requires assumptions on the abstraction of the system of which security
+since requires non-justifiable assumptions on the abstraction of the system of which security
 is verified. As an example, for the formal security verification of the system
@@
 the abstraction of the cryptographic primitives (e.g.  ProVerif vs
-CryptoVerif\autocite{Blanchet2017symbolic}).  Some of the choices will
-completely change the results of the formal verification of the system. For
+CryptoVerif\autocite{Blanchet2017symbolic}).  While many tools and theories
+improves issues when discovered, there is no agreement on which should be
+the definitive approach if any and, more importantly, some of the choices made by the modelers/engineers using those formal apporaches may
+completely change the results of the formal verification of the system (an interesting example on how difficult it is to 
+even just compare the approaches is given in \autocite{Cremers2009comparing}). For
 example, under the perfect cryptography assumption\footnote{As defined
@@
 this highlights the false sense of security that may derive from a
-non-scientific theory of system security\footnote{``To the superficial
+non-falsifiable theory of system security\footnote{``To the superficial
 observer, the analysis of these forms seems to turn upon minutiae. It does in
@@
 belief\footnote{A belief has to be intended as a proposition which is supposed
-to be true by the majority of humans in our society without a scientific
-underlying theory but based on partial empirical evidences or inductive proofs.} that
+to be true by the majority of people in our society, without a scientific
+underlying theory but based on partial empirical evidences or inductive reasoning based on partial empirical evidences.} that
 safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
@@
 what to do.} (accidents are unfortunate while incidents are not) of the causes
-that generates the threat; namely, nature is believed not to have malicious
+that generates a threat; namely, nature is believed not to have malicious
 intents (but unfortunate causes-effects) while threats generated by mankind are
@@
 afterwards in the article) malicious, and then we consider any vulnerability to
-be malicious (without loss of generality) even if due to the lack of skills.}.
+be malicious even if due to the lack of skills.}.
 An overview on the aforementioned aspects of safety and security is depicted in
@@
 				considering those terms to be semantically
-				equivalent.  This ``shallowness'' raise form the
+				equivalent.  This ``shallowness'' has raised form the
 				necessity of properly citing the different sources, but,
@@
 				\emph{threat agents} and an \emph{attacker} as
-				a specific malicious threat agent which materialize a
+				a specific malicious threat agent which materializes a
 				threat.
@@
 		highlighting how commonly this word is used without a proper
-		supporting semantics}, as defined in\autocite{cnssi20104009}
+		supporting semantics.}, as defined in\autocite{cnssi20104009}
 		(and adopted in\autocite{nist2013800-53}), is ``weakness in an
@@
 		and Mankind, not differentiating between safety and security.
-		In Definition~\ref{def:vulnerability}, we provide a formal
+		\fix{mr}{In Definition~\ref{def:vulnerability}}, we provide a formal
 		theory of vulnerability (so that the scientific community can
@@
 	\includegraphics[width=\textwidth]{safety-security.pdf}
-	\caption{Overview security and safety keywords}
+	\caption{Overview of keywords related to security and safety}
 	\label{fig:safety-security}
@@
 safety-critical cyber-physical systems (such as elevators and aircraft), upon
-which safety requirements (e.g. in standards such as the IEC 61508 or 61511\autocite{IEC201761511}) are defined, have
-been defined following empirical tests and measurements. While reasoning by
+which safety requirements are defined (e.g. in standards such as the IEC 61508 or 61511\autocite{IEC201761511}), 
+are based on empirical tests and measurements (therefore should be considered hypothesis and not definitions). While reasoning by
 induction based on the empirical observation should be avoided, since it may
-easily lead to false beliefs instead of scientific theories, this approach is
+easily lead to false beliefs, this approach is
 often justified by the supposed impossibility of defining a theory that
@@
 correlation between predictability of environment and believed unpredictability
-of attackers (i.e. a malicious
-environment) has not been correlated to a theory on cybersecurity. 
+of attackers has not been correlated to a theory on cybersecurity. 
 Therefore, inductive research efforts in
@@
 environment (e.g. due to humidity, dust, heat \&c) is defined from empirical
-evidences and processes have been standardized to test qualities of hardware components
-%convinced by the induction principle. What is already a
-%weak argument in general, 
+evidences and processes have been standardized to test qualities of hardware components.
 This process completely breaks down when a malicious environment (i.e. an attacker)
@@
 natural environment. Therefore, the same approach that is in use for safety,
-seems not to be applicable to test security.
+seems not to be applicable for security (e.g. for security testing).
 
@@
 channel that allows the violation of the security policy
-[\ldots]''\autocite{cnssi20104009} (e.g. a software or procedure) that takes
-advantage of the vulnerability causing an \emph{attack} to the system, which may
+[\ldots]''\autocite{cnssi20104009}. For example, a software or procedure that takes
+advantage of the vulnerability causing an \emph{attack} to the system may
 result in several correlated incidents and threats.  The process of
@@
 		(the effect). The essence of causality is the generation and
-		determination of one phenomenon by another. In this respect
+		determination of one phenomenon by another. In this respect,
 		causality differs from various other kinds of connection, for
@@
 	\item An \emph{Exploit}\footnote{We note that the term exploit is only
-		used as a verb in\autocite{ISO2009information}} is ``An exploit
+		used as a verb in\autocite{ISO2009information}} ``[\ldots]
 		(from the English verb to exploit, meaning to use something to
@@
 		In fact, we consider the term \emph{hack} as
-		non-malicious (see Hacker\autocite{Stallman2002hacker}).
+		non-malicious (as, e.g. in \autocite{Stallman2002hacker}).
 	\item A \emph{Threat}, as defined in\autocite{cnssi20104009}, is ``Any

### 2020-04-02 engineering.tex
partial review (up to sec 3.4) of sec 3

@@
 
+\begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
+	A system security design is given by a system security specification, i.e. a set of
+	requirements on the integrity and availability over the physical and
+	functional architectures, that uniquely defines the designs built on
+	top of those requirements.
+\end{hypothesis}
+
+\begin{hypothesis}{\bf System Insecurity Design --}\label{hyp:insecurity}
+	If, given a system security specification as a collection of
+	requirements on the physical and functional architectures, there exist
+	a non-unique design that respect those requirements, the number of
+	equivalent designs that fullfills the requirements quantitatively
+	defines the magnitute of insecurity of a system design.
+\end{hypothesis}
+
 \fix{mr}{

### 2020-04-02 formal_glossary.tex
partial review (up to sec 3.4) of sec 3

@@
 where $\wedge$, $\neg$, $\vee$, and $\Rightarrow$ are connectives representing
-conjunction, negation, disjunction, and (material) implication respectively; while
-$\forall$ and $\exists$ represent the standard universal and existential
-quantifiers (resp.). Finally, the symbol ``$.$'' is just syntactic sugar. 
-We consider $\interpretation\subset\Phi\times\{\top,\bot\}$ as the interpretation
-function, where $\Phi$ is any collection of sentences in $\follang$ 
-and $\top$ and $\bot$ represent the concepts of ``Tautology/True''
-and ``Contradiction/False'' respectively.
+conjunction, negation, disjunction, and (material) implication respectively;
+while $\forall$ and $\exists$ represent the standard universal and existential
+quantifiers (resp.). Finally, the symbol ``$.$'' is just syntactic sugar.  We
+consider $\interpretation\subset\Phi\times\{\top,\bot\}$ as the interpretation
+function, where $\Phi$ is any collection of sentences in $\follang$ and $\top$
+and $\bot$ represent the concepts of ``Tautology/True'' and
+``Contradiction/False'' respectively.
 
@@
 
-\subsubsection{Causality Principle}\label{sec:causality}
+\subsection{Causality Principle}\label{sec:causality}
 We formalize the Causality Principle starting from a K modal
@@
 
-%\begin{definition}{\bf Causality --}\label{def:causality}
-%	The Causality principle is defined over a K Modal Logic (i.e. without
-%	restrictions on the causality relation between worlds) as in
-%	Definition~\ref{def:modallogic}, with the standard interpretation
-%	function defined in~\ref{def:modalinterpretation}
-%\end{definition}
+The causality principle has been defined in its generic form. In fact, the
+accessibility relation $\modalrelation$ is free from any axiomatic restriction
+(e.g. it's not reflexive nor anti-reflexive).  We will focus in
+Section~\ref{sec:engineering} (for our tests) on the application of our theory
+to CPS system engineering.  More detailed case studies (i.e. \fix{mr}{a CPS as
+a smart power-grid}) will be defined in Section~\ref{sec:engineering}, where
+the accessibility relation defines in details how the system itself evolves,
+due to causal relation (i.e. by restricting the causality principle only to
+those cause-effect that defines the system). Therefore, the definition of
+$\modalrelation$ will be specialized in a more strict way based on the
+application domains and case study.
 
-The causality principle has been defined in its
-generic form. In fact, the accessibility relation $\modalrelation$ is free from
-any axiomatic restriction (e.g. it's not reflexive nor anti-reflexive).  
-We will focus in Section~\ref{sec:process} (for our tests) on 
-the application of our theory to CPS system engineering.
-More detailed case studies (i.e. \fix{mr}{a CPS as a smart power-grid}) 
-will be defined in Section~\ref{sec:prototest}, where
-the accessibility relation defines in details how the system itself evolves, due to causal relation 
-(i.e. by restricting the causality principle only to those cause-effect that defines
-the system). Therefore, the definition of $\modalrelation$
-will be specialized in a more strict way based on the application domains and case study.
+We note that we have defined the Causality principle without considering it a
+\emph{threat source} since we lack the concept of intent and maliciousness.
+Similarly, in the next sections we won't discriminate between Nature and
+Mankind until we introduce the concepts of maliciousness in
+Section~\ref{sec:theory} and then formally define a threat agent in the same
+section.
 
-We note that we have defined the
-Causality principle without considering it a \emph{threat source} since 
-we lack the concept of intent and maliciousness. Similarly, in the 
-next sections we won't discriminate between Nature and Mankind until we introduce 
-the concepts of maliciousness in Section~\ref{sec:theory} 
-and then formally define a threat agent
-in the same section.
-
-\subsubsection{Agents: Mankind and Nature}\label{sec:mankind-nature}
+\subsection{Agents: Mankind and Nature}\label{sec:mankind-nature}
 \begin{table}[t]
@@
 
-Mankind and Nature are defined in Section~\ref{sec:problem} as two abstract
-agents, both as collections (i.e. an abstract type that does not imply a
+Mankind and Nature are considered as two abstract
+agents, both defined as collections (i.e. an abstract type that does not imply a
 specific implementation) of their sub-agents (i.e. humans for Mankind and
-plants, animals, \&c.  for Nature). Similarly to\autocite{Santaca2016abf}, 
-we define Mankind and Nature, and any
-other agent in the reminder of this article, as a meronomy (an hierarchy of
-Part-Whole relations) based on a standard definition of mereology, i.e. based
-on the definition of Parthood relation between \emph{Parts}.  
-However, shall 
-consider different types of Part; so, we extend the mereology to a
+plants, animals, \&c.  for Nature). Similarly to\autocite{Santaca2016abf}, we
+define Mankind and Nature, and any other agent in the reminder of this article,
+as a meronomy (an hierarchy of Part-Whole relations) based on a standard
+definition of mereology, i.e. based on the definition of Parthood relation
+between \emph{Parts}.  However, we consider different types of Part; so, we
+extend the mereology to a
 mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
-to increase the number of different types considered and to generalize the relations between Parts (as
-in Table~\ref{tab:rcc358}).
-For the sake of readability, we use the term \emph{Region} both to refer to a
-mereological Part and to a topological Region.
-The choice of mereotopology is also correlated to the objective of defining a formal
-ontology, which we use to define the (formal) semantics of the terms (Parts) in
-Section~\ref{sec:problem}, and of the concepts of safety and security
-(whole). We aim at creating a meronomy instead of the taxonomies such as the one provided
+to increase the number of different types considered and to generalize the
+relations between Parts (as in Table~\ref{tab:rcc358}).  For the sake of
+readability, we use the term \emph{Region} both to refer to a mereological Part
+and to a topological Region.  The choice of mereotopology is also correlated to
+the objective of defining a formal ontology, which we use to define the
+(formal) semantics of the terms (Parts) in Section~\ref{sec:problem}, and of
+the concepts of safety and security (whole). We aim at creating a meronomy
+instead of the taxonomies such as the one provided
 in\autocite{NIST2020NVD,MITRE2020CVE} or instead of the poorly justified
@@
 
-As later defined in Definition~\ref{def:agent}, 
-the Parthood relation orders a universe of agents $\agentuniverse$ by defining
-the so called \emph{Connects with} (see in Table~\ref{tab:rcc358}) relation 
-between Regions. We want this universe $\agentuniverse$ to be 
-expressible in FOL. In this way, we
-can reason both on the constituent of security (i.e. its terms and agents
-defining a system where security needs to be considered), and on 
-the evolution of those constituent w.r.t. cause-effects relations according
-to the modal structure of causality we defined in Section~\ref{sec:causality}.
-This will allow us (in Section~\ref{sec:process} and Section~\ref{sec:prototest})
-a better\fixnote{mr}{Can we prove this by construction?} 
-positioning w.r.t. risk assessment technologies
-(which most often reason on the constituent of a system design), and
-protocol verification tools (which requires some formalization of a cause-effect relation, e.g.
-in Linear Temporal Logic).
+The Parthood relation orders a universe of agents $\agentuniverse$ (later defined in
+Definition~\ref{def:mankind-nature}) by defining the so called
+\emph{Connects with} (see in Table~\ref{tab:rcc358}) relation between Regions.
+We want this universe $\agentuniverse$ to be expressible in FOL. In this way,
+we can reason both on the constituent of security, and on the evolution of
+those constituent w.r.t. cause-effects relations according to the modal
+structure of causality we defined.  This will allow us (in
+Section~\ref{sec:engineering} and Section~\ref{sec:tool}) a
+better\fixnote{mr}{Can we prove this by construction?} positioning w.r.t. risk
+assessment technologies (which most often reason on the constituent of a system
+design), and protocol verification tools (which requires some formalization of
+a cause-effect relation, e.g.  in Linear Temporal Logic).
 
-As we argued in Section~\ref{sec:problem}, we must correlate the definition of
-agent (i.e. Mankind and Nature) in Definition~\ref{def:mankind-nature} to the
-mathematical structure of the logic that defines them.  We express Mankind and
+In order to correlate the definition of
+agent (i.e. Mankind and Nature) to the
+mathematical structure of the logic that defines them, we express Mankind and
 Nature as formulas over the theory of mereology and then in terms of
@@
 in~\cite{bennettLogics,improvingRCC}, to provide an axiomatization of the
-spatial concepts and relations in first-order logic to correlate the algebraic
+spatial concepts and relations in FOL to correlate the algebraic
 structure to mereology. In its broader definition, the RCC theory is composed
@@
 Regions. We discuss the choice of RCC5 in more detail in
-Appendix~\ref{app:rcc}. In Table~\ref{tab:rcc}, we
+Section~\ref{sec:engsystemstate}. In Table~\ref{tab:rcc}, we
 summarize the axioms of the Region Connection Calculus (see, e.g., \autocite{Grutter2008rcc}).
@@
 
-\begin{definition}{\bf Agent: Mankind or Nature --}\label{def:agent}
+\begin{definition}{\bf Agent: Mankind or Nature --}\label{def:mankind-nature}
 	An agent $\agent\in\agentuniverse$ is a tuple
@@
 \end{definition}
-
-Currently, we do not distinguish between Mankind and Nature (since we still
-lack of the definition of ``malicious intent'', which is defined in
-Section~\ref{sec:theory}) and we have defined them as two generic
-agents.
 
@@
 
-\subsubsection{Regions: Vulnerability and Defect (and Weakness)}\label{sec:vulnerabilitydefect}
-As defined in Section~\ref{sec:problem}, a Vulnerability or a Defect is a
+\subsection{Regions: Vulnerability and Defect (and Weakness)}\label{sec:vulnerabilitydefect}
+As informally defined in Section~\ref{sec:problem}, a Vulnerability or a Defect is a
 \emph{Weakness}.  As an example, a categorization of weaknesses is given
@@
 $\Weakness,\Vulnerability$ are Regions of Weaknesses and Vulnerabilities (resp.);
-accepting the hierarchy in the CWE\autocite{CWE} as ground truth. 
-Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability) 
-or the CVE reported in the NVD as a ground truth.
+accepting (for now) the hierarchy in the CWE\autocite{CWE} as ground truth. 
+Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability),
+or the CVE reported in the NVD, as a ground truth.
 
 \begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
-	A Region $\region\subseteq\agent$ of an agent
-	$\agent\in\agentuniverse$, is defined as Weakness $\Weakness$ (i.e.
-	representing a weakness introduced by the agent into any phase of
+	A Region $\Weakness\subseteq\agent$ of an agent
+	$\agent\in\agentuniverse$, is defined as Weakness\footnote{i.e.
+	representing an error introduced by the agent into any phase of
 	production, e.g.  of the secure process development life-cycle, of any
-	system or subsystem) iff there exists $\world\in\possibleworlds$ such
-	that $\world,\Weakness\models\Diamond\exists\Weakness',\Vulnerability'.~
-	\rcc(\Weakness',\Vulnerability')\wedge\neg\dr{\Weakness'}{\Vulnerability'}$, 
-	where $\Weakness'\subseteq\Weakness$, and
-	$\Vulnerability'\subseteq\Vulnerability$ is a Region of an agent $\agent'\in\agentuniverse$.
+	system or subsystem} iff there exists $\world\in\possibleworlds$ such
+	that $\world\models\Weakness$, $\world\models\Diamond\Vulnerability$,
+	and $\neg\dr{\Weakness}{\Vulnerability}$; where $\Vulnerability$ is a
+	Region of an agent $\agent\in\agentuniverse$.
 \end{definition}
@@
 
-\subsubsection{Process: Incidents and Accidents}\label{sec:incidentaccident}
+\subsection{Process: Incidents and Accidents}\label{sec:incidentaccident}
 In our informal definition depicted in Figure~\ref{fig:safety-security}, an
@@
 	\item in Section~\ref{sec:theory}, formally define a system as a structure of agents, and
-	\item in Section~\ref{sec:process}, correlate the formal definition of agents (i.e. the
+	\item in Section~\ref{sec:engineering}, correlate the formal definition of agents (i.e. the
 		Definition~\ref{def:agent}) with the Causality principle (i.e.

### 2020-04-02 problem_statement.tex
partial review (up to sec 3.4) of sec 3

@@
 \end{itemize}
+We now proceed to the formalization of the security concepts introduced in this
+section. The main outcome of our formalization is that there is no evidence
+that security must necessarily by linked to the malicious intent of an agent.
+While this is the (only?) hypothesis behind security models and engineering
+processes we start from a different hypothesis; system \emph{in-security} and
+\emph{security} should not be searched on those events who maliciously produces
+negative effects on a systems (incidents), but in the potential designs and
+implementations due to vague specifications. Therefore, we will show that
+security-by-design with a requirement-based engineering process is to be
+preferred.
+
+The line of reasoning that we apply works as follows. The
+epistemologic concept of Knowledge can be equated to the concept of Truth.
+Anything that is known is, therefore, also true. It is also believed that
+Knowledge, defined as we did, is supposed to be inapprehensible. In contrast,
+Beliefs are potentially true and leads new hypothesis. Due to the structure of
+deductive reasoning we can only start from hypothesis and test the deductive
+conclusions on the reality. Whatever a scientific-security-theory is, it must
+be defined starting from hypothesis.  A sound reasoning must lead to conclusion
+that shall be tested on a real system. Thefore, only at the end of the reasoning
+we will be able to test, and judge, the quality of the theory itself. The identification
+of hypothesis should start from probably true beliefs (which means from
+Information), and then from the state of the art. Instead of having a ``related work''
+section, we review and cite the literature throughout the paper.

### 2020-04-02 theory.tex
partial review (up to sec 3.4) of sec 3

@@
 
-\begin{definition}{\bf Operational System State --}\label{def:system}
+\begin{definition}{\bf Operational System State --}\label{def:opsystem}
 	An operational system (or a sub-system) is represented as an agent (see Definition~\ref{def:agent}) and then defined by the resulting state space,

### 2020-04-02 formal_glossary.tex
first full review (adding comments) of sec 3

@@
 or the CVE reported in the NVD, as a ground truth. 
+\fix{mr}{it needs to be stressed that without the formalization of
+the concept of error, we cannot properly formalize the concept of weakness
+which remains just a region as any other region with the only 
+constraint that it must be connected to another region called Vulnerability (which 
+in turn depend on the definition of error)}
 
+\fix{mr}{the following definition should be moved (and detailed)
+to section~\ref{engineering} when the concept of weakness is defined}
 \begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
@@
 
-\fixnote{mr}{shouldn't we consider the agent who introduces the weakness as dishonest? don't we say this in part 1?}
+\fix{mr}{shouldn't we consider the agent who introduces the weakness as dishonest? don't we say this in part 1?}
 \end{example}
@@
 
+\fix{mr}{similar to weakness and vuln, the following definition should be kept here just to show that
+it is impossible to defined those concept without the proper definition of the concept of error. and the definition should
+be introduced in section~\ref{engineering} when the concept of weakness is defined}
 \begin{definition}{\bf Region: Vulnerability and Defect --}\label{def:vulnerability-defect}
@@
 	\item in Section~\ref{sec:engineering}, correlate the formal definition of agents (i.e. the
-		Definition~\ref{def:agent}) with the Causality principle (i.e.
+		Definition~\ref{def:mankind-nature}) with the Causality principle (i.e.
 		the formalization of the Kripke structure in

### 2020-04-06 formal_glossary.tex
full review of the formalization section

--- 2020-04-02 formal_glossary.tex
@@
-We now define a formalizations of the concepts described in
-Section~\ref{sec:problem} and depicted in Figure~\ref{fig:safety-security}. We
-base our formalization on first order logic (FOL) with the standard
-truth-value semantics. The choice of this logic is required by the 
-semantics of the concepts formalized afterwards in this section.
+A formalization requires a formal logic and a formal language.  We base our
+formalization on first order logic (FOL) with the standard truth-value
+semantics. The choice of this logic is usefull for the fomal treatment of the
+theory we'll define in Section~\ref{sec:theory}.  We do not provide a proof
+that the choice of FOL doesn't impede the formulation of a correct security
+theory, nor that the semantics of English isn't an issue. So, we believe that
+this doesn't pose an issue for the formulation of the seucirty hypotheses since
+the hypothesis requires to be able to define \emph{an agent and its
+honest/dishonest behaviors in such a way that an agent can be used to define a
+system}. To the best of our knowledge, there is no result that shows that FOL
+cannot be used to do so; on the contrary, using classical logic such as FOL or
+Propositional logic is rather standard (along with Kripke structures for the
+definition of the underlying causality principle).
 
@@
 We formalize the Causality Principle starting from a K modal
-logic\autocite{Garson2018modal} (i.e. without restrictions on the causality
-relation between worlds). The standard definition of K modal logic is given in
+logic\autocite{Garson2018modal} so that (i) we use a de-facto standard
+way of defining the Causality Principle, and (ii) we do not restrict the causality
+relation (between worlds) in any way (so we are as general and abstract as possible).
+The standard definition of K modal logic is given in
 Definition~\ref{def:modallogic} in terms of an interpretation function (which
@@
 
-The causality principle has been defined in its generic form. In fact, the
-accessibility relation $\modalrelation$ is free from any axiomatic restriction
-(e.g. it's not reflexive nor anti-reflexive).  We will focus in
-Section~\ref{sec:engineering} (for our tests) on the application of our theory
-to CPS system engineering.  More detailed case studies (i.e. \fix{mr}{a CPS as
-a smart power-grid}) will be defined in Section~\ref{sec:engineering}, where
-the accessibility relation defines in details how the system itself evolves,
-due to causal relation (i.e. by restricting the causality principle only to
-those cause-effect that defines the system). Therefore, the definition of
-$\modalrelation$ will be specialized in a more strict way based on the
-application domains and case study.
+As mentioned before, the accessibility relation $\modalrelation$ is free from
+any axiomatic restriction (e.g. it's not reflexive nor anti-reflexive).  We
+note that we have defined the Causality principle without considering it a
+\emph{threat source} since, given that this is the first and outermost concept
+in Figure~\ref{fig:security-safety}, we lack the concept of intent and
+maliciousness (which are not inherent to the Causality Principle). 
+In other words, the Causality principle or the Kripke structure
+doesn't correlate cause-effect (or worlds) due to any other principle but
+causality itself; therefore, in order to distinguish nominal or honest cause-effects
+and non-nominal or dishonest ones we would need to add additional constraints, defining
+what honesty is in a factious way.
 
-We note that we have defined the Causality principle without considering it a
-\emph{threat source} since we lack the concept of intent and maliciousness.
-Similarly, in the next sections we won't discriminate between Nature and
-Mankind until we introduce the concepts of maliciousness in
-Section~\ref{sec:theory} and then formally define a threat agent in the same
-section.
+Similarly, in the next sections we won't be able to discriminate between Nature
+and Mankind due to their honesty. We argue that the concept of honesty, even if
+we do not state that it is not intrinsic to all the definition of Nature or
+Mankind, requires additional constraints or statements in order to be defined.
 
@@
 definition of mereology, i.e. based on the definition of Parthood relation
-between \emph{Parts}.  However, we consider different types of Part; so, we
+between \emph{Parts}.  Due to the necessity of considering different types of Part (as we'll show afterwards) we
 extend the mereology to a
@@
 
+We have now defined the concept of agent, and defined Mankind and Nature as
+agents. Given that, as we stated beforehand, they are defined by their
+subagents, and that we are not interested in a general definition of Mankind and
+Nature but in the one that serves our purpose (i.e. defining a security
+theory), in the following sections we focus on those terms identified in the
+literature review.
+
 \subsection{Regions: Vulnerability and Defect (and Weakness)}\label{sec:vulnerabilitydefect}
@@
 Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability),
-or the CVE reported in the NVD, as a ground truth.
+or the CVE reported in the NVD, as a ground truth in the case of Vulnerability. 
+It is then obvious that without the formalization of
+the concept of error,
+we cannot properly formalize the concepts of Weakness (and Vulnerability);
+which remains one of the sub-region of errors with the only constraint that it
+must be connected to another region called Vulnerability (which in turn depend
+on the definition of error).
 
 \begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
-	A Region $\Weakness\subseteq\agent$ of an agent
-	$\agent\in\agentuniverse$, is defined as Weakness\footnote{i.e.
+	A Region $\Weakness\subseteq\Error$ of a region $\Error$ of errors,
+	is defined as Weakness\footnote{i.e.
 	representing an error introduced by the agent into any phase of
@@
 	that $\world\models\Weakness$, $\world\models\Diamond\Vulnerability$,
-	and $\neg\dr{\Weakness}{\Vulnerability}$; where $\Vulnerability$ is a
-	Region of an agent $\agent\in\agentuniverse$.
+	and $\world\models\Diamond\neg\dr{\Weakness}{\Vulnerability}$.
 \end{definition}
@@
 Figure~\ref{fig:rcc5_CWE-116}, or EC connection in RCC8 (Figure~\ref{fig:rcc8_CWE-116}).
-
-\fixnote{mr}{shouldn't we consider the agent who introduces the weakness as dishonest? don't we say this in part 1?}
 \end{example}
@@
 details on the implementation/realization in Figure~\ref{fig:safety-security})
-that we analyze in the next section, Section~\ref{sec:incidentaccident}.
+that we analyze in Section~\ref{sec:incidentaccident}.
+
+\Paragraph{Attacker as a Dis-honest Agent}
+As we mentioned in Section~\ref{sec:problem_statement} when we postulated the
+security hypothesis, Weaknesses and Vulnerability are the two core security
+concepts of the current understanding of what security is. However, we see two
+(opposed in terms of abstraction) possibility to better defined those concepts:
+defining how an agent (Mankind or Nature) can maliciously use Weaknesses to
+produce Vulnerability (and, in turn, Exploits), or how Errors can be defined
+(since Error is a term from which Weakness are, in turn, defined).  The former,
+in our current formalization, implies detailing and constraining the Causality
+Principle by introducing constraints on how agents acts. For example,
+introducing the concept of intentionality, motivation, or desire (see the BDI
+framework defined in \autocite{BDI} for a general overview of those concepts).
+The symbolic approach (also called Dolev-Yao model) in the formal verification
+of security and communication protocols, can be used in this spirit. The
+attacker can be seen as an additional agent (often identified with the network
+where the messages are exchanged) that sees the network traffic and manipulates
+it. The actions are generic (the attacker can combine two messages, split a
+message into its components, or decrypt/encrypt if it has the proper security
+key) but not something that a dishonest agent cannot do. The main difference is
+that an honest agent follows the nominal specification of the protocol while
+the attacker doesn't. The attacker non-deterministically opens as many new
+sessions as it wants and the causality structure in which this scenario
+(protocol + attacker) has been defined, is searched for a state that is
+considered erroneous with respect to some security requirements decided a
+priori. Therefore, as we mentioned in Section~\ref{sec:problem_statement}, the
+formal verification of the security of protocols can be considered a scientific
+security theory; the implementations resulting from the formalized (and
+verified) design can be tested against the formalized security requirements.
+So, it is not completely true that ``Thus, formal approaches offer no escape
+from our basic problem; only by making unfalsifiable assumptions (about what an
+attacker can do) will they allow derivation of a necessary condition'' (as
+Herley describes in\autocite{Herley2009so}) rather, the
+unfalsifiability-related issue seems to stem from how security conditions are
+defined, i.e. how security requirements are defined. One can easily see that
+this is not just an issue in the formal protocol verification community but the
+very foundations of the concept of Weakness and Vulnerability are defined in a
+way that makes it impossible (they are necessary conditions) to make a general
+security statement on any system (even if we tested every single Weakness on
+the CWE on our system it may be the case that a Weakness han yet to be
+discovered). Therefore, we hypothesize that a security theory should define in
+a complete way how errors (and then Weaknesses, Vulnerability, and Exploits)
+may be introduced.
+
+We now conclude (for completeness) the analysis of the various security-related terms we identified
+in the literature review, without formally defining what Errors are (which will be defined
+in the reminder of the paper).
 
@@
 \begin{enumerate}
-	\item An agent (Mankind or Nature) defines a system, e.g.
-by following an engineering process for creating a new system, or for redesigning or improving an existing system (e.g.
-a component or a collection of components)
-	\item The dogmatic definition of a system contains a Weakness (i.e.
-the agent, Mankind or Nature, introduces an unintended behavior), for example as an un-intended behavior or as an architectural fallacy.
-The Weakness can be applied to the abstract definition of the system.
-	\item The Weakness can be defined as a malicious process for a system, called Vulnerability (or Defect for Nature)
-	\item The process expressed by the Vulnerability can be made operational for a specific implementation of the system (i.e. implemented by Mankind or realized by Nature) 
+	\item An agent (Mankind or Nature) defines a system, e.g. by following
+		an engineering process for creating a new system, or for
+		redesigning or improving an existing system (e.g. a component
+		or a collection of components) 
+	\item The axiomatic/dogmatic definition of a system contains a Weakness (i.e.
+		the agent, Mankind or Nature, introduces an unintended
+		behavior), for example as an un-intended behavior or as an
+		architectural fallacy, and this Weakness can be applied to the
+		abstract definition of the system.
+	\item The Weakness can be defined as a malicious process for a system,
+		called Vulnerability (or Defect for Nature)
+	\item The process expressed by the Vulnerability can be made
+		operational for a specific implementation of the system (i.e.
+		implemented by Mankind or realized by Nature) 
 \end{enumerate}
@@
 	\item While understanding the exploitation process is the aim of the
-		CWE and CVE initiatives, understanding the hacking process is
+		CWE and CVE initiatives, understanding the malicious (or black-hat) hacking process is
 		the aim of the CAPEC initiative.
@@
 	\item An agent (Mankind\footnote{For the sake of readability we focus
-		only on Mankind as an agent}) the exploitation process can be
-		used to hack a system. We use the term hack (as hacking) to
-		stress that nothing prevents this process from being honest,
-		however, for the sake of simplicity, we focus on Attacks which
-		are malicious by definition.
+		only on Mankind as an agent}) starts the exploitation process
+		to hack a system. We use the term hack (as hacking) to stress
+		that nothing prevents this process from being honest (i.e. the
+		so called honest-but-curious attacker), however, for the sake
+		of simplicity, we focus on Attacks which are malicious by
+		definition.
 	\item An Attack is applied to a system with malicious intent by an
@@
 \end{itemize}
+
+The main outcome of this section is that there is no evidence that properly
+considering the malicious intent of an agent necessarily results in a security
+scientific theory.  On the other hand, the complete definition of Errors,
+predicted by a (sound w.r.t. Errors) security theory seems to be a proper
+research direction.  In other words, system \emph{in-security} and
+\emph{security} should not be searched on those events who maliciously produces
+negative effects on a systems (incidents), but in the potential designs and
+implementations due to vague specifications. Therefore, we will consider a
+security-by-design approach based on a requirement-based engineering process.

### 2020-04-06 introduction.tex
full review of the formalization section

--- 2020-02-20 introduction.tex
@@
 \epigraph{Humanum est errare}{{\itshape Seneca the Elder}}
-The European Commission states in\autocite{EU2019market} that: ``Cybersecurity
-is one of the priority areas [\ldots] of the Commission initiative on ICT
-Standards, which is part of the Digitising European
-Industry\autocite{EU2019standard} strategy launched on 19 April 2016.  The aim
-is to identify the essential ICT standards and present measures to accelerate
-their development in support of digital innovations across the economy''. The
-same document (i.e.\autocite{EU2019market}) states that ``The EU will invest up
-to \euro450 million [\ldots], under its research and innovation programme
-Horizon 2020''. The EU, in 2016 published a press release\autocite{EU2016press}
-in which they present a strategy to invest \euro1.8 \emph{billion} to
-``increase measures to address cyber threats''. The EU is not the only investor
-in cybersecurity, most of the developed countries and several companies are
-investing enormous amount of money towards various aspects of cybersecurity
-(e.g. The US vulnerability databases\autocite{NIST2020NVD} maintained by the
-National Institute of Standards and Technologies, i.e. NIST, of the US
-Department of Commerce).
-
-The cybersecurity industry is growing fast, e.g. as reported
-in\autocite{Nasdaq2018market}. For example, in\autocite{Forbes2017market},
-published by the Forbes, is stated that \euro5.3 billion of funding where
-poured by venture capitalist into cybersecurity companies in 2018. The Forbes,
-in the same article, also highlights another peculiar (as seemly contradictory)
-trend: ``[\ldots] during the same time period, the number of cybersecurity
-breaches increased exponentially''. The data reported by the NIST 
-through the official CPE (Common Platform Enumeration) Dictionary Statistics 
-on the NVD websites in \autocite{NIST2020CPEstatistics}, show that in 2016 the
-number of reported vulnerabilities reported where around $~6000$ while in 2019
-the number of vulnerabilities was above $16000$.  The scientific community also
-reports similar findings.  In fact, in\autocite{Herley2009so}, Cormac Herley
-(Microsoft Research) shows how basic cybersecurity principles (such as the
-confidentiality benefit over the clear text for passwords typed into forms,
+The cybersecurity industry is growing fast (see Appendix~\ref{app:economy} for
+a longer economic motivation), e.g. as reported in\autocite{Nasdaq2018market}.
+For example, in\autocite{Forbes2017market}, published by the Forbes, is stated
+that \euro5.3 billion of funding where poured by venture capitalist into
+cybersecurity companies in 2018. The Forbes, in the same article, also
+highlights another peculiar (as seemly contradictory) trend: ``[\ldots] during
+the same time period, the number of cybersecurity breaches increased
+exponentially''. The data reported by the NIST through the official CPE (Common
+Platform Enumeration) Dictionary Statistics on the NVD websites in
+\autocite{NIST2020CPEstatistics}, show that in 2016 the number of reported
+vulnerabilities reported where around $~6000$ while in 2019 the number of
+vulnerabilities was above $16000$.  The scientific community also reports
+seemingly unjustifiable findings.  In fact, in\autocite{Herley2009so}, Cormac
+Herley (Microsoft Research) shows how basic cybersecurity principles (such as
+the confidentiality benefit over the clear text for passwords typed into forms,
 e.g. for logins in websites) are not fully understood or shared between the
@@
 cybersecurity, as already highlighted by Cormac Herley
-in\autocite{Herley2016unfalsifiability}.
+in\autocite{Herley2016unfalsifiability}.  An important issue raised by Herley
+is that the methodology applied to by the security research community is quite
+often implicit and not always scientifically ``correct''. Herley equates
+correct to what Popper defined in \autocite{popper1962conjectures} even if, we
+believe, the comments made by Hintikka in \autocite{Hintikka1993Information}
+should be taken into account. Therefore, without discussing what correct and
+non-correct is, we start by explicitly mention our method of enquiry.
 
-Another way of looking at the same problem is by analyzing the different
-definitions of security with respect to the scientific method of enquiry
-applied to get to the definition itself. In the following we categorize the
-related work into three categories, as defined by Sextus Empiricus in
-\autocite{Empiricus1990Pyrrhonism}.  To the best of our knowledge, there is no
-evidence that this categorization isn't complete.  ``The natural result of any
-investigation is that the investigators either discover the object of search or
-deny that it is discoverable and confess it to be in-apprehensible or persist
-in their search. [\ldots] This is probably
-why''\autocite{Empiricus1990Pyrrhonism}: 
-\begin{itemize}
-	\item The \emph{dogmatists} ``have claimed to have discovered the truth'' on what cybersecurity is
-		\begin{itemize}
-			\item Wikipedia defines cybersecurity in
-				\autocite{wiki-cybersecurity} as the protection
-				of computer systems and networks from the theft
-				of or damage to their hardware, software, or
-				electronic data, as well as from the disruption
-				or misdirection of the services they provide
-		\end{itemize}
-	\item The \emph{academics} ``have asserted that it cannot be apprehended''
-		\begin{itemize}
-			\item Eugene H. Spafford, Professor at Purdue
-				University, defines cybersecurity as follow.
-				``The only truly secure system is one that is
-				powered off, cast in a block of concrete and
-				sealed in a lead-lined room with armed guards —
-				and even then I have my doubts.''
-				\autocite{Spafford2019Quotes}
-		\end{itemize}
-	\item The \emph{skeptics} ``go on inquiring''
-		\begin{itemize}
-			\item Cormac Herley reaches the conclusion that
-				cybersecurity has no definition. ``There is an
-				inherent asymmetry in computer security: things
-				can be declared insecure by observation, but
-				not the reverse. There is no test that allows
-				us to declare an arbitrary system or technique
-				secure. This implies that claims of necessary
-				conditions for security are unfalsifiable. ''
-				\autocite{Herley2016unfalsifiability}
-		\end{itemize}
-\end{itemize}
-
-In this article, we give the first scientific theory (to the best of our
-knowledge) on security.
-
-\Paragraph{Structure} In Section~\ref{sec:problem} we define and formalize the
-problem statement.  In Section~\ref{sec:theory} we outline our security theory,
-and in Section~\ref{sec:prototest} we describe the implementation of the theory
-and some empirical tests of the theory.  Finally, in Section~\ref{sec:related}
-we conclude the paper with an overview of the related work.
+\Paragraph{Methodology} The method applied in our line of research follows what
+(to the best of our knowledge) the scientific method mandates. We start, in
+Section~\ref{sec:problem}, by detailing the problem statement, reporting a
+literature review on the main concepts and definitions related to security.  We
+formulate (inductively)\footnote{``So, whenever they argue ``Every man is an
+animal and Socrates is a man; therefore Socrates is an animal,'' proposing to
+deduce from the universal proposition ``every man is an animal'' the particular
+proposition ``Socrates therefore is an animal,'' which in fact goes (as we have
+mentioned) to enstablish by way of induction the universal proposition, the
+fall into the error of circular reasoning, since they are enstablishing the
+universal proposition inductively by means of each of the particulars and
+deducing the particular proposition from the universal syllogistically.''
+Sextus Empiricus, Outlines of Pyrrhonism II-195
+\autocite{Empiricus1990Pyrrhonism}} a security hypothesis, based on our review,
+in Section~\ref{sec:glossary}; which we use to propose a theory on system security in
+Section~\ref{sec:theory}. We apply our theory to an abstract and standard (i.e.
+based on standards) secure process development lifecycle, following the
+principle of requirement engineering in Section~\ref{sec:engineering} (with an
+ad-hoc running example, based on the SWaT testbed \autocite{Mathur2016swat}).
+We implement, in Section~\ref{sec:tool}, the theory in a prototype tool that can be used for the risk
+assessment of a design of cyber-physical systems.

### 2020-04-06 problem_statement.tex
full review of the formalization section

--- 2020-04-02 problem_statement.tex
@@
+Before going into the details of the literature review, we show that the lack
+of agreement on what security is can actually be found in the literature, and
+can be categorized with respect to the method of enquiry.  In the following, we
+provide three examples based on the categorization provided by Sextus Empiricus
+in \autocite{Empiricus1990Pyrrhonism}, since it appear to us to be
+comprehensive. ``The natural result of any investigation is that the
+investigators either discover the object of search or deny that it is
+discoverable and confess it to be in-apprehensible or persist in their search.
+[\ldots] This is probably why''\autocite{Empiricus1990Pyrrhonism}: 
+\begin{itemize}
+	\item The \emph{dogmatists} ``have claimed to have discovered the truth''
+		\begin{itemize}
+			\item Wikipedia defines cybersecurity in
+				\autocite{wiki-cybersecurity} as the protection
+				of computer systems and networks from the theft
+				of or damage to their hardware, software, or
+				electronic data, as well as from the disruption
+				or misdirection of the services they provide.
+		\end{itemize}
+	\item The \emph{academics} ``have asserted that it cannot be apprehended''
+		\begin{itemize}
+			\item Eugene H. Spafford, Professor at Purdue
+				University, defines cybersecurity as follow.
+				``The only truly secure system is one that is
+				powered off, cast in a block of concrete and
+				sealed in a lead-lined room with armed guards —
+				and even then I have my doubts.''
+				\autocite{Spafford2019Quotes}
+		\end{itemize}
+	\item The \emph{skeptics} ``go on inquiring''
+		\begin{itemize}
+			\item Cormac Herley reaches the conclusion that
+				cybersecurity has no definition. ``There is an
+				inherent asymmetry in computer security: things
+				can be declared insecure by observation, but
+				not the reverse. There is no test that allows
+				us to declare an arbitrary system or technique
+				secure. This implies that claims of necessary
+				conditions for security are unfalsifiable. ''
+				\autocite{Herley2016unfalsifiability}. Therefore, we propose our investigation.
+		\end{itemize}
+\end{itemize}
+
 \begin{figure}[t]
@@
 in such approaches is not the definition of what cybersecurity is, but 
-generality of the results since they takle a specific step (and not the first) of the engineering of system.
+the generality of the results since they takle a specific step (and not the first) of the engineering process (of system).
 Therefore, the theories underlying the verification are based on 
 assumptions which non-evidently apply to a general security theory (since, as mentioned before,
-there is no sucha theory).
+there is no such a theory).
 As an example, the so called Dolev-Yao attacker model\footnote{For the sake of
@@
 since requires non-justifiable assumptions on the abstraction of the system of which security
-is verified. As an example, for the formal security verification of the system
+is verified. More specifically, for the formal security verification of the system
 in Figure~\ref{fig:protocol-example}, a formalized scenario needs to be defined
@@
 CryptoVerif\autocite{Blanchet2017symbolic}).  While many tools and theories
-improves issues when discovered, there is no agreement on which should be
+improves the issues we discussed, there is no agreement on which should be
 the definitive approach if any and, more importantly, some of the choices made by the modelers/engineers using those formal apporaches may
@@
 
-\subsection{Sicurezza: Safety and Security}\label{sec:sicurezza}
+\subsection{Terminology: Safety or Security}\label{sec:sicurezza}
 In most of the natural languages, and in Italian too, the concepts of safety
@@
 \end{itemize}
+
+Either a valid security hypothesis can be formulated based on one or more of
+the concepts reviewed (and depicted in Figure~\ref{fig:safety-security}) or it
+can't.  Since, in the next Section, we will define our security hypothesis (on
+top of which we'll deveop a security theory) on those terms, if the second case
+is true we will not formulate a valid security theory. However, while we want
+to highlight that the incompleteness of this review may be a potential issue;
+on the other hand, due to the nature of a literature review, it seems to us
+that only an inductive approach can be used and then any literature review will
+be incomplete.  If the first case applies, in one of those terms one shall find
+a valid ground on top of which formulate a security hypothesis (assuming that a
+valid security hypothesis exists).
+
+Our literature review shows that most of the definitions relates insecurity to
+dis-honesty (also called maliciousness or adversarial) of an agent (often
+called adversary or attacker).  If those correlation are important for the
+insecurity of systems, it is plausible that they are important for the opposite
+concept, security.  Therefore, we define the following hypothesis.
+\begin{hypothesis}{\bf Security as Honesty and Insecurity as Dishonesty --}
+	\begin{itemize}
+		\item a closed system can be considered \emph{secure} if there
+			is no dishonest agent
+		\item a closed system can be considered \emph{insecure} if
+			there is a dishonest agent 
+	\end{itemize}
+	where an agent is any virtual or physical entity of the system or using
+	the system (e.g. a device, a software, or a human being) and dishonesty
+	is not necessarily related to malicious motivation but also to
+	incompetence or lack of skills.
+\end{hypothesis}
+As in the Dolev-Yao theory, we may correlate ``being dishonest'' to ``not
+following the intended behavior/rules''.  In the case of a generic system, a
+dishonest agent is therefore, any agent that doesn't follow the intended
+behavior (or functionality or logic) of the system.  Given the generality of
+the definition, and its high level of abstraction, we may
+conclude that the hypothesis seems evident. For example, a software can be
+considered an agent of the system and whenever it has a bug, it can be
+exploited causing an Incident. However, this hypothesis has lead to the
+un-testable conclusion that the dishonest behaviors of agents cannot be defined
+in general (e.g. due to the heterogeneity of agents and systems) 
+and that huge repository of dishonest behaviors should be kept as
+definition, such as CWE\autocite{MITRE2020CWEresearch}, CVE\autocite{CVE},
+CAPEC\autocite{MITRE2020CAPEC}, NVD\autocite{NIST2020NVD} and many many other;
+or that dishonesty of an agent with respect to a security protocol should be
+defined as a number of predefined actions as in
+\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
+(to name a few).
+
+We now proceed with a more detailed analysis of the terms
+and with the re-formulation of the security hypothesis.

### 2020-04-06 theory.tex
full review of the formalization section

@@
 
+\fix{mr}{
+\Paragraph{Methodology}
+We now conclude this section by detailign the line of reasoning we want to
+apply to our investigation.
+The epistemologic concept of Knowledge can be equated to the
+concept of Truth.  Anything that is known is, therefore, also true. It is also
+believed that Knowledge, defined as we did, is supposed to be inapprehensible.
+In contrast, Beliefs are potentially true and leads new hypothesis. Due to the
+structure of deductive reasoning we can only start from hypothesis and test the
+deductive conclusions on the reality. Whatever a scientific-security-theory is,
+it must be defined starting from hypothesis.  A sound reasoning must lead to
+conclusion that shall be tested on a real system. Thefore, only at the end of
+the reasoning we will be able to test, and judge, the quality of the theory
+itself. The identification of hypothesis should start from probably true
+beliefs (which means from Information), and then from the state of the art.
+Instead of having a ``related work'' section, we review and cite the literature
+throughout the paper.
 If we consider \autocite{Empiricus1990Pyrrhonism}, the author applies a
@@
 (ii) knowledge require truth, (iii) knowledge must be \emph{properly justified}.
+}
 

### 2020-04-07 formal_glossary.tex
review of Theory section

@@
 The main outcome of this section is that there is no evidence that properly
-considering the malicious intent of an agent necessarily results in a security
-scientific theory.  On the other hand, the complete definition of Errors,
+considering the malicious intent of an agent necessarily results in a falsifiable or general security
+theory.  On the other hand, the complete definition of Errors,
 predicted by a (sound w.r.t. Errors) security theory seems to be a proper
-research direction.  In other words, system \emph{in-security} and
-\emph{security} should not be searched on those events who maliciously produces
+research goal.  In other words, system \emph{in-security} and
+\emph{security} should not be searched on those malicious events which produces
 negative effects on a systems (incidents), but in the potential designs and

### 2020-04-07 theory.tex
review of Theory section

--- 2020-03-31 theory.tex
@@
-In order to address the problem raised by Herley, 
-we shall define how to distinguish between a secure and an insecure system.
-While most of the literature (and this paper so far) have
-correlated the problem of in-security to the maliciousness of
-agents interacting with the system, we show that security
-doesn't seem to stem from a malicious nature but, rather, insecurity 
-raises from the lack of well-defined security requirements for the
-design process of a system. We argue that the high number
-of security vulnerability reported today are simply the realization
-of potential system configurations, which deviate from the nominal
-behavior just because the \emph{intended behavior}\fixnote{mr}{maybe ``nominal'' instead of ``intended'' is better and behavior can be confused with the
-relation between assertions and beliefs which is not what we mean} of the system is not precisely 
-defined in the specification at the very early stages of the engineering process. 
-A reference engineering process (so called, system security lifecycle of a product)
-is described, in details, in Section~\ref{sec:process}. However,
-for the sake of simplicity, the reader can assume for this section that 
-we are currently dealing with just the relation between 
-the specification of a system and its design.
-
-Given that we apply our theory to \emph{systems engineering} and that
-the two foundational processes (exploitation and hacking) are based on the
-concept of a system, we shall begin by defining what a system is, in its
-abstract form.
+In order to address the problem raised by Herley, we shall define how to
+distinguish between a secure and an insecure system.  While most of the
+literature have correlated the problem of in-security
+to the maliciousness of agents interacting with the system, we show that
+security doesn't seem to stem from a malicious nature but, rather, insecurity
+raises from the lack of well-defined security requirements for the design
+process of a system. We argue that the high number of security vulnerability
+reported today are simply the realization of potential system configurations,
+which deviate from the nominal behavior just because the \emph{intended (or
+nominal) behavior} of the system is not precisely defined in the specification
+at the very early stages of the engineering process.
 
 \subsection{A Reference Model for Cybersecurity Engineering}\label{sec:vmodel}
-In order to describe the $\abf$-theory we briefly introduce the 
-engineering concepts (detailed and extended in Section~\ref{sec:engineering})
-underlying the $\abf$-theory; namely, the first three stages of the, so called,
-engineering V-Model. Our definition is partial and we only use it for the 
-sake of simplicity and to better present the theory.
+A reference engineering process (so called, system security lifecycle of a
+product) is described, in Section~\ref{sec:engineering}.  For the sake of
+simplicity, we now give an high-level overview of a reference engineering model
+focusing on the first three stages of the, so called, engineering V-Model. Our
+definition is partial and we only use it for the sake of simplicity and to
+better present the theory.
 
@@
 should follow an engineering model. Our reference engineering process 
-(inspired by the waterfall V-Model\autocite{waterfallvmodel}), as depicted in
+(inspired by the waterfall V-Model), as depicted in
 Figure~\ref{fig:simplifiedvmodel}, is structured as the following ordered list
@@
 	\item The process then continues with the definition of a design where
-		the requirements of the previous phase holds; detailing the
-		logic of the physical (e.g. the physical-layer encoding of the
-		information) and functional architectures (e.g. the protocol
+		the requirements of the previous phase holds; structuring the
+		physical architecture along with its logic (e.g. the physical-layer encoding of the
+		information), and the functional architectures and its logic (e.g. the protocol
 		logic of communication over a number of potential sessions).
@@
 
+Given that we apply our theory to \emph{systems engineering} and that
+the two foundational processes (exploitation and hacking) are based on the
+concept of a system, we shall begin by defining what a system is, in its
+abstract form.
+
 \subsection{Abstract System}\label{sec:systemstate}
@@
 system as an agent and then extend the definition to a ``combination of
-interacting'' agents.  We use the same process of \autocite{Santaca2016abf} but
-over slightly different terms. The differences and commonalities with
-\autocite{Santaca2016abf} are described in Appendix~\ref{app:abf}.
+interacting'' agents.
 
@@
 		of Information and the difference with Knowledge and Belief.
-	\item In \autocite{Empiricus1990Pyrrhonism}, the author states: ``The
-		logical criterion also may be used in three senses -- of the
-		agent, or the instrument, or the ``according to what''; the
-		agent, for instance, may be a man, the instrument either sense
-		perception or intelligence, and the ``according to what'' the
-		application of the impression ``according to'' which the man
-		proceeds to judge by means of one of the aforesaid
-		instruments.'' 
+	%\item In \autocite{Empiricus1990Pyrrhonism}, the author states: ``The
+	%	logical criterion also may be used in three senses -- of the
+	%	agent, or the instrument, or the ``according to what''; the
+	%	agent, for instance, may be a man, the instrument either sense
+	%	perception or intelligence, and the ``according to what'' the
+	%	application of the impression ``according to'' which the man
+	%	proceeds to judge by means of one of the aforesaid
+	%	instruments.'' 
 	\item In \autocite{Santaca2016abf}, the authors defines an agent as a
@@
 
-If we consider \autocite{Empiricus1990Pyrrhonism}, the author applies a
-``logical criterion'' to an agent, for instance a man, for instance a device in
-a system (or a system) as in \autocite{Santaca2016abf}.  The instrument identified is sense
-perception, because a man is considered as example, but we can abstract from the
-concept of instrument and consider any received information, where information is
-considered as in \autocite{Hintikka1993Information}; which can be
-summarized by the following axiom: ``Information is specified by specifying
+\Paragraph{Methodology}
+Before going into the details of the various terms and their applications for
+the definition of our theory, we 
+detail our methodology.
+The concept of Knowledge \autocite{Hintikka1962knowledge} (from an
+epistemological point of view) can be defined as the way to get to the Truth.
+Anything that is known is considered true\footnote{It is also
+believed that Knowledge, defined as we did, is in-apprehensible \autocite{Empiricus1990Pyrrhonism}, but this shouldn't be an issue since it is a common problem to every scientific search}
+In contrast, Beliefs \autocite{Hintikka1962knowledge} are potentially true and
+leads to new hypothesis. Due to the
+structure of deductive reasoning we can only start from unproven hypothesis and test the
+deductive conclusions on the reality. Therefore, whatever a scientific-security-theory may be,
+it can only be defined starting from hypotheses.  A sound reasoning should lead to
+conclusion that shall be tested on a real system. Therefore, only at the end of
+the reasoning we will be able to test, and judge, the quality of the theory
+itself. The identification of hypotheses should start from probably true
+beliefs (e.g. from Information \autocite{Hintikka1993Information}), and then from the state of the art.
+
+For our argument, 
+as in \autocite{Santaca2016abf}, an agent is composed by its knowledge, beliefs, and the information or assertions
+it provides; where knowledge is defined, as in \autocite{Steup2020epistemology}, as
+a set of proposition known by an agent, such that: (i) knowledge requires belief,
+(ii) knowledge require truth, (iii) knowledge must be \emph{properly justified}, and
+the only objective of Information is to exchange beliefs between agents.
+As defined in \autocite{Hintikka1993Information}, ``Information is specified by specifying
 which alternatives concerning the reality it admits and which alternatives
-excludes''. This means, for our argument, that considering a propositional
-variable, which admits the two alternatives True/False, its information is
+excludes''. This means that if we consider a propositional
+variable (which admits the two alternatives True/False) its information is
 defined as Believed to be True/False and not Believed to be the opposite. 
-%Similarly, when reasoning on a
-%mereotopological region w.r.t. the RCC (e.g. RCC3 admits two regions to be
-%equal, disjoint, or overlapping) information on this region should exclude
-%some of the alternatives. 
-Finally, in \autocite{Empiricus1990Pyrrhonism}
-the author states that the logical criterium can be judged according to a frame of reference.
-For our argument, as in \autocite{Santaca2016abf}, we define this concept
-as Knowledge (or Facts); where knowledge is defined, as in \autocite{Steup2020epistemology}, as
-a set of proposition known by an agent, such that: (i) knowledge requires belief,
-(ii) knowledge require truth, (iii) knowledge must be \emph{properly justified}.
-
-Before detailing what ``properly justified'' means, we discuss the mentioned
-relation between agent and belief and we introduce a more formal definition
-of knowledge and belief. For our argument, the only objective of
-Information is to exchange beliefs between agents.
 Due to the definition of Information and then with its relation to the probabilistic
@@
 information in relation with an agent's beliefs. Similarly, we consider
-beliefs to define the actual behavior of an agent or a system. On the other hand,
-Knowledge drives the nominal behavior of an agent or a system. This view
-is not considered in most (if not all) the approaches to protocol verification
-and the Dolev-Yao theory is usually applied to a representation/abstraction
+beliefs to define the actual behavior of an agent or a system. On the 
+contrary, this view is not considered in most (if not all) the approaches to formal protocol verification
+in the symbolic model. In fact, the security reasoning is usually applied to a representation/abstraction
 of a protocol that only considers knowledge and transfer of knowledge (i.e.
 if an agent sends a message, the recipient knows that message). However,
-the security of a system is tightly related to the difference between
-the nominal behavior and the actual (e.g. the implemented) one.
+we believe that the security of a system is tightly related to the differences between
+the nominal behavior and the actual (e.g. the implemented) one and only in the
+nominal one (as hypothetical) one should consider agent's knowledge.
 
 \begin{definition}{\bf System State and State-Space --}\label{def:system}
-	The state of a system (or a sub-system) $s$ is identified with an
-	instance of an agent state (see Definition~\ref{def:agent}) such that
+	The state of a system (or a sub-system or an agent) $s$ is defined as the tuple 
 	$s=\langle\rcc(\knowledgeRegion,\beliefRegion),\rcc(\knowledgeRegion,\informationRegion),\rcc(\beliefRegion,\informationRegion)\rangle$,
@@
 	of Knowledge, Beliefs, and Information respectively, and $\rcc$ is a
-	specific relation between the Regions it predicates on.  The
+	specific relation between Regions.  The
 	state-space of a system is then defined by the different $\rcc$
@@
 space) through all the admissible RCC relations. 
-\fixnote{mr}{are the next lines redundant?}A \emph{system state} (as an instance of the system state space)
+A \emph{system state} (as an instance of the system state space)
 is a specific configuration of
@@
 	of predicates known by an agent $\knowledge{a}=\bigcup_\Phi \knows{a}{\varphi}$ 
-	(where $\Phi$ is the collection of all the propositions known by $a$).
+	(for all $\varphi\in\Phi$ where $\Phi$ is the collection of all the propositions known by $a$).
 	Given a proposition $P$, we extend the semantics of the Causality structure with:
@@
 considering abstract and general concepts such as knowledge, belief, or
-information may result to be too abstract for the objectives of the overall engineering
+information may result to be too abstract 
+for the objectives of the overall engineering
 process (e.g. verification and validation).
@@
 				subsystem
-			\item Other important aspects of the system that we
+			\item Other aspects of the system, related to the two previous points, that we
 				describe afterwards in this paper, such as
@@
 \begin{definition}{\bf Assertion -- }\label{def:assertion}
-	An assertion (\fix{mr}{subtype of Information}) of  is an intended transfer of beliefs between
+	An assertion is an intended transfer of beliefs between
 	two agents $a$ and $b$ such that
@@
 \begin{definition}{\bf Behavior -- }\label{def:behavior}
-	The behavior (\fix{mr}{subtype of Beliefs}) of an agent is defined as the transformation process (e.g. defined as a
+	The behavior of an agent is defined as the transformation process (e.g. defined as a
 	protocol, or a functional architecture) that determines the
@@
 	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation20)$] $\world\models\bigcup_{\Phi}\believe{s}{\varphi}$ and $\world\models\behave{a}{[\bigcup_{s\in S}\believe{s}{P_{s}}]}$ iff
-			$\world\models\belief{a}{P'}$,
-
-		where $a$ is an agent, $S$ is a set of agents such that $a\neg\in S$, and the calculation of $P'$ is deteministic.
+		\item[$(\interpretation20)$] $\world\models\bigcup\behaviorRegion(\beliefRegion_I,\beliefRegion_O)$ and 
+			$\world\models\beliefRegion_I\wedge\beliefRegion_O$,
+		where $a$ is an agent, $\beliefRegion_I$ and $\beliefRegion_O$ are input and output-beliefs respectively.
 	\end{enumerate}
@@
 \begin{definition}{\bf Fact -- }\label{def:fact}
-	Facts are a \fix{mr}{sub-region/subtype} of Knowledge $\factRegion = \bigcup_\Phi \knows{a}{\varphi}$
+	Facts are defined as $\factRegion = \bigcup_\Phi \knows{a}{\varphi}$
 	that predicate in a factive way, e.g.
@@
 
-\begin{definition}{\bf Operational System State --}\label{def:system}
-	An operational system (or a sub-system) is represented as an agent (see Definition~\ref{def:agent}) and then defined by the resulting state space,
-	such that
+\begin{definition}{\bf Operational System State --}\label{def:opsystem}
+	An operational system (or a sub-system) state is defined as the tuple
 	$s=\langle\rcc(\factRegion,\behaviorRegion),\rcc(\factRegion,\assertionRegion),\rcc(\behaviorRegion,\assertionRegion)\rangle$,
@@
 representation of the system. In Section~\ref{sec:engineering} we
-detail this numeric evaluation to faitfully represent the
-number of insecurity configurations.
+detail this numeric evaluation to faithfully represent the
+number of insecurity configurations, while in this section we consider
+the general abstract definition given in Defintion~\ref{def:opsystem}.
 
@@
 us to better qualify an Asset, we consider the fact of ``being an asset'' as
-a predicate \fix{mr}{(or a true function)} of a component of a system.
+a predicate of a component of a system.
 
@@
 
-\subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}
+\subsection{Qualitative Evaluation of Agent Space in $\abftheory$}\label{sec:agentspace}
 While a quantitative analysis reveals how many possible configurations of an
-agent (i.e. a system) exist w.r.t. the $\abf$ theory (i.e. $54/125$ in RCC5), a
+agent (i.e. a system) exist w.r.t. the $\abftheory$ theory (i.e. $54/125$ in RCC5), a
 qualitative analysis of the different configurations describes
-the configurations allowed by the $\abf$ theory, and how those configurations can be
+the configurations allowed by the $\abftheory$ theory, and how those configurations can be
 categorized.  In Table~\ref{tab:5com} we provide the generic composition table
@@
 
-A detailed qualitative categorization of agents over a theory similar to the
-$\abf$ one, is presented in this article and described in
-\autocite{Santaca2016abf}. The differences in the formalization and of the
-application requires another in-depth analysis. As depicted in
+As depicted in
 Figure~\ref{fig:soundness}, the relation between Facts, and Assertions and Behavior
@@
 For the sake of simplicity, soundness is opposed to non-soundness in the following, however,
-with the RCC we consider different ``degrees'' of soundness. For example,
+with the RCC one should consider different ``degrees'' of non-soundness. For example,
 in RCC5, if we consider EQ between two Regions as representing soundness, DR over the same Regions
-represents non-soundness while PP, PO, PPi represents the different degrees of soundness.
+represents non-soundness; while PP, PO, PPi represents the different degrees of non-soundness.
+A similar argument can be done for completeness.
 
@@
 		handled in the functional specification (behavior) the
-		specification is sound.
+		specification is complete.
 
@@
 Assertions and Behavior with Facts determines the quality with respect to the
-nominal (specified) system.  Given that Facts defines what is true in the
-system (e.g. the level of water in a tank may be considered as a Fact, while
-the reading of a sensor as a Belief), the relation of Assertions an Facts
+nominal (specified) system.  Given that Facts defines what must be true in the
+system, the relation of Assertions an Facts
 determines the degree of quality between the information circulating in a
@@
 circulating false information discerns a dishonest and an incompetent agent,
-however, we consider honesty related to truth rather than related to what is
-true\footnote{``[\ldots] truth exists only in the good man, but the true in the bad
-man as well; for it is possible for the bad man to utter something true.''
-Sextus Empiricus, Outlines of Pyrrhonism, II-83\autocite{Empiricus1990Pyrrhonism}}.  This implies that the
-relation between Beliefs and Facts determines the competence (on the subjected
+however, we consider honesty related to sharing truths\footnote{``[\ldots] truth
+exists only in the good man, but the true in the bad man as well; for it is
+possible for the bad man to utter something true.''
+Sextus Empiricus, Outlines of Pyrrhonism, II-83\autocite{Empiricus1990Pyrrhonism}}.  The
+relation between Beliefs and Facts determines the competence (on the subjects
 defined by the Facts) of an agent (i.e. the more competent an agent is, the
@@
 
-\subsubsection{$\abf$ Security Enumeration (SE)} The following security requirement for a CPS specification can be summarized:
+\subsubsection{$\abftheory$ Security Enumeration (SE)} The following security requirement for a CPS specification can be summarized:
 \begin{enumerate}
-	\item[SE-$1$] Proper interaction between multiple correctly-behaving agents (in
+	\item[SE-$1$] Proper interaction between correctly-behaving agents (in
 		contrast with the ``Improper Interaction Between Multiple
 		Correctly-Behaving Entities'' defined by the CWE--$435$ as one of the top ``view'' of the ``research concepts'' in
-		\autocite{MITRE2020CWEresearch}) is defined as $\eq{\assertionRegion_a}{\behaviorRegion_a}$ for an agent $a$
+		\autocite{MITRE2020CWEresearch}) is defined as $\eq{\assertionRegion_a}{\behaviorRegion_a}$ for an agent $a$ while can be detailed as follows when multiple agents are considered.
 	\begin{enumerate}
@@
 		in\autocite{MITRE2020CWEresearch}) is defined as
-		$\eq{\beliefRegion}{\factRegion}$.
+		$\eq{\behaviorRegion}{\factRegion}$.
 \end{enumerate}
 
-We are now in the position to define what a secure system (with respect to the $\abf$-theory) is
+We are now in the position to define what a secure system is (with respect to the $\abftheory$-theory) 
 and, based on that definition, what the security risk is and how to quantify it in a risk matrix.
@@
 The whole space of potential designs of a specification with respect to the
-$\abf$-theory.
+$\abftheory$-theory.
 \end{definition}
-The definition of Risk leads to the risk matrix in Figure~\autoref{fig:quantitative}, defined as follows.
+The definition of Risk leads to the risk matrix in Figure~\ref{fig:quantitative}, defined as follows.
 \begin{definition}{\bf Risk Matrix --}

### 2020-04-07 engineering.tex
partial review of engineering section

@@
 calculated based on the $\abftheory$-theory. An iterative process between 
-the user (e.g. the engineer) and the \secramod module of the \abftool allows the user to move to the design phase
+the user (e.g. the engineer) and the \secramod{} module of the \abftool{} allows the user to move to the design phase
 when the risk level is considered acceptable. The design phase starts by automatically mapping
-the Weaknesses into Vulnerabilities which are, in turn, fed into the \designverifmod
+the Weaknesses into Vulnerabilities which are, in turn, fed into the \designverifmod{}
 module along with the final specification. The Module returns a list of Vulnerabilities and
 potential Mitigations. The user then procees in implementing the CPS and the SW/HW choices
-can be tested by the \atgmod.
+can be tested by the \atgmod. \fix{mr}{While we have described the overall view, from
+specificaiton to implementation, in this text we only consider the process from specificaiton to design.}
 
@@
 		\item a tank that is filled with raw water coming from the inlet pipe 
-		\item a motorized valve (a valve with an actuator) that regulates that
+		\item a motorized valve (a valve with an actuator) such that:
 			\begin{itemize}
@@
 		\item a sensor that reads the meters of raw water in the tank
-		\item a PLC controller, connected to the sensor and the actuator, that receives the readings from the sensor and regulates the quantity of water in the tank by communicating a change in the state of the actuator based on the following logi
+		\item a PLC controller, connected to the sensor and the
+			actuator, that receives the readings from the sensor
+			and regulates the quantity of water in the tank by
+			communicating a change in the state of the actuator
+			based on the following logic
 			\begin{itemize}
-				\item if the readings from the sensor reports a water level in the tank above (or equal) to 10m, the actuator shall close the valve
-				\item if th readings are below 10m the actuator shall open the valve.
+				\item if the readings from the sensor reports a
+					water level in the tank above (or
+					equal) to 10m, the actuator shall close
+					the valve
+				\item if the readings are below 10m the actuator
+					shall open the valve.
 			\end{itemize}
@@
 The first step of the SPDL is the definition of the requirements and then, for
-the purpose of this paper of the (i) security requirements, (ii) architectural
+the purpose of this paper, of (i) security requirements, (ii) architectural
 (physical and functional) requirements.  Our focus in mainly on security
@@
 	\item the sensor communicates the readings to the controller ($\text{sensor}\rightarrow \text{controller}$)
-	\item the controller calculates the valve state and communicates it to the actuator ($\text{controller}\rightarrow \text{actuator}$)
-	\item the actuator translates (digital to analog) the states received and communicates\footnote{we do not distinguish, for the sake of simplicity between different types of channels and assume that communicating to the valve produces the expected change of physical state} it to the valve
+	\item the controller calculates the state of the valve and communicates it to the actuator ($\text{controller}\rightarrow \text{actuator}$)
+	\item the actuator translates (digital to analog) the state received and communicates\footnote{we do not distinguish, for the sake of simplicity between different types of channels and assume that communicating to the valve produces the expected change of physical state} it to the valve
 \end{enumerate}
@@
 
-\Paragraph{Input and Output Ports}
+\subsubsection{Input and Output Ports}
 Since the $\abftheory$-theory is a theory of agents, we consider ports as agents that
@@
 assertions received or sent and the belief, i.e. $\Rcc{\assertionRegion_s}{\beliefRegion_a}$ for the input-port
-or $\Rcc{\assertionRegion_r}{\beliefRegion_a}$.
+or $\Rcc{\assertionRegion_r}{\beliefRegion_a}$. We only define secure input port but the
+definition of a secure output port is symmetrical.
 
 \begin{definition}{\bf Secure Input Port --}\label{def:secport}
-	A secure input port allows information as incoming assertions to flow
+	A secure input port ($s\secureport a$) allows information as incoming assertions to flow
 	from a sender $s$ to the behavior of the recipient $a$ (agent).  If $p_I$ is a secure input port,
-	$\world\models\varphi$, and $\varphi\in\assertionRegion_s$ where $s$
-	has the only input port $p_I$, $\world'\models\beliefRegion_a$ and $\varphi\in\beliefRegion_a$.
+	$\world\models\varphi$, $\varphi\in\rassert{s}{a}$, and $a$
+	has the only input port $p_I$, $\world'\models\beliefRegion_a$ and $\varphi\in\beliefRegion_a$ for any
+	$\world\modalrelation\world'$.
 \end{definition}
@@
 
-\Paragraph{Communication Channels}
+\subsubsection{Communication Channels}
 We start by considering the difference between a (communication)
@@
 \begin{enumerate}[start=5, label={W\arabic*)}]
-	\item \emph{secure input port} and \emph{output drop port} ($s\sdcom r$), 
-	\item \emph{secure input port} and \emph{output insertion port} ($s\sicom r$), 
-	\item \emph{secure input port} and \emph{output injection port} ($s\sjcom r$), 
-	\item \emph{secure input port} and \emph{output selective drop port} ($s\ssdcom r$), 
-	\item \emph{secure input port} and \emph{output selective insertion port} ($s\ssicom r$), 
-	\item \emph{secure input port} and \emph{output selective injection port} ($s\ssjcom r$),
+	\item \emph{secure output port} and \emph{input drop port} ($s\sdcom r$), 
+	\item \emph{secure output port} and \emph{input insertion port} ($s\sicom r$), 
+	\item \emph{secure output port} and \emph{input injection port} ($s\sjcom r$), 
+	\item \emph{secure output port} and \emph{input selective drop port} ($s\ssdcom r$), 
+	\item \emph{secure output port} and \emph{input selective insertion port} ($s\ssicom r$), 
+	\item \emph{secure output port} and \emph{input selective injection port} ($s\ssjcom r$),
 
-	\item \emph{input drop port} and \emph{output drop port} ($s\ddcom r$)
-	\item \emph{input drop port} and \emph{output insertion port} ($s\dicom r$)
-	\item \emph{input drop port} and \emph{output secure port} ($s\dscom r$)
+	\item \emph{output drop port} and \emph{input drop port} ($s\ddcom r$)
+	\item \emph{output drop port} and \emph{input insertion port} ($s\dicom r$)
+	\item \emph{output drop port} and \emph{input secure port} ($s\dscom r$)
 
-	\item \emph{input insert port} and \emph{output drop port} ($s\idcom r$)
-	\item \emph{input insert port} and \emph{output insert port} ($s\iicom r$)
-	\item \emph{input insert port} and \emph{output injection port} ($s\ijcom r$)
-	\item \emph{input insert port} and \emph{output selective drop port} ($s\isdcom r$)
-	\item \emph{input insert port} and \emph{output selective insertion port} ($s\isicom r$)
-	\item \emph{input insert port} and \emph{output selective injection port} ($s\isjcom r$)
-	\item \emph{input insert port} and \emph{output secure port} ($s\iscom r$)
+	\item \emph{output insert port} and \emph{input drop port} ($s\idcom r$)
+	\item \emph{output insert port} and \emph{input insert port} ($s\iicom r$)
+	\item \emph{output insert port} and \emph{input injection port} ($s\ijcom r$)
+	\item \emph{output insert port} and \emph{input selective drop port} ($s\isdcom r$)
+	\item \emph{output insert port} and \emph{input selective insertion port} ($s\isicom r$)
+	\item \emph{output insert port} and \emph{input selective injection port} ($s\isjcom r$)
+	\item \emph{output insert port} and \emph{input secure port} ($s\iscom r$)
 
-	\item \emph{input injection port} and \emph{output drop port} ($s\jdcom r$)
-	\item \emph{input injection port} and \emph{output insertion port} ($s\jicom r$)
-	\item \emph{input injection port} and \emph{output injection port} ($s\jjcom r$)
-	\item \emph{input injection port} and \emph{output selective drop port} ($s\jsdcom r$)
-	\item \emph{input injection port} and \emph{output selective insertion port} ($s\jsicom r$)
-	\item \emph{input injection port} and \emph{output selective injection port} ($s\jsjcom r$)
-	\item \emph{input injection port} and \emph{output secure port} ($s\jscom r$)
+	\item \emph{output injection port} and \emph{input drop port} ($s\jdcom r$)
+	\item \emph{output injection port} and \emph{input insertion port} ($s\jicom r$)
+	\item \emph{output injection port} and \emph{input injection port} ($s\jjcom r$)
+	\item \emph{output injection port} and \emph{input selective drop port} ($s\jsdcom r$)
+	\item \emph{output injection port} and \emph{input selective insertion port} ($s\jsicom r$)
+	\item \emph{output injection port} and \emph{input selective injection port} ($s\jsjcom r$)
+	\item \emph{output injection port} and \emph{input secure port} ($s\jscom r$)
 
-	\item \emph{input selective drop port} and \emph{output drop port} ($s\sddcom r$)
-	\item \emph{input selective drop port} and \emph{output insertion port} ($s\sdicom r$)
-	\item \emph{input selective drop port} and \emph{output injection port} ($s\sdjcom r$)
-	\item \emph{input selective drop port} and \emph{output selective drop port} ($s\sdsdcom r$)
-	\item \emph{input selective drop port} and \emph{output selective insertion port} ($s\sdsicom r$)
-	\item \emph{input selective drop port} and \emph{output selective injection port} ($s\sdsjcom r$)
-	\item \emph{input selective drop port} and \emph{output secure port} ($s\sdscom r$)
+	\item \emph{output selective drop port} and \emph{input drop port} ($s\sddcom r$)
+	\item \emph{output selective drop port} and \emph{input insertion port} ($s\sdicom r$)
+	\item \emph{output selective drop port} and \emph{input injection port} ($s\sdjcom r$)
+	\item \emph{output selective drop port} and \emph{input selective drop port} ($s\sdsdcom r$)
+	\item \emph{output selective drop port} and \emph{input selective insertion port} ($s\sdsicom r$)
+	\item \emph{output selective drop port} and \emph{input selective injection port} ($s\sdsjcom r$)
+	\item \emph{output selective drop port} and \emph{input secure port} ($s\sdscom r$)
 
-	\item \emph{input selective insertion port} and \emph{output drop port} ($s\sidcom r$)
-	\item \emph{input selective insertion port} and \emph{output insertion port} ($s\siicom r$)
-	\item \emph{input selective insertion port} and \emph{output injection port} ($s\sijcom r$)
-	\item \emph{input selective insertion port} and \emph{output selective drop port} ($s\sisdcom r$)
-	\item \emph{input selective insertion port} and \emph{output selective insertion port} ($s\sisicom r$)
-	\item \emph{input selective insertion port} and \emph{output selective injection port} ($s\sisjcom r$)
-	\item \emph{input selective insertion port} and \emph{output secure port} ($s\siscom r$)
+	\item \emph{output selective insertion port} and \emph{input drop port} ($s\sidcom r$)
+	\item \emph{output selective insertion port} and \emph{input insertion port} ($s\siicom r$)
+	\item \emph{output selective insertion port} and \emph{input injection port} ($s\sijcom r$)
+	\item \emph{output selective insertion port} and \emph{input selective drop port} ($s\sisdcom r$)
+	\item \emph{output selective insertion port} and \emph{input selective insertion port} ($s\sisicom r$)
+	\item \emph{output selective insertion port} and \emph{input selective injection port} ($s\sisjcom r$)
+	\item \emph{output selective insertion port} and \emph{input secure port} ($s\siscom r$)
 
-	\item \emph{input selective injection port} and \emph{output drop port} ($s\sjdcom r$)
-	\item \emph{input selective injection port} and \emph{output insertion port} ($s\sjicom r$)
-	\item \emph{input selective injection port} and \emph{output injection port} ($s\sjjcom r$)
-	\item \emph{input selective injection port} and \emph{output selective drop port} ($s\sjsdcom r$)
-	\item \emph{input selective injection port} and \emph{output selective insertion port} ($s\sjsicom r$)
-	\item \emph{input selective injection port} and \emph{output selective injection port} ($s\sjsjcom r$)
-	\item \emph{input selective injection port} and \emph{output secure port}  ($s\sjcom r$)
+	\item \emph{output selective injection port} and \emph{input drop port} ($s\sjdcom r$)
+	\item \emph{output selective injection port} and \emph{input insertion port} ($s\sjicom r$)
+	\item \emph{output selective injection port} and \emph{input injection port} ($s\sjjcom r$)
+	\item \emph{output selective injection port} and \emph{input selective drop port} ($s\sjsdcom r$)
+	\item \emph{output selective injection port} and \emph{input selective insertion port} ($s\sjsicom r$)
+	\item \emph{output selective injection port} and \emph{input selective injection port} ($s\sjsjcom r$)
+	\item \emph{output selective injection port} and \emph{input secure port}  ($s\sjcom r$)
 \end{enumerate}
@@
 
-\Paragraph{Functional Block}
+\subsubsection{Functional Block}
 A Functional Architecture, as already described beforehand, takes Information as input-Beliefs
@@
 
-\Paragraph{Facts}
+\Paragraph{Requirements as Facts}
 During the specification phase, for any agent, channel, port, functional block
@@
 input-Beliefs formulas and the symmetric formulas in the Fact region
-(equivalently for output-Beliefs). Therefore, the quality of the functional
-architecture is not a purely architectural property but relies on the logic
-implemented in the functional blocks (which we abstract away at this stage of
-the engineering model).
+(equivalently for output-Beliefs).
+
 \begin{definition}{\bf Secure Functional Block and Architecture --}\label{def:secfun}
@@
 the corresponding Facts (i.e. what really happens). 
-\begin{enumerate}
-	\item $\dr{\beliefRegion_I}{\factRegion_I}$ and
-		$\dr{\beliefRegion_O}{\factRegion_O}$, the input beliefs are
-		misinterpreted and this result in the complete alteration of the
-		output beliefs
-	\item \fix{mr}{TBD}
+\begin{itemize}
+	\item $\eq{\beliefRegion_I}{\factRegion_I}$ and then only specified
+		inputs can be fed to the functional block but:
+	\begin{enumerate}[start=49, label={W\arabic*)}]
+		\item $\dr{\beliefRegion_O}{\factRegion_O}$, not all the specified
+			outputs are produced
+		\item $\po{\beliefRegion_O}{\factRegion_O}$ , only some but not all of the specified
+			outputs are produced along with some non-specified output are produced
+		\item $\pp{\beliefRegion_O}{\factRegion_O}$ , not all the specified outputs are produced 
+		\item $\ppi{\beliefRegion_O}{\factRegion_O}$ , not all the output have been specified
 	\end{enumerate}
+	\item $\po{\beliefRegion_I}{\factRegion_I}$, not all the specified inputs can be fed into the functional block
+	\begin{enumerate}[start=53, label={W\arabic*)}]
+		\item $\dr{\beliefRegion_O}{\factRegion_O}$
+		\item $\po{\beliefRegion_O}{\factRegion_O}$
+		\item $\pp{\beliefRegion_O}{\factRegion_O}$
+		\item $\ppi{\beliefRegion_O}{\factRegion_O}$
+		\item $\eq{\beliefRegion_O}{\factRegion_O}$ this case implies
+			that somehow there is no input that generates an
+			non-specified output but this is not because the inputs
+			are secure 
+	\end{enumerate}
+	\item $\pp{\beliefRegion_I}{\factRegion_I}$, not all the specified
+		inputs can be fed into the block
+	\begin{enumerate}[start=58, label={W\arabic*)}]
+		\item $\dr{\beliefRegion_O}{\factRegion_O}$
+		\item $\po{\beliefRegion_O}{\factRegion_O}$
+		\item $\pp{\beliefRegion_O}{\factRegion_O}$
+		\item $\ppi{\beliefRegion_O}{\factRegion_O}$
+		\item $\eq{\beliefRegion_O}{\factRegion_O}$
+	\end{enumerate}
+	\item $\ppi{\beliefRegion_I}{\factRegion_I}$, not all the inputs that
+		can be fed into the functional block follows the specification
+	\begin{enumerate}[start=63, label={W\arabic*)}]
+		\item $\dr{\beliefRegion_O}{\factRegion_O}$
+		\item $\po{\beliefRegion_O}{\factRegion_O}$
+		\item $\pp{\beliefRegion_O}{\factRegion_O}$
+		\item $\ppi{\beliefRegion_O}{\factRegion_O}$
+		\item $\eq{\beliefRegion_O}{\factRegion_O}$
+	\end{enumerate}
+	\item $\dr{\beliefRegion_I}{\factRegion_I}$, no specified inputs can be
+		fed into the functional block
+	\begin{enumerate}[start=68, label={W\arabic*)}]
+		\item $\dr{\beliefRegion_O}{\factRegion_O}$
+		\item $\po{\beliefRegion_O}{\factRegion_O}$
+		\item $\pp{\beliefRegion_O}{\factRegion_O}$
+		\item $\ppi{\beliefRegion_O}{\factRegion_O}$
+		\item $\eq{\beliefRegion_O}{\factRegion_O}$
+	\end{enumerate}
+\end{itemize}
 
-\Paragraph{Security Requirements} 
+\subsubsection{Security Requirements} 
 While standards (such
@@
 
-\subsubsection{Availability and Integrity (Architecture Security)}
+\Paragraph{Availability and Integrity (Architecture Security)}
 \begin{figure}[t]
@@
 
-\subsubsection{Secrecy, Authenticity, and Key Agreement (Cryptographic Security)}\label{sec:encryption}
+\Paragraph{Secrecy, Authenticity, and Key Agreement (Cryptographic Security)}\label{sec:encryption}
 As Maurer puts it (\autocite{Maurer1999information}), there exist two types of

### 2020-04-08 engineering.tex
full review of engineering section

@@
 \end{enumerate}
-Companies who develop CPS (e.g. aerospace systems or elevator systems), must adhere
+Companies which develop CPS (e.g. aerospace systems or elevator systems), must adhere
 to strict regulations and quality assurances processes. Introducing tools and procedures 
@@
 
+As depicted in Figure~\ref{fig:atom}, so far, we have considered Information
+generated by a Port $P_I$ and then sent through a channel $C$ to another
+(recipient) port $P_O$. In this scenario where Ports and then Channels are
+atomic (otherwise raising infinite regress), we can only  consider the
+relations between Ports and Channel; considering both input-port to channel and
+channel to output-port. In fact, the Weaknesses of a channel are defined in terms
+of the Weaknesses of Ports. Our intention is to introduce now the concept of
+functional architecture and functional blocks composing the functional architecture.
+Io order to define a functional block without encountering an infinite definition,
+we must reach the same conclusions as for the Channel. Therefore, describing
+the information as flowing over a channel or in a functional block is purely
+syntactic sugar.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.7\textwidth]{engineering_relations.pdf}
+	\caption{Relations between Ports and Channel}
+	\label{fig:atom}
+\end{figure}
+
 \subsubsection{Functional Block}
@@
 \end{definition}
-The quality of a functional block or architecture cannot be expressed as a relation 
-between input and output belief since there is a transformation process in the functional
-block which will be defined during the design and implementation phase, but not
-necessarily known in details in the specification phase. In order to
-define the quality of functional block we first need to introduce the engineering concept of
-Facts.
+The quality of a functional block or architecture can be expressed as a
+relation between input and output belief, similarly to the quality of a channel
+but inverting the flow (input$\rightarrow$output instead of
+output$\rightarrow$input).  One may, of course, argue that since there is a
+transformation process in the functional block the quality of that
+transformation process is what the theory should consider. This can
+be argued not just for functional blocks (Behaviors) but also for Channels
+(Assertions). As we detail afterwards, the relations considered so
+far revolves around the concept of integrity or availability, while the reminder of this section 
+will be focused on a broader analysis of security properties and requirements
+which will also consider other security properties such as availability and
+confidentiality.  In order to define the quality of the internals of a functional block or channel, we first
+need to relate the concept of Facts to engineering requirements.
 
@@
 During the specification phase, for any agent, channel, port, functional block
-and architecture, there exist a Fact predicating over them.
+and architecture, there may exist a factorial requirement (Fact) predicating over them.
 In other words, any requirement is defined as a Fact since they must be true in
@@
 
+\Paragraph{Methodology}
+We have now mapped the building blocks of our theory, namely Information,
+Belief, Assertion, Behavior, and Fact. One may notice that the concept of
+Knowledge hasn't been discussed in this section so far. Our argument is that
+knowledge, as an epistemological concept cannot be defined due to several
+implication that this would have (see, for example,
+\autocite{Empiricus1990Pyrrhonism}). On the other hand, the factorial way in
+which requirements are imposed may be true only in the specification, while not
+necessarily true in the final implementation. While Knowledge may not be
+apprehensible in general, we can define tests to see if Facts holds in a
+design or implementation. What is of our interest is to guarantee security in a
+system so we now correlate the potential behaviors with the standard (de-facto)
+understanding of security; so that we can identify test that, at the end of the
+engineering process, may not give Knowledge but, at last, tested Fact.
+
+\subsubsection{Security Requirements} 
+While standards such as IEC 62443-1-3 (the Industrial communication networks -
+Network and system security -- Part 3-3: System security requirements and
+security levels) defines requirements as ``confidentiality of information in
+transit/at-rest'', we believe that, at specification level, a mapping
+between the Infosec security concepts of the well-known CIA (Confidentiality,
+Integrity, Availability) triad to the system components can be used to specify
+a list of security requirements.
+
+Infosec, or Information Security, defines the security risks related to
+information as the CIA triad, which has been often criticized
+(as we show afterwards) for being too general or non-adequate (e.g. by
+adding authenticity which is often used as a building block for confidentiality
+and integrity) to be effectively applied to the engineering of systems. Due to
+this, many researchers and organization tried to improve the CIA triad.  The
+evolutions/extensions of the CIA triad has been documented, e.g., in
+\autocite{Samonas2014cia} and are summarized in Table~\ref{tab:ciaevolution}.
+
+\begin{table}[h]
+\centering
+%\setlength{\tabcolsep}{3.5pt}
+%\renewcommand{\arraystretch}{1}
+\small
+\begin{tabular}{rll} 
+	{\bf Year} & {\bf Definition} & {\bf Legend}\\
+	\hline
+	1970s & infosec = CIA & Confidentiality, Integrity, Availability\\
+	1980s & infosec += (Au, nR) & Authenticity and non-Repudiation\\
+	1990s & infosec += CSpec & Correctness in Specification\\
+	2000s & infosec += RITE & Responsibility, Integrity of people, Trust, Ethicality
+\end{tabular}
+\caption{Chronological progression of the CIA triad~\label{tab:ciaevolution}}
+\end{table}
+
+A representative example is the effort made by the OECD (Organisation for Economic
+Co-operation and Development) in \autocite{OECD2013guidelines} to define
+security guidelines for information system (1992) based on new principles such
+as ``awareness, ethics, risk assessment'' and maintain those improving the document with
+subsequent revisions 
+(e.g.  following the multistakeholder expert consultation in 2013).
+Other similar efforts focused on defining entirely new principles or extending
+the CIA triad, such as \autocite{NISTSP800-160}. 
+%What is of interest for our
+%argument is that, to the best of our knowledge, all of the approaches that aim
+%at improving the CIA triad are besed on motivations related to empirical evidences. On the contrary,
+%We now define (in Section~\ref{sec:properties}) the CIA triad for an agent
+%defined in the $\abftheory$-theory and we test if (i) other properties are
+%allowed in the $\abftheory$-theory, and (ii) if and how the CIA triad can be
+%detailed in the $\abftheory$-theory.
+In \autocite{Anderson1972report,Samonas2014cia}, CIA are defined as follows.
+\begin{itemize}
+	\item \emph{Unauthorized information release}: an unauthorized person is able
+		to read and take  advantage  of  information  stored  in  the
+		computer.  This  category of concern  sometimes  extends  to
+		``traffic  analysis,''  in  which  the intruder  only observes
+		the  patterns  of  information  use.  From  those patterns,
+		the  intruder can   infer   some   information   content.
+		This category   also   includes   the unauthorized use of a
+		proprietary program.  (Confidentiality or Secrecy) 
+	\item \emph{Unauthorized  information  modification}:  an  unauthorized person
+		is  able  to make changes in stored information -- a form  of
+		sabotage.  It should be noted that in the case of this kind of
+		violation, the intruder does not necessarily see the
+		information he has changed.  (Integrity)
+	\item \emph{Unauthorized  denial  of  use}:  an  intruder  can  prevent  an
+		authorized  user  from referring to, or from modifying
+		information, even though the intruder may not be able to refer
+		to, neither modify the information themselves. (Availability)
+\end{itemize}
+
+We now want to generalize the CIA triad to agents and systems in the
+$\abftheory$-theory (and not just Information). Given that we are currently
+dealing with the step that moves from specification to design, we consider
+architectural properties. For the sake of simplicity, we revert the order and
+start from availability.
+
+\Paragraph{Availability and Integrity (Architecture Security)}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\textwidth]{client-server.pdf}
+	\caption{Client-Server paradigm in $\abftheory$-theory}
+	\label{fig:client-server}
+\end{figure}
+Availability, again, is the denial of use of a resource. In the $\abftheory$-theory, 
+in order to prevent a resource to be available, there
+exists only (i.e. inclusively) the following cases
+that we illustrate using a client-server paradigm (depicted in Figure~\ref{fig:client-server}).
+\begin{enumerate}
+	\item\label{case:srv1} The resource on the server cannot be found, so that it is
+		possible for the client to contact the server and for the server
+		to send a response back, but the output-port does not receive the
+		necessary information. This may be due to a software error such
+		that the functional architecture doesn't output a response, or
+		a problem in the flow of information (i.e. the functional
+		architecture doesn't receive the necessary information from
+		other resources), or an hardware failure \&c. This can be formalized
+		as an issue to be found in the relation between the beliefs produced by the functional
+		architecture and the ones received by the output port, i.e.
+		$\rcc(\behaviorRegion_s,\{\text{res}_0\})$ or 
+		$\rcc(\beliefRegion_s,\{\text{res}_0\})$ 
+	\item\label{case:srv2} The server hosting the resource is down, and then the server
+		cannot output responses (i.e. insecure output port due to drop-related issue) or receive
+		requests (insecure input port due to drop-related issue) or both.
+	\item\label{case:srv3} The client cannot properly compute request (symmetric to Case~\ref{case:srv1})
+	\item The client is down, symmetric to Case~\ref{case:srv2}
+	\item The channel is down due to a drop related issue in one of the two channels 
+\end{enumerate}
+%All the cases but Case~\ref{case:srv1} and Case~\ref{case:srv3}, tell us how to map the
+%property of availability over the weaknesses of ports and communications.  The
+%Cases~\ref{case:srv1} and~\ref{case:srv3}, are exceptions because the formulas $res$ and $req$ 
+%detail the specification with a singleton instead of a region of formulas (or, better, $|res|$ is equal to the one of the
+%base).
+%In other words, a channel, a port, an entity, or a functional block are
+%Fact (stated in a factorial way) and not Beliefs; what they produce or 
+%their non-specified internals are Beliefs.  In fact, we never assume an
+%entity not to be in the design implementing the specification.  Similarly, when a
+%functional block produces some information it produces a Belief, unless this Belief if detailed in the specification.
+%For example, the server produces the Belief $res$ but the Belief given that can only be in $EQ$
+%relation with the Fact $res$ (in contrast other Belief may not); therefore, $res$ is not really a belief but a Fact. 
+%$res$ as a Fact is something that
+%exists in the specification, while the Belief of $res$ in one of the possible
+%designs may not exist. 
+We shall, then, distinguish between two cases, as we mentioned beforehand:
+agent (or system) and Information (which we call Infosec) availability. 
+\begin{definition}{\bf Agent-Availability --}\label{def:agent-a}
+when the property of availability is expressed over an agent or a system, the
+	list of possible designs do not include those in which there exist:
+	drop ports, or drop channels/functional-blocks.
+\end{definition}
+\begin{definition}{\bf Infosec-Availability --}\label{def:infosec-c}
+when the property of availability is expressed
+	over a formula, the list of possible designs that violates the
+	availability property is reduced to the part of the system that
+	does not influence that specific formula.
+\end{definition}
+
+In a similar way we can define integrity since this requirement may 
+be expressed over an agent/system or a specific formula (i.e. element of a Region).
+
+\begin{definition}{\bf Agent-Integrity --}\label{def:agent-a}
+when the property of integrity is expressed over
+	an agent or a system, the list of possible designs 
+	do not include those in which there exist: insert or inject ports, 
+	or inert or inject channels/functional-blocks. 
+\end{definition}
+\begin{definition}{\bf Infosec-Integrity --}\label{def:infosec-c}
+when the property of integrity is expressed
+	over a formula, the list of possible designs that violates the
+	integrity property is reduced to the part of the system that
+	influences that specific variable.
+\end{definition}
+
+While availability and integrity are architectural property, meaning that they
+can be expressed over the behavior of a port or a channel (as constituent of
+the physical or functional architecture), confidentiality is not.
+Confidentiality states that an information shall not be understood and not that
+the information shall not be retrieved. So, if a port leaks an entire message,
+there is no confidentiality issue in general, unless that message was not
+encrypted at all, in which case the confidentiality leakage is straightforward. 
+On the other hand, if the
+information is not provided or altered there is an availability or integrity
+issue. A similar argument can be proposed for integrity sub-properties such as
+authenticity or non-repudiation. No wonder why it is often stated that
+availability and integrity are the first properties to be considered in CPS
+while confidentiality takes the first place when dealing with security
+protocols. In fact, integrity is often assumed by pattern-matching (e.g. in
+AVANTSSAR \autocite{avantssar}) and availability is hardly ever considered in
+protocol verification. The question is, can we express properties such as
+confidentiality, authentication, or non-repudiation as architectural properties
+in $\abftheory$-theory? Confidentiality is based on the concept 
+of secrecy so we shall first start from describing what secrecy is and how it
+can be used to guarantee confidentiality in a system.
+
+\Paragraph{Secrecy, Authenticity, and Key Agreement (Cryptographic Security)}\label{sec:encryption}
+As Maurer puts it (\autocite{Maurer1999information}), there exist two types of
+cryptographic security (i.e. scientific theories aiming at guaranteeing
+security through the use of cryptographic functions): ``The security  of a
+cryptographic system can rely either on the computational infeasibility of
+breaking it (computational security), or on the theoretical impossibility of
+breaking it, even using in infinite computing power (information-theoretic or
+unconditional security)''. We are not currently interested in detailing the
+difference between the two and we simplify it as follows. A
+computationally-secure cryptographic function (usually called encryption
+function) $E(\varphi)=e$ (where $\varphi$ is a plain-text and $e$ a
+cipher-text) can be considered perfectly secure if a malicious agent that
+obtains the result of the cryptographic function ($e$) has bounded
+computational power, while unconditional secure cryptographic functions are
+considered perfectly secure. In this context, security means that is is
+impossible to calculate $\varphi$ from $e$, therefore impossible to compute
+$D(e)=\varphi$.  As one can evince from \autocite{Shannon1948information}, the
+commonality between the two types of cryptographic functions is that they both
+rely on the assumption about the probabilistic behavior of the universe. They
+do so, for the random (i.e. with high entropy
+\autocite{Shannon1948information}) generation of the so called \emph{key}, a
+piece of information, parametric to the cryptographic function, used to
+guarantee the security. Therefore, we can assume that it is feasible for an
+agent that knows $k$ to decrypt the cipher-text obtaining the clear-text (i.e. $D(k, E(k,\varphi))=\varphi$)
+while it is impossible for any agent that has not the proper key. In this
+case, the information is considered to be \emph{secret or confidential}.
+
+\begin{definition}{\bf Infosec-Secrecy or Confidentiality --}\label{def:confidentiality}
+	When an Assertion has been generated as an output by an
+	agent, port, or functional block such that it is said (in a
+	factorial way) to encrypt the input, that Assertion is considered
+	confidential.
+\end{definition}
+The quality of the confidentiality of information (Assertions or Beliefs depending on 
+the subsystem considered) is inversely proportional to the number of 
+integrity issues of the subsystem.
+We now discuss more in details what secrecy of Information is when 
+dealing with functional blocks (Behavior) and Beliefs but a similar treatment can be done 
+in the case of Channels and Assertions.
+
 The quality of a Functional Block is determined by the relation between the
@@
 
-\begin{definition}{\bf Secure Functional Block and Architecture --}\label{def:secfun}
-	A functional block is secure if the input beliefs and output beliefs
-	are in $EQ$ relation with the corresponding Facts mandated by requirements.
-	A functional architecture composed exclusively by secure functional blocks
-	is considered secure.
+\begin{definition}{\bf Secrecy of Functional Block and Architecture --}\label{def:secfun}
+	Secrecy is preserved in a functional block if the input beliefs and output beliefs
+	are in $EQ$ relation with the corresponding Facts mandated by the confidentiality/secrecy requirements.
+	A functional architecture composed exclusively by confidential/secrecy-preserving functional blocks
+	is considered confidential/secrecy-preserving.
 \end{definition}
 
-Therefore, there exists the following weaknesses in the functional
+\Paragraph{Confidentiality Weaknesses}
+In a functional block, there exists the following weaknesses in the functional
 architecture, where $\beliefRegion_I$ and $\beliefRegion_O$ are input and
 output Beliefs respectively, and $\factRegion_I$ and $\factRegion_O$ are
-the corresponding Facts (i.e. what really happens). 
+the corresponding Facts. 
 \begin{itemize}
@@
 	\begin{enumerate}[start=49, label={W\arabic*)}]
-		\item $\dr{\beliefRegion_O}{\factRegion_O}$, not all the specified
-			outputs are produced
-		\item $\po{\beliefRegion_O}{\factRegion_O}$ , only some but not all of the specified
-			outputs are produced along with some non-specified output are produced
-		\item $\pp{\beliefRegion_O}{\factRegion_O}$ , not all the specified outputs are produced 
-		\item $\ppi{\beliefRegion_O}{\factRegion_O}$ , not all the output have been specified
+		\item\label{w:fb_eqdr} $\dr{\beliefRegion_O}{\factRegion_O}$, not all the specified
+			outputs are produced. This may produce a secrecy issue if, for example,
+			the functional block is responsible for generating an encryption key since
+			if it is not produced the encryption of the Information may not be performed
+		\item\label{w:fb_eqpo} $\po{\beliefRegion_O}{\factRegion_O}$, only some but not all of the specified
+			outputs are produced along with some non-specified output are produced. This
+			strightforwardly generates a secrecy when the functionality is 
+			related to secrecy functional blocks since some outputs doesn't follow
+			the requirements.
+		\item $\pp{\beliefRegion_O}{\factRegion_O}$, not all the specified outputs are produced.
+			This may pose a secrecy issue as in \ref{w:fb_eqdr}
+		\item $\ppi{\beliefRegion_O}{\factRegion_O}$, not all the output have been specified.
+			This may pose a secrecy issue as in \ref{w:fb_eqpo}
 	\end{enumerate}
@@
 
-\subsubsection{Security Requirements} 
-While standards (such
-as IEC62443-1-3\fixnote{mr}{check the correct version}) defines requirements as
-``information in transit/at-rest should be encrypted'' we believe that, at
-specification level, security requirements can be formulated as subcategories of
-the well-known CIA triad (where CIA stands for Confidentiality, Integrity,
-Availability).
-Infosec, or Information Security, defines the security risks related to
-information with the de-facto standard CIA triad, which is often criticized
-\autocite{CIAcriticismCPS} for being too general or non-adequate (e.g. by
-adding authenticity which is often used as a building block for confidentiality
-and integrity) to be effectively applied to the engineering of systems. Due to
-this, many researchers and organization tried to improve the CIA triad.  The
-evolutions/extensions of the CIA triad has been documented, e.g., in
-\autocite{Samonas2014cia} and is summarized in Table~\ref{tab:ciaevolution}.
-
-\begin{table}[h]
-\centering
-%\setlength{\tabcolsep}{3.5pt}
-%\renewcommand{\arraystretch}{1}
-\small
-\begin{tabular}{rll} 
-	{\bf Year} & {\bf Definition} & {\bf Legend}\\
-	\hline
-	1970s & infosec = CIA & Confidentiality, Integrity, Availability\\
-	1980s & infosec += (Au, nR) & Authenticity and non-Repudiation\\
-	1990s & infosec += CSpec & Correctness in Specification\\
-	2000s & infosec += RITE & Responsibility, Integrity of people, Trust, Ethicality
-\end{tabular}
-\caption{Chronological progression of the CIA triad~\label{tab:ciaevolution}}
-\end{table}
-
-An representative example is the effort made by the OECD (Organisation for Economic
-Co-operation and Development) in \autocite{OECD2013guidelines} to define
-security guidelines for information system (1992) based on new principles such
-as ``awareness, ethics, risk assessment'' and maintain those revising the
-document (e.g.  following the multistakeholder expert consultation in 2013).
-Other similar efforts focused on defining entirely new principles or extending
-the CIA triad, such as \autocite{NISTSP800-160}. What is of interest for our
-argument is that, to the best of our knowledge, all of the approaches that aim
-at improving the CIA triad are besed on motivations related to empirical evidences. On the contrary,
-we now define (in Section~\ref{sec:properties}) the CIA triad for an agent
-defined in the $\abftheory$-theory and we test if (i) other properties are
-allowed in the $\abftheory$-theory, and (ii) if and how the CIA triad can be
-detailed in the $\abftheory$-theory.
-
-In \autocite{Anderson1972report,Samonas2014cia}, CIA are defined as follows.
-\begin{itemize}
-	\item Unauthorized information release: an unauthorized person is able
-		to read and take  advantage  of  information  stored  in  the
-		computer.  This  category of concern  sometimes  extends  to
-		``traffic  analysis,''  in  which  the intruder  only observes
-		the  patterns  of  information  use.  From  those patterns,
-		the  intruder can   infer   some   information   content.
-		This category   also   includes   the unauthorized use of a
-		proprietary program.  (Confidentiality or Secrecy) 
-	\item Unauthorized  information  modification:  an  unauthorized person
-		is  able  to make changes in stored information -- a form  of
-		sabotage.  It should be noted that in the case of this kind of
-		violation, the intruder does not necessarily see the
-		information he has changed.  (Integrity)
-	\item Unauthorized  denial  of  use:  an  intruder  can  prevent  an
-		authorized  user  from referring to, or from modifying
-		information, even though the intruder may not be able to refer
-		to, neither modify the information themselves. (Availability)
-\end{itemize}
-
-We now want to generalize the CIA triad to agents and not just 
-information and map it to the $\abftheory$-theory. For the sake of simplicity,
-we revert the order and start from availability.
-
-\Paragraph{Availability and Integrity (Architecture Security)}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{client-server.pdf}
-	\caption{Client-Server paradigm in $\abftheory$-theory}
-	\label{fig:client-server}
-\end{figure}
-Availability, again, is the denial of use of a resource. In the $\abftheory$-theory, 
-in order to prevent a resource to be available, there
-exists only (i.e. inclusively) the following mutually exclusive cases
-that we illustrate using a client-server paradigm (depicted in Figure~\ref{fig:client-server}).
-\begin{enumerate}
-	\item\label{case:srv1} The resource on the server cannot be found, so that it is
-		possible for the client to contact the server and the server
-		sends a response back, but the output-port does not receive the
-		necessary information. This may be due to a software error such
-		that the functional architecture doesn't output a response, or
-		a problem in the flow of information (i.e. the functional
-		architecture doesn't receive the necessary information from
-		other resources), or an hardware failure \&c. This is formalized
-		as the relation between the beliefs produced by the functional
-		architecture and the ones received by the output port, i.e.
-		$\rcc(\behaviorRegion_s,\{\text{res}\})$ or 
-		$\rcc(\beliefRegion_s,\{\text{res}\})$ 
-	\item\label{case:srv2} The server hosting the resource is down, and then the server
-		cannot output responses (i.e. insecure output port due to drop-related issue) or receive
-		requests (insecure input port due to drop-related issue) or both.
-	\item\label{case:srv3} The client cannot properly compute request (simmetrically to Case~\ref{case:srv1})
-	\item The client is down, symmetric to Case~\ref{case:srv2}
-	\item The channel is down due to a drop related issue in one of the two channels 
-\end{enumerate}
-%All the cases but Case~\ref{case:srv1} and Case~\ref{case:srv3}, tell us how to map the
-%property of availability over the weaknesses of ports and communications.  The
-%Cases~\ref{case:srv1} and~\ref{case:srv3}, are exceptions because the formulas $res$ and $req$ 
-%detail the specification with a singleton instead of a region of formulas (or, better, $|res|$ is equal to the one of the
-%base).
-%In other words, a channel, a port, an entity, or a functional block are
-%Fact (stated in a factorial way) and not Beliefs; what they produce or 
-%their non-specified internals are Beliefs.  In fact, we never assume an
-%entity not to be in the design implementing the specification.  Similarly, when a
-%functional block produces some information it produces a Belief, unless this Belief if detailed in the specification.
-%For example, the server produces the Belief $res$ but the Belief given that can only be in $EQ$
-%relation with the Fact $res$ (in contrast other Belief may not); therefore, $res$ is not really a belief but a Fact. 
-%$res$ as a Fact is something that
-%exists in the specification, while the Belief of $res$ in one of the possible
-%designs may not exist. 
-We shall, then, distinguish between two cases:
-\begin{definition}{\bf Agent-Availability --}\label{def:agent-a}
-when the property of availability is expressed over
-	an agent or a system, the list of possible designs violating the
-	availability property can be shown as the designs that allow drop ports
-	or drop channels. 
-\end{definition}\fixnote{mr}{all security definitions defines when the property is
-not preserved instead of when it's preserved}
-\begin{definition}{\bf Infosec-Availability --}\label{def:infosec-c}
-when the property of availability is expressed
-	over a formula, the list of possible designs that violates the
-	availability property is reduced to the part of the system that
-	influences that specific variable.
-\end{definition}
-
-In a similar way we can define integrity since this requirement may 
-be expressed over an agent/system or a specific formula (i.e. element of a Region).
-
-\begin{definition}{\bf Agent-Integrity --}\label{def:agent-a}
-when the property of integrity is expressed over
-	an agent or a system, the list of possible designs violating the
-	integrity property can be shown as the designs that allow insert or inject ports,
-	or inert or inject channels. 
-\end{definition}
-\begin{definition}{\bf Infosec-Integrity --}\label{def:infosec-c}
-when the property of integrity is expressed
-	over a formula, the list of possible designs that violates the
-	integrity property is reduced to the part of the system that
-	influences that specific variable.
-\end{definition}
-
-While availability and integrity are architectural property, meaning that they
-can be expressed over the behavior of a port or a channel (as constituent of
-the physical or functional architecture), confidentiality is not.
-Confidentiality states that an information shall not be understood and not that
-the information shall not be retrieved. So, if a port leaks an entire message,
-there is no confidentiality issue in general, unless that message was not
-encrypted at all, in which case the confidentiality leakage is straightforward. 
-On the other hand, if the
-information is not provided or altered there is an availability or integrity
-issue. A similar argument can be proposed for integrity sub-properties such as
-authenticity or non-repudiation. No wonder why it is often stated that
-availability and integrity are the first properties to be considered in CPS
-while confidentiality takes the first place when dealing with security
-protocols. In fact, integrity is often assumed by pattern-matching (e.g. in
-AVANTSSAR \autocite{avantssar}) and availability is hardly ever considered in
-protocol verification. The question is, can we express properties such as
-confidentiality, authentication, or non-repudiation as architectural properties
-in $\abftheory$-theory? Confidentiality (and Availability) is based on the concept 
-of secrecy so we shall first start from describing what secrecy is and how it
-can be used to guarantee confidentiality in a system.
-
-\Paragraph{Secrecy, Authenticity, and Key Agreement (Cryptographic Security)}\label{sec:encryption}
-As Maurer puts it (\autocite{Maurer1999information}), there exist two types of
-cryptographic security (i.e. scientific theories aiming at guaranteeing
-security through the use of cryptographic functions): `` The  security  of a
-cryptographic system can rely either on the computational infeasibility of
-breaking it (computational security), or on the theoretical impossibility of
-breaking it, even using in infinite computing power (information-theoretic or
-unconditional security)''. We are not currently interested in detailing the
-difference between the two and we simplify it as follows. A
-computationally-secure cryptographic function (usually called encryption
-function) $E(\varphi)=e$ (where $\varphi$ is a plain-text and $e$ a
-cipher-text) can be considered perfectly secure if a malicious agent that
-obtains the result of the cryptographic function ($e$) has bounded
-computational power, while a unconditional secure cryptographic functions are
-considered perfectly secure. In this context, security means that is is
-impossible to calculate $\varphi$ from $e$, therefore impossible to compute
-$D(e)=\varphi$.  As one can evince from \autocite{Shannon1948information}, the
-commonality between the two types of cryptographic functions is that they both
-rely on the assumption about the probabilistic behavior of the universe. They
-do so, for the random (i.e. with high entropy
-\autocite{Shannon1948information}) generation of the so called \emph{key}, a
-piece of information, parametric to the cryptographic function, used to
-guarantee the security. Therefore, we can conclude that it is feasible for an
-agent who knows $k$ to decrypt the cipher-text obtaining the clear-text (i.e. $D(k, E(k,\varphi))=\varphi$)
-while it is impossible for any agent who has not the proper key. In this
-case, the information is considered to be \emph{secret or confidential}.
-
-\begin{definition}{Infosec-Secrecy or Confidentiality --}\label{def:confidentiality}
-	When an Assertion has been generated as an output of a functional or
-	physical agent, port, or functional block such that it is said (in a
-	factorial way) to encrypt the input, that Assertion is considered
-	confidential.
-\end{definition}
-The quality of the confidentiality of information (Assertions or Beliefs depending on 
-the subsystem considered) are inversely proportional to the number of 
-integrity issues of the subsystem.
-
 \begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
-	A system security design is given by a system security specification, i.e. a set of
-	requirements on the integrity and availability over the physical and
-	functional architectures, that uniquely defines the designs built on
-	top of those requirements.
+	A system security design is given by a precise system security
+	specification over the physical and functional architectures, that
+	uniquely, i.e.  tested against the range of design possible in the
+	$\abftheory$-theory, defines the designs built on top of those
+	requirements.
 \end{hypothesis}
@@
 	If, given a system security specification as a collection of
-	requirements on the physical and functional architectures, there exist
+	requirements on the system, there exist
 	a non-unique design that respect those requirements, the number of
@@
 
-\fix{mr}{
-Authenticity: assurance that the information is from the source it claims to be
-from. Authenticity involves proof of identity. Authenticity is verified through
-authentication.  If information X arrives from A to B, A sent information X and
-B can verify it.  ??Authenticity implies integrity but Integrity doesn't imply
-authenticity??  ??Authenticity doesn't imply confidentiality but
-confidentiality does imply authenticity??
-
-Non-repudiation: the ability to ensure that a party to a contract or a
-communication cannot deny the authenticity of their signature on information
-(or the sending of) that they originated.  If information X arrives from A, A
-sent it and cannot deny it (partial overlap??).  If sender A sends information
-X, it cannot say he didn't do it.  Non-repudiation implies authenticity and
-integrity. Authenticity doesn't imply non-repudiation (replay attacks?).
-Integrity doesn't imply non-repudiation (replay attacks?). 
-}
 \subsection{Abstraction Level of a Specification in $\abftheory$-theory}
-As highlighted in \autocite{}, ``The level of abstraction is always a concern
-when it come sto the definition of a system model.'' (as we already mentioned in Section~\ref{sec:problem}).
+The level of abstraction is always a concern when it comes to the definition of
+a system model.  In this paper, we justified our assumptions on the atomicity
+of ports, channels, and functional blocks but nothing prevents a modeler to
+consider other atomic element of the system specification and design.  Those
+additional elements may detail the system and may change the syntactic sugar we
+defined; however, it won't change the underlying (semantic) reasoning unless
+another calculus is used (instead of RCC).  Furthermore, other Infosec
+properties may be considered, such as authenticity or non-repudiation. We,
+however, leave them as future development.

### 2020-04-08 formal_glossary.tex
overall review (minor changes)

@@
 \emph{threat source} since, given that this is the first and outermost concept
-in Figure~\ref{fig:security-safety}, we lack the concept of intent and
+in Figure~\ref{fig:safety-security}, we lack the concept of intent and
 maliciousness (which are not inherent to the Causality Principle). 
@@
 considered erroneous with respect to some security requirements decided a
-priori. Therefore, as we mentioned in Section~\ref{sec:problem_statement}, the
+priori. Therefore, as we mentioned in Section~\ref{sec:problem}, the
 formal verification of the security of protocols can be considered a scientific

### 2020-04-08 formal_glossary.tex
minor changes

@@
 \Paragraph{Attacker as a Dis-honest Agent}
-As we mentioned in Section~\ref{sec:problem_statement} when we postulated the
+As we mentioned in Section~\ref{sec:problem} when we postulated the
 security hypothesis, Weaknesses and Vulnerability are the two core security
@@
 introducing the concept of intentionality, motivation, or desire (see the BDI
-framework defined in \autocite{BDI} for a general overview of those concepts).
+framework defined in \autocite{Rao1995semantics} for a general overview of those concepts).
 The symbolic approach (also called Dolev-Yao model) in the formal verification

### 2020-04-16 tool.tex
add modelio project final draft and XMI parser

\begin{itemize}
\item structured in package where the \emph{architecture} is
\begin{itemize}
\item is defined by an object diagram where the \emph{physical architecture} is structured on dependency connections and
\item the \emph{functional architecture} is linked to the agents and defined as nested/inner object diagrams
\item the information flow between functional or physical components is explicit 
\end{itemize}
\end{itemize}

### 2020-04-16 prototypes/read_xmi.py

@@ -1,0 +1,93 @@
+import pprint
+import re 
+import xml.etree.ElementTree as ET
+tree = ET.parse('Engineering.xmi')
+root = tree.getroot()
+
+cps_spec_name="UC1-CPS"
+
+#{id:{name,type}} for agents, ports, and functional blocks
+components={}
+#{id1:id2} for information flows (id---message--->id)
+flows={}
+
+#this dictionary is only used to create
+#a "fake flow" from outer-port to inner-port (socket)
+#we could also merge based on name
+ports={}
+
+ID={}
+schema="{http://schema.omg.org/spec/XMI/2.1}"
+
+for child in root:
+    if(len(ID)==3):
+        break
+    if(child.tag == "packagedElement"):
+        if(child.attrib[schema+'type'] == "uml:Class"):
+            if(child.attrib['name'] == "InputPort"):
+                ID[child.attrib[schema+'id']]='InputPort'
+            if(child.attrib['name'] == "OutputPort"):
+                ID[child.attrib[schema+'id']]='OutputPort'
+            if(child.attrib['name'] == "FunctionalBlock"):
+                ID[child.attrib[schema+'id']]='FunctionalBlock'
+
+for child in root:
+        if(child.tag == "packagedElement" and child.attrib['name'] == cps_spec_name):
+            for innerchild in child:
+                #print(innerchild.tag, innerchild.attrib)
+                #find agents as Nodes in deployment diagram
+                if(innerchild.attrib[schema+'type'] == "uml:Node"):
+                    #print(innerchild.tag, innerchild.attrib)
+                    components[innerchild.attrib[schema+'id']]={'name':innerchild.attrib['name']}
+                    components[innerchild.attrib[schema+'id']]['type']='agent'
+                    components[innerchild.attrib[schema+'id']]['owner']='root'
+                    for attribute in innerchild:
+                        #print(attribute.tag, attribute.attrib)
+                        if(attribute.attrib[schema+'type'] == "uml:Port"):
+                            ports[attribute.attrib['name']]=attribute.attrib[schema+'id']
+                            components[attribute.attrib[schema+'id']]={'name':attribute.attrib['name']}
+                            components[attribute.attrib[schema+'id']]['type']='port'
+                            components[attribute.attrib[schema+'id']]['owner']=innerchild.attrib['name']
+                        if(attribute.attrib[schema+'type'] == "uml:Property"):
+                            components[attribute.attrib[schema+'id']]={'name':attribute.attrib['name']}
[…]

### 2020-04-17 tool.tex
add sockets, multiple-flows, channels, and bugfix

@@
 \end{itemize}
+\item port to socket must exist and have the same name
+\item socket to port may not (case loopback)
+\item a base defines the granularity. if a base is a bit, it faithfully? represents information. This is important 
+	for the definitino of Vulnerability. If a port has all the weaknesses but only 1 base is transferred over that port,
+		then ony the DR can be exploited into a vulnerability since a PO would be equal to EQ
 \end{itemize}

### 2020-04-17 prototypes/parse_model.py

@@ -1,0 +1,154 @@
+import sys
+import pprint
+import re 
+import xml.etree.ElementTree as ET
+tree = ET.parse('Engineering.xmi')
+root = tree.getroot()
+
+#CONSTANTS
+cps_spec_name="UC1-CPS"
+schema="{http://schema.omg.org/spec/XMI/2.1}"
+abf_theory_package="ABFTheory"
+
+#{id:{name,type}} for agents, ports, and functional blocks
+components={}
+#adjacency list
+#{id1:[id2]} for information flows (id1->id2)
+#{id1:[id2, id3]} for information flows (id1->id2 and id1->id3)
+flows={}
+
+
+ID={}
+
+for child in root:
+    if(child.tag == "packagedElement"):
+        if(child.attrib[schema+'type'] == "uml:Package" and child.attrib['name'] == abf_theory_package):
+            for elem in child:
+                if(elem.attrib['name'] == "Base"):
+                    ID[elem.attrib[schema+'id']]='Base'
+                    break
+        elif(child.attrib[schema+'type'] == "uml:Class"):
+            if(child.attrib['name'] == "InputPort"):
+                ID[child.attrib[schema+'id']]='InputPort'
+            elif(child.attrib['name'] == "OutputPort"):
+                ID[child.attrib[schema+'id']]='OutputPort'
+            elif(child.attrib['name'] == "FunctionalBlock"):
+                ID[child.attrib[schema+'id']]='FunctionalBlock'
+
+        #this is just a speedup but we need to increase this is we add support for other attributes 
+        if(len(ID)==4):
+            break
+
+for child in root:
+        if(child.tag == "packagedElement" and child.attrib['name'] == cps_spec_name):
+            for innerchild in child:
+                #print(innerchild.tag, innerchild.attrib)
+                #find agents as Nodes in deployment diagram
+                if(innerchild.attrib[schema+'type'] == "uml:Node"):
+                    #print(innerchild.tag, innerchild.attrib)
+                    components[innerchild.attrib[schema+'id']]={'name':innerchild.attrib['name']}
+                    components[innerchild.attrib[schema+'id']]['type']='agent'
+                    components[innerchild.attrib[schema+'id']]['owner']='root'
+                    for attribute in innerchild:
+                        print(attribute.tag, attribute.attrib)
+                        #print(attribute.attrib[schema+'type']) 
+                        if(attribute.attrib[schema+'type'] == "uml:Port"):
+                            components[attribute.attrib[schema+'id']]={'name':attribute.attrib['name']}
[…]

### 2020-04-17 prototypes/read_xmi.py

@@ -1,93 +1,0 @@
-import pprint
-import re 
-import xml.etree.ElementTree as ET
-tree = ET.parse('Engineering.xmi')
-root = tree.getroot()
-
-cps_spec_name="UC1-CPS"
-
-#{id:{name,type}} for agents, ports, and functional blocks
-components={}
-#{id1:id2} for information flows (id---message--->id)
-flows={}
-
-#this dictionary is only used to create
-#a "fake flow" from outer-port to inner-port (socket)
-#we could also merge based on name
-ports={}
-
-ID={}
-schema="{http://schema.omg.org/spec/XMI/2.1}"
-
-for child in root:
-    if(len(ID)==3):
-        break
-    if(child.tag == "packagedElement"):
-        if(child.attrib[schema+'type'] == "uml:Class"):
-            if(child.attrib['name'] == "InputPort"):
-                ID[child.attrib[schema+'id']]='InputPort'
-            if(child.attrib['name'] == "OutputPort"):
-                ID[child.attrib[schema+'id']]='OutputPort'
-            if(child.attrib['name'] == "FunctionalBlock"):
-                ID[child.attrib[schema+'id']]='FunctionalBlock'
-
-for child in root:
-        if(child.tag == "packagedElement" and child.attrib['name'] == cps_spec_name):
-            for innerchild in child:
-                #print(innerchild.tag, innerchild.attrib)
-                #find agents as Nodes in deployment diagram
-                if(innerchild.attrib[schema+'type'] == "uml:Node"):
-                    #print(innerchild.tag, innerchild.attrib)
-                    components[innerchild.attrib[schema+'id']]={'name':innerchild.attrib['name']}
-                    components[innerchild.attrib[schema+'id']]['type']='agent'
-                    components[innerchild.attrib[schema+'id']]['owner']='root'
-                    for attribute in innerchild:
-                        #print(attribute.tag, attribute.attrib)
-                        if(attribute.attrib[schema+'type'] == "uml:Port"):
-                            ports[attribute.attrib['name']]=attribute.attrib[schema+'id']
-                            components[attribute.attrib[schema+'id']]={'name':attribute.attrib['name']}
-                            components[attribute.attrib[schema+'id']]['type']='port'
-                            components[attribute.attrib[schema+'id']]['owner']=innerchild.attrib['name']
-                        if(attribute.attrib[schema+'type'] == "uml:Property"):
-                            components[attribute.attrib[schema+'id']]={'name':attribute.attrib['name']}
[…]

### 2020-04-18 prototypes/parse_model.py
structure into functions

@@ -2,112 +2,118 @@ import sys
 import pprint
 import re 
 import xml.etree.ElementTree as ET
-tree = ET.parse('Engineering.xmi')
-root = tree.getroot()
 
-#CONSTANTS
-cps_spec_name="UC1-CPS"
-schema="{http://schema.omg.org/spec/XMI/2.1}"
-abf_theory_package="ABFTheory"
-
-#{id:{name,type}} for agents, ports, and functional blocks
-components={}
-#adjacency list
-#{id1:[id2]} for information flows (id1->id2)
-#{id1:[id2, id3]} for information flows (id1->id2 and id1->id3)
-flows={}
-
-
-ID={}
-
-for child in root:
-    if(child.tag == "packagedElement"):
-        if(child.attrib[schema+'type'] == "uml:Package" and child.attrib['name'] == abf_theory_package):
-            for elem in child:
-                if(elem.attrib['name'] == "Base"):
-                    ID[elem.attrib[schema+'id']]='Base'
-                    break
-        elif(child.attrib[schema+'type'] == "uml:Class"):
-            if(child.attrib['name'] == "InputPort"):
-                ID[child.attrib[schema+'id']]='InputPort'
-            elif(child.attrib['name'] == "OutputPort"):
-                ID[child.attrib[schema+'id']]='OutputPort'
-            elif(child.attrib['name'] == "FunctionalBlock"):
-                ID[child.attrib[schema+'id']]='FunctionalBlock'
-
-        #this is just a speedup but we need to increase this is we add support for other attributes 
-        if(len(ID)==4):
-            break
-
-for child in root:
-        if(child.tag == "packagedElement" and child.attrib['name'] == cps_spec_name):
-            for innerchild in child:
-                #print(innerchild.tag, innerchild.attrib)
-                #find agents as Nodes in deployment diagram
-                if(innerchild.attrib[schema+'type'] == "uml:Node"):
+#given as inputs:
+# - xmi: v-research xmi containing the whole Engineering of:
+# -- the ABF-theory (of which the package name needs to be specified in abf_theory_package)
+# -- the package of the cps spec (in cps_spec_name)
+# -- and the scheme of the xmi (of which we support 2.1 even if we export 2.4.1 with modelio
+#returns as output a structure of components as a dictionary with two entries:
+# 1. components: {ID:{'name': 'sensor', 'owner': 'root', 'type': 'agent'}}
[…]

### 2020-04-20 engineering.tex
overall review (minor changes)
spellcheck
add features considered of a functional block

@@
 	every open region can be written as union of elements of that base.  If
-	a communication occurred and results in
+	a communication occurred and resulted in
 	$\dr{\assertionRegion}{\beliefRegion}$, there exist only two mutually
@@
 but inverting the flow (input$\rightarrow$output instead of
-output$\rightarrow$input).  One may, of course, argue that since there is a
+output$\rightarrow$input). In other words, we consider two properties:
+how inputs are interpreted, and how outputs are produced.
+One may, of course, argue that since there is a
 transformation process in the functional block the quality of that
@@
 protocols. In fact, integrity is often assumed by pattern-matching (e.g. in
-AVANTSSAR \autocite{avantssar}) and availability is hardly ever considered in
+AVANTSSAR \autocite{Armando2012avantssar}) and availability is hardly ever considered in
 protocol verification. The question is, can we express properties such as

### 2020-04-20 prototypes/parse_model.py
concluded generation of Regions, beginning of constraints

@@ -1,5 +1,7 @@
+#!/usr/bin/python3
+#author: Marco Rocchetto @V-Research
+
 import sys
-import pprint
 import re 
 import xml.etree.ElementTree as ET
 
@@ -26,9 +28,14 @@ def get_components_from_xmi(cps_spec_name,abf_theory_package="ABFTheory",schema=
         if(child.tag == "packagedElement"):
             if(child.attrib[schema+'type'] == "uml:Package" and child.attrib['name'] == abf_theory_package):
                 for elem in child:
-                    if(elem.attrib['name'] == "Base"):
+                    if(elem.attrib['name'] == "Fact"):
+                        ID[elem.attrib[schema+'id']]='Fact'
+                    elif(elem.attrib['name'] == "Belief"):
+                        ID[elem.attrib[schema+'id']]='Belief'
+                    elif(elem.attrib['name'] == "Assertion"):
+                        ID[elem.attrib[schema+'id']]='Assertion'
+                    elif(elem.attrib['name'] == "Base"):
                         ID[elem.attrib[schema+'id']]='Base'
-                        break
             elif(child.attrib[schema+'type'] == "uml:Class"):
                 if(child.attrib['name'] == "InputPort"):
                     ID[child.attrib[schema+'id']]='InputPort'
@@ -72,6 +79,12 @@ def get_components_from_xmi(cps_spec_name,abf_theory_package="ABFTheory",schema=
                                     components[attribute.attrib[schema+'id']]['type']='funblock'
                                 elif(ID[attribute.attrib['type']]=="Base"):
                                     components[attribute.attrib[schema+'id']]['type']='base'
+                                elif(ID[attribute.attrib['type']]=="Fact"):
+                                    components[attribute.attrib[schema+'id']]['type']='fact'
+                                elif(ID[attribute.attrib['type']]=="Belief"):
+                                    components[attribute.attrib[schema+'id']]['type']='belief'
+                                elif(ID[attribute.attrib['type']]=="Assertion"):
+                                    components[attribute.attrib[schema+'id']]['type']='assertion'
                                 else:
                                     print("DEBUG Attribute not supported")
                                     print("DEBUG name: ",attribute.attrib['name'])
@@ -82,13 +95,18 @@ def get_components_from_xmi(cps_spec_name,abf_theory_package="ABFTheory",schema=
                         else:
[…]

### 2020-04-20 prototypes/secra.py

@@ -1,0 +1,343 @@
+#!/usr/bin/python3
+#author: Marco Rocchetto @V-Research
+
+#requirements
+# 1) z3 python API (https://github.com/Z3Prover/z3/blob/master/README.md)
+
+import sys
+from z3 import *
+import scipy.special
+import itertools
+from parse_model import get_components_from_xmi
+
+################################
+# since we unfold the quantifiers we have to calculate
+# the (minimal) number of subregions of A, B, and F that allows
+# Z3 to find a model that satisfies the agent configuration
+# 
+# For example, if we have A,B,F and no subregions of those 3 regions 
+# there is no model that satisfies PO(A,B) /\ PO(B,F) /\ PO(A,F)
+# since PO(A,B) <-> PP(A,B) if A has no subregions 
+#
+# requires scipy for binomial calculation: pip install scipy
+# otherwise download wheel from https://pypi.python.org/pypi/scipy
+# and install it with pip install <downloaded-file>
+################################
+def calculate_regions(regions, Region):
+    region_of_subregions=[]
+    for k in range(1,len(regions)):
+        counter=0
+        for i in range(int(scipy.special.binom(len(regions), (k+1)))):
+            subregion="subregion_%s_%d"%((k+1), counter)
+            region_of_subregions.append(Const(subregion, Region))
+            #print("%d - (%d %d)=%d"%(i,len(regions),(k+1),int(scipy.special.binom(len(regions), (k+1)))))
+            counter+=1
+        
+    return region_of_subregions
+
+def topology(solver, regions_and_subregions, P):
+    ################################
+    # PART OF
+    ################################
+    
+    for s in regions_and_subregions:
+        #solver.add(P(s,s))
+        solver.assert_and_track(P(s,s), str("reflexivity(%s)"%s))
+    
+    for s1 in regions_and_subregions:
+        for s2 in regions_and_subregions:
+            #solver.add(Implies(And(P(s1,s2),P(s2,s1)), s1==s2))
+            solver.assert_and_track(Implies(And(P(s1,s2),P(s2,s1)), s1==s2), str("asymmetry(%s,%s)"%(s1,s2)))
+    
+    for s1 in regions_and_subregions:
+        for s2 in regions_and_subregions:
+            for s3 in regions_and_subregions:
+               #solver.add(Implies(And(P(s1,s2),P(s2,s3)), P(s1,s3)))
[…]

### 2020-04-20 prototypes/parse_model.py
add regions in structure components

@@ -95,18 +95,22 @@ def get_components_from_xmi(cps_spec_name,abf_theory_package="ABFTheory",schema=
                         else:
                             flows[innerchild.attrib['informationSource']]=[innerchild.attrib['informationTarget']]
     
-    #we create a channel per each flow between ports, and we remove the flow
+    #we create a channel per each flow between ports, and we update the flow f1->f2 to f1->channel channel->f2
     flow2del={}
+    #flow2add={}
     for f1k,f1v in flows.items():
         if(components[f1k]['type']=="outputport"):
             for target in f1v:
                 if(components[target]['type']=="inputport"):
-                    components[f1k+target]={'name':components[f1k]['name']+"2"+components[target]['name'],'owner':"root",'type':"channel",'source':f1k,'target':target}
+                    components[f1k+target]={'name':components[f1k]['name']+"2"+components[target]['name'],'owner':"root",'type':"channel"}#,'source':f1k,'target':target}
                     flow2del[f1k]=target
+                    #flow2add[f1k]=f1k+target
+                    #flow2add[f1k+target]=target
+
     for f2dk,f2dv in flow2del.items():
         flows[f2dk].remove(f2dv)
-        if(len(flows[f2dk])==0):
-            del flows[f2dk]
+        flows[f2dk].append(f2dk+f2dv)
+        flows[f2dk+f2dv]=[f2dv]
     
     #we create a "fake flow" from port-socket 
     #for each port there must be a socket, the opposite may not be true
@@ -156,9 +160,9 @@ def dot(cps_spec_name, components, flows):
                         f.write(" [style=dotted]")
                     f.write(";")
             f.write("\n\t}")
-        elif(v['type']=="channel"):
-            f.write("\n%s_%s -> %s_%s"%(components[v['source']]['name'],components[v['source']]['type'],components[v['target']]['name'],components[v['target']]['type']))
-            f.write(" [penwidth=5, arrowhead=none];")
+        #elif(v['type']=="channel"):
+        #    f.write("\n%s_%s -> %s_%s"%(components[v['source']]['name'],components[v['source']]['type'],components[v['target']]['name'],components[v['target']]['type']))
+        #    f.write(" [penwidth=5, arrowhead=none];")
 
     for f1,f2 in flows.items():
         for e in f2:

### 2020-04-20 prototypes/secra.py

@@ -152,83 +152,82 @@ def rcc_five(solver, regions_and_subregions, C, P, O, EQ, PP, PO, PPi, DR):
 #input
 # -spec: string with the name of the package of the spec
 #output
-# -region_mapping is a dictionary with two entries:
-# --regions (assertions, beliefs, and facts)
-# --map containing the mapping between components and regions
+# -component_constraints is a dictionary with two entries:
+# --components updated with regions (assertions, beliefs, and facts)
 # --constraints defining an equality constraint between the LHS and RHS of each flow, except for the flow in the channel
 def create_regions_from_xmi(spec):
     components_flows=get_components_from_xmi(spec)
     components=components_flows['components']
     flows=components_flows['flows']
 
-    #DEBUG
-    import pprint
-    pp=pprint.PrettyPrinter(indent=0)
-    pp.pprint(components)
-    pp.pprint(flows)
-
+    region2components={}
     region_id=0
-    region2components={}    
-    regions={}
     for ck,cv in components.items():
         if(cv['type']=="agent"):
             region2components[ck]=region_id
-            regions["A"+str(region_id)]=Const("A"+str(region_id),Region)
-            regions["B"+str(region_id)]=Const("B"+str(region_id),Region)
-            regions["F"+str(region_id)]=Const("F"+str(region_id),Region)
-        elif(cv['type']=="inputport" or cv['type']=="outputport"):
+            cv['regions']={}
+            cv['regions']['assertion']=Const("A"+str(region_id),Region)
+            cv['regions']['belief']=Const("B"+str(region_id),Region)
+            cv['regions']['fact']=Const("F"+str(region_id),Region)
+        elif(cv['type']=="inputport"):
             region2components[ck]=region_id
-            regions["A"+str(region_id)]=Const("A"+str(region_id),Region)
-            regions["B"+str(region_id)]=Const("B"+str(region_id),Region)
+            cv['regions']={}
+            cv['regions']['input']=Const("A"+str(region_id),Region)
+            cv['regions']['output']=Const("B"+str(region_id),Region)
+        elif(cv['type']=="outputport"):
+            region2components[ck]=region_id
+            cv['regions']={}
+            cv['regions']['input']=Const("B"+str(region_id),Region)
+            cv['regions']['output']=Const("A"+str(region_id),Region)
         elif(cv['type']=="funblock" or cv['type']=="inputsocket" or cv['type']=="outputsocket"):
[…]

### 2020-04-21 prototypes/secra.py
create solver with regions,constraints,topology,rcc5

@@ -55,43 +55,11 @@ def topology(solver, regions_and_subregions, P):
                #solver.add(Implies(And(P(s1,s2),P(s2,s3)), P(s1,s3)))
                solver.assert_and_track(Implies(And(P(s1,s2),P(s2,s3)), P(s1,s3)), str("transitivity(%s,%s,%s)"%(s1,s2,s3)))
 
-################################
-# with mereotopology we first define connects with and then part of
-################################
-def mereotopology(solver, regions_and_subregions, C, P):
-    ################################
-    # CONNECTS WITH
-    # define here the relation C (primitive binary relation called connects with)  
-    # which is reflexive and symmetric 
-    ################################
-    
-    for s in regions_and_subregions:
-        #solver.add(C(s,s))
-        solver.assert_and_track(C(s,s), str("riflexivity(%s,%s)"%(s,s)))
-    
-    for s1 in regions_and_subregions:
-        for s2 in regions_and_subregions:
-            #solver.add(C(s1,s2)==C(s2,s1))
-            solver.assert_and_track(C(s1,s2)==C(s2,s1), str("symmetry(%s,%s)"%(s1,s2)))
-    
-    ################################
-    # PART OF
-    # P(X,Y) : forall Z C(Z,X) => C(Z,Y)
-    ################################
-    
-    for s1 in regions_and_subregions:
-        for s2 in regions_and_subregions:
-            array=[]
-            for s3 in regions_and_subregions:
-                array.append(Implies(C(s3,s1), C(s3,s2)))
-            #solver.add(P(s1,s2) == Implies(C(s3,s1), C(s3,s2))) 
-            solver.assert_and_track(P(s1,s2) == And(array), str("P(%s,%s) and Z=%s"%(s1,s2,s3))) 
-
 
 ################################
 # add the 5 relation of rcc5 to the solver
 ################################
-def rcc_five(solver, regions_and_subregions, C, P, O, EQ, PP, PO, PPi, DR):
+def rcc_five(solver, regions_and_subregions, P, O, EQ, PP, PO, PPi, DR):
     ################################
     # OVERLAPS 
     # O(X,Y) : exists Z P(Z, X) /\ P(Z, Y) 
@@ -149,64 +117,58 @@ def rcc_five(solver, regions_and_subregions, C, P, O, EQ, PP, PO, PPi, DR):
[…]

### 2020-04-21 prototypes/secra.py
add full prototype (generates and verify cases)

@@ -233,8 +233,6 @@ print("add rcc5")
 rcc_five(solver, regions, P, O, EQ, PP, PO, PPi, DR)
 print("done")
 
-sys.exit(1)
-
 ################################
 # AGENT
 ################################

### 2020-04-21 prototypes/secra.py
optimized number of regions, add progress bar

@@ -181,7 +181,8 @@ def create_regions_from_xmi(spec):
     #each flow equates beliefs
     for fk,fv in flows.items():
         for r in fv: #fk->r is a flow
-            constraints.append(components[fk]['regions']['output']==components[r]['regions']['input'])
+            #constraints.append(components[fk]['regions']['output']==components[r]['regions']['input'])
+            components[fk]['regions']['output']=components[r]['regions']['input']
 
     #each agent's beliefs encompass the beliefs resulting from its components and the assertions of its ports
     for c in components.values():
@@ -193,6 +194,27 @@ def create_regions_from_xmi(spec):
 
     return {'components':components,'constraints':constraints}
 
+# Print iterations progress
+def printProgressBar (iteration, total, prefix = '', suffix = '', decimals = 1, length = 100, fill = '█', printEnd = "\r"):
+    """
+    Call in a loop to create terminal progress bar
+    @params:
+        iteration   - Required  : current iteration (Int)
+        total       - Required  : total iterations (Int)
+        prefix      - Optional  : prefix string (Str)
+        suffix      - Optional  : suffix string (Str)
+        decimals    - Optional  : positive number of decimals in percent complete (Int)
+        length      - Optional  : character length of bar (Int)
+        fill        - Optional  : bar fill character (Str)
+        printEnd    - Optional  : end character (e.g. "\r", "\r\n") (Str)
+    """
+    percent = ("{0:." + str(decimals) + "f}").format(100 * (iteration / float(total)))
+    filledLength = int(length * iteration // total)
+    bar = fill * filledLength + '-' * (length - filledLength)
+    print('\r%s |%s| %s%% %s' % (prefix, bar, percent, suffix), end = printEnd)
+    # Print New Line on Complete
+    if iteration == total:
+        print()
 
 spec="UC1-CPS"
 solver=Solver()
@@ -208,15 +230,19 @@ PO = Function('PO', Region, Region, BoolSort())
 PP = Function('PP', Region, Region, BoolSort())
 PPi= Function('Pi', Region, Region, BoolSort())
 
-print("parse package %s in XMI and calculate Regions",spec)
+print("parse package %s in XMI and calculate Regions"%spec)
 components_constraints=create_regions_from_xmi(spec)
 print("done")
 
+# create list of unique regions (and subregions) of the spec
[…]

### 2020-04-21 prototypes/secra.py
print results to file

@@ -195,6 +195,8 @@ def create_regions_from_xmi(spec):
     return {'components':components,'constraints':constraints}
 
 # Print iterations progress
+# taken from (thanks Greenstick):
+# https://stackoverflow.com/questions/3173320/text-progress-bar-in-the-console
 def printProgressBar (iteration, total, prefix = '', suffix = '', decimals = 1, length = 100, fill = '█', printEnd = "\r"):
     """
     Call in a loop to create terminal progress bar
@@ -232,7 +234,7 @@ PPi= Function('Pi', Region, Region, BoolSort())
 
 print("parse package %s in XMI and calculate Regions"%spec)
 components_constraints=create_regions_from_xmi(spec)
-print("done")
+print("done\n")
 
 # create list of unique regions (and subregions) of the spec
 # as a (time) speedup this can be an output of create_regions_from_xmi()
@@ -247,25 +249,24 @@ print("there are %s different Regions in %s"%(len(regions),spec))
 print("add constraints on regions")
 for c in components_constraints['constraints']:
     solver.add(c)
-print("done")
+    print(c)
+print("done\n")
 
 # add topology to solver
 print("add topology")
 topology(solver, regions, P)
-print("done")
+print("done\n")
 
 # add rcc5 to solver
 print("add rcc5")
 rcc_five(solver, regions, P, O, EQ, PP, PO, PPi, DR)
-print("done")
+print("done\n")
 
 #generates all the possible pairs of regions 
 pairs_regions=[]
 for i in range(len(regions)):
     for j in range(i+1,len(regions)):
         pairs_regions.append([regions[i],regions[j]])
-
-print(len(pairs_regions))
 
 #create a list of rcc5 relation per each pair of regions
 rcc5=[EQ,PP,PPi,PO,DR]
@@ -286,15 +287,19 @@ print("pairs of regions: ",len(pairs_regions))
 print("possible scenarios: ",len(itertables)**len(rcc5))
 #itertools creates the cartesian product on the vector of rcc5 relations
 #total_scenarios=len(list(itertools.product(*itertables)))
+f=open(spec+".out","w+")
+f.write("spec: %s\n"%spec)
+f.write("pairs of regions: %s\n"%str(len(pairs_regions)))
+f.write("possible scenarios: %s\n\n"%str(len(itertables)**len(rcc5)))
 for t in itertools.product(*itertables):
[…]

### 2020-04-22 engineering.tex
minor changes -- functional block

@@
 channel to output-port. In fact, the Weaknesses of a channel are defined in terms
-of the Weaknesses of Ports. Our intention is to introduce now the concept of
-functional architecture and functional blocks composing the functional architecture.
-Io order to define a functional block without encountering an infinite definition,
+of the Weaknesses of Ports. We now introduce the concept of
+functional architecture and its decomposition into functional blocks. 
+Io order to define a functional block without encountering a infinitely recursive definition,
 we must reach the same conclusions as for the Channel. Therefore, describing
@@
 occurring in an infinite regress. Therefore, we consider Functional Blocks as 
-correctly executing their abstract behavior, correctly transforming inputs into outputs (Beliefs).
+executing an abstract undefined behavior, of which we only observe the
+inputs and the resulting outputs (Beliefs).
 
 \begin{definition}{\bf Functional Block and Architecture --}\label{def:funblock}
-	A functional block of an agent $a$ 
+	A functional block of an agent 
 	takes a region of input-Beliefs and outputs a region of output-Beliefs. 
@@
 be argued not just for functional blocks (Behaviors) but also for Channels
-(Assertions). As we detail afterwards, the relations considered so
+(Assertions). \fix{mr}{The verification of the semantics of functional blocks and of the 
+protocols executed over the channels will be considered at design level while
+now we are still defining the specification}.
+As we detail afterwards, the relations considered so
 far revolves around the concept of integrity or availability, while the reminder of this section 
-will be focused on a broader analysis of security properties and requirements
-which will also consider other security properties such as availability and
+will be focused on a broader analysis of security properties and requirements;
+which will also consider other security properties, such as availability and
 confidentiality.  In order to define the quality of the internals of a functional block or channel, we first

### 2020-04-22 prototypes/secra.py
add function to print poset

@@ -177,13 +177,14 @@ def create_regions_from_xmi(spec):
 
         region_id+=1
 
-    constraints=[]    
     #each flow equates beliefs
     for fk,fv in flows.items():
         for r in fv: #fk->r is a flow
             #constraints.append(components[fk]['regions']['output']==components[r]['regions']['input'])
             components[fk]['regions']['output']=components[r]['regions']['input']
 
+    poset_graph={}
+    constraints=[]    
     #each agent's beliefs encompass the beliefs resulting from its components and the assertions of its ports
     for c in components.values():
         if(c['type']=="agent"):
@@ -191,8 +192,15 @@ def create_regions_from_xmi(spec):
                 if(c1['owner']==c['name']):
                     for r in c1['regions'].values():
                         constraints.append(P(c['regions'][get_region_type(r)],r))
+                        #DEBUG
+                        if(c['regions'][get_region_type(r)] in poset_graph):
+                            poset_graph[c['regions'][get_region_type(r)]].add(r)
+                        else:
+                            s=set()
+                            s.add(r)
+                            poset_graph[c['regions'][get_region_type(r)]]=s
 
-    return {'components':components,'constraints':constraints}
+    return {'components':components,'constraints':constraints,'poset':poset_graph}
 
 # Print iterations progress
 # taken from (thanks Greenstick):
@@ -217,6 +225,25 @@ def printProgressBar (iteration, total, prefix = '', suffix = '', decimals = 1, 
     # Print New Line on Complete
     if iteration == total:
         print()
+
+def print_poset(regions,poset):
+    f=open("poset.dot","w+")
+    f.write("digraph G {\n")
+
+    for r in regions:
+        if(r in poset.keys()):
+            continue
+        elif(r in poset.items()):
+            continue
+        else:
+            f.write("%s\n"%r)
+
+    for k,v in poset.items():
+        for v1 in v:
+            f.write("%s -> %s\n"%(k,v1))
+
+    f.write("\n}")
+    f.close()
 
 spec="UC1-CPS"
 solver=Solver()
@@ -245,6 +272,11 @@ for c in components_constraints['components'].values():
 #this may not be the most elegant solution...
 regions=list(regions)
 print("there are %s different Regions in %s"%(len(regions),spec))
+
+#import pprint
+#pprint.pprint(components_constraints)
+#print_poset(regions,components_constraints['poset'])
+#sys.exit(1)
 
 print("add constraints on regions")
 for c in components_constraints['constraints']:

### 2020-04-25 prototypes/parse_model.py
slides per demo

@@ -67,10 +67,10 @@ def get_components_from_xmi(cps_spec_name,abf_theory_package="ABFTheory",schema=
                                     components[attribute.attrib[schema+'id']]['type']='inputport'
                                 elif(ID[attribute.attrib['type']]=="OutputPort"):
                                     components[attribute.attrib[schema+'id']]['type']='outputport'
-                                components[attribute.attrib[schema+'id']]['owner']=innerchild.attrib['name']
+                                components[attribute.attrib[schema+'id']]['owner']=innerchild.attrib[schema+'id']
                             elif(attribute.attrib[schema+'type'] == "uml:Property"):
                                 components[attribute.attrib[schema+'id']]={'name':attribute.attrib['name']}
-                                components[attribute.attrib[schema+'id']]['owner']=innerchild.attrib['name']
+                                components[attribute.attrib[schema+'id']]['owner']=innerchild.attrib[schema+'id']
                                 if(ID[attribute.attrib['type']]=="InputPort"):
                                     components[attribute.attrib[schema+'id']]['type']='inputsocket'
                                 elif(ID[attribute.attrib['type']]=="OutputPort"):
@@ -148,7 +148,7 @@ def dot(cps_spec_name, components, flows):
         if(v['type']=="agent"):
             f.write("\n\tsubgraph cluster_%s {\n\t\tlabel=\"%s\";\n\t\tlabeljust=\"l\""%(v['name'],v['name']))
             for k2,v2 in components.items():
-                if(v2['owner']==v['name'] and v2['type']!="inputport" and v2['type']!="outputport"):
+                if(v2['owner']==k and v2['type']!="inputport" and v2['type']!="outputport"):
                     f.write("\n\t\t%s_%s"%(v2['name'],v2['type']))
                     if(v2['type']=="fact"):
                         f.write(" [style=dashed]")

### 2020-04-25 prototypes/secra.py

@@ -9,6 +9,8 @@ from z3 import *
 import scipy.special
 import itertools
 from parse_model import get_components_from_xmi
+#DEBUG
+import pprint
 
 ################################
 # since we unfold the quantifiers we have to calculate
@@ -23,17 +25,17 @@ from parse_model import get_components_from_xmi
 # otherwise download wheel from https://pypi.python.org/pypi/scipy
 # and install it with pip install <downloaded-file>
 ################################
-def calculate_regions(regions, Region):
-    region_of_subregions=[]
-    for k in range(1,len(regions)):
-        counter=0
-        for i in range(int(scipy.special.binom(len(regions), (k+1)))):
-            subregion="subregion_%s_%d"%((k+1), counter)
-            region_of_subregions.append(Const(subregion, Region))
-            #print("%d - (%d %d)=%d"%(i,len(regions),(k+1),int(scipy.special.binom(len(regions), (k+1)))))
-            counter+=1
-        
-    return region_of_subregions
+#def calculate_regions(regions, Base):
+#    region_of_subregions=[]
+#    for k in range(1,len(regions)):
+#        counter=0
+#        for i in range(int(scipy.special.binom(len(regions), (k+1)))):
+#            subregion="subregion_%s_%d"%((k+1), counter)
+#            region_of_subregions.append(Const(subregion, Base))
+#            #print("%d - (%d %d)=%d"%(i,len(regions),(k+1),int(scipy.special.binom(len(regions), (k+1)))))
+#            counter+=1
+#        
+#    return region_of_subregions
 
 def topology(solver, regions_and_subregions, P):
     ################################
@@ -117,24 +119,26 @@ def rcc_five(solver, regions_and_subregions, P, O, EQ, PP, PO, PPi, DR):
             #solver.add(PPi(s1,s2) == PP(s2, s1)) #  And(P(s2,s1), Not(P(s1,s2))))
             solver.assert_and_track(PPi(s1,s2) == PP(s2, s1), str("PPi(%s,%s)"%(s1,s2))) #  And(P(s2,s1), Not(P(s1,s2))))
 
-
-def get_region_type(region):
-    if(str(region).startswith("A")):
-        regiontype="assertion"
-    elif(str(region).startswith("B")):
-        regiontype="belief"
-    elif(str(region).startswith("F")):
-        regiontype="fact"
+#should we create objects with method returning a constant for Z3 of type Base? Maybe... maybe not
+def get_base_type(base):
[…]

### 2020-04-26 prototypes/secra.py
change relations after meeting with Chicco. add subregions

@@ -25,7 +25,7 @@ import pprint
 # otherwise download wheel from https://pypi.python.org/pypi/scipy
 # and install it with pip install <downloaded-file>
 ################################
-#def calculate_regions(regions, Base):
+#def calculate_subregions(regions, Base):
 #    region_of_subregions=[]
 #    for k in range(1,len(regions)):
 #        counter=0
@@ -36,6 +36,31 @@ import pprint
 #            counter+=1
 #        
 #    return region_of_subregions
+
+# each region in a pair needs to have
+# a subregion to share and one to keep
+# to be able to use RCC5. This would not
+# be necessary if we didn't unfold quantifiers;
+# but unfolding prevents Z3 from unswering unknown 
+# instead of sat/unsat
+def add_minimal_subregions(pairs,solver):
+    subregion=0
+    regions=set()
+    for p1,p2 in pairs:
+        s="S"+str(p1)+"_"+str(subregion)
+        solver.assert_and_track(P(Const(s,Base),p1), str("subregion(%s,%s)"%(s,p1)))
+        subregion+=1
+        s="S"+str(p1)+"_"+str(subregion)
+        solver.assert_and_track(P(Const(s,Base),p1), str("subregion(%s,%s)"%(s,p1)))
+        subregion+=1
+
+        s="S"+str(p2)+"_"+str(subregion)
+        solver.assert_and_track(P(Const(s,Base),p2), str("subregion(%s,%s)"%(s,p2)))
+        subregion+=1
+        s="S"+str(p2)+"_"+str(subregion)
+        solver.assert_and_track(P(Const(s,Base),p2), str("subregion(%s,%s)"%(s,p2)))
+        subregion+=1
+    return subregion
 
 def topology(solver, regions_and_subregions, P):
     ################################
@@ -174,6 +199,7 @@ def create_regions_from_xmi(spec):
         elif(cv['type']=="base"):
             cv['regions']={}
             cv['regions']['belief']=Const("B"+str(region_id),Base)
+            #cv['regions']['fact']=Const("F"+str(region_id),Base)
         else:
             continue
 
@@ -185,7 +211,7 @@ def create_regions_from_xmi(spec):
             for r in fv: #fk->r is a flow
                 if(components[fk]['type']=="base"):
                     components[fk]['regions']['belief']=components[r]['regions']['input']
-                if(components[r]['type']=="base"):
+                elif(components[r]['type']=="base"):
[…]

### 2020-04-27 prototypes/secra.py
full calculation of configurations

@@ -5,12 +5,37 @@
 # 1) z3 python API (https://github.com/Z3Prover/z3/blob/master/README.md)
 
 import sys
+import time
 from z3 import *
 import scipy.special
 import itertools
 from parse_model import get_components_from_xmi
 #DEBUG
 import pprint
+
+# Print iterations progress
+# taken from (thanks Greenstick):
+# https://stackoverflow.com/questions/3173320/text-progress-bar-in-the-console
+def printProgressBar (iteration, total, prefix = '', suffix = '', decimals = 1, length = 100, fill = '█', printEnd = "\r"):
+    """
+    Call in a loop to create terminal progress bar
+    @params:
+        iteration   - Required  : current iteration (Int)
+        total       - Required  : total iterations (Int)
+        prefix      - Optional  : prefix string (Str)
+        suffix      - Optional  : suffix string (Str)
+        decimals    - Optional  : positive number of decimals in percent complete (Int)
+        length      - Optional  : character length of bar (Int)
+        fill        - Optional  : bar fill character (Str)
+        printEnd    - Optional  : end character (e.g. "\r", "\r\n") (Str)
+    """
+    percent = ("{0:." + str(decimals) + "f}").format(100 * (iteration / float(total)))
+    filledLength = int(length * iteration // total)
+    bar = fill * filledLength + '-' * (length - filledLength)
+    print('\r%s |%s| %s%% %s' % (prefix, bar, percent, suffix), end = printEnd)
+    # Print New Line on Complete
+    if iteration == total:
+        print()
 
 ################################
 # since we unfold the quantifiers we have to calculate
@@ -44,57 +69,76 @@ import pprint
 # but unfolding prevents Z3 from unswering unknown 
 # instead of sat/unsat
 def add_minimal_subregions(pairs,solver):
-    subregion=0
+    num_subreg=0
     regions=set()
+    subregions=set()
     for p1,p2 in pairs:
-        s="S"+str(p1)+"_"+str(subregion)
-        solver.assert_and_track(P(Const(s,Base),p1), str("subregion(%s,%s)"%(s,p1)))
-        subregion+=1
-        s="S"+str(p1)+"_"+str(subregion)
[…]

### 2020-05-19 prototypes/secra.py
add weakness-risk relation in Excel

@@ -347,9 +347,9 @@ def generate_graph(components):
     f.close()
     return {'pairs':pairs,'num_pairs':num_pairs}
 
-def write_report(path,spec_package,risk_structure,components):
+def write_report(path,spec_package,risk_structure,cyclic_risk_struct,components):
     workbook = xlsxwriter.Workbook(os.path.join(path,spec_package+"_securityAssessment.xlsx"))
-
+    relation2row = {}
     #ARCHITECTURE (HW/SW requirements) sheet
     #create png and add to 
     architecture_sheet = workbook.add_worksheet("Architecture")
@@ -358,9 +358,10 @@ def write_report(path,spec_package,risk_structure,components):
     architecture_sheet.insert_image('B2', os.path.join(path,spec_package+"_model.png"))
 
     #WEAKNESSES sheet
-    first_row=["ID","Agent","Component","Comp. Type","Weakness","Mitigation","Weight","Indirect Weight","Status","Assegnee"]
+    first_row=["ID","Agent","Component","Comp. Type","Weakness","Mitigation","Weight","Status","Assegnee","Subgraph","Relation Type","Relation"]
 
-    weak_sheet = workbook.add_worksheet("Weaknesses")
+    weak_sheet_name="Weaknesses"
+    weak_sheet = workbook.add_worksheet(weak_sheet_name)
 
     weak_sheet.set_column(1, 8, 30)
     cell_format={}
@@ -399,8 +400,8 @@ def write_report(path,spec_package,risk_structure,components):
     #   forall c in Comp:
     #       if A1 is In(c) and A2 is Out(c) or viceversa:
     #           weakness_semantics(c,R,weak_channel)
-    risk_subgraphs={}
     weak_id=1
+    risk_sheet_struct={}
     for rel,pairs_weight in risk_structure.items():
         for pair,weight in pairs_weight.items():
             left=pair[0]
@@ -469,7 +470,9 @@ def write_report(path,spec_package,risk_structure,components):
                     weakness['agent']=owner_tmp['name']
                     break
 
-            weak_sheet.write(weak_id, 0, weak_id, cell_format['all_weak'])
+            relation2row[str(rel).lower()+str(pair).lower()]=weak_id+1
+            #weak_sheet.write(weak_id, 0, weak_id, cell_format['all_weak'])
[…]

### 2020-05-21 engineering.tex
add other fixnotes
add security equation. review based on tests

@@
 can be tested by the \atgmod. \fix{mr}{While we have described the overall view, from
-specificaiton to implementation, in this text we only consider the process from specificaiton to design.}
+specification to implementation, in this text we only consider the process from specification to design.}
 
@@
 
+\fixnote{mr}{this proof needs to be restructured as in slides v-research-demo}
 \begin{proof}
@@
 \end{definition}
-The quality of a functional block or architecture can be expressed as a
-relation between input and output belief, similarly to the quality of a channel
-but inverting the flow (input$\rightarrow$output instead of
-output$\rightarrow$input). In other words, we consider two properties:
-how inputs are interpreted, and how outputs are produced.
+It is evident that the quality of a functional block cannot be determined
+by the difference between its inputs and outputs (as we did for
+ports and channels). The behavior of a functional block
+cannot be determined in general; since any functional block will have 
+its own purpose based on functional requirements. 
+In other words, a functional block is a generic agent with no pre-defined general
+behavior (while ports and channels have a pre-defined behavior).
+Therefore, we can consider the relation between 
+the outputs of a functional block with respect to the requirements.
 One may, of course, argue that since there is a
@@
 be argued not just for functional blocks (Behaviors) but also for Channels
-(Assertions). \fix{mr}{The verification of the semantics of functional blocks and of the 
+(Assertions). The verification of the semantics of functional blocks and of the 
 protocols executed over the channels will be considered at design level while
-now we are still defining the specification}.
+now we are still defining the specification.
+
+\begin{itemize}
+	\item $\po{\behaviorRegion}{\factRegion}$ the component has a Byzantine behavior where occasionally outputs the expected
+	        output given the correct inputs. Not all the inputs are handled
+	        properly, nor all the expected outputs are always generated
+	        when correct inputs are given.
+	\item $\pp{\behaviorRegion}{\factRegion}$ part of the expected outputs are not
+	        generated in response to the correct
+	        inputs
+	\item $\ppi{\behaviorRegion}{\factRegion}$ the components
+	        correctly performs the expected behavior when the correct
+	        inputs are provided but is subject to input
+	        injections
+	\item $\dr{\behaviorRegion}{\factRegion}$ the component
+		never performs the expected behavior (e.g. physical
+		damage)
+\end{itemize}
+These weaknesses will be detailed afterwards in this section.
+
 As we detail afterwards, the relations considered so
@@
 \end{enumerate}
-%All the cases but Case~\ref{case:srv1} and Case~\ref{case:srv3}, tell us how to map the
-%property of availability over the weaknesses of ports and communications.  The
-%Cases~\ref{case:srv1} and~\ref{case:srv3}, are exceptions because the formulas $res$ and $req$ 
-%detail the specification with a singleton instead of a region of formulas (or, better, $|res|$ is equal to the one of the
-%base).
-%In other words, a channel, a port, an entity, or a functional block are
-%Fact (stated in a factorial way) and not Beliefs; what they produce or 
-%their non-specified internals are Beliefs.  In fact, we never assume an
-%entity not to be in the design implementing the specification.  Similarly, when a
-%functional block produces some information it produces a Belief, unless this Belief if detailed in the specification.
-%For example, the server produces the Belief $res$ but the Belief given that can only be in $EQ$
-%relation with the Fact $res$ (in contrast other Belief may not); therefore, $res$ is not really a belief but a Fact. 
-%$res$ as a Fact is something that
-%exists in the specification, while the Belief of $res$ in one of the possible
-%designs may not exist. 
 We shall, then, distinguish between two cases, as we mentioned beforehand:
@@
 	do not include those in which there exist: insert or inject ports, 
-	or inert or inject channels/functional-blocks. 
+	or insert or inject channels/functional-blocks. 
 \end{definition}
@@
 there is no confidentiality issue in general, unless that message was not
-encrypted at all, in which case the confidentiality leakage is straightforward. 
+encrypted, in which case the confidentiality leakage is straightforward. 
 On the other hand, if the
@@
 authenticity or non-repudiation. No wonder why it is often stated that
-availability and integrity are the first properties to be considered in CPS
+availability and integrity are the first properties to be considered in CPS secure engineering
 while confidentiality takes the first place when dealing with security
@@
 	a non-unique design that respect those requirements, the number of
-	equivalent designs that fullfills the requirements quantitatively
-	defines the magnitute of insecurity of a system design.
+	equivalent designs that fulfills the requirements quantitatively
+	defines the magnitude of insecurity of a system design.
 \end{hypothesis}
+
+Based on these hypothesis, security ($S$) can be expressed as the following equation.
+\begin{displaymath}
+	S=-((\prod_{a\in Ag} r^{\binom{n}{k}})-\delta_T)
+\end{displaymath}
+where $r^{\binom{n}{k}}$ represents the number of possible configurations (permutations) 
+of an agent based on the number $r$ of relations (e.g. EQ, DR, PO) with arity $k$ between
+$n$ diffrent sets in the $\abftheory$ theory. $\delta_T$ represents the 
+number of configurations which are not satisfiable with respect to the logical theory
+defining the 
+algebraic structure (topology) and constraints of the calculus (RCC). $Ag$ is the 
+set of all the agents in the system under evaluation.
 

### 2020-05-21 theory.tex
spellcheck
add fixnotes
add security equation. review based on tests

@@
 \Paragraph{Methodology}
-Before going into the details of the various terms and their applications for
-the definition of our theory, we 
-detail our methodology.
 The concept of Knowledge \autocite{Hintikka1962knowledge} (from an
@@
 				subsystem
-			\item Other aspects of the system, related to the two previous points, that we
-				describe afterwards in this paper, such as
-				the assets and security properties 
+			\item Assets and security properties
 		\end{itemize}
@@
 space for a single agent. The color coding of the table represents the 
-risk function as a risk matrix as we detail afterwards in this section.
+security risk related to a generic agent.
 
@@
 A similar argument can be done for completeness.
-
+\fix{mr}{it's missing the relation with equivalence. missing parallelism with 
+knowledge-belief/assertion (see slide v-research-demo)}
 \begin{figure}[t]
@@
 \end{definition}
-We note that often the risk matrix is defined by the relation between likelihood and impact.
-We argue that the likelihood is related to the number of possible non secure (i.e. non maximum secure) 
-configurations of a system and, similarly, the impact.
+We note that often the risk matrix is defined by the relation between likelihood and impact which we
+discuss in the next section.
 
@@
 \hline %eq
-	\eq{$\assertionRegion$}{$\factRegion$} & 
+	\eq{$\behaviorRegion$}{$\factRegion$} & 
 	\cellcolor{abf-rg-4}\dr{$\assertionRegion$}{$\factRegion$} & 

### 2020-05-25 reports/journal_paper/economy.tex
create report structure

--- 2020-02-19 v-research_experiment-alpha.tex
@@
-\documentclass{article}
-\usepackage{mathtools}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{enumitem}
-\usepackage{multicol}
-\usepackage[toc,page]{appendix}
-\usepackage[gen]{eurosym}
-\usepackage{tabularx}
-\usepackage{epigraph}
-\usepackage{textcomp}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{csquotes}% Recommended
-\usepackage{rotating}
-\usepackage[style=numeric]{biblatex}
-% \bibliography{<mybibfile>}% ONLY selects .bib file; syntax for version <= 1.1b
-\addbibresource{literature.bib}% Syntax for version >= 1.2
-
-\usepackage{graphicx}
-\graphicspath{{./images/}}
-
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\lhead{}
-\chead{}
-\rhead{}
-\lfoot{}
-\cfoot{\thepage}
-\rfoot{Property of V-Research S.r.l.}
-\renewcommand\headrulewidth{0pt}
-\renewcommand\footrulewidth{0pt}
-
-\usepackage[colorinlistoftodos]{todonotes}
-
-\newcommand{\Paragraph}[1]{\smallskip\noindent{\bf #1.}}
-\newcommand{\fixnote}[2]{\textbf{\color{red}{FIX}}\footnote{{\bf #1:} #2}}
-\newcommand{\fix}[2]{{\color{red} {\bf tofix:} #2}}
-% \renewcommand{\fixnote}[2]{}
-% \renewcommand{\fix}[2]{}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\theoremstyle{corollary}
-\newtheorem{corollary}{Corollary}[section]
-\theoremstyle{lemma}
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{theorem}
-\newtheorem{theorem}{Theorem}[section]
-\theoremstyle{theorem}
-\newtheorem{example}{Example}
-
-\usepackage{authblk}
-\date{}                     %% if you don't need date to appear
-\setcounter{Maxaffil}{0}
-\renewcommand\Affilfont{\itshape\small}
-
-% for fancy table with rotated column titles
-\usepackage{adjustbox}
-\newcolumntype{R}[2]{%
-    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
-    l%
-    <{\egroup}%
-}
-\newcommand*\rot{\multicolumn{1}{R{45}{0.01em}}}% no optional argument here, please!
-\newcommand*\rota{\multicolumn{1}{R{90}{0.01em}}}% no optional argument here, please!
-\def\hb{\hbox to 10.7 cm{}}
-
-% circles
-\usepackage{wasysym}
-\newcommand{\Tdot}{$\CIRCLE$}
-\newcommand{\Thdot}{$\LEFTcircle$}
-\newcommand{\Twdot}{$\Circle$}
-
-% formulas
-\newcommand{\kframe}{\mathbf{K}}
-\newcommand{\bframe}{\mathbf{B_K}}
-\newcommand{\possibleworlds}{G}
-\newcommand{\modalrelation}{R}
-\newcommand{\actualworld}{\omega*}
-\newcommand{\world}{\omega}
-\newcommand{\World}{\Omega}
-\newcommand{\kmodel}{M}
-\newcommand{\interpretation}{\sigma}
-\newcommand{\knowledge}[1]{\mathbb{K}_{#1}}
-\newcommand{\knows}[2]{K_{#1}#2}
-\newcommand{\belief}[1]{\mathbb{B}_{#1}}
-\newcommand{\believe}[2]{B_{#1}#2}
-\newcommand{\region}{\chi}
-\newcommand{\Region}{\mathrm{X}}
-\newcommand{\agentuniverse}{\mathcal{A}}
-\newcommand{\agent}{a}
-\newcommand{\follang}{\mathcal{L}}
-\newcommand{\weakness}{w}
-\newcommand{\Weakness}{W}
-\newcommand{\vulnerability}{v}
-\newcommand{\Vulnerability}{V}
-%rcc5
-\newcommand{\rcc}{rcc}
-\newcommand{\connects}[2]{C(#1,#2)}
-\newcommand{\disconnected}[2]{\neg C(#1,#2)}
-\newcommand{\partof}[2]{P(#1,#2)}
-\newcommand{\overlaps}[2]{O(#1,#2)}
-\newcommand{\eq}[2]{EQ(#1,#2)}
-\newcommand{\pp}[2]{PP(#1,#2)}
-\newcommand{\po}[2]{PO(#1,#2)}
-\newcommand{\ppi}[2]{PPi(#1,#2)}
-\newcommand{\dr}[2]{DR(#1,#2)}
-
-\makeindex
-
-\begin{document}
-
-\title{On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}}
-\author[1]{Francesco Beltramini}
-\author[1]{Marco Rocchetto}
-\author[2]{Luca Vigan\`o}
-\affil[1]{V-Research, Verona, Italy}
-\affil[2]{King's College London, UK}
-
-\maketitle
-
-\begin{abstract}
-	The objective of this research is to develop of a theory that defines
-	(all and only) the possible insecurity and security configurations of
-	any abstract system. The theory is structured upon other theories that 
-	defines how a component of a system can be abstracted into an agent, defining how
-	agents can be formalized (both syntactically and semantically) to
-	describe an abstract system, such as a graph. Some of these theories
-	(e.g. used for the semantic
-	definition of the abstract system) are the epistemological definition
-	of knowledge, the Belief-Desire-Intent and the Assertion-Belief-Fact 
-	framework of reference, 
-	mereology, and topological structure. We argue that a mereology is the
-	most appropriate abstract underlying structure, do to its generality,
-	for defining the expressiveness of the system abstraction.  Furthermore, a
-	mereology allows us to define an ontology rather than a taxonomy.  We
-	also correlate different abstractions of the system to the TRL and the
-	engineering V-model.
-
-	We implemented a formal theory (of axioms) of a mereotopology, and of
-	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
-	uses the Z3 SMT solver. The results show that a single component (i.e.
-	agent) of an abstract system has a definite number of  different
-	insecurity configurations (e.g. 53 using RCC5 over a topological
-	structure) and only 1 secure (i.e. expected) configurations. The
-	configurations are reported as models satisfying the abstract system
-	semantics.
-
-	We considered the philosophical definition of truth behind our
-	approach, rejecting ``proof'' by induction from partial empirical evidences.
-	Our theory can be applied to system engineering and we show a concrete
-	application of our theory to the risk assessment of an ad-hoc system.
-	Finally, we provide a number of ideas to support the engineering of
-	secure systems (e.g. purely cyber or cyber-physical).
-\end{abstract}
-\newpage
-
-\section{Introduction}\label{sec:intro}
-\epigraph{Humanum est errare}{{\itshape Seneca the Elder}}
 The European Commission states in\autocite{EU2019market} that: ``Cybersecurity
@@
 Department of Commerce).
-
-The cybersecurity industry is growing fast, e.g. as reported
-in\autocite{Nasdaq2018market}. For example, in\autocite{Forbes2017market},
-published by the Forbes, is stated that \euro5.3 billion of funding where
-poured by venture capitalist into cybersecurity companies in 2018. The Forbes,
-in the same article, also highlights another peculiar (as seemly contradictory)
-trend: ``[\ldots] during the same time period, the number of cybersecurity
-breaches increased exponentially''. The data reported by the NIST 
-through the official CPE (Common Platform Enumeration) Dictionary Statistics 
-on the NVD websites in \autocite{NIST2020CPEstatistics}, show that in 2016 the
-number of reported vulnerabilities reported where around $~6000$ while in 2019
-the number of vulnerabilities was above $16000$.  The scientific community also
-reports similar findings.  In fact, in\autocite{Herley2009so}, Cormac Herley
-(Microsoft Research) shows how basic cybersecurity principles (such as the
-confidentiality benefit over the clear text for passwords typed into forms,
-e.g. for logins in websites) are not fully understood or shared between the
-cybersecurity research community\autocite{Nielsen2009stop}.  The lack of
-understanding of basic security principle, the inverse proportionality between
-investments in cybersecurity and the number of reported vulnerabilities year
-after year, can be linked to the lack of a foundational theory on
-cybersecurity, as already highlighted by Cormac Herley
-in\autocite{Herley2016unfalsifiability}.
-
-In this article, we give the first scientific theory (to the best of our knowledge)
-on security.
-
-\Paragraph{Structure} In Section~\ref{sec:problem} we define and formalize the problem statement. 
-In Section~\ref{sec:theory} we outline our security theory, and in Section~\ref{sec:prototest}
-we describe the implementation of the theory and some empirical tests of the theory.
-Finally, in Section~\ref{sec:related} we conclude the paper with an overview of the related work.
-
-\section{Problem Statement}\label{sec:problem}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.6\textwidth]{protocol-example.pdf}
-	\caption{Abstraction of an ad-hoc esemplificative protocol execution}
-	\label{fig:protocol-example}
-\end{figure}
-
-In\autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
-``an asymmetry in computer security'', which he defines as follows: ``Things
-can be declared insecure by observation, but not the reverse. There is no
-observation that allows us to declare an arbitrary system or technique
-secure''. Herley then uses this argument to show that ``claims that any measure
-is necessary for security are empirically unfalsifiable''. Given that, any
-theory which is not falsifiable by an empirical experiment is well
-known\footnote{``A theory which is not refutable by any conceivable event is
-nonscientific. Irrefutability is not a virtue of a theory (as people often
-think) but a vice.'' -- Karl Popper, Conjectures and
-Refutations\autocite{popper1962conjectures}} to be nonscientific (i.e.
-unfalsifiability is a fallacy of a theory), Herley concludes that there is no
-scientific theory on cybersecurity; which means that cybersecurity lays in the
-realm of pseudo-sciences\autocite{Herley2016usenixvideo}.  Herley, e.g.
-in\autocite{Herley2017justifying}, discusses the implications of a
-nonscientific approach to cybersecurity, and highlights the tremendous impact on
-all the scientific research and engineering of systems; leading often to
-terrorism and wars, and wasting of resources in useless protections or
-overspending.  While the criticism is investigated
-in\autocite{Herley2016unfalsifiability}, no solution is provided.  On the
-contrary, the goal of this work \emph{is} to lay the foundations of a
-scientific cybersecurity theory. Furthermore, in Section~\ref{sec:sicurezza},
-we consider the problem raised by Herley not confined to ``computer security''
-but to any abstract system (so that our theory may hold for any sound
-implementation such as networks, mechanical, cyber, or cyber-physical system,
-or even a single computer or a single device such as an hard-drive).  There is
-also an apparent inconsistency in\autocite{Herley2016unfalsifiability} that we
-seek to clarify before following (as we agree) the scientific path draw by
-Herley: cybersecurity is defined as an abstract property in many formal
-approaches to the investigation of the security of systems, and the security of the design of a
-formally verified protocol is indeed falsifiable.  For example, in the protocol
-verification community, security is often defined as a formalization of the
-high-level properties confidentiality, integrity, and availability. The problem
-in such approaches is not the definition of what cybersecurity is, but the use
-of theories (such as the Dolev-Yao attacker model\footnote{For the sake of
-simplicity, the Dolev-Yao attacker can be considered as an abstraction of an
-active attacker who controls the network but cannot break
-cryptography.}\autocite{Dolev1983security})
-that only applies to specific instances (often called scenarios) and
-abstraction of the protocol. This, in turn, creates a false sense of security
-since requires assumptions on the abstraction of the system of which security
-is verified. As an example, for the formal security verification of the system
-in Figure~\ref{fig:protocol-example}, a formalized scenario needs to be defined
-by a modeler who chooses (among others): (i) a scope of the formalization (e.g.
-excluding the server that distributes the public key is often done when
-verifying the security of authentication protocols), (ii) the number of
-sessions (even tough some approaches do reason on an infinite number of
-sessions such as\autocite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
-the peers (e.g.  in the ASLan++ language\autocite{Oheimb2010aslan++}), and (iv)
-the abstraction of the cryptographic primitives (e.g.  ProVerif vs
-CryptoVerif\autocite{Blanchet2017symbolic}).  Some of the choices will
-completely change the results of the formal verification of the system. For
-example, under the perfect cryptography assumption\footnote{As defined
-in\autocite{Rocchetto2016cpdy}: ``In the so called perfect cryptography
-assumption, the security encryption scheme is suppose to be perfect, without
-any exploitable flaw, and so the only way for the attacker to decrypt a message
-is by using the proper key. That assumption is widely accepted in the security
-protocol community, and most of the formal reasoning tools for the analysis of
-security protocols abstract away the mathematical and implementation details of
-the encryption
-scheme\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
-and assuming that no violation to any security property is done after message
-I); in Figure~\ref{fig:protocol-example}, the freedom of choosing the scope
-determines that the flaws related to the dishonest impersonation of the Server
-may or may not be considered in the verification process.  This choice has
-tremendous impact on the focus and findings of the verification of the security
-of the protocol.  While this may seem to turn upon minutiae and foreseeable,
-this highlights the false sense of security that may derive from a
-non-scientific theory of system security\footnote{``To the superficial
-observer, the analysis of these forms seems to turn upon minutiae. It does in
-fact deal with minutiae, but they are of the same order as those dealt with in
-microscopic anatomy.'' -- Karl Marx, Capital Volume 1, 1867}.
-
-\subsection{Sicurezza: Safety and Security}\label{sec:sicurezza}
-In most of the natural languages, and in Italian too, the concepts of safety
-and security are not syntactically differentiated and both terms (safety and
-security) are expressed by the same word, e.g. sicurezza in Italian.  A
-semantic distinction between safety and security is correlated to a
-belief\footnote{A belief has to be intended as a proposition which is supposed
-to be true by the majority of humans in our society without a scientific
-underlying theory but based on partial empirical evidences or inductive proofs.} that
-safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
-natural environment (e.g. natural events such as wearing of hardware
-components) while security deals with \emph{incidents} posed by mankind (e.g.
-attackers and bugs).  The fundamental difference between nature and mankind (and,
-in turn, between safety and cybersecurity) is believed to be on the different
-intents\footnote{``The belief–desire–intention software model (BDI) is a
-software model developed for programming intelligent
-agents.''\autocite{wiki-bdi}. In the BDI model, the intents represents the
-deliberative state of an agent which determines the choice of that agent on
-what to do.} (accidents are unfortunate while incidents are not) of the causes
-that generates the threat; namely, nature is believed not to have malicious
-intents (but unfortunate causes-effects) while threats generated by mankind are
-believed to be malicious\footnote{Of course, logical flaws or bugs may be
-introduced by other means (e.g. ignorance) without explicit malicious intents,
-but the exploitation of those flaws is considered (for now, and detailed
-afterwards in the article) malicious, and then we consider any vulnerability to
-be malicious (without loss of generality) even if due to the lack of skills.}.
-An overview on the aforementioned aspects of safety and security is depicted in
-Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
-the terms that structure our current understanding of safety and security. 
-\begin{itemize}
-	\item \emph{Mankind} ``refers collectively to humans''
-\autocite{wiki-mankind}, while the concept of \emph{Nature} is
-		related ``to the intrinsic characteristics that plants,
-		animals, and other features of the world develop of their own
-		accord'' (e.g. the physical universe)\autocite{wiki-nature}. 
-		\begin{itemize}
-			\item So far, we have used several terms to refer to an
-				\emph{attacker}, i.e. threat agent or threat source,
-				considering those terms to be semantically
-				equivalent.  This ``shallowness'' raise form the
-				necessity of properly citing the different sources, but,
-				in the reminder of this paper, we consider the
-				Causality principle to be the \emph{threat
-				source}, Nature or Mankind to be the
-				\emph{threat agents} and an \emph{attacker} as
-				a specific malicious threat agent which materialize a
-				threat.
-		\end{itemize}
-	\item \emph{Vulnerability}\footnote{The term vulnerability is not
-		present in the Encyclopedia of Cryptography and Security, while
-		it is used in 12 entries (such as in the definition of
-		``penetration testing''\autocite{caddy2005pentest})
-		highlighting how commonly this word is used without a proper
-		supporting semantics}, as defined in\autocite{cnssi20104009}
-		(and adopted in\autocite{nist2013800-53}), is ``weakness in an
-		information system, system security procedures, internal
-		controls, or implementation that could be exploited by a threat
-		source''. On the one hand, the definition is broad to enclose
-		as much causes (that generates a vulnerability) as possible; on
-		the other hand, it derives from empirical evidences (which
-		should be considered beliefs\footnote{``For this view, that
-		\emph{That Which Is Not} exists, can never predominate. You
-		must debar your thought from this way of search, nor let
-		ordinary experience in its variety force you along this way,
-		(namely, that of allowing) the eye, sightless as it is, and the
-		ear, full of sound, and the tongue, to rule; but (you must)
-		judge by means of the Reason (Logos) the much-contested proof
-		which is expounded by me.'' -- Parmenides of Elea, On Nature
-		(circa 500 B.C.), fragments B7.1–8.2
-		\autocite{Hakim2016philosophy}} since they are partial results in nature) 
-		while a vulnerability should
-		be defined in a way that is empirically falsifiable. This means
-		that the term vulnerability should have a complete and sound
-		definition, so that no other causes (e.g.  other sources) but
-		the ones in the definition are responsible for a vulnerability.
-		Furthermore, the term ``threat sources'' used in the definition
-		in\autocite{cnssi20104009} may be identified with both Nature
-		and Mankind, not differentiating between safety and security.
-		In Definition~\ref{def:vulnerability}, we provide a formal
-		theory of vulnerability (so that the scientific community can
-		identify tests for the completeness and soundness of the
-		definition itself).
-\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{safety-security.pdf}
-	\caption{Overview security and safety keywords}
-	\label{fig:safety-security}
-\end{figure}
-
-Most of the safety-preserving principles in the field of engineering of
-safety-critical cyber-physical systems (such as elevators and aircraft), upon
-which safety requirements (e.g. in standards such as the IEC 61508 or 61511\autocite{IEC201761511}) are defined, have
-been defined following empirical tests and measurements. While reasoning by
-induction based on the empirical observation should be avoided, since it may
-easily lead to false beliefs instead of scientific theories, this approach is
-often justified by the supposed impossibility of defining a theory that
-correctly predicts failures which, in turn, pose hazards to a system. 
-To the best of our knowledge, and supported by\autocite{Herley2016unfalsifiability}, the
-correlation between predictability of environment and believed unpredictability
-of attackers (i.e. a malicious
-environment) has not been correlated to a theory on cybersecurity. 
-Therefore, inductive research efforts in
-predicting malicious effects are accepted (and published) in scientific
-conferences (e.g.~\autocite{Rocchetto2014CSRF}). A failure of a wire due to
-environment (e.g. due to humidity, dust, heat \&c) is defined from empirical
-evidences and processes have been standardized to test qualities of hardware components
-%convinced by the induction principle. What is already a
-%weak argument in general, 
-This process completely breaks down when a malicious environment (i.e. an attacker)
-is considered instead of the (supposedly honest and predictable)
-natural environment. Therefore, the same approach that is in use for safety,
-seems not to be applicable to test security.
-
-Going back to Figure~\ref{fig:safety-security}, a vulnerability does not
-necessarily become a threat for the system, unless exploited ``through a
-channel that allows the violation of the security policy
-[\ldots]''\autocite{cnssi20104009} (e.g. a software or procedure) that takes
-advantage of the vulnerability causing an \emph{attack} to the system, which may
-result in several correlated incidents and threats.  The process of
-exploitation of a defect as a vulnerability is reported in
-Figure~\ref{fig:safety-security} such that the difference between exploit and failure,
-and attack and accident is to be found just in the maliciousness of the intents
-that causes this process (i.e. excluding the intent, the terms are just syntactic transformation from a vulnerability to defect, from
-accident to incident). In the following, we conclude the informal definition of
-the terms that we used in this section and in Figure~\ref{fig:safety-security}.
-
-\begin{itemize}
-	\item \emph{Causality} refers to the causality principle; defined
-		in\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
-		connection of phenomena through which one thing (the cause)
-		under certain conditions gives rise to, causes something else
-		(the effect). The essence of causality is the generation and
-		determination of one phenomenon by another. In this respect
-		causality differs from various other kinds of connection, for
-		example, the simple temporal sequence of phenomena, of the
-		regularities of accompanying processes''.
-	\item An \emph{Exploit}\footnote{We note that the term exploit is only
-		used as a verb in\autocite{ISO2009information}} is ``An exploit
-		(from the English verb to exploit, meaning to use something to
-		one’s own advantage) is a piece of software, a chunk of data,
-		or a sequence of commands that takes advantage of a bug or
-		vulnerability to cause unintended or unanticipated behavior to
-		occur on computer software, hardware, or something electronic
-		(usually computerized).''\autocite{wiki-exploit}.
-	\item An \emph{Attack}, as defined by the International Standard
-		ISO/IEC 27000 is an ``attempt to destroy, expose, alter,
-		disable, steal or gain unauthorized access to or make
-		unauthorized use of an asset''; where an \emph{Asset} is
-		``anything that has value to the organization''. We note that for
-		the purpose of this article, we do not want to focus on a specific
-		organization or business to define asset but, in general, on any 
-		abstract organization (e.g. a company or a society).
-		We do not consider ethical hackers as attacking a system. 
-		In fact, we consider the term \emph{hack} as
-		non-malicious (see Hacker\autocite{Stallman2002hacker}).
-	\item A \emph{Threat}, as defined in\autocite{cnssi20104009}, is ``Any
-		circumstance or event with the potential to adversely impact
-		organizational operations (including mission, functions, image,
-		or reputation), organizational assets, individuals, other
-		organizations, or the Nation through an information system via
-		unauthorized access, destruction, disclosure, modification of
-		information, and/or denial of service''.
-	\item \emph{Defect}, ``anything that renders the product not reasonably
-		safe''\autocite{Robinson2019writing} (i.e. a characteristic of
-		an object which hinders its proper usability).
-	\item \emph{Failure}, as defined in\autocite{Merriam2020failure} as ``a state of
-		inability to perform a normal function''. The term is
-		structured and detailed in
-		\autocite{cnssi20104009,iet2017glossary} but relying on an
-		abstract notion of failure without a specific definition.
-	\item \emph{Hazard}, ``a potential source of
-		harm''\autocite{iet2017glossary}.
-\end{itemize}
-
-\subsection{Glossary -- A Formalization}\label{sec:glossary}
-We now define a formalizations of the concepts described in
-Section~\ref{sec:problem} and depicted in Figure~\ref{fig:safety-security}. We
-base our formalization on first order logic (FOL) with the standard
-truth-value semantics. The choice of this logic is required by the 
-semantics of the concepts formalized afterwards in this section.
-
-We consider a first-order language $\follang$ over a
-signature $\Sigma_P$ where $P,P',\ldots,P^n$ represent terms, and
-$\varphi,\varphi',\ldots,\varphi^n$ represent formulas.
-The syntax is defined as follows.
-\begin{displaymath}
-	\varphi := P~|~\neg\varphi~|~\varphi\wedge\varphi~|~\varphi\vee\varphi~|~\varphi\Rightarrow\varphi'~|~\forall P.~\varphi~|~\exists P.~\varphi
-\end{displaymath}
-where $\wedge$, $\neg$, $\vee$, and $\Rightarrow$ are connectives representing
-conjunction, negation, disjunction, and (material) implication respectively; while
-$\forall$ and $\exists$ represent the standard universal and existential
-quantifiers (resp.). Finally, the symbol ``$.$'' is just syntactic sugar. 
-We consider $\interpretation\subset\Phi\times\{\top,\bot\}$ as the interpretation
-function, where $\Phi$ is any collection of sentences in $\follang$ 
-and $\top$ and $\bot$ represent the concepts of ``Tautology/True''
-and ``Contradiction/False'' respectively.
-
-Following Figure~\ref{fig:safety-security}, we start by formalizing the outermost term: Causality.
-
-\subsubsection{Causality Principle}\label{sec:causality}
-We formalize the Causality Principle starting from a K modal
-logic\autocite{Garson2018modal} (i.e. without restrictions on the causality
-relation between worlds). The standard definition of K modal logic is given in
-Definition~\ref{def:modallogic} in terms of an interpretation function (which
-we named $\sigma$ with a slight abuse of notation) defined in
-Definition~\ref{def:modalinterpretation}. 
-\begin{definition}{\bf K Modal Logic --}\label{def:modallogic}
-A K-frame is a frame $ \kframe=\langle \possibleworlds,\modalrelation \rangle $
-	in the K modal logic where $ \modalrelation $ is the binary relation
-	(i.e. a set of ordered pairs) between possible worlds
-	$\modalrelation\subseteq\possibleworlds\times\possibleworlds$, where
-	$\possibleworlds$ represents the possible worlds, and
-	$\possibleworlds\neq\emptyset$. An actual world
-	$\actualworld\in\possibleworlds$ is assumed. For any proposition $P$,
-	an interpretation function $\interpretation(\world,P)$ returns the
-	truth value of $P$; e.g. $\interpretation(\world,P)=\top$ means that
-	$P$ holds in $\world$. A model is defined as the tuple
-	$\kmodel=\langle\possibleworlds,\modalrelation,\interpretation\rangle$.
-\end{definition}
-
-\begin{definition}{\bf Causality as K Modal Interpretation Function --}\label{def:modalinterpretation}
-	Causality is (recursively) defined as the modal interpretation function $\interpretation$, as follows. 
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation0)$] if $\interpretation(\world,P)=\top$ then $\world\models P$
-		\item[$(\interpretation1)$] $\world\models\neg P$ iff $\world\not\models P$
-		\item[$(\interpretation2)$] $\world\models P \wedge Q$ iff $\world\models P$ and $\world\models Q$
-		\item[$(\interpretation3)$] $\world\models\Box P$ iff for any world $\world'\in\possibleworlds$ if $\world \modalrelation\world'$ then $\world'\models P$
-		\item[$(\interpretation4)$] $\world\models\Diamond P$ iff there exists a set of worlds $\World'\subset\possibleworlds$ such that for any $\world'\in\World'$, if $\world\modalrelation\world'$ then $\world'\models P$
-		\item[$(\interpretation5)$] $\models P$ iff $\actualworld\models P$
-	\end{enumerate}
-	where truth is defined as necessary with $\Box$ and possible with $\Diamond$.
-\end{definition}
-
-%\begin{definition}{\bf Causality --}\label{def:causality}
-%	The Causality principle is defined over a K Modal Logic (i.e. without
-%	restrictions on the causality relation between worlds) as in
-%	Definition~\ref{def:modallogic}, with the standard interpretation
-%	function defined in~\ref{def:modalinterpretation}
-%\end{definition}
-
-The causality principle has been defined in its
-generic form. In fact, the accessibility relation $\modalrelation$ is free from
-any axiomatic restriction (e.g. it's not reflexive nor anti-reflexive).  
-We will focus in Section~\ref{sec:process} (for our tests) on 
-the application of our theory to CPS system engineering.
-More detailed case studies (i.e. \fix{mr}{a CPS as a smart power-grid}) 
-will be defined in Section~\ref{sec:prototest}, where
-the accessibility relation defines in details how the system itself evolves, due to causal relation 
-(i.e. by restricting the causality principle only to those cause-effect that defines
-the system). Therefore, the definition of $\modalrelation$
-will be specialized in a more strict way based on the application domains and case study.
-
-We note that we have defined the
-Causality principle without considering it a \emph{threat source} since 
-we lack the concept of intent and maliciousness. Similarly, in the 
-next sections we won't discriminate between Nature and Mankind until we introduce 
-the concepts of maliciousness in Section~\ref{sec:theory} 
-and then formally define a threat agent
-in the same section.
-
-\subsubsection{Agents: Mankind and Nature}\label{sec:mankind-nature}
-\begin{table}[t]
-\centering
-\setlength{\tabcolsep}{3.5pt}
-\renewcommand{\arraystretch}{1}
-\scriptsize
-\caption{RCC3, RCC5, and RCC8 relations between Regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
-\begin{tabular}{ccclll} 
-\rota{\textbf{RCC3}}&\rota{\textbf{RCC5}}&\rota{\textbf{RCC8}}&\textbf{Name} & \textbf{Notation} & \textbf{Definition} \\
-\hline
-&&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& $\mathit{X}\subseteq \mathit{Y}$ \\
-&&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& $\mathit{X}\not\subseteq \mathit{Y}$\\
-&&&Part of				& $\mathit{P}(\mathit{X},\mathit{Y})$		& $\forall \mathit{Z} ~\mathit{C}(\mathit{Z},\mathit{X}) \rightarrow \mathit{C}(\mathit{Z},\mathit{Y})$\\
-&&&Overlaps			& $\mathit{O}(\mathit{X},\mathit{Y})$		& $\exists \mathit{Z} ~\mathit{P}(\mathit{Z},\mathit{X})\wedge \mathit{P}(\mathit{Z},\mathit{Y})$\\
-\Tdot&&&  \textbf{Overlaps Not Equal} 	& $\mathit{ONE}(\mathit{X},\mathit{Y})$		& $\mathit{O}(\mathit{X},\mathit{Y}) \land \neg \mathit{EQ}(\mathit{X},\mathit{Y})$ \\
-\Tdot&\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
-\Tdot&\Tdot&\Tdot& \textbf{DiscRete from} 		& $\mathit{DR}(\mathit{X},\mathit{Y})$		& $\neg \mathit{O}(\mathit{X},\mathit{Y})$\\
-&\Tdot&\Tdot&\textbf{Partial-Overlap}	& $\mathit{PO}(\mathit{X},\mathit{Y})$ 		& $\mathit{O}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{X},\mathit{Y}) \wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\
-&\Tdot&&\textbf{Proper-Part-of} 	& $\mathit{PP}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\ 
-	&\Tdot&&\textbf{Proper-Part-of-\textit{\textbf{i}}nverse} & $\mathit{PPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{Y},\mathit{X}) \wedge \neg \mathit{P}(\mathit{X},\mathit{Y})$\\
-&&\Tdot&\textbf{Externally Connected} 	& $\mathit{EC}(\mathit{X},\mathit{Y})$ 		& $\mathit{C}(\mathit{X},\mathit{Y}) \wedge \neg\mathit{O}(\mathit{X},\mathit{Y})$\\ 
-&&\Tdot&\textbf{Tangential PP} 	& $\mathit{TPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Tangential PPi} 	& $\mathit{TPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{TPP}(\mathit{Y},\mathit{X})$\\ 
-&&\Tdot&\textbf{Non-Tangential PP} 	& $\mathit{NTPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\neg\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Non-Tangential PPi} 	& $\mathit{NTPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{NTPP}(\mathit{Y},\mathit{X})$\\ 
-\end{tabular}
-\end{table}
-
-Mankind and Nature are defined in Section~\ref{sec:problem} as two abstract
-agents, both as collections (i.e. an abstract type that does not imply a
-specific implementation) of their sub-agents (i.e. humans for Mankind and
-plants, animals, \&c.  for Nature). Similarly to\autocite{Santaca2016abf}, 
-we define Mankind and Nature, and any
-other agent in the reminder of this article, as a meronomy (an hierarchy of
-Part-Whole relations) based on a standard definition of mereology, i.e. based
-on the definition of Parthood relation between \emph{Parts}.  
-However, shall 
-consider different types of Part; so, we extend the mereology to a
-mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
-to increase the number of different types considered and to generalize the relations between Parts (as
-in Table~\ref{tab:rcc358}).
-For the sake of readability, we use the term \emph{Region} both to refer to a
-mereological Part and to a topological Region.
-The choice of mereotopology is also correlated to the objective of defining a formal
-ontology, which we use to define the (formal) semantics of the terms (Parts) in
-Section~\ref{sec:problem}, and of the concepts of safety and security
-(whole). We aim at creating a meronomy instead of the taxonomies such as the one provided
-in\autocite{NIST2020NVD,MITRE2020CVE} or instead of the poorly justified 
-CVSS\autocite{Mell2007CVSS} scoring system.
-
-A mereotopology, as defined e.g. in\autocite{Rachavelpula2017mereotopology},
-is an ordered mathematical structure where the basic relation between Regions is the
-reflexive and symmetric\fixnote{mr}{I guess it must be monotonic as defined in\autocite{Rachavelpula2017mereotopology} but I don't find it consistently in other papers.} 
-\emph{Parthood} relation $\subseteq$. 
-\begin{definition}{\bf Parthood --}\label{def:parthood}
-Given any pair of mereotopological Regions $X$ and $Y$,
-	\begin{enumerate}[noitemsep]
-		\item Reflexivity: $\forall X.~ (X\subseteq X)$
-		\item Symmetry: $\forall X, Y.~ (X\subseteq Y \Rightarrow Y\subseteq X)$
-	\end{enumerate}
-\end{definition}
-
-As later defined in Definition~\ref{def:agent}, 
-the Parthood relation orders a universe of agents $\agentuniverse$ by defining
-the so called \emph{Connects with} (see in Table~\ref{tab:rcc358}) relation 
-between Regions. We want this universe $\agentuniverse$ to be 
-expressible in FOL. In this way, we
-can reason both on the constituent of security (i.e. its terms and agents
-defining a system where security needs to be considered), and on 
-the evolution of those constituent w.r.t. cause-effects relations according
-to the modal structure of causality we defined in Section~\ref{sec:causality}.
-This will allow us (in Section~\ref{sec:process} and Section~\ref{sec:prototest})
-a better\fixnote{mr}{Can we prove this by construction?} 
-positioning w.r.t. risk assessment technologies
-(which most often reason on the constituent of a system design), and
-protocol verification tools (which requires some formalization of a cause-effect relation, e.g.
-in Linear Temporal Logic).
-
-As we argued in Section~\ref{sec:problem}, we must correlate the definition of
-agent (i.e. Mankind and Nature) in Definition~\ref{def:mankind-nature} to the
-mathematical structure of the logic that defines them.  We express Mankind and
-Nature as formulas over the theory of mereology and then in terms of
-mereotopological Regions, extending the interpretation function
-$\interpretation$, to include a formal theory of mereotopology.  We use the
-Region Connection Calculus (RCC), as defined
-in~\cite{bennettLogics,improvingRCC}, to provide an axiomatization of the
-spatial concepts and relations in first-order logic to correlate the algebraic
-structure to mereology. In its broader definition, the RCC theory is composed
-by eight axioms, and is known as RCC8. In the text, for brevity, 
-we will often focus only on RCC5 (without
-loss of generality) by not considering tangential connections between spatial
-Regions. We discuss the choice of RCC5 in more detail in
-Appendix~\ref{app:rcc}. In Table~\ref{tab:rcc}, we
-summarize the axioms of the Region Connection Calculus (see, e.g., \autocite{Grutter2008rcc}).
-
-\begin{definition}{\bf RCC axiomatization --}
-	For any $X,Y$ pair of Regions in a mereotopology:
-	\begin{itemize}[noitemsep]
-	\item[$(\interpretation6)$] $\interpretation(X\subseteq Y)$ iff [$\interpretation(X\subseteq X)=\top$, and $\interpretation(X\subseteq Y)=\bot$ or $\interpretation(Y\subseteq X)=\top$]\fixnote{mr}{would it be better/clearer or just correct to write $\sigma(X\subseteq Y)$ iff $\sigma(X\subseteq X\wedge [X\subseteq Y \vee Y\subseteq X])=\top$}
-	\item[$(\interpretation7)$] $\world\models\connects{X}{Y}$ iff $\world\models X\subseteq Y$ 
-	\item[$(\interpretation8)$] $\world\models\disconnected{X}{Y}$ iff  $\world\not\models X\subseteq Y$
-	\item[$(\interpretation9)$] $\world\models\partof{X}{Y}$ iff $\world\models\forall Z.~\connects{Z}{X}\Rightarrow\connects{Z}{Y}$
-	\item[$(\interpretation10)$] $\world\models\overlaps{X}{Y}$ iff $\world\models\exists Z.~\partof{Z}{X}\wedge\partof{Z}{Y}$
-	\item[$(\interpretation11)$] $\world\models\eq{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\partof{Y}{X}$
-	\item[$(\interpretation12)$] $\world\models\dr{X}{Y}$ iff $\world\models\neg\overlaps{X}{Y}$
-	\item[$(\interpretation13)$] $\world\models\po{X}{Y}$ iff $\world\models\overlaps{X}{Y}\wedge\neg\partof{X}{Y}\wedge\neg\partof{Y}{X}$
-	\item[$(\interpretation14)$] $\world\models\pp{X}{Y}$ iff $\world\models\partof{X}{Y}\wedge\neg\partof{Y}{X}$
-	\item[$(\interpretation15)$] $\world\models\ppi{X}{Y}$ iff $\world\models\partof{Y}{X}\wedge\neg\partof{X}{Y}$
-\end{itemize}
-where $Z$ is a mereotopological Region.
-\end{definition}
-
-\begin{definition}{\bf Agent: Mankind or Nature --}\label{def:agent}
-	An agent $\agent\in\agentuniverse$ is a tuple
-	$\langle\rcc(\region',\region''),\ldots,\rcc(\region^{n-1},\region^n)\rangle$
-	of RCC relations $\rcc$ over mereotopological Regions
-	${\region',\ldots,\region^n}\subseteq\Region$. 
-\end{definition}
-
-Currently, we do not distinguish between Mankind and Nature (since we still
-lack of the definition of ``malicious intent'', which is defined in
-Section~\ref{sec:theory}) and we have defined them as two generic
-agents.
-
-As depicted in Figure~\ref{fig:safety-security}, Causality, Mankind, and Nature
-have a dashed border representing their correlation in terms of cause-effect
-and then in terms of formal structure which defines them: Modal Logic.
-Vulnerability, Defect, Incident, Accident, and Threat, similarly, 
-are correlated (depicted as a solid border) in terms of underlying 
-formal structure: Mereotopology.
-
-\subsubsection{Regions: Vulnerability and Defect (and Weakness)}\label{sec:vulnerabilitydefect}
-As defined in Section~\ref{sec:problem}, a Vulnerability or a Defect is a
-\emph{Weakness}.  As an example, a categorization of weaknesses is given
-in\autocite{MITRE2020CWEresearch} with 808 weaknesses categorized as ``Research
-Concepts'', distributed as follows:
-\begin{itemize}[noitemsep]
-	\item Incorrect Calculation - ($682$)
-	\item Incorrect Access of Indexable Resource (``Range Error'') - ($118$)
-	\item Use of Insufficiently Random Values - ($330$)
-	\item Improper Interaction Between Multiple Correctly-Behaving Entities - ($435$)
-	\item Improper Control of a Resource Through its Lifetime - ($664$)
-	\item Insufficient Control Flow Management - ($691$)
-	\item Protection Mechanism Failure - ($693$)
-	\item Incorrect Comparison - ($697$)
-	\item Improper Check or Handling of Exceptional Conditions - ($703$)
-	\item Improper Enforcement of Message or Data Structure - ($707$)
-	\item Improper Adherence to Coding Standards - ($710$)
-\end{itemize}
-
-The definition given by the MITRE in\autocite{MITRE2020CWEweakness} of
-weakness is: ``Software weaknesses are errors that can lead to software
-vulnerabilities. A software vulnerability, such as those enumerated on the
-Common Vulnerabilities and Exposures (CVE) List, is a mistake in software that
-can be directly used by a hacker to gain access to a system or network''.
-The definition is circular if we interpret the word ``error'' and
-``mistake'' with the same semantics: a weakness is an error that leads to a
-vulnerability and a vulnerability is a mistake which, in turn, is a weakness.
-The only difference (between weakness and vulnerability) seems to be
-that one can consider weakness as a ground term and state that a
-vulnerability is caused by a weakness, i.e. 
-$\World,\Weakness\models\Diamond\Vulnerability\wedge\Weakness$ where
-$\Weakness,\Vulnerability$ are Regions of Weaknesses and Vulnerabilities (resp.);
-accepting the hierarchy in the CWE\autocite{CWE} as ground truth. 
-Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability) 
-or the CVE reported in the NVD as a ground truth.
-
-\begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
-	A Region $\region\subseteq\agent$ of an agent
-	$\agent\in\agentuniverse$, is defined as Weakness $\Weakness$ (i.e.
-	representing a weakness introduced by the agent into any phase of
-	production, e.g.  of the secure process development life-cycle, of any
-	system or subsystem) iff there exists $\world\in\possibleworlds$ such
-	that $\world,\Weakness\models\Diamond\exists\Weakness',\Vulnerability'.~
-	\rcc(\Weakness',\Vulnerability')\wedge\neg\dr{\Weakness'}{\Vulnerability'}$, 
-	where $\Weakness'\subseteq\Weakness$, and
-	$\Vulnerability'\subseteq\Vulnerability$ is a Region of an agent $\agent'\in\agentuniverse$.
-\end{definition}
-
-\begin{example}{CWE-116: Improper Encoding or Escaping of Output\autocite{CWE-116} --}
-	\begin{itemize} 
-		\item Description: The software prepares a structured message
-		for communication with another component, but encoding or
-		escaping of the data is either missing or done incorrectly. As
-		a result, the intended structure of the message is not
-		preserved. 
-		\item Example: This code displays an email
-		address that was submitted as part of a form. Example language JSP.
-		\begin{verbatim} 
-		<% String email = request.getParameter("email"); %> 
-		...
-		Email Address: <%= email %>
-		\end{verbatim}
-		The value read from the form parameter is
-		reflected back to the client browser without
-		having been encoded prior to output, allowing
-		various XSS attacks (CWE-79).
-		\item Observed Examples
-			\begin{itemize}
-			\item CVE-2008-4636\autocite{CVE-2008-4636}: OS command injection in backup
-				software using shell metacharacters in a
-				filename; correct behavior would
-				require that this filename could not be
-				changed.
-			\item CVE-2008-0769\autocite{CVE-2008-0769}: Web application does not set the
-				charset when sending a page to a browser,
-				allowing for XSS exploitation when a
-				browser chooses an unexpected encoding.
-			\item CVE-2008-0005\autocite{CVE-2008-0005}: Program does not set the charset
-				when sending a page to a browser, allowing for
-				XSS exploitation when a browser chooses
-				an unexpected encoding.
-			\item CVE-2008-5573\autocite{CVE-2008-5573}: SQL injection via password
-				parameter; a strong password might contain \&
-			\item CVE-2008-3773\autocite{CVE-2008-3773}: Cross-site scripting in chat
-				application via a message subject, which
-				normally might contain \& and other
-				XSS-related characters.
-			\item CVE-2008-0757\autocite{CVE-2008-0757}: Cross-site scripting in chat
-				application via a message, which normally might
-				be allowed to contain arbitrary
-				content.
-			\end{itemize}
-\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\begin{subfigure}[b]{.5\textwidth}
-		\centering
-		\includegraphics[width=.7\textwidth]{rcc5_CWE-116.pdf}
-		\caption{Representation with RCC3}
-		\label{fig:rcc5_CWE-116}
-	\end{subfigure}\begin{subfigure}[b]{.5\textwidth}
-		\centering
-		\includegraphics[width=.7\textwidth]{rcc8_CWE-116.pdf}
-		\caption{Representation with RCC8}
-		\label{fig:rcc8_CWE-116}
-	\end{subfigure}
-	\begin{subfigure}[b]{\textwidth}
-		\centering
-		\includegraphics[width=\textwidth]{KML_CWE-116.pdf}
-		\caption{Representation in K-Modal Logic}
-		\label{fig:KML_CWE-116}
-	\end{subfigure}
-	\caption{Relations between CWE-116 and correlated CVEs}
-\end{figure}
-Definition~\ref{def:weakness} states that CWE-116 is a weakness iff there exists a world in
-K Modal Logic, representing the system in which this weakness exists, 
-such that the natural evolution of this system (i.e. formalized by the causality principle) 
-make it possible to reach another state of the system (i.e. another, accessible world)
-where there exist a vulnerability that can be implemented from CWE-116 (and this relation 
-is not DR). The CWE website proposes the connection between CWE-116 and, for example, 
-CVE-2008-5573; a vulnerability of the login sub-system 
-of the Poll Pro v2.0\autocite{pollpro} system.
-The formal relation between the two is given as a link (i.e. URI) between the CWE and
-the CVE, the description of the relation is not defined but it is supposed to be
-inferred from the descriptions of the CWE-CVE. 
-We can formally represent this link as the ONE connection in RCC3, depicted in 
-Figure~\ref{fig:rcc5_CWE-116}, or EC connection in RCC8 (Figure~\ref{fig:rcc8_CWE-116}).
-
-\fixnote{mr}{shouldn't we consider the agent who introduces the weakness as dishonest? don't we say this in part 1?}
-\end{example}
-
-It is interesting to note that the CWE-CVE relation expresses the correlation
-between weaknesses and vulnerabilities in the most simple form, such that we
-can formalize all the relation as ONE in RCC3 or EC in RCC8.  To express a more
-complex relation between the two we shall analyze the definition of CWE-116.
-This is related to the Weakness-Vulnerability-Incident process (i.e. to the
-details on the implementation/realization in Figure~\ref{fig:safety-security})
-that we analyze in the next section, Section~\ref{sec:incidentaccident}.
-
-\begin{definition}{\bf Region: Vulnerability and Defect --}\label{def:vulnerability-defect}
-A region $\region\subseteq\agent$ of an agent $\agent\in\agentuniverse$ is
-	called \emph{Vulnerability} if the agent $\agent$ is referred to as Mankind, 
-	\emph{Defect} if the agent is referred to as Nature.
-\end{definition}
-
-\subsubsection{Process: Incidents and Accidents}\label{sec:incidentaccident}
-In our informal definition depicted in Figure~\ref{fig:safety-security}, an
-Incident is generated through a process caused by a Vulnerability (which,
-in turn, is caused by a Weakness). Symmetrically, Nature has a process from Defect
-to Accidents. We start by analyzing the process generated by Mankind, which is
-divided into the following phases (explained in the remainder of this sub-Section):
-%Agent$\rightarrow$Vulnerability$\rightarrow$Exploit$\rightarrow$Attack$\rightarrow$Incident.
-\begin{enumerate}[noitemsep]
-	\item An agent (Mankind or Nature) defines a system, e.g.
-by following an engineering process for creating a new system, or for redesigning or improving an existing system (e.g.
-a component or a collection of components)
-	\item The dogmatic definition of a system contains a Weakness (i.e.
-the agent, Mankind or Nature, introduces an unintended behavior), for example as an un-intended behavior or as an architectural fallacy.
-The Weakness can be applied to the abstract definition of the system.
-	\item The Weakness can be defined as a malicious process for a system, called Vulnerability (or Defect for Nature)
-	\item The process expressed by the Vulnerability can be made operational for a specific implementation of the system (i.e. implemented by Mankind or realized by Nature) 
-\end{enumerate}
-Those additional details are depicted in Figure~\ref{fig:safety-security_2}, an update
-of Figure~\ref{fig:safety-security}.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{safety-security_2.pdf}
-	\caption{Overview security and safety keywords}
-	\label{fig:safety-security_2}
-\end{figure}
-
-\Paragraph{Exploitation Flow} An \emph{Exploit} is an \emph{implementation} of
-a Vulnerability, where a Vulnerability is one of the many ways to transform a
-Weakness into a malicious process that can be applied\footnote{Meaning that the
-malicious process, i.e.  any structured procedure such as a protocol, can be
-somehow made operational, e.g. implemented} to one or many systems. So, the act
-of an agent (Mankind) of ``exploiting a Vulnerability'' is the process of
-making the Vulnerability operational, by implementing the Vulnerability w.r.t.
-a specific target system.  Whenever the agent is Nature, the implementation is
-more broadly considered a realization\footnote{``The state of being
-realized''\autocite{Merriam2020realization}}
-
-The Common Attack Pattern Enumeration and
-Classification\autocite{MITRE2020CAPEC} (CAPEC) online service provides a
-``community resource'' which aims at ``identifying and understanding attacks''.
-The MITRE writes on the CAPEC homepage that ``Understanding how the adversary
-operates is essential to effective cyber security. CAPEC helps by providing a
-comprehensive dictionary of known patterns of attack employed by adversaries to
-exploit known weaknesses in cyber-enabled capabilities. It can be used by
-analysts, developers, testers, and educators to advance community understanding
-and enhance defenses.''. We can extrapolate two major concepts: 
-\begin{enumerate}
-	\item while understanding the exploitation process is the aim of the
-		CWE and CVE initiatives, understanding the hacking process is
-		the aim of the CAPEC initiative.
-	\item CAPEC provides a \emph{dictionary} of \emph{known} attack patterns.
-		As stated largely in Section~\ref{sec:}, this dictionary contains
-		empirical evidences and should not be used to induce a cybersecurity 
-		theory; but it can be used to validate a theory.
-\end{enumerate}
-
-The process continues as follows.
-\begin{enumerate}[noitemsep]
-\setcounter{enumi}{4}
-	\item an agent (Mankind\footnote{For the sake of readability we focus only on Mankind as an agent}) the exploitation flow can be used to hack a system.
-We use the term hack (as hacking) to stress that nothing prevents this process from being honest, however, for the sake of simplicity, we focus on Attacks which are malicious by definition.
-	\item An Attack is applied to a system with malicious intent by an agent
-	\item The application of the attack results in an Incident which impacts, i.e. poses a Threat to, an Asset
-\end{enumerate}
-
-\Paragraph{Hacking Flow}
-
-\begin{example}
-given in natural language in the description of the 
-CWE-116 which we can analyze as follows:
-\begin{itemize}
-	\item ``the software prepares a structured message''
-	\item the message is intended ``for communication with another component''
-	\item ``but''
-	\item ``encoding or escaping of the data is either missing or done incorrectly''
-	\item ``[\ldots] result, the intended structure of the message is not preserved''
-\end{itemize}
-\end{example}
-
-In this Vulnerability is seen as a process through which an agent (i.e. Mankind)
-generates an Incident, i.e.
-
-We describe processes, in our current formalization, in terms of the K-Modal
-Structure as in Definition~\ref{def:modalinterpretation} (i.e. in terms of
-causality principle).
-
-\begin{definition}{\bf Vulnerability-Incident Process --}\label{def:vulnerability-incident}
-	Given two 
-\end{definition}
-
-The whole-part relation \fix{mr}{(i.e.  HAS-A\footnote{HAS-A is just an hint for the
-mechanization of this theory.  We'll come back to this in
-Section~\ref{sec:implementation}})} is given, in the following, over the K Modal
-logic described beforehand. To do so, we first need to formally define the
-concept of \emph{agent} in its abstract form. We build it following the same
-structure of\autocite{Santaca2016abf} but over the terms used in
-Section~\ref{sec:problem}, i.e. knowledge, and belief as epistemic concepts. We'll then 
-introduce the concept of \emph{intent} to define a \emph{threat agent}.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.8\textwidth]{knowledge-belief.pdf}
-	\caption{Informal representation of Knowledge and Belief}
-	\label{fig:knowledge-belief}
-\end{figure}
-
-\subsubsection{Region: Threat and Asset}\label{sec:threatasset}
-The difference between Knowledge and Belief is depicted in
-Figure~\ref{fig:knowledge-belief} (see \autocite{wiki-knowledgebelief}).  However,
-according to\autocite{Gettier2012knowledge},
-Knowledge\footnote{``\emph{Theaetetus}: [\ldots] He said that knowledge was
-true opinion accompanied by reason, but that unreasoning true opinion was
-outside of the sphere of knowledge; and matters of which there is not a
-rational explanation are unknowable -- yes, that is what he called them -- and
-those of which there is are knowable. [\ldots] \emph{Socrates}: [\ldots] the
-primary elements of which we and all else are composed admit of no rational
-explanation; for each alone by itself can only be named, and no qualification
-can be added, neither that it is nor that it is not, for that would at once be
-adding to it existence or non-existence, whereas we must add nothing to it, if
-we are to speak of that itself alone.  [\ldots]'' Plato -- Theaetetus 201
-\autocite{Plato1914Plato}} as an epistemological concept is difficult (and
-sometimes believed impossible\autocite{citation}) to formally define. In this
-work, we are not interested in 
-how knowledge can be precisely formalized from an epistemic standpoint. 
-We assume that a semantic of a correct definition of epistemic knowledge exists,
-for example the one given in\autocite{Hintikka1962knwoledge} by Hintikka, and we then define 
-knowledge in terms of the Kripke structure defined in Definition~\ref{def:causality}.
-
-\begin{definition}{\bf Knowledge --}\label{def:knowledge}
-Given an abstract collection of Agents $Ag$, and the modal operator
-	$\knows{a}{}$ (where $a\in Ag$), Knowledge is defined as a collection
-	of predicates known by an agent $\knowledge{a}=\bigcup_\Phi \knows{a}{\varphi}$ 
-	(where $\Phi$ is the collection of all the propositions known by $a$).
-	Given a proposition $P$, we extend the semantics of the Causality structure with:
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation6)$] $\world\models\knows{a}{P}$ iff
-			$\world'\models P$ for all $\world'$ such that
-			$\world\modalrelation\world'$.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Belief --}\label{def:belief}
-	\fix{mr}{we should help the reader in understanding the upgrades of the logic.
-We used propositional with modal operators with a K Kripke semantics and we are now introducing Belief as in Doxastic logic?
-and we have epistemological operators as knowledge}
-	Given an abstract collection of Agents $Ag$, and the modal operator
-	$\believe{a}{}$ (where $a\in Ag$), Belief is defined as a collection
-	of predicates believed by an agent $\belief{a}=\bigcup_\Phi \believe{a}{\varphi}$
-	(where $\Phi$ is the collection of all the propositions believed by $a$).
-	Given a proposition $P$, we extend the semantics of the Causality structure with:
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation7)$] $\world\models\believe{a}{P}$ iff
-			$\world\models\neg\knows{a}{\neg P}$ (i.e. the agent $a$ considers $P$ possible) 
-			and $\world'\models P$ for all
-			$\world'$ such that $\world\modalrelation\world'$.
-	\end{enumerate}
-\end{definition}
-
-\begin{definition}{\bf Mankind/Nature --}\label{def:mankind-nature}
-Mankind/Nature is represented as an agent.  
-	%countable logical signature $\Sigma_M$/$\Sigma_N$ over an
-	%abstract algebraic structure.  We also consider a countable signature
-	%of variables, disjoint from $\Sigma_M$/$\Sigma_N$, and write
-	%$\Sigma_{M_n}$/$\Sigma_{N_n}$ for the symbols of $\Sigma_M$/$\Sigma_N$
-	%with arity $n$. Thus, $\Sigma_0$ is a collection of constants, which we
-	%assume to have disjoint sub-collections that we refer to as
-	%\emph{sub-agents} of Mankind or Nature (i.e. humans, plants, \&c).
-	%The variables are untyped and can be instantiated with arbitrary types,
-	%yielding an untyped model.
-\end{definition}
-
-\subsubsection{Mereotopological Representation of Safety and Security}\label{sec:formalGlossary}
-The objectives of this section are:
-\begin{enumerate}[noitemsep]
-	\item formalize the abstract algebraic structure (as a mereotopology)
-		on top of which we have formalized the terms in
-		Figure~\ref{fig:safety-security} so far, i.e. Mankind, Nature,
-		Truths, Beliefs, and Knowledge.
-	\item correlate mereological structure to Causality and then Kripke
-		structure by defining the relation $R$ over the Kripke
-		structure as the mereotopological basic relation connects with,
-		i.e. reflexive and symmetric
-	\item we map signature to mereotopological Regions to express the
-		logical formulation of terms (e.g. Mankind, Nature, Truths,
-		Beliefs, Knowledge, Vulnerabilities) over the mereotopological
-		structure
-	\item we formalize the relations between terms using the RCC(5)
-\end{enumerate}
-
-\fix{mr}{change propositions in propositional language\url{http://logic.stanford.edu/intrologic/lectures/lecture_02.pdf}}
-
-Given that the concepts of Truths, Beliefs, and Knowledge are positioned over
-an underlying structure of Propositions, the first concept to formalize is that
-of Propositions. We do not restrict Propositions to a specific structure but,
-as for Mankind and Nature, we refer to them as a collection.  We express the
-generality of a collection as a (propositional) signature over an abstract
-algebraic structure. In our formalization, we use mereotopology as algebraic
-structure (due to its generality, as discussed in Section~\ref{sec:problem})
-and the RCC to formalize concepts over a mereotopology.
-
-We can now define the possible mereotopological relations between Belief and
-Knowledge (informally depicted in Figure~\ref{fig:knowledge-belief}) and then
-define the peculiarity of the believed difference between Mankind and Nature 
-(i.e. maliciousness as intent). 
-We first start from the formalization of Figure~\ref{fig:knowledge-belief}, as
-follows.
-
-\begin{definition}{\bf Information}\label{def:information}
-\end{definition}
-\begin{definition}{\bf Desire}\label{def:desire}
-\end{definition}
-\begin{definition}{\bf Intent}\label{def:intent}
-\end{definition}
-
-\subsection{Vulnerability and Defect}\label{sec:vulnerability-defect}
-In order to formalize the semantics of the concepts (informally) described in
-Section~\ref{sec:problem},
-
-\begin{definition}{\bf Vulnerability}
-\end{definition}
-
-\subsection{Cybersecurity Engineering Process}\label{sec:process}
-%\epigraph{La differenza tra idee e realt\`a \`e la differenza tra
-%filosofia e ingegneria. Il lavoro per trasformare l'una nell'altra \`e ricerca
-%scientifica
-\epigraph{The difference between ideas and reality is the
-difference between philosophy and engineering. The work to transform one into
-the other is scientific research}{{\itshape V-Research}}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{vmodel.pdf}
-	\caption{Cybersecurity Engineering Life-cycle}
-	\label{fig:knowledge-belief}
-\end{figure}
-
-\subsubsection{Relevant Standards}\label{sec:standards}
-\begin{enumerate}[noitemsep]
-	\item DO-326A
-\end{enumerate}
-
-\section{Honesty and Maliciousness -- A Cybersecurity Theory}\label{sec:theory}
-\begin{enumerate}[noitemsep]
-	\item cybersecurity as a requirement
-	\item vulnerability\label{def:vulnerability}
-\end{enumerate}
-
-\section{Prototype and Empirical Tests}\label{sec:prototest}
-\subsection{Prototype Implementation}\label{sec:implementation}
-\subsection{Empirical Tests}\label{sec:tests}
-
-\section{Conclusion and Related Work}\label{sec:related}
-
-\printbibliography
-
-\begin{appendices}
-	\section{RCC3, RCC5, or RCC8}\label{app:rcc}
-	\section{Wrong idea 1}
-In order to apply consistently the mereotopological structure not just to
-reason the agents but also on their evolution, we extend the properties of the
-Parthood relation (reflexivity and symmetry) to the modal relation
-$\modalrelation$ of the K Modal logic, which in turns defines the causality
-principle. This updates the K Modal logic to a B (Brouwer) Modal Logic\autocite{Garson2018modal}.
-
-\begin{definition}{\bf B Modal Logic --}\label{thm:bk-parthood}
-	The B Modal Logic (defined, with a slight abuse of notation as
-	$\bframe=\langle\possibleworlds,\modalrelation\rangle$) is an extension
-	of the K modal logic with the following axioms
-	\begin{enumerate}[noitemsep]
-		\item[$(\interpretation8)$] $\world\models P\Rightarrow Q$ iff
-			$\world\not\models P$ or $\world\models Q$ (logical implication)~\footnote{In
-			Definition~\ref{def:modallogic} we didn't specify the
-			semantics in the case of the connective $\Rightarrow$
-			(i.e. logical implication), so we state it here. Given
-			that $P\Rightarrow Q$ is equisatisfiable to
-			$\neg(P\wedge \neg Q)$, for any proposition $P$ and
-			$Q$, $\world\models P\Rightarrow Q$ is equisatisfiable
-			to $\world\models \neg(P\wedge \neg Q)$. If
-			$\world\models \neg(P\wedge \neg Q)$ then
-			$\world\not\models P\wedge\neg Q$ and then
-			$\world\models \neg P$ or $\world\models Q$.}
-		\item[$(\interpretation9)$] $\Box P\Rightarrow P$ (which ``claims that whatever is necessary is the case''\autocite{Garson2018modal}), and
-		\item[$(\interpretation10)$] $P\Rightarrow\Box\Diamond P$
-			(which says that if $P$ is the case, then $P$ is
-			necessarily possible\autocite{Garson2018modal})
-	\end{enumerate}
-\end{definition}
-
-\begin{theorem}{\bf Causality, Parthood, and B Modal Logic}\label{thm:bk-parthood}
-	\begin{enumerate}[noitemsep]
-		\item (Reflexivity) $\bframe\models\Box P\Rightarrow P$ ($\interpretation8$) iff $\modalrelation$ is reflexive
-		\item (Symmetry) $\bframe\models P\Rightarrow\Box\Diamond$ ($\interpretation9$) iff $\modalrelation$ is symmetric
-		\item (Consistency) $\world\modalrelation\world'$ iff $\connects{\region}{\region'}$
-	\end{enumerate}
-	whenever $\world$ and $\world'$, and $\region$ and $\region'$ are logically equivalent ($\equiv$).
-\end{theorem}
-\begin{proof}
-	The proof is exhaustive over all possible cases.
-	\begin{enumerate}[noitemsep]
-		\item(Reflexivity) proven in\autocite{}
-		\item(Symmetry) proven in\autocite{}
-		\item(Equivalence) \fix{mr}{todo} If
-			$\world\modalrelation\world'$ and
-			$\world\equiv\region$, then if
-			it doesn't hold that $\connects{\region,\region'}$ it
-			follows that there is no reflexive and transitive
-			relation that correlates $\region$ to $\region'$. Given 
-			that Regions are defined as signatures
-	\end{enumerate}
-\end{proof}
-
-\end{appendices}
-
-\end{document}

### 2020-05-25 reports/journal_paper/formal_glossary.tex
create report structure

--- 2020-04-08 formal_glossary.tex
@@
 formalization on first order logic (FOL) with the standard truth-value
-semantics. The choice of this logic is usefull for the fomal treatment of the
+semantics\footnote{The choice of this logic is usefull for the fomal treatment of the
 theory we'll define in Section~\ref{sec:theory}.  We do not provide a proof
@@
 theory, nor that the semantics of English isn't an issue. So, we believe that
-this doesn't pose an issue for the formulation of the seucirty hypotheses since
-the hypothesis requires to be able to define \emph{an agent and its
-honest/dishonest behaviors in such a way that an agent can be used to define a
-system}. To the best of our knowledge, there is no result that shows that FOL
-cannot be used to do so; on the contrary, using classical logic such as FOL or
-Propositional logic is rather standard (along with Kripke structures for the
-definition of the underlying causality principle).
+this doesn't pose an issue for the formulation of the security hypotheses.}.
 
@@
 A K-frame is a frame $ \kframe=\langle \possibleworlds,\modalrelation \rangle $
-	in the K modal logic where $ \modalrelation $ is the binary relation
+	in the K modal logic where $ \modalrelation $ is a binary relation
 	(i.e. a set of ordered pairs) between possible worlds
@@
 what honesty is in a factious way.
-
 Similarly, in the next sections we won't be able to discriminate between Nature
@@
 Mankind, requires additional constraints or statements in order to be defined.
+\fixnote{mr}{is connects with an equivalence? can it be correctly represented by $\subseteq$?}
 
@@
 mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
-to increase the number of different types considered and to generalize the
-relations between Parts (as in Table~\ref{tab:rcc358}).  For the sake of
+considering the relations in Table~\ref{tab:rcc358}.
+For the sake of
 readability, we use the term \emph{Region} both to refer to a mereological Part
@@
 In order to correlate the definition of
-agent (i.e. Mankind and Nature) to the
+agents (i.e. Mankind and Nature) to the
 mathematical structure of the logic that defines them, we express Mankind and
@@
 mereotopological Regions, extending the interpretation function
-$\interpretation$, to include a formal theory of mereotopology.  We use the
+$\interpretation$ to include a formal theory of mereotopology.  We use the
 Region Connection Calculus (RCC), as defined
@@
 message into its components, or decrypt/encrypt if it has the proper security
-key) but not something that a dishonest agent cannot do. The main difference is
+key) but not something that an honest agent cannot do. The main difference is
 that an honest agent follows the nominal specification of the protocol while
@@
 sessions as it wants and the causality structure in which this scenario
-(protocol + attacker) has been defined, is searched for a state that is
+(protocol + attacker) has been defined, can be analyzed for a state that is
 considered erroneous with respect to some security requirements decided a

### 2020-05-25 reports/journal_paper/problem_statement.tex
create report structure

--- 2020-04-06 problem_statement.tex
@@
 high-level properties confidentiality, integrity, and availability. The problem
-in such approaches is not the definition of what cybersecurity is, but 
-the generality of the results since they takle a specific step (and not the first) of the engineering process (of system).
+in such approaches is not the definition of what cybersecurity is but 
+the generality of the results, since they takle a specific step (and not the first) of the engineering process (of system).
 Therefore, the theories underlying the verification are based on 
-assumptions which non-evidently apply to a general security theory (since, as mentioned before,
-there is no such a theory).
+assumptions which non-evidently apply to a general security theory.
 As an example, the so called Dolev-Yao attacker model\footnote{For the sake of
@@
 active attacker who controls the network but cannot break
-cryptography.}\autocite{Dolev1983security})
+cryptography.}\autocite{Dolev1983security}
 that only applies to specific instances (often called scenarios) and
@@
 		\autocite{Hakim2016philosophy}} since they are partial results in nature) 
-		while a vulnerability should
-		be defined in a way that is empirically falsifiable. This means
-		that the term vulnerability should have a complete and sound
+		%while a vulnerability should
+		%be defined in a way that is empirically falsifiable. This means
+		%that 
+		On the other hand, the term vulnerability should have a complete and sound
 		definition, so that no other causes (e.g.  other sources) but
@@
 		and Mankind, not differentiating between safety and security.
-		\fix{mr}{In Definition~\ref{def:vulnerability}}, we provide a formal
-		theory of vulnerability (so that the scientific community can
-		identify tests for the completeness and soundness of the
-		definition itself).
 \end{itemize}
@@
 often justified by the supposed impossibility of defining a theory that
-correctly predicts failures which, in turn, pose hazards to a system. 
-To the best of our knowledge, and supported by\autocite{Herley2016unfalsifiability}, the
-correlation between predictability of environment and believed unpredictability
-of attackers has not been correlated to a theory on cybersecurity. 
-Therefore, inductive research efforts in
+correctly predicts failures.
+%which, in turn, pose hazards to a system. 
+To the best of our knowledge, and supported by\autocite{Herley2016unfalsifiability}, %the
+there is no scientific theory that defines what a secure system is.
+%correlation between predictability of environment and believed unpredictability
+%of attackers has not been correlated to a theory on cybersecurity. 
+Therefore, 
+(inductive\footnote{``So, whenever they argue ``Every man is an
+animal and Socrates is a man; therefore Socrates is an animal,'' proposing to
+deduce from the universal proposition ``every man is an animal'' the particular
+proposition ``Socrates therefore is an animal,'' which in fact goes (as we have
+mentioned) to enstablish by way of induction the universal proposition, the
+fall into the error of circular reasoning, since they are enstablishing the
+universal proposition inductively by means of each of the particulars and
+deducing the particular proposition from the universal syllogistically.''
+Sextus Empiricus, Outlines of Pyrrhonism II-195
+\autocite{Empiricus1990Pyrrhonism}})
+research efforts in
 predicting malicious effects are accepted (and published) in scientific
@@
 
-Either a valid security hypothesis can be formulated based on one or more of
-the concepts reviewed (and depicted in Figure~\ref{fig:safety-security}) or it
-can't.  Since, in the next Section, we will define our security hypothesis (on
-top of which we'll deveop a security theory) on those terms, if the second case
-is true we will not formulate a valid security theory. However, while we want
-to highlight that the incompleteness of this review may be a potential issue;
-on the other hand, due to the nature of a literature review, it seems to us
-that only an inductive approach can be used and then any literature review will
-be incomplete.  If the first case applies, in one of those terms one shall find
-a valid ground on top of which formulate a security hypothesis (assuming that a
-valid security hypothesis exists).
+%Either a valid security hypothesis can be formulated based on one or more of
+%the concepts reviewed (and depicted in Figure~\ref{fig:safety-security}) or it
+%can't.  Since, in the next Section, we will define our security hypothesis (on
+%top of which we'll develop a security theory) on those terms, if the second case
+%is true we will not formulate a valid security theory. However, while we want
+%to highlight that the incompleteness of this review may be a potential issue;
+%on the other hand, due to the nature of a literature review, it seems to us
+%that only an inductive approach can be used and then any literature review will
+%be incomplete.  If the first case applies, in one of those terms one shall find
+%a valid ground on top of which formulate a security hypothesis (assuming that a
+%valid security hypothesis exists).
 
@@
 dis-honesty (also called maliciousness or adversarial) of an agent (often
-called adversary or attacker).  If those correlation are important for the
-insecurity of systems, it is plausible that they are important for the opposite
-concept, security.  Therefore, we define the following hypothesis.
+called adversary or attacker). We can synthesize this belief in the
+following hypothesis.
+
 \begin{hypothesis}{\bf Security as Honesty and Insecurity as Dishonesty --}
@@
 following the intended behavior/rules''.  In the case of a generic system, a
-dishonest agent is therefore, any agent that doesn't follow the intended
+dishonest agent is, therefore, any agent that doesn't follow the intended
 behavior (or functionality or logic) of the system.  Given the generality of
@@
 definition, such as CWE\autocite{MITRE2020CWEresearch}, CVE\autocite{CVE},
-CAPEC\autocite{MITRE2020CAPEC}, NVD\autocite{NIST2020NVD} and many many other;
+CAPEC\autocite{MITRE2020CAPEC}, NVD\autocite{NIST2020NVD};
 or that dishonesty of an agent with respect to a security protocol should be

### 2020-05-25 reports/journal_paper/tool.tex
create report structure

--- 2020-04-17 tool.tex
@@
 \end{itemize}
+
+\subsection{Impact and Likelihood}\label{sec:impact}
+As described in~\autocite{CORASMethod}, in order to quantitatively estimate the
+cybersecurity \emph{risk} of a CPS design, one needs to estimate both the
+\emph{likelihood} of \emph{attack paths} (attack from now on)\fixnote{mr}{check terminology
+consistency} along with their (negative) \emph{impact} on a requirement which, in
+turn, is related to some \emph{company assets}. An attack that has a negative impact
+on a requirement invalidates the requirement, leading to one or more \emph{incidents}.

### 2020-05-25 reports/journal_paper/engineering.tex
initial draft tool description

--- 2020-05-21 engineering.tex
@@
 functional architecture and physical boundaries that defines its physical
-architecture, as follows.
-\begin{enumerate}[start=1, label={req\arabic*)}]
-	\item the tank has no functional architecture, it is a purely physical component which has the only purpose of containing water
-	\item the communication is unidirectional from the tank to the input port of the sensor
-	\item the sensor has an input-port that receives only incoming communication from the tank
-	\item the sensor has two functional blocks
-		\begin{enumerate}[label={req4.\arabic*)}]
-			\item read\_water\_level, accepts inputs only from the input-port and outputs a new reading every $t$ seconds
-			\item comm\_water\_level, accepts inputs only from the read\_water\_level and communicates the readings to the output-port
-		\end{enumerate}
-	\item the sensor has an output-port that only outputs the messages generated by the comm\_water\_level to the input-port of the controller
-	\item the controller has in input-port that only receives incoming communications from the sensor and sends the messages to the functional architecture
-	\item the functional architecture of the controller is composed by three functional blocks
-		\begin{enumerate}[label={req7.\arabic*)}]
-			\item get\_water\_level, accepts inputs only from the input-port and outputs the water level to the main control function
-			\item control, calculate the next state of the valve based on the inputs receive from the get\_water\_level
-			\item send\_state, receives the output of the control function and communicates it to the output-port
-		\end{enumerate}
-	\item the controller has an output-port that only outputs the messages generated by the send\_state to the input-port of the actuator 
-	\item the actuator has an input-port that receives only incoming communication from the controller 
-	\item the functional architecture of the actuator is composed by two functional blocks
-		\begin{enumerate}[label={req10.\arabic*)}]
-			\item read\_change\_state, accepts inputs only from the input-port and outputs the next state of the valve to the close\_open\_valve function
-			\item close\_open\_valve, converts the digital next state received from the read\_change\_state an analog signal and communicates it the output-port
-		\end{enumerate}
-	\item the valve accepts analog messages from the output port of the actuator and changes its state based on those messages
-\end{enumerate}
+architecture, as detailed in Appendix~\ref{app:requirements}.
 
@@
 	\begin{enumerate}[start=1, label={W4.\arabic*)}]
-		\item dropped ($s\selectivedropport a$), 
-		\item inserted ($s\selectiveinsertport a$), or
-		\item injected ($s\selectiveinjectionport a$)
+		\item drop ($s\selectivedropport a$), 
+		\item insert ($s\selectiveinsertport a$), or
+		\item injection ($s\selectiveinjectionport a$)
 \end{enumerate}
@@
 	are in $EQ$ relation with the corresponding Facts mandated by the confidentiality/secrecy requirements.
-	A functional architecture composed exclusively by confidential/secrecy-preserving functional blocks
+	A functional architecture where all the secrecy-preserving functional blocks
+	are in $EQ$ relation with the corresponding Facts
 	is considered confidential/secrecy-preserving.
@@
 \begin{displaymath}
-	S=-((\prod_{a\in Ag} r^{\binom{n}{k}})-\delta_T)
+	S=1-(\prod_{a\in Ag} r^{\binom{n}{k}}-\delta_T)
 \end{displaymath}
@@
 set of all the agents in the system under evaluation.
+The constant 1 represents the nominal configuration of the system and then the
+only secure configuration. The product represents all the unintended, insecure, configurations.
 

### 2020-05-25 reports/journal_paper/tool.tex
initial draft tool description

@@
+In order to test our theory we implemented a tool-chain for
+the identification of weaknesses and the precise calculation of 
+potential insecure configurations. In this section, we describe
+the tool-chain and the concepts related to the risk estimantion, 
+such as likelihood and impact of attack paths.
+
+The engineering of the $\abftheory$ is summarized in the UML Class
+diagram in Figure~\ref{fig:secraclassdiagram}. The Class diagram
+represents the concepts introduced in Section~\ref{sec:engineering}.
+\begin{sidewaysfigure}
+	\centering
+	\includegraphics[width=\textwidth]{secra_classDiagram.png}
+	\caption{Security Risk Assessment -- Class Diagram}
+	\label{fig:secraclassdiagram}
+\end{sidewaysfigure}
+This allows us to define a CPS as a composition of the two (symmetric)
+structures depicted in Figure~\ref{fig:molecules}. A system can be
+designed in $\abftheory$ as the composition of structures formed
+by 3 elements: functional architecture, ports, and channels.
+If the information flows from left to right the port is an output port,
+otherwise (right-to-left) is an input port.
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{molecule.pdf}
+	\caption{Summary of $\abf$ Engineering Model}
+	\label{fig:molecules}
+\end{figure}
+
+\subsection{V-SecRA Tool-chain}
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{v-secra.pdf}
+	\caption{V-SecRA}
+	\label{fig:v-secra}
+\end{figure}
+The security risk assessment process starts with the definition
+of the use cases (see Figure~\ref{fig:usecase}) and architectural requirements 
+(see Appendix~\ref{app:requirements}); which we call specification.
+In our process, the specification is translated into a UML design where:
 \begin{itemize}
-\item structured in package where the \emph{architecture} is
-\begin{itemize}
-\item is defined by an object diagram where the \emph{physical architecture} is structured on dependency connections and
-\item the \emph{functional architecture} is linked to the agents and defined as nested/inner object diagrams
-\item the information flow between functional or physical components is explicit 
-\end{itemize}
-\item port to socket must exist and have the same name
-\item socket to port may not (case loopback)
-\item a base defines the granularity. if a base is a bit, it faithfully? represents information. This is important 
-	for the definitino of Vulnerability. If a port has all the weaknesses but only 1 base is transferred over that port,
-		then ony the DR can be exploited into a vulnerability since a PO would be equal to EQ
+	\item a \emph{Deployment Diagram} describes the \emph{Physical
+		Architecture}. Each agent is defined as a Node with (physical)
+		Ports, and Agent's ports are connected via Information Flow
+		connectors, representing the physical channel.
+	\item a \emph{Functional Architecture} is linked to each agent in the
+		deployment diagram and is defined by an \emph{Object Diagram}.
+		The Object Diagram is composed by Instances of Functional
+		Blocks, connected via Information Flow connectors.
+	\item the connection between the two diagrams is implemented by
+		``sockets'', functional blocks with the same name of the
+		corresponding physical port.
 \end{itemize}

### 2020-05-29 reports/journal_paper/theory.tex

--- 2020-05-21 theory.tex
@@
 security doesn't seem to stem from a malicious nature but, rather, insecurity
-raises from the lack of well-defined security requirements for the design
+raises from the lack of well-defined security requirements for the development
 process of a system. We argue that the high number of security vulnerability
 reported today are simply the realization of potential system configurations,
-which deviate from the nominal behavior just because the \emph{intended (or
+which deviate from the nominal behavior because the \emph{intended (or
 nominal) behavior} of the system is not precisely defined in the specification

### 2020-06-01 reports/white_paper/abstract.tex
cybersecurity requirements improvements

--- 2020-03-24 abstract.tex
@@
-The objective of this research is to develop of a theory that defines
-(all and only) the possible insecurity and security configurations of
-any abstract system. The theory is structured upon other theories that 
-defines how a component of a system can be abstracted into an agent, defining how
-agents can be formalized (both syntactically and semantically) to
-describe an abstract system, such as a graph. Some of these theories
-(e.g. used for the semantic
-definition of the abstract system) are the epistemological definition
-of knowledge, the Belief-Desire-Intent and the Assertion-Belief-Fact 
-framework of reference, 
-mereology, and topological structure. We argue that a mereology is the
-most appropriate abstract underlying structure, due to its generality,
-for defining the expressiveness of the system abstraction.  Furthermore, a
-mereology allows us to define an ontology rather than a taxonomy.  We
-also correlate different abstractions of the system to the TRL and the
-engineering V-model.
+The objective of this research is the development of a theory that defines (all
+and only) the possible insecurity and security configurations of any abstract
+system. The theory is structured upon other theories that defines how a
+component of a system can be abstracted into an agent, defining how agents can
+be formalized (both syntactically and semantically) to describe an abstract
+system, such as a graph. Some of these theories (e.g. used for the semantic
+definition of the abstract system) are the epistemological definition of
+knowledge, the Belief-Desire-Intent and the Assertion-Belief-Fact framework of
+reference, mereology, and topological structure. We argue that a mereology is
+the most appropriate abstract underlying structure, due to its generality, for
+defining the expressiveness of the system abstraction.  Furthermore, a
+mereology allows us to define an ontology rather than a taxonomy.  We also
+correlate different abstractions of the system to the TRL and the engineering
+V-model.
 
-We implemented a formal theory (of axioms) of a mereotopology, and of
-the Region Connection Calculus (RCC3 and RCC5) in a Python program that
-uses the Z3 SMT solver. The results show that a single component (i.e.
-agent) of an abstract system has a definite number of  different
-insecurity configurations (e.g. 53 using RCC5 over a topological
-structure) and only 1 secure (i.e. expected) configurations. The
-configurations are reported as models satisfying the abstract system
-semantics.
-
-We considered the philosophical definition of truth behind our
-approach, rejecting ``proof'' by induction from partial empirical evidences.
-Our theory can be applied to system engineering and we show a concrete
-application of our theory to the risk assessment of an ad-hoc system.
-Finally, we provide a number of ideas to support the engineering of
-secure systems (e.g. purely cyber or cyber-physical).
+We implemented a formal theory (of axioms) of a mereotopology, and of the
+Region Connection Calculus (RCC3 and RCC5) in a Python program that uses the Z3
+SMT solver. The results show that a single component (i.e.  agent) of an
+abstract system has a definite number of  different insecurity configurations
+(e.g. 53 using RCC5 over a topological structure) and only 1 secure (i.e.
+expected) configurations. The configurations are reported as models satisfying
+the abstract system semantics. The implementation allows us to  apply our
+theory to system engineering and showing concrete applications of our theory to
+the risk assessment of an ad-hoc system.

### 2020-06-01 reports/white_paper/introduction.tex
cybersecurity requirements improvements

--- 2020-04-06 introduction.tex
@@
 \autocite{NIST2020CPEstatistics}, show that in 2016 the number of reported
-vulnerabilities reported where around $~6000$ while in 2019 the number of
+vulnerabilities were around $~6000$ while in 2019 the number of
 vulnerabilities was above $16000$.  The scientific community also reports
@@
 in\autocite{Herley2016unfalsifiability}.  An important issue raised by Herley
-is that the methodology applied to by the security research community is quite
+is that the methodology applied by the security research community is quite
 often implicit and not always scientifically ``correct''. Herley equates
@@
 believe, the comments made by Hintikka in \autocite{Hintikka1993Information}
-should be taken into account. Therefore, without discussing what correct and
-non-correct is, we start by explicitly mention our method of enquiry.
+should be taken into account. Therefore, we start by explicitly mention our method of enquiry.
 
@@
 literature review on the main concepts and definitions related to security.  We
-formulate (inductively)\footnote{``So, whenever they argue ``Every man is an
-animal and Socrates is a man; therefore Socrates is an animal,'' proposing to
-deduce from the universal proposition ``every man is an animal'' the particular
-proposition ``Socrates therefore is an animal,'' which in fact goes (as we have
-mentioned) to enstablish by way of induction the universal proposition, the
-fall into the error of circular reasoning, since they are enstablishing the
-universal proposition inductively by means of each of the particulars and
-deducing the particular proposition from the universal syllogistically.''
-Sextus Empiricus, Outlines of Pyrrhonism II-195
-\autocite{Empiricus1990Pyrrhonism}} a security hypothesis, based on our review,
-in Section~\ref{sec:glossary}; which we use to propose a theory on system security in
-Section~\ref{sec:theory}. We apply our theory to an abstract and standard (i.e.
-based on standards) secure process development lifecycle, following the
-principle of requirement engineering in Section~\ref{sec:engineering} (with an
-ad-hoc running example, based on the SWaT testbed \autocite{Mathur2016swat}).
-We implement, in Section~\ref{sec:tool}, the theory in a prototype tool that can be used for the risk
-assessment of a design of cyber-physical systems.
+formulate 
+a security hypothesis %, based on our review,
+in Section~\ref{sec:glossary}; which we use to propose a theory on system
+security in Section~\ref{sec:theory}. In Section~\ref{sec:engineering}, we
+apply our theory to an abstract and standard (i.e.  based on standards) secure
+process development lifecycle, following the principle of requirement
+engineering  (with an ad-hoc running example, based on the SWaT testbed
+\autocite{Mathur2016swat}). This application shows how our theory can be used
+to predict all of the possible security weaknesses of a system, allowing the
+falsification of our theory.  In fact, if any security weaknesses were to be
+found in a system and not predicted by our theory, the theory could be declared
+incomplete.  Similarly, if a security weakness would be predicted by our theory
+but found to be impossible to have our theory could be declared as wrong.  In
+order to apply our theory we implemented, as described in
+Section~\ref{sec:tool}, the theory in a prototype tool that can be used for the
+risk assessment of a design of cyber-physical systems.

### 2020-06-01 reports/white_paper/requirements.tex
cybersecurity requirements improvements

--- 2020-04-20 engineering.tex
@@
-\epigraph{The difference between ideas and reality is like the
-difference between philosophy and engineering. The work to transform one into
-the other is scientific research}{{\itshape V-Research}}
-
-Several standards mandates a secure-by-design approach in which 
-cybersecurity shall be considered at the very early stages of the 
-design process. For example,
-\begin{enumerate}[noitemsep]
-	\item DO-326A -- ``Airworthiness Security Process Specification''
-		requires a cybersecurity risk assessment of the design and is
-		the ``are the only Acceptable Means of Compliance (AMC) by FAA
-		\& EASA for aviation cyber-security airworthiness certification,
-		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
-	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
-		Control System (ICS) Security''
-	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
-		for Cyber-Physical Vehicle Systems'' defines ``set of
-		high-level guiding principles for Cybersecurity as it relates
-		to cyber-physical vehicle systems'' and states that
-		``incorporate Cybersecurity into cyber-physical vehicle systems
-		from concept phase through production, operation, service, and
-		decommissioning''
-\end{enumerate}
-Companies which develop CPS (e.g. aerospace systems or elevator systems), must adhere
-to strict regulations and quality assurances processes. Introducing tools and procedures 
-in their processes requires a detailed and justified overview on how those
-tools and procedures can be used. In Figure~\ref{fig:spdl}, we provide an overview
-of a process that consider cybersecurity in the specification, design, and implementation
-stages. The overall process start with the specification of the architectural (both physical
-and functional) requirements from which the Weaknesses are automatically identified.
-Based on the Assets, architectures, and the Weaknesses the cybersecurity risk is
-calculated based on the $\abftheory$-theory. An iterative process between 
-the user (e.g. the engineer) and the \secramod{} module of the \abftool{} allows the user to move to the design phase
-when the risk level is considered acceptable. The design phase starts by automatically mapping
-the Weaknesses into Vulnerabilities which are, in turn, fed into the \designverifmod{}
-module along with the final specification. The Module returns a list of Vulnerabilities and
-potential Mitigations. The user then procees in implementing the CPS and the SW/HW choices
-can be tested by the \atgmod. \fix{mr}{While we have described the overall view, from
-specificaiton to implementation, in this text we only consider the process from specificaiton to design.}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{spdl.pdf}
-	\caption{Secure-by-design System Development Life-cycle}
-	\label{fig:spdl}
-\end{figure}
-
-In order to explain in details the SPDL, we use the following running example.
-\begin{example}
-	As depicted in Figure~\ref{fig:usecase}, the use case of the running example
-	(a small part of a subprocess of the SwAT testbed \autocite{Mathur2016swat}), shows
-	\begin{itemize}[noitemsep]
-		\item a tank that is filled with raw water coming from the inlet pipe 
-		\item a motorized valve (a valve with an actuator) such that:
-			\begin{itemize}
-				\item if opened, allows the raw water to flow into the tank
-				\item if closed, stops the raw water to flow into the tank
-			\end{itemize}
-		\item a sensor that reads the meters of raw water in the tank
-		\item a PLC controller, connected to the sensor and the
-			actuator, that receives the readings from the sensor
-			and regulates the quantity of water in the tank by
-			communicating a change in the state of the actuator
-			based on the following logic
-			\begin{itemize}
-				\item if the readings from the sensor reports a
-					water level in the tank above (or
-					equal) to 10m, the actuator shall close
-					the valve
-				\item if the readings are below 10m the actuator
-					shall open the valve.
-			\end{itemize}
-	\end{itemize}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.55\textwidth]{usecase.pdf}
-	\caption{Use Case -- Running Example}
-	\label{fig:usecase}
-\end{figure}
-\end{example}
-
-\subsection{Specification -- Elicitation of Security Requirements}\label{sec:properties}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{spec.pdf}
-	\caption{Specification -- Running Example}
-	\label{fig:spec}
-\end{figure}
-The first step of the SPDL is the definition of the requirements and then, for
-the purpose of this paper, of (i) security requirements, (ii) architectural
-(physical and functional) requirements.  Our focus in mainly on security
-requirements but they are necessarily intertwined with the architectural
-requirements, as we will show afterwards in this section.  A complete study on
-how to properly define the architectural requirements is outside the scope of
-this paper so we propose a method that is convenient for the purpose of our
-assessment.
-
-\Paragraph{System Requirements}
-In our running example, there is a system (a CPS) composed by $5$ main
-subsystem which we consider atomic (i.e. not themself composed by subsystems)
-and then agents: the tank, the sensor, the controller, the actuator, and the
-motorized valve (valve from now on).  The whole specification (which we now
-describe) of the physical and functional architecture is depicted in
-Figure~\ref{fig:spec}.  In the use case, the communication flow is as follows:
-\begin{enumerate}
-	\item the sensor reads the level of water in the tank ($\text{tank}\rightarrow \text{sensor}$) \footnote{this communication depends on the technology in use for the readings.
-We assume a unidirectional communication without loss of generality}
-	\item the sensor communicates the readings to the controller ($\text{sensor}\rightarrow \text{controller}$)
-	\item the controller calculates the state of the valve and communicates it to the actuator ($\text{controller}\rightarrow \text{actuator}$)
-	\item the actuator translates (digital to analog) the state received and communicates\footnote{we do not distinguish, for the sake of simplicity between different types of channels and assume that communicating to the valve produces the expected change of physical state} it to the valve
-\end{enumerate}
-Each agent has an input or output port from which it communicates with other
-agents.  It is also composed by the functional blocks that defines its
-functional architecture and physical boundaries that defines its physical
-architecture, as follows.
 \begin{enumerate}[start=1, label={req\arabic*)}]
@@
 \end{enumerate}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.9\textwidth]{IOports.pdf}
-	\caption{Input and Output Port of an agent}
-	\label{fig:ioports}
-\end{figure}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.9\textwidth]{channel.pdf}
-	\caption{Communication over a Mono-directional Channel}
-	\label{fig:channel}
-\end{figure}
-
-While \emph{agents} and \emph{communication} between agents are described and
-defined in Section~\ref{sec:theory}, the concepts of \emph{port} (depicted in
-Figure~\ref{fig:ioports}), \emph{channel} (depicted in
-Figure~\ref{fig:channel}), \emph{functional and physical architecture}, and
-\emph{functional block} have not yet been defined with respect to the
-$\abftheory$-theory.  Informally, a channel\footnote{for now, we rely on the
-intuitive notion of channel as ``something'' that transfers information as
-Assertions between output and input port, i.e. mono-directional} identifies the
-place where the communication takes place.  In this work, we only consider
-mono-directional channels and communication but the extension to bi-directional
-channel can be considered as the union of two unidirectional channels. A
-mono-directional channel is defined by the assertions sent or received (over
-the channel).
-
-\subsubsection{Input and Output Ports}
-Since the $\abftheory$-theory is a theory of agents, we consider ports as agents that
-allows the exchange of information between a channel and another
-agent.  However, a port must be considered as a special type of agent to avoid
-an \emph{infinite regress} in which a port needs a port to transfer information
-between the outside of the port to the inside of itself.
-
-\begin{definition}{\bf Input or Output Port --}\label{def:port} 
-	Ports are agents with the following predefined behavior: an input-port changes
-	the type of information from assertion from a sender $s$ to an agent $a$,
-	$\rassert{s}{a}$ to belief of the agent $a$ ($\beliefRegion_a$) while an
-	output-port from belief to assertion; forwarding information from the outside
-	of an agent's boundary to the inside (input-port) or vice versa (output-port). 
-\end{definition}
-The \emph{quality of a port} is determined by the $\rcc$ relation between the 
-assertions received or sent and the belief, i.e. $\Rcc{\assertionRegion_s}{\beliefRegion_a}$ for the input-port
-or $\Rcc{\assertionRegion_r}{\beliefRegion_a}$. We only define secure input port but the
-definition of a secure output port is symmetrical.
-
-\begin{definition}{\bf Secure Input Port --}\label{def:secport}
-	A secure input port ($s\secureport a$) allows information as incoming assertions to flow
-	from a sender $s$ to the behavior of the recipient $a$ (agent).  If $p_I$ is a secure input port,
-	$\world\models\varphi$, $\varphi\in\rassert{s}{a}$, and $a$
-	has the only input port $p_I$, $\world'\models\beliefRegion_a$ and $\varphi\in\beliefRegion_a$ for any
-	$\world\modalrelation\world'$.
-\end{definition}
-
-\Paragraph{Port Weaknesses}
-From this definition, it follows that there exist only the following \emph{six} types of weaknesses, generating six types of insecure port in RCC5 (the notation is reported for input-ports, i.e.
-on the lef-hand side of the arrow):
-\begin{enumerate}[start=1, label={W\arabic*)}]
-	\item \emph{drop port} ($s\dropport a$), where assertions reaches the port but do not pass the boundary of the agent (i.e. do not become belief of the agent)
-	\item \emph{insertion port} ($s\insertport a$), where some new information is believed by $a$ as incoming from the port but $s$ didn't send it
-	\item \emph{injection port} ($s\injectionport a$), where information coming from $s$ is substituted with new information which becomes believed by $a$
-	\item \emph{selective port}, where some information passes the port and part is either:
-	\begin{enumerate}[start=1, label={W4.\arabic*)}]
-		\item dropped ($s\selectivedropport a$), 
-		\item inserted ($s\selectiveinsertport a$), or
-		\item injected ($s\selectiveinjectionport a$)
-\end{enumerate}
-\end{enumerate}
-
-\begin{proof}
-An input port is, in the $\abftheory$-theory, defined secure as long as the relation
-	between the two regions of input assertions $\assertionRegion$ and
-	output beliefs $\beliefRegion$ are equal, i.e.
-	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
-	should result in a weakness (related to an insecurity flaw) of that
-	input port.  Using RCC5, there exist exactly other $4$ different type
-	of relations, one of which is the discrete-from (DR) relation, i.e.
-	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
-	the DR relations, they have no subregion in common. Lets
-	define a function weight $|X|$ such as, for any region $X$, it
-	represents the smallest possible cardinality of a (mereo)topological base for
-	$X$; where a base is a collection of regions in a (mereo)topology such that
-	every open region can be written as union of elements of that base.  If
-	a communication occurred and resulted in
-	$\dr{\assertionRegion}{\beliefRegion}$, there exist only two mutually
-	exclusive options: either $|\assertionRegion|=|\beliefRegion|$ or
-	$|\assertionRegion|\neq|\beliefRegion|$. 
-	\begin{itemize}
-		\item If $|\assertionRegion|\neq|\beliefRegion|$ (at the end of
-			a communication through the input-port), the two
-			regions have a different number of subregions, and,
-			then, there exist only two mutually exclusive options: 
-			\begin{itemize}
-				\item either $|\assertionRegion|>|\beliefRegion|$, where one
-					ore more Assertions didn't become Beliefs, which we
-					call \emph{drop port} ($s\dropport a$) since its behavior drops and
-					prevents incoming communications,
-				\item or $|\assertionRegion|<|\beliefRegion|$, where there are
-					one ore more Beliefs which have not been sent as
-					input-Assertions, which we call \emph{insertion
-					port} ($s\insertport a$) since its behavior generates new Beliefs
-					unrelated to the incoming communication. 
-			\end{itemize}
-		\item If $|\assertionRegion|=|\beliefRegion|$, supposing an increasing monotonic entropy on the information (Assertions) exchanged, 
-				there exist only two mutually exclusive options: 
-			\begin{itemize}
-				\item either information has been generated as Belief and transferred as Assertions, increasing the cardinality of both
-					$\assertionRegion$ and $\beliefRegion$, i.e.
-					$|\assertionRegion_{t^0}|>|\assertionRegion_{t^1}|$ and
-					$|\beliefRegion_{t^0}|>|\beliefRegion_{t^1}|$ (where
-					$t^0$ and $t^1$ represent time at the beginning of
-					communication and right after, respectively); which
-					means that Assertions have reached the port and new
-					Beliefs have been generated by the port, but no
-					correlation between the elements of the two regions exist. 
-					We call this \emph{injection port} ($s\injectionport a$) since it has modified
-					the information carried by the Assertions into new
-					corresponding Beliefs.
-				\item or $|\assertionRegion_{t^0}|=|\assertionRegion_{t^1}|$
-					and $|\beliefRegion_{t^0}|=|\beliefRegion_{t^1}|$ but
-					this implies that no communication occurred, while we
-					assumed that a communication was taking place (absurd).
-			\end{itemize}
-	\end{itemize}
-	In other words, $\dr{\assertionRegion}{\beliefRegion}\equiv\neg\overlaps{\assertionRegion}{\beliefRegion}\equiv\neg(\varphi\subseteq\assertionRegion\wedge\varphi\subseteq\beliefRegion)\equiv\varphi\not\subseteq\assertionRegion\vee\varphi\not\subseteq\beliefRegion$
-	Similarly, 
-	\begin{itemize}
-		\item given that $\po{\assertionRegion}{\beliefRegion}$ is equivalent to
-			$\dr{\assertionRegion}{\beliefRegion}$ except for the
-			overlapping part (see Appendix~\ref{app:selectiveinjectionport}) in which case is
-			$\eq{\assertionRegion}{\beliefRegion}$ (see
-			\autocite{Santaca2016abf}),
-			we call it a \emph{selective injection port} ($s\selectiveinjectionport a$); otherwise
-		\item $\pp{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Leftarrow\beliefRegion$ and
-			$\ppi{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Rightarrow\beliefRegion$ are subcases of
-			$\eq{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Leftrightarrow\beliefRegion$, and then
-		\begin{itemize}
-			\item if $\pp{\assertionRegion}{\beliefRegion}$, we have that
-				some Beliefs are not correlated to incoming Assertions
-				and we call it \emph{selective insertion port} ($s\selectiveinsertport a$),
-			\item if $\ppi{\assertionRegion}{\beliefRegion}$ we have that
-				some incoming Assertions are not correlated to Belief 
-				and we call it \emph{selective drop port} ($s\selectivedropport a$)
-		\end{itemize}
-	\end{itemize}
-\end{proof}
-
-\subsubsection{Communication Channels}
-We start by considering the difference between a (communication)
-mono-directional channel (channel from now on) and an agent, as we did for the
-ports, since the $\abftheory$-theory is a theory of agents.  In fact, if a channel
-were considered an agent (channel-agent) then the question would be how an
-agent would transfer its Assertions to the channel-agent. If the channel
-between the agent and the channel-agent is again an agent, we would generate an
-\emph{infinite regress}. Therefore, we do allow channel-agents but we assume a
-finite depth (of detail) for a channel, where there exists a bottom-channel
-which is not an agent. For now, we do not constrain a channel-agent in any way
-so there is no difference between a channel agent and agent. Therefore, for the
-sake of simplicity, we consider, in this text, channels to be bottom-channels,
-defined as agents with the pre-defined behavior (i.e. defined in a dogmatic
-way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
-prevents us from introducing additional constraints to the channel as storing
-Assertions that are transferred over the channel, or filter out some
-input-assertions.}.
-
-\begin{definition}{\bf Secure Mono-directional Channel (bottom-channel) --}\label{def:monochannel}
-	A mono-directional channel between the two agents ($s \rightarrow r$) is defined as 
-	an agent which behavior is (dogmatically) defined as: to forward any Assertion received from $s$ over an input-port, to  
-	the output-port where $r$ is listening to.
-\end{definition}
-The \emph{quality of a mono-directional} channel is defined as the 
-$\rcc$ relation between the Assertions made by the sender and the ones received by the receiver, i.e. $\Rcc{\assertionRegion_s}{\assertionRegion_r}$.
-
-\Paragraph{Channel Weaknesses}
-Given that a mono-directional bottom-channel is assumed to be perfectly
-forwarding any Assertion from its input-port to its output-port, there is no
-insecure behavior but only the combination of the weaknesses of the input and
-output port; therefore there exists $(7^2)-2=47$ theoretical configurations
-($7^2$ because there are $6$ insecure types of port, plus $1$ secure type, on both input and oupput
-side; and we exclude the configuration with $2$ secure types as input
-and ouput, $-2$); where only 43 are possible.
-
-\begin{enumerate}[start=5, label={W\arabic*)}]
-	\item \emph{secure output port} and \emph{input drop port} ($s\sdcom r$), 
-	\item \emph{secure output port} and \emph{input insertion port} ($s\sicom r$), 
-	\item \emph{secure output port} and \emph{input injection port} ($s\sjcom r$), 
-	\item \emph{secure output port} and \emph{input selective drop port} ($s\ssdcom r$), 
-	\item \emph{secure output port} and \emph{input selective insertion port} ($s\ssicom r$), 
-	\item \emph{secure output port} and \emph{input selective injection port} ($s\ssjcom r$),
-
-	\item \emph{output drop port} and \emph{input drop port} ($s\ddcom r$)
-	\item \emph{output drop port} and \emph{input insertion port} ($s\dicom r$)
-	\item \emph{output drop port} and \emph{input secure port} ($s\dscom r$)
-
-	\item \emph{output insert port} and \emph{input drop port} ($s\idcom r$)
-	\item \emph{output insert port} and \emph{input insert port} ($s\iicom r$)
-	\item \emph{output insert port} and \emph{input injection port} ($s\ijcom r$)
-	\item \emph{output insert port} and \emph{input selective drop port} ($s\isdcom r$)
-	\item \emph{output insert port} and \emph{input selective insertion port} ($s\isicom r$)
-	\item \emph{output insert port} and \emph{input selective injection port} ($s\isjcom r$)
-	\item \emph{output insert port} and \emph{input secure port} ($s\iscom r$)
-
-	\item \emph{output injection port} and \emph{input drop port} ($s\jdcom r$)
-	\item \emph{output injection port} and \emph{input insertion port} ($s\jicom r$)
-	\item \emph{output injection port} and \emph{input injection port} ($s\jjcom r$)
-	\item \emph{output injection port} and \emph{input selective drop port} ($s\jsdcom r$)
-	\item \emph{output injection port} and \emph{input selective insertion port} ($s\jsicom r$)
-	\item \emph{output injection port} and \emph{input selective injection port} ($s\jsjcom r$)
-	\item \emph{output injection port} and \emph{input secure port} ($s\jscom r$)
-
-	\item \emph{output selective drop port} and \emph{input drop port} ($s\sddcom r$)
-	\item \emph{output selective drop port} and \emph{input insertion port} ($s\sdicom r$)
-	\item \emph{output selective drop port} and \emph{input injection port} ($s\sdjcom r$)
-	\item \emph{output selective drop port} and \emph{input selective drop port} ($s\sdsdcom r$)
-	\item \emph{output selective drop port} and \emph{input selective insertion port} ($s\sdsicom r$)
-	\item \emph{output selective drop port} and \emph{input selective injection port} ($s\sdsjcom r$)
-	\item \emph{output selective drop port} and \emph{input secure port} ($s\sdscom r$)
-
-	\item \emph{output selective insertion port} and \emph{input drop port} ($s\sidcom r$)
-	\item \emph{output selective insertion port} and \emph{input insertion port} ($s\siicom r$)
-	\item \emph{output selective insertion port} and \emph{input injection port} ($s\sijcom r$)
-	\item \emph{output selective insertion port} and \emph{input selective drop port} ($s\sisdcom r$)
-	\item \emph{output selective insertion port} and \emph{input selective insertion port} ($s\sisicom r$)
-	\item \emph{output selective insertion port} and \emph{input selective injection port} ($s\sisjcom r$)
-	\item \emph{output selective insertion port} and \emph{input secure port} ($s\siscom r$)
-
-	\item \emph{output selective injection port} and \emph{input drop port} ($s\sjdcom r$)
-	\item \emph{output selective injection port} and \emph{input insertion port} ($s\sjicom r$)
-	\item \emph{output selective injection port} and \emph{input injection port} ($s\sjjcom r$)
-	\item \emph{output selective injection port} and \emph{input selective drop port} ($s\sjsdcom r$)
-	\item \emph{output selective injection port} and \emph{input selective insertion port} ($s\sjsicom r$)
-	\item \emph{output selective injection port} and \emph{input selective injection port} ($s\sjsjcom r$)
-	\item \emph{output selective injection port} and \emph{input secure port}  ($s\sjcom r$)
-\end{enumerate}
-In Appendix~\ref{app:insecurecom}, we report the proof which is exhaustive on all
-possible combinations.
-
-As depicted in Figure~\ref{fig:atom}, so far, we have considered Information
-generated by a Port $P_I$ and then sent through a channel $C$ to another
-(recipient) port $P_O$. In this scenario where Ports and then Channels are
-atomic (otherwise raising infinite regress), we can only  consider the
-relations between Ports and Channel; considering both input-port to channel and
-channel to output-port. In fact, the Weaknesses of a channel are defined in terms
-of the Weaknesses of Ports. Our intention is to introduce now the concept of
-functional architecture and functional blocks composing the functional architecture.
-Io order to define a functional block without encountering an infinite definition,
-we must reach the same conclusions as for the Channel. Therefore, describing
-the information as flowing over a channel or in a functional block is purely
-syntactic sugar.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.7\textwidth]{engineering_relations.pdf}
-	\caption{Relations between Ports and Channel}
-	\label{fig:atom}
-\end{figure}
-
-\subsubsection{Functional Block}
-A Functional Architecture, as already described beforehand, takes Information as input-Beliefs
-and transforms the Information into output-Beliefs. Those transformations 
-occurs within the Functional Architecture, where Functional Blocks transforms Beliefs into 
-other Beliefs. Similarly to channels, we could consider a Functional Block as a Functional Architecture
-occurring in an infinite regress. Therefore, we consider Functional Blocks as 
-correctly executing their abstract behavior, correctly transforming inputs into outputs (Beliefs).
-
-\begin{definition}{\bf Functional Block and Architecture --}\label{def:funblock}
-	A functional block of an agent $a$ 
-	takes a region of input-Beliefs and outputs a region of output-Beliefs. 
-	A functional architecture is an interconnected system of functional blocks.
-\end{definition}
-The quality of a functional block or architecture can be expressed as a
-relation between input and output belief, similarly to the quality of a channel
-but inverting the flow (input$\rightarrow$output instead of
-output$\rightarrow$input). In other words, we consider two properties:
-how inputs are interpreted, and how outputs are produced.
-One may, of course, argue that since there is a
-transformation process in the functional block the quality of that
-transformation process is what the theory should consider. This can
-be argued not just for functional blocks (Behaviors) but also for Channels
-(Assertions). As we detail afterwards, the relations considered so
-far revolves around the concept of integrity or availability, while the reminder of this section 
-will be focused on a broader analysis of security properties and requirements
-which will also consider other security properties such as availability and
-confidentiality.  In order to define the quality of the internals of a functional block or channel, we first
-need to relate the concept of Facts to engineering requirements.
-
-\Paragraph{Requirements as Facts}
-During the specification phase, for any agent, channel, port, functional block
-and architecture, there may exist a factorial requirement (Fact) predicating over them.
-In other words, any requirement is defined as a Fact since they must be true in
-any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
-Figure~\ref{fig:soundness}, Facts are strategic rules that defines how
-the system shall behave (by specification), while reality may be shown to
-be insecure.
-
-\Paragraph{Methodology}
-We have now mapped the building blocks of our theory, namely Information,
-Belief, Assertion, Behavior, and Fact. One may notice that the concept of
-Knowledge hasn't been discussed in this section so far. Our argument is that
-knowledge, as an epistemological concept cannot be defined due to several
-implication that this would have (see, for example,
-\autocite{Empiricus1990Pyrrhonism}). On the other hand, the factorial way in
-which requirements are imposed may be true only in the specification, while not
-necessarily true in the final implementation. While Knowledge may not be
-apprehensible in general, we can define tests to see if Facts holds in a
-design or implementation. What is of our interest is to guarantee security in a
-system so we now correlate the potential behaviors with the standard (de-facto)
-understanding of security; so that we can identify test that, at the end of the
-engineering process, may not give Knowledge but, at last, tested Fact.
-
-\subsubsection{Security Requirements} 
-While standards such as IEC 62443-1-3 (the Industrial communication networks -
-Network and system security -- Part 3-3: System security requirements and
-security levels) defines requirements as ``confidentiality of information in
-transit/at-rest'', we believe that, at specification level, a mapping
-between the Infosec security concepts of the well-known CIA (Confidentiality,
-Integrity, Availability) triad to the system components can be used to specify
-a list of security requirements.
-
-Infosec, or Information Security, defines the security risks related to
-information as the CIA triad, which has been often criticized
-(as we show afterwards) for being too general or non-adequate (e.g. by
-adding authenticity which is often used as a building block for confidentiality
-and integrity) to be effectively applied to the engineering of systems. Due to
-this, many researchers and organization tried to improve the CIA triad.  The
-evolutions/extensions of the CIA triad has been documented, e.g., in
-\autocite{Samonas2014cia} and are summarized in Table~\ref{tab:ciaevolution}.
-
-\begin{table}[h]
-\centering
-%\setlength{\tabcolsep}{3.5pt}
-%\renewcommand{\arraystretch}{1}
-\small
-\begin{tabular}{rll} 
-	{\bf Year} & {\bf Definition} & {\bf Legend}\\
-	\hline
-	1970s & infosec = CIA & Confidentiality, Integrity, Availability\\
-	1980s & infosec += (Au, nR) & Authenticity and non-Repudiation\\
-	1990s & infosec += CSpec & Correctness in Specification\\
-	2000s & infosec += RITE & Responsibility, Integrity of people, Trust, Ethicality
-\end{tabular}
-\caption{Chronological progression of the CIA triad~\label{tab:ciaevolution}}
-\end{table}
-
-A representative example is the effort made by the OECD (Organisation for Economic
-Co-operation and Development) in \autocite{OECD2013guidelines} to define
-security guidelines for information system (1992) based on new principles such
-as ``awareness, ethics, risk assessment'' and maintain those improving the document with
-subsequent revisions 
-(e.g.  following the multistakeholder expert consultation in 2013).
-Other similar efforts focused on defining entirely new principles or extending
-the CIA triad, such as \autocite{NISTSP800-160}. 
-%What is of interest for our
-%argument is that, to the best of our knowledge, all of the approaches that aim
-%at improving the CIA triad are besed on motivations related to empirical evidences. On the contrary,
-%We now define (in Section~\ref{sec:properties}) the CIA triad for an agent
-%defined in the $\abftheory$-theory and we test if (i) other properties are
-%allowed in the $\abftheory$-theory, and (ii) if and how the CIA triad can be
-%detailed in the $\abftheory$-theory.
-In \autocite{Anderson1972report,Samonas2014cia}, CIA are defined as follows.
-\begin{itemize}
-	\item \emph{Unauthorized information release}: an unauthorized person is able
-		to read and take  advantage  of  information  stored  in  the
-		computer.  This  category of concern  sometimes  extends  to
-		``traffic  analysis,''  in  which  the intruder  only observes
-		the  patterns  of  information  use.  From  those patterns,
-		the  intruder can   infer   some   information   content.
-		This category   also   includes   the unauthorized use of a
-		proprietary program.  (Confidentiality or Secrecy) 
-	\item \emph{Unauthorized  information  modification}:  an  unauthorized person
-		is  able  to make changes in stored information -- a form  of
-		sabotage.  It should be noted that in the case of this kind of
-		violation, the intruder does not necessarily see the
-		information he has changed.  (Integrity)
-	\item \emph{Unauthorized  denial  of  use}:  an  intruder  can  prevent  an
-		authorized  user  from referring to, or from modifying
-		information, even though the intruder may not be able to refer
-		to, neither modify the information themselves. (Availability)
-\end{itemize}
-
-We now want to generalize the CIA triad to agents and systems in the
-$\abftheory$-theory (and not just Information). Given that we are currently
-dealing with the step that moves from specification to design, we consider
-architectural properties. For the sake of simplicity, we revert the order and
-start from availability.
-
-\Paragraph{Availability and Integrity (Architecture Security)}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\textwidth]{client-server.pdf}
-	\caption{Client-Server paradigm in $\abftheory$-theory}
-	\label{fig:client-server}
-\end{figure}
-Availability, again, is the denial of use of a resource. In the $\abftheory$-theory, 
-in order to prevent a resource to be available, there
-exists only (i.e. inclusively) the following cases
-that we illustrate using a client-server paradigm (depicted in Figure~\ref{fig:client-server}).
-\begin{enumerate}
-	\item\label{case:srv1} The resource on the server cannot be found, so that it is
-		possible for the client to contact the server and for the server
-		to send a response back, but the output-port does not receive the
-		necessary information. This may be due to a software error such
-		that the functional architecture doesn't output a response, or
-		a problem in the flow of information (i.e. the functional
-		architecture doesn't receive the necessary information from
-		other resources), or an hardware failure \&c. This can be formalized
-		as an issue to be found in the relation between the beliefs produced by the functional
-		architecture and the ones received by the output port, i.e.
-		$\rcc(\behaviorRegion_s,\{\text{res}_0\})$ or 
-		$\rcc(\beliefRegion_s,\{\text{res}_0\})$ 
-	\item\label{case:srv2} The server hosting the resource is down, and then the server
-		cannot output responses (i.e. insecure output port due to drop-related issue) or receive
-		requests (insecure input port due to drop-related issue) or both.
-	\item\label{case:srv3} The client cannot properly compute request (symmetric to Case~\ref{case:srv1})
-	\item The client is down, symmetric to Case~\ref{case:srv2}
-	\item The channel is down due to a drop related issue in one of the two channels 
-\end{enumerate}
-%All the cases but Case~\ref{case:srv1} and Case~\ref{case:srv3}, tell us how to map the
-%property of availability over the weaknesses of ports and communications.  The
-%Cases~\ref{case:srv1} and~\ref{case:srv3}, are exceptions because the formulas $res$ and $req$ 
-%detail the specification with a singleton instead of a region of formulas (or, better, $|res|$ is equal to the one of the
-%base).
-%In other words, a channel, a port, an entity, or a functional block are
-%Fact (stated in a factorial way) and not Beliefs; what they produce or 
-%their non-specified internals are Beliefs.  In fact, we never assume an
-%entity not to be in the design implementing the specification.  Similarly, when a
-%functional block produces some information it produces a Belief, unless this Belief if detailed in the specification.
-%For example, the server produces the Belief $res$ but the Belief given that can only be in $EQ$
-%relation with the Fact $res$ (in contrast other Belief may not); therefore, $res$ is not really a belief but a Fact. 
-%$res$ as a Fact is something that
-%exists in the specification, while the Belief of $res$ in one of the possible
-%designs may not exist. 
-We shall, then, distinguish between two cases, as we mentioned beforehand:
-agent (or system) and Information (which we call Infosec) availability. 
-\begin{definition}{\bf Agent-Availability --}\label{def:agent-a}
-when the property of availability is expressed over an agent or a system, the
-	list of possible designs do not include those in which there exist:
-	drop ports, or drop channels/functional-blocks.
-\end{definition}
-\begin{definition}{\bf Infosec-Availability --}\label{def:infosec-c}
-when the property of availability is expressed
-	over a formula, the list of possible designs that violates the
-	availability property is reduced to the part of the system that
-	does not influence that specific formula.
-\end{definition}
-
-In a similar way we can define integrity since this requirement may 
-be expressed over an agent/system or a specific formula (i.e. element of a Region).
-
-\begin{definition}{\bf Agent-Integrity --}\label{def:agent-a}
-when the property of integrity is expressed over
-	an agent or a system, the list of possible designs 
-	do not include those in which there exist: insert or inject ports, 
-	or inert or inject channels/functional-blocks. 
-\end{definition}
-\begin{definition}{\bf Infosec-Integrity --}\label{def:infosec-c}
-when the property of integrity is expressed
-	over a formula, the list of possible designs that violates the
-	integrity property is reduced to the part of the system that
-	influences that specific variable.
-\end{definition}
-
-While availability and integrity are architectural property, meaning that they
-can be expressed over the behavior of a port or a channel (as constituent of
-the physical or functional architecture), confidentiality is not.
-Confidentiality states that an information shall not be understood and not that
-the information shall not be retrieved. So, if a port leaks an entire message,
-there is no confidentiality issue in general, unless that message was not
-encrypted at all, in which case the confidentiality leakage is straightforward. 
-On the other hand, if the
-information is not provided or altered there is an availability or integrity
-issue. A similar argument can be proposed for integrity sub-properties such as
-authenticity or non-repudiation. No wonder why it is often stated that
-availability and integrity are the first properties to be considered in CPS
-while confidentiality takes the first place when dealing with security
-protocols. In fact, integrity is often assumed by pattern-matching (e.g. in
-AVANTSSAR \autocite{Armando2012avantssar}) and availability is hardly ever considered in
-protocol verification. The question is, can we express properties such as
-confidentiality, authentication, or non-repudiation as architectural properties
-in $\abftheory$-theory? Confidentiality is based on the concept 
-of secrecy so we shall first start from describing what secrecy is and how it
-can be used to guarantee confidentiality in a system.
-
-\Paragraph{Secrecy, Authenticity, and Key Agreement (Cryptographic Security)}\label{sec:encryption}
-As Maurer puts it (\autocite{Maurer1999information}), there exist two types of
-cryptographic security (i.e. scientific theories aiming at guaranteeing
-security through the use of cryptographic functions): ``The security  of a
-cryptographic system can rely either on the computational infeasibility of
-breaking it (computational security), or on the theoretical impossibility of
-breaking it, even using in infinite computing power (information-theoretic or
-unconditional security)''. We are not currently interested in detailing the
-difference between the two and we simplify it as follows. A
-computationally-secure cryptographic function (usually called encryption
-function) $E(\varphi)=e$ (where $\varphi$ is a plain-text and $e$ a
-cipher-text) can be considered perfectly secure if a malicious agent that
-obtains the result of the cryptographic function ($e$) has bounded
-computational power, while unconditional secure cryptographic functions are
-considered perfectly secure. In this context, security means that is is
-impossible to calculate $\varphi$ from $e$, therefore impossible to compute
-$D(e)=\varphi$.  As one can evince from \autocite{Shannon1948information}, the
-commonality between the two types of cryptographic functions is that they both
-rely on the assumption about the probabilistic behavior of the universe. They
-do so, for the random (i.e. with high entropy
-\autocite{Shannon1948information}) generation of the so called \emph{key}, a
-piece of information, parametric to the cryptographic function, used to
-guarantee the security. Therefore, we can assume that it is feasible for an
-agent that knows $k$ to decrypt the cipher-text obtaining the clear-text (i.e. $D(k, E(k,\varphi))=\varphi$)
-while it is impossible for any agent that has not the proper key. In this
-case, the information is considered to be \emph{secret or confidential}.
-
-\begin{definition}{\bf Infosec-Secrecy or Confidentiality --}\label{def:confidentiality}
-	When an Assertion has been generated as an output by an
-	agent, port, or functional block such that it is said (in a
-	factorial way) to encrypt the input, that Assertion is considered
-	confidential.
-\end{definition}
-The quality of the confidentiality of information (Assertions or Beliefs depending on 
-the subsystem considered) is inversely proportional to the number of 
-integrity issues of the subsystem.
-We now discuss more in details what secrecy of Information is when 
-dealing with functional blocks (Behavior) and Beliefs but a similar treatment can be done 
-in the case of Channels and Assertions.
-
-The quality of a Functional Block is determined by the relation between the
-input-Beliefs formulas and the symmetric formulas in the Fact region
-(equivalently for output-Beliefs).
-
-\begin{definition}{\bf Secrecy of Functional Block and Architecture --}\label{def:secfun}
-	Secrecy is preserved in a functional block if the input beliefs and output beliefs
-	are in $EQ$ relation with the corresponding Facts mandated by the confidentiality/secrecy requirements.
-	A functional architecture composed exclusively by confidential/secrecy-preserving functional blocks
-	is considered confidential/secrecy-preserving.
-\end{definition}
-
-\Paragraph{Confidentiality Weaknesses}
-In a functional block, there exists the following weaknesses in the functional
-architecture, where $\beliefRegion_I$ and $\beliefRegion_O$ are input and
-output Beliefs respectively, and $\factRegion_I$ and $\factRegion_O$ are
-the corresponding Facts. 
-\begin{itemize}
-	\item $\eq{\beliefRegion_I}{\factRegion_I}$ and then only specified
-		inputs can be fed to the functional block but:
-	\begin{enumerate}[start=49, label={W\arabic*)}]
-		\item\label{w:fb_eqdr} $\dr{\beliefRegion_O}{\factRegion_O}$, not all the specified
-			outputs are produced. This may produce a secrecy issue if, for example,
-			the functional block is responsible for generating an encryption key since
-			if it is not produced the encryption of the Information may not be performed
-		\item\label{w:fb_eqpo} $\po{\beliefRegion_O}{\factRegion_O}$, only some but not all of the specified
-			outputs are produced along with some non-specified output are produced. This
-			strightforwardly generates a secrecy when the functionality is 
-			related to secrecy functional blocks since some outputs doesn't follow
-			the requirements.
-		\item $\pp{\beliefRegion_O}{\factRegion_O}$, not all the specified outputs are produced.
-			This may pose a secrecy issue as in \ref{w:fb_eqdr}
-		\item $\ppi{\beliefRegion_O}{\factRegion_O}$, not all the output have been specified.
-			This may pose a secrecy issue as in \ref{w:fb_eqpo}
-	\end{enumerate}
-	\item $\po{\beliefRegion_I}{\factRegion_I}$, not all the specified inputs can be fed into the functional block
-	\begin{enumerate}[start=53, label={W\arabic*)}]
-		\item $\dr{\beliefRegion_O}{\factRegion_O}$
-		\item $\po{\beliefRegion_O}{\factRegion_O}$
-		\item $\pp{\beliefRegion_O}{\factRegion_O}$
-		\item $\ppi{\beliefRegion_O}{\factRegion_O}$
-		\item $\eq{\beliefRegion_O}{\factRegion_O}$ this case implies
-			that somehow there is no input that generates an
-			non-specified output but this is not because the inputs
-			are secure 
-	\end{enumerate}
-	\item $\pp{\beliefRegion_I}{\factRegion_I}$, not all the specified
-		inputs can be fed into the block
-	\begin{enumerate}[start=58, label={W\arabic*)}]
-		\item $\dr{\beliefRegion_O}{\factRegion_O}$
-		\item $\po{\beliefRegion_O}{\factRegion_O}$
-		\item $\pp{\beliefRegion_O}{\factRegion_O}$
-		\item $\ppi{\beliefRegion_O}{\factRegion_O}$
-		\item $\eq{\beliefRegion_O}{\factRegion_O}$
-	\end{enumerate}
-	\item $\ppi{\beliefRegion_I}{\factRegion_I}$, not all the inputs that
-		can be fed into the functional block follows the specification
-	\begin{enumerate}[start=63, label={W\arabic*)}]
-		\item $\dr{\beliefRegion_O}{\factRegion_O}$
-		\item $\po{\beliefRegion_O}{\factRegion_O}$
-		\item $\pp{\beliefRegion_O}{\factRegion_O}$
-		\item $\ppi{\beliefRegion_O}{\factRegion_O}$
-		\item $\eq{\beliefRegion_O}{\factRegion_O}$
-	\end{enumerate}
-	\item $\dr{\beliefRegion_I}{\factRegion_I}$, no specified inputs can be
-		fed into the functional block
-	\begin{enumerate}[start=68, label={W\arabic*)}]
-		\item $\dr{\beliefRegion_O}{\factRegion_O}$
-		\item $\po{\beliefRegion_O}{\factRegion_O}$
-		\item $\pp{\beliefRegion_O}{\factRegion_O}$
-		\item $\ppi{\beliefRegion_O}{\factRegion_O}$
-		\item $\eq{\beliefRegion_O}{\factRegion_O}$
-	\end{enumerate}
-\end{itemize}
-
-\begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
-	A system security design is given by a precise system security
-	specification over the physical and functional architectures, that
-	uniquely, i.e.  tested against the range of design possible in the
-	$\abftheory$-theory, defines the designs built on top of those
-	requirements.
-\end{hypothesis}
-
-\begin{hypothesis}{\bf System Insecurity Design --}\label{hyp:insecurity}
-	If, given a system security specification as a collection of
-	requirements on the system, there exist
-	a non-unique design that respect those requirements, the number of
-	equivalent designs that fullfills the requirements quantitatively
-	defines the magnitute of insecurity of a system design.
-\end{hypothesis}
-
-\subsection{Abstraction Level of a Specification in $\abftheory$-theory}
-The level of abstraction is always a concern when it comes to the definition of
-a system model.  In this paper, we justified our assumptions on the atomicity
-of ports, channels, and functional blocks but nothing prevents a modeler to
-consider other atomic element of the system specification and design.  Those
-additional elements may detail the system and may change the syntactic sugar we
-defined; however, it won't change the underlying (semantic) reasoning unless
-another calculus is used (instead of RCC).  Furthermore, other Infosec
-properties may be considered, such as authenticity or non-repudiation. We,
-however, leave them as future development.

### 2020-06-01 reports/white_paper/compile.sh

@@ -1,0 +1,27 @@
+#!/bin/bash
+latexOutputDir="latex-output-dir"
+latexInputFile="main.tex"
+pdfOutputDir="${latexOutputDir}/${latexInputFile%.*}.pdf"
+
+if [[ ! -d $latexOutputDir ]]; then 
+	mkdir $latexOutputDir
+fi
+
+if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
+then 
+	if [ "$1" != "-short" ] && [ "$1" != "-s" ] 
+	then 
+		if biber --nodieonerror --input-directory=$latexOutputDir --output-directory=$latexOutputDir ${latexInputFile%.*}  
+		then
+			if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
+			then 
+				pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile 
+			fi
+		fi
+	fi
+
+	if [[ -f $pdfOutputDir ]]
+	then
+		mv $pdfOutputDir `dirname $0`
+	fi
+fi

### 2020-06-05 reports/white_paper/formal_glossary.tex
minor changes

--- 2020-05-25 reports/journal_paper/formal_glossary.tex
@@
 
-\Paragraph{Attacker as a Dis-honest Agent}
+\paragraph{Attacker as a Dis-honest Agent}
 As we mentioned in Section~\ref{sec:problem} when we postulated the

### 2020-06-05 reports/white_paper/theory.tex
minor changes

--- 2020-05-29 reports/journal_paper/theory.tex
@@
 
-\Paragraph{Methodology}
+\paragraph{Methodology}
 The concept of Knowledge \autocite{Hintikka1962knowledge} (from an
@@
 
-\Paragraph{Number of different configurations of a system}
+\paragraph{Number of different configurations of a system}
 The general formula to calculate the number of different types of agents is
@@
 
-\Paragraph{\Rcc{$\assertionRegion$}{$\behaviorRegion$} -- Collaboration}
+\paragraph{\Rcc{$\assertionRegion$}{$\behaviorRegion$} -- Collaboration}
 In order to reason on the relation between assertions and behavior we first
@@
 
-\Paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and \Rcc{$\behaviorRegion$}{$\factRegion$} -- Honesty and Competence}
+\paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and \Rcc{$\behaviorRegion$}{$\factRegion$} -- Honesty and Competence}
 While Assertions and Beliefs generated by the Behavior determines how a system should work, and the

### 2020-06-05 reports/conference_paper/main.tex
conference paper (intro + literature review)

\documentclass[conference]{IEEEtran}
\usepackage{xcolor}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[style=numeric]{biblatex}
\addbibresource{../literature.bib}% Syntax for version >= 1.2
\usepackage{graphicx}
\graphicspath{{../images/}}

\newcommand{\fixnote}[2]{\textbf{\color{red}{FIX}}\footnote{{\bf #1:} #2}}
\newcommand{\fix}[2]{{\color{red} {\bf tofix:} #2}}
% \renewcommand{\fixnote}[2]{}
% \renewcommand{\fix}[2]{}

\begin{document}
\title{On Etiology of Cybersecurity}

\author{\IEEEauthorblockN{Marco Rocchetto}
\IEEEauthorblockA{V-Research\\
Verona, Italy\\
Email: marco@v-research.it}
\and
\IEEEauthorblockN{Francesco Beltramini}
\IEEEauthorblockA{V-Research\\
Verona, Italy\\
Email: francesco@v-research.it}
}

\maketitle

\begin{abstract}
The abstract goes here.
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}\label{sec:intro}
In \cite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
``an asymmetry in computer security'', which he defines as follows: ``Things
can be declared insecure by observation, but not the reverse. There is no
observation that allows us to declare an arbitrary system or technique
secure''. Herley then uses this argument to show that ``claims that any measure
is necessary for security are empirically unfalsifiable''. Given that, any
theory which is not falsifiable by an empirical experiment is well
known\footnote{``A theory which is not refutable by any conceivable event is
nonscientific. Irrefutability is not a virtue of a theory (as people often
think) but a vice.'' -- Karl Popper, Conjectures and
Refutations \cite{popper1962conjectures}} to be nonscientific (i.e.
unfalsifiability is a fallacy of a theory), Herley concludes that there is no
scientific theory on cybersecurity; which means that cybersecurity lays in the
realm of pseudo-sciences \cite{Herley2016usenixvideo}.  Herley, e.g.
in \cite{Herley2017justifying}, discusses the implications of a
nonscientific approach to cybersecurity, and highlights the tremendous impact on
all the scientific research and engineering of systems; leading often to
terrorism and wars, and wasting of resources in useless protections or
overspending.  While the criticism is investigated
in \cite{Herley2016unfalsifiability}, no solution is provided nor envisioned.  On the
contrary, the goal of this work is to lay the foundations of a
scientific cybersecurity theory.

We consider the problem raised by Herley not confined to
``computer security'' but to any abstract system (so that our theory may hold
for any sound implementation such as networks, mechanical, cyber, or
cyber-physical system, or even a single computer or a single device such as an
hard-drive).  There is also an apparent inconsistency in
\cite{Herley2016unfalsifiability} that we seek to clarify before following (as
we agree) the scientific path draw by Herley: cybersecurity is defined as an
abstract property in many formal approaches to the investigation of the
security of systems, and the security of the design of a formally verified
protocol is indeed falsifiable (against the security properties verified).  For
example, in the protocol verification community, security is often defined as a
formalization of the high-level properties confidentiality, integrity, and
availability. The problem in such approaches is not the definition of what
cybersecurity is but the generality of the results, since they tackle a specific
step (and not the first) of the engineering process (of systems).  Therefore,
the theories underlying the verification are based on assumptions which
non-evidently apply to a general security theory.  As an example, the so called
Dolev-Yao attacker model\footnote{For the sake of simplicity, the Dolev-Yao
attacker can be considered as an abstraction of an active attacker who controls
the network but cannot break cryptography.} \cite{Dolev1983security} only
applies to specific instances (often called scenarios) and abstraction of the
protocol. This, in turn, creates a false sense of security since requires
non-justifiable assumptions on the abstraction of the system of which security
is verified. More specifically, for the formal security verification of a
system in which two or more agents are communicating, a formalized scenario
needs to be defined by a modeler who chooses (among others): (i) a scope of the
formalization (e.g.  excluding the server that distributes the public key is
often done when verifying the security of authentication protocols), (ii) the
number of sessions (even tough some approaches do reason on an infinite number
of sessions such as \cite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
the peers (e.g.  in the ASLan++ language \cite{Oheimb2010aslan++}), and (iv)
the abstraction of the cryptographic primitives (e.g.  ProVerif vs CryptoVerif
\cite{Blanchet2017symbolic}).  While many tools and theories improves those 
issues, there is no agreement on which should be the definitive
approach (if any). More importantly, some of the choices made by the
modelers/engineers using those formal approaches may completely change the
results of the formal verification of the system (an interesting example on how
difficult it is to even just compare the approaches is given in
\cite{Cremers2009comparing}). 
\begin{figure}[t]
	\centering
	\includegraphics[width=.7\columnwidth]{protocol-example.pdf}
	\caption{Abstraction of an ad-hoc esemplificative protocol execution}
	\label{fig:protocol-example}
\end{figure}
For example, in Figure~\ref{fig:protocol-example}, under the perfect
cryptography assumption\footnote{As defined in \cite{Rocchetto2016cpdy}: ``In
the so called perfect cryptography assumption, the security encryption scheme
is suppose to be perfect, without any exploitable flaw, and so the only way for
the attacker to decrypt a message is by using the proper key. That assumption
is widely accepted in the security protocol community, and most of the formal
reasoning tools for the analysis of security protocols abstract away the
mathematical and implementation details of the encryption scheme
\cite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
and assuming that no violation to any security property is done after message
I), the freedom of choosing the scope
determines that the flaws related to the dishonest impersonation of the Server
may or may not be considered in the verification process.  This choice has
tremendous impact on the focus and findings of the verification of the security
of the protocol.  While this may seem to turn upon minutiae and foreseeable,
this highlights the false sense of security that may derive from a
non-falsifiable theory of system security\footnote{``To the superficial
observer, the analysis of these forms seems to turn upon minutiae. It does in
fact deal with minutiae, but they are of the same order as those dealt with in
microscopic anatomy.'' -- Karl Marx, Capital Volume 1, 1867}.

\paragraph{Structure} We start, in
Section~\ref{sec:literature}, by detailing the problem statement, reporting a
literature review on the main concepts and definitions related to security.  We
formulate a security hypothesis in Section~\ref{sec:hypothesis}; which we use
to propose a theory on system security in Section~\ref{sec:theory}. In
Section~\ref{sec:secra}, we apply our theory for the tool-assisted security
risk assessment of a CPS (with an ad-hoc example, based on the SWaT testbed
\cite{Mathur2016swat}).  This application shows how our theory can be used to
predict all of the possible security weaknesses of a system, allowing the
falsification of our theory.  In fact, if any security weaknesses were to be
found in a system and not predicted by our theory, the theory could be declared
incomplete.  Similarly, if a security weakness would be predicted by our theory
but found to be impossible to have our theory could be declared as wrong.

\section{Literature Review}\label{sec:literature}
\begin{figure*}[t]
	\centering
	\includegraphics[width=\textwidth]{safety-security_3.pdf}
	\caption{Overview of keywords related to security and safety}
	\label{fig:safety-security}
\end{figure*}

\fixnote{mr}{add weakness-error in picture and text}
In most of the natural languages the concepts of safety
and security are not syntactically differentiated and both terms (safety and
security) are expressed by the same word, e.g. sicurezza in Italian.  A
semantic distinction between safety and security is correlated to a
belief\footnote{A belief has to be intended as a proposition which is supposed
to be true by the majority of people in our society, without a scientific
underlying theory but based on partial empirical evidences or inductive reasoning based on partial empirical evidences.} that
safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
natural environment (e.g. natural events such as wearing of hardware
components) while security deals with \emph{incidents} posed by mankind (e.g.
attackers and bugs).  The fundamental difference between nature and mankind (and,
in turn, between safety and cybersecurity) is believed to be on the different
intents\footnote{``The belief–desire–intention software model (BDI) is a
software model developed for programming intelligent
agents.''\cite{wiki-bdi}. In the BDI model, the intents represents the
deliberative state of an agent which determines the choice of that agent on
what to do.} (accidents are unfortunate while incidents are not) of the causes
that generates a threat; namely, nature is believed not to have malicious
intents (but unfortunate causes-effects) while threats generated by mankind are
malicious\footnote{Of course, logical flaws or bugs may be
introduced by other means (e.g. ignorance) without explicit malicious intents,
but the exploitation of those flaws is considered (for now, and detailed
afterwards in the article) malicious, and then we consider any vulnerability to
be malicious even if due to the lack of skills.}.
This conclusion seems to be against a general formulation of a cybersecurity theory;
how can we define a theory that predicts what a human being will do?
The cybersecurity attacks seems to be related to the creativity of the attacker
and then unpredictable. Currently, our understanding of cybersecurity is
stored into a network of databases of weaknesses (e.g.
CWE\cite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE\cite{CVE},
NVD\cite{NIST2020NVD}),
and attacks (e.g. CAPEC\cite{MITRE2020CAPEC}, ATTACK\cite{attack}). Systems are
nowadays tested against 
known attacks ignoring the fact that the considering a system secure
if no known attack is possible, defines an unfalsifiable claim.
As an example, if we have a system which is secure against all of the
vulnerabilities in the CVE we just need to wait for a month to have 1000+ more
vulnerabilities to test \cite{newVulns}. If we suppose that one of those new vulnerabilities
affects our system, that should be the falsification of the hypothesis that a
system is secure if no known attacks apply. So, even tough the field 
of security protocols is still lacking of a general cybersecurity theory
the secure-by-design approach they apply can be considered as going
into the opposite direction of the ``implement and test'' approach.

Most of the safety-preserving principles in the field of engineering of
safety-critical cyber-physical systems (such as elevators and aircraft), upon
which safety requirements are defined (e.g. in standards such as the IEC 61508
or 61511\cite{IEC201761511}), are based on empirical tests and measurements
(therefore should be considered hypothesis and not definitions). While
reasoning by induction\footnote{``So, whenever they argue ``Every man is an
animal and Socrates is a man; therefore Socrates is an animal,'' proposing to
deduce from the universal proposition ``every man is an animal'' the particular
proposition ``Socrates therefore is an animal,'' which in fact goes (as we have
mentioned) to establish by way of induction the universal proposition, the
fall into the error of circular reasoning, since they are establishing the
universal proposition inductively by means of each of the particulars and
deducing the particular proposition from the universal syllogistically.''
Sextus Empiricus, Outlines of Pyrrhonism II-195 \cite{Empiricus1990Pyrrhonism}}
based on the empirical observation should be avoided, since it may easily lead
to false beliefs, this approach is often justified by the supposed
impossibility of defining a theory that correctly predicts failures.  A failure
of a wire due to environment (e.g. due to humidity, dust, heat \&c) is defined
from empirical evidences and processes have been standardized to test qualities
of hardware components.  This process completely breaks down when a malicious
environment (i.e. an attacker) is considered instead of the (supposedly honest
and predictable) natural environment. Therefore, the same approach that is in
use for safety, seems not to be applicable for security (e.g. for security
testing).
An overview on the aforementioned aspects of safety and security is depicted in
Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
the terms that structure our current understanding of security. 
\begin{itemize}
	\item \emph{Mankind} ``refers collectively to humans''
\cite{wiki-mankind}, while the concept of \emph{Nature} is
		related ``to the intrinsic characteristics that plants,
		animals, and other features of the world develop of their own
		accord'' (e.g. the physical universe)\cite{wiki-nature}. 
		\begin{itemize}
			\item There exist several terms to refer to an
				\emph{attacker}, i.e. threat agent or threat source,
				considering those terms to be semantically
				equivalent.  From now on, we consider the
				Causality principle to be the \emph{threat
				source}, Nature or Mankind to be the
				\emph{threat agents} and an \emph{attacker} as
				a specific malicious threat agent which materializes a
				threat.
		\end{itemize}
	\item \emph{Vulnerability}\footnote{The term vulnerability is not
		present in the Encyclopedia of Cryptography and Security, while
		it is used in 12 entries (such as in the definition of
		``penetration testing'' \cite{caddy2005pentest})
		highlighting how commonly this word is used without a proper
		supporting semantics.}, as defined in \cite{cnssi20104009}
		(and adopted in \cite{nist2013800-53}), is ``weakness in an
		information system, system security procedures, internal
		controls, or implementation that could be exploited by a threat
		source''. On the one hand, the definition is broad to enclose
		as much causes (that generates a vulnerability) as possible
		On the other hand, the term vulnerability should have a complete and sound
		definition, so that no other causes (e.g.  other sources) but
		the ones in the definition are responsible for a vulnerability
		\footnote{``For this view, that
		\emph{That Which Is Not} exists, can never predominate. You
		must debar your thought from this way of search, nor let
		ordinary experience in its variety force you along this way,
		(namely, that of allowing) the eye, sightless as it is, and the
		ear, full of sound, and the tongue, to rule; but (you must)
		judge by means of the Reason (Logos) the much-contested proof
		which is expounded by me.'' -- Parmenides of Elea, On Nature
		(circa 500 B.C.), fragments B7.1–8.2
		\cite{Hakim2016philosophy}}.
		Furthermore, the term ``threat sources'' used in the definition
		in\cite{cnssi20104009} may be identified with both Nature
		and Mankind, not differentiating between safety and security.
\end{itemize}

As depicted in Figure~\ref{fig:safety-security}, a vulnerability does not
necessarily become a threat for the system, unless exploited ``through a
channel that allows the violation of the security policy
[\ldots]''\cite{cnssi20104009}. For example, a software or procedure that takes
advantage of the vulnerability causing an \emph{attack} to the system may
result in several correlated incidents and threats.  The process of
exploitation of a defect as a vulnerability is reported in
Figure~\ref{fig:safety-security} such that the difference between exploit and failure,
and attack and accident, is to be found just in the maliciousness of the intents
that causes this process (i.e. excluding the intent, the terms are just syntactic transformation from a vulnerability to defect, from
accident to incident).

\begin{itemize}
	\item \emph{Weakness}. The definition given by the MITRE
		in \cite{MITRE2020CWEweakness} of weakness is: ``Software
		weaknesses are errors that can lead to software
		vulnerabilities. A software vulnerability, such as those
		enumerated on the Common Vulnerabilities and Exposures (CVE)
		List, is a mistake in software that can be directly used by a
		hacker to gain access to a system or network''.  The definition
		is circular if we interpret the word ``error'' and ``mistake''
		with the same semantics: a weakness is an error that leads to a
		vulnerability and a vulnerability is a mistake which, in turn,
		is a weakness.  
\end{itemize}

The only difference between a weakness and vulnerability seems to be that one
can consider weakness as a ground term and state that a vulnerability is caused
by a weakness.

\begin{itemize}
	\item \emph{Causality} refers to the causality principle; defined
		in\cite{Spirkin1983Dialectical} as ``Causality is a genetic
		connection of phenomena through which one thing (the cause)
		under certain conditions gives rise to, causes something else
		(the effect). The essence of causality is the generation and
		determination of one phenomenon by another. In this respect,
		causality differs from various other kinds of connection, for
		example, the simple temporal sequence of phenomena, of the
		regularities of accompanying processes''. One can consider the
		causality principle to be formalized by the K Modal Logic.
	\item An \emph{Exploit}\footnote{We note that the term exploit is only
		used as a verb in\cite{ISO2009information}} ``[\ldots]
		(from the English verb to exploit, meaning to use something to
		one’s own advantage) is a piece of software, a chunk of data,
		or a sequence of commands that takes advantage of a bug or
		vulnerability to cause unintended or unanticipated behavior to
		occur on computer software, hardware, or something electronic
		(usually computerized).''\cite{wiki-exploit}.
	\item An \emph{Attack}, as defined by the International Standard
		ISO/IEC 27000 is an ``attempt to destroy, expose, alter,
		disable, steal or gain unauthorized access to or make
		unauthorized use of an asset''; where an \emph{Asset} is
		``anything that has value to the organization''. We note that for
		the purpose of this article, we do not want to focus on a specific
		organization or business to define asset but, in general, on any 
		abstract organization (e.g. a company or a society).
		We do not consider ethical hackers as attacking a system. 
		In fact, we consider the term \emph{hack} as
		non-malicious (as, e.g. in \cite{Stallman2002hacker}).
	\item A \emph{Threat}, as defined in\cite{cnssi20104009}, is ``Any
		circumstance or event with the potential to adversely impact
		organizational operations (including mission, functions, image,
		or reputation), organizational assets, individuals, other
		organizations, or the Nation through an information system via
		unauthorized access, destruction, disclosure, modification of
		information, and/or denial of service''.
	\item \emph{Defect}, ``anything that renders the product not reasonably
		safe''\cite{Robinson2019writing} (i.e. a characteristic of
		an object which hinders its proper usability).
	\item \emph{Failure}, as defined in\cite{Merriam2020failure} as ``a state of
		inability to perform a normal function''. The term is
		structured and detailed in
		\cite{cnssi20104009,iet2017glossary} but relying on an
		abstract notion of failure without a specific definition.
	\item \emph{Hazard}, ``a potential source of
		harm''\cite{iet2017glossary}.
\end{itemize}

Our literature review shows that most of the definitions relates insecurity to
dis-honesty (also called maliciousness or adversarial) of an agent (often
called adversary or attacker). This, however, just moves the problem of
defining what cybersecurity is as the problem of defining what an dishonest
agent can or cannot do in a system\footnote{where an agent is any virtual or
physical entity of the system or using the system (e.g. a device, a software,
or a human being) and dishonesty is not necessarily related to malicious
motivation but also to incompetence or lack of skills.}
As in the Dolev-Yao theory, we may correlate ``being dishonest'' to ``not
following the intended behavior/rules''.  In the case of a generic system, a
dishonest agent is, therefore, any agent that doesn't follow the intended
behavior (or functionality or logic) of the system.  Given the generality of
the definition, and its high level of abstraction, we may
conclude that the hypothesis seems evident. For example, a software can be
considered an agent of the system and whenever it has a bug, it can be
exploited causing an Incident. However, this hypothesis has lead to the
un-testable conclusion that the dishonest behaviors of agents cannot be defined
in general (e.g. due to the heterogeneity of agents and systems) 
and that huge repository of dishonest behaviors should be kept as
definition, such as 
CAPEC\cite{MITRE2020CAPEC};
or that dishonesty of an agent with respect to a security protocol should be
defined as a number of predefined actions as in
\cite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
(to name a few). In order to define a theory on cybersecurity and on 
the malicious behaviors of agents we noticed that Attack Patterns (or Attacks)
are caused-by the presence of Vulnerabilities in the system. Those 
Vulnerabilities are, in turn, cause-by the presence of Weaknesses in the system.
Weaknesses are Errors in the design or implementation of a system.
Therefore, a theory on cybersecurity should first predict the Errors 
in a system design.

\section{Security Hypothesis -- ABF Theory of Agents}\label{sec:hypothesis}
Hypothesis formulated by looking at the causality principle and CWE/CVE/CAPEC repositories.
SUmmarize formal\_glossary justifying the choice of constituent of hyp.

\section{Prediction of Security Weaknesses}\label{sec:theory}
\paragraph{Fun, Behavior, Channels}

\section{Security Risk Assessment}\label{sec:secra}
\paragraph{Automated Security Requirements Extraction}

\section{Discussions}\label{sec:discussion}
\paragraph{Framework for Theory Falsification}
\paragraph{SPDL, Testing, and Falsification}
\paragraph{Relation with DY}

\section{Conclusion and Future Works}
\begin{itemize}
	\item Verification
	\item Testing
\end{itemize}

\section*{Acknowledgment}
We thank Katia Santac\`a for the development of the initial ideas, and for
the helpful discussions that allowed us to make this step forward.

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

%\bibliographystyle{IEEEtran}
%\bibliography{../literature}
\printbibliography

\end{document}

### 2020-06-06 reports/conference_paper/main.tex
section 3 (sec hyp) draft

@@
 \documentclass[conference]{IEEEtran}
-\usepackage{xcolor}
+\usepackage[table]{xcolor}
 \usepackage[english]{babel}
@@
 \usepackage[style=numeric]{biblatex}
+\usepackage{amsmath}
 \addbibresource{../literature.bib}% Syntax for version >= 1.2
@@
 \graphicspath{{../images/}}
+\usepackage{multirow}
 
+% circles
+\usepackage{wasysym}
+\newcommand{\Tdot}{$\CIRCLE$}
+\newcommand{\Thdot}{$\LEFTcircle$}
+\newcommand{\Twdot}{$\Circle$}
+
+% for fancy table with rotated column titles
+\usepackage{tabularx}
+\usepackage{adjustbox}
+\newcolumntype{R}[2]{%
+    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
+    l%
+    <{\egroup}%
+}
+\newcommand*\rot{\multicolumn{1}{R{45}{0.01em}}}% no optional argument here, please!
+\newcommand*\rota{\multicolumn{1}{R{90}{0.01em}}}% no optional argument here, please!
+\def\hb{\hbox to 10.7 cm{}}
+
+%colors in tables
+\definecolor{abfred}{RGB}{255, 215, 215}
+\definecolor{abf-rg-1}{RGB}{251,213,242}
+\definecolor{abf-rg-2}{RGB}{243,212,249}
+\definecolor{abf-rg-3}{RGB}{213,210,245}
+\definecolor{abf-rg-4}{RGB}{208,231,242}
+\definecolor{abf-rg-5}{RGB}{206,238,223}
+\definecolor{abf-rg-6}{RGB}{205,236,210}
+\definecolor{abf-rg-7}{RGB}{211,234,204}
+\definecolor{abfgreen}{RGB}{221,232,203}
+\definecolor{abforange}{RGB}{255, 245, 206}
+
+%fix commands
 \newcommand{\fixnote}[2]{\textbf{\color{red}{FIX}}\footnote{{\bf #1:} #2}}
@@
 % \renewcommand{\fix}[2]{}
+
+\newcommand{\assertionRegion}{\mathcal{A}}
+\newcommand{\beliefRegion}{\mathcal{B}}
+\newcommand{\factRegion}{\mathcal{F}}
+\newcommand{\rcc}{rcc}
+\newcommand{\agentuniverse}{Ag}
+\newcommand{\abf}{\assertionRegion,\beliefRegion,\factRegion}
+\newcommand{\Rcc}[2]{rcc(#1,#2)}
+\newcommand{\connects}[2]{C(#1,#2)}
+\newcommand{\disconnected}[2]{\neg C(#1,#2)}
+\newcommand{\partof}[2]{P(#1,#2)}
+\newcommand{\overlaps}[2]{O(#1,#2)}
+\newcommand{\eq}[2]{EQ(#1,#2)}
+\newcommand{\pp}[2]{PP(#1,#2)}
+\newcommand{\po}[2]{PO(#1,#2)}
+\newcommand{\ppi}[2]{PPi(#1,#2)}
+\newcommand{\dr}[2]{DR(#1,#2)}
+\newcommand{\all}[2]{T(#1,#2)}
+\newcommand{\rassert}[3]{\mathcal{A}_{#1\rightarrow #2}#3}
+\newcommand{\lassert}[3]{\mathcal{A}_{#1\leftarrow #2}#3}
+
+\newtheorem{definition}{Definition}%[section]
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{safety-security_3.pdf}
+	\includegraphics[width=.9\textwidth]{safety-security_3.pdf}
 	\caption{Overview of keywords related to security and safety}
@@
 
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.7\columnwidth]{causality.pdf}
+	\caption{Etiology of cybersecurity}
+	\label{fig:causality}
+\end{figure}
+
 Our literature review shows that most of the definitions relates insecurity to
@@
 or a human being) and dishonesty is not necessarily related to malicious
-motivation but also to incompetence or lack of skills.}
-As in the Dolev-Yao theory, we may correlate ``being dishonest'' to ``not
-following the intended behavior/rules''.  In the case of a generic system, a
-dishonest agent is, therefore, any agent that doesn't follow the intended
-behavior (or functionality or logic) of the system.  Given the generality of
-the definition, and its high level of abstraction, we may
-conclude that the hypothesis seems evident. For example, a software can be
-considered an agent of the system and whenever it has a bug, it can be
-exploited causing an Incident. However, this hypothesis has lead to the
-un-testable conclusion that the dishonest behaviors of agents cannot be defined
-in general (e.g. due to the heterogeneity of agents and systems) 
-and that huge repository of dishonest behaviors should be kept as
-definition, such as 
-CAPEC\cite{MITRE2020CAPEC};
-or that dishonesty of an agent with respect to a security protocol should be
-defined as a number of predefined actions as in
+motivation but also to incompetence or lack of skills.} As in the Dolev-Yao
+theory, we may correlate ``being dishonest'' to ``not following the intended
+behavior/rules''.  In the case of a generic system, a dishonest agent is,
+therefore, any agent that doesn't follow the intended behavior (or
+functionality or logic) of the system.  Given the generality of the definition,
+and its high level of abstraction, we may conclude that the hypothesis seems
+evident. For example, a software can be considered an agent of the system and
+whenever it has a bug, it can be exploited causing an Incident. However, this
+hypothesis has lead to the un-testable conclusion that the dishonest behaviors
+of agents cannot be defined in general (e.g. due to the heterogeneity of agents
+and systems) and that huge repository of dishonest behaviors should be kept as
+definition, such as CAPEC\cite{MITRE2020CAPEC}; or that dishonesty of an agent
+with respect to a security protocol should be defined as a number of predefined
+actions as in
 \cite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
-(to name a few). In order to define a theory on cybersecurity and on 
-the malicious behaviors of agents we noticed that Attack Patterns (or Attacks)
-are caused-by the presence of Vulnerabilities in the system. Those 
-Vulnerabilities are, in turn, cause-by the presence of Weaknesses in the system.
-Weaknesses are Errors in the design or implementation of a system.
-Therefore, a theory on cybersecurity should first predict the Errors 
-in a system design.
+(to name a few). As depicted in Figure~\ref{fig:causality}, in order to define
+a theory on cybersecurity and on the malicious behaviors of agents we noticed
+that Attack Patterns (or Attacks) are caused-by the presence of Vulnerabilities
+in the system. Those Vulnerabilities are, in turn, cause-by the presence of
+Weaknesses in the system.  Weaknesses are Errors in the design or
+implementation of a system.  Therefore, a theory on cybersecurity should first
+predict the Errors in a system design.
 
 \section{Security Hypothesis -- ABF Theory of Agents}\label{sec:hypothesis}
-Hypothesis formulated by looking at the causality principle and CWE/CVE/CAPEC repositories.
-SUmmarize formal\_glossary justifying the choice of constituent of hyp.
+In order to address the problem raised by Herley, we shall define how to
+distinguish between a secure and an insecure system. While most of the
+literature have correlated the problem of in-security
+to the maliciousness of agents interacting with the system, we show that
+security doesn't seem to stem from a malicious nature but, rather, insecurity
+raises from the lack of well-defined security requirements for the development
+process of a system. We argue that the high number of security vulnerability
+reported today are simply the realization of potential system configurations,
+which deviate from the nominal behavior because the \emph{intended (or
+nominal) behavior} of the system is not precisely defined in the specification
+at the very early stages of the engineering process.
+
+As a reference, we consider the following steps of the engineering process
+and our objective is to be able to predict the security requirements
+instead of axiomatically assume them.
+\begin{enumerate}
+	\item \emph{System Specification}, where the \emph{functional and physical requirements} are defined
+	\item \emph{Architecture Design}, where the specification is structured into \emph{functional and physical architectures}
+	\item \emph{Security Risk Assessment}, where the potential \emph{weaknesses} (errors)
+		are identified and, consequently, \emph{security requirements} are
+		captured.
+	%\item \emph{Design Refinement}, where the physical and functional
+	%	architectures are extended with \emph{mitigations} for the
+	%	weaknesses. Mitigations at this step details the semantics of
+	%	the functional blocks (behaviors), the data flow, and communication logic.
+	%\item \emph{Verification of Security Requirements}, where the data flow, behaviors, and communication protocols are verified against the security requirements.
+\end{enumerate}
+Given that, in our investigation, we changed the focus from the attacker to the
+potential designs of a system, we start by defining a general framework for the definition
+of a system. On top of this framework we will identify system weaknesses as potential
+design errors.
+
+\subsection{Systems and Multi-Agent Systems}\label{sec:system}
+The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
+of system as ``A combination of interacting elements organized to achieve one
+or more stated purposes.''\autocite{ISO201515288}.  Therefore, a system can be
+considered as a single agent where its interacting elements are the
+constituents of the agent itself. For the sake of simplicity we first define a
+system as an agent and then extend the definition to a ``combination of
+interacting'' agents.
+
+There is no agreement between the research communities (e.g.
+Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent
+as a system. However, the same ideas revolved around for thousands of years.
+Some relevant examples for our objective are the following\footnote{
+It is interesting to notice that 
+in \autocite{Empiricus1990Pyrrhonism}, the author states: ``The
+ logical criterion also may be used in three senses -- of the
+ agent, or the instrument, or the ``according to what''; the
+ agent, for instance, may be a man, the instrument either sense
+ perception or intelligence, and the ``according to what'' the
+ application of the impression ``according to'' which the man
+ proceeds to judge by means of one of the aforesaid
+ instruments.''. The author than proceeds showing that
+ even in this framework, knowledge still is non-apprehensible.
+}.
+\begin{itemize}
+	\item In \autocite{Hintikka1962knowledge}, Hintikka describes the
+		difference between Knowledge and Belief (as epistemological
+		concepts), and the whole Doxastic logic defines in details how
+		Beliefs can be formalized.
+	\item In \autocite{Hintikka1993Information}, Hintikka describes the concept
+		of Information and the difference with Knowledge and Belief.
+	\item In \autocite{Santaca2016abf}, the authors defines an agent as a
+		tuple of Assertions, Beliefs, and Facts.
+\end{itemize}
+For our argument, as in \autocite{Santaca2016abf}, an agent is composed by its
+knowledge, beliefs, and the information or assertions it provides; where
+knowledge is defined, as in \autocite{Steup2020epistemology}, as a set of
+proposition known by an agent, such that: (i) knowledge requires belief, (ii)
+knowledge require truth, (iii) knowledge must be \emph{properly justified}, and
+the only objective of Information is to exchange beliefs between agents.  As
+defined in \autocite{Hintikka1993Information}, ``Information is specified by
+specifying which alternatives concerning the reality it admits and which
+alternatives excludes''. This means that if we consider a propositional
+variable (which admits the two alternatives True/False) its information is
+defined as Believed to be True/False and not Believed to be the opposite.  Due
+to the definition of Information and then with its relation to the
+probabilistic correlation to truth/reality, we consider information in relation
+with an agent's beliefs. Similarly, we consider beliefs to define the actual
+behavior of an agent or a system.
+
+The difference between Knowledge and Information as used by Hintikka and Facts
+and Assertions as used in \autocite{Santaca2016abf} are not precisely discussed
+in \autocite{Santaca2016abf}.  Furthermore, those epistemological concepts are
+difficult to define
+\autocite{Gettier2012knowledge}\footnote{``\emph{Theaetetus}: [\ldots] He said
+that knowledge was true opinion accompanied by reason, but that unreasoning
+true opinion was outside of the sphere of knowledge; and matters of which there
+is not a rational explanation are unknowable -- yes, that is what he called
+them -- and those of which there is are knowable. [\ldots] \emph{Socrates}:
+[\ldots] the primary elements of which we and all else are composed admit of no
+rational explanation; for each alone by itself can only be named, and no
+qualification can be added, neither that it is nor that it is not, for that
+would at once be adding to it existence or non-existence, whereas we must add
+nothing to it, if we are to speak of that itself alone.  [\ldots]'' Plato --
+Theaetetus 201 \autocite{Plato1914Plato}}.  For example, Hintikka in
+\autocite{Hintikka1993Information} states that ``A purely logical definition of
+information is impossible''. A detailed formalization of those concepts is out
+of the scope of this paper. We now define them abstractly, inspired by the
+ABF-theory  defined in \autocite{Santaca2016abf} and, in
+Section~\ref{sec:theory} we detail them as engineering concepts.
+\begin{enumerate}
+	\item We consider Information only when the intention of exchanging 
+		that Information is from a sender to 
+		recipient is defined; and we call it \emph{Assertion}  
+	\item Similarly, the portion of Beliefs we consider for system
+		engineering is the one that builds (input) or describes
+		(output) the behavior of an agent (strategy rules
+		\footnote{``The logical structure of information is one of the
+		most basic and one of the most basic and one of the simplest
+		thing in the wide and wonderful world of logical analysis. This
+		point can be put in a deeper perspective. A distinction
+		[\ldots] ought to be made [--] between two kinds of rules (or
+		principles) in any strategic activity like knowledge seeking.
+		On the one hand you have the rules that define the game, e.g.
+		how chessmen are moved on a board. The can be called
+		\emph{definitory} rules.  They must be distinguished from rules
+		[\ldots] that deal with what is better and what is worse in the
+		game in question.  Definitory rules do not say anything about
+		this subject. Rules which do can be called \emph{strategic
+		rules}'' -- Hintikka in \autocite{Hintikka1993Information}}),
+		and
+	\item We consider a set of axiomatic \emph{Facts} (definitory
+		rules\fixnote{mr}{facts are definitory rule, as they don't
+		define how a real system is finally implemented (currently) but
+		how it should, trough a series of requirements which may not be
+		respected, through insecurity, in the implemented system})
+		instead of considering the more general epistemic definition of
+		Knowledge. Specifically, Facts describes:
+		\begin{itemize}
+			\item The functional architecture of each agent
+			\item The physical/structural (HW/SW) architecture of each
+				subsystem of agents and agent within a
+				subsystem
+		\end{itemize}
+\end{enumerate}
+We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}
+\fixnote{mr}{remove ``functional behavior'' in the figure and leave only ``behavior''}.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.9\columnwidth]{system-agent.pdf}
+	\caption{Example of system and agent structure}
+	\label{fig:system-agent}
+\end{figure}
+
+\subsection{Mereo-topological Reasoning}
+\begin{table*}[t]
+\centering
+\begin{tabular}{ccclll} 
+\rota{\textbf{RCC3}}&\rota{\textbf{RCC5}}&\rota{\textbf{RCC8}}&\textbf{Terminology} & \textbf{Notation} & \textbf{Definition} \\
+\hline
+&&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& $\mathit{X}\subseteq \mathit{Y}$ \\
+&&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& $\mathit{X}\not\subseteq \mathit{Y}$\\
+&&&Part of				& $\mathit{P}(\mathit{X},\mathit{Y})$		& $\forall \mathit{Z} ~\mathit{C}(\mathit{Z},\mathit{X}) \rightarrow \mathit{C}(\mathit{Z},\mathit{Y})$\\
+&&&Overlaps			& $\mathit{O}(\mathit{X},\mathit{Y})$		& $\exists \mathit{Z} ~\mathit{P}(\mathit{Z},\mathit{X})\wedge \mathit{P}(\mathit{Z},\mathit{Y})$\\
+\Tdot&&&  \textbf{Overlaps Not Equal} 	& $\mathit{ONE}(\mathit{X},\mathit{Y})$		& $\mathit{O}(\mathit{X},\mathit{Y}) \land \neg \mathit{EQ}(\mathit{X},\mathit{Y})$ \\
+\Tdot&\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
+\Tdot&\Tdot&\Tdot& \textbf{DiscRete from} 		& $\mathit{DR}(\mathit{X},\mathit{Y})$		& $\neg \mathit{O}(\mathit{X},\mathit{Y})$\\
+&\Tdot&\Tdot&\textbf{Partial-Overlap}	& $\mathit{PO}(\mathit{X},\mathit{Y})$ 		& $\mathit{O}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{X},\mathit{Y}) \wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\
+&\Tdot&&\textbf{Proper-Part-of} 	& $\mathit{PP}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\ 
+	&\Tdot&&\textbf{Proper-Part-of-\textit{\textbf{i}}nverse} & $\mathit{PPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{Y},\mathit{X}) \wedge \neg \mathit{P}(\mathit{X},\mathit{Y})$\\
+&&\Tdot&\textbf{Externally Connected} 	& $\mathit{EC}(\mathit{X},\mathit{Y})$ 		& $\mathit{C}(\mathit{X},\mathit{Y}) \wedge \neg\mathit{O}(\mathit{X},\mathit{Y})$\\ 
+&&\Tdot&\textbf{Tangential PP} 	& $\mathit{TPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
+&&\Tdot&\textbf{Tangential PPi} 	& $\mathit{TPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{TPP}(\mathit{Y},\mathit{X})$\\ 
+&&\Tdot&\textbf{Non-Tangential PP} 	& $\mathit{NTPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\neg\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
+&&\Tdot&\textbf{Non-Tangential PPi} 	& $\mathit{NTPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{NTPP}(\mathit{Y},\mathit{X})$\\ 
+\end{tabular}
+\caption{RCC3, RCC5, and RCC8 relations between Regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
+\end{table*}
+
+Similarly to\autocite{Santaca2016abf}, we define agents as a meronomy (an
+hierarchy of Part-Whole relations) over the constituent that we previously defined
+(assertions, beliefs, and facts), based on a standard definition of mereology,
+i.e. based on the definition of Parthood relation between \emph{Parts}.  Due to
+the necessity of considering different types of Part (as we'll show afterwards)
+we extend the mereology to a
+mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
+considering the relations in Table~\ref{tab:rcc358}.  For the sake of
+readability, we use the term \emph{Region} both to refer to a mereological Part
+and to a topological Region.  Our aim is to create a meronomy
+instead of the taxonomies such as the one provided
+in\autocite{NIST2020NVD,MITRE2020CVE} with the
+CVSS\autocite{Mell2007CVSS} scoring system.
+
+A mereotopology, as defined e.g. in\autocite{Rachavelpula2017mereotopology}, is
+an ordered mathematical structure where the basic relation between Regions is
+the reflexive and symmetric\fixnote{mr}{I guess it must be monotonic as defined
+in\autocite{Rachavelpula2017mereotopology} but I don't find it consistently in
+other papers.} \emph{Parthood} relation $\subseteq$.  
+\begin{definition}{\bf Parthood --}\label{def:parthood}
+Given any pair of mereotopological Regions $X$ and $Y$,
+	\begin{enumerate}[noitemsep]
+		\item Reflexivity: $\forall X.~ (X\subseteq X)$
+		\item Symmetry: $\forall X, Y.~ (X\subseteq Y \Rightarrow Y\subseteq X)$
+	\end{enumerate}
+\end{definition}
+
+The Parthood relation orders a universe of agents $\agentuniverse$ by defining the so called
+\emph{Connects with} (see in Table~\ref{tab:rcc358}) relation between Regions.
+We use the
+Region Connection Calculus (RCC), as defined
+in~\cite{bennettLogics,improvingRCC}, to provide an axiomatization of the
+mereo-topological concepts. In its broader definition, the RCC theory is composed
+by eight axioms, and is known as RCC8. In the text, for brevity, 
+we will often focus only on RCC5 (without
+loss of generality) by not considering tangential connections between spatial
+Regions. In Table~\ref{tab:rcc}, we
+summarize the axioms of the Region Connection Calculus (see, e.g., \autocite{Grutter2008rcc}).
+We can now define a system over the mereotopology using the RCC calculus, as follows.
+
+\begin{definition}{\bf System State --}\label{def:opsystem}
+	A CPS system (or a sub-system) state is defined as the tuple
+	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
+	where $\assertionRegion,\beliefRegion,\factRegion$, are regions of
+	assertions, beliefs (i.e. the beliefs generated by the behavior), and
+	facts expressed as requirements respectively.
+\end{definition}
+
+As already presented in \autocite{Santaca2016abf} it follows that, defining
+a system (or an operational system) with a fixed number of regions, there exist
+an upper-bound to the number of possible configuration of a system, defined by
+the possible relations between the different regions.
+For completeness, we report in the next paragraph 
+the calculation done in \autocite{Santaca2016abf}.
+
+\paragraph{Number of different configurations of a system}
+The general formula to calculate the number of different types of agents is
+$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
+between $n$ different sets, where $r^e$ is the number of permutation of $r$
+relations over $e$ elements with repetitions, with $e$ being the number of
+$k$-ary combinations of $n$ sets, $\binom{n}{k}$.
+In our case, $\binom{n}{k}=3$ since we consider $3$ sets
+($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
+considered in the RCC are binary.  Hence, using RCC5 (with five different
+spatial relations) over three sets, we can theoretically define up to 125
+different type of agents. However, only 54 of the 125 (as showed in
+\cite{improvingRCC}) combinations are topologically correct with respect to
+the definition of the relations of RCC5. Generalizing to all the RCCs:
+
+\begin{itemize}%[nosep]
+\item \emph{RCC3} --- theoretical: $3^3=27$,  correct: 15 
+\item \emph{RCC5} --- theoretical: $5^3=125$, correct: 54
+\item \emph{RCC8} --- theoretical: $8^3=512$, correct: 193
+\end{itemize}
+
+Hence, even if considering a different number of sets than the three
+$\assertionRegion$, $\beliefRegion$ and $\factRegion$ exponentially affects
+the number of theoretical agents, the application of RCC downscales that number
+of a factor that ranges from 1.8 to 2.5. In addition, using RCC5 we consider
+3.6 times more (different) types of agents than RCC3, but using RCC8 would
+allow us to consider 3.5 times more different agents.
+In the quantitative evaluation of a single agent, depicted in Figure~\ref{fig:quantitative},
+we argue that only 1 configuration represents the nominal (expected) behavior 
+of the agent while the other configurations are either impossible to 
+implement or diverge from the intended nominal behavior. We note 
+that, the numbers reported here do not consider the details of the
+engineering process and should be considered a limit of an abstract 
+representation of the system.
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{quantitative.pdf}
+	\caption{Cybersecurity risk for a single agent}
+	\label{fig:quantitative}
+\end{figure}
+
+\subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}
+While a quantitative analysis reveals how many possible configurations of an
+agent (i.e. a system) exist w.r.t. the $\abf$ theory (i.e. $54/125$ in RCC5), a
+qualitative analysis of the different configurations describes the
+configurations allowed by the $\abf$ theory, and how those configurations can
+be categorized.  In Table~\ref{tab:5com}, we provide the generic composition
+table of RCC5 over 3 regions instantiated over $\abf$, which shows the whole
+state space for a single agent. The color coding of the table represents the
+security risk related to a generic agent.
+
+As depicted in Figure~\ref{fig:soundness}, the relation between Facts, and
+Assertions and Behavior defines the soundness of the design (admissible
+configurations of Assertions, Behaviors and, in turn, Beliefs) w.r.t. the
+specification (what the specification mandates, such as by the nature of the
+physical/functional architectural models). We categorize agents by first
+analyzing the relations between each pair of Regions defining an agent (i.e.
+$\abf$), and then we categorize the different agents as tuple of the three
+Regions.  For the sake of simplicity, soundness is opposed to non-soundness in
+the following, however, with the RCC one should consider different ``degrees''
+of non-soundness. For example, in RCC5, if we consider EQ between two Regions
+as representing soundness, DR over the same Regions represents non-soundness;
+while PP, PO, PPi represents the different degrees of non-soundness.  A similar
+argument can be done for completeness.  \fix{mr}{it's missing the relation with
+equivalence. missing parallelism with knowledge-belief/assertion (see slide
+v-research-demo)}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{soundness.pdf}
+	\caption{Example relation between Facts, and Assertions and Beliefs}
+	\label{fig:soundness}
+\end{figure}
+
+\paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration} In order
+to reason on the relation between assertions and behavior we first need to
+consider that, by definition, assertions are defined as transfer of information
+between two agents, e.g., $a$ and $b$.  Therefore, as depicted in
+Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
+input and output assertions.  Given an agent $a$ and a collection of asserted
+predicates $\Phi$, the Input assertions are those received by $a$ from an agent
+$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
+are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
+two pairs of regions\footnote{``I am not asserting, as Lotze did, that a
+relation between $X$ and $Y$ consists of a quality in $X$ and a quality in $Y$
+-- a view which I regard as quite indefensible. -- I assert that a relation $Z$
+between $X$ and $Y$ involves the existence in $X$ of the quality ``having the
+relation $Z$ to $Y$'' so that a difference of relations always involves a
+difference in quality, and a change of relations always involves a change of
+quality.'' --  Ellis J. McTaggart, The Unreality of
+Time\autocite{Mctaggart1908unreality}}: 
+\begin{itemize}
+	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\beliefRegion$},
+		where the relation between Input-assertions and behavior
+		describes the soundness of the execution of the functional
+		architecture w.r.t. input elicitation. With more details, the
+		ideal specification of the functional architecture, along with
+		the expected inputs, defines the functional behavior of an
+		ideal system.  If all the inputs (Assertions) are correctly
+		handled in the functional specification (behavior) the
+		specification is complete.
+
+	\item \Rcc{$\assertionRegion_{a\rightarrow r}$}{$\beliefRegion$},
+		where the relation between behavior and outputs describes the
+		completeness of the behavior defined in the specification
+		w.r.t. the input elicitation.  With more details, if all the
+		outputs (assertions) of the functional architecture can be
+		produced, the functional architecture is complete.
+\end{itemize}
+
+\paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and
+\Rcc{$\beliefRegion$}{$\factRegion$} -- Honesty and Competence} While
+Assertions and Beliefs generated by the Behavior determines how a system should
+work, and the relation between the two defines the quality (i.e. the
+completeness w.r.t the specification) of the agent, the relation of Assertions
+and Behavior with Facts determines the quality with respect to the nominal
+(specified) system.  Given that Facts defines what must be true in the system,
+the relation of Assertions an Facts determines the degree of quality between
+the information circulating in a system (or within an agent) and
+reality\fixnote{mr}{assertions are reality while facts are dogmatic assertions
+at specification level, i.e. requirements}.  Since the transfer of information
+through Assertions generates Beliefs, a dishonest agent may circulate false
+information, generating false Beliefs.  We note that it is often implied that
+the intention behind circulating false information discerns a dishonest and an
+incompetent agent, however, we consider honesty related to sharing
+truths\footnote{``[\ldots] truth exists only in the good man, but the true in
+the bad man as well; for it is possible for the bad man to utter something
+true.'' Sextus Empiricus, Outlines of Pyrrhonism,
+II-83\autocite{Empiricus1990Pyrrhonism}}.  The relation between Beliefs and
+Facts determines the competence (on the subjects defined by the Facts) of an
+agent (i.e. the more competent an agent is, the more likely a Belief of that
+agent is true).
+
+\subsubsection{$\abf$ Security Enumeration (SE)} The following security
+requirement for a CPS specification can be summarized:
+\begin{enumerate}
+	\item[SE-$1$] Proper interaction between correctly-behaving agents (in
+		contrast with the ``Improper Interaction Between Multiple
+		Correctly-Behaving Entities'' defined by the CWE--$435$ as one
+		of the top ``view'' of the ``research concepts'' in
+		\autocite{MITRE2020CWEresearch}) is defined as
+		$\eq{\assertionRegion_a}{\beliefRegion_a}$ for an agent $a$
+		while can be detailed as follows when multiple agents are
+		considered.
+	\begin{enumerate}
+		\item[SE-$1.1$] The equality relation
+			$\eq{\assertionRegion_{s\rightarrow
+			a}}{\beliefRegion_a}$ describes the intended secure
+			behavior as: the beliefs generated by the behavior of
+			the functional architecture shall be complete w.r.t.
+			the specified inputs of the agent. Therefore, \emph{the
+			assertions received by an agent or a system shall be
+			compliant with the expected inputs of the functional
+			architecture}. For example, the inputs of the user of a
+			SW must be sanitized to exclude deviations w.r.t. the
+			expected inputs of the functions implemented in the SW.
+			Another example is the type checking between allowed
+			inputs and expected inputs.
+		\item[SE-$1.2$] Similarly, the equality relation
+			$\eq{\assertionRegion_{a\rightarrow
+			r}}{\beliefRegion_a}$ defines that the outputs of an
+			agent $a$ shall be the outputs of the functional
+			architecture.
+	\end{enumerate}
+	\item[SE-$2$]Sufficient Control Flow Management (in contrast with
+		the ``Insufficient Control Flow Management'' defined by MITRE
+		in the CWE--$691$ as one of the top ``view'' of the ``research
+		concepts'' in\autocite{MITRE2020CWEresearch}) is defined as
+		$\eq{\assertionRegion}{\factRegion}$.
+	\item[SE-$3$] Correct Calculation (in contrast with the ``Improper
+		Calculation'' defined by MITRE in the CWE--$682$ as one of the
+		top ``view'' of the ``research concepts''
+		in\autocite{MITRE2020CWEresearch}) is defined as
+		$\eq{\beliefRegion}{\factRegion}$.
+\end{enumerate}
+
+We are now in the position to define what a secure system is (with respect to
+the $\abf$-theory) and, based on that definition, what the security risk is and
+how to quantify it in a risk matrix.
+\begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
+	A secure system is a system where SE-$1$, SE-$2$, and SE-$3$ holds for
+	each agent composing the system.
+\end{definition}
+The ISO 31000 consider risk as the ``effect of uncertainty on objectives'' and
+the refers both of positive and negative consequences of uncertainty.
+Accordingly, we consider risk as follows. 
+\begin{definition}{\bf Risk --}
+The whole space of potential designs of a specification with respect to the
+	$\abf$-theory.
+\end{definition}
+The definition of Risk leads to the risk matrix in
+Figure~\ref{fig:quantitative}, defined as follows.
+\begin{definition}{\bf Risk Matrix --}
+	The risk matrix is a function between the three relations
+	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
+	where the maximum risk is defined by the $DR$ relation between the
+	three groups of Regions, and the minimum risk by the $EQ$ relation over
+	the same Regions. In between the two extremes, the granularity of
+	possible intermediate configuration is defined by the calculus used
+	(RCC5 in our case).
+\end{definition}
+We note that often the risk matrix is defined by the relation between
+likelihood and impact which we discuss in the next section.
+
+\begin{table*}[t]
+\centering
+\begin{adjustbox}{width=.6\textwidth}
+\begin{tabular}{r||c|c|c|c|c} 
+& \dr{$\assertionRegion$}{$\beliefRegion$} & 
+	\po{$\assertionRegion$}{$\beliefRegion$}& 
+	\pp{$\assertionRegion$}{$\beliefRegion$} &
+	\ppi{$\assertionRegion$}{$\beliefRegion$} & 
+	\eq{$\assertionRegion$}{$\beliefRegion$} \\
+\hline
+\hline %dr
+ \multirow{3}{*}{\dr{$\beliefRegion$}{$\factRegion$}} & 
+	\cellcolor{abfred} & %dr
+	\cellcolor{abf-rg-1}\dr{$\assertionRegion$}{$\factRegion$} & %po
+	\cellcolor{abf-rg-2}\multirow{3}{*}{\dr{$\assertionRegion$}{$\factRegion$}} & %pp
+	\cellcolor{abf-rg-3} \dr{$\assertionRegion$}{$\factRegion$}& % ppi
+	 \cellcolor{abf-rg-4} \\ % eq
+& \cellcolor{abfred}\all{$\assertionRegion$}{$\factRegion$}& %dr
+	\cellcolor{abf-rg-1}\po{$\assertionRegion$}{$\factRegion$} & %po
+	\cellcolor{abf-rg-2}\dr{$\assertionRegion$}{$\factRegion$} & %pp
+	\cellcolor{abf-rg-3}\po{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-4}\dr{$\assertionRegion$}{$\factRegion$}\\ % e q
+ & \cellcolor{abfred}& %dr
+	\cellcolor{abf-rg-1}\pp{$\assertionRegion$}{$\factRegion$} & %po
+	\cellcolor{abf-rg-2}  & %pp
+	\cellcolor{abf-rg-3}\pp{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-4} \\ %eq
+\hline %po
+ \multirow{3}{*}{\po{$\beliefRegion$}{$\factRegion$}} &
+	\cellcolor{abf-rg-1}\dr{$\assertionRegion$}{$\factRegion$} & %dr
+	\cellcolor{abf-rg-2} & %po
+	\cellcolor{abf-rg-3}\dr{$\assertionRegion$}{$\factRegion$} & %pp
+	\cellcolor{abf-rg-4}\po{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-5} \\ %eq
+ & \cellcolor{abf-rg-1}\po{$\assertionRegion$}{$\factRegion$} & %dr
+	\cellcolor{abf-rg-2} \all{$\assertionRegion$}{$\factRegion$} & %po 
+	\cellcolor{abf-rg-3}\po{$\assertionRegion$}{$\factRegion$} & %pp
+	\cellcolor{abf-rg-4}\ppi{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-5} \po{$\assertionRegion$}{$\factRegion$}\\%eq
+ & \cellcolor{abf-rg-1}\pp{$\assertionRegion$}{$\factRegion$} & %dr
+	\cellcolor{abf-rg-2} &  %po
+	\cellcolor{abf-rg-3}\pp{$\assertionRegion$}{$\factRegion$} & %pp
+	\cellcolor{abf-rg-4}& %ppi
+	\cellcolor{abf-rg-5}\\ %eq
+\hline %pp
+ \multirow{4}{*}{\pp{$\beliefRegion$}{$\factRegion$}} &
+	\cellcolor{abf-rg-2}\dr{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-3}&
+	\cellcolor{abf-rg-4} & %pp
+	\cellcolor{abf-rg-5}\po{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-6}\\ %eq
+ & \cellcolor{abf-rg-2}\po{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-3}\po{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-4}\pp{$\assertionRegion$}{$\factRegion$} & %pp
+	\cellcolor{abf-rg-5}\eq{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-6}\pp{$\assertionRegion$}{$\factRegion$} \\ %eq
+ & \cellcolor{abf-rg-2}\pp{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-3}\pp{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-4} & %pp
+	\cellcolor{abf-rg-5}\pp{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-6} \\ %eq
+ & \cellcolor{abf-rg-2}&
+ 	\cellcolor{abf-rg-3} & 
+	\cellcolor{abf-rg-4}& %pp
+	\cellcolor{abf-rg-5}\ppi{$\assertionRegion$}{$\factRegion$} & %ppi
+	\cellcolor{abf-rg-6} \\ %eq
+\hline %ppi
+ \multirow{3}{*}{\ppi{$\beliefRegion$}{$\factRegion$}} &
+ 	\cellcolor{abf-rg-3}\multirow{3}{*}{} &
+ 	\cellcolor{abf-rg-4}\dr{$\assertionRegion$}{$\factRegion$} &
+ 	\cellcolor{abf-rg-5}& %pp
+ 	\cellcolor{abf-rg-6}& %ppi
+ 	\cellcolor{abf-rg-7} \\ %eq
+& \cellcolor{abf-rg-3}\dr{$\assertionRegion$}{$\factRegion$} &
+ 	\cellcolor{abf-rg-4}\po{$\assertionRegion$}{$\factRegion$} & 
+ 	\cellcolor{abf-rg-5}\all{$\assertionRegion$}{$\factRegion$}& %pp
+ 	\cellcolor{abf-rg-6}\ppi{$\assertionRegion$}{$\factRegion$}& %ppi
+ 	\cellcolor{abf-rg-7}\ppi{$\assertionRegion$}{$\factRegion$}\\ %eq
+& \cellcolor{abf-rg-3} & 
+	\cellcolor{abf-rg-4}\ppi{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-5} & %pp
+	\cellcolor{abf-rg-6}& %ppi
+	\cellcolor{abf-rg-7}\\ %eq
+\hline %eq
+	\eq{$\beliefRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-4}\dr{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-5} \po{$\assertionRegion$}{$\factRegion$} & 
+	\cellcolor{abf-rg-6}\pp{$\assertionRegion$}{$\factRegion$} & %pp
+	\cellcolor{abf-rg-7} \ppi{$\assertionRegion$}{$\factRegion$} & %ppi 
+	\cellcolor{abfgreen} \eq{$\assertionRegion$}{$\factRegion$}  %eq
+\end{tabular}
+\end{adjustbox}
+\caption{RCC5 composition table over 3 regions. The results show that there
+exist 54 possible relations and the coloring anticipates the ideal risk matrix
+(green the secure state with low risk, red the high risk state, and a gradient
+of medium risk states). 
+\all{$\assertionRegion$}{$\factRegion$} = \{\dr{$\assertionRegion$}{$\factRegion$}, \po{$\assertionRegion$}{$\factRegion$}, \pp{$\assertionRegion$}{$\factRegion$}, \ppi{$\assertionRegion$}{$\factRegion$}, \eq{$\assertionRegion$}{$\factRegion$}\}
+\label{tab:5com}}
+\end{table*}
 
 \section{Prediction of Security Weaknesses}\label{sec:theory}
+\begin{figure*}[t]
+	\centering
+	\includegraphics[width=.9\textwidth]{mashyp.pdf}
+	\caption{asd}
+	\label{fig:mashyp}
+\end{figure*}
+
 \paragraph{Fun, Behavior, Channels}
@@
 \paragraph{Relation with DY}
+\paragraph{Abstraction Level}
 

### 2020-06-08 reports/conference_paper/main.tex
first full draft of the conference paper

@@
 \usepackage{multirow}
+\usepackage{enumitem}
+\usepackage{tikz}
+\usepackage{amsthm}
+\usepackage{rotating}
 
@@
 \newcommand{\abf}{\assertionRegion,\beliefRegion,\factRegion}
+\newcommand{\abftheory}{\assertionRegion\beliefRegion\factRegion}
 \newcommand{\Rcc}[2]{rcc(#1,#2)}
@@
 
+\newcommand*\halfbullet[1][0.4ex]{%
+  \begin{tikzpicture}
+	  \raisebox{0.6pt}{\draw[fill] (0,0)-- (90:#1) arc (90:270:#1) -- cycle ;
+	  \draw (0,0) circle (#1);}
+  \end{tikzpicture}}
+
+\newcommand{\fakedmapsfromchar}{%
+  \mathrel{\mathpalette\reflectmathchar\mapstochar}%
+}
+\makeatletter
+\newcommand{\reflectmathchar}[2]{%
+  \reflectbox{$\m@th#1#2$}%
+}
+\makeatother
+\newcommand{\rightbararrow}{\rightarrow\fakedmapsfromchar}
+\newcommand{\leftbararrow}{\mapstochar\rightarrow}
+\newcommand{\rightleftbararrow}{\mapstochar\rightarrow\fakedmapsfromchar}
+\setlength{\fboxrule}{0.1pt}
+\setlength{\fboxsep}{1.5pt}
+\newcommand{\secureportelem}{\mathrel{\rightarrow}}
+\newcommand{\secureport}{\fbox{\ensuremath \secureportelem}}
+\newcommand{\dropportelem}{\mathrel{\circ\!\!\rightarrow}}
+\newcommand{\dropport}{\fbox{\ensuremath \dropportelem}}
+\newcommand{\insertportelem}{\mathrel{\bullet\!\!\rightarrow}}
+\newcommand{\insertport}{\fbox{\ensuremath \insertportelem}}
+\newcommand{\injectionportelem}{\mathrel{\halfbullet\!\!\rightarrow}}
+\newcommand{\injectionport}{\fbox{\ensuremath \injectionportelem}}
+\newcommand{\selectiveportelem}{\mathrel{\leftbararrow}}
+\newcommand{\selectiveport}{\fbox{\ensuremath \selectiveportelem}}
+\newcommand{\selectiveinsertportelem}{\mathrel{\bullet\!\!\leftbararrow}}
+\newcommand{\selectiveinsertport}{\fbox{\ensuremath \selectiveinsertportelem}}
+\newcommand{\selectiveinjectionportelem}{\mathrel{\halfbullet\!\leftbararrow}}
+\newcommand{\selectiveinjectionport}{\fbox{\ensuremath \selectiveinjectionportelem}}
+\newcommand{\selectivedropportelem}{\mathrel{\circ\!\!\leftbararrow}}
+\newcommand{\selectivedropport}{\fbox{\ensuremath \selectivedropportelem}}
+%communication channels
+%secure input port
+\newcommand{\sdcom}{\mathrel{\rightarrow\!\!\circ}}
+\newcommand{\sicom}{\mathrel{\rightarrow\!\!\bullet}}
+\newcommand{\sjcom}{\mathrel{\rightarrow\!\!\halfbullet}}
+\newcommand{\ssdcom}{\mathrel{\rightbararrow\!\!\circ}}
+\newcommand{\ssicom}{\mathrel{\rightbararrow\!\!\bullet}}
+\newcommand{\ssjcom}{\mathrel{\rightbararrow\!\halfbullet}}
+%drop input port
+\newcommand{\ddcom}{\mathrel{\circ\!\!\rightarrow\!\!\circ}}
+\newcommand{\dicom}{\mathrel{\circ\!\!\rightarrow\!\!\bullet}}
+\newcommand{\dscom}{\mathrel{\circ\!\!\rightarrow}}
+%insert input port
+\newcommand{\idcom}{\mathrel{\bullet\!\!\rightarrow\!\!\circ}}
+\newcommand{\iicom}{\mathrel{\bullet\!\!\rightarrow\!\!\bullet}}
+\newcommand{\ijcom}{\mathrel{\bullet\!\!\rightarrow\!\!\halfbullet}}
+\newcommand{\isdcom}{\mathrel{\bullet\!\!\rightbararrow\!\!\circ}}
+\newcommand{\isicom}{\mathrel{\bullet\!\!\rightbararrow\!\!\bullet}}
+\newcommand{\isjcom}{\mathrel{\bullet\!\!\rightbararrow\!\halfbullet}}
+\newcommand{\iscom}{\mathrel{\bullet\!\!\rightarrow}}
+%injection input
+\newcommand{\jdcom}{\mathrel{\halfbullet\!\!\rightarrow\!\!\circ}}
+\newcommand{\jicom}{\mathrel{\halfbullet\!\!\rightarrow\!\!\bullet}}
+\newcommand{\jjcom}{\mathrel{\halfbullet\!\!\rightarrow\!\!\halfbullet}}
+\newcommand{\jsdcom}{\mathrel{\halfbullet\!\!\rightbararrow\!\!\circ}}
+\newcommand{\jsicom}{\mathrel{\halfbullet\!\!\rightbararrow\!\!\bullet}}
+\newcommand{\jsjcom}{\mathrel{\halfbullet\!\!\rightbararrow\!\halfbullet}}
+\newcommand{\jscom}{\mathrel{\halfbullet\!\!\rightarrow}}
+%selective drop input
+\newcommand{\sddcom}{\mathrel{\circ\!\!\leftbararrow\!\!\circ}}
+\newcommand{\sdicom}{\mathrel{\circ\!\!\leftbararrow\!\!\bullet}}
+\newcommand{\sdjcom}{\mathrel{\circ\!\!\leftbararrow\!\!\halfbullet}}
+\newcommand{\sdsdcom}{\mathrel{\circ\!\!\rightleftbararrow\!\!\circ}}
+\newcommand{\sdsicom}{\mathrel{\circ\!\!\rightleftbararrow\!\!\bullet}}
+\newcommand{\sdsjcom}{\mathrel{\circ\!\!\rightleftbararrow\!\halfbullet}}
+\newcommand{\sdscom}{\mathrel{\circ\!\!\leftbararrow}}
+%selective insertion
+\newcommand{\sidcom}{\mathrel{\bullet\!\!\leftbararrow\!\!\circ}}
+\newcommand{\siicom}{\mathrel{\bullet\!\!\leftbararrow\!\!\bullet}}
+\newcommand{\sijcom}{\mathrel{\bullet\!\!\leftbararrow\!\!\halfbullet}}
+\newcommand{\sisdcom}{\mathrel{\bullet\!\!\rightleftbararrow\!\!\circ}}
+\newcommand{\sisicom}{\mathrel{\bullet\!\!\rightleftbararrow\!\!\bullet}}
+\newcommand{\sisjcom}{\mathrel{\bullet\!\!\rightleftbararrow\!\halfbullet}}
+\newcommand{\siscom}{\mathrel{\bullet\!\!\leftbararrow}}
+%selective injection
+\newcommand{\sjdcom}{\mathrel{\halfbullet\!\!\leftbararrow\!\!\circ}}
+\newcommand{\sjicom}{\mathrel{\halfbullet\!\!\leftbararrow\!\!\bullet}}
+\newcommand{\sjjcom}{\mathrel{\halfbullet\!\!\leftbararrow\!\!\halfbullet}}
+\newcommand{\sjsdcom}{\mathrel{\halfbullet\!\!\rightleftbararrow\!\!\circ}}
+\newcommand{\sjsicom}{\mathrel{\halfbullet\!\!\rightleftbararrow\!\!\bullet}}
+\newcommand{\sjsjcom}{\mathrel{\halfbullet\!\!\rightleftbararrow\!\halfbullet}}
+\newcommand{\sjscom}{\mathrel{\halfbullet\!\!\leftbararrow}}
+
 \newtheorem{definition}{Definition}%[section]
+\newtheorem{hypothesis}{Hypothesis}%[section]
 
@@
 \begin{abstract}
-The abstract goes here.
+The objective of this research is the development of a theory that defines (all
+	and only) the possible insecurity and security configurations of any
+	abstract system. The theory is structured upon other theories that
+	defines how a component of a system can be abstracted into an agent,
+	defining how agents can be formalized (both syntactically and
+	semantically) to describe an abstract system, such as a graph. The core
+	theories are the epistemological definition of knowledge, Beliefs, and
+	Information, the Assertion-Belief-Fact framework, and mereotopology.
+	We implemented a formal theory (of axioms) of a mereotopology, and of
+	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
+	uses the Z3 SMT solver. The results show that a single component (i.e.
+	agent) of an abstract system has a definite number of  different
+	insecurity configurations (e.g. 53 using RCC5 over a topological
+	structure) and only 1 secure (i.e.  expected) configuration. The
+	configurations are reported as models satisfying the abstract system
+	semantics. We use these results to predict all possible security
+	weaknesses of a system, instead of relying on databases of known
+	weaknesses, and to lay the foundations of a scientific theory of
+	cybersecurity.  We show a concrete applications of our theory to the
+	risk assessment of an ad-hoc system.
 \end{abstract}
@@
 
-\subsection{Systems and Multi-Agent Systems}\label{sec:system}
+\subsection{Systems as Multi-Agent Systems}\label{sec:system}
 The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
@@
 \subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}
-While a quantitative analysis reveals how many possible configurations of an
-agent (i.e. a system) exist w.r.t. the $\abf$ theory (i.e. $54/125$ in RCC5), a
-qualitative analysis of the different configurations describes the
-configurations allowed by the $\abf$ theory, and how those configurations can
-be categorized.  In Table~\ref{tab:5com}, we provide the generic composition
-table of RCC5 over 3 regions instantiated over $\abf$, which shows the whole
-state space for a single agent. The color coding of the table represents the
-security risk related to a generic agent.
-
-As depicted in Figure~\ref{fig:soundness}, the relation between Facts, and
-Assertions and Behavior defines the soundness of the design (admissible
-configurations of Assertions, Behaviors and, in turn, Beliefs) w.r.t. the
-specification (what the specification mandates, such as by the nature of the
-physical/functional architectural models). We categorize agents by first
-analyzing the relations between each pair of Regions defining an agent (i.e.
-$\abf$), and then we categorize the different agents as tuple of the three
-Regions.  For the sake of simplicity, soundness is opposed to non-soundness in
-the following, however, with the RCC one should consider different ``degrees''
-of non-soundness. For example, in RCC5, if we consider EQ between two Regions
-as representing soundness, DR over the same Regions represents non-soundness;
-while PP, PO, PPi represents the different degrees of non-soundness.  A similar
-argument can be done for completeness.  \fix{mr}{it's missing the relation with
-equivalence. missing parallelism with knowledge-belief/assertion (see slide
-v-research-demo)}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\columnwidth]{soundness.pdf}
-	\caption{Example relation between Facts, and Assertions and Beliefs}
-	\label{fig:soundness}
-\end{figure}
-
-\paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration} In order
-to reason on the relation between assertions and behavior we first need to
-consider that, by definition, assertions are defined as transfer of information
-between two agents, e.g., $a$ and $b$.  Therefore, as depicted in
-Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
-input and output assertions.  Given an agent $a$ and a collection of asserted
-predicates $\Phi$, the Input assertions are those received by $a$ from an agent
-$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
-are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
-two pairs of regions\footnote{``I am not asserting, as Lotze did, that a
-relation between $X$ and $Y$ consists of a quality in $X$ and a quality in $Y$
--- a view which I regard as quite indefensible. -- I assert that a relation $Z$
-between $X$ and $Y$ involves the existence in $X$ of the quality ``having the
-relation $Z$ to $Y$'' so that a difference of relations always involves a
-difference in quality, and a change of relations always involves a change of
-quality.'' --  Ellis J. McTaggart, The Unreality of
-Time\autocite{Mctaggart1908unreality}}: 
-\begin{itemize}
-	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\beliefRegion$},
-		where the relation between Input-assertions and behavior
-		describes the soundness of the execution of the functional
-		architecture w.r.t. input elicitation. With more details, the
-		ideal specification of the functional architecture, along with
-		the expected inputs, defines the functional behavior of an
-		ideal system.  If all the inputs (Assertions) are correctly
-		handled in the functional specification (behavior) the
-		specification is complete.
-
-	\item \Rcc{$\assertionRegion_{a\rightarrow r}$}{$\beliefRegion$},
-		where the relation between behavior and outputs describes the
-		completeness of the behavior defined in the specification
-		w.r.t. the input elicitation.  With more details, if all the
-		outputs (assertions) of the functional architecture can be
-		produced, the functional architecture is complete.
-\end{itemize}
-
-\paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and
-\Rcc{$\beliefRegion$}{$\factRegion$} -- Honesty and Competence} While
-Assertions and Beliefs generated by the Behavior determines how a system should
-work, and the relation between the two defines the quality (i.e. the
-completeness w.r.t the specification) of the agent, the relation of Assertions
-and Behavior with Facts determines the quality with respect to the nominal
-(specified) system.  Given that Facts defines what must be true in the system,
-the relation of Assertions an Facts determines the degree of quality between
-the information circulating in a system (or within an agent) and
-reality\fixnote{mr}{assertions are reality while facts are dogmatic assertions
-at specification level, i.e. requirements}.  Since the transfer of information
-through Assertions generates Beliefs, a dishonest agent may circulate false
-information, generating false Beliefs.  We note that it is often implied that
-the intention behind circulating false information discerns a dishonest and an
-incompetent agent, however, we consider honesty related to sharing
-truths\footnote{``[\ldots] truth exists only in the good man, but the true in
-the bad man as well; for it is possible for the bad man to utter something
-true.'' Sextus Empiricus, Outlines of Pyrrhonism,
-II-83\autocite{Empiricus1990Pyrrhonism}}.  The relation between Beliefs and
-Facts determines the competence (on the subjects defined by the Facts) of an
-agent (i.e. the more competent an agent is, the more likely a Belief of that
-agent is true).
-
-\subsubsection{$\abf$ Security Enumeration (SE)} The following security
-requirement for a CPS specification can be summarized:
-\begin{enumerate}
-	\item[SE-$1$] Proper interaction between correctly-behaving agents (in
-		contrast with the ``Improper Interaction Between Multiple
-		Correctly-Behaving Entities'' defined by the CWE--$435$ as one
-		of the top ``view'' of the ``research concepts'' in
-		\autocite{MITRE2020CWEresearch}) is defined as
-		$\eq{\assertionRegion_a}{\beliefRegion_a}$ for an agent $a$
-		while can be detailed as follows when multiple agents are
-		considered.
-	\begin{enumerate}
-		\item[SE-$1.1$] The equality relation
-			$\eq{\assertionRegion_{s\rightarrow
-			a}}{\beliefRegion_a}$ describes the intended secure
-			behavior as: the beliefs generated by the behavior of
-			the functional architecture shall be complete w.r.t.
-			the specified inputs of the agent. Therefore, \emph{the
-			assertions received by an agent or a system shall be
-			compliant with the expected inputs of the functional
-			architecture}. For example, the inputs of the user of a
-			SW must be sanitized to exclude deviations w.r.t. the
-			expected inputs of the functions implemented in the SW.
-			Another example is the type checking between allowed
-			inputs and expected inputs.
-		\item[SE-$1.2$] Similarly, the equality relation
-			$\eq{\assertionRegion_{a\rightarrow
-			r}}{\beliefRegion_a}$ defines that the outputs of an
-			agent $a$ shall be the outputs of the functional
-			architecture.
-	\end{enumerate}
-	\item[SE-$2$]Sufficient Control Flow Management (in contrast with
-		the ``Insufficient Control Flow Management'' defined by MITRE
-		in the CWE--$691$ as one of the top ``view'' of the ``research
-		concepts'' in\autocite{MITRE2020CWEresearch}) is defined as
-		$\eq{\assertionRegion}{\factRegion}$.
-	\item[SE-$3$] Correct Calculation (in contrast with the ``Improper
-		Calculation'' defined by MITRE in the CWE--$682$ as one of the
-		top ``view'' of the ``research concepts''
-		in\autocite{MITRE2020CWEresearch}) is defined as
-		$\eq{\beliefRegion}{\factRegion}$.
-\end{enumerate}
-
-We are now in the position to define what a secure system is (with respect to
-the $\abf$-theory) and, based on that definition, what the security risk is and
-how to quantify it in a risk matrix.
-\begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
-	A secure system is a system where SE-$1$, SE-$2$, and SE-$3$ holds for
-	each agent composing the system.
-\end{definition}
-The ISO 31000 consider risk as the ``effect of uncertainty on objectives'' and
-the refers both of positive and negative consequences of uncertainty.
-Accordingly, we consider risk as follows. 
-\begin{definition}{\bf Risk --}
-The whole space of potential designs of a specification with respect to the
-	$\abf$-theory.
-\end{definition}
-The definition of Risk leads to the risk matrix in
-Figure~\ref{fig:quantitative}, defined as follows.
-\begin{definition}{\bf Risk Matrix --}
-	The risk matrix is a function between the three relations
-	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
-	where the maximum risk is defined by the $DR$ relation between the
-	three groups of Regions, and the minimum risk by the $EQ$ relation over
-	the same Regions. In between the two extremes, the granularity of
-	possible intermediate configuration is defined by the calculus used
-	(RCC5 in our case).
-\end{definition}
-We note that often the risk matrix is defined by the relation between
-likelihood and impact which we discuss in the next section.
-
 \begin{table*}[t]
@@
 \end{table*}
+While a quantitative analysis reveals how many possible configurations of an
+agent (i.e. a system) exist w.r.t. the $\abf$ theory (i.e. $54/125$ in RCC5), a
+qualitative analysis of the different configurations describes the
+configurations allowed by the $\abf$ theory, and how those configurations can
+be categorized.  In Table~\ref{tab:5com}, we provide the generic composition
+table of RCC5 over 3 regions instantiated over $\abf$, which shows the whole
+state space for a single agent. The color coding of the table represents the
+security risk related to a generic agent.
+
+As depicted in Figure~\ref{fig:soundness}, the relation between Facts, and
+Assertions and Behavior defines the soundness of the design (admissible
+configurations of Assertions, Behaviors and, in turn, Beliefs) w.r.t. the
+specification (what the specification mandates, such as by the nature of the
+physical/functional architectural models). We categorize agents by first
+analyzing the relations between each pair of Regions defining an agent (i.e.
+$\abf$), and then we categorize the different agents as tuple of the three
+Regions.  For the sake of simplicity, soundness is opposed to non-soundness in
+the following, however, with the RCC one should consider different ``degrees''
+of non-soundness. For example, in RCC5, if we consider EQ between two Regions
+as representing soundness, DR over the same Regions represents non-soundness;
+while PP, PO, PPi represents the different degrees of non-soundness.  A similar
+argument can be done for completeness.  \fix{mr}{it's missing the relation with
+equivalence. missing parallelism with knowledge-belief/assertion (see slide
+v-research-demo)}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{soundness.pdf}
+	\caption{Example relation between Facts, and Assertions and Beliefs}
+	\label{fig:soundness}
+\end{figure}
+
+\paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration} In order
+to reason on the relation between assertions and behavior we first need to
+consider that, by definition, assertions are defined as transfer of information
+between two agents, e.g., $a$ and $b$.  Therefore, as depicted in
+Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
+input and output assertions.  Given an agent $a$ and a collection of asserted
+predicates $\Phi$, the Input assertions are those received by $a$ from an agent
+$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
+are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
+two pairs of regions\footnote{``I am not asserting, as Lotze did, that a
+relation between $X$ and $Y$ consists of a quality in $X$ and a quality in $Y$
+-- a view which I regard as quite indefensible. -- I assert that a relation $Z$
+between $X$ and $Y$ involves the existence in $X$ of the quality ``having the
+relation $Z$ to $Y$'' so that a difference of relations always involves a
+difference in quality, and a change of relations always involves a change of
+quality.'' --  Ellis J. McTaggart, The Unreality of
+Time\autocite{Mctaggart1908unreality}}: 
+\begin{itemize}
+	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\beliefRegion$},
+		where the relation between Input-assertions and behavior
+		describes the soundness of the execution of the functional
+		architecture w.r.t. input elicitation. With more details, the
+		ideal specification of the functional architecture, along with
+		the expected inputs, defines the functional behavior of an
+		ideal system.  If all the inputs (Assertions) are correctly
+		handled in the functional specification (behavior) the
+		specification is complete.
+
+	\item \Rcc{$\assertionRegion_{a\rightarrow r}$}{$\beliefRegion$},
+		where the relation between behavior and outputs describes the
+		completeness of the behavior defined in the specification
+		w.r.t. the input elicitation.  With more details, if all the
+		outputs (assertions) of the functional architecture can be
+		produced, the functional architecture is complete.
+\end{itemize}
+
+\paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and
+\Rcc{$\beliefRegion$}{$\factRegion$} -- Honesty and Competence} While
+Assertions and Beliefs generated by the Behavior determines how a system should
+work, and the relation between the two defines the quality (i.e. the
+completeness w.r.t the specification) of the agent, the relation of Assertions
+and Behavior with Facts determines the quality with respect to the nominal
+(specified) system.  Given that Facts defines what must be true in the system,
+the relation of Assertions an Facts determines the degree of quality between
+the information circulating in a system (or within an agent) and
+reality\fixnote{mr}{assertions are reality while facts are dogmatic assertions
+at specification level, i.e. requirements}.  Since the transfer of information
+through Assertions generates Beliefs, a dishonest agent may circulate false
+information, generating false Beliefs.  We note that it is often implied that
+the intention behind circulating false information discerns a dishonest and an
+incompetent agent, however, we consider honesty related to sharing
+truths\footnote{``[\ldots] truth exists only in the good man, but the true in
+the bad man as well; for it is possible for the bad man to utter something
+true.'' Sextus Empiricus, Outlines of Pyrrhonism,
+II-83\autocite{Empiricus1990Pyrrhonism}}.  The relation between Beliefs and
+Facts determines the competence (on the subjects defined by the Facts) of an
+agent (i.e. the more competent an agent is, the more likely a Belief of that
+agent is true).
+
+\subsubsection{$\abf$ Security Enumeration (SE)} The following security
+requirement for a CPS specification can be summarized:
+\begin{enumerate}
+	\item[SE-$1$] Proper interaction between correctly-behaving agents (in
+		contrast with the ``Improper Interaction Between Multiple
+		Correctly-Behaving Entities'' defined by the CWE--$435$ as one
+		of the top ``view'' of the ``research concepts'' in
+		\autocite{MITRE2020CWEresearch}) is defined as
+		$\eq{\assertionRegion_a}{\beliefRegion_a}$ for an agent $a$
+		while can be detailed as follows when multiple agents are
+		considered.
+	\begin{enumerate}
+		\item[SE-$1.1$] The equality relation
+			$\eq{\assertionRegion_{s\rightarrow
+			a}}{\beliefRegion_a}$ describes the intended secure
+			behavior as: the beliefs generated by the behavior of
+			the functional architecture shall be complete w.r.t.
+			the specified inputs of the agent. Therefore, \emph{the
+			assertions received by an agent or a system shall be
+			compliant with the expected inputs of the functional
+			architecture}. For example, the inputs of the user of a
+			SW must be sanitized to exclude deviations w.r.t. the
+			expected inputs of the functions implemented in the SW.
+			Another example is the type checking between allowed
+			inputs and expected inputs.
+		\item[SE-$1.2$] Similarly, the equality relation
+			$\eq{\assertionRegion_{a\rightarrow
+			r}}{\beliefRegion_a}$ defines that the outputs of an
+			agent $a$ shall be the outputs of the functional
+			architecture.
+	\end{enumerate}
+	\item[SE-$2$]Sufficient Control Flow Management (in contrast with
+		the ``Insufficient Control Flow Management'' defined by MITRE
+		in the CWE--$691$ as one of the top ``view'' of the ``research
+		concepts'' in\autocite{MITRE2020CWEresearch}) is defined as
+		$\eq{\assertionRegion}{\factRegion}$.
+	\item[SE-$3$] Correct Calculation (in contrast with the ``Improper
+		Calculation'' defined by MITRE in the CWE--$682$ as one of the
+		top ``view'' of the ``research concepts''
+		in\autocite{MITRE2020CWEresearch}) is defined as
+		$\eq{\beliefRegion}{\factRegion}$.
+\end{enumerate}
+
+We are now in the position to define what a secure system is (with respect to
+the $\abf$-theory) and, based on that definition, what the security risk is and
+how to quantify it in a risk matrix.
+\begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
+	A secure system is a system where SE-$1$, SE-$2$, and SE-$3$ holds for
+	each agent composing the system.
+\end{definition}
+The ISO 31000 consider risk as the ``effect of uncertainty on objectives'' and
+the refers both of positive and negative consequences of uncertainty.
+Accordingly, we consider risk as follows. 
+\begin{definition}{\bf Risk --}
+The whole space of potential designs of a specification with respect to the
+	$\abf$-theory.
+\end{definition}
+The definition of Risk leads to the risk matrix in
+Figure~\ref{fig:quantitative}, defined as follows.
+\begin{definition}{\bf Risk Matrix --}
+	The risk matrix is a function between the three relations
+	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
+	where the maximum risk is defined by the $DR$ relation between the
+	three groups of Regions, and the minimum risk by the $EQ$ relation over
+	the same Regions. In between the two extremes, the granularity of
+	possible intermediate configuration is defined by the calculus used
+	(RCC5 in our case).
+\end{definition}
+We note that often the risk matrix is defined by the relation between
+likelihood and impact which we discuss in the next section.
 
 \section{Prediction of Security Weaknesses}\label{sec:theory}
+Several standards mandates a secure-by-design approach in which cybersecurity
+shall be considered at the very early stages of the design process.  For
+example,
+\begin{enumerate}[noitemsep]
+	\item DO-326A -- ``Airworthiness Security Process Specification''
+		requires a cybersecurity risk assessment of the design and is
+		the ``are the only Acceptable Means of Compliance (AMC) by FAA
+		\& EASA for aviation cyber-security airworthiness certification,
+		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
+	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
+		Control System (ICS) Security''
+	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
+		for Cyber-Physical Vehicle Systems'' defines ``set of
+		high-level guiding principles for Cybersecurity as it relates
+		to cyber-physical vehicle systems'' and states that
+		``incorporate Cybersecurity into cyber-physical vehicle systems
+		from concept phase through production, operation, service, and
+		decommissioning''
+\end{enumerate}
+However, standards do not describe in detail how to perform a security risk 
+assessment and only vaguely define the overall objective. The overall
+objective of this first step of the cybersecurity engineering process
+is to provide an understanding of the potential security risks.
+Roughly speaking, a security risk assessment (a.g. \autocite{coras,secram}) 
+methodology starts from the (i) identification of assets, (ii) determines
+the threat scenarios (correlating vulnerabilities, threats and incidents) e.g. 
+threat diagrams in the CORAS approach, (iii) calculate the risk as a function
+between impact and likelihood of threat scenarios, and (iv) finally provides
+treatments (mandating a partial re-design) and residual risks.
+All the methodologies and tools we reviewed (Threatmodeler, CORAS, SECRAM, ...)
+relies on the expertise of the person who performs the risk assessment for
+the identification of threats and for the quantitative estimation of risks.
+As Herley puts it \autocite{cite}, ``we don't have a test for security and
+we don't have a metric for security'' while standards requires a security risk assessment
+which, in turn, mandates a metric to evaluate the security risk.
+
+So far, we have reasoned on systems in the abstract, considering
+generic MAS; mapping Epistemological concepts
+to MAS. We now map MAS to CPS and show how this mapping allows us
+to predict security weaknesses in a system.
+
+\subsection{From Multi-Agent to Cyber-Physical Systems}
 \begin{figure*}[t]
@@
 
-\paragraph{Fun, Behavior, Channels}
-
-\section{Security Risk Assessment}\label{sec:secra}
-\paragraph{Automated Security Requirements Extraction}
+As depicted in Figure~\ref{fig:mashyp}, we relate MAS and CPS as follows.
+\begin{itemize}
+	\item We consider a System as an hierarchy of agents. So, we map agents
+		to \emph{systems, sub-systems, or devices}, depending on the
+		granularity of the design. For example, a modeler can model a
+		specific device, as in Figure~\ref{fig:mashyp}. In this case,
+		the system is not decomposed into sub-systems or devices and
+		the device is considered as an agent.
+	\item Agents reason over Beliefs (i.e. transform Beliefs into other
+		Beliefs) and for each \emph{component of a CPS} (system, sub-system,
+		or device) is composed by a \emph{functional architecture} that
+		transforms input-Beliefs into output-Beliefs.
+	\item Components of a CPS have \emph{ports} to exchange information
+		with the outer environment (which may be a sub-system),
+		similarly to Agents. The transfer of information in a CPS is
+		defined by \emph{channels}.
+	\item The concept of Knowledge is related to the \emph{requirements}
+		that, in a factorial way, describe how the CPS shall behave and
+		its \emph{physical architecture}.
+
+\end{itemize}
+
+\paragraph{Input and Output Ports}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{IOports.pdf}
+	\caption{Input and Output Port of an agent}
+	\label{fig:ioports}
+\end{figure}
+Since the $\abf$-theory is a theory of agents, we could consider ports as
+agents that allows the exchange of information between a channel and another
+agent.  However, we considered a port as a special type of agent that to avoid
+an \emph{infinite regress} in which a port needs a port to transfer information
+between the outside of the port to the inside of itself.
+
+\begin{definition}{\bf Input or Output Port --}\label{def:port} 
+	A port forwards information from the outside of an agent's boundary to
+	the inside (input-port) or vice versa (output-port).  There exist two
+	types of ports with the following behavior: an input-port changes
+	transforms Assertions from a sender $s$ ($\rassert{s}{a}$) to Beliefs
+	of an agent $a$ ($\beliefRegion_a$), while an output-port from belief
+	to assertion.
+\end{definition}
+The \emph{quality of a port} is determined by the $\rcc$ relation between the
+assertions received or sent and the belief, i.e.
+$\Rcc{\assertionRegion_s}{\beliefRegion_a}$ for the input-port or
+$\Rcc{\assertionRegion_r}{\beliefRegion_a}$. A port is, in fact, syntactic sugar to express the 
+relation between Assertions and Beliefs. We only define secure input port
+but the definition of a secure output port is symmetrical.
+
+\begin{definition}{\bf Secure Input Port --}\label{def:secport}
+	A secure input port ($s\secureport a$) always allow information as
+	incoming assertions to flow from a sender $s$ to the behavior of the
+	recipient $a$ (agent).  
+	%If $p_I$ is a secure input port,
+	%$\world\models\varphi$, $\varphi\in\rassert{s}{a}$, and $a$ has the
+	%only input port $p_I$, $\world'\models\beliefRegion_a$ and
+	%$\varphi\in\beliefRegion_a$ for any $\world\modalrelation\world'$.
+\end{definition}
+
+\paragraph{Port Weaknesses}
+From our definition of ports, it follows that there exist only the following \emph{six}
+types of weaknesses, generating six types of insecure port in RCC5 (the
+notation is reported for input-ports, i.e. on the lef-hand side of the arrow):
+\begin{enumerate}[start=1, label={W\arabic*)}]
+	\item \emph{replace port} ($s\dropport a$), where assertions reaches
+		the port but do not pass the boundary of the agent (i.e. do not
+		become belief of the agent)
+	\item \emph{drop port} ($s\dropport a$), where assertions reaches the
+		port but do not pass the boundary of the agent (i.e. do not
+		become belief of the agent)
+	\item \emph{insertion port} ($s\insertport a$), where some new
+		information is believed by $a$ as incoming from the port but
+		$s$ didn't send it
+	\item \emph{injection port} ($s\injectionport a$), where information
+		coming from $s$ is substituted with new information which
+		becomes believed by $a$
+	\item \emph{selective port}, where some information passes the port and
+		part is either:
+	\begin{enumerate}[start=1, label={W4.\arabic*)}]
+		\item drop ($s\selectivedropport a$), 
+		\item drop+insert ($s\selectiveinsertport a$).
+	\end{enumerate}
+\end{enumerate}
+
+\fixnote{mr}{this proof needs to be restructured as in slides v-research-demo}
+\begin{proof}
+An input port is, in the $\abf$-theory, defined secure as long as the relation
+	between the two regions of input assertions $\assertionRegion$ and
+	output beliefs $\beliefRegion$ are equal, i.e.
+	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
+	should result in a weakness (related to an insecurity flaw) of that
+	input port.  Using RCC5, there exist exactly other $4$ different type
+	of relations, one of which is the discrete-from (DR) relation, i.e.
+	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
+	the DR relations, they have no subregion in common. Lets
+	define a function weight $|X|$ such as, for any region $X$, it
+	represents the smallest possible cardinality of a (mereo)topological base for
+	$X$; where a base is a collection of regions in a (mereo)topology such that
+	every open region can be written as union of elements of that base. 
+	We distinguish between Regions that contain Information and Regions that
+	don't by writing the latter as $\emptyset$.
+	\begin{enumerate}
+		\item if $\eq{\assertionRegion}{\beliefRegion})$ then either
+			$\assertionRegion=\beliefRegion=\emptyset$ (no
+			communication) or
+			$\assertionRegion=\beliefRegion\neq\emptyset$ (forward
+			communication)
+		\item if $\dr{\assertionRegion}{\beliefRegion})$ then
+			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
+			(we call \emph{insert} the former and \emph{full drop}
+			the latter case), or
+			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
+			which we call \emph{replace} (i.e. drop and insert)
+		\item if $\pp{\assertionRegion}{\beliefRegion})$ then
+			$\beliefRegion$ contains and extend $\assertionRegion$
+			which we call \emph{injection}
+		\item if $\ppi{\assertionRegion}{\beliefRegion})$ then
+			$\assertionRegion$ contains and extend $\beliefRegion$
+			which we call \emph{selective drop}
+		\item if $\po{\assertionRegion}{\beliefRegion})$ then a part of
+			the $\assertionRegion$ is contained in the
+			$\beliefRegion$ which is a combination of
+			\emph{selective drop and generation}
+	\end{enumerate}
+\end{proof}
+
+\paragraph{Communication Channels}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{channel.pdf}
+	\caption{Communication over a Mono-directional Channel}
+	\label{fig:channel}
+\end{figure}
+In this work, we only consider
+mono-directional channels and communication but the extension to bi-directional
+channel can be considered as the union of two unidirectional channels. A
+mono-directional channel is defined by the assertions sent or received (over
+the channel).
+
+We start by considering the difference between a (communication)
+mono-directional channel (channel from now on) and an agent, as we did for the
+ports, since the $\abf$-theory is a theory of agents.  In fact, if a channel
+were considered an agent (channel-agent) then the question would be how an
+agent would transfer its Assertions to the channel-agent. If the channel
+between the agent and the channel-agent is again an agent, we would generate an
+\emph{infinite regress}. Therefore, we do allow channel-agents but we assume a
+finite depth (of detail) for a channel, where there exists a bottom-channel
+which is not an agent. For now, we do not constrain a channel-agent in any way
+so there is no difference between a channel agent and agent. Therefore,
+we consider channels to be bottom-channels,
+defined as agents with the pre-defined behavior (i.e. defined in a axiomatic
+way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
+prevents us from introducing additional constraints to the channel as storing
+Assertions that are transferred over the channel, or filter out some
+input-assertions.}.
+
+\begin{definition}{\bf Secure Mono-directional Channel (bottom-channel) --}\label{def:monochannel}
+	A mono-directional channel between the two agents ($s \rightarrow r$) is defined as 
+	an agent which behavior is (dogmatically) defined as: to forward any Assertion received from $s$ over an input-port, to  
+	the output-port where $r$ is listening to.
+\end{definition}
+The \emph{quality of a mono-directional} channel is defined as the $\rcc$
+relation between the Assertions made by the sender and the ones received by the
+receiver, i.e. $\Rcc{\assertionRegion_s}{\assertionRegion_r}$.
+
+\paragraph{Channel Weaknesses}
+Given that a mono-directional bottom-channel is assumed to be perfectly
+forwarding any Assertion from its input-port to its output-port, there is no
+insecure behavior but only the combination of the weaknesses of the input and
+output port; therefore there exists $(7^2)-2=47$ theoretical configurations
+($7^2$ because there are $6$ insecure types of port, plus $1$ secure type, on
+both input and output side; and we exclude the configuration with $2$ secure
+types as input and output, $-2$); where only 43 are possible. For the sake of
+readability, we report 6 examples in the following but the proof by exhaustion
+over all the possible cases is straightforward.
+
+\begin{enumerate}[start=6, label={W\arabic*)}]
+	\item \emph{secure output port} and \emph{input drop port} ($s\sdcom r$), 
+	\item \emph{secure output port} and \emph{input insertion port} ($s\sicom r$), 
+	\item \emph{output drop port} and \emph{input drop port} ($s\ddcom r$)
+	\item \emph{output drop port} and \emph{input insertion port} ($s\dicom r$)
+	\item \emph{output drop port} and \emph{input secure port} ($s\dscom r$)
+	\item \emph{output injection port} and \emph{input secure port} ($s\jscom r$)
+\end{enumerate}
+
+\paragraph{Security Weaknesses} 
+It is important to note that all the results of the application of the
+$\abf$-theory to channels (the analysis of the relation
+$\Rcc{\assertionRegion_o}{\assertionRegion_i}$) lead to the same results of the
+analysis of a pair of an input and output port (i.e.
+$\Rcc{\beliefRegion_1}{\assertionRegion_2}$ and
+$\Rcc{\assertionRegion_2}{\beliefRegion_3}$).  The same result can be obtained
+by analyzing the relation between the outputs of a functional block and the
+inputs of another functional block, where functional blocks are constituents of
+the functional architecture as described afterwards.  As depicted in
+Figure~\ref{fig:atom}, so far, we have considered Information generated by a
+Port $P_I$ and then sent through a channel $C$ to another (recipient) port
+$P_O$. In this scenario where Ports and then Channels are atomic (otherwise
+raising infinite regress), we can only  consider the relations between Ports
+and Channel; considering both input-port to channel and channel to output-port.
+In fact, the Weaknesses of a channel are defined in terms of the Weaknesses of
+Ports. We now introduce the concept of functional architecture and its
+decomposition into functional blocks.  Io order to define a functional block
+without encountering a infinitely recursive definition, we must reach the same
+conclusions as for the Channel. Therefore, describing the information as
+flowing over a channel or in a functional block is purely syntactic sugar.
+
+We can summarize these results by saying that the relations between Assertions
+and Beliefs, Output Assertions of an Agent and Input Assertions of another
+Agent, or Output Beliefs of a Functional Block and Input Beliefs of another
+Block can \emph{only} be affected by the following weaknesses: replace, drop,
+injection, insertion, selective drop, and selective drop + insertion.  We can,
+then, deduce the following security properties to mitigate security weaknesses
+of a port or a channel (between ports, functional blocks, or both).
+\begin{itemize}
+	\item \emph{Order-preserving} -- it shall be known if Information is \emph{replaced}
+	\item \emph{Availability} -- it shall be known if Information is \emph{dropped} or \emph{selectively dropped}
+	\item \emph{Integrity} -- it shall be known if Information is \emph{injected}
+	\item \emph{Authentication} -- it shall be known if Information is \emph{inserted}
+\end{itemize}
+%We can also generalize the weaknesses related to the data flow (i.e. the flow
+%of Information as Assertions or Beliefs) as the relation between output-input
+%Beliefs between functional blocks, output-input Assertions of a channel,
+%inputs-outputs of a port, as follows.
+%\begin{itemize}
+%	\item \emph{equal}, expected nominal behavior
+%	\item \emph{discrete-from}, drops all the inputs and inserts new malicious data
+%	\item \emph{proper part}, selectively drops inputs
+%	\item \emph{proper part inverse}, forwards all the inputs but crafts and
+%	      inserts new malicious data
+%	\item \emph{partial overlap}, selectively drops inputs and inserts new data.
+%\end{itemize}
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.9\columnwidth]{engineering_relations.pdf}
+	\caption{Relations between Ports and Channel}
+	\label{fig:atom}
+\end{figure}
+
+\paragraph{Functional Architecture}
+A Functional Architecture takes Information as input-Beliefs and transforms the
+Information into output-Beliefs. Those transformations occurs within the
+Functional Architecture, where Functional Blocks transforms Beliefs into other
+Beliefs. Similarly to channels, we could consider a Functional Block as a
+Functional Architecture occurring in an infinite regress. Therefore, we
+consider Functional Blocks as executing an abstract undefined behavior, of
+which we only observe the inputs and the resulting outputs (Beliefs).
+
+\begin{definition}{\bf Functional Block and Architecture --}\label{def:funblock}
+	A functional block of an agent 
+	takes a region of input-Beliefs and outputs a region of output-Beliefs. 
+	A functional architecture is an interconnected system of functional blocks.
+\end{definition}
+It is evident that the quality of a functional block cannot be determined
+by the difference between its inputs and outputs (as we did for
+ports and channels). The behavior of a functional block
+cannot be determined in general; since any functional block will have 
+its own purpose based on functional requirements. 
+Therefore, while a port semantics is determined by the relation 
+between Assertions and Beliefs, the semantics of a Functional Block 
+is determined by the relation between Facts/Requirements and I/O Beliefs.
+
+In other words, a functional block is a generic agent with no pre-defined general
+behavior (while ports and channels have a pre-defined behavior).
+
+\begin{itemize}
+	\item $\po{\beliefRegion}{\factRegion}$ the component has a Byzantine behavior where occasionally outputs the expected
+	        output given the correct inputs. Not all the inputs are handled
+	        properly, nor all the expected outputs are always generated
+	        when correct inputs are given.
+	\item $\pp{\beliefRegion}{\factRegion}$ part of the expected outputs are not
+	        generated in response to the correct
+	        inputs
+	\item $\ppi{\beliefRegion}{\factRegion}$ the components
+	        correctly performs the expected behavior when the correct
+	        inputs are provided but is subject to input
+	        injections
+	\item $\dr{\beliefRegion}{\factRegion}$ the component
+		never performs the expected behavior (e.g. physical
+		damage)
+\end{itemize}
+
+\paragraph{Requirements as Facts}
+During the specification phase, for any agent, channel, port, functional block
+and architecture, there may exist a factorial requirement (Fact) predicating over them.
+In other words, any requirement is defined as a Fact since they must be true in
+any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
+Figure~\ref{fig:soundness}, Facts are strategic rules that defines how
+the system shall behave (by specification), while reality may be shown to
+be insecure.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.4\columnwidth]{sum.pdf}
+	\caption{Example of a Functional Block}
+	\label{fig:sum}
+\end{figure}
+
+As an example, consider a functional block that performs the summation of two
+inputs, as in Figure~\ref{fig:sum}, defined by the requirement $r := b_3 = b_1
++ b_2$. The possible relations between the behavior of the functional block and
+the requirements (i.e. $\Rcc{\beliefRegion}{\factRegion}$ is determined by
+the relations between the I/O Beliefs $sum(b_1,b_2)=b_3$ and the requirement
+$sum(b_1,b_2)=b_1+b_2$, as follows.
+\begin{itemize}
+	\item $\eq{\beliefRegion}{\factRegion}=\eq{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, 
+		where $\beliefRegion_3$ represents the Region of all the
+		outputs of the functional block $sum$ while the Region
+		$\beliefRegion_1+\beliefRegion_2$ represents the expected
+		outputs of an ideal implementation of the requirement $r$.
+		Therefore, the functional block correctly implements the
+		requirements.
+	\item
+		$\dr{\beliefRegion}{\factRegion}=\dr{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$,
+		the function block never implements the requirements.
+	\item
+		$\pp{\beliefRegion}{\factRegion}=\pp{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exist some inputs for which
+		the output is incorrect.
+	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exists some outputs that are not the result of the summation of the inputs but given the correct inputs the function always outputs correctly.
+	\item
+		$\po{\beliefRegion}{\factRegion}=\po{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$,
+		the component has a Byzantine behavior where occasionally
+		outputs the expected output given the correct inputs. Not all
+		the inputs are handled properly, nor all the expected outputs
+		are always generated when correct inputs are given.
+\end{itemize}
+
+\subsection{Security and Insecurity of a System}
+\begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
+	A system security design is given by a precise system 
+	specification over the physical and functional architectures, that
+	uniquely, i.e.  tested against the range of design possible in the
+	$\abftheory$-theory, defines the design built on top of those
+	requirements.
+\end{hypothesis}
+
+\begin{hypothesis}{\bf System Insecurity Design --}\label{hyp:insecurity}
+	If, given a system specification as a collection of
+	requirements on the system, there exist
+	a non-unique design with respect to those requirements, the number of
+	equivalent designs that fulfills the requirements quantitatively
+	defines the magnitude of insecurity of a system design.
+\end{hypothesis}
+
+Based on these hypothesis, security ($S$) can be expressed as the following equation.
+\begin{displaymath}
+	S=1-(\prod_{a\in Ag} r^{\binom{n}{k}}-\delta_T)
+\end{displaymath}
+where $r^{\binom{n}{k}}$ represents the number of possible configurations (permutations) 
+of an agent based on the number $r$ of relations (e.g. EQ, DR, PO) with arity $k$ between
+$n$ different sets in the $\abftheory$ theory. $\delta_T$ represents the 
+number of configurations which are not satisfiable with respect to the logical theory
+defining the 
+algebraic structure (topology) and constraints of the calculus (RCC). $Ag$ is the 
+set of all the agents in the system under evaluation.
+The constant 1 represents the nominal configuration of the system and then the
+only secure configuration. The product represents all the unintended, insecure, configurations.
+
+\subsection{Security Risk Assessment}\label{sec:secra}
+In order to test our theory we implemented a tool-chain (open-source under the
+AGPLv3 license and available at \autocite{cybersecurityGitHubRepo}) for the
+identification of weaknesses and the precise calculation of potential insecure
+configurations. The engineering of the $\abftheory$ is summarized in the UML
+Class diagram in Figure~\ref{fig:secraclassdiagram} (in appendix).
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{v-secra.pdf}
+	\caption{Security Risk Assessment Tool}
+	\label{fig:secra}
+\end{figure}
+As depicted in Figure~\ref{fig:secra}, the security risk assessment process
+starts with the definition of the use cases and architectural requirements; the
+specification of a system.  In our process, the specification is manually
+translated into a UML design where:
+\begin{itemize}
+	\item a \emph{Deployment Diagram} describes the \emph{Physical
+		Architecture}. Each agent is defined as a Node with (physical)
+		Ports, and Agent's ports are connected via Information Flow
+		connectors, representing the physical channel.
+	\item a \emph{Functional Architecture} is linked to each agent in the
+		deployment diagram and is defined by an \emph{Object Diagram}.
+		The Object Diagram is composed by Instances of Functional
+		Blocks, connected via Information Flow connectors.
+	\item the connection between the two diagrams is implemented by
+		``sockets'', functional blocks connected to a 
+		physical port.
+\end{itemize}
+
+The tool generates a graph-like internal structure which represents
+the specification (called ABF graph). The graph is translated into 
+a logical formula that represents the specification in the $\abftheory$-theory and,
+along with the axiomatization of the RCC5 calculus, is given as input to the Z3 SMT solver.
+The solver is used to identify all possible configurations of the system and, in turn,
+to identify all potential weaknesses. 
+The internal structures of the tool
+can be viewed as PDF and the results are reported into an Excel file.
+The Excel file also report the total number of configurations as indicating
+the security risk associated to the specification. A user can change the status
+of each weaknesses in the Excel file, from the initial (default) status (open)
+to ``mitigated''. As a consequence, the risk is re-calculated on-the-fly, i.e.
+without the need of running the tool again, based on annotations and formulas 
+in the Excel file.
+
+We note that, in this approach, security requirements are not imposed by the
+specification but are automatically extracted by our tool as potential weaknesses.
+
+\paragraph{Case Studies}
+We report the results of the evaluation of a water level reader (sensor ad-hoc example). 
+\begin{figure}
+	\centering
+	\includegraphics[width=.7\columnwidth]{eng_cs1.png}
+	\caption{Water Level Reader Deployment Diagram}
+	\label{fig:eng_cs1}
+\end{figure}
+As in Figure~\ref{fig:eng_cs1}, we defined 2 agents: sensorInTank and sensorBoard,
+which represent the physical reader that needs to be placed in a tank of water, and 
+the board that interprets the readings and outputs them as digital signals.
+The two components are connected by a wire.
+\begin{figure*}
+	\centering
+	\includegraphics[width=.7\textwidth]{internal_cs1.png}
+	\caption{Sensor Board Object Diagram}
+	\label{fig:int_cs1}
+\end{figure*}
+In Figure~\ref{fig:int_cs1} we report the functional architecture that
+receives the incoming communications from the sensor in the tank, and 
+communicates them encrypted.
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{results_excel.pdf}
+	\caption{Partial View of the results in the Excel file}
+	\label{fig:results}
+\end{figure*}
+The results, summarized in Figure~\ref{fig:results} reports more than 16 millions
+possible scenarios in which at least one component diverge from the specification.
 
 \section{Discussions}\label{sec:discussion}
-\paragraph{Framework for Theory Falsification}
-\paragraph{SPDL, Testing, and Falsification}
+\paragraph{Abstraction Level}
+The level of abstraction is always a concern when it comes to the definition of
+a system model.  In this paper, we justified our assumptions on the atomicity
+of ports, channels, and functional blocks but nothing prevents a modeler to
+consider other atomic element of the system specification and design.  Those
+additional elements may detail the system and may change the syntactic sugar we
+defined; however, it won't change the underlying (semantic) reasoning unless
+another calculus is used (instead of RCC).  Furthermore, other Infosec
+properties may be considered, such as authenticity or non-repudiation. We,
+however, leave them as future development.
+
+\paragraph{$\abftheory$ Falsification} The theory we developed can be falsified
+by providing an experiment that shows a weaknesses not predicted by the
+$\abftheory$-theory, or by showing that the weaknesses predicted for a certain
+experiment are impossible to realize.  This theory, however, needs to be
+detailed to be applicable to a full engineering process (we only cover the
+first steps).  We propose the following complete process as a draft for
+potential next steps in this direction.
+\begin{enumerate}
+\item System Specification
+\item Architecture Design
+\item Security Risk Assessment and Security Requirement Extraction
+\item Design Refinement, where the semantics of the functional behavior and
+	communication protocols are defined 
+\item Security Requirement Extension, based on the design refinement
+\item Verification of Security Requirements (data flow, behaviors, \& protocol
+	logic)
+\item Software Prototype Implementation/Synthesis
+\item Automated Security Testing (abstract test-suite generation \&
+	concretization)
+\item Hardware Prototype Implementation
+\item Automated Security Testing
+\item Penetration Testing
+  \end{enumerate}
+
 \paragraph{Relation with DY}
-\paragraph{Abstraction Level}
+The DY attacker model has been extensively
+used for the identification of logical weaknesses in security protocols.
+Our theory should predict all possible weaknesses which, in turn,
+should enable all the attacks identified by the DY model. 
+A proof that the two theories/models correctly predicts
+(or lacks in predicting) weaknesses and related protocol flaws would
+connect risk assessment and protocol verification, the first
+two steps of the secure engineering of systems.
 
 \section{Conclusion and Future Works}
-\begin{itemize}
-	\item Verification
-	\item Testing
-\end{itemize}
+We proposed a foundational theory on security, arguing that
+security-related issues are not related to the maliciousness of an agent but to
+the vagueness of the security controls on the engineering processes.
+The current state-of-the-art processes allows, given a specification,
+an engineer to design the system in such a way that security issues arise due to the lack
+of proper security risk assessment processes.
+Those design, again, lack of security verification processes based on a solid 
+security foundational theory and then permit the generation of insecure implementation.
+The security verification and test-case generation will be the focus of our next steps.
+
+We conclude that the problem of security is a problem related to the 
+many possible design and, in turn, implementation given a specification.
+The problem is correlated to the epistemological search for truth,
+where the challenge is to relate Information and Belief to Human Knowledge.
+Scientifically, generalizing the Human to an Agent, the problem is
+to relate Assertions and Behaviors, to Facts. From an engineering standpoint,
+by generalizing the concept of Agents as architectural subsystems, the
+problem is to link Channels (and Ports) and Functional Architectures to 
+Requirements.
 
 \section*{Acknowledgment}
-We thank Katia Santac\`a for the development of the initial ideas, and for
-the helpful discussions that allowed us to make this step forward.
+We thank Katia Santac\`a for the help in the development of the initial ideas, and for
+the helpful discussions that allowed us to make this step.
 
@@
 
+\appendix
+\begin{sidewaysfigure*}
+	\centering
+	\includegraphics[width=\textwidth]{secra_classDiagram.png}
+	\caption{Security Risk Assessment -- Class Diagram}
+	\label{fig:secraclassdiagram}
+\end{sidewaysfigure*}
+
 \end{document}

### 2020-06-09 reports/conference_paper/main.tex
minor change in conf paper
review of conference paper final draft

@@
 \section{Introduction}\label{sec:intro}
-In \cite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
+In \autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
 ``an asymmetry in computer security'', which he defines as follows: ``Things
@@
 think) but a vice.'' -- Karl Popper, Conjectures and
-Refutations \cite{popper1962conjectures}} to be nonscientific (i.e.
+Refutations \autocite{popper1962conjectures}} to be nonscientific (i.e.
 unfalsifiability is a fallacy of a theory), Herley concludes that there is no
 scientific theory on cybersecurity; which means that cybersecurity lays in the
-realm of pseudo-sciences \cite{Herley2016usenixvideo}.  Herley, e.g.
-in \cite{Herley2017justifying}, discusses the implications of a
+realm of pseudo-sciences \autocite{Herley2016usenixvideo}.  Herley, e.g.
+in \autocite{Herley2017justifying}, discusses the implications of a
 nonscientific approach to cybersecurity, and highlights the tremendous impact on
@@
 overspending.  While the criticism is investigated
-in \cite{Herley2016unfalsifiability}, no solution is provided nor envisioned.  On the
+in \autocite{Herley2016unfalsifiability}, no solution is provided nor envisioned.  On the
 contrary, the goal of this work is to lay the foundations of a
@@
 hard-drive).  There is also an apparent inconsistency in
-\cite{Herley2016unfalsifiability} that we seek to clarify before following (as
+\autocite{Herley2016unfalsifiability} that we seek to clarify before following (as
 we agree) the scientific path draw by Herley: cybersecurity is defined as an
@@
 attacker can be considered as an abstraction of an active attacker who controls
-the network but cannot break cryptography.} \cite{Dolev1983security} only
+the network but cannot break cryptography.} \autocite{Dolev1983security} only
 applies to specific instances (often called scenarios) and abstraction of the
@@
 number of sessions (even tough some approaches do reason on an infinite number
-of sessions such as \cite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
-the peers (e.g.  in the ASLan++ language \cite{Oheimb2010aslan++}), and (iv)
+of sessions such as \autocite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
+the peers (e.g.  in the ASLan++ language \autocite{Oheimb2010aslan++}), and (iv)
 the abstraction of the cryptographic primitives (e.g.  ProVerif vs CryptoVerif
-\cite{Blanchet2017symbolic}).  While many tools and theories improves those 
+\autocite{Blanchet2017symbolic}).  While many tools and theories improves those 
 issues, there is no agreement on which should be the definitive
@@
 difficult it is to even just compare the approaches is given in
-\cite{Cremers2009comparing}). 
+\autocite{Cremers2009comparing}). 
 \begin{figure}[t]
@@
 For example, in Figure~\ref{fig:protocol-example}, under the perfect
-cryptography assumption\footnote{As defined in \cite{Rocchetto2016cpdy}: ``In
+cryptography assumption\footnote{As defined in \autocite{Rocchetto2016cpdy}: ``In
 the so called perfect cryptography assumption, the security encryption scheme
@@
 mathematical and implementation details of the encryption scheme
-\cite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
+\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
 and assuming that no violation to any security property is done after message
@@
 this highlights the false sense of security that may derive from a
-non-falsifiable theory of system security\footnote{``To the superficial
-observer, the analysis of these forms seems to turn upon minutiae. It does in
-fact deal with minutiae, but they are of the same order as those dealt with in
-microscopic anatomy.'' -- Karl Marx, Capital Volume 1, 1867}.
+non-falsifiable theory of system security
+%\footnote{``To the superficial
+%observer, the analysis of these forms seems to turn upon minutiae. It does in
+%fact deal with minutiae, but they are of the same order as those dealt with in
+%microscopic anatomy.'' -- Karl Marx, Capital Volume 1, 1867}.
 
@@
 risk assessment of a CPS (with an ad-hoc example, based on the SWaT testbed
-\cite{Mathur2016swat}).  This application shows how our theory can be used to
+\autocite{Mathur2016swat}).  This application shows how our theory can be used to
 predict all of the possible security weaknesses of a system, allowing the
@@
 incomplete.  Similarly, if a security weakness would be predicted by our theory
-but found to be impossible to have our theory could be declared as wrong.
+but found to be impossible to realize, our theory could be declared as wrong.
 
@@
 
-\fixnote{mr}{add weakness-error in picture and text}
 In most of the natural languages the concepts of safety
@@
 software model developed for programming intelligent
-agents.''\cite{wiki-bdi}. In the BDI model, the intents represents the
+agents.''\autocite{wiki-bdi}. In the BDI model, the intents represents the
 deliberative state of an agent which determines the choice of that agent on
-what to do.} (accidents are unfortunate while incidents are not) of the causes
+what to do.} (incidents are intentional while accidents are not) of the causes
 that generates a threat; namely, nature is believed not to have malicious
@@
 how can we define a theory that predicts what a human being will do?
-The cybersecurity attacks seems to be related to the creativity of the attacker
+Cybersecurity attacks seems to be related to the creativity of the attacker
 and then unpredictable. Currently, our understanding of cybersecurity is
 stored into a network of databases of weaknesses (e.g.
-CWE\cite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE\cite{CVE},
-NVD\cite{NIST2020NVD}),
-and attacks (e.g. CAPEC\cite{MITRE2020CAPEC}, ATTACK\cite{attack}). Systems are
+CWE\autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE\autocite{CVE},
+NVD\autocite{NIST2020NVD}),
+and attacks (e.g. CAPEC\autocite{MITRE2020CAPEC}, ATT\&CK\autocite{MITRE2020ATTACK}). Systems are
 nowadays tested against 
-known attacks ignoring the fact that the considering a system secure
-if no known attack is possible, defines an unfalsifiable claim.
+known attacks, ignoring that considering a system secure
+if no known attack is possible is against the scientific method.
 As an example, if we have a system which is secure against all of the
 vulnerabilities in the CVE we just need to wait for a month to have 1000+ more
-vulnerabilities to test \cite{newVulns}. If we suppose that one of those new vulnerabilities
-affects our system, that should be the falsification of the hypothesis that a
-system is secure if no known attacks apply. So, even tough the field 
-of security protocols is still lacking of a general cybersecurity theory
+vulnerabilities to test \autocite{NVD2020dashboard}. If even just one of those new vulnerabilities
+affects our system, that falsify the hypothesis that a
+system is secure if no known attacks apply. 
+We argue that this ``implement and test'' approach should be changed.
+Even tough the field 
+of security protocols is still lacking of a general cybersecurity theory,
 the secure-by-design approach they apply can be considered as going
@@
 
+\subsection{Terminology}
 Most of the safety-preserving principles in the field of engineering of
@@
 which safety requirements are defined (e.g. in standards such as the IEC 61508
-or 61511\cite{IEC201761511}), are based on empirical tests and measurements
+or 61511\autocite{IEC201761511}), are based on empirical tests and measurements
 (therefore should be considered hypothesis and not definitions). While
@@
 deducing the particular proposition from the universal syllogistically.''
-Sextus Empiricus, Outlines of Pyrrhonism II-195 \cite{Empiricus1990Pyrrhonism}}
+Sextus Empiricus, Outlines of Pyrrhonism II-195 \autocite{Empiricus1990Pyrrhonism}}
 based on the empirical observation should be avoided, since it may easily lead
@@
 	\item \emph{Mankind} ``refers collectively to humans''
-\cite{wiki-mankind}, while the concept of \emph{Nature} is
+\autocite{wiki-mankind}, while the concept of \emph{Nature} is
 		related ``to the intrinsic characteristics that plants,
 		animals, and other features of the world develop of their own
-		accord'' (e.g. the physical universe)\cite{wiki-nature}. 
+		accord'' (e.g. the physical universe)\autocite{wiki-nature}. 
 		\begin{itemize}
@@
 		it is used in 12 entries (such as in the definition of
-		``penetration testing'' \cite{caddy2005pentest})
+		``penetration testing'' \autocite{caddy2005pentest})
 		highlighting how commonly this word is used without a proper
-		supporting semantics.}, as defined in \cite{cnssi20104009}
-		(and adopted in \cite{nist2013800-53}), is ``weakness in an
+		supporting semantics.}, as defined in \autocite{cnssi20104009}
+		(and adopted in \autocite{nist2013800-53}), is ``weakness in an
 		information system, system security procedures, internal
@@
 		source''. On the one hand, the definition is broad to enclose
-		as much causes (that generates a vulnerability) as possible
-		On the other hand, the term vulnerability should have a complete and sound
+		as much causes (that generates a vulnerability) as possible,
+		on the other hand the term vulnerability should have a complete and sound
 		definition, so that no other causes (e.g.  other sources) but
 		the ones in the definition are responsible for a vulnerability
-		\footnote{``For this view, that
-		\emph{That Which Is Not} exists, can never predominate. You
-		must debar your thought from this way of search, nor let
-		ordinary experience in its variety force you along this way,
-		(namely, that of allowing) the eye, sightless as it is, and the
-		ear, full of sound, and the tongue, to rule; but (you must)
-		judge by means of the Reason (Logos) the much-contested proof
-		which is expounded by me.'' -- Parmenides of Elea, On Nature
-		(circa 500 B.C.), fragments B7.1–8.2
-		\cite{Hakim2016philosophy}}.
+		%\footnote{``For this view, that
+		%\emph{That Which Is Not} exists, can never predominate. You
+		%must debar your thought from this way of search, nor let
+		%ordinary experience in its variety force you along this way,
+		%(namely, that of allowing) the eye, sightless as it is, and the
+		%ear, full of sound, and the tongue, to rule; but (you must)
+		%judge by means of the Reason (Logos) the much-contested proof
+		%which is expounded by me.'' -- Parmenides of Elea, On Nature
+		%(circa 500 B.C.), fragments B7.1–8.2
+		%\autocite{Hakim2016philosophy}}.
 		Furthermore, the term ``threat sources'' used in the definition
-		in\cite{cnssi20104009} may be identified with both Nature
+		in\autocite{cnssi20104009} may be identified with both Nature
 		and Mankind, not differentiating between safety and security.
@@
 channel that allows the violation of the security policy
-[\ldots]''\cite{cnssi20104009}. For example, a software or procedure that takes
+[\ldots]''\autocite{cnssi20104009}. For example, a software or procedure that takes
 advantage of the vulnerability causing an \emph{attack} to the system may
@@
 	\item \emph{Weakness}. The definition given by the MITRE
-		in \cite{MITRE2020CWEweakness} of weakness is: ``Software
+		in \autocite{MITRE2020CWEweakness} of weakness is: ``Software
 		weaknesses are errors that can lead to software
@@
 	\item \emph{Causality} refers to the causality principle; defined
-		in\cite{Spirkin1983Dialectical} as ``Causality is a genetic
+		in\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
 		connection of phenomena through which one thing (the cause)
@@
 	\item An \emph{Exploit}\footnote{We note that the term exploit is only
-		used as a verb in\cite{ISO2009information}} ``[\ldots]
+		used as a verb in\autocite{ISO2009information}} ``[\ldots]
 		(from the English verb to exploit, meaning to use something to
@@
 		occur on computer software, hardware, or something electronic
-		(usually computerized).''\cite{wiki-exploit}.
+		(usually computerized).''\autocite{wiki-exploit}.
 	\item An \emph{Attack}, as defined by the International Standard
@@
 		In fact, we consider the term \emph{hack} as
-		non-malicious (as, e.g. in \cite{Stallman2002hacker}).
-	\item A \emph{Threat}, as defined in\cite{cnssi20104009}, is ``Any
+		non-malicious (as, e.g. in \autocite{Stallman2002hacker}).
+	\item A \emph{Threat}, as defined in\autocite{cnssi20104009}, is ``Any
 		circumstance or event with the potential to adversely impact
@@
 	\item \emph{Defect}, ``anything that renders the product not reasonably
-		safe''\cite{Robinson2019writing} (i.e. a characteristic of
+		safe''\autocite{Robinson2019writing} (i.e. a characteristic of
 		an object which hinders its proper usability).
-	\item \emph{Failure}, as defined in\cite{Merriam2020failure} as ``a state of
+	\item \emph{Failure}, as defined in\autocite{Merriam2020failure} as ``a state of
 		inability to perform a normal function''. The term is
 		structured and detailed in
-		\cite{cnssi20104009,iet2017glossary} but relying on an
+		\autocite{cnssi20104009,iet2017glossary} but relying on an
 		abstract notion of failure without a specific definition.
 	\item \emph{Hazard}, ``a potential source of
-		harm''\cite{iet2017glossary}.
+		harm''\autocite{iet2017glossary}.
 \end{itemize}
@@
 and systems) and that huge repository of dishonest behaviors should be kept as
-definition, such as CAPEC\cite{MITRE2020CAPEC}; or that dishonesty of an agent
+definition, such as CAPEC\autocite{MITRE2020CAPEC}; or that dishonesty of an agent
 with respect to a security protocol should be defined as a number of predefined
 actions as in
-\cite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
+\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
 (to name a few). As depicted in Figure~\ref{fig:causality}, in order to define
@@
 behavior of an agent or a system.
-
-The difference between Knowledge and Information as used by Hintikka and Facts
-and Assertions as used in \autocite{Santaca2016abf} are not precisely discussed
+The difference between Knowledge and Information as used by Hintikka, and Facts
+and Assertions as used in \autocite{Santaca2016abf}, are not precisely discussed
 in \autocite{Santaca2016abf}.  Furthermore, those epistemological concepts are
@@
 \autocite{Hintikka1993Information} states that ``A purely logical definition of
-information is impossible''. A detailed formalization of those concepts is out
-of the scope of this paper. We now define them abstractly, inspired by the
+information is impossible''. However, a detailed formalization of those concepts is out
+of the scope of this paper.
+
+We now define them abstractly, inspired by the
 ABF-theory  defined in \autocite{Santaca2016abf} and, in
@@
 		recipient is defined; and we call it \emph{Assertion}  
-	\item Similarly, the portion of Beliefs we consider for system
+	\item Similarly, the portion of \emph{Beliefs} we consider for system
 		engineering is the one that builds (input) or describes
@@
 	\item We consider a set of axiomatic \emph{Facts} (definitory
-		rules\fixnote{mr}{facts are definitory rule, as they don't
-		define how a real system is finally implemented (currently) but
-		how it should, trough a series of requirements which may not be
-		respected, through insecurity, in the implemented system})
+		rules\footnote{Facts are definitory rule as they don't
+		define how a real system is finally implemented but
+		how it should be, trough a series of requirements. Those requirements 
+		may not hold, through insecurity, in the implemented system.})
 		instead of considering the more general epistemic definition of
@@
 We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}
-\fixnote{mr}{remove ``functional behavior'' in the figure and leave only ``behavior''}.
 
@@
 in\autocite{NIST2020NVD,MITRE2020CVE} with the
-CVSS\autocite{Mell2007CVSS} scoring system.
+so that we don't need to rely on a scoring system such as the 
+CVSS\autocite{Mell2007CVSS}. Instead, we want to precisely calculate
+the number of insecurity configurations as a scoring system.
 
@@
 an ordered mathematical structure where the basic relation between Regions is
-the reflexive and symmetric\fixnote{mr}{I guess it must be monotonic as defined
-in\autocite{Rachavelpula2017mereotopology} but I don't find it consistently in
-other papers.} \emph{Parthood} relation $\subseteq$.  
+the reflexive and symmetric
+\emph{Parthood} relation $\subseteq$.  
 \begin{definition}{\bf Parthood --}\label{def:parthood}
@@
 Region Connection Calculus (RCC), as defined
-in~\cite{bennettLogics,improvingRCC}, to provide an axiomatization of the
+in~\autocite{bennettLogics,improvingRCC}, to provide an axiomatization of the
 mereo-topological concepts. In its broader definition, the RCC theory is composed
@@
 As already presented in \autocite{Santaca2016abf} it follows that, defining
-a system (or an operational system) with a fixed number of regions, there exist
+a system with a fixed number of regions, there exist
 an upper-bound to the number of possible configuration of a system, defined by
@@
 $r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
-between $n$ different sets, where $r^e$ is the number of permutation of $r$
+between $n$ different regions, where $r^e$ is the number of permutation of $r$
 relations over $e$ elements with repetitions, with $e$ being the number of
-$k$-ary combinations of $n$ sets, $\binom{n}{k}$.
-In our case, $\binom{n}{k}=3$ since we consider $3$ sets
+$k$-ary combinations of $n$ regions, $\binom{n}{k}$.
+In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
 ($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
 considered in the RCC are binary.  Hence, using RCC5 (with five different
-spatial relations) over three sets, we can theoretically define up to 125
+spatial relations) over three regions, we can theoretically define up to 125
 different type of agents. However, only 54 of the 125 (as showed in
-\cite{improvingRCC}) combinations are topologically correct with respect to
+\autocite{improvingRCC}) combinations are topologically correct with respect to
 the definition of the relations of RCC5. Generalizing to all the RCCs:
@@
 
-Hence, even if considering a different number of sets than the three
-$\assertionRegion$, $\beliefRegion$ and $\factRegion$ exponentially affects
+Hence, even if considering a different number of regions than the three
+$\assertionRegion$, $\beliefRegion$, and $\factRegion$ exponentially affects
 the number of theoretical agents, the application of RCC downscales that number
@@
 while PP, PO, PPi represents the different degrees of non-soundness.  A similar
-argument can be done for completeness.  \fix{mr}{it's missing the relation with
-equivalence. missing parallelism with knowledge-belief/assertion (see slide
-v-research-demo)}
+argument can be done for completeness.  
 \begin{figure}[t]
@@
 \paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and
-\Rcc{$\beliefRegion$}{$\factRegion$} -- Honesty and Competence} While
-Assertions and Beliefs generated by the Behavior determines how a system should
-work, and the relation between the two defines the quality (i.e. the
-completeness w.r.t the specification) of the agent, the relation of Assertions
+\Rcc{$\beliefRegion$}{$\factRegion$} -- Honesty and Competence} 
+%While
+%Assertions and Beliefs generated by the Behavior determines how a system should
+%work, and the relation between the two defines the quality (i.e. the
+%completeness w.r.t the specification) of the agent, 
+The relation of Assertions
 and Behavior with Facts determines the quality with respect to the nominal
-(specified) system.  Given that Facts defines what must be true in the system,
+(specified) system.  Given that Facts defines what needs to be true in the system,
 the relation of Assertions an Facts determines the degree of quality between
-the information circulating in a system (or within an agent) and
-reality\fixnote{mr}{assertions are reality while facts are dogmatic assertions
-at specification level, i.e. requirements}.  Since the transfer of information
+the real information circulating in a system (or within an agent) and
+the one specified.  Since the transfer of information
 through Assertions generates Beliefs, a dishonest agent may circulate false
@@
 the intention behind circulating false information discerns a dishonest and an
-incompetent agent, however, we consider honesty related to sharing
+incompetent agent. However, we consider honesty related to sharing
 truths\footnote{``[\ldots] truth exists only in the good man, but the true in
@@
 The ISO 31000 consider risk as the ``effect of uncertainty on objectives'' and
-the refers both of positive and negative consequences of uncertainty.
+refers both to positive and negative consequences of uncertainty.
 Accordingly, we consider risk as follows. 
@@
 \end{definition}
-We note that often the risk matrix is defined by the relation between
-likelihood and impact which we discuss in the next section.
+The risk matrix is often defined as a function between 
+likelihood and impact of attacks. We suggest
+that security weaknesses are all equally likely to be exploited
+if there's a connection between the weakness and the asset that
+an attacker wants to impact. Therefore, a risk matrix should
+capture the number of insecure configurations of a system, rather
+than predicating over likelihoods of weaknesses exploitation.
 
@@
 is to provide an understanding of the potential security risks.
-Roughly speaking, a security risk assessment (a.g. \autocite{coras,secram}) 
+Roughly speaking, a security risk assessment (e.g. CORAS\autocite{Lund2010model}) 
 methodology starts from the (i) identification of assets, (ii) determines
 the threat scenarios (correlating vulnerabilities, threats and incidents) e.g. 
-threat diagrams in the CORAS approach, (iii) calculate the risk as a function
+threat diagrams in the CORAS
+approach, (iii) calculate the risk as a function
 between impact and likelihood of threat scenarios, and (iv) finally provides
 treatments (mandating a partial re-design) and residual risks.
-All the methodologies and tools we reviewed (Threatmodeler, CORAS, SECRAM, ...)
+All the methodologies and tools we reviewed (e.g. Threatmodeler \autocite{Threatmodeler},
+CORAS \autocite{Lund2010model}, SECRAM \autocite{De2015role})
 relies on the expertise of the person who performs the risk assessment for
 the identification of threats and for the quantitative estimation of risks.
-As Herley puts it \autocite{cite}, ``we don't have a test for security and
+As Herley puts it \autocite{Herley2016usenixvideo}, ``we don't have a test for security and
 we don't have a metric for security'' while standards requires a security risk assessment
@@
 
-\fixnote{mr}{this proof needs to be restructured as in slides v-research-demo}
 \begin{proof}
@@
 
-\begin{itemize}
+\begin{enumerate}[start=50, label={W\arabic*)}]
 	\item $\po{\beliefRegion}{\factRegion}$ the component has a Byzantine behavior where occasionally outputs the expected
@@
 		damage)
-\end{itemize}
+\end{enumerate}
 
@@
 of an agent based on the number $r$ of relations (e.g. EQ, DR, PO) with arity $k$ between
-$n$ different sets in the $\abftheory$ theory. $\delta_T$ represents the 
+$n$ different regions in the $\abftheory$ theory. $\delta_T$ represents the 
 number of configurations which are not satisfiable with respect to the logical theory
@@
 In order to test our theory we implemented a tool-chain (open-source under the
-AGPLv3 license and available at \autocite{cybersecurityGitHubRepo}) for the
+AGPLv3 license and available at \autocite{v-research2020cybersecurity}) for the
 identification of weaknesses and the precise calculation of potential insecure
@@
 The tool generates a graph-like internal structure which represents
-the specification (called ABF graph). The graph is translated into 
+the specification (called ABF graph). The ABF graph defines the system as
+a number of regions of assertions, beliefs, and facts. Those regions
+are connected by a generic relation which is evaluated as follows.
+The graph is translated into 
 a logical formula that represents the specification in the $\abftheory$-theory and,
 along with the axiomatization of the RCC5 calculus, is given as input to the Z3 SMT solver.
-The solver is used to identify all possible configurations of the system and, in turn,
+The solver identifies all possible configurations of the system and, in turn,
 to identify all potential weaknesses. 
@@
 
-We note that, in this approach, security requirements are not imposed by the
+In our approach, security requirements are not imposed by the
 specification but are automatically extracted by our tool as potential weaknesses.
+Those weaknesses are related to the different insecure configurations 
+of the specified system.
+
+The risk matrix is often defined as a function between 
+likelihood and impact of attacks. We suggest
+that security weaknesses are all equally likely to be exploited
+if there's a connection between the weakness and the asset that
+an attacker wants to impact.
 
@@
 %\bibliography{../literature}
-\printbibliography
-
 \appendix
@@
 \end{sidewaysfigure*}
+\printbibliography
 

### 2020-06-09 reports/conference_paper/compile.sh
add compile script for conference paper

@@ -1,0 +1,29 @@
+#!/bin/bash
+latexOutputDir="latex-output-dir"
+latexInputFile="main.tex"
+pdfOutputDir="${latexOutputDir}/${latexInputFile%.*}.pdf"
+
+
+#if [[ ! -d $latexOutputDir ]]; then 
+rm -rf $latexOutputDir
+mkdir $latexOutputDir
+#fi
+
+if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
+then 
+	if [ "$1" != "-short" ] && [ "$1" != "-s" ] 
+	then 
+		if biber --nodieonerror --input-directory=$latexOutputDir --output-directory=$latexOutputDir ${latexInputFile%.*}  
+		then
+			if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
+			then 
+				pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile 
+			fi
+		fi
+	fi
+
+	if [[ -f $pdfOutputDir ]]
+	then
+		mv $pdfOutputDir `dirname $0`
+	fi
+fi

### 2020-06-09 reports/confidential/engineering.tex
mark white paper as confidential

--- 2020-05-25 reports/journal_paper/engineering.tex
@@
 
-\Paragraph{System Requirements}
+\paragraph{System Requirements}
 In our running example, there is a system (a CPS) composed by $5$ main
@@
 
-\Paragraph{Port Weaknesses}
+\paragraph{Port Weaknesses}
 From this definition, it follows that there exist only the following \emph{six} types of weaknesses, generating six types of insecure port in RCC5 (the notation is reported for input-ports, i.e.
@@
 \begin{enumerate}[start=1, label={W\arabic*)}]
+	\item \emph{replace port} ($s\dropport a$), where assertions reaches the port but do not pass the boundary of the agent (i.e. do not become belief of the agent)
 	\item \emph{drop port} ($s\dropport a$), where assertions reaches the port but do not pass the boundary of the agent (i.e. do not become belief of the agent)
@@
 		\item drop ($s\selectivedropport a$), 
-		\item insert ($s\selectiveinsertport a$), or
-		\item injection ($s\selectiveinjectionport a$)
+		\item drop+insert ($s\selectiveinsertport a$), or
+		%\item injection ($s\selectiveinjectionport a$)
 \end{enumerate}
@@
 	$X$; where a base is a collection of regions in a (mereo)topology such that
-	every open region can be written as union of elements of that base.  If
-	a communication occurred and resulted in
-	$\dr{\assertionRegion}{\beliefRegion}$, there exist only two mutually
-	exclusive options: either $|\assertionRegion|=|\beliefRegion|$ or
-	$|\assertionRegion|\neq|\beliefRegion|$. 
-	\begin{itemize}
-		\item If $|\assertionRegion|\neq|\beliefRegion|$ (at the end of
-			a communication through the input-port), the two
-			regions have a different number of subregions, and,
-			then, there exist only two mutually exclusive options: 
-			\begin{itemize}
-				\item either $|\assertionRegion|>|\beliefRegion|$, where one
-					ore more Assertions didn't become Beliefs, which we
-					call \emph{drop port} ($s\dropport a$) since its behavior drops and
-					prevents incoming communications,
-				\item or $|\assertionRegion|<|\beliefRegion|$, where there are
-					one ore more Beliefs which have not been sent as
-					input-Assertions, which we call \emph{insertion
-					port} ($s\insertport a$) since its behavior generates new Beliefs
-					unrelated to the incoming communication. 
-			\end{itemize}
-		\item If $|\assertionRegion|=|\beliefRegion|$, supposing an increasing monotonic entropy on the information (Assertions) exchanged, 
-				there exist only two mutually exclusive options: 
-			\begin{itemize}
-				\item either information has been generated as Belief and transferred as Assertions, increasing the cardinality of both
-					$\assertionRegion$ and $\beliefRegion$, i.e.
-					$|\assertionRegion_{t^0}|>|\assertionRegion_{t^1}|$ and
-					$|\beliefRegion_{t^0}|>|\beliefRegion_{t^1}|$ (where
-					$t^0$ and $t^1$ represent time at the beginning of
-					communication and right after, respectively); which
-					means that Assertions have reached the port and new
-					Beliefs have been generated by the port, but no
-					correlation between the elements of the two regions exist. 
-					We call this \emph{injection port} ($s\injectionport a$) since it has modified
-					the information carried by the Assertions into new
-					corresponding Beliefs.
-				\item or $|\assertionRegion_{t^0}|=|\assertionRegion_{t^1}|$
-					and $|\beliefRegion_{t^0}|=|\beliefRegion_{t^1}|$ but
-					this implies that no communication occurred, while we
-					assumed that a communication was taking place (absurd).
-			\end{itemize}
-	\end{itemize}
-	In other words, $\dr{\assertionRegion}{\beliefRegion}\equiv\neg\overlaps{\assertionRegion}{\beliefRegion}\equiv\neg(\varphi\subseteq\assertionRegion\wedge\varphi\subseteq\beliefRegion)\equiv\varphi\not\subseteq\assertionRegion\vee\varphi\not\subseteq\beliefRegion$
-	Similarly, 
-	\begin{itemize}
-		\item given that $\po{\assertionRegion}{\beliefRegion}$ is equivalent to
-			$\dr{\assertionRegion}{\beliefRegion}$ except for the
-			overlapping part (see Appendix~\ref{app:selectiveinjectionport}) in which case is
-			$\eq{\assertionRegion}{\beliefRegion}$ (see
-			\autocite{Santaca2016abf}),
-			we call it a \emph{selective injection port} ($s\selectiveinjectionport a$); otherwise
-		\item $\pp{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Leftarrow\beliefRegion$ and
-			$\ppi{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Rightarrow\beliefRegion$ are subcases of
-			$\eq{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Leftrightarrow\beliefRegion$, and then
-		\begin{itemize}
-			\item if $\pp{\assertionRegion}{\beliefRegion}$, we have that
-				some Beliefs are not correlated to incoming Assertions
-				and we call it \emph{selective insertion port} ($s\selectiveinsertport a$),
-			\item if $\ppi{\assertionRegion}{\beliefRegion}$ we have that
-				some incoming Assertions are not correlated to Belief 
-				and we call it \emph{selective drop port} ($s\selectivedropport a$)
-		\end{itemize}
-	\end{itemize}
+	every open region can be written as union of elements of that base. 
+	We distinguish between Regions that contain Information and Regions that
+	don't by writing the latter as $\emptyset$.
+	\begin{enumerate}
+		\item if $\eq{\assertionRegion}{\beliefRegion})$ then either $\assertionRegion=\beliefRegion=\emptyset$ (no communication) or $\assertionRegion=\beliefRegion\neq\emptyset$ (forward communication)
+		\item if $\dr{\assertionRegion}{\beliefRegion})$ then $\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$ (we call \emph{insert} the former and \emph{full drop} the latter case), or $\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$ which we call \emph{replace} (i.e.
+drop and insert)
+		\item if $\pp{\assertionRegion}{\beliefRegion})$ then $\beliefRegion$ contains and extend $\assertionRegion$ which we call \emph{injection}
+		\item if $\ppi{\assertionRegion}{\beliefRegion})$ then $\assertionRegion$ contains and extend $\beliefRegion$ which we call \emph{selective drop}
+		\item if $\po{\assertionRegion}{\beliefRegion})$ then a part of the $\assertionRegion$ is contained in the $\beliefRegion$ which is a combination of \emph{selective drop and generation}
+	\end{enumerate}
+
+	%	If
+	%	a communication occurred and resulted in
+	%	$\dr{\assertionRegion}{\beliefRegion}$, there exist only two mutually
+	%	exclusive options: either $|\assertionRegion|=|\beliefRegion|$ or
+	%	$|\assertionRegion|\neq|\beliefRegion|$. 
+	%	\begin{itemize}
+	%		\item If $|\assertionRegion|\neq|\beliefRegion|$ (at the end of
+	%			a communication through the input-port), the two
+	%			regions have a different number of subregions, and,
+	%			then, there exist only two mutually exclusive options: 
+	%			\begin{itemize}
+	%				\item either $|\assertionRegion|>|\beliefRegion|$, where one
+	%					ore more Assertions didn't become Beliefs, which we
+	%					call \emph{drop port} ($s\dropport a$) since its behavior drops and
+	%					prevents incoming communications,
+	%				\item or $|\assertionRegion|<|\beliefRegion|$, where there are
+	%					one ore more Beliefs which have not been sent as
+	%					input-Assertions, which we call \emph{insertion
+	%					port} ($s\insertport a$) since its behavior generates new Beliefs
+	%					unrelated to the incoming communication. 
+	%			\end{itemize}
+	%		\item If $|\assertionRegion|=|\beliefRegion|$, supposing an increasing monotonic entropy on the information (Assertions) exchanged, 
+	%				there exist only two mutually exclusive options: 
+	%			\begin{itemize}
+	%				\item either information has been generated as Belief and transferred as Assertions, increasing the cardinality of both
+	%					$\assertionRegion$ and $\beliefRegion$, i.e.
+	%					$|\assertionRegion_{t^0}|>|\assertionRegion_{t^1}|$ and
+	%					$|\beliefRegion_{t^0}|>|\beliefRegion_{t^1}|$ (where
+	%					$t^0$ and $t^1$ represent time at the beginning of
+	%					communication and right after, respectively); which
+	%					means that Assertions have reached the port and new
+	%					Beliefs have been generated by the port, but no
+	%					correlation between the elements of the two regions exist. 
+	%					We call this \emph{injection port} ($s\injectionport a$) since it has modified
+	%					the information carried by the Assertions into new
+	%					corresponding Beliefs.
+	%				\item or $|\assertionRegion_{t^0}|=|\assertionRegion_{t^1}|$
+	%					and $|\beliefRegion_{t^0}|=|\beliefRegion_{t^1}|$ but
+	%					this implies that no communication occurred, while we
+	%					assumed that a communication was taking place (absurd).
+	%			\end{itemize}
+	%\end{itemize}
+	%In other words, $\dr{\assertionRegion}{\beliefRegion}\equiv\neg\overlaps{\assertionRegion}{\beliefRegion}\equiv\neg(\varphi\subseteq\assertionRegion\wedge\varphi\subseteq\beliefRegion)\equiv\varphi\not\subseteq\assertionRegion\vee\varphi\not\subseteq\beliefRegion$
+	%Similarly, 
+	%\begin{itemize}
+	%	\item given that $\po{\assertionRegion}{\beliefRegion}$ is equivalent to
+	%		$\dr{\assertionRegion}{\beliefRegion}$ except for the
+	%		overlapping part (see Appendix~\ref{app:selectiveinjectionport}) in which case is
+	%		$\eq{\assertionRegion}{\beliefRegion}$ (see
+	%		\autocite{Santaca2016abf}),
+	%		we call it a \emph{selective injection port} ($s\selectiveinjectionport a$); otherwise
+	%	\item $\pp{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Leftarrow\beliefRegion$ and
+	%		$\ppi{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Rightarrow\beliefRegion$ are subcases of
+	%		$\eq{\assertionRegion}{\beliefRegion}\equiv\assertionRegion\Leftrightarrow\beliefRegion$, and then
+	%	\begin{itemize}
+	%		\item if $\pp{\assertionRegion}{\beliefRegion}$, we have that
+	%			some Beliefs are not correlated to incoming Assertions
+	%			and we call it \emph{selective insertion port} ($s\selectiveinsertport a$),
+	%		\item if $\ppi{\assertionRegion}{\beliefRegion}$ we have that
+	%			some incoming Assertions are not correlated to Belief 
+	%			and we call it \emph{selective drop port} ($s\selectivedropport a$)
+	%	\end{itemize}
+	%\end{itemize}
 \end{proof}
@@
 sake of simplicity, we consider, in this text, channels to be bottom-channels,
-defined as agents with the pre-defined behavior (i.e. defined in a dogmatic
+defined as agents with the pre-defined behavior (i.e. defined in a axiomatic
 way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
@@
 Assertions that are transferred over the channel, or filter out some
-input-assertions.}.
+input-assertions.}. A port is, in fact, syntactic sugar to express the 
+relation between Assertions and Beliefs.
 
@@
 
-\Paragraph{Channel Weaknesses}
+\paragraph{Channel Weaknesses}
 Given that a mono-directional bottom-channel is assumed to be perfectly
@@
 
-\begin{enumerate}[start=5, label={W\arabic*)}]
+\begin{enumerate}[start=6, label={W\arabic*)}]
 	\item \emph{secure output port} and \emph{input drop port} ($s\sdcom r$), 
@@
 
+\paragraph{Ports and Channels -- Discussion} 
+It is important to note that all the results of the application of the
+$\abftheory$ to channels is the analysis of the relation
+$\rcc{\assertionRegion_o}{\assertionRegion_i}$ leads to the same results of the
+analysis of a pair one input and one output port (i.e.
+$\rcc{\beliefRegion_1}{\assertionRegion_2}\wedge\rcc{\assertionRegion_2}{\beliefRegion_3}$
+The same result can be obtained by analyzing the relation
+between the outputs of a functional block and the inputs of
+another functional block.
 As depicted in Figure~\ref{fig:atom}, so far, we have considered Information
@@
 syntactic sugar.
+
+We can summarize these results by saying that the relations
+between Assertions and Beliefs, Output Assertions of an Agent and 
+Input Assertions of another Agent, or Output Beliefs of a 
+Functional Block and Input Beliefs of another Block can \emph{only}
+be affected by the following weaknesses: replace,
+drop, injection, insertion, selective drop, and selective drop + insertion.
+We can, then, propose the following security properties:
+\begin{itemize}
+	\item \emph{Order-preserving} -- it shall be known if Information is \emph{replaced}
+	\item \emph{Availability} -- it shall be known if Information is \emph{dropped} or \emph{selectively dropped}
+	\item \emph{Integrity} -- it shall be known if Information is \emph{injected}
+	\item \emph{Authentication} -- it shall be known if Information is \emph{inserted}
+\end{itemize}
 
@@
 its own purpose based on functional requirements. 
+Therefore, while a port semantics is determined by the relation 
+between Assertions and Beliefs, the semantics of a Functional Block 
+is determined by the relation between Facts/Requirements and I/O Beliefs.
+
 In other words, a functional block is a generic agent with no pre-defined general
 behavior (while ports and channels have a pre-defined behavior).
-Therefore, we can consider the relation between 
-the outputs of a functional block with respect to the requirements.
-One may, of course, argue that since there is a
-transformation process in the functional block the quality of that
-transformation process is what the theory should consider. This can
-be argued not just for functional blocks (Behaviors) but also for Channels
-(Assertions). The verification of the semantics of functional blocks and of the 
-protocols executed over the channels will be considered at design level while
-now we are still defining the specification.
 
@@
 \end{itemize}
-These weaknesses will be detailed afterwards in this section.
 
-As we detail afterwards, the relations considered so
-far revolves around the concept of integrity or availability, while the reminder of this section 
-will be focused on a broader analysis of security properties and requirements;
-which will also consider other security properties, such as availability and
-confidentiality.  In order to define the quality of the internals of a functional block or channel, we first
-need to relate the concept of Facts to engineering requirements.
-
-\Paragraph{Requirements as Facts}
+\paragraph{Requirements as Facts}
 During the specification phase, for any agent, channel, port, functional block
@@
 
-\Paragraph{Methodology}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.3\textwidth]{sum.pdf}
+	\caption{Example of a Functional Block}
+	\label{fig:sum}
+\end{figure}
+
+As an example, consider a functional block that performs the summation of two
+inputs, as in Figure~\ref{fig:sum}, defined by the requirement $r := b_3 = b_1
++ b_2$. The possible relations between the behavior of the functional block and
+the requirements (i.e. $\Rcc{\behaviorRegion}{\factRegion}$ is determined by
+the relations between the I/O Beliefs $sum(b_1,b_2)=b_3$ and the requirement
+$sum(b_1,b_2)=b_1+b_2$, as follows.
+\begin{itemize}
+	\item $\eq{\beliefRegion}{\factRegion}=\eq{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, 
+		where $\beliefRegion_3$ represents the Region of all the
+		outputs of the functional block $sum$ while the Region
+		$\beliefRegion_1+\beliefRegion_2$ represents the expected
+		outputs of an ideal implementation of the requirement $r$.
+		Therefore, the functional block correctly implements the
+		requirements.
+	\item
+		$\dr{\beliefRegion}{\factRegion}=\dr{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$,
+		the function block never implements the requirements.
+	\item
+		$\pp{\beliefRegion}{\factRegion}=\pp{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exist some inputs for which
+		the output is incorrect.
+	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exists some outputs that are not the result of the summation of the inputs but given the correct inputs the function always outputs correctly.
+	\item
+		$\po{\beliefRegion}{\factRegion}=\po{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$,
+		the component has a Byzantine behavior where occasionally
+		outputs the expected output given the correct inputs. Not all
+		the inputs are handled properly, nor all the expected outputs
+		are always generated when correct inputs are given.
+\end{itemize}
+
+\paragraph{Methodology}
 We have now mapped the building blocks of our theory, namely Information,
@@
 
-\Paragraph{Availability and Integrity (Architecture Security)}
+\paragraph{Availability and Integrity (Architecture Security)}
 \begin{figure}[t]
@@
 
-\Paragraph{Secrecy, Authenticity, and Key Agreement (Cryptographic Security)}\label{sec:encryption}
+\paragraph{Secrecy, Authenticity, and Key Agreement (Cryptographic Security)}\label{sec:encryption}
 As Maurer puts it (\autocite{Maurer1999information}), there exist two types of
@@
 
-\Paragraph{Confidentiality Weaknesses}
+\paragraph{Confidentiality Weaknesses}
 In a functional block, there exists the following weaknesses in the functional
@@
 of an agent based on the number $r$ of relations (e.g. EQ, DR, PO) with arity $k$ between
-$n$ diffrent sets in the $\abftheory$ theory. $\delta_T$ represents the 
+$n$ different sets in the $\abftheory$ theory. $\delta_T$ represents the 
 number of configurations which are not satisfiable with respect to the logical theory

### 2020-06-09 reports/confidential/insecureport.tex
mark white paper as confidential

--- 2020-03-30 insecureport.tex
@@
-\fix{mr}{this proof needs to be done}
+\fix{mr}{this proof needs to be done. The current text
+is not sufficient} 
 \begin{proof}
-In Appendix~\ref{app:commproof}, we report the proof which is exhaustive on all
-possible combinations.  In the following we report only the excluded cases and
-assume the other to be strightforwad (as a proof sketch).
 \begin{enumerate}

### 2020-06-09 reports/confidential/problem_statement.tex
mark white paper as confidential

--- 2020-05-25 reports/journal_paper/problem_statement.tex
@@
 
-%Either a valid security hypothesis can be formulated based on one or more of
-%the concepts reviewed (and depicted in Figure~\ref{fig:safety-security}) or it
-%can't.  Since, in the next Section, we will define our security hypothesis (on
-%top of which we'll develop a security theory) on those terms, if the second case
-%is true we will not formulate a valid security theory. However, while we want
-%to highlight that the incompleteness of this review may be a potential issue;
-%on the other hand, due to the nature of a literature review, it seems to us
-%that only an inductive approach can be used and then any literature review will
-%be incomplete.  If the first case applies, in one of those terms one shall find
-%a valid ground on top of which formulate a security hypothesis (assuming that a
-%valid security hypothesis exists).
-
 Our literature review shows that most of the definitions relates insecurity to
 dis-honesty (also called maliciousness or adversarial) of an agent (often
-called adversary or attacker). We can synthesize this belief in the
-following hypothesis.
-
-\begin{hypothesis}{\bf Security as Honesty and Insecurity as Dishonesty --}
-	\begin{itemize}
-		\item a closed system can be considered \emph{secure} if there
-			is no dishonest agent
-		\item a closed system can be considered \emph{insecure} if
-			there is a dishonest agent 
-	\end{itemize}
-	where an agent is any virtual or physical entity of the system or using
-	the system (e.g. a device, a software, or a human being) and dishonesty
-	is not necessarily related to malicious motivation but also to
-	incompetence or lack of skills.
-\end{hypothesis}
+called adversary or attacker). This, however, just moves the problem of
+defining what cybersecurity is as the problem of defining what an dishonest
+agent can or cannot do in a system\footnote{where an agent is any virtual or
+physical entity of the system or using the system (e.g. a device, a software,
+or a human being) and dishonesty is not necessarily related to malicious
+motivation but also to incompetence or lack of skills.}
 As in the Dolev-Yao theory, we may correlate ``being dishonest'' to ``not
@@
 (to name a few).
-
-We now proceed with a more detailed analysis of the terms
-and with the re-formulation of the security hypothesis.

### 2020-06-09 reports/confidential/selectiveinjectionport.tex
mark white paper as confidential

\begin{displaymath}
\po{x}{y}=\overlaps{x}{y}\wedge\neg\partof{x}{y}\wedge\neg\partof{y}{x}\\
\end{displaymath}
and
\begin{align*}
\overlaps{x}{y}=\exists z.~ &\partof{z}{x} &\wedge& \partof{z}{y}\\
&\forall z'.~z'\subseteq z\Rightarrow z'\subseteq x && \forall z''.~z''\subseteq z\Rightarrow z'\subseteq y
\end{align*}
and
\begin{equation*}
\begin{split}
\neg\partof{x}{y}=&\neg[\forall\varphi.~(\varphi\subseteq x)\Rightarrow(\varphi\subseteq y)]\\
&\neg[\forall\varphi.~\neg(\varphi\subseteq x)\vee(\varphi\subseteq y)]\\
&\exists\varphi.~(\varphi\subseteq x)\wedge(\varphi\not\subseteq y)
\end{split}
\end{equation*}
Therefore
\begin{displaymath}
\po{\assertionRegion}{\beliefRegion}=(\varphi\subseteq\assertionRegion\wedge\varphi\subseteq\beliefRegion)\wedge(\varphi'\subseteq\assertionRegion\wedge\varphi'\not\subseteq\beliefRegion)\wedge(\varphi''\subseteq\assertionRegion\wedge\varphi''\subseteq\beliefRegion)
\end{displaymath}
Which means that, when a communication over an input-port results in $\po{\assertionRegion}{\beliefRegion}$, the port is:
\begin{itemize}
\item correctly receiving incoming communication as Assertion and sending it to the recipient agent as Beliefs, \emph{and} ($s\rightarrow a$)
\item incorrectly, dropping some input communication (Assertions) ($s\dropport a$), \emph{and}
\item incorrectly, inserting some new Beliefs ($s\insertport a$)
\end{itemize}
Which we abbreviate as a \emph{selective injection} port: $s\selectiveinjectionport a$.

### 2020-06-11 prototypes/alpha_0/2a2w.py
remove test code

@@ -1,271 +1,0 @@
-#!/usr/bin/python3
-#author: Marco Rocchetto @V-Research
-
-#requirements
-# 1) z3 python API (https://github.com/Z3Prover/z3/blob/master/README.md)
-
-from z3 import *
-import scipy.special
-import itertools
-from multiprocessing import Process
-from multiprocessing import Queue 
-
-################################
-# since we unfold the quantifiers we have to calculate
-# the (minimal) number of subsets of A, B, and F that allows
-# Z3 to find a model that satisfies the agent configuration
-# 
-# For example, if we have A,B,F and no subsets of those 3 sets 
-# there is no model that satisfies PO(A,B) /\ PO(B,F) /\ PO(A,F)
-# since PO(A,B) <-> PP(A,B) if A has no subsets 
-#
-# requires scipy for binomial calculation: pip install scipy
-# otherwise download wheel from https://pypi.python.org/pypi/scipy
-# and install it with pip install <downloaded-file>
-################################
-def calculate_sets(sets, Set):
-    set_of_subsets=[]
-    for k in range(1,len(sets)):
-        counter=0
-        for i in range(int(scipy.special.binom(len(sets), (k+1)))):
-            subset="subset_%s_%d"%((k+1), counter)
-            set_of_subsets.append(Const(subset, Set))
-            #print("%d - (%d %d)=%d"%(i,len(sets),(k+1),int(scipy.special.binom(len(sets), (k+1)))))
-            counter+=1
-        
-    return set_of_subsets
-
-def topology(solver, sets_and_subsets, P):
-    ################################
-    # PART OF
-    ################################
-    
-    for s in sets_and_subsets:
-        #solver.add(P(s,s))
-        solver.assert_and_track(P(s,s), str("reflexivity(%s)"%s))
-    
-    for s1 in sets_and_subsets:
-        for s2 in sets_and_subsets:
-            #solver.add(Implies(And(P(s1,s2),P(s2,s1)), s1==s2))
-            solver.assert_and_track(Implies(And(P(s1,s2),P(s2,s1)), s1==s2), str("asymmetry(%s,%s)"%(s1,s2)))
-    
-    for s1 in sets_and_subsets:
-        for s2 in sets_and_subsets:
-            for s3 in sets_and_subsets:
[…]

### 2020-06-11 prototypes/alpha_0/2a2w_sequential.py

@@ -1,252 +1,0 @@
-#!/usr/bin/python3
-#author: Marco Rocchetto @V-Research
-
-#requirements
-# 1) z3 python API (https://github.com/Z3Prover/z3/blob/master/README.md)
-
-from z3 import *
-import scipy.special
-import itertools
-
-################################
-# since we unfold the quantifiers we have to calculate
-# the (minimal) number of subsets of A, B, and F that allows
-# Z3 to find a model that satisfies the agent configuration
-# 
-# For example, if we have A,B,F and no subsets of those 3 sets 
-# there is no model that satisfies PO(A,B) /\ PO(B,F) /\ PO(A,F)
-# since PO(A,B) <-> PP(A,B) if A has no subsets 
-#
-# requires scipy for binomial calculation: pip install scipy
-# otherwise download wheel from https://pypi.python.org/pypi/scipy
-# and install it with pip install <downloaded-file>
-################################
-def calculate_sets(sets, Set):
-    set_of_subsets=[]
-    for k in range(1,len(sets)):
-        counter=0
-        for i in range(int(scipy.special.binom(len(sets), (k+1)))):
-            subset="subset_%s_%d"%((k+1), counter)
-            set_of_subsets.append(Const(subset, Set))
-            #print("%d - (%d %d)=%d"%(i,len(sets),(k+1),int(scipy.special.binom(len(sets), (k+1)))))
-            counter+=1
-        
-    return set_of_subsets
-
-def topology(solver, sets_and_subsets, P):
-    ################################
-    # PART OF
-    ################################
-    
-    for s in sets_and_subsets:
-        #solver.add(P(s,s))
-        solver.assert_and_track(P(s,s), str("reflexivity(%s)"%s))
-    
-    for s1 in sets_and_subsets:
-        for s2 in sets_and_subsets:
-            #solver.add(Implies(And(P(s1,s2),P(s2,s1)), s1==s2))
-            solver.assert_and_track(Implies(And(P(s1,s2),P(s2,s1)), s1==s2), str("asymmetry(%s,%s)"%(s1,s2)))
-    
-    for s1 in sets_and_subsets:
-        for s2 in sets_and_subsets:
-            for s3 in sets_and_subsets:
-               #solver.add(Implies(And(P(s1,s2),P(s2,s3)), P(s1,s3)))
[…]

### 2020-06-12 reports/paper_0/main.tex
change ref to images in paper 0
add weakness table and a-f relation to report_0

--- 2020-06-09 reports/conference_paper/main.tex
@@
 \usepackage{graphicx}
-\graphicspath{{../images/}}
+\graphicspath{{./images/}}
 \usepackage{multirow}
@@
 
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+
 \begin{document}
@@
 	\centering
-	\includegraphics[width=.9\textwidth]{safety-security_3.pdf}
+	\includegraphics[width=.8\textwidth]{safety-security_3.pdf}
 	\caption{Overview of keywords related to security and safety}
@@
 	\centering
-	\includegraphics[width=.7\columnwidth]{causality.pdf}
+	\includegraphics[width=.6\columnwidth]{causality.pdf}
 	\caption{Etiology of cybersecurity}
@@
 difficult to define
-\autocite{Gettier2012knowledge}\footnote{``\emph{Theaetetus}: [\ldots] He said
-that knowledge was true opinion accompanied by reason, but that unreasoning
-true opinion was outside of the sphere of knowledge; and matters of which there
-is not a rational explanation are unknowable -- yes, that is what he called
-them -- and those of which there is are knowable. [\ldots] \emph{Socrates}:
-[\ldots] the primary elements of which we and all else are composed admit of no
-rational explanation; for each alone by itself can only be named, and no
-qualification can be added, neither that it is nor that it is not, for that
-would at once be adding to it existence or non-existence, whereas we must add
-nothing to it, if we are to speak of that itself alone.  [\ldots]'' Plato --
-Theaetetus 201 \autocite{Plato1914Plato}}.  For example, Hintikka in
+\autocite{Gettier2012knowledge}.
+%\footnote{``\emph{Theaetetus}: [\ldots] He said
+%that knowledge was true opinion accompanied by reason, but that unreasoning
+%true opinion was outside of the sphere of knowledge; and matters of which there
+%is not a rational explanation are unknowable -- yes, that is what he called
+%them -- and those of which there is are knowable. [\ldots] \emph{Socrates}:
+%[\ldots] the primary elements of which we and all else are composed admit of no
+%rational explanation; for each alone by itself can only be named, and no
+%qualification can be added, neither that it is nor that it is not, for that
+%would at once be adding to it existence or non-existence, whereas we must add
+%nothing to it, if we are to speak of that itself alone.  [\ldots]'' Plato --
+%Theaetetus 201 \autocite{Plato1914Plato}}.
+For example, Hintikka in
 \autocite{Hintikka1993Information} states that ``A purely logical definition of
@@
 	\centering
-	\includegraphics[width=\columnwidth]{soundness.pdf}
+	\includegraphics[width=.9\columnwidth]{soundness.pdf}
 	\caption{Example relation between Facts, and Assertions and Beliefs}
@@
 	\centering
-	\includegraphics[width=.9\textwidth]{mashyp.pdf}
-	\caption{asd}
+	\includegraphics[width=.8\textwidth]{mashyp.pdf}
+	\caption{Mapping Epistemological Concepts to (Cyber-Physical) Systems Engineering}
 	\label{fig:mashyp}
@@
 by analyzing the relation between the outputs of a functional block and the
-inputs of another functional block, where functional blocks are constituents of
+inputs of another functional block ($\Rcc{\beliefRegion_o}{\beliefRegion_i}$), 
+where functional blocks are constituents of
 the functional architecture as described afterwards.  As depicted in
@@
 
+\paragraph{Assertions and Facts}
+The whole reasoning on the relation between Beliefs and Facts can be duplicated
+for the relation between Assertions and Facts; we cannot appreciate the
+difference at this level of abstraction.  If the functional architecture would
+be extended to capture the semantics (i.e. the logic) of the communication and
+security protocols, with the relation between Assertions and Facts we would
+compare protocol logics with the requirements. We consider the verification of
+the functional architecture and protocol logic in subsequent steps of the
+engineering process (as we describe in Section~\ref{sec:discussion}).
+Therefore, we can summarize our results by categorizing the weaknesses
+predicted by our theory into: data-flow-related and functionality-related
+weaknesses; as in Table~\ref{tab:weaknesses}.
+\begin{table*}[t]
+\centering
+	\begin{tabular}{c C{.35\textwidth} C{.5\textwidth}} 
+		\textbf{RCC5} & \textbf{Quantity: Data Flow} & \textbf{Quality: Requirements Adherence}\\
+		\hline 
+		EQ & Expected/Nominal & Expected/Nominal\\[.1cm]
+	DR & Drops all the inputs and inserts new malicious data & The component never performs/carries the expected behavior/information\\[.1cm]
+	PP & Selectively drops inputs & Part of the expected outputs are not generated in response to the correct inputs \\[.1cm]
+	PPi& Forwards all the inputs but crafts and inserts new malicious data & The components correctly performs/carries the expected behavior/information when the correct inputs are provided but is subject to input injections \\[.1cm]
+	PO & Selectively drops inputs and inserts new data & Byzantine behavior - occasionally outputs the expected output given the correct inputs.
+Not all the inputs are handled properly, nor all the expected outputs are always generated when correct inputs are given
+\end{tabular}
+\caption{Weaknesses Categorized~\label{tab:weaknesses}}
+\end{table*}
+
 \subsection{Security and Insecurity of a System}
@@
 	\centering
-	\includegraphics[width=.7\columnwidth]{eng_cs1.png}
+	\includegraphics[width=.6\columnwidth]{eng_cs1.png}
 	\caption{Water Level Reader Deployment Diagram}
@@
 	\centering
-	\includegraphics[width=.7\textwidth]{internal_cs1.png}
+	\includegraphics[width=.6\textwidth]{internal_cs1.png}
 	\caption{Sensor Board Object Diagram}
@@
 	\centering
-	\includegraphics[width=\textwidth]{results_excel.pdf}
+	\includegraphics[width=.7\textwidth]{results_excel.pdf}
 	\caption{Partial View of the results in the Excel file}
@@
 
+\appendix
+%\section{Engineering $\abftheory$-theory}
+The Class Diagram for the Engineering of the $\abftheory$-theory is reported in
+Figure~\ref{fig:secraclassdiagram}. A \emph{specification} of a CPS is viewed
+as an aggregation of \emph{architectures} which can describe the functional or
+physical requirements. The physical components of the architecture are
+input/output \emph{ports} and \emph{channels} (aggregations of pairs of ports)
+while \emph{functional blocks} are the only constituents of the functional
+architecture. All of the classes are abstract except input/output ports and
+functional blocks. Therefore, agents (which represents sub-systems or
+components) are composed by ports and functional blocks, as an aggregation of 
+architectures.
+
+\begin{sidewaysfigure*}
+	\centering
+	\includegraphics[width=\textwidth]{secra_classDiagram.png}
+	\caption{Security Risk Assessment -- Class Diagram}
+	\label{fig:secraclassdiagram}
+\end{sidewaysfigure*}
+\printbibliography
+
 % trigger a \newpage just before the given reference
@@
 %\bibliography{../literature}
-\appendix
-\begin{sidewaysfigure*}
-	\centering
-	\includegraphics[width=\textwidth]{secra_classDiagram.png}
-	\caption{Security Risk Assessment -- Class Diagram}
-	\label{fig:secraclassdiagram}
-\end{sidewaysfigure*}
-\printbibliography
 

### 2020-06-25 reports/paper_0/main.tex
implement Luca's feedback

@@
 \newtheorem{hypothesis}{Hypothesis}%[section]
+\newtheorem{theorem}{Theorem}%[section]
 
@@
 \section{Introduction}\label{sec:intro}
+A scientific theory is an explanation of a phenomenon such that the explanation
+follows the scientific method. The scientific method is an \emph{empirical}
+method that aims at mitigating potential fallacies in theories.  K. Popper
+famously argued (e.g. in \autocite{Popper1959logic}) that a scientific theory
+can never been verified but only falsified, that a theory should not be
+conceived by using the principle of induction\footnote{A.  Einstein wrote to K.
+Popper: ``[...] and I think (like you, by the way) that theory cannot be
+fabricated out of the results of observation, but that it can only be
+invented.''\autocite{Popper1959logic}}, and that empirical experiments should
+be considered as the only evidence to support the non-falseness of a scientific
+theory. If we consider a natural (real-world) phenomenon, such as gravity, we
+can think of a scientific theory as an \emph{abstract} mathematical formula (or
+a set of formulas) that axiomatically define a certain (natural) phenomenon in
+a specific algebraic structure. While such a theory can be proved to be correct
+within the mathematical setting it has been conceived in (e.g. with respect to
+other axioms of the algebraic structure where it has been postulated, or with
+respect to other background theories), no proof (for soundness or completeness)
+can be carried out to verify that the theory correctly abstracts  the
+phenomenon it describes. Similarly, the results of the application of a
+scientific theory (i.e. its predictions) can be verified in the abstract
+mathematical structure but how can we be sure that the theory correctly
+characterize the phenomenon it describes? With his famous ``demarcation
+principle'', K. Popper draw a line between scientific and pseudo-scientific
+theories by proposing that \emph{any} scientific theory can be considered
+correct as long as no empirical experiment falsify it. For example, the theory
+of gravity, in its mathematical setting, can predict how object attracts with
+each other, and those predictions can be tested by empirical experiment. If the
+theory correctly abstracts the phenomenon no empirical experiment should
+falsify the predictions of that theory.
+
 In \autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
@@
 observation that allows us to declare an arbitrary system or technique
-secure''. Herley then uses this argument to show that ``claims that any measure
+secure''. His intuition is that there is no scientific theory that can predict
+the security of a system, nor a theory that can predict all possible
+insecurity of a system (which, by negation it would become a theory on security). 
+Herley then uses this argument to show that ``claims that any measure
 is necessary for security are empirically unfalsifiable''. Given that, any
@@
 
-We consider the problem raised by Herley not confined to
-``computer security'' but to any abstract system (so that our theory may hold
-for any sound implementation such as networks, mechanical, cyber, or
-cyber-physical system, or even a single computer or a single device such as an
-hard-drive).  There is also an apparent inconsistency in
-\autocite{Herley2016unfalsifiability} that we seek to clarify before following (as
-we agree) the scientific path draw by Herley: cybersecurity is defined as an
-abstract property in many formal approaches to the investigation of the
-security of systems, and the security of the design of a formally verified
-protocol is indeed falsifiable (against the security properties verified).  For
-example, in the protocol verification community, security is often defined as a
-formalization of the high-level properties confidentiality, integrity, and
-availability. The problem in such approaches is not the definition of what
-cybersecurity is but the generality of the results, since they tackle a specific
-step (and not the first) of the engineering process (of systems).  Therefore,
-the theories underlying the verification are based on assumptions which
+We consider the problem raised by Herley not confined to ``computer security''
+but to any abstract system (so that our theory may hold for any sound
+implementation such as networks, mechanical, cyber, or cyber-physical system,
+or even a single computer or a single device such as an hard-drive).  There is
+also an apparent inconsistency in \autocite{Herley2016unfalsifiability} that we
+seek to clarify before following (as we agree) the scientific path draw by
+Herley: cybersecurity is defined as an abstract property in many formal
+approaches to the investigation of the security of systems, and the security of
+the design of a formally verified protocol is indeed falsifiable (against the
+security properties verified).  For example, in the protocol verification
+community, security is often defined as a formalization of the high-level
+properties confidentiality, integrity, and authentication. The problem in such
+approaches is not the definition of what cybersecurity is but the generality of
+the results, since they tackle a specific step\footnote{In fact, many standards
+(e.g. ) used for the engineering of systems starts with the requirements
+definition, and physical and functional architecture design, before the logic
+of protocols is even considered.} of the engineering process (of systems)
+without considering many aspects related to the implementation of a protocol
+that may generate security vulnerabilities later in the development process. An
+infamous example is the KRAK attack on the WPA2 4-way handshake protocol where
+the authors wrote\autocite{krak2020web}: ``our attacks do not violate the security properties proven
+in formal analysis of the 4-way handshake'' but ``the problem is that the
+proofs do not model key installation.''.
+The theories underlying the verification of security protocols are based on assumptions which
 non-evidently apply to a general security theory.  As an example, the so called
@@
 the $\abf$-theory) and, based on that definition, what the security risk is and
-how to quantify it in a risk matrix.
+how to quantify it in a risk matrix (where a risk matrix is usually defined as
+a function between the likelihood and the impact of all potential attacks).
 \begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
@@
 \end{definition}
-The risk matrix is often defined as a function between 
-likelihood and impact of attacks. We suggest
+While a risk matrix is often defined as a function between 
+likelihood and impact of attacks, we suggest
 that security weaknesses are all equally likely to be exploited
@@
 
-\paragraph{Port Weaknesses}
+\begin{theorem}{Port Weaknesses.}
 From our definition of ports, it follows that there exist only the following \emph{six}
 types of weaknesses, generating six types of insecure port in RCC5 (the
-notation is reported for input-ports, i.e. on the lef-hand side of the arrow):
+notation is reported for input-ports, i.e. on the left-hand side of the arrow):
 \begin{enumerate}[start=1, label={W\arabic*)}]
-	\item \emph{replace port} ($s\dropport a$), where assertions reaches
+	\item \emph{replace port}, where assertions reaches
 		the port but do not pass the boundary of the agent (i.e. do not
 		become belief of the agent)
-	\item \emph{drop port} ($s\dropport a$), where assertions reaches the
+	\item \emph{drop port}, where assertions reaches the
 		port but do not pass the boundary of the agent (i.e. do not
 		become belief of the agent)
-	\item \emph{insertion port} ($s\insertport a$), where some new
+	\item \emph{insertion port}, where some new
 		information is believed by $a$ as incoming from the port but
 		$s$ didn't send it
-	\item \emph{injection port} ($s\injectionport a$), where information
+	\item \emph{injection port}, where information
 		coming from $s$ is substituted with new information which
@@
 		part is either:
-	\begin{enumerate}[start=1, label={W4.\arabic*)}]
-		\item drop ($s\selectivedropport a$), 
-		\item drop+insert ($s\selectiveinsertport a$).
+	\begin{enumerate}[start=1, label={W5.\arabic*)}]
+		\item drop, 
+		\item drop+insert.
 	\end{enumerate}
 \end{enumerate}
+\end{theorem}
 
@@
 \begin{enumerate}[start=6, label={W\arabic*)}]
-	\item \emph{secure output port} and \emph{input drop port} ($s\sdcom r$), 
-	\item \emph{secure output port} and \emph{input insertion port} ($s\sicom r$), 
-	\item \emph{output drop port} and \emph{input drop port} ($s\ddcom r$)
-	\item \emph{output drop port} and \emph{input insertion port} ($s\dicom r$)
-	\item \emph{output drop port} and \emph{input secure port} ($s\dscom r$)
-	\item \emph{output injection port} and \emph{input secure port} ($s\jscom r$)
+	\item \emph{secure output port} and \emph{input drop port}
+	\item \emph{secure output port} and \emph{input insertion port}
+	\item \emph{output drop port} and \emph{input drop port}
+	\item \emph{output drop port} and \emph{input insertion port}
+	\item \emph{output drop port} and \emph{input secure port}
+	\item \emph{output injection port} and \emph{input secure port}
 \end{enumerate}
@@
 We thank Katia Santac\`a for the help in the development of the initial ideas, and for
-the helpful discussions that allowed us to make this step.
+the helpful discussions that allowed us to make this step forward.
 

### 2020-07-13 reports/paper_0/main.tex
Luca

@@
 
+\newcommand{\fixn}[2]{{\bf FIX}\footnote{#1: #2}}
+
 \newtheorem{definition}{Definition}%[section]
@@
 \begin{abstract}
+\fixn{Luca}{Per il titolo, meglio: ``On the Etiology of Cybersecurity'' oppure ``The Etiology of Cybersecurity''.
+}
 The objective of this research is the development of a theory that defines (all
@@
 	abstract system. The theory is structured upon other theories that
-	defines how a component of a system can be abstracted into an agent,
+	define how a component of a system can be abstracted into an agent,
 	defining how agents can be formalized (both syntactically and
-	semantically) to describe an abstract system, such as a graph. The core
+	semantically) to describe an abstract system, such as a graph.\fixn{Luca}{la frase subordinata ``defining...'' non si capisce a cosa si riferisca. Troppi ``define'' e troppo lunga la frase.} The core
 	theories are the epistemological definition of knowledge, Beliefs, and
 	Information, the Assertion-Belief-Fact framework, and mereotopology.
-	We implemented a formal theory (of axioms) of a mereotopology, and of
+	We\fixn{Luca}{Non si capisce come questa frase si colleghi a quanto sopra. E se e come questa formal theory sia related a sopra} implemented a formal theory (of axioms) of a mereotopology, and of
 	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
-	uses the Z3 SMT solver. The results show that a single component (i.e.
+	uses the Z3 SMT solver. The results\fixn{Luca}{The results of what?} show that a single component (i.e.
 	agent) of an abstract system has a definite number of  different
@@
 	configurations are reported as models satisfying the abstract system
-	semantics. We use these results to predict all possible security
+	semantics.\fixn{Luca}{Non si capisce la relazione tra agent e abstract system semantics} We use these results to predict all possible security
 	weaknesses of a system, instead of relying on databases of known
 	weaknesses, and to lay the foundations of a scientific theory of
-	cybersecurity.  We show a concrete applications of our theory to the
-	risk assessment of an ad-hoc system.
+	cybersecurity.
+We show a concrete applications of our theory for the risk assessment of an ad-hoc system.\fixn{Luca}{Io continuo a pensare che non si capisca bene cosa stiate facendo.
+\`E ancora scritto in maniera troppo complicata e vengono dette troppe cose.
+Io accorcerei e semplificherei.}
 \end{abstract}
@@
 \section{Introduction}\label{sec:intro}
-A scientific theory is an explanation of a phenomenon such that the explanation
-follows the scientific method. The scientific method is an \emph{empirical}
-method that aims at mitigating potential fallacies in theories.  K. Popper
+\paragraph{Context and motivation} 
+A \emph{scientific theory} is an explanation of a phenomenon such that the explanation
+follows the scientific method. The \emph{scientific method} is an \emph{empirical}
+method that aims at mitigating potential fallacies in theories.  Karl Popper
 famously argued (e.g. in \autocite{Popper1959logic}) that a scientific theory
-can never been verified but only falsified, that a theory should not be
-conceived by using the principle of induction\footnote{A.  Einstein wrote to K.
+can never be verified but only falsified, that a theory should not be
+conceived by using the principle of induction\footnote{Albert  Einstein wrote to 
 Popper: ``[...] and I think (like you, by the way) that theory cannot be
 fabricated out of the results of observation, but that it can only be
-invented.''\autocite{Popper1959logic}}, and that empirical experiments should
+invented.''~\autocite{Popper1959logic}}, and that empirical experiments should
 be considered as the only evidence to support the non-falseness of a scientific
-theory. If we consider a natural (real-world) phenomenon, such as gravity, we
+theory.
+
+If we consider a natural (real-world) phenomenon, such as gravity, we
 can think of a scientific theory as an \emph{abstract} mathematical formula (or
-a set of formulas) that axiomatically define a certain (natural) phenomenon in
+a set of formulas) that defines axiomatically a certain (natural) phenomenon in
 a specific algebraic structure. While such a theory can be proved to be correct
@@
 can be carried out to verify that the theory correctly abstracts the
-phenomenon it describes. Similarly, the results of the application of a
+phenomenon it describes.\fixn{Luca}{Troppi correct. Meglio chiarire bene cosa sono correctness, soundness e completeness} Similarly, the results of the application of a
 scientific theory (i.e. its predictions) can be verified in the abstract
 mathematical structure but how can we be sure that the theory correctly
-characterize the phenomenon it describes? With his famous ``demarcation
-principle'', K. Popper draw a line between scientific and pseudo-scientific
+characterizes the phenomenon it describes?
+
+Popper's famous ``demarcation
+principle'' draws a line between scientific and pseudo-scientific
 theories by proposing that \emph{any} scientific theory can be considered
 correct as long as no empirical experiment falsify it. For example, the theory
-of gravity, in its mathematical setting, can predict how object attracts with
+of gravity, in its mathematical setting, can predict how objects attract
 each other, and those predictions can be tested by empirical experiment. If the
-theory correctly abstracts the phenomenon no empirical experiment should
+theory correctly abstracts the phenomenon, no empirical experiment should
 falsify the predictions of that theory.
@@
 the security of a system, nor a theory that can predict all possible
-insecurity of a system (which, by negation it would become a theory on security). 
+insecurities of a system (which, by negation would become a theory on security). 
 Herley then uses this argument to show that ``claims that any measure
-is necessary for security are empirically unfalsifiable''. Given that, any
-theory which is not falsifiable by an empirical experiment is well
+is necessary for security are empirically unfalsifiable''.\fixn{Luca}{Whay if I have a system that I have proved to be secure under a set of $n$ assumptions.
+Now I remove those $n$ assumptions one by one and run the proof again.
+What if the proof works whenever I remove one of the first $n-1$ assumptions, but fails (only) if I remove the $n$th assumption.
+Would the counterexample, which can be applied empirically, not be evidence that the assumption removed was a necessary measure?} Given that any
+theory that is not falsifiable by an empirical experiment is well
 known\footnote{``A theory which is not refutable by any conceivable event is
@@
 think) but a vice.'' -- Karl Popper, Conjectures and
-Refutations \autocite{popper1962conjectures}} to be nonscientific (i.e.
+Refutations~\autocite{popper1962conjectures}.} to be nonscientific (i.e.
 unfalsifiability is a fallacy of a theory), Herley concludes that there is no
 scientific theory on cybersecurity; which means that cybersecurity lays in the
-realm of pseudo-sciences \autocite{Herley2016usenixvideo}.  Herley, e.g.
+realm of pseudo-sciences~\autocite{Herley2016usenixvideo}.\fixn{Luca}{Theory ON cybersecurity or theory OF cybersecurity. A me viene da dire OF}  Herley, e.g.
 in \autocite{Herley2017justifying}, discusses the implications of a
 nonscientific approach to cybersecurity, and highlights the tremendous impact on
-all the scientific research and engineering of systems; leading often to
+all the scientific research and engineering of systems; leading\fixn{Luca}{Questo suona un po' esagerato, Meglio chiarire.} often to
 terrorism and wars, and wasting of resources in useless protections or
 overspending.  While the criticism is investigated
-in \autocite{Herley2016unfalsifiability}, no solution is provided nor envisioned.  On the
-contrary, the goal of this work is to lay the foundations of a
+in \autocite{Herley2016unfalsifiability}, no solution is provided nor envisioned.
+
+\paragraph{Contributions} 
+The goal of this paper is to lay the foundations of a
 scientific cybersecurity theory. 
-
 We consider the problem raised by Herley not confined to ``computer security''
@@
 implementation such as networks, mechanical, cyber, or cyber-physical system,
-or even a single computer or a single device such as an hard-drive).  There is
+or even a single computer or a single device such as a hard-drive).\fixn{Luca}{Non si capisce ``any'' system.
+Quindi anche un sistema che non ha nulla a che fare con security?
+La frase ``not confined to...
+but to...
+'' non \`e grammaticamente corretta.
+Il verbo non pu\`o rimanere ``confined'', casomai ``not confined to ``computer security''
+but rather we consider any abstract system'' o una cosa del genere}
+
+There is
 also an apparent inconsistency in \autocite{Herley2016unfalsifiability} that we
-seek to clarify before following (as we agree) the scientific path draw by
+seek to clarify before following (as we agree) the scientific path drawn by
 Herley: cybersecurity is defined as an abstract property in many formal
@@
 the design of a formally verified protocol is indeed falsifiable (against the
-security properties verified).  For example, in the protocol verification
+security properties verified).\fixn{Luca}{Questa frase non la capisco.
+Innanzitutto non capisco la differenza tra cybersecurity e security.
+E poi cybersecurity non \`e una property ma un set of properties.
+E non si capisce come si passa da un system ad un protocol...
+sembrano due frasi diverse.
+Io le spezzerei, parlando del problema in generale, e poi dicendo, As a concrete example, consider the case of security protocols...}  For example, in the protocol verification
 community, security is often defined as a formalization of the high-level
@@
 the results, since they tackle a specific step\footnote{In fact, many standards
-(e.g. ) used for the engineering of systems starts with the requirements
+(e.g. \fixn{Luca}{manca ref}) used for the engineering of systems start with the requirements
 definition, and physical and functional architecture design, before the logic
@@
 without considering many aspects related to the implementation of a protocol
-that may generate security vulnerabilities later in the development process. An
+that may generate security vulnerabilities later in the development process.\fixn{Luca}{Qui si cercano di dire troppe cose.
+Improvvisamente, mentre si parlare di security definition, si parla del software life-cycle.
+Va spiegato molto meglio (e in maniera pi\`u semplice.) Inoltre, non \`e proprio corretto.
+Che dire di model-based testing in cui si fa analysis della specifica e dell'implementazione?} An
 infamous example is the KRAK attack on the WPA2 4-way handshake protocol where
-the authors wrote\autocite{krak2020web}: ``our attacks do not violate the security properties proven
+the authors\fixn{Luca}{The authors of what? Solo Vanhoef?} wrote~\autocite{krak2020web}: ``our attacks do not violate the security properties proven
 in formal analysis of the 4-way handshake'' but ``the problem is that the
-proofs do not model key installation.''.
+proofs do not model key installation.''\fixn{Luca}{Questo mi sembra un modo confuso e un po' specioso di dire un'ovviet\`a: i risultati di un'analisi formale sono buoni tanto quanto il modello.
+Se il modello non modella una cosa, allora non si possono dedurre conclusioni su quella cosa.
+Questo \`e un problema generico della modellizzazione di qualunque cosa, non solo di security.
+E non direi che questo sia un esempio infamous, visto che ne esistono molti altri molto pi\`u importanti.}
 The theories underlying the verification of security protocols are based on assumptions which
-non-evidently apply to a general security theory.  As an example, the so called
+non-evidently apply to a general security theory.\fixn{Luca}{Non capisco ``which
+non-evidently apply to a general security theory''}  As an example, the so called
 Dolev-Yao attacker model\footnote{For the sake of simplicity, the Dolev-Yao
@@
 applies to specific instances (often called scenarios) and abstraction of the
-protocol. This, in turn, creates a false sense of security since requires
+protocol.\fix{Luca}{Io lo direi diversamente: ``Dolev-Yao attacker model'' is an abstract model of an attacker that can be used for the analysis of protocol specifications} This, in turn, creates a false sense of security since it requires
 non-justifiable assumptions on the abstraction of the system of which security
-is verified. More specifically, for the formal security verification of a
+is verified.\fixn{Luca}{Frase scorretta grammaticamente ma anche molto confusa.
+Io direi: a proof of security of the protocol specification under the Dolev-Yao attacker model may instill a false sense of security of the whole protocol, whereas we have only verified it for the speicification.
+O una cosa del genere.
+In generale, ci sono troppo salti logici nei claims, dettati dalla voglia di dire troppe cose subito.} More specifically, for the formal security verification of a
 system in which two or more agents are communicating, a formalized scenario
@@
 formalization (e.g.  excluding the server that distributes the public key is
-often done when verifying the security of authentication protocols), (ii) the
+often done when verifying the security of authentication protocols),fixn{Luca}{Nowadays, most tools don't remove these} (ii) the
 number of sessions (even tough some approaches do reason on an infinite number
-of sessions such as \autocite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
-the peers (e.g.  in the ASLan++ language \autocite{Oheimb2010aslan++}), and (iv)
+of sessions such as \autocite{Escobar2007maudenpa}),\fixn{Luca}{Nowadays, most tools work for infinite sessions: Maude-NPA, Tamarin, ProVerif, OFMC.} (iii) honesty/dishonesty of
+the peers (e.g.  in the ASLan++ language \autocite{Oheimb2010aslan++}),\fixn{Luca}{Nowadays, most tools allow for easy switch of the resulting sessions} and (iv)
 the abstraction of the cryptographic primitives (e.g.  ProVerif vs CryptoVerif
-\autocite{Blanchet2017symbolic}).  While many tools and theories improves those 
+\autocite{Blanchet2017symbolic}).\fixn{Luca}{This is the only claim that, more or less, still holds, even though more and more tools do cryptoanalysis.
+The good news is that, I believe, you do not need (i)-(iv) at all, but you can simply say that by the nature of modeling a specification, you are abstracting away certain things (e.g., cryptography, actual message content, PKI, etc.), and the vulnerabilities might be hidden there} While many tools and theories improved those
 issues, there is no agreement on which should be the definitive
-approach (if any). More importantly, some of the choices made by the
+approach (if any).\fixn{Luca}{I don't think that this is important.
+In many fields of research you don't have a single definitive approach and it is good that it is so.
+I would remove this sentence, but only focus on the next point, which is the only important one of this discussion of protocols} More importantly, some of the choices made by the
 modelers/engineers using those formal approaches may completely change the
@@
 the so called perfect cryptography assumption, the security encryption scheme
-is suppose to be perfect, without any exploitable flaw, and so the only way for
+is supposed to be perfect, without any exploitable flaw, and so the only way for
 the attacker to decrypt a message is by using the proper key. That assumption
@@
 mathematical and implementation details of the encryption scheme
-\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
-and assuming that no violation to any security property is done after message
+\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}\fixn{Luca}{I would not quote your own paper for something for which there are hundreds of other equivalent quotes.
+Just say ``under the perfect
+cryptography assumption of the Dolev-Yao attacker model''}
+and assuming that no violation to any security property occurs after message
 I), the freedom of choosing the scope
-determines that the flaws related to the dishonest impersonation of the Server
-may or may not be considered in the verification process.  This choice has
+determines that the flaws related to the dishonest impersonification of the Server
+may or may not be considered in the verification process.  This choice has a
 tremendous impact on the focus and findings of the verification of the security
@@
 this highlights the false sense of security that may derive from a
-non-falsifiable theory of system security
+non-falsifiable theory of system security.\fixn{Luca}{Frase troppo complicata, alla fine di un paragrafo alquanto confuso e ridondante.
+Secondo me tutto questo discorso si pu\`o ridurre a tre frasi, come ho suggerito sopra.'}
 %\footnote{``To the superficial
@@
 
-\paragraph{Structure} We start, in
+\paragraph{Structure} 
+We start, in
 Section~\ref{sec:literature}, by detailing the problem statement, reporting a

### 2020-07-14 reports/paper_0/main.tex
implement Luca's feedback on intro. remove abstract

--- 2020-06-25 reports/paper_0/main.tex
@@
 \begin{document}
-\title{On Etiology of Cybersecurity}
+\title{The Etiology of Cybersecurity}
 
@@
 \begin{abstract}
-The objective of this research is the development of a theory that defines (all
-	and only) the possible insecurity and security configurations of any
-	abstract system. The theory is structured upon other theories that
-	defines how a component of a system can be abstracted into an agent,
-	defining how agents can be formalized (both syntactically and
-	semantically) to describe an abstract system, such as a graph. The core
-	theories are the epistemological definition of knowledge, Beliefs, and
-	Information, the Assertion-Belief-Fact framework, and mereotopology.
-	We implemented a formal theory (of axioms) of a mereotopology, and of
-	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
-	uses the Z3 SMT solver. The results show that a single component (i.e.
-	agent) of an abstract system has a definite number of  different
-	insecurity configurations (e.g. 53 using RCC5 over a topological
-	structure) and only 1 secure (i.e.  expected) configuration. The
-	configurations are reported as models satisfying the abstract system
-	semantics. We use these results to predict all possible security
-	weaknesses of a system, instead of relying on databases of known
-	weaknesses, and to lay the foundations of a scientific theory of
-	cybersecurity.  We show a concrete applications of our theory to the
-	risk assessment of an ad-hoc system.
 \end{abstract}
@@
 \section{Introduction}\label{sec:intro}
-A scientific theory is an explanation of a phenomenon such that the explanation
-follows the scientific method. The scientific method is an \emph{empirical}
-method that aims at mitigating potential fallacies in theories.  K. Popper
-famously argued (e.g. in \autocite{Popper1959logic}) that a scientific theory
-can never been verified but only falsified, that a theory should not be
-conceived by using the principle of induction\footnote{A.  Einstein wrote to K.
-Popper: ``[...] and I think (like you, by the way) that theory cannot be
-fabricated out of the results of observation, but that it can only be
-invented.''\autocite{Popper1959logic}}, and that empirical experiments should
-be considered as the only evidence to support the non-falseness of a scientific
-theory. If we consider a natural (real-world) phenomenon, such as gravity, we
-can think of a scientific theory as an \emph{abstract} mathematical formula (or
-a set of formulas) that axiomatically define a certain (natural) phenomenon in
-a specific algebraic structure. While such a theory can be proved to be correct
+\paragraph{Context and motivation} 
+A \emph{scientific theory} is an explanation of a phenomenon such that the
+explanation follows the scientific method. The \emph{scientific method} is an
+\emph{empirical} method that aims at mitigating potential fallacies in
+theories.  Karl Popper famously argued (e.g. in \autocite{Popper1959logic})
+that a scientific theory can never be verified but only falsified, that a
+theory should not be conceived by using the principle of
+induction\footnote{Albert  Einstein wrote to Popper: ``[...] and I think (like
+you, by the way) that theory cannot be fabricated out of the results of
+observation, but that it can only be invented.''~\autocite{Popper1959logic}},
+and that empirical experiments should be considered as the only evidence to
+support the non-falseness of a scientific theory.
+
+If we consider a natural (real-world) phenomenon, such as gravity, we can think
+of a scientific theory as an \emph{abstract} mathematical formula (or a set of
+formulas) that defines axiomatically a certain (natural) phenomenon in a
+specific algebraic structure. While such a theory can be proved to be correct
 within the mathematical setting it has been conceived in (e.g. with respect to
 other axioms of the algebraic structure where it has been postulated, or with
-respect to other background theories), no proof (for soundness or completeness)
-can be carried out to verify that the theory correctly abstracts  the
-phenomenon it describes. Similarly, the results of the application of a
-scientific theory (i.e. its predictions) can be verified in the abstract
-mathematical structure but how can we be sure that the theory correctly
-characterize the phenomenon it describes? With his famous ``demarcation
-principle'', K. Popper draw a line between scientific and pseudo-scientific
-theories by proposing that \emph{any} scientific theory can be considered
-correct as long as no empirical experiment falsify it. For example, the theory
-of gravity, in its mathematical setting, can predict how object attracts with
-each other, and those predictions can be tested by empirical experiment. If the
-theory correctly abstracts the phenomenon no empirical experiment should
-falsify the predictions of that theory.
+respect to other background theories), no proof can be carried out to verify
+that the theory properly abstracts the phenomenon it describes (i.e. that the
+theory is a sound or complete approximation of the real phenomenon).
+Similarly, the results of the application of a scientific theory (i.e. its
+predictions) can be verified in the abstract mathematical structure but how can
+we be sure that the theory correctly characterizes the phenomenon it describes?
+
+Popper's famous ``demarcation principle'' draws a line between scientific and
+pseudo-scientific theories by proposing that \emph{any} scientific theory can
+be considered correct as long as no empirical experiment falsify it. For
+example, the theory of gravity, in its mathematical setting, can predict how
+objects attract each other, and those predictions can be tested by empirical
+experiment. If the theory correctly abstracts the phenomenon, no empirical
+experiment should falsify the predictions of that theory.
 
@@
 the security of a system, nor a theory that can predict all possible
-insecurity of a system (which, by negation it would become a theory on security). 
-Herley then uses this argument to show that ``claims that any measure
-is necessary for security are empirically unfalsifiable''. Given that, any
-theory which is not falsifiable by an empirical experiment is well
-known\footnote{``A theory which is not refutable by any conceivable event is
-nonscientific. Irrefutability is not a virtue of a theory (as people often
-think) but a vice.'' -- Karl Popper, Conjectures and
-Refutations \autocite{popper1962conjectures}} to be nonscientific (i.e.
+insecurities of a system (which, by negation would become a theory on
+security).  Herley then uses this argument to show that ``claims that any
+measure is necessary for security are empirically unfalsifiable''. An example
+is whether or not password masking makes a system secure. Password masking
+prevents a complete leakage of the password and a system in which password
+masking is implemented is believed to be secure because an intruder won't be
+seeing the password by ``shoulder surging''.  However, this reasoning is
+circular and it is always true that a system in which password masking is
+implemented is a system in which a password is masked. No experiment can ever
+falsify this statement.  Given that any theory that is not falsifiable by an
+empirical experiment is well known\footnote{``A theory which is not refutable
+by any conceivable event is nonscientific. Irrefutability is not a virtue of a
+theory (as people often think) but a vice.'' -- Karl Popper, Conjectures and
+Refutations~\autocite{popper1962conjectures}.} to be nonscientific (i.e.
 unfalsifiability is a fallacy of a theory), Herley concludes that there is no
-scientific theory on cybersecurity; which means that cybersecurity lays in the
-realm of pseudo-sciences \autocite{Herley2016usenixvideo}.  Herley, e.g.
-in \autocite{Herley2017justifying}, discusses the implications of a
-nonscientific approach to cybersecurity, and highlights the tremendous impact on
-all the scientific research and engineering of systems; leading often to
-terrorism and wars, and wasting of resources in useless protections or
-overspending.  While the criticism is investigated
-in \autocite{Herley2016unfalsifiability}, no solution is provided nor envisioned.  On the
-contrary, the goal of this work is to lay the foundations of a
-scientific cybersecurity theory.
-
-We consider the problem raised by Herley not confined to ``computer security''
-but to any abstract system (so that our theory may hold for any sound
-implementation such as networks, mechanical, cyber, or cyber-physical system,
-or even a single computer or a single device such as an hard-drive).  There is
-also an apparent inconsistency in \autocite{Herley2016unfalsifiability} that we
-seek to clarify before following (as we agree) the scientific path draw by
-Herley: cybersecurity is defined as an abstract property in many formal
-approaches to the investigation of the security of systems, and the security of
-the design of a formally verified protocol is indeed falsifiable (against the
-security properties verified).  For example, in the protocol verification
-community, security is often defined as a formalization of the high-level
-properties confidentiality, integrity, and authentication. The problem in such
-approaches is not the definition of what cybersecurity is but the generality of
-the results, since they tackle a specific step\footnote{In fact, many standards
-(e.g. ) used for the engineering of systems starts with the requirements
-definition, and physical and functional architecture design, before the logic
-of protocols is even considered.} of the engineering process (of systems)
-without considering many aspects related to the implementation of a protocol
-that may generate security vulnerabilities later in the development process. An
-infamous example is the KRAK attack on the WPA2 4-way handshake protocol where
-the authors wrote\autocite{krak2020web}: ``our attacks do not violate the security properties proven
-in formal analysis of the 4-way handshake'' but ``the problem is that the
-proofs do not model key installation.''.
-The theories underlying the verification of security protocols are based on assumptions which
-non-evidently apply to a general security theory.  As an example, the so called
-Dolev-Yao attacker model\footnote{For the sake of simplicity, the Dolev-Yao
-attacker can be considered as an abstraction of an active attacker who controls
-the network but cannot break cryptography.} \autocite{Dolev1983security} only
-applies to specific instances (often called scenarios) and abstraction of the
-protocol. This, in turn, creates a false sense of security since requires
-non-justifiable assumptions on the abstraction of the system of which security
-is verified. More specifically, for the formal security verification of a
-system in which two or more agents are communicating, a formalized scenario
-needs to be defined by a modeler who chooses (among others): (i) a scope of the
-formalization (e.g.  excluding the server that distributes the public key is
-often done when verifying the security of authentication protocols), (ii) the
-number of sessions (even tough some approaches do reason on an infinite number
-of sessions such as \autocite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
-the peers (e.g.  in the ASLan++ language \autocite{Oheimb2010aslan++}), and (iv)
-the abstraction of the cryptographic primitives (e.g.  ProVerif vs CryptoVerif
-\autocite{Blanchet2017symbolic}).  While many tools and theories improves those 
-issues, there is no agreement on which should be the definitive
-approach (if any). More importantly, some of the choices made by the
+scientific theory of cybersecurity; which means that cybersecurity lays in the
+realm of pseudo-sciences~\autocite{Herley2016usenixvideo}. Herley, e.g.  in
+\autocite{Herley2017justifying}, discusses the implications of a nonscientific
+approach to cybersecurity, and highlights the tremendous impact on all the
+scientific research and engineering of systems. While the criticism is
+investigated in \autocite{Herley2016unfalsifiability}, no solution is provided
+nor envisioned.
+
+\paragraph{Contributions} 
+The goal of this paper is to lay the foundations of a scientific cybersecurity
+theory.  We consider the problem raised by Herley not confined to ``computer
+security'' but rather we reason on any abstract system (so that our theory may
+hold for any sound implementation such as networks, mechanical, cyber, or
+cyber-physical system, or even a single computer or a single device such as a
+hard-drive).  There is also an apparent inconsistency in
+\autocite{Herley2016unfalsifiability} that we seek to clarify before following
+(as we agree) the scientific path drawn by Herley: cybersecurity is defined as
+a set of abstract properties in many formal approaches to the investigation of
+the security of systems.  Let's, for the sake of simplicity, consider the case
+of a system implementing a security protocol (e.g. an authentication protocol).
+The proof of security obtained from the formal verification of the design of
+the protocol (e.g. with respect to a formalization of the authentication
+property) is indeed falsifiable against the security properties verified.  For
+example, in the protocol verification community, cybersecurity is often defined
+as a formalization of the high-level properties confidentiality, integrity, and
+authentication. The problem in such approaches is not the definition of what
+cybersecurity is but the generality of the results, since they tackle a
+specific step\footnote{In fact, engineering models such as the V-model start
+with the requirements definition, and physical and functional architecture
+design, before the logic of protocols is even considered.} of the engineering
+process (of systems) without considering many aspects related to the
+implementation of a protocol that may generate security vulnerabilities later
+in the development process.  An infamous example is the KRAK attack on the WPA2
+4-way handshake protocol where the author wrote~\autocite{krak2020web}: ``our
+attacks do not violate the security properties proven in formal analysis of the
+4-way handshake'' but ``the problem is that the proofs do not model key
+installation.'' The theories underlying the verification of security protocols
+are based on assumptions which non-evidently apply to a general security
+theory; otherwise the assumptions would clearly state the necessary element
+that needs to be considered (e.g. modeled) in order to obtain a secure system.
+As an example, the so called Dolev-Yao attacker model\footnote{The Dolev-Yao
+attacker model can be considered as abstract model of an attacker that can be
+used for the analysis of protocol specifications.} \autocite{Dolev1983security}
+only applies to specific instances (often called scenarios) and abstraction of
+the protocol.  This, in turn, creates a false sense of security since it
+requires non-justifiable assumptions on the abstraction of the system under
+verification.
+%More specifically, for the formal security verification of a system in which
+%two or more agents are communicating, a formalized scenario needs to be
+%defined by a modeler who chooses (among others): (i) a scope of the
+%formalization (e.g.  excluding the server that distributes the public key is
+%often done when verifying the security of authentication protocols), (ii) the
+%number of sessions (even tough some approaches do reason on an infinite number
+%of sessions such as \autocite{Escobar2007maudenpa}),\fixn{Luca}{Nowadays, most
+%tools work for infinite sessions: Maude-NPA, Tamarin, ProVerif, OFMC.} (iii)
+%honesty/dishonesty of the peers (e.g.  in the ASLan++ language
+%\autocite{Oheimb2010aslan++}),\fixn{Luca}{Nowadays, most tools allow for easy
+%switch of the resulting sessions} and (iv) the abstraction of the
+%cryptographic primitives (e.g.  ProVerif vs CryptoVerif
+%\autocite{Blanchet2017symbolic}).\fixn{Luca}{This is the only claim that, more
+%or less, still holds, even though more and more tools do cryptoanalysis. The
+%good news is that, I believe, you do not need (i)-(iv) at all, but you can
+%simply say that by the nature of modeling a specification, you are abstracting
+%away certain things (e.g., cryptography, actual message content, PKI, etc.),
+%and the vulnerabilities might be hidden there} While many tools and theories
+%improved those issues, there is no agreement on which should be the definitive
+%approach (if any).\fixn{Luca}{I don't think that this is important. In many
+%fields of research you don't have a single definitive approach and it is good
+%that it is so. I would remove this sentence, but only focus on the next point,
+%which is the only important one of this discussion of protocols} 
+More importantly, some of the choices made by the
 modelers/engineers using those formal approaches may completely change the
@@
 \end{figure}
-For example, in Figure~\ref{fig:protocol-example}, under the perfect
-cryptography assumption\footnote{As defined in \autocite{Rocchetto2016cpdy}: ``In
-the so called perfect cryptography assumption, the security encryption scheme
-is suppose to be perfect, without any exploitable flaw, and so the only way for
-the attacker to decrypt a message is by using the proper key. That assumption
-is widely accepted in the security protocol community, and most of the formal
-reasoning tools for the analysis of security protocols abstract away the
-mathematical and implementation details of the encryption scheme
-\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
-and assuming that no violation to any security property is done after message
-I), the freedom of choosing the scope
-determines that the flaws related to the dishonest impersonation of the Server
-may or may not be considered in the verification process.  This choice has
-tremendous impact on the focus and findings of the verification of the security
-of the protocol.  While this may seem to turn upon minutiae and foreseeable,
-this highlights the false sense of security that may derive from a
-non-falsifiable theory of system security
-%\footnote{``To the superficial
-%observer, the analysis of these forms seems to turn upon minutiae. It does in
-%fact deal with minutiae, but they are of the same order as those dealt with in
-%microscopic anatomy.'' -- Karl Marx, Capital Volume 1, 1867}.
-
-\paragraph{Structure} We start, in
-Section~\ref{sec:literature}, by detailing the problem statement, reporting a
-literature review on the main concepts and definitions related to security.  We
-formulate a security hypothesis in Section~\ref{sec:hypothesis}; which we use
-to propose a theory on system security in Section~\ref{sec:theory}. In
-Section~\ref{sec:secra}, we apply our theory for the tool-assisted security
-risk assessment of a CPS (with an ad-hoc example, based on the SWaT testbed
-\autocite{Mathur2016swat}).  This application shows how our theory can be used to
-predict all of the possible security weaknesses of a system, allowing the
-falsification of our theory.  In fact, if any security weaknesses were to be
-found in a system and not predicted by our theory, the theory could be declared
-incomplete.  Similarly, if a security weakness would be predicted by our theory
-but found to be impossible to realize, our theory could be declared as wrong.
+As an illustrative example, in Figure~\ref{fig:protocol-example}, under the perfect
+cryptography assumption (i.e. assuming that the cryptographic primitives have
+no security flaw), and assuming that no violation to any security property
+occurs after message I), the freedom of choosing the scope determines that the
+flaws related to the dishonest impersonification of the Server may or may not
+be considered in the verification process.  This choice has a tremendous impact
+on the focus and findings of the verification of the security of the protocol.
+While this may seem to turn upon minutiae and foreseeable, this highlights the
+false sense of security that may derive from a non-falsifiable theory of system
+security.
+
+\paragraph{Structure} 
+We start, in Section~\ref{sec:literature}, by detailing the problem statement,
+reporting a literature review on the main concepts and definitions related to
+security.  We formulate a security hypothesis in Section~\ref{sec:hypothesis};
+which we use to propose a theory on system security in
+Section~\ref{sec:theory}. In Section~\ref{sec:secra}, we apply our theory for
+the tool-assisted security risk assessment of a CPS (with an ad-hoc example,
+based on the SWaT testbed \autocite{Mathur2016swat}).  This application shows
+how our theory can be used to predict all of the possible security weaknesses
+of a system, allowing the falsification of our theory.  In fact, if any
+security weaknesses were to be found in a system and not predicted by our
+theory, the theory could be declared incomplete.  Similarly, if a security
+weakness would be predicted by our theory but found to be impossible to
+realize, our theory could be declared as wrong.
 

### 2020-07-15 reports/paper_0/main.tex
add abstract on paper_0

@@
 \begin{abstract}
+	The objective of this research is the development of a theory that
+	determines (all and only) the possible insecurity and security
+	configurations of any abstract system.  Our hypothesis is that errors
+	are at the bottom of the causality chain that leads to cybersecurity
+	attacks. Based on this hypothesis, we formulate a scientific
+	(falsifiable by experiments) theory that predicts all the errors that lead to
+	the presence of cybersecurity weaknesses in the architectural design of
+	an abstract system.  Our theory is based on a correlation between the
+	epistemological concepts of knowledge, beliefs, and information with
+	the engineering concepts of requirements, functional architectures, and
+	channels.  Our theory allows the definition of the security of a system
+	as a mathematical formula correlating the security and insecurity
+	configurations of a system.  We implemented a prototype cybersecurity
+	risk assessment tool that, based on our theory, predicts weaknesses in
+	a UML model of a (cyber-physical) system.
 \end{abstract}

### 2020-07-15 reports/paper_0/main.tex
implement Volpe's review (part 1)

@@
 
+\fix{mr}{add description of our approach and theory}
+
 \paragraph{Structure} 
@@
 in turn, between safety and cybersecurity) is believed to be on the different
-intents\footnote{``The belief–desire–intention software model (BDI) is a
-software model developed for programming intelligent
-agents.''\autocite{wiki-bdi}. In the BDI model, the intents represents the
-deliberative state of an agent which determines the choice of that agent on
-what to do.} (incidents are intentional while accidents are not) of the causes
+intents (incidents are intentional while accidents are not) of the causes
 that generates a threat; namely, nature is believed not to have malicious
@@
 (therefore should be considered hypothesis and not definitions). While
-reasoning by induction\footnote{``So, whenever they argue ``Every man is an
-animal and Socrates is a man; therefore Socrates is an animal,'' proposing to
-deduce from the universal proposition ``every man is an animal'' the particular
-proposition ``Socrates therefore is an animal,'' which in fact goes (as we have
-mentioned) to establish by way of induction the universal proposition, the
-fall into the error of circular reasoning, since they are establishing the
-universal proposition inductively by means of each of the particulars and
-deducing the particular proposition from the universal syllogistically.''
-Sextus Empiricus, Outlines of Pyrrhonism II-195 \autocite{Empiricus1990Pyrrhonism}}
+reasoning by induction
+%\footnote{``So, whenever they argue ``Every man is an
+%animal and Socrates is a man; therefore Socrates is an animal,'' proposing to
+%deduce from the universal proposition ``every man is an animal'' the particular
+%proposition ``Socrates therefore is an animal,'' which in fact goes (as we have
+%mentioned) to establish by way of induction the universal proposition, the
+%fall into the error of circular reasoning, since they are establishing the
+%universal proposition inductively by means of each of the particulars and
+%deducing the particular proposition from the universal syllogistically.''
+%Sextus Empiricus, Outlines of Pyrrhonism II-195 \autocite{Empiricus1990Pyrrhonism}}
 based on the empirical observation should be avoided, since it may easily lead
@@
 		example, the simple temporal sequence of phenomena, of the
-		regularities of accompanying processes''. One can consider the
-		causality principle to be formalized by the K Modal Logic.
+		regularities of accompanying processes''. 
+		%One can consider the
+		%causality principle to be formalized by the K Modal Logic.
 	\item An \emph{Exploit}\footnote{We note that the term exploit is only
@@
 
-\subsection{Systems as Multi-Agent Systems}\label{sec:system}
+\subsection{Multi-Agent Systems}\label{sec:system}
 The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
@@
 or more stated purposes.''\autocite{ISO201515288}.  Therefore, a system can be
-considered as a single agent where its interacting elements are the
-constituents of the agent itself. For the sake of simplicity we first define a
+considered as a network of interacting elements.
+\fixnote{mv}{il ragionamento dell'intro, in particolare il "Therefore", non mi convince del tutto; dalla definizione mi verrebbe da pensare un sistema come un insieme di agenti che interagiscono piuttosto che come un agente singolo}
+%single agent where its interacting elements are the
+%constituents of the agent itself. 
+For the sake of simplicity we first define a
 system as an agent and then extend the definition to a ``combination of
@@
 \end{enumerate}
-We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}
+We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}.
+The notation will be introduced in the next section but, intuitively,
+Figure~\ref{fig:system-agent} shows a system composed by two agents. The system
+has to ports denoted by ``I'' (with incoming assertions) and ``O'' (with outgoing assertions).
+The bottom part of Figure~\ref{fig:system-agent} details the structure of a single
+agent which is not decomposed into other (sub-)agents. In this case, ports translates
+incoming assertions into beliefs that feeds the behavior block, and outgoing beliefs
+from the behavior block are then translated back into (outgoing) assertions.
 
@@
 an ordered mathematical structure where the basic relation between Regions is
-the reflexive and symmetric
-\emph{Parthood} relation $\subseteq$.  
-\begin{definition}{\bf Parthood --}\label{def:parthood}
-Given any pair of mereotopological Regions $X$ and $Y$,
-	\begin{enumerate}[noitemsep]
-		\item Reflexivity: $\forall X.~ (X\subseteq X)$
-		\item Symmetry: $\forall X, Y.~ (X\subseteq Y \Rightarrow Y\subseteq X)$
-	\end{enumerate}
-\end{definition}
-
+the reflexive, antisymmetric, and transitive \emph{Parthood} relation $\subseteq$.  
 The Parthood relation orders a universe of agents $\agentuniverse$ by defining the so called
@@
 by eight axioms, and is known as RCC8. In the text, for brevity, 
-we will often focus only on RCC5 (without
-loss of generality) by not considering tangential connections between spatial
-Regions. In Table~\ref{tab:rcc}, we
+we will often focus only on RCC5 without
+loss of generality. Using RCC5 instead of RCC8 prevents us from 
+considering tangential connections between spatial
+Regions. However, tangential connections in RCC8 can be considered as 
+special cases of the more general spatial relations considered in RCC5.
+
+In Table~\ref{tab:rcc}, we
 summarize the axioms of the Region Connection Calculus (see, e.g., \autocite{Grutter2008rcc}).
-We can now define a system over the mereotopology using the RCC calculus, as follows.
+We can now define a system over the mereotopology using the RCC calculus, as follows, 
+where $\rcc(X,Y)$ on two generic regions $X,Y$ represents one of the possible RCC relations
+between $X$ and $Y$.
 
@@
 $r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
-between $n$ different regions, where $r^e$ is the number of permutation of $r$
-relations over $e$ elements with repetitions, with $e$ being the number of
-$k$-ary combinations of $n$ regions, $\binom{n}{k}$.
+between $n$ different regions. Therefore, $r^{\binom{n}{k}}$
+expresses the number of permutation of $r$
+relations over ${\binom{n}{k}}$ elements with repetitions, 
+with ${\binom{n}{k}}$ being the number of
+$k$-ary combinations of $n$ regions.
 In our case, $\binom{n}{k}=3$ since we consider $3$ regions 

### 2020-07-15 reports/paper_0/main.tex
implement Volpe's feedback (part 2)

@@
 \newcommand{\fixnote}[2]{\textbf{\color{red}{FIX}}\footnote{{\bf #1:} #2}}
-\newcommand{\fix}[2]{{\color{red} {\bf tofix:} #2}}
+\newcommand{\fix}[2]{{\color{red}{\bf tofix:} #2}}
 % \renewcommand{\fixnote}[2]{}
@@
 prevents a complete leakage of the password and a system in which password
-masking is implemented is believed to be secure because an intruder won't be
+masking is implemented is believed to be secure because an attacker won't be
 seeing the password by ``shoulder surging''.  However, this reasoning is
@@
 
-\fix{mr}{add description of our approach and theory}
+Our hypothesis is that cybersecurity attacks are caused by 
+the presence of errors in the design or implementation of a system.
+For example, a design or implementation error in a sanitization function
+may allow an attacker to perform SQL-injections. If errors are necessary
+to enable cybersecurity attacks, a theory that predicts all potential 
+errors (or source of errors) can be effectively used to predict the
+security of a system. With our approach, a list of weaknesses emerges
+from the mathematical formulation of the system and predicts 4 main classes
+of weaknesses. Those classes of weaknesses are used to calculate
+all the insecurity configurations of all the components of a system,
+obtaining a precise estimation of all potential cybersecurity-related risks
+in any given system.
 
@@
 
-\subsection{Multi-Agent Systems}\label{sec:system}
+\subsection{Epistemology and Multi-Agent Systems}\label{sec:system}
 The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
 of system as ``A combination of interacting elements organized to achieve one
-or more stated purposes.''\autocite{ISO201515288}.  Therefore, a system can be
-considered as a network of interacting elements.
-\fixnote{mv}{il ragionamento dell'intro, in particolare il "Therefore", non mi convince del tutto; dalla definizione mi verrebbe da pensare un sistema come un insieme di agenti che interagiscono piuttosto che come un agente singolo}
-%single agent where its interacting elements are the
-%constituents of the agent itself. 
+or more stated purposes.''\autocite{ISO201515288}. If a system is structured
+into multiple sub-systems, a sub-system can be defines in the same way as a system (i.e. as
+a composition of interacting elements). A hierarchy of sub-systems 
+can be limited by a ``bottom-subsystem'' often called device (or element in the
+aforementioned definition) which is not itself composed by 
+other elements but is considered atomic.         
 For the sake of simplicity we first define a
-system as an agent and then extend the definition to a ``combination of
-interacting'' agents.
+system as composed by a single element and then extend the definition to a ``combination of
+interacting'' elements. Given that we use concepts from the multi-agent system
+research field, instead of using the word element we use agent.
 
@@
 readability, we use the term \emph{Region} both to refer to a mereological Part
-and to a topological Region.  Our aim is to create a meronomy
-instead of the taxonomies such as the one provided
-in\autocite{NIST2020NVD,MITRE2020CVE} with the
-so that we don't need to rely on a scoring system such as the 
-CVSS\autocite{Mell2007CVSS}. Instead, we want to precisely calculate
-the number of insecurity configurations as a scoring system.
+and to a topological Region.  Our aim is to create a meronomy (hierarchy of part-whole
+relations)
+instead of the taxonomies (categorization based on discrete sets) 
+such as the one provided
+in\autocite{NIST2020NVD,MITRE2020CVE} 
+so that we don't need to rely on a scoring system (such as the 
+CVSS\autocite{Mell2007CVSS}) to assign a quantitative evaluation 
+of the security of each entry. Instead, we want a precise calculation
+of the number of insecurity configurations to emerge from the mathematical
+formulation of our cybersecurity theory.
 

### 2020-07-16 reports/paper_0/main.tex
review Volpe's feedbacks on 3.C (part 1)

@@
 other elements but is considered atomic.         
-For the sake of simplicity we first define a
+For the sake of simplicity, we consider a device as a system with only one element.
+We then first define a
 system as composed by a single element and then extend the definition to a ``combination of
 interacting'' elements. Given that we use concepts from the multi-agent system
-research field, instead of using the word element we use agent.
+research field, instead of using the word element we will use the word agent.
 
@@
 For our argument, as in \autocite{Santaca2016abf}, an agent is composed by its
-knowledge, beliefs, and the information or assertions it provides; where
+knowledge, beliefs, and the information or assertions\fixnote{mv}{
+In generale, farei attenzione a uniformare upper/lower case per assertion, behavior, beliefs, ecc.
+} it provides; where
 knowledge is defined, as in \autocite{Steup2020epistemology}, as a set of
@@
 	\includegraphics[width=.9\columnwidth]{system-agent.pdf}
-	\caption{Example of system and agent structure}
+	\caption{xample of system and agent structure}
 	\label{fig:system-agent}
@@
 representation of the system.
+\fixnote{mv}{In Fig. 5 e 6, i simboli A, B e F hanno un font diverso dal solito}
 \begin{figure}[t]
@@
 \subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}
-\begin{table*}[t]
+\begin{table}[t]
 \centering
-\begin{adjustbox}{width=.6\textwidth}
+\begin{adjustbox}{width=\columnwidth}
 \begin{tabular}{r||c|c|c|c|c} 
@@
 	\cellcolor{abf-rg-4}\dr{$\assertionRegion$}{$\factRegion$}\\ % e q
- & \cellcolor{abfred}& %dr
-	\cellcolor{abf-rg-1}\pp{$\assertionRegion$}{$\factRegion$} & %po
-	\cellcolor{abf-rg-2}  & %pp
-	\cellcolor{abf-rg-3}\pp{$\assertionRegion$}{$\factRegion$} & %ppi
-	\cellcolor{abf-rg-4} \\ %eq
+% & \cellcolor{abfred}& %dr
+%	\cellcolor{abf-rg-1}& %\pp{$\assertionRegion$}{$\factRegion$} & %po
+%	\cellcolor{abf-rg-2}  & %pp
+%	\cellcolor{abf-rg-3} & %\pp{$\assertionRegion$}{$\factRegion$} & %ppi
+%	\cellcolor{abf-rg-4} \\ %eq
 \hline %po
@@
 \label{tab:5com}}
-\end{table*}
+\end{table}
 While a quantitative analysis reveals how many possible configurations of an
 agent (i.e. a system) exist w.r.t. the $\abf$ theory (i.e. $54/125$ in RCC5), a
-qualitative analysis of the different configurations describes the
+qualitative analysis of the different configurations describe the
 configurations allowed by the $\abf$ theory, and how those configurations can
@@
 state space for a single agent. The color coding of the table represents the
-security risk related to a generic agent.
+security risk related to a generic agent, the risk is highest on the top left
+corner of the matrix, lowest on the bottom right corner.
 
 As depicted in Figure~\ref{fig:soundness}, the relation between Facts, and
-Assertions and Behavior defines the soundness of the design (admissible
+Assertions and Behavior\fixnote{mv}{io ho inteso A B e F come Assertions, Beliefs e Facts, rispettivamente; di tanto in tanto tornate a usare il termine Behavior e questo secondo me confonde} 
+defines the soundness of the design (admissible
 configurations of Assertions, Behaviors and, in turn, Beliefs) w.r.t. the
@@
 the following, however, with the RCC one should consider different ``degrees''
-of non-soundness. For example, in RCC5, if we consider EQ between two Regions
+of non-soundness\fixnote{mv}{In particolare, spiegherei meglio Fig. 6 e il discorso della soundness}. 
+For example, in RCC5, if we consider EQ between two Regions
 as representing soundness, DR over the same Regions represents non-soundness;
@@
 \Rcc{$\beliefRegion$}{$\factRegion$} -- Honesty and Competence} 
-%While
-%Assertions and Beliefs generated by the Behavior determines how a system should
-%work, and the relation between the two defines the quality (i.e. the
-%completeness w.r.t the specification) of the agent, 
 The relation of Assertions
@@
 \begin{enumerate}
-	\item[SE-$1$] Proper interaction between correctly-behaving agents (in
-		contrast with the ``Improper Interaction Between Multiple
-		Correctly-Behaving Entities'' defined by the CWE--$435$ as one
-		of the top ``view'' of the ``research concepts'' in
-		\autocite{MITRE2020CWEresearch}) is defined as
-		$\eq{\assertionRegion_a}{\beliefRegion_a}$ for an agent $a$
-		while can be detailed as follows when multiple agents are
-		considered.
+	\item[SE-$1$] Proper interaction between correctly-behaving agents is
+		defined as $\eq{\assertionRegion_a}{\beliefRegion_a}$ for an
+		agent $a$ while can be detailed as follows when multiple agents
+		are considered.
 	\begin{enumerate}
@@
 	\end{enumerate}
-	\item[SE-$2$]Sufficient Control Flow Management (in contrast with
-		the ``Insufficient Control Flow Management'' defined by MITRE
-		in the CWE--$691$ as one of the top ``view'' of the ``research
-		concepts'' in\autocite{MITRE2020CWEresearch}) is defined as
+\item[SE-$2$]\fixnote{mv}{Are agents missing in the notation A,B,F for SE-2 and SE-3?}
+	Sufficient Control Flow Management  is defined as
 		$\eq{\assertionRegion}{\factRegion}$.
-	\item[SE-$3$] Correct Calculation (in contrast with the ``Improper
-		Calculation'' defined by MITRE in the CWE--$682$ as one of the
-		top ``view'' of the ``research concepts''
-		in\autocite{MITRE2020CWEresearch}) is defined as
+	\item[SE-$3$] Correct Calculation  is defined as
 		$\eq{\beliefRegion}{\factRegion}$.
 \end{enumerate}
+
+We note that our SE define the properties of a secure system, and correlated
+weaknesses can be found in the CWE dataset.  For example, SE-$1$ can be sees as
+correlated to the weakness class of ``Improper Interaction Between Multiple
+Correctly-Behaving Entities'' defined by the CWE--$435$, SE-$2$ with the
+``Insufficient Control Flow Management'' defined by MITRE in the CWE--$691$,
+and SE-$3$ with the ``Improper Calculation'' defined by MITRE in the
+CWE--$682$. All those CWE are in the top ``view'' of the ``research concepts''
+in\autocite{MITRE2020CWEresearch}, while the other classes of weaknesses do not
+have a direct counterpart in our theory; we believe they can be sees as sub-classes,
+but a full comparison with the CWE is out of the scope of this article.
 
@@
 how to quantify it in a risk matrix (where a risk matrix is usually defined as
-a function between the likelihood and the impact of all potential attacks).
+a function of the likelihood and the impact of all potential attacks).
 \begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
@@
 \begin{definition}{\bf Risk --}
-The whole space of potential designs of a specification with respect to the
+A risk is the whole space of potential designs of a specification with respect to the
 	$\abf$-theory.
@@
 \begin{definition}{\bf Risk Matrix --}
-	The risk matrix is a function between the three relations
+	The risk matrix, as summarized in Table~\ref{tab:5com}, is a function of the three relations
 	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
@@
 \end{definition}
-While a risk matrix is often defined as a function between 
+While a risk matrix is often defined as a function of 
 likelihood and impact of attacks, we suggest
@@
 approach, (iii) calculate the risk as a function
-between impact and likelihood of threat scenarios, and (iv) finally provides
+of impact and likelihood of threat scenarios, and (iv) finally provides
 treatments (mandating a partial re-design) and residual risks.
@@
 	\centering
-	\includegraphics[width=0.4\columnwidth]{sum.pdf}
+	\includegraphics[width=0.3\columnwidth]{sum.pdf}
 	\caption{Example of a Functional Block}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{v-secra.pdf}
+	\includegraphics[width=.9\columnwidth]{v-secra.pdf}
 	\caption{Security Risk Assessment Tool}
@@
 
-The risk matrix is often defined as a function between 
+The risk matrix is often defined as a function of 
 likelihood and impact of attacks. We suggest
@@
 The two components are connected by a wire.
-\begin{figure*}
+\begin{figure}
 	\centering
-	\includegraphics[width=.6\textwidth]{internal_cs1.png}
+	\includegraphics[width=.9\columnwidth]{internal_cs1.png}
 	\caption{Sensor Board Object Diagram}
 	\label{fig:int_cs1}
-\end{figure*}
+\end{figure}
 In Figure~\ref{fig:int_cs1} we report the functional architecture that
@@
 
-\section{Discussions}\label{sec:discussion}
-\paragraph{Abstraction Level}
-The level of abstraction is always a concern when it comes to the definition of
-a system model.  In this paper, we justified our assumptions on the atomicity
-of ports, channels, and functional blocks but nothing prevents a modeler to
-consider other atomic element of the system specification and design.  Those
-additional elements may detail the system and may change the syntactic sugar we
-defined; however, it won't change the underlying (semantic) reasoning unless
-another calculus is used (instead of RCC).  Furthermore, other Infosec
-properties may be considered, such as authenticity or non-repudiation. We,
-however, leave them as future development.
-
-\paragraph{$\abftheory$ Falsification} The theory we developed can be falsified
-by providing an experiment that shows a weaknesses not predicted by the
-$\abftheory$-theory, or by showing that the weaknesses predicted for a certain
-experiment are impossible to realize.  This theory, however, needs to be
-detailed to be applicable to a full engineering process (we only cover the
-first steps).  We propose the following complete process as a draft for
-potential next steps in this direction.
-\begin{enumerate}
-\item System Specification
-\item Architecture Design
-\item Security Risk Assessment and Security Requirement Extraction
-\item Design Refinement, where the semantics of the functional behavior and
-	communication protocols are defined 
-\item Security Requirement Extension, based on the design refinement
-\item Verification of Security Requirements (data flow, behaviors, \& protocol
-	logic)
-\item Software Prototype Implementation/Synthesis
-\item Automated Security Testing (abstract test-suite generation \&
-	concretization)
-\item Hardware Prototype Implementation
-\item Automated Security Testing
-\item Penetration Testing
-  \end{enumerate}
-
-\paragraph{Relation with DY}
-The DY attacker model has been extensively
-used for the identification of logical weaknesses in security protocols.
-Our theory should predict all possible weaknesses which, in turn,
-should enable all the attacks identified by the DY model. 
-A proof that the two theories/models correctly predicts
-(or lacks in predicting) weaknesses and related protocol flaws would
-connect risk assessment and protocol verification, the first
-two steps of the secure engineering of systems.
+%\section{Discussions}\label{sec:discussion}
+%\paragraph{Abstraction Level}
+%The level of abstraction is always a concern when it comes to the definition of
+%a system model.  In this paper, we justified our assumptions on the atomicity
+%of ports, channels, and functional blocks but nothing prevents a modeler to
+%consider other atomic element of the system specification and design.  Those
+%additional elements may detail the system and may change the syntactic sugar we
+%defined; however, it won't change the underlying (semantic) reasoning unless
+%another calculus is used (instead of RCC).  Furthermore, other Infosec
+%properties may be considered, such as authenticity or non-repudiation. We,
+%however, leave them as future development.
+%
+%\paragraph{$\abftheory$ Falsification} The theory we developed can be falsified
+%by providing an experiment that shows a weaknesses not predicted by the
+%$\abftheory$-theory, or by showing that the weaknesses predicted for a certain
+%experiment are impossible to realize.  This theory, however, needs to be
+%detailed to be applicable to a full engineering process (we only cover the
+%first steps).  We propose the following complete process as a draft for
+%potential next steps in this direction.
+%\begin{enumerate}
+%\item System Specification
+%\item Architecture Design
+%\item Security Risk Assessment and Security Requirement Extraction
+%\item Design Refinement, where the semantics of the functional behavior and
+%	communication protocols are defined 
+%\item Security Requirement Extension, based on the design refinement
+%\item Verification of Security Requirements (data flow, behaviors, \& protocol
+%	logic)
+%\item Software Prototype Implementation/Synthesis
+%\item Automated Security Testing (abstract test-suite generation \&
+%	concretization)
+%\item Hardware Prototype Implementation
+%\item Automated Security Testing
+%\item Penetration Testing
+%  \end{enumerate}
+%
+%\paragraph{Relation with DY}
+%The DY attacker model has been extensively
+%used for the identification of logical weaknesses in security protocols.
+%Our theory should predict all possible weaknesses which, in turn,
+%should enable all the attacks identified by the DY model. 
+%A proof that the two theories/models correctly predicts
+%(or lacks in predicting) weaknesses and related protocol flaws would
+%connect risk assessment and protocol verification, the first
+%two steps of the secure engineering of systems.
 

### 2020-07-16 reports/paper_0/main.tex
review fonts in images and add nonce example in paper_0

@@
 We now define them abstractly, inspired by the
-ABF-theory  defined in \autocite{Santaca2016abf} and, in
+$\abf$-framework  defined in \autocite{Santaca2016abf} and, in
 Section~\ref{sec:theory} we detail them as engineering concepts.
@@
 representation of the system.
-\fixnote{mv}{In Fig. 5 e 6, i simboli A, B e F hanno un font diverso dal solito}
 \begin{figure}[t]
@@
 	\end{enumerate}
-\item[SE-$2$]\fixnote{mv}{Are agents missing in the notation A,B,F for SE-2 and SE-3?}
-	Sufficient Control Flow Management  is defined as
-		$\eq{\assertionRegion}{\factRegion}$.
-	\item[SE-$3$] Correct Calculation  is defined as
-		$\eq{\beliefRegion}{\factRegion}$.
+\item[SE-$2$] The proper adherence of the data transmitted between agents with respect to
+	requirements, is defined as $\eq{\assertionRegion}{\factRegion}$. 
+	\item[SE-$3$] The proper adherence of the behavior with respect to
+		requirements is defined as $\eq{\beliefRegion}{\factRegion}$.
 \end{enumerate}
@@
 We are now in the position to define what a secure system is (with respect to
-the $\abf$-theory) and, based on that definition, what the security risk is and
-how to quantify it in a risk matrix (where a risk matrix is usually defined as
-a function of the likelihood and the impact of all potential attacks).
+the $\abf$-framework) and, based on that definition, what the security risk is and
+how to quantify it in a risk matrix.
 \begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
@@
 A risk is the whole space of potential designs of a specification with respect to the
-	$\abf$-theory.
+	$\abf$-framework.
 \end{definition}
@@
 \end{definition}
-While a risk matrix is often defined as a function of 
-likelihood and impact of attacks, we suggest
+While a risk matrix is often defined as a function of the 
+likelihood and impact of attacks (based on quantitative ad-hoc estimation
+of how likely it is that an attacker will exploit one or more vulnerabilities
+and what is the magnitude of the incidents produced by this exploitation), we suggest
 that security weaknesses are all equally likely to be exploited
@@
 However, standards do not describe in detail how to perform a security risk 
-assessment and only vaguely define the overall objective. The overall
-objective of this first step of the cybersecurity engineering process
-is to provide an understanding of the potential security risks.
+assessment and only vaguely define the overall objective, which
+can be summarized as to provide an understanding of the potential security risks.
 Roughly speaking, a security risk assessment (e.g. CORAS\autocite{Lund2010model}) 
@@
 \end{figure}
-Since the $\abf$-theory is a theory of agents, we could consider ports as
+Since the $\abf$-framework is a theory of agents, we could consider ports as
 agents that allows the exchange of information between a channel and another
@@
 \begin{definition}{\bf Secure Input Port --}\label{def:secport}
-	A secure input port ($s\secureport a$) always allow information as
+	A secure input port always allow information as
 	incoming assertions to flow from a sender $s$ to the behavior of the
@@
 \begin{proof}
-An input port is, in the $\abf$-theory, defined secure as long as the relation
+An input port is, in the $\abf$-framework, defined secure as long as the relation
 	between the two regions of input assertions $\assertionRegion$ and
@@
 mono-directional channel (channel from now on) and an agent, as we did for the
-ports, since the $\abf$-theory is a theory of agents.  In fact, if a channel
+ports, since the $\abf$-framework is a theory of agents.  In fact, if a channel
 were considered an agent (channel-agent) then the question would be how an
@@
 
-\paragraph{Security Weaknesses} 
+\paragraph{Security Weaknesses -- The RIDI-Hypothesis} 
 It is important to note that all the results of the application of the
-$\abf$-theory to channels (the analysis of the relation
+$\abf$-framework to channels (the analysis of the relation
 $\Rcc{\assertionRegion_o}{\assertionRegion_i}$) lead to the same results of the
@@
 the functional architecture as described afterwards.  As depicted in
-Figure~\ref{fig:atom}, so far, we have considered Information generated by a
+Figure~\ref{fig:atom}, so far we have considered Information generated by a
 Port $P_I$ and then sent through a channel $C$ to another (recipient) port
@@
 In fact, the Weaknesses of a channel are defined in terms of the Weaknesses of
-Ports. We now introduce the concept of functional architecture and its
-decomposition into functional blocks.  Io order to define a functional block
+Ports.  Io order to define a functional block
 without encountering a infinitely recursive definition, we must reach the same
@@
 Block can \emph{only} be affected by the following weaknesses: replace, drop,
-injection, insertion, selective drop, and selective drop + insertion.  We can,
+injection, insertion, selective drop, and selective drop + insertion. 
+We call this the RIDI-Hypothesis, being the
+four main categories of weaknesse: Replace, Insertion, Drop, Injection. We can,
 then, deduce the following security properties to mitigate security weaknesses
@@
 security protocols, with the relation between Assertions and Facts we would
-compare protocol logics with the requirements. We consider the verification of
-the functional architecture and protocol logic in subsequent steps of the
-engineering process (as we describe in Section~\ref{sec:discussion}).
-Therefore, we can summarize our results by categorizing the weaknesses
+compare protocol logics with the requirements. We won't consider the verification of
+the functional architecture and protocol logic in this paper since we
+focus on the architecture specification step of the engineering process
+without going into the design of the behavior of agents. We can give an example
+but research is needed to apply our theory
+to the behavior of agents.
+
+If we consider a functional block that generates nonces (i.e. number used
+only once), the semantics of this block must define a procedure such that 
+a freshness requirement holds. By the RIDI-Hypothesis we have that
+the following weaknesses may occur:
+\begin{itemize}
+	\item Replace: the output-nonces of the block are replaced with
+		non-fresh numbers
+	\item Insert: the output-nonces are concatenated with other numbers,
+		resulting in non-fresh nonces. For example, the nonces 110 and
+		111 are modified by inserting a 1 and a 0 at the beginning and
+		end respectively, resulting in the two identical (and the
+		non-fresh) nonces 1110 and 1110
+	\item Delete: part of output-nonce is dropped. For example, the last bit of 1101 and 1100 is dropped, resulting in the two equal and non-fresh nonces: 110 and 110.
+	\item Inject: substitution at bit level results in non-fresh nonces. For example, 110 becomes 111 becomes equal by flipping either the last 0 of the former nonce or the last 1 of the latter.
+\end{itemize}
+
+We summarize our results by categorizing the weaknesses
 predicted by our theory into: data-flow-related and functionality-related
@@
 	uniquely, i.e.  tested against the range of design possible in the
-	$\abftheory$-theory, defines the design built on top of those
+	$\abftheory$-framework, defines the design built on top of those
 	requirements.
@@
 The graph is translated into 
-a logical formula that represents the specification in the $\abftheory$-theory and,
+a logical formula that represents the specification in the $\abftheory$-framework and,
 along with the axiomatization of the RCC5 calculus, is given as input to the Z3 SMT solver.
@@
 %by providing an experiment that shows a weaknesses not predicted by the
-%$\abftheory$-theory, or by showing that the weaknesses predicted for a certain
+%$\abftheory$-framework, or by showing that the weaknesses predicted for a certain
 %experiment are impossible to realize.  This theory, however, needs to be
@@
 \appendix
-%\section{Engineering $\abftheory$-theory}
-The Class Diagram for the Engineering of the $\abftheory$-theory is reported in
+%\section{Engineering $\abftheory$-framework}
+The Class Diagram for the Engineering of the $\abftheory$-framework is reported in
 Figure~\ref{fig:secraclassdiagram}. A \emph{specification} of a CPS is viewed

### 2020-07-16 reports/paper_0/main.tex
implement Volpe's feedback on 3.C (part 2)

@@
 or more stated purposes.''\autocite{ISO201515288}. If a system is structured
-into multiple sub-systems, a sub-system can be defines in the same way as a system (i.e. as
-a composition of interacting elements). A hierarchy of sub-systems 
-can be limited by a ``bottom-subsystem'' often called device (or element in the
-aforementioned definition) which is not itself composed by 
-other elements but is considered atomic.         
-For the sake of simplicity, we consider a device as a system with only one element.
-We then first define a
-system as composed by a single element and then extend the definition to a ``combination of
-interacting'' elements. Given that we use concepts from the multi-agent system
-research field, instead of using the word element we will use the word agent.
+into multiple sub-systems, a sub-system can be defined in the same way as a
+system is defined (i.e. as a composition of interacting elements). A hierarchy
+of sub-systems can be limited by a ``bottom-subsystem'' often called device (or
+element in the aforementioned definition) which is not itself composed by other
+elements but is considered atomic.         For the sake of simplicity, we
+consider a device as a system with only one element: itself.  We then first define a
+system as composed by a single element and then extend the definition to a
+``combination of interacting'' elements. Given that we use concepts from the
+multi-agent system research field, instead of using the word element we will
+use the word agent.
 
 There is no agreement between the research communities (e.g.
-Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent
-as a system. However, the same ideas revolved around for thousands of years.
+Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent.
+However, the same ideas revolved around for thousands of years.
 Some relevant examples for our objective are the following\footnote{
@@
 	\item In \autocite{Hintikka1962knowledge}, Hintikka describes the
-		difference between Knowledge and Belief (as epistemological
+		difference between knowledge and belief (as epistemological
 		concepts), and the whole Doxastic logic defines in details how
-		Beliefs can be formalized.
+		beliefs can be formalized.
 	\item In \autocite{Hintikka1993Information}, Hintikka describes the concept
-		of Information and the difference with Knowledge and Belief.
+		of information and the difference with knowledge and belief.
 	\item In \autocite{Santaca2016abf}, the authors defines an agent as a
-		tuple of Assertions, Beliefs, and Facts.
+		tuple of assertions, beliefs, and facts.
 \end{itemize}
 For our argument, as in \autocite{Santaca2016abf}, an agent is composed by its
-knowledge, beliefs, and the information or assertions\fixnote{mv}{
-In generale, farei attenzione a uniformare upper/lower case per assertion, behavior, beliefs, ecc.
-} it provides; where
+knowledge, beliefs, and the information or assertions it provides; where
 knowledge is defined, as in \autocite{Steup2020epistemology}, as a set of
@@
 knowledge require truth, (iii) knowledge must be \emph{properly justified}, and
-the only objective of Information is to exchange beliefs between agents.  As
-defined in \autocite{Hintikka1993Information}, ``Information is specified by
+the only objective of information is to exchange beliefs between agents.  As
+defined in \autocite{Hintikka1993Information}, ``information is specified by
 specifying which alternatives concerning the reality it admits and which
@@
 defined as Believed to be True/False and not Believed to be the opposite.  Due
-to the definition of Information and then with its relation to the
+to the definition of information and then with its relation to the
 probabilistic correlation to truth/reality, we consider information in relation
 with an agent's beliefs. Similarly, we consider beliefs to define the actual
-behavior of an agent or a system.
-The difference between Knowledge and Information as used by Hintikka, and Facts
-and Assertions as used in \autocite{Santaca2016abf}, are not precisely discussed
+behavior of an agent or a system; i.e. we consider behaviors as belief revision
+systems.
+The difference between knowledge and information as used by Hintikka, and facts
+and assertions as used in \autocite{Santaca2016abf}, are not precisely discussed
 in \autocite{Santaca2016abf}.  Furthermore, those epistemological concepts are
@@
 
-We now define them abstractly, inspired by the
+We now define those concepts abstractly, inspired by the
 $\abf$-framework  defined in \autocite{Santaca2016abf} and, in
@@
 \begin{enumerate}
-	\item We consider Information only when the intention of exchanging 
-		that Information is from a sender to 
-		recipient is defined; and we call it \emph{Assertion}  
-	\item Similarly, the portion of \emph{Beliefs} we consider for system
+	\item We consider information only when the intention of exchanging 
+		that information is from a sender to 
+		recipient is defined; and we call it \emph{assertion}  
+	\item Similarly, the portion of \emph{beliefs} we consider for system
 		engineering is the one that builds (input) or describes
@@
 		and
-	\item We consider a set of axiomatic \emph{Facts} (definitory
+	\item We consider a set of axiomatic \emph{facts} (definitory
 		rules\footnote{Facts are definitory rule as they don't
@@
 		instead of considering the more general epistemic definition of
-		Knowledge. Specifically, Facts describes:
+		knowledge. Specifically, facts describes:
 		\begin{itemize}
@@
 
-As depicted in Figure~\ref{fig:soundness}, the relation between Facts, and
-Assertions and Behavior\fixnote{mv}{io ho inteso A B e F come Assertions, Beliefs e Facts, rispettivamente; di tanto in tanto tornate a usare il termine Behavior e questo secondo me confonde} 
-defines the soundness of the design (admissible
-configurations of Assertions, Behaviors and, in turn, Beliefs) w.r.t. the
-specification (what the specification mandates, such as by the nature of the
-physical/functional architectural models). We categorize agents by first
+In Figure~\ref{fig:soundness},
+the relation between facts, and
+assertions and beliefs (as inputs and outputs of the Behavior of an agent)
+is illustrated. assertions and beliefs generated by the design of a system
+may not be exactly aligned with what the facts mandate (i.e. what the specification mandates.
+The relation between facts, and assertions and beliefs 
+can be used as a metric to determine the soundess of the 
+design with respect to the specification.
+
+We categorize agents by first
 analyzing the relations between each pair of Regions defining an agent (i.e.
@@
 	\includegraphics[width=.9\columnwidth]{soundness.pdf}
-	\caption{Example relation between Facts, and Assertions and Beliefs}
+	\caption{Example relation between facts, and assertions and beliefs}
 	\label{fig:soundness}
@@
 \paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration} In order
-to reason on the relation between assertions and behavior we first need to
+to reason on the relation between assertions and beliefs we first need to
 consider that, by definition, assertions are defined as transfer of information
@@
 	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\beliefRegion$},
-		where the relation between Input-assertions and behavior
+		where the relation between Input-assertions and beliefs
 		describes the soundness of the execution of the functional
@@
 		the expected inputs, defines the functional behavior of an
-		ideal system.  If all the inputs (Assertions) are correctly
-		handled in the functional specification (behavior) the
+		ideal system.  If all the inputs (assertions) are correctly
+		handled in the functional specification (beliefs) the
 		specification is complete.
@@
 \Rcc{$\beliefRegion$}{$\factRegion$} -- Honesty and Competence} 
-The relation of Assertions
-and Behavior with Facts determines the quality with respect to the nominal
-(specified) system.  Given that Facts defines what needs to be true in the system,
-the relation of Assertions an Facts determines the degree of quality between
+The relation of assertions
+and beliefs with facts determines the quality with respect to the nominal
+(specified) system.  Given that facts defines what needs to be true in the system,
+the relation of assertions an facts determines the degree of quality between
 the real information circulating in a system (or within an agent) and
 the one specified.  Since the transfer of information
-through Assertions generates Beliefs, a dishonest agent may circulate false
-information, generating false Beliefs.  We note that it is often implied that
+through assertions generates beliefs, a dishonest agent may circulate false
+information, generating false beliefs.  We note that it is often implied that
 the intention behind circulating false information discerns a dishonest and an
@@
 true.'' Sextus Empiricus, Outlines of Pyrrhonism,
-II-83\autocite{Empiricus1990Pyrrhonism}}.  The relation between Beliefs and
-Facts determines the competence (on the subjects defined by the Facts) of an
-agent (i.e. the more competent an agent is, the more likely a Belief of that
+II-83\autocite{Empiricus1990Pyrrhonism}}.  The relation between beliefs and
+facts determines the competence (on the subjects defined by the facts) of an
+agent (i.e. the more competent an agent is, the more likely a belief of that
 agent is true).
@@
 	requirements, is defined as $\eq{\assertionRegion}{\factRegion}$. 
-	\item[SE-$3$] The proper adherence of the behavior with respect to
+\item[SE-$3$] The proper adherence of the behavior (in terms of input and output beliefs) with respect to
 		requirements is defined as $\eq{\beliefRegion}{\factRegion}$.
@@
 		the device is considered as an agent.
-	\item Agents reason over Beliefs (i.e. transform Beliefs into other
-		Beliefs) and for each \emph{component of a CPS} (system, sub-system,
+	\item Agents reason over beliefs (i.e. transform beliefs into other
+		beliefs) and for each \emph{component of a CPS} (system, sub-system,
 		or device) is composed by a \emph{functional architecture} that
-		transforms input-Beliefs into output-Beliefs.
+		transforms input-beliefs into output-beliefs.
 	\item Components of a CPS have \emph{ports} to exchange information
 		with the outer environment (which may be a sub-system),
-		similarly to Agents. The transfer of information in a CPS is
+		similarly to agents. The transfer of information in a CPS is
 		defined by \emph{channels}.
-	\item The concept of Knowledge is related to the \emph{requirements}
+	\item The concept of knowledge is related to the \emph{requirements}
 		that, in a factorial way, describe how the CPS shall behave and
@@
 	types of ports with the following behavior: an input-port changes
-	transforms Assertions from a sender $s$ ($\rassert{s}{a}$) to Beliefs
+	transforms assertions from a sender $s$ ($\rassert{s}{a}$) to beliefs
 	of an agent $a$ ($\beliefRegion_a$), while an output-port from belief
@@
 $\Rcc{\assertionRegion_r}{\beliefRegion_a}$. A port is, in fact, syntactic sugar to express the 
-relation between Assertions and Beliefs. We only define secure input port
+relation between assertions and beliefs. We only define secure input port
 but the definition of a secure output port is symmetrical.
@@
 	every open region can be written as union of elements of that base. 
-	We distinguish between Regions that contain Information and Regions that
+	We distinguish between Regions that contain information and Regions that
 	don't by writing the latter as $\emptyset$.
@@
 were considered an agent (channel-agent) then the question would be how an
-agent would transfer its Assertions to the channel-agent. If the channel
+agent would transfer its assertions to the channel-agent. If the channel
 between the agent and the channel-agent is again an agent, we would generate an
@@
 prevents us from introducing additional constraints to the channel as storing
-Assertions that are transferred over the channel, or filter out some
+assertions that are transferred over the channel, or filter out some
 input-assertions.}.
@@
 	A mono-directional channel between the two agents ($s \rightarrow r$) is defined as 
-	an agent which behavior is (dogmatically) defined as: to forward any Assertion received from $s$ over an input-port, to  
+	an agent which behavior is (dogmatically) defined as: to forward any assertion received from $s$ over an input-port, to  
 	the output-port where $r$ is listening to.
@@
 The \emph{quality of a mono-directional} channel is defined as the $\rcc$
-relation between the Assertions made by the sender and the ones received by the
+relation between the assertions made by the sender and the ones received by the
 receiver, i.e. $\Rcc{\assertionRegion_s}{\assertionRegion_r}$.
@@
 Given that a mono-directional bottom-channel is assumed to be perfectly
-forwarding any Assertion from its input-port to its output-port, there is no
+forwarding any assertion from its input-port to its output-port, there is no
 insecure behavior but only the combination of the weaknesses of the input and
@@
 the functional architecture as described afterwards.  As depicted in
-Figure~\ref{fig:atom}, so far we have considered Information generated by a
-Port $P_I$ and then sent through a channel $C$ to another (recipient) port
-$P_O$. In this scenario where Ports and then Channels are atomic (otherwise
-raising infinite regress), we can only  consider the relations between Ports
-and Channel; considering both input-port to channel and channel to output-port.
+Figure~\ref{fig:atom}, so far we have considered information generated by a
+port $P_I$ and then sent through a channel $C$ to another (recipient) port
+$P_O$. In this scenario where ports and then channels are atomic (otherwise
+raising infinite regress), we can only  consider the relations between ports
+and channel; considering both input-port to channel and channel to output-port.
 In fact, the Weaknesses of a channel are defined in terms of the Weaknesses of
-Ports.  Io order to define a functional block
+ports.  Io order to define a functional block
 without encountering a infinitely recursive definition, we must reach the same
-conclusions as for the Channel. Therefore, describing the information as
+conclusions as for the channel. Therefore, describing the information as
 flowing over a channel or in a functional block is purely syntactic sugar.
 
-We can summarize these results by saying that the relations between Assertions
-and Beliefs, Output Assertions of an Agent and Input Assertions of another
-Agent, or Output Beliefs of a Functional Block and Input Beliefs of another
-Block can \emph{only} be affected by the following weaknesses: replace, drop,
+We can summarize these results by saying that the relations between assertions
+and beliefs, Output assertions of an agent and Input assertions of another
+agent, or Output beliefs of a functional block and Input beliefs of another
+block can \emph{only} be affected by the following weaknesses: replace, drop,
 injection, insertion, selective drop, and selective drop + insertion. 
 We call this the RIDI-Hypothesis, being the
-four main categories of weaknesse: Replace, Insertion, Drop, Injection. We can,
+four main categories of weakness: Replace, Insertion, Drop, Injection. We can,
 then, deduce the following security properties to mitigate security weaknesses
@@
 \begin{itemize}
-	\item \emph{Order-preserving} -- it shall be known if Information is \emph{replaced}
-	\item \emph{Availability} -- it shall be known if Information is \emph{dropped} or \emph{selectively dropped}
-	\item \emph{Integrity} -- it shall be known if Information is \emph{injected}
-	\item \emph{Authentication} -- it shall be known if Information is \emph{inserted}
+	\item \emph{Order-preserving} -- it shall be known if information is \emph{replaced}
+	\item \emph{Availability} -- it shall be known if information is \emph{dropped} or \emph{selectively dropped}
+	\item \emph{Integrity} -- it shall be known if information is \emph{injected}
+	\item \emph{Authentication} -- it shall be known if information is \emph{inserted}
 \end{itemize}
 %We can also generalize the weaknesses related to the data flow (i.e. the flow
-%of Information as Assertions or Beliefs) as the relation between output-input
-%Beliefs between functional blocks, output-input Assertions of a channel,
+%of information as assertions or beliefs) as the relation between output-input
+%beliefs between functional blocks, output-input assertions of a channel,
 %inputs-outputs of a port, as follows.
@@
 \paragraph{Functional Architecture}
-A Functional Architecture takes Information as input-Beliefs and transforms the
-Information into output-Beliefs. Those transformations occurs within the
-Functional Architecture, where Functional Blocks transforms Beliefs into other
-Beliefs. Similarly to channels, we could consider a Functional Block as a
-Functional Architecture occurring in an infinite regress. Therefore, we
-consider Functional Blocks as executing an abstract undefined behavior, of
-which we only observe the inputs and the resulting outputs (Beliefs).
+A functional architecture takes information as input-beliefs and transforms the
+information into output-beliefs. Those transformations occurs within the
+functional architecture, where functional blocks transforms beliefs into other
+beliefs. Similarly to channels, we could consider a functional block as a
+functional architecture occurring in an infinite regress. Therefore, we
+consider functional blocks as executing an abstract undefined behavior, of
+which we only observe the inputs and the resulting outputs (beliefs).
 
@@
 	A functional block of an agent 
-	takes a region of input-Beliefs and outputs a region of output-Beliefs. 
+	takes a region of input-beliefs and outputs a region of output-beliefs. 
 	A functional architecture is an interconnected system of functional blocks.
@@
 Therefore, while a port semantics is determined by the relation 
-between Assertions and Beliefs, the semantics of a Functional Block 
-is determined by the relation between Facts/Requirements and I/O Beliefs.
+between assertions and beliefs, the semantics of a functional block 
+is determined by the relation between facts/requirements and I/O beliefs.
 
@@
 During the specification phase, for any agent, channel, port, functional block
-and architecture, there may exist a factorial requirement (Fact) predicating over them.
-In other words, any requirement is defined as a Fact since they must be true in
+and architecture, there may exist a factorial requirement (fact) predicating over them.
+In other words, any requirement is defined as a fact since they must be true in
 any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
-Figure~\ref{fig:soundness}, Facts are strategic rules that defines how
+Figure~\ref{fig:soundness}, facts are strategic rules that defines how
 the system shall behave (by specification), while reality may be shown to
@@
 the requirements (i.e. $\Rcc{\beliefRegion}{\factRegion}$ is determined by
-the relations between the I/O Beliefs $sum(b_1,b_2)=b_3$ and the requirement
+the relations between the I/O beliefs $sum(b_1,b_2)=b_3$ and the requirement
 $sum(b_1,b_2)=b_1+b_2$, as follows.
@@
 \paragraph{Assertions and Facts}
-The whole reasoning on the relation between Beliefs and Facts can be duplicated
-for the relation between Assertions and Facts; we cannot appreciate the
+The whole reasoning on the relation between beliefs and facts can be duplicated
+for the relation between assertions and facts; we cannot appreciate the
 difference at this level of abstraction.  If the functional architecture would
 be extended to capture the semantics (i.e. the logic) of the communication and
-security protocols, with the relation between Assertions and Facts we would
+security protocols, with the relation between assertions and facts we would
 compare protocol logics with the requirements. We won't consider the verification of
@@
 		Architecture}. Each agent is defined as a Node with (physical)
-		Ports, and Agent's ports are connected via Information Flow
+		ports, and agent's ports are connected via information Flow
 		connectors, representing the physical channel.
@@
 		deployment diagram and is defined by an \emph{Object Diagram}.
-		The Object Diagram is composed by Instances of Functional
-		Blocks, connected via Information Flow connectors.
+		The Object Diagram is composed by Instances of functional
+		blocks, connected via information Flow connectors.
 	\item the connection between the two diagrams is implemented by
@@
 The problem is correlated to the epistemological search for truth,
-where the challenge is to relate Information and Belief to Human Knowledge.
-Scientifically, generalizing the Human to an Agent, the problem is
-to relate Assertions and Behaviors, to Facts. From an engineering standpoint,
-by generalizing the concept of Agents as architectural subsystems, the
-problem is to link Channels (and Ports) and Functional Architectures to 
-Requirements.
+where the challenge is to relate information and belief to human knowledge.
+Scientifically, generalizing the human to an agent, the problem is
+to relate assertions and beliefs, to facts. From an engineering standpoint,
+by generalizing the concept of agents as architectural subsystems, the
+problem is to link channels (and ports) and functional architectures to 
+requirements.
 

### 2020-07-19 reports/paper_0/main.tex
Luca

@@
 %number of sessions (even tough some approaches do reason on an infinite number
-%of sessions such as \autocite{Escobar2007maudenpa}),\fixn{Luca}{Nowadays, most
+%of sessions such as \autocite{Escobar2007maudenpa}),\fixnote{Luca}{Nowadays, most
 %tools work for infinite sessions: Maude-NPA, Tamarin, ProVerif, OFMC.} (iii)
 %honesty/dishonesty of the peers (e.g.  in the ASLan++ language
-%\autocite{Oheimb2010aslan++}),\fixn{Luca}{Nowadays, most tools allow for easy
+%\autocite{Oheimb2010aslan++}),\fixnote{Luca}{Nowadays, most tools allow for easy
 %switch of the resulting sessions} and (iv) the abstraction of the
 %cryptographic primitives (e.g.  ProVerif vs CryptoVerif
-%\autocite{Blanchet2017symbolic}).\fixn{Luca}{This is the only claim that, more
+%\autocite{Blanchet2017symbolic}).\fixnote{Luca}{This is the only claim that, more
 %or less, still holds, even though more and more tools do cryptoanalysis. The
@@
 %improved those issues, there is no agreement on which should be the definitive
-%approach (if any).\fixn{Luca}{I don't think that this is important. In many
+%approach (if any).\fixnote{Luca}{I don't think that this is important. In many
 %fields of research you don't have a single definitive approach and it is good
@@
 \section{Literature Review}\label{sec:literature}
+\fixnote{Luca}{I think that this section needs thorough rewriting.
+First of all, you are trying to say too many things, many of which are not really relevant for what you want to say.
+Second, you are not just reviewing but also introducing some new things that are part of your contributions and this is quite confusing.
+Third, there are some claims that, in my view, are wrong or at best misleading.
+Shorten, simplify, aiming to distill really only what is relevant.
+As it is now, the text is quite confused and confusing.}
 \begin{figure*}[t]
@@
 natural environment (e.g. natural events such as wearing of hardware
-components) while security deals with \emph{incidents} posed by mankind (e.g.
+components), whereas security deals with \emph{incidents} posed by mankind (e.g.
 attackers and bugs). The fundamental difference between nature and mankind (and,
-in turn, between safety and cybersecurity) is believed to be on the different
-intents (incidents are intentional while accidents are not) of the causes
-that generates a threat; namely, nature is believed not to have malicious
-intents (but unfortunate causes-effects) while threats generated by mankind are
-malicious\footnote{Of course, logical flaws or bugs may be
+in turn, between safety and cybersecurity\fixnote{Luca}{You spoke of security before, which is much wieder than cybersecurity.
+This is something you have to be very careful with...
+political scientists would correctly claim that security is much more than just cyber and that their term is the correct one to contrappose to safety}) is believed to be on the different
+intents (incidents are intentional, whereas accidents are not) of the causes
+that generates a threat. Namely, nature is believed not to have malicious
+intents (but unfortunate causes-effects), whereas threats generated by mankind are
+malicious.\footnote{Of course, logical flaws or bugs may be
 introduced by other means (e.g. ignorance) without explicit malicious intents,
 but the exploitation of those flaws is considered (for now, and detailed
-afterwards in the article) malicious, and then we consider any vulnerability to
-be malicious even if due to the lack of skills.}.
+afterwards in the article) malicious, and thus we consider any vulnerability to
+be malicious even if it is due to the lack of skills.}.\fixnote{Luca}{Secondo me tutta questa prima discussione \`e alquanto inutile e si pu\`o riassumere in un'unica frase.}
 This conclusion seems to be against a general formulation of a cybersecurity theory;
 how can we define a theory that predicts what a human being will do?
-Cybersecurity attacks seems to be related to the creativity of the attacker
-and then unpredictable. Currently, our understanding of cybersecurity is
+Cybersecurity attacks seem to be related to the creativity of the attacker
+and thus unpredictable. Currently, our understanding of cybersecurity is
 stored into a network of databases of weaknesses (e.g.
-CWE\autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE\autocite{CVE},
-NVD\autocite{NIST2020NVD}),
-and attacks (e.g. CAPEC\autocite{MITRE2020CAPEC}, ATT\&CK\autocite{MITRE2020ATTACK}). Systems are
+CWE~\autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE~\autocite{CVE},
+NVD~\autocite{NIST2020NVD}),
+and attacks (e.g.
+CAPEC~\autocite{MITRE2020CAPEC}, ATT\&CK~\autocite{MITRE2020ATTACK}).\fixnote{Luca}{Come detto, sono fortemente in disaccordo con questa frase.
+Parte del nostro understanding \`e l\`i, ma non tutto.
+Tutto il lavoro teorica e formale non \`e contenuto l\`i.
+Mi sembra tra l'altro un'affermazione alquanto inutilmente provocatoria} Systems are
 nowadays tested against 
 known attacks, ignoring that considering a system secure
-if no known attack is possible is against the scientific method.
+if no known attack is possible is against the scientific method.\fixnote{Luca}{No, non \`e una scelta ma \`e semplicemente ignoranza di chi fa i test.
+Chi fa i test seriamente questo lo sa benissimo.
+E non solo perch\'e \`e contro lo scientific methos ma anche perch\'e testing it not proving}
 As an example, if we have a system which is secure against all of the
-vulnerabilities in the CVE we just need to wait for a month to have 1000+ more
+vulnerabilities in the CVE we often just need to wait for a month to have 1000+ more
 vulnerabilities to test \autocite{NVD2020dashboard}. If even just one of those new vulnerabilities
-affects our system, that falsify the hypothesis that a
-system is secure if no known attacks apply. 
+affects our system, that falsifies the hypothesis that a
+system is secure if no known attacks apply.\fixnote{Luca}{But, again, the hypothesis was utterly wrong from the start, even without considering the scientific method.
+No person who is really knowledgeable of testing would ever claim that testing shows the absence of all attacks.
+Maybe this is what is done in industry, but then industry is simply wrong (it would not be the first time) and people in academia have been saying so for ages.
+It is not you argueing it..
+I believe that building the argument as you do is wrong.
+I would suggest: (1) Industry uses testing believing that it is enough.
+(2) It is well know that testing cannot show the absence of attacks.
+(3) The scientific method makes this even more evident and in fact we show ...}
 We argue that this ``implement and test'' approach should be changed.
@@
 the secure-by-design approach they apply can be considered as going
-into the opposite direction of the ``implement and test'' approach.
+into the opposite direction of the ``implement and test'' approach.\fixnote{Luca}{Questa ultima frase \`e appesa qui senza una vera chiara connessione a quanto detto finora.}
 
@@
 or 61511\autocite{IEC201761511}), are based on empirical tests and measurements
-(therefore should be considered hypothesis and not definitions). While
+(therefore should be considered hypotheses and not definitions).\fixnote{Luca}{Hypotheses and definitions don't type check. Moreover, many standards require formal methods proofs, don't they?} While
 reasoning by induction
@@
 to false beliefs, this approach is often justified by the supposed
-impossibility of defining a theory that correctly predicts failures.  A failure
+impossibility of defining a theory that correctly predicts failures.\fixnote{Luca}{This sentence does not follow grammatically}  A failure
 of a wire due to environment (e.g. due to humidity, dust, heat \&c) is defined
@@
 Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
-the terms that structure our current understanding of security. 
+the terms that structure our current understanding of security.\fixnote{Luca}{I disagree.
+It should be cybersecurity, not security.
+Moreover, as I mentioned, the figure is misleading when it talks about blac-hat hacking.
+I also don't like ``Manking'', as mankind does many of the things, not just designing systems}
 \begin{itemize}
@@
 		animals, and other features of the world develop of their own
-		accord'' (e.g. the physical universe)\autocite{wiki-nature}. 
+		accord'' (e.g. the physical universe)~\autocite{wiki-nature}. 
 		\begin{itemize}
@@
 				considering those terms to be semantically
-				equivalent.  From now on, we consider the
+				equivalent.\fixnote{Luca}{Grammar issues}  From now on, we consider the
 				Causality principle to be the \emph{threat
@@
 		supporting semantics.}, as defined in \autocite{cnssi20104009}
-		(and adopted in \autocite{nist2013800-53}), is ``weakness in an
+		(and adopted in \autocite{nist2013800-53}), is a ``weakness in an
 		information system, system security procedures, internal
@@
 		Furthermore, the term ``threat sources'' used in the definition
-		in\autocite{cnssi20104009} may be identified with both Nature
+		in~\autocite{cnssi20104009} may be identified with both Nature
 		and Mankind, not differentiating between safety and security.
@@
 channel that allows the violation of the security policy
-[\ldots]''\autocite{cnssi20104009}. For example, a software or procedure that takes
+[\ldots]''~\autocite{cnssi20104009}. For example, a software or procedure that takes
 advantage of the vulnerability causing an \emph{attack} to the system may
@@
 		weaknesses are errors that can lead to software
-		vulnerabilities. A software vulnerability, such as those
+		vulnerabilities.\fixnote{Luca}{Worth to say that many security issues are not software weakenesses, e.g.
+cryptography issues or the way cryptography is used.
+Moreover, in cyber-physical systems, security issues might be due to hardware issues.
+Again, I think that you are trying to say too many things.
+Focus on what you really want to say.} A software vulnerability, such as those
 		enumerated on the Common Vulnerabilities and Exposures (CVE)
@@
 	\item \emph{Causality} refers to the causality principle; defined
-		in\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
+		in~\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
 		connection of phenomena through which one thing (the cause)
@@
 	\item An \emph{Exploit}\footnote{We note that the term exploit is only
-		used as a verb in\autocite{ISO2009information}} ``[\ldots]
+		used as a verb in~\autocite{ISO2009information}}\fixnote{Luca}{So what? Is it The use in~\autocite{ISO2009information} is incomplete? } ``[\ldots]
 		(from the English verb to exploit, meaning to use something to
@@
 		occur on computer software, hardware, or something electronic
-		(usually computerized).''\autocite{wiki-exploit}.
+		(usually computerized).''~\autocite{wiki-exploit}.
 	\item An \emph{Attack}, as defined by the International Standard
-		ISO/IEC 27000 is an ``attempt to destroy, expose, alter,
+		ISO/IEC 27000, is an ``attempt to destroy, expose, alter,
 		disable, steal or gain unauthorized access to or make
@@
 		In fact, we consider the term \emph{hack} as
-		non-malicious (as, e.g. in \autocite{Stallman2002hacker}).
-	\item A \emph{Threat}, as defined in\autocite{cnssi20104009}, is ``Any
+		non-malicious (as, e.g. in~\autocite{Stallman2002hacker}).
+	\item A \emph{Threat}, as defined in~\autocite{cnssi20104009}, is ``Any
 		circumstance or event with the potential to adversely impact
@@
 		information, and/or denial of service''.
-	\item \emph{Defect}, ``anything that renders the product not reasonably
-		safe''\autocite{Robinson2019writing} (i.e. a characteristic of
+	\item \emph{Defect}: ``anything that renders the product not reasonably
+		safe''~\autocite{Robinson2019writing} (i.e. a characteristic of
 		an object which hinders its proper usability).
-	\item \emph{Failure}, as defined in\autocite{Merriam2020failure} as ``a state of
+	\item \emph{Failure}, as defined in\autocite{Merriam2020failure}, is ``a state of
 		inability to perform a normal function''. The term is
 		structured and detailed in
-		\autocite{cnssi20104009,iet2017glossary} but relying on an
+		\autocite{cnssi20104009,iet2017glossary}, but relying on an
 		abstract notion of failure without a specific definition.
 	\item \emph{Hazard}, ``a potential source of
-		harm''\autocite{iet2017glossary}.
+		harm''~\autocite{iet2017glossary}.
 \end{itemize}
@@
 
-Our literature review shows that most of the definitions relates insecurity to
+Our literature review shows that most of the definitions relate insecurity to
 dis-honesty (also called maliciousness or adversarial) of an agent (often
-called adversary or attacker). This, however, just moves the problem of
-defining what cybersecurity is as the problem of defining what an dishonest
-agent can or cannot do in a system\footnote{where an agent is any virtual or
+called adversary or attacker).\fixnote{Luca}{This is slightly misleading, as you have not discussed this above.
+You mean in the review that you did you discovered that...} This, however, just moves the problem of
+defining what cybersecurity is as the problem of defining what a dishonest
+agent can or cannot do in a system, where an agent is any virtual or
 physical entity of the system or using the system (e.g. a device, a software,
 or a human being) and dishonesty is not necessarily related to malicious
-motivation but also to incompetence or lack of skills.} As in the Dolev-Yao
+motivation but also to incompetence or lack of skills. As in the Dolev-Yao
 theory, we may correlate ``being dishonest'' to ``not following the intended
@@
 and its high level of abstraction, we may conclude that the hypothesis seems
-evident. For example, a software can be considered an agent of the system and
+evident.\fixnote{Luca}{Which hypothesis?} For example, a software can be considered an agent of the system and
 whenever it has a bug, it can be exploited causing an Incident. However, this
@@
 and systems) and that huge repository of dishonest behaviors should be kept as
-definition, such as CAPEC\autocite{MITRE2020CAPEC}; or that dishonesty of an agent
+definition, such as CAPEC~\autocite{MITRE2020CAPEC}; or that dishonesty of an agent
 with respect to a security protocol should be defined as a number of predefined
@@
 \autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
-(to name a few). As depicted in Figure~\ref{fig:causality}, in order to define
+(to name a few).\fixnote{Luca}{I foudn this sentence very hard to follow.
+Moreover, I would be careful to speak of ``predefined actions'' as basically the DY can do anything so it is not cleaer what you are referring to.
+Of course, you could this that this excludes, e.g., human errors (say as in the work I did with Diego) or more, but then you should clarify.} As depicted in Figure~\ref{fig:causality}, in order to define
 a theory on cybersecurity and on the malicious behaviors of agents we noticed
 that Attack Patterns (or Attacks) are caused-by the presence of Vulnerabilities
-in the system. Those Vulnerabilities are, in turn, cause-by the presence of
+in the system.\fixnote{``cause-by'' is grammatically wrong, but more importantly the presence of a vulnerability does not necessarily cause an attack, but allows it.
+Moreover, you speak of Patterns for the first time.} Those Vulnerabilities are, in turn, cause-by the presence of
 Weaknesses in the system.  Weaknesses are Errors in the design or

### 2020-07-20 reports/paper_0/main.tex
fixed Luca's comments on intro of sec 2 (literature review)

@@
 	the engineering concepts of requirements, functional architectures, and
-	channels.  Our theory allows the definition of the security of a system
-	as a mathematical formula correlating the security and insecurity
-	configurations of a system.  We implemented a prototype cybersecurity
+	channels.  Ultimately, our theory allows the definition of cybersecurity of a system
+	as a mathematical formula.  We implemented a prototype cybersecurity
 	risk assessment tool that, based on our theory, predicts weaknesses in
@@
 we be sure that the theory correctly characterizes the phenomenon it describes? 
-
 Popper's famous ``demarcation principle'' draws a line between scientific and
@@
 observation that allows us to declare an arbitrary system or technique
-secure''. His intuition is that there is no scientific theory that can predict
+secure''. With security, Herley (as in the reminder of this paper) only 
+focuses on cybersecurity and his intuition is that there is no scientific theory that can predict
 the security of a system, nor a theory that can predict all possible
-insecurities of a system (which, by negation would become a theory on
+insecurities of a system (which, by negation may be used as a theory of
 security).  Herley then uses this argument to show that ``claims that any
@@
 \section{Literature Review}\label{sec:literature}
-\fixnote{Luca}{I think that this section needs thorough rewriting.
-First of all, you are trying to say too many things, many of which are not really relevant for what you want to say.
-Second, you are not just reviewing but also introducing some new things that are part of your contributions and this is quite confusing.
-Third, there are some claims that, in my view, are wrong or at best misleading.
-Shorten, simplify, aiming to distill really only what is relevant.
-As it is now, the text is quite confused and confusing.}
 \begin{figure*}[t]
@@
 attackers and bugs). The fundamental difference between nature and mankind (and,
-in turn, between safety and cybersecurity\fixnote{Luca}{You spoke of security before, which is much wieder than cybersecurity.
-This is something you have to be very careful with...
-political scientists would correctly claim that security is much more than just cyber and that their term is the correct one to contrappose to safety}) is believed to be on the different
+in turn, between safety and cybersecurity) is believed to be on the different
 intents (incidents are intentional, whereas accidents are not) of the causes
@@
 intents (but unfortunate causes-effects), whereas threats generated by mankind are
-malicious.\footnote{Of course, logical flaws or bugs may be
+malicious\footnote{Of course, logical flaws or bugs may be
 introduced by other means (e.g. ignorance) without explicit malicious intents,
@@
 afterwards in the article) malicious, and thus we consider any vulnerability to
-be malicious even if it is due to the lack of skills.}.\fixnote{Luca}{Secondo me tutta questa prima discussione \`e alquanto inutile e si pu\`o riassumere in un'unica frase.}
+be malicious even if it is due to the lack of skills.}.
 This conclusion seems to be against a general formulation of a cybersecurity theory;
@@
 Cybersecurity attacks seem to be related to the creativity of the attacker
-and thus unpredictable. Currently, our understanding of cybersecurity is
-stored into a network of databases of weaknesses (e.g.
+and thus unpredictable.
+
+Currently, the most complete understanding of insecurity issues, is stored into
+a network of databases of weaknesses (e.g.
 CWE~\autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE~\autocite{CVE},
-NVD~\autocite{NIST2020NVD}),
-and attacks (e.g.
-CAPEC~\autocite{MITRE2020CAPEC}, ATT\&CK~\autocite{MITRE2020ATTACK}).\fixnote{Luca}{Come detto, sono fortemente in disaccordo con questa frase.
-Parte del nostro understanding \`e l\`i, ma non tutto.
-Tutto il lavoro teorica e formale non \`e contenuto l\`i.
-Mi sembra tra l'altro un'affermazione alquanto inutilmente provocatoria} Systems are
-nowadays tested against 
-known attacks, ignoring that considering a system secure
-if no known attack is possible is against the scientific method.\fixnote{Luca}{No, non \`e una scelta ma \`e semplicemente ignoranza di chi fa i test.
-Chi fa i test seriamente questo lo sa benissimo.
-E non solo perch\'e \`e contro lo scientific methos ma anche perch\'e testing it not proving}
-As an example, if we have a system which is secure against all of the
-vulnerabilities in the CVE we often just need to wait for a month to have 1000+ more
-vulnerabilities to test \autocite{NVD2020dashboard}. If even just one of those new vulnerabilities
-affects our system, that falsifies the hypothesis that a
-system is secure if no known attacks apply.\fixnote{Luca}{But, again, the hypothesis was utterly wrong from the start, even without considering the scientific method.
-No person who is really knowledgeable of testing would ever claim that testing shows the absence of all attacks.
-Maybe this is what is done in industry, but then industry is simply wrong (it would not be the first time) and people in academia have been saying so for ages.
-It is not you argueing it..
-I believe that building the argument as you do is wrong.
-I would suggest: (1) Industry uses testing believing that it is enough.
-(2) It is well know that testing cannot show the absence of attacks.
-(3) The scientific method makes this even more evident and in fact we show ...}
-We argue that this ``implement and test'' approach should be changed.
-Even tough the field 
-of security protocols is still lacking of a general cybersecurity theory,
-the secure-by-design approach they apply can be considered as going
-into the opposite direction of the ``implement and test'' approach.\fixnote{Luca}{Questa ultima frase \`e appesa qui senza una vera chiara connessione a quanto detto finora.}
+NVD~\autocite{NIST2020NVD}), and attacks (e.g. CAPEC~\autocite{MITRE2020CAPEC},
+ATT\&CK~\autocite{MITRE2020ATTACK}).  Those insecurity issues can be related to
+the violation of one or more requirements (explicit or implicit) in the
+specification, design or implementation of a system. The correlation 
+between insecurity flaws and security requirements has been used to define
+standards such as the IEC 62443-1-3 (the Industrial communication networks -
+Network and system security -- Part 3-3: System security requirements and
+security levels) defines requirements as ``confidentiality of information
+in transit/at-rest''. More generally, the idea of defining security requirements
+as properties of a system was initially defined in 1970s with the 
+CIA-triad (Confidentiality, Integrity, Availability) and refined 
+over the decades introducing related concepts such as authenticity or
+non-repudiation, or introducing new ones such as ``responsibility'' in the
+RITE approach (see~\autocite{Samonas2014cia} for an overview of the evolution of the CIA-triad).
+
+Systems are
+nowadays tested against known or unknown attacks,
+or verified against security requirements such as confidentiality.
+However, no scientific falsifiable theory correlate
+security requirements to necessary and sufficient conditions (e.g. mitigations) 
+to declare a system secure.
 

### 2020-07-21 reports/paper_0/main.tex

@@
 
-Our hypothesis is that cybersecurity attacks are caused by 
+Instead of reasoning of what makes a system secure or insecure,
+we reason on what causes insecurities.
+Our hypothesis is that cybersecurity attacks are only caused by 
 the presence of errors in the design or implementation of a system.
@@
 obtaining a precise estimation of all potential cybersecurity-related risks
-in any given system.
+in any given system. Our hypothesis can be falsified by means of experiments,
+testing if all the predicted weaknesses are present in the system under
+test, or testing if other (not predicted by our hypothesis) weaknesses are
+present.
 
@@
 Network and system security -- Part 3-3: System security requirements and
-security levels) defines requirements as ``confidentiality of information
+security levels) which defines requirements as ``confidentiality of information
 in transit/at-rest''. More generally, the idea of defining security requirements
 as properties of a system was initially defined in 1970s with the 
-CIA-triad (Confidentiality, Integrity, Availability) and refined 
+CIA triad (Confidentiality, Integrity, Availability) and refined 
 over the decades introducing related concepts such as authenticity or
 non-repudiation, or introducing new ones such as ``responsibility'' in the
-RITE approach (see~\autocite{Samonas2014cia} for an overview of the evolution of the CIA-triad).
-
-Systems are
-nowadays tested against known or unknown attacks,
-or verified against security requirements such as confidentiality.
-However, no scientific falsifiable theory correlate
+RITE approach (see~\autocite{Samonas2014cia} for an overview of the evolution of the CIA triad).
+The link between security requirements and vulnerabilities is reported 
+in the NVD databases by the CVSS\autocite{Mell2007CVSS}) scoring system.
+The CVSS evaluates of the severity of a vulnerability
+by means of different metrics (such as attack complexity and user interaction)
+and quantitatively evaluates the impact on the CIA triad.
+While security requirements, weaknesses, vulnerabilities, and attacks
+have been extensively studied and implemented both in academia and industry
+to provide tools for the testing or verification of systems, 
+no scientific falsifiable theory correlate
 security requirements to necessary and sufficient conditions (e.g. mitigations) 
-to declare a system secure.
+to declare a system secure \autocite{Herley2016unfalsifiability}.
+Nonetheless, the extensive body of literature have scientific foundations,
+for example, providing either providing necessary or sufficient conditions
+for cybersecurity.
+As a driver for our argumentation, we start by reviewing the key concepts
+in the cybersecurity domain.
 
@@
 which safety requirements are defined (e.g. in standards such as the IEC 61508
-or 61511\autocite{IEC201761511}), are based on empirical tests and measurements
-(therefore should be considered hypotheses and not definitions).\fixnote{Luca}{Hypotheses and definitions don't type check. Moreover, many standards require formal methods proofs, don't they?} While
-reasoning by induction
+or 61511\autocite{IEC201761511}), are based on empirical tests and measurements.
+While reasoning by induction
 %\footnote{``So, whenever they argue ``Every man is an
@@
 %Sextus Empiricus, Outlines of Pyrrhonism II-195 \autocite{Empiricus1990Pyrrhonism}}
-based on the empirical observation should be avoided, since it may easily lead
-to false beliefs, this approach is often justified by the supposed
-impossibility of defining a theory that correctly predicts failures.\fixnote{Luca}{This sentence does not follow grammatically}  A failure
-of a wire due to environment (e.g. due to humidity, dust, heat \&c) is defined
-from empirical evidences and processes have been standardized to test qualities
-of hardware components.  This process completely breaks down when a malicious
-environment (i.e. an attacker) is considered instead of the (supposedly honest
-and predictable) natural environment. Therefore, the same approach that is in
-use for safety, seems not to be applicable for security (e.g. for security
-testing).
-An overview on the aforementioned aspects of safety and security is depicted in
+based on the empirical observation should be avoided in general,
+since it may easily lead to false beliefs \autocite{Popper1959logic}, 
+inducing general principle by means of tests
+is often justified by the supposed impossibility of defining a theory
+that correctly predicts failures.
+A failure of a wire due to environment (e.g. due to humidity,
+dust, heat \&c) is defined from empirical evidences and processes have been
+standardized to test qualities of hardware components.  This process completely
+breaks down when a malicious environment (i.e. an attacker) is considered
+instead of the (supposedly honest and predictable) natural environment.
+Therefore, the same approach that is in use for safety, seems not to be
+applicable for cybersecurity (e.g. for cybersecurity testing).  An overview on the
+aforementioned aspects of safety and cybersecurity is depicted in
 Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
-the terms that structure our current understanding of security.\fixnote{Luca}{I disagree.
-It should be cybersecurity, not security.
-Moreover, as I mentioned, the figure is misleading when it talks about blac-hat hacking.
-I also don't like ``Manking'', as mankind does many of the things, not just designing systems}
+the terms that structure our current understanding of cybersecurity.
+
 \begin{itemize}
@@
 			\item There exist several terms to refer to an
-				\emph{attacker}, i.e. threat agent or threat source,
-				considering those terms to be semantically
-				equivalent.\fixnote{Luca}{Grammar issues}  From now on, we consider the
+				\emph{attacker}, e.g. threat agent or threat source,
+				but, from now on, we consider the
 				Causality principle to be the \emph{threat
@@
 \begin{itemize}
-	\item \emph{Weakness}. The definition given by the MITRE
-		in \autocite{MITRE2020CWEweakness} of weakness is: ``Software
-		weaknesses are errors that can lead to software
-		vulnerabilities.\fixnote{Luca}{Worth to say that many security issues are not software weakenesses, e.g.
-cryptography issues or the way cryptography is used.
-Moreover, in cyber-physical systems, security issues might be due to hardware issues.
-Again, I think that you are trying to say too many things.
-Focus on what you really want to say.} A software vulnerability, such as those
-		enumerated on the Common Vulnerabilities and Exposures (CVE)
-		List, is a mistake in software that can be directly used by a
-		hacker to gain access to a system or network''.  The definition
-		is circular if we interpret the word ``error'' and ``mistake''
-		with the same semantics: a weakness is an error that leads to a
-		vulnerability and a vulnerability is a mistake which, in turn,
-		is a weakness.  
-\end{itemize}
-
-The only difference between a weakness and vulnerability seems to be that one
-can consider weakness as a ground term and state that a vulnerability is caused
-by a weakness.
-
-\begin{itemize}
+	\item \emph{Weakness}. The definition given by the MITRE in
+		\autocite{MITRE2020CWEweakness} of weakness is: `` a type of
+		mistake that, in proper conditions, could contribute to the
+		introduction of vulnerabilities within that product. This term
+		applies to mistakes regardless of whether they occur in
+		implementation, design, or other phases of a product
+		lifecycle.'' A vulnerability, such as those enumerated on the
+		Common vulnerabilities and Exposures (CVE) List, is a mistake
+		that can be directly used by a hacker to gain access to a
+		system or network.  The definition is circular if we interpret
+		the word ``error'' and ``mistake'' with the same semantics: a
+		weakness is an error that leads to a vulnerability and a
+		vulnerability is a mistake which, in turn, is a weakness. The
+		only difference between a weakness and vulnerability seems to
+		be that one can consider weakness as a ground term and state
+		that a vulnerability is caused by a weakness.
 	\item \emph{Causality} refers to the causality principle; defined
@@
 		%causality principle to be formalized by the K Modal Logic.
-	\item An \emph{Exploit}\footnote{We note that the term exploit is only
-		used as a verb in~\autocite{ISO2009information}}\fixnote{Luca}{So what? Is it The use in~\autocite{ISO2009information} is incomplete? } ``[\ldots]
+	\item An \emph{Exploit} ``[\ldots]
 		(from the English verb to exploit, meaning to use something to
@@
 dis-honesty (also called maliciousness or adversarial) of an agent (often
-called adversary or attacker).\fixnote{Luca}{This is slightly misleading, as you have not discussed this above.
-You mean in the review that you did you discovered that...} This, however, just moves the problem of
+called adversary or attacker). In fact, weaknesses become vulnerability if an
+attacker exploits them in an attack.  This, however, just moves the problem of
 defining what cybersecurity is as the problem of defining what a dishonest
-agent can or cannot do in a system, where an agent is any virtual or
-physical entity of the system or using the system (e.g. a device, a software,
-or a human being) and dishonesty is not necessarily related to malicious
-motivation but also to incompetence or lack of skills. As in the Dolev-Yao
-theory, we may correlate ``being dishonest'' to ``not following the intended
-behavior/rules''.  In the case of a generic system, a dishonest agent is,
-therefore, any agent that doesn't follow the intended behavior (or
-functionality or logic) of the system.  Given the generality of the definition,
-and its high level of abstraction, we may conclude that the hypothesis seems
-evident.\fixnote{Luca}{Which hypothesis?} For example, a software can be considered an agent of the system and
+agent can or cannot do in a system, where an agent is any virtual or physical
+entity of the system or using the system (e.g. a device, a software, or a human
+being) and dishonesty is not necessarily related to malicious motivation but
+also to incompetence or lack of skills. As in the Dolev-Yao theory, we may
+correlate ``being dishonest'' to ``not following the intended behavior/rules''.
+In the case of a generic system, a dishonest agent is, therefore, any agent
+that doesn't follow the intended behavior (or functionality or logic) of the
+system.  For example, a software can be considered an agent of the system and
 whenever it has a bug, it can be exploited causing an Incident. However, this
-hypothesis has lead to the un-testable conclusion that the dishonest behaviors
-of agents cannot be defined in general (e.g. due to the heterogeneity of agents
-and systems) and that huge repository of dishonest behaviors should be kept as
-definition, such as CAPEC~\autocite{MITRE2020CAPEC}; or that dishonesty of an agent
-with respect to a security protocol should be defined as a number of predefined
-actions as in
+may lead to the conclusion that the dishonest behaviors of agents cannot be
+defined in general (e.g. due to the heterogeneity of agents and systems) and
+that huge repository of dishonest behaviors should be kept as definition, such
+as CAPEC~\autocite{MITRE2020CAPEC}; or that dishonesty of an agent with respect
+to a security protocol should be defined as a number of predefined actions as
+in
 \autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
-(to name a few).\fixnote{Luca}{I foudn this sentence very hard to follow.
-Moreover, I would be careful to speak of ``predefined actions'' as basically the DY can do anything so it is not cleaer what you are referring to.
-Of course, you could this that this excludes, e.g., human errors (say as in the work I did with Diego) or more, but then you should clarify.} As depicted in Figure~\ref{fig:causality}, in order to define
-a theory on cybersecurity and on the malicious behaviors of agents we noticed
-that Attack Patterns (or Attacks) are caused-by the presence of Vulnerabilities
-in the system.\fixnote{``cause-by'' is grammatically wrong, but more importantly the presence of a vulnerability does not necessarily cause an attack, but allows it.
-Moreover, you speak of Patterns for the first time.} Those Vulnerabilities are, in turn, cause-by the presence of
-Weaknesses in the system.  Weaknesses are Errors in the design or
-implementation of a system.  Therefore, a theory on cybersecurity should first
-predict the Errors in a system design.
+(to name a few)\footnote{In formal protocol verifiers, in the DY model, the
+attacker has a set of hardcoded inference rules that defines how the attacker
+deduces new messages based on the observed message exchange.}.  As depicted in
+Figure~\ref{fig:causality}, in order to define a theory on cybersecurity we may
+say that the presence of vulnerabilities is a necessary condition in order to
+cause an Attack in the system.  Those vulnerabilities are, in turn, caused by
+the presence of weaknsesses in the system.  Weaknesses are errors in the design
+or implementation of a system.  Therefore, a theory on cybersecurity should
+first predict the errors in a system design.
 
@@
 so that we don't need to rely on a scoring system (such as the 
-CVSS\autocite{Mell2007CVSS}) to assign a quantitative evaluation 
+CVSS) to assign a quantitative evaluation 
 of the security of each entry. Instead, we want a precise calculation
@@
 and channel; considering both input-port to channel and channel to output-port.
-In fact, the Weaknesses of a channel are defined in terms of the Weaknesses of
+In fact, the weaknesses of a channel are defined in terms of the weaknesses of
 ports.  Io order to define a functional block

### 2020-07-22 reports/paper_0/main.tex
bugfix and add Volpe's comments

@@
 Section~\ref{sec:theory}. In Section~\ref{sec:secra}, we apply our theory for
-the tool-assisted security risk assessment of a CPS (with an ad-hoc example,
-based on the SWaT testbed \autocite{Mathur2016swat}).  This application shows
+the tool-assisted security risk assessment of a Cyber-Physical System (CPS), with an ad-hoc example,
+based on the SWaT testbed \autocite{Mathur2016swat}.  This application shows
 how our theory can be used to predict all of the possible security weaknesses
@@
 Figure~\ref{fig:system-agent} shows a system composed by two agents. The system
-has to ports denoted by ``I'' (with incoming assertions) and ``O'' (with outgoing assertions).
+has two ports denoted by ``I'' (with incoming assertions) and ``O'' (with outgoing assertions).
 The bottom part of Figure~\ref{fig:system-agent} details the structure of a single
@@
 \end{tabular}
-\caption{RCC3, RCC5, and RCC8 relations between Regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
+\caption{RCC3, RCC5, and RCC8 relations between regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
 \end{table*}
 
-Similarly to\autocite{Santaca2016abf}, we define agents as a meronomy (an
+Similarly to \autocite{Santaca2016abf}, we define agents as a meronomy (an
 hierarchy of Part-Whole relations) over the constituent that we previously defined
 (assertions, beliefs, and facts), based on a standard definition of mereology,
-i.e. based on the definition of Parthood relation between \emph{Parts}.  Due to
-the necessity of considering different types of Part (as we'll show afterwards)
+i.e. based on the definition of parthood relation between \emph{Parts}.  Due to
+the necessity of considering different relations between parts (as we'll show afterwards)
 we extend the mereology to a
-mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
+mereo-topology \autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
 considering the relations in Table~\ref{tab:rcc358}.  For the sake of
-readability, we use the term \emph{Region} both to refer to a mereological Part
-and to a topological Region.  Our aim is to create a meronomy (hierarchy of part-whole
+readability, we use the term \emph{region} both to refer to a mereological Part
+and to a topological region.  Our aim is to create a meronomy (hierarchy of part-whole
 relations)
@@
 such as the one provided
-in\autocite{NIST2020NVD,MITRE2020CVE} 
+in \autocite{NIST2020NVD,MITRE2020CVE} 
 so that we don't need to rely on a scoring system (such as the 
@@
 
-A mereotopology, as defined e.g. in\autocite{Rachavelpula2017mereotopology}, is
-an ordered mathematical structure where the basic relation between Regions is
-the reflexive, antisymmetric, and transitive \emph{Parthood} relation $\subseteq$.  
-The Parthood relation orders a universe of agents $\agentuniverse$ by defining the so called
-\emph{Connects with} (see in Table~\ref{tab:rcc358}) relation between Regions.
-We use the
-Region Connection Calculus (RCC), as defined
-in~\autocite{bennettLogics,improvingRCC}, to provide an axiomatization of the
-mereo-topological concepts. In its broader definition, the RCC theory is composed
-by eight axioms, and is known as RCC8. In the text, for brevity, 
-we will often focus only on RCC5 without
-loss of generality. Using RCC5 instead of RCC8 prevents us from 
-considering tangential connections between spatial
-Regions. However, tangential connections in RCC8 can be considered as 
-special cases of the more general spatial relations considered in RCC5.
-
-In Table~\ref{tab:rcc}, we
-summarize the axioms of the Region Connection Calculus (see, e.g., \autocite{Grutter2008rcc}).
-We can now define a system over the mereotopology using the RCC calculus, as follows, 
-where $\rcc(X,Y)$ on two generic regions $X,Y$ represents one of the possible RCC relations
-between $X$ and $Y$.
+A mereotopology, as defined e.g. in \autocite{Rachavelpula2017mereotopology},
+is an ordered mathematical structure where the basic relation between regions
+is the reflexive and symmetric relation \emph{Connects With} $\subseteq$,
+that we use to order a universe of agents
+$\agentuniverse$ (see in Table~\ref{tab:rcc358}).  We use the Region Connection
+Calculus (RCC), as defined in~\autocite{bennettLogics,improvingRCC}, to provide
+an axiomatization of the mereo-topological concepts. In its broader definition,
+the RCC theory is composed by eight axioms, and is known as RCC8. In the text,
+for brevity, we will often focus only on RCC5 without loss of generality. Using
+RCC5 instead of RCC8 prevents us from considering tangential connections
+between spatial regions. However, tangential connections in RCC8 can be
+considered as special cases of the more general spatial relations considered in
+RCC5.
+
+In Table~\ref{tab:rcc}, we summarize the axioms of the RCC (see, e.g., \autocite{Grutter2008rcc}).  We can now define a system
+over the mereotopology using the RCC calculus, as follows, where $\rcc(X,Y)$ on
+two generic regions $X,Y$ represents one of the possible RCC relations between
+$X$ and $Y$.
 
@@
 the relation between facts, and
-assertions and beliefs (as inputs and outputs of the Behavior of an agent)
+assertions and beliefs (as inputs and outputs of the behavior of an agent)
 is illustrated. assertions and beliefs generated by the design of a system
@@
 We categorize agents by first
-analyzing the relations between each pair of Regions defining an agent (i.e.
+analyzing the relations between each pair of regions defining an agent (i.e.
 $\abf$), and then we categorize the different agents as tuple of the three
-Regions.  For the sake of simplicity, soundness is opposed to non-soundness in
+regions.  For the sake of simplicity, soundness is opposed to non-soundness in
 the following, however, with the RCC one should consider different ``degrees''
 of non-soundness\fixnote{mv}{In particolare, spiegherei meglio Fig. 6 e il discorso della soundness}. 
-For example, in RCC5, if we consider EQ between two Regions
-as representing soundness, DR over the same Regions represents non-soundness;
+For example, in RCC5, if we consider EQ between two regions
+as representing soundness, DR over the same regions represents non-soundness;
 while PP, PO, PPi represents the different degrees of non-soundness.  A similar
@@
 	where the maximum risk is defined by the $DR$ relation between the
-	three groups of Regions, and the minimum risk by the $EQ$ relation over
-	the same Regions. In between the two extremes, the granularity of
+	three groups of regions, and the minimum risk by the $EQ$ relation over
+	the same regions. In between the two extremes, the granularity of
 	possible intermediate configuration is defined by the calculus used
@@
 \section{Prediction of Security Weaknesses}\label{sec:theory}
-Several standards mandates a secure-by-design approach in which cybersecurity
+Several standards mandate a secure-by-design approach in which cybersecurity
 shall be considered at the very early stages of the design process.  For
@@
 	\item DO-326A -- ``Airworthiness Security Process Specification''
-		requires a cybersecurity risk assessment of the design and is
-		the ``are the only Acceptable Means of Compliance (AMC) by FAA
+		requires a cybersecurity risk assessment of the design and
+		``are the only Acceptable Means of Compliance (AMC) by FAA
 		\& EASA for aviation cyber-security airworthiness certification,
@@
 threat diagrams in the CORAS
-approach, (iii) calculate the risk as a function
+approach, (iii) calculates the risk as a function
 of impact and likelihood of threat scenarios, and (iv) finally provides
@@
 CORAS \autocite{Lund2010model}, SECRAM \autocite{De2015role})
-relies on the expertise of the person who performs the risk assessment for
+rely on the expertise of the person who performs the risk assessment for
 the identification of threats and for the quantitative estimation of risks.
 As Herley puts it \autocite{Herley2016usenixvideo}, ``we don't have a test for security and
-we don't have a metric for security'' while standards requires a security risk assessment
+we don't have a metric for security'' while standards require a security risk assessment
 which, in turn, mandates a metric to evaluate the security risk.
@@
 So far, we have reasoned on systems in the abstract, considering
-generic MAS; mapping Epistemological concepts
+generic Multi-Agent System (MAS); mapping Epistemological concepts
 to MAS. We now map MAS to CPS and show how this mapping allows us
-to predict security weaknesses in a system.
+to predict security weaknesses in a system\fixnote{mv}{Se capisco bene, questa intro ha di nuovo un valore di motivation, in particolare mostrando i problemi degli approcci correnti; mi aspetterei un paio di frasi in più alla fine per spiegare perché il vostro approccio risolva il problema}.
 
@@
 \begin{itemize}
-	\item We consider a System as an hierarchy of agents. So, we map agents
+	\item We consider a System as a hierarchy of agents. So, we map agents
 		to \emph{systems, sub-systems, or devices}, depending on the
@@
 	\item Agents reason over beliefs (i.e. transform beliefs into other
-		beliefs) and for each \emph{component of a CPS} (system, sub-system,
+		beliefs) and each \emph{component of a CPS} (system, sub-system,
 		or device) is composed by a \emph{functional architecture} that
@@
 an \emph{infinite regress} in which a port needs a port to transfer information
-between the outside of the port to the inside of itself.
+between the outside of the port to the inside of itself \fixnote{mv}{non chiaro, riformulare}.
 
@@
 	the inside (input-port) or vice versa (output-port).  There exist two
-	types of ports with the following behavior: an input-port changes
+	types of ports with the following behavior: an input-port 
 	transforms assertions from a sender $s$ ($\rassert{s}{a}$) to beliefs
-	of an agent $a$ ($\beliefRegion_a$), while an output-port from belief
-	to assertion.
+	of an agent $a$ ($\beliefRegion_a$), while an output-port transforms belief
+	into assertion.
 \end{definition}
@@
 assertions received or sent and the belief, i.e.
-$\Rcc{\assertionRegion_s}{\beliefRegion_a}$ for the input-port or
-$\Rcc{\assertionRegion_r}{\beliefRegion_a}$. A port is, in fact, syntactic sugar to express the 
+$\Rcc{\rassert{s}{a}}{\beliefRegion_a}$ for the input-port or
+$\Rcc{\rassert{r}{*}}{\beliefRegion_a}$ for the output-port. 
+A port is, in fact, syntactic sugar to express the 
 relation between assertions and beliefs. We only define secure input port
@@
 
-\begin{definition}{\bf Secure Input Port --}\label{def:secport}
-	A secure input port always allow information as
-	incoming assertions to flow from a sender $s$ to the behavior of the
-	recipient $a$ (agent).  
+\begin{definition}{\bf Secure Input Port \fixnote{mv}{chiarirei meglio intuitivamente, magari fuori della definizione, quale sia la differenza tra port e secure port.
+La differenza è in quell'always?}--}\label{def:secport}
+	A secure input port always allows information as incoming assertions to
+	flow from a sender $s$ to the beliefs (generated by the behavior) of
+	the recipient $a$ (agent).  
 	%If $p_I$ is a secure input port,
@@
 
+\paragraph{Port Weaknesses}
 \begin{theorem}{Port Weaknesses.}
@@
 \begin{enumerate}[start=1, label={W\arabic*)}]
-	\item \emph{replace port}, where assertions reaches
-		the port but do not pass the boundary of the agent (i.e. do not
-		become belief of the agent)
-	\item \emph{drop port}, where assertions reaches the
+	\item \emph{replace port}, where assertions reach
+		the port but is replaced with different and un-related information
+		before passing the boundary 
+	\item \emph{drop port}, where assertions reach the
 		port but do not pass the boundary of the agent (i.e. do not
@@
 	should result in a weakness (related to an insecurity flaw) of that
-	input port.  Using RCC5, there exist exactly other $4$ different type
+	input port.  Using RCC5, there exist exactly other $4$ different types
 	of relations, one of which is the discrete-from (DR) relation, i.e.
 	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
-	the DR relations, they have no subregion in common. Lets
-	define a function weight $|X|$ such as, for any region $X$, it
+	the DR relations, they have no subregion in common. Let us
+	define a function weight $|X|$ such that, for any region $X$, it
 	represents the smallest possible cardinality of a (mereo)topological base for
@@
 	every open region can be written as union of elements of that base. 
-	We distinguish between Regions that contain information and Regions that
+	We distinguish between regions that contain information and regions that
 	don't by writing the latter as $\emptyset$.
@@
 we consider channels to be bottom-channels,
-defined as agents with the pre-defined behavior (i.e. defined in a axiomatic
+defined as agents with the pre-defined behavior (i.e. defined in an axiomatic
 way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
@@
 \begin{definition}{\bf Secure Mono-directional Channel (bottom-channel) --}\label{def:monochannel}
-	A mono-directional channel between the two agents ($s \rightarrow r$) is defined as 
-	an agent which behavior is (dogmatically) defined as: to forward any assertion received from $s$ over an input-port, to  
-	the output-port where $r$ is listening to.
+	A mono-directional channel between two agents ($s \rightarrow r$) is
+	defined as an agent whose behavior is (dogmatically) defined as: to
+	forward any assertion received from $s$ over an input-port, to  the
+	output-port where $r$ is listening to.
 \end{definition}
@@
 \paragraph{Channel Weaknesses}
-Given that a mono-directional bottom-channel is assumed to be perfectly
+Given that a mono-directional bottom-channel is assumed\fixnote{mv}{C'è un motivo per cui fate l'assunzione che il canale sia sicuro mentre non lo facevate per le port?} to be perfectly
 forwarding any assertion from its input-port to its output-port, there is no
 insecure behavior but only the combination of the weaknesses of the input and
-output port; therefore there exists $(7^2)-2=47$ theoretical configurations
+output port; therefore there exist $(7^2)-2=47$ theoretical configurations
 ($7^2$ because there are $6$ insecure types of port, plus $1$ secure type, on
@@
 $\Rcc{\beliefRegion_1}{\assertionRegion_2}$ and
-$\Rcc{\assertionRegion_2}{\beliefRegion_3}$).  The same result can be obtained
+$\Rcc{\assertionRegion_2}{\beliefRegion_3}$)\fixnote{mv}{Come dicevo in un'altra nota, userei sempre la freccia nei pedici delle asserzioni (oppure mai, ma insomma in maniera uniforme).
+Inoltre l'ordine degli argomenti della relazione rcc è rilevante?
+O rcc è simmetrica?
+In ogni caso, da qualche parte lo direi}.
+The same result can be obtained
 by analyzing the relation between the outputs of a functional block and the
 inputs of another functional block ($\Rcc{\beliefRegion_o}{\beliefRegion_i}$),
-where functional blocks are constituents of
-the functional architecture as described afterwards.  As depicted in
-Figure~\ref{fig:atom}, so far we have considered information generated by a
-port $P_I$ and then sent through a channel $C$ to another (recipient) port
-$P_O$. In this scenario where ports and then channels are atomic (otherwise
-raising infinite regress), we can only  consider the relations between ports
-and channel; considering both input-port to channel and channel to output-port.
-In fact, the weaknesses of a channel are defined in terms of the weaknesses of
-ports.  Io order to define a functional block
-without encountering a infinitely recursive definition, we must reach the same
-conclusions as for the channel. Therefore, describing the information as
-flowing over a channel or in a functional block is purely syntactic sugar.
+where functional blocks are constituents of the functional architecture as
+described afterwards\fixnote{mv}{Questo è difficile da seguire.
+Non si può spostare a dopo aver definito i functional blocks e il modo in cui li interpretate nel vostro MAS?}.
+As depicted in Figure~\ref{fig:atom}, so far we have
+considered information generated by a port $P_I$ and then sent through a
+channel $C$ to another (recipient) port $P_O$. In this scenario where ports and
+then channels are atomic (otherwise raising infinite regress), we can only
+consider the relations between ports and channel; considering both input-port
+to channel and channel to output-port.  In fact, the weaknesses of a channel
+are defined in terms of the weaknesses of ports.  Io order to define a
+functional block without encountering an infinitely recursive definition, we
+must reach the same conclusions as for the channel. Therefore, describing the
+information as flowing over a channel or in a functional block is purely
+syntactic sugar.
 
@@
 We call this the RIDI-Hypothesis, being the
-four main categories of weakness: Replace, Insertion, Drop, Injection. We can,
+four main categories of weaknesses: Replace, Insertion, Drop, Injection. We can,
 then, deduce the following security properties to mitigate security weaknesses
@@
 	A functional block of an agent 
-	takes a region of input-beliefs and outputs a region of output-beliefs. 
+	takes a region \fixnote{mv}{Perché qui specificate region mentre nelle definizioni di port e channel non lo facevate?} of input-beliefs and outputs a region of output-beliefs. 
 	A functional architecture is an interconnected system of functional blocks.
@@
 \begin{enumerate}[start=50, label={W\arabic*)}]
-	\item $\po{\beliefRegion}{\factRegion}$ the component has a Byzantine behavior where occasionally outputs the expected
+	\item $\po{\beliefRegion}{\factRegion}$\fixnote{mv}{Dalla def.
+sopra un functional block ha un input e un output belief.
+Capisco che ora entrino in gioco i fact ma non capisco perché qui nelle weakness consideriate solo una belief (ed è quella di input o di output?).
+Oppure qui B sta per behavior?
+In ogni caso, chiarirei} the component has a Byzantine behavior where occasionally outputs the expected
 	        output given the correct inputs. Not all the inputs are handled
@@
 During the specification phase, for any agent, channel, port, functional block
-and architecture, there may exist a factorial requirement (fact) predicating over them.
+and architecture, there may exist a factorial\fixnote{mv}{chiarire} requirement (fact) predicating over them.
 In other words, any requirement is defined as a fact since they must be true in
 any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
-Figure~\ref{fig:soundness}, facts are strategic rules that defines how
+Figure~\ref{fig:soundness}, facts are strategic rules that define how
 the system shall behave (by specification), while reality may be shown to
@@
 	\item $\eq{\beliefRegion}{\factRegion}=\eq{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, 
-		where $\beliefRegion_3$ represents the Region of all the
-		outputs of the functional block $sum$ while the Region
+		\fixnote{mv}{Is this the sum of two regions?
+How is it defined?
+If it's just the name of the region corresponding to the facts, I would use a different name: I find this confusing} where $\beliefRegion_3$ represents the region of all the
+		outputs\fixnote{mv}{Non mi è chiarissimo cosa intendiate con ``all the outputs''} of the functional block $sum$ while the region
 		$\beliefRegion_1+\beliefRegion_2$ represents the expected
@@
 		the output is incorrect.
-	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exists some outputs that are not the result of the summation of the inputs but given the correct inputs the function always outputs correctly.
+	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exist some outputs that are not the result of the summation of the inputs but given the correct inputs the function always outputs correctly.
 	\item
@@
 \subsection{Security and Insecurity of a System}
+\fixnote{mv}{Questo paragrafo mi fa un po' paura :) Non capisco del tutto da dove venga la formula.
+Per ogni agente: in r(n,k) ho tutte le possibili configurazioni e da queste poi sottraggo quelle che non si possono verificare?
+Ma tra tutte le possibili non ce ne sono anche di buone?
+Ad esempio quelle con EQ non sono in genere buone?
+Il risultato è compreso tra 0 e 1 o può essere anche negativo?}
 \begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
@@
 The solver identifies all possible configurations of the system and, in turn,
-to identify all potential weaknesses. 
+identifies all potential weaknesses. 
 The internal structures of the tool
 can be viewed as PDF and the results are reported into an Excel file.
-The Excel file also report the total number of configurations as indicating
+The Excel file also reports the total number of configurations as indicating
 the security risk associated to the specification. A user can change the status
-of each weaknesses in the Excel file, from the initial (default) status (open)
+of each weakness in the Excel file, from the initial (default) status (open)
 to ``mitigated''. As a consequence, the risk is re-calculated on-the-fly, i.e.
@@
 \section{Conclusion and Future Works}
+\fixnote{mv}{Se ho capito bene, detto a parole mie, quello che il vostro
+approccio permette di fare, dati requisiti e architettura del sistema, è di
+elencare in maniera sistematica tutti i "punti" in cui potrebbero sorgere dei
+problemi. Spiegherei un po' meglio, qui nelle conclusioni o nella sezione
+precedente, come questo approccio dovrebbe essere usato nella pratica... Voglio
+dire che non mi sembra necessariamente in opposizione ad altri approcci ma
+magari può integrarsi a questi.  Inoltre quando parlate dei passi successivi,
+ad esempio il considerare la generazione di test cases, se avete già qualche
+idea su come procedere, potreste discuterne brevemente.}
 We proposed a foundational theory on security, arguing that
@@
 the vagueness of the security controls on the engineering processes.
-The current state-of-the-art processes allows, given a specification,
+The current state-of-the-art processes allow, given a specification,
 an engineer to design the system in such a way that security issues arise due to the lack
@@
 many possible design and, in turn, implementation given a specification.
-The problem is correlated to the epistemological search for truth,
+The problem is related to the epistemological search for truth,
 where the challenge is to relate information and belief to human knowledge.

### 2020-07-22 reports/paper_0/main.tex
rewrite of insecurity formula and fix Volpe's feedback

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
 \addbibresource{../literature.bib}% Syntax for version >= 1.2
@@
 We now define those concepts abstractly, inspired by the
-$\abf$-framework  defined in \autocite{Santaca2016abf} and, in
+$\abftheory$-framework  defined in \autocite{Santaca2016abf} and, in
 Section~\ref{sec:theory} we detail them as engineering concepts.
@@
 two generic regions $X,Y$ represents one of the possible RCC relations between
-$X$ and $Y$.
+$X$ and $Y$. We note that all the RCC relations are symmetric with the exception
+of those that has an explicit (related) inverse.
 
@@
 
-In Figure~\ref{fig:soundness},
-the relation between facts, and
-assertions and beliefs (as inputs and outputs of the behavior of an agent)
-is illustrated. assertions and beliefs generated by the design of a system
-may not be exactly aligned with what the facts mandate (i.e. what the specification mandates.
-The relation between facts, and assertions and beliefs 
-can be used as a metric to determine the soundess of the 
-design with respect to the specification.
-
-We categorize agents by first
-analyzing the relations between each pair of regions defining an agent (i.e.
-$\abf$), and then we categorize the different agents as tuple of the three
-regions.  For the sake of simplicity, soundness is opposed to non-soundness in
-the following, however, with the RCC one should consider different ``degrees''
-of non-soundness\fixnote{mv}{In particolare, spiegherei meglio Fig. 6 e il discorso della soundness}. 
-For example, in RCC5, if we consider EQ between two regions
-as representing soundness, DR over the same regions represents non-soundness;
-while PP, PO, PPi represents the different degrees of non-soundness.  A similar
-argument can be done for completeness.  
+In Figure~\ref{fig:soundness}, the relation between facts, and assertions and
+beliefs (as inputs and outputs of the behavior of an agent) is illustrated.
+Assertions and beliefs generated by the design of a system may not be exactly
+aligned with what the facts mandate (i.e. what the specification mandates).
+The relation between facts, and assertions and beliefs can be used as a metric
+to determine the soundness of the design with respect to the specification.
+
+We categorize agents by first analyzing the relations between each pair of
+regions defining an agent (i.e.  $\abf$), and then we categorize the different
+agents as tuple of the three regions.  For the sake of simplicity, soundness is
+opposed to non-soundness in the following, however, with the RCC one should
+consider different ``degrees'' of non-soundness.  For example, in RCC5, if we
+consider EQ between two regions as representing soundness, DR over the same
+regions represents non-soundness; while PP, PO, PPi represents the different
+degrees of non-soundness.  A similar argument can be done for completeness.  
 \begin{figure}[t]
@@
 We are now in the position to define what a secure system is (with respect to
-the $\abf$-framework) and, based on that definition, what the security risk is and
+the $\abftheory$-framework) and, based on that definition, what the security risk is and
 how to quantify it in a risk matrix.
@@
 A risk is the whole space of potential designs of a specification with respect to the
-	$\abf$-framework.
+	$\abftheory$-framework.
 \end{definition}
@@
 can be summarized as to provide an understanding of the potential security risks.
-Roughly speaking, a security risk assessment (e.g. CORAS\autocite{Lund2010model}) 
+Roughly speaking, a security risk assessment (e.g. CORAS \autocite{Lund2010model}) 
 methodology starts from the (i) identification of assets, (ii) determines
@@
 which, in turn, mandates a metric to evaluate the security risk. 
+In contrast, in this section, we define how to specify a CPS in the $\abftheory$-framework
+and we identify a security metric.
 
@@
 to MAS. We now map MAS to CPS and show how this mapping allows us
-to predict security weaknesses in a system\fixnote{mv}{Se capisco bene, questa intro ha di nuovo un valore di motivation, in particolare mostrando i problemi degli approcci correnti; mi aspetterei un paio di frasi in più alla fine per spiegare perché il vostro approccio risolva il problema}.
+to predict security weaknesses in a system.
 
@@
 \end{figure}
-Since the $\abf$-framework is a theory of agents, we could consider ports as
+Since the $\abftheory$-framework is a theory of agents, we could consider ports as
 agents that allows the exchange of information between a channel and another
-agent.  However, we considered a port as a special type of agent that to avoid
-an \emph{infinite regress} in which a port needs a port to transfer information
-between the outside of the port to the inside of itself \fixnote{mv}{non chiaro, riformulare}.
+agent.  However, we considered a port as a special type of agent to avoid
+an \emph{infinite regress}. While a channel transfers information between agents and
+a functional architecture elaborate information, a port is simply a connector between
+a channel and a functional architecture. One, however, may argue that a similar
+connect is needed between a channel and the port itself. While this is not 
+excluded by the $\abftheory$-framework, would obviously lead to an infinite
+nested structure of ports. In order to avoid this, we assume
+that a port doesn't require any other means to transfer information from/to a channel
+or from/to a functional architecture.
 
@@
 $\Rcc{\rassert{s}{a}}{\beliefRegion_a}$ for the input-port or
-$\Rcc{\rassert{r}{*}}{\beliefRegion_a}$ for the output-port. 
-A port is, in fact, syntactic sugar to express the 
-relation between assertions and beliefs. We only define secure input port
-but the definition of a secure output port is symmetrical.
-
-\begin{definition}{\bf Secure Input Port \fixnote{mv}{chiarirei meglio intuitivamente, magari fuori della definizione, quale sia la differenza tra port e secure port.
+$\Rcc{\rassert{r}{*}}{\beliefRegion_a}$ for the output-port.  A port is, in
+fact, syntactic sugar to express the relation between assertions and beliefs.
+We only define secure input port but the definition of a secure output port is
+symmetrical.  Moreover, we note that the definition of an input/output port can
+be considered ``secure'', meaning that we implicitly formulated the requirement
+that a port always forwards the information without modifying it in any way.
+This is assumed because we define a port as if the RCC equality relation holds
+between the input/output assertions and the input/output beliefs.  In contrast,
+assuming any RCC relation, between inputs and outputs of a port, which is not
+an equality can be considered as generating a weaknesses, as we describe in
+more details in the next paragraph.
+
+%\begin{definition}{\bf Secure Input Port \fixnote{mv}{chiarirei meglio intuitivamente, magari fuori della definizione, quale sia la differenza tra port e secure port.
 La differenza è in quell'always?}--}\label{def:secport}
-	A secure input port always allows information as incoming assertions to
-	flow from a sender $s$ to the beliefs (generated by the behavior) of
-	the recipient $a$ (agent).  
-	%If $p_I$ is a secure input port,
-	%$\world\models\varphi$, $\varphi\in\rassert{s}{a}$, and $a$ has the
-	%only input port $p_I$, $\world'\models\beliefRegion_a$ and
-	%$\varphi\in\beliefRegion_a$ for any $\world\modalrelation\world'$.
-\end{definition}
+%	A secure input port always allows information as incoming assertions to
+%	flow from a sender $s$ to the beliefs (generated by the behavior) of
+%	the recipient $a$ (agent).  
+%	%If $p_I$ is a secure input port,
+%	%$\world\models\varphi$, $\varphi\in\rassert{s}{a}$, and $a$ has the
+%	%only input port $p_I$, $\world'\models\beliefRegion_a$ and
+%	%$\varphi\in\beliefRegion_a$ for any $\world\modalrelation\world'$.
+%\end{definition}
 
@@
 \begin{proof}
-An input port is, in the $\abf$-framework, defined secure as long as the relation
+An input port is, in the $\abftheory$-framework, defined secure as long as the relation
 	between the two regions of input assertions $\assertionRegion$ and
@@
 mono-directional channel (channel from now on) and an agent, as we did for the
-ports, since the $\abf$-framework is a theory of agents.  In fact, if a channel
+ports, since the $\abftheory$-framework is a theory of agents.  In fact, if a channel
 were considered an agent (channel-agent) then the question would be how an
@@
 
-\begin{definition}{\bf Secure Mono-directional Channel (bottom-channel) --}\label{def:monochannel}
+\begin{definition}{\bf Mono-directional Channel (bottom-channel) --}\label{def:monochannel}
 	A mono-directional channel between two agents ($s \rightarrow r$) is
@@
 \paragraph{Channel Weaknesses}
-Given that a mono-directional bottom-channel is assumed\fixnote{mv}{C'è un motivo per cui fate l'assunzione che il canale sia sicuro mentre non lo facevate per le port?} to be perfectly
-forwarding any assertion from its input-port to its output-port, there is no
-insecure behavior but only the combination of the weaknesses of the input and
-output port; therefore there exist $(7^2)-2=47$ theoretical configurations
-($7^2$ because there are $6$ insecure types of port, plus $1$ secure type, on
-both input and output side; and we exclude the configuration with $2$ secure
-types as input and output, $-2$); where only 43 are possible. For the sake of
-readability, we report 6 examples in the following but the proof by exhaustion
-over all the possible cases is straightforward.
+Given that a mono-directional bottom-channel is assumed to be perfectly
+forwarding any assertion (as we assumed for ports) from its input-port to its
+output-port, there is no insecure behavior but only the combination of the
+weaknesses of the input and output port; therefore there exist $(7^2)-2=47$
+theoretical configurations ($7^2$ because there are $6$ insecure types of port,
+plus $1$ secure type, on both input and output side; and we exclude the
+configuration with $2$ secure types as input and output, $-2$); where only 43
+are possible. For the sake of readability, we report 6 examples in the
+following but the proof by exhaustion over all the possible cases is
+straightforward.
 
@@
 It is important to note that all the results of the application of the
-$\abf$-framework to channels (the analysis of the relation
-$\Rcc{\assertionRegion_o}{\assertionRegion_i}$) lead to the same results of the
-analysis of a pair of an input and output port (i.e.
-$\Rcc{\beliefRegion_1}{\assertionRegion_2}$ and
-$\Rcc{\assertionRegion_2}{\beliefRegion_3}$)\fixnote{mv}{Come dicevo in un'altra nota, userei sempre la freccia nei pedici delle asserzioni (oppure mai, ma insomma in maniera uniforme).
-Inoltre l'ordine degli argomenti della relazione rcc è rilevante?
-O rcc è simmetrica?
-In ogni caso, da qualche parte lo direi}.
+$\abftheory$-framework to channels (the analysis of the RCC relations
+between output and input assertions of an agent)
+%$\Rcc{\assertionRegion_o}{\assertionRegion_i}$) 
+lead to the same results of the
+analysis of a pair of an input and output port.
+%(i.e.
+%$\Rcc{\beliefRegion_1}{\assertionRegion_2}$ and
+%$\Rcc{\assertionRegion_2}{\beliefRegion_3}$)
 The same result can be obtained
 by analyzing the relation between the outputs of a functional block and the
-inputs of another functional block ($\Rcc{\beliefRegion_o}{\beliefRegion_i}$),
+inputs of another functional block, %($\Rcc{\beliefRegion_o}{\beliefRegion_i}$),
 where functional blocks are constituents of the functional architecture as
-described afterwards\fixnote{mv}{Questo è difficile da seguire.
-Non si può spostare a dopo aver definito i functional blocks e il modo in cui li interpretate nel vostro MAS?}.
+described afterwards\fixnote{mv}{Questo è difficile da seguire. Non si può spostare a dopo aver definito i functional blocks e il modo in cui li interpretate nel vostro MAS?}.  
 As depicted in Figure~\ref{fig:atom}, so far we have
 considered information generated by a port $P_I$ and then sent through a
-channel $C$ to another (recipient) port $P_O$. In this scenario where ports and
-then channels are atomic (otherwise raising infinite regress), we can only
+channel $C$ to another (recipient) port $P_O$. In this scenario, where ports and
+channels are atomic (otherwise raising infinite regress), we can only
 consider the relations between ports and channel; considering both input-port
@@
 behavior (while ports and channels have a pre-defined behavior).
+In the following, for the sake of simplicity, 
+we use the generic region $\beliefRegion$ to refer to the behavior (i.e.
+the beliefs generated by the behavior).
 
 \begin{enumerate}[start=50, label={W\arabic*)}]
-	\item $\po{\beliefRegion}{\factRegion}$\fixnote{mv}{Dalla def.
-sopra un functional block ha un input e un output belief.
-Capisco che ora entrino in gioco i fact ma non capisco perché qui nelle weakness consideriate solo una belief (ed è quella di input o di output?).
-Oppure qui B sta per behavior?
-In ogni caso, chiarirei} the component has a Byzantine behavior where occasionally outputs the expected
-	        output given the correct inputs. Not all the inputs are handled
-	        properly, nor all the expected outputs are always generated
-	        when correct inputs are given.
+	\item $\po{\beliefRegion}{\factRegion}$ the component has a Byzantine
+		behavior where occasionally outputs the expected output given
+		the correct inputs. Not all the inputs are handled properly,
+		nor all the expected outputs are always generated when correct
+		inputs are given.
 	\item $\pp{\beliefRegion}{\factRegion}$ part of the expected outputs are not
@@
 During the specification phase, for any agent, channel, port, functional block
-and architecture, there may exist a factorial\fixnote{mv}{chiarire} requirement (fact) predicating over them.
+and architecture, there may exist a requirement (fact) predicating over them.
 In other words, any requirement is defined as a fact since they must be true in
@@
 the system shall behave (by specification), while reality may be shown to
-be insecure.
+be insecure (i.e. diverging from the expected behavior).
 
@@
 inputs, as in Figure~\ref{fig:sum}, defined by the requirement $r := b_3 = b_1
-+ b_2$. The possible relations between the behavior of the functional block and
++ b_2$ for any $b_1$ and $b_2$. The possible relations between the beliefs generated by the behavior of the functional block and
 the requirements (i.e. $\Rcc{\beliefRegion}{\factRegion}$ is determined by
@@
 \begin{itemize}
-	\item $\eq{\beliefRegion}{\factRegion}=\eq{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, 
-		\fixnote{mv}{Is this the sum of two regions?
-How is it defined?
-If it's just the name of the region corresponding to the facts, I would use a different name: I find this confusing} where $\beliefRegion_3$ represents the region of all the
-		outputs\fixnote{mv}{Non mi è chiarissimo cosa intendiate con ``all the outputs''} of the functional block $sum$ while the region
-		$\beliefRegion_1+\beliefRegion_2$ represents the expected
+	\item $\eq{\beliefRegion}{\factRegion}=\eq{\beliefRegion_3}{\beliefRegion_{1+2}}$, 
+		where $\beliefRegion_3$ represents the region of the
+		outputs of the functional block $sum$ while the region
+		$\beliefRegion_{1+2}$ represents the expected
 		outputs of an ideal implementation of the requirement $r$.
@@
 	\item
-		$\dr{\beliefRegion}{\factRegion}=\dr{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$,
+		$\dr{\beliefRegion}{\factRegion}=\dr{\beliefRegion_3}{\beliefRegion_{1+2}}$,
 		the function block never implements the requirements.
 	\item
-		$\pp{\beliefRegion}{\factRegion}=\pp{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exist some inputs for which
+		$\pp{\beliefRegion}{\factRegion}=\pp{\beliefRegion_3}{\beliefRegion_{1+2}}$, there exist some inputs for which
 		the output is incorrect.
-	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exist some outputs that are not the result of the summation of the inputs but given the correct inputs the function always outputs correctly.
+	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_{1+2}}$, there exist some outputs that are not the result of the summation of the inputs but given the correct inputs the function always outputs correctly.
 	\item
-		$\po{\beliefRegion}{\factRegion}=\po{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$,
+		$\po{\beliefRegion}{\factRegion}=\po{\beliefRegion_3}{\beliefRegion_{1+2}}$,
 		the component has a Byzantine behavior where occasionally
@@
 \subsection{Security and Insecurity of a System}
-\fixnote{mv}{Questo paragrafo mi fa un po' paura :) Non capisco del tutto da dove venga la formula.
-Per ogni agente: in r(n,k) ho tutte le possibili configurazioni e da queste poi sottraggo quelle che non si possono verificare?
-Ma tra tutte le possibili non ce ne sono anche di buone?
-Ad esempio quelle con EQ non sono in genere buone?
-Il risultato è compreso tra 0 e 1 o può essere anche negativo?}
 \begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
@@
 
-Based on these hypothesis, security ($S$) can be expressed as the following equation.
+Based on these hypothesis, we can formulate the concepts of security and
+insecurity (in the $\abftheory$-framework) as mathematical equations.  Let us consider a CPS
+$S$, represented as a graph $G=\langle V,E\rangle$ where $V$ represents the
+set of functional blocks and ports of $S$, and $E\subseteq V\times E$ is the
+set of pairs representing the channels and connections (data flows) between
+functional blocks. We define $R\subseteq V\times F$, where $F$ is the set
+of all the requirements of $S$, and extend $G$ as $G'=\langle E'\times V'\rangle$
+with $E'= E\cup R$ and $V'=V\cup F$. Let $\pi: E'\rightarrow RCC$ (where $RCC$ is
+the set of relations in the RCC) be
+the total function associating an RCC relation to each edge in $G'$, and
+$\Pi$ be the set of all different permutations of RCC relations over $E'$
+(i.e. $\Pi=\{\langle\pi(e_0)=EQ,\ldots,\pi(e_n)=EQ\rangle,\ldots,\langle \pi(e_0)=DR,\ldots,\pi(e_n)=DR\rangle\}$). If $\sigma:\Pi\rightarrow\{0,1\}$ is an evaluation function
+such that $\sigma(p)=1$ (where $p\in\Pi$) iff the input configuration is satisfiable with respect to
+the logical theory defining the algebraic structure (mereotopology) and constraints
+of the calculus RCC (otherwise $\sigma$ returns $0$), we can define
 \begin{displaymath}
-	S=1-(\prod_{a\in Ag} r^{\binom{n}{k}}-\delta_T)
+	I=\sum_{p\in\{\Pi\setminus\pi_{EQ}\}} \sigma(p)
 \end{displaymath}
-where $r^{\binom{n}{k}}$ represents the number of possible configurations (permutations) 
-of an agent based on the number $r$ of relations (e.g. EQ, DR, PO) with arity $k$ between
-$n$ different regions in the $\abftheory$ theory. $\delta_T$ represents the 
-number of configurations which are not satisfiable with respect to the logical theory
-defining the 
-algebraic structure (topology) and constraints of the calculus (RCC). $Ag$ is the 
-set of all the agents in the system under evaluation.
-The constant 1 represents the nominal configuration of the system and then the
-only secure configuration. The product represents all the unintended, insecure, configurations.
+where $\pi_{EQ}\in\Pi$ is the output of the function $\pi$ that associates only EQ relations
+to any $e\in E'$, and $I\in\mathbb{N}$ represents all the insecurity configurations
+of the CPS $S$ where, at least, one of the RCC relations isn't EQ.
+In other words, we consider $\pi_{EQ}$ as the only secure configuration, and 
+we characterize security as follows.
+
+%the insecurity of a system 
+%($I$) can be expressed as the following equation.
+%\begin{displaymath}
+%	I=\prod_{a\in Ag} r_i^{\binom{n}{k}}-\delta_T
+%\end{displaymath}
+%where $r_i^{\binom{n}{k}}$ represents the number of possible configurations
+%(permutations) of an agent (defined as a tuple of RCC relations as in
+%Definition~\ref{def:opsystem}) in which at least one of the relation is not EQ.
+%$r_i$ represents the RCC relations with arity $k$ between $n$ different regions
+%in the $\abftheory$ theory. $\delta_T$ represents the number of configurations
+%which are not satisfiable with respect to the logical theory defining the
+%algebraic structure (topology) and constraints of the calculus (RCC). $Ag$ is
+%the set of all the agents in the system under evaluation and the product
+%represents all the unintended, insecure, configurations.  Given that there is
+%only $1$ nominal configuration of the system where all the agents are defined
+%only with EQ relations (and then the only secure configuration), which can be
+%expressed as $sad$ it follows that security can be defined as
+%\begin{displaymath}
+%	I=\prod_{a\in Ag} r_i^{\binom{n}{k}}-\delta_T
+%\end{displaymath}
+%
+%if $I=0$ the system can be considered secure, 
+%otherwise $I$ represents the number of possible insecure configurations, where
+%$I\in\mathbb{N}$.
 
@@
 \appendix
-%\section{Engineering $\abftheory$-framework}
 The Class Diagram for the Engineering of the $\abftheory$-framework is reported in

### 2020-07-23 reports/paper_0/main.tex
minor change
minor change in equation,add explanations in intro

@@
 pseudo-scientific theories by proposing that \emph{any} scientific theory can
-be considered correct as long as no empirical experiment falsify it. For
+be considered correct (or, better, acceptable) as long as no empirical experiment falsify it. For
 example, the theory of gravity, in its mathematical setting, can predict how
@@
 security. 
-
-Instead of reasoning of what makes a system secure or insecure,
+\fix{mr}{In other words, if (as Herley mentions in \autocite{herley}) 
+the problem is in the way we reason on cybersecurity, we should
+see it in the way we reply to simple cybersecurity (foundamental) questions.
+Let us consider the foundamental question ``what is cybersecurity?'',
+as in \autocite{hintikka}, we can translate the wh-question into an equivalent
+existentially quantified formula. 
+But if the formula is the following second-order formula: $\exists P. P(s)\rightarrow Secure(s)$,
+(where we ask if there exists a predicate $P$ such that, given any arbitrary 
+system $s$, when $P$ holds for $s$ then $s$ can be considered secure),
+when the answer is (for example) ``if the system is confidential, the
+system is secure'' the formula becomes tautological. In fact, $P$ and $Secure$ can 
+be considered as semantically equivalent (security is confidentiality and vice-versa).}
+Instead of starting from reasoning on what makes a system secure or insecure,
 we reason on what causes insecurities.
+\fix{mr}{We focus on insecurities only caused by the exploitation of
+cybersecurity attacks, and we assume that achieving security means preventing all those attacks 
+from being exploitable or exploited.}
 Our hypothesis is that cybersecurity attacks are only caused by 
@@
 errors (or source of errors) can be effectively used to predict the
-security of a system. With our approach, a list of weaknesses emerges
+security of a system.
+\fix{mr}{Therefore, 
+a possible answer to the initial question can be that cybersecurity is the 
+absence of attacks and the absence of errors implies cybersecurity.
+This, however, raises another question: ``which are all the errors
+in (the engineering of) a system that make it possible for a cybersecurity attack
+to happen?''. More rigorusly, which scientific theory predicts all
+possible errors that causes cybersecurity attacks? 
+}
+
+With our approach, a list of weaknesses emerges
 from the mathematical formulation of the system and predicts 4 main classes
@@
 	\includegraphics[width=.9\columnwidth]{system-agent.pdf}
-	\caption{xample of system and agent structure}
+	\caption{Example of system and agent structure}
 	\label{fig:system-agent}
@@
 \begin{displaymath}
-	I=\sum_{p\in\{\Pi\setminus\pi_{EQ}\}} \sigma(p)
+	I=\sum_{p\in\{\Pi\setminus\pi_{eq}\}} \sigma(p)
 \end{displaymath}
-where $\pi_{EQ}\in\Pi$ is the output of the function $\pi$ that associates only EQ relations
+where $\pi_{eq}\in\Pi$ is the output of the function $\pi$ that associates only EQ relations
 to any $e\in E'$, and $I\in\mathbb{N}$ represents all the insecurity configurations
 of the CPS $S$ where, at least, one of the RCC relations isn't EQ.
-In other words, we consider $\pi_{EQ}$ as the only secure configuration, and 
-we characterize security as follows.
+In other words, we consider $\pi_{eq}$ as the only secure configuration.
 

### 2020-07-23 reports/paper_0/main.tex
sec 4b minor review

@@
 Based on these hypothesis, we can formulate the concepts of security and
-insecurity (in the $\abftheory$-framework) as mathematical equations.  Let us consider a CPS
-$S$, represented as a graph $G=\langle V,E\rangle$ where $V$ represents the
-set of functional blocks and ports of $S$, and $E\subseteq V\times E$ is the
-set of pairs representing the channels and connections (data flows) between
-functional blocks. We define $R\subseteq V\times F$, where $F$ is the set
-of all the requirements of $S$, and extend $G$ as $G'=\langle E'\times V'\rangle$
-with $E'= E\cup R$ and $V'=V\cup F$. Let $\pi: E'\rightarrow RCC$ (where $RCC$ is
-the set of relations in the RCC) be
-the total function associating an RCC relation to each edge in $G'$, and
-$\Pi$ be the set of all different permutations of RCC relations over $E'$
-(i.e. $\Pi=\{\langle\pi(e_0)=EQ,\ldots,\pi(e_n)=EQ\rangle,\ldots,\langle \pi(e_0)=DR,\ldots,\pi(e_n)=DR\rangle\}$). If $\sigma:\Pi\rightarrow\{0,1\}$ is an evaluation function
-such that $\sigma(p)=1$ (where $p\in\Pi$) iff the input configuration is satisfiable with respect to
-the logical theory defining the algebraic structure (mereotopology) and constraints
-of the calculus RCC (otherwise $\sigma$ returns $0$), we can define
+insecurity (in the $\abftheory$-framework) as mathematical equations.  Let us
+consider a CPS $S$, represented as a graph $G=\langle V,E\rangle$ where $V$
+represents the set of functional blocks and ports of $S$, and $E\subseteq
+V\times E$ is the set of pairs representing the channels and connections (data
+flows) between functional blocks. We define $R\subseteq V\times F$, where $F$
+is the set of all the requirements of $S$, and extend $G$ as $G'=\langle
+E'\times V'\rangle$ with $E'= E\cup R$ and $V'=V\cup F$. Let $\pi:
+E'\rightarrow RCC$ (where $RCC$ is the set of relations in the RCC) be the
+total function associating an RCC relation to each edge in $G'$, and $\Pi$ be
+the set of all different permutations of RCC relations over $E'$ (i.e.
+$\Pi=\{\langle\pi(e_0)=EQ,\ldots,\pi(e_n)=EQ\rangle,\ldots,\langle
+\pi(e_0)=DR,\ldots,\pi(e_n)=DR\rangle\}$ where $e_i\in E'$ for $0\leq i \leq n$ and $|E'|=n$). If $\sigma:\Pi\rightarrow\{0,1\}$ is
+an evaluation function such that $\sigma(p)=1$ (where $p\in\Pi$) iff the input
+configuration is satisfiable with respect to the logical theory defining the
+algebraic structure (mereotopology) and constraints of the calculus RCC
+(otherwise $\sigma$ returns $0$), we can define
 \begin{displaymath}
@@
 
-\subsection{Security Risk Assessment}\label{sec:secra}
+\subsection{Cybersecurity Risk Assessment}\label{sec:secra}
 In order to test our theory we implemented a tool-chain (open-source under the

### 2020-07-29 reports/paper_0/main.tex

@@
 \begin{abstract}
-	The objective of this research is the development of a theory that
-	determines (all and only) the possible insecurity and security
-	configurations of any abstract system.  Our hypothesis is that errors
-	are at the bottom of the causality chain that leads to cybersecurity
-	attacks. Based on this hypothesis, we formulate a scientific
-	(falsifiable by experiments) theory that predicts all the errors that lead to
+	The objective of this research is lay the foundation for the
+	development of a theory that determines (all and only) the possible
+	insecurity and security configurations of any abstract system.  Our
+	hypothesis is that cybersecurity weaknesses (i.e. errors) are at the
+	bottom of the causality chain that leads to cybersecurity attacks.
+	Based on this hypothesis, we formulate a scientific (falsifiable by
+	experiments) hypothesis that predicts all the weaknesses that lead to
 	the presence of cybersecurity weaknesses in the architectural design of
-	an abstract system.  Our theory is based on a correlation between the
-	epistemological concepts of knowledge, beliefs, and information with
-	the engineering concepts of requirements, functional architectures, and
-	channels.  Ultimately, our theory allows the definition of cybersecurity of a system
-	as a mathematical formula.  We implemented a prototype cybersecurity
-	risk assessment tool that, based on our theory, predicts weaknesses in
-	a UML model of a (cyber-physical) system.
+	an abstract system. The mathematical formulation of our hypothesis is
+	based on a correlation between the epistemological concepts of
+	knowledge, beliefs, and information with the engineering concepts of
+	requirements, functional architectures, and channels.  Ultimately, our
+	hypothesis allows the definition of the cybersecurity of a system as a
+	mathematical formula.  We implemented a prototype cybersecurity risk
+	assessment tool that, based on our hypothesis, predicts weaknesses in a UML
+	model of a (cyber-physical) system.
 \end{abstract}
@@
 you, by the way) that theory cannot be fabricated out of the results of
-observation, but that it can only be invented.''~\autocite{Popper1959logic}},
+observation, but that it can only be invented.'' \autocite{Popper1959logic}},
 and that empirical experiments should be considered as the only evidence to
@@
 Popper's famous ``demarcation principle'' draws a line between scientific and
-pseudo-scientific theories by proposing that \emph{any} scientific theory can
-be considered correct (or, better, acceptable) as long as no empirical experiment falsify it. For
+pseudo-scientific theories by proposing that a scientific theory can
+be considered acceptable as long as no empirical experiment falsify it. For
 example, the theory of gravity, in its mathematical setting, can predict how
@@
 theory (as people often think) but a vice.'' -- Karl Popper, Conjectures and
-Refutations~\autocite{popper1962conjectures}.} to be nonscientific (i.e.
+Refutations \autocite{popper1962conjectures}.} to be nonscientific (i.e.
 unfalsifiability is a fallacy of a theory), Herley concludes that there is no
 scientific theory of cybersecurity; which means that cybersecurity lays in the
-realm of pseudo-sciences~\autocite{Herley2016usenixvideo}. Herley, e.g.  in
+realm of pseudo-sciences \autocite{Herley2016usenixvideo}. Herley, e.g.  in
 \autocite{Herley2017justifying}, discusses the implications of a nonscientific
@@
 theory.  We consider the problem raised by Herley not confined to ``computer
-security'' but rather we reason on any abstract system (so that our theory may
-hold for any sound implementation such as networks, mechanical, cyber, or
+security'' but rather we reason on any abstract system (so that our hypothesis may
+be tested in any sound implementation such as networks, mechanical, cyber, or
 cyber-physical system, or even a single computer or a single device such as a
@@
 design, before the logic of protocols is even considered.} of the engineering
-process (of systems) without considering many aspects related to the
+process (of systems) without considering many aspects related to the design or
 implementation of a protocol that may generate security vulnerabilities later
 in the development process.  An infamous example is the KRAK attack on the WPA2
-4-way handshake protocol where the author wrote~\autocite{krak2020web}: ``our
+4-way handshake protocol where the author wrote \autocite{krak2020web}: ``our
 attacks do not violate the security properties proven in formal analysis of the
@@
 security. 
-\fix{mr}{In other words, if (as Herley mentions in \autocite{herley}) 
+In other words, if (as Herley mentions in \autocite{Herley2016usenixvideo}) 
 the problem is in the way we reason on cybersecurity, we should
@@
 Let us consider the foundamental question ``what is cybersecurity?'',
-as in \autocite{hintikka}, we can translate the wh-question into an equivalent
+as in \autocite{hintikka2013inquiry}, we can translate the wh-question into an equivalent
 existentially quantified formula. 
-But if the formula is the following second-order formula: $\exists P. P(s)\rightarrow Secure(s)$,
+But if the formula is $\exists P. P(s)\rightarrow Secure(s)$,
 (where we ask if there exists a predicate $P$ such that, given any arbitrary 
@@
 system is secure'' the formula becomes tautological. In fact, $P$ and $Secure$ can 
-be considered as semantically equivalent (security is confidentiality and vice-versa).}
+be considered as semantically equivalent (security is confidentiality and vice-versa).
 Instead of starting from reasoning on what makes a system secure or insecure,
 we reason on what causes insecurities.
-\fix{mr}{We focus on insecurities only caused by the exploitation of
+We focus on insecurities only caused by the exploitation of
 cybersecurity attacks, and we assume that achieving security means preventing all those attacks 
-from being exploitable or exploited.}
+from being exploitable or exploited.
 Our hypothesis is that cybersecurity attacks are only caused by 
-the presence of errors in the design or implementation of a system.
-For example, a design or implementation error in a sanitization function
-may allow an attacker to perform SQL-injections. If errors are necessary
+the presence of errors in the design or implementation of a system (i.e. cybersecurity weaknesses).
+For example, a design or implementation weakness in a sanitization function
+may allow an attacker to perform SQL-injections. If weaknesses are necessary
 to enable cybersecurity attacks, a theory that predicts all potential 
-errors (or source of errors) can be effectively used to predict the
-security of a system.
-\fix{mr}{Therefore, 
-a possible answer to the initial question can be that cybersecurity is the 
-absence of attacks and the absence of errors implies cybersecurity.
-This, however, raises another question: ``which are all the errors
+weaknesses can be effectively used to predict the
+(in)security of a system.
+Therefore, 
+a possible answer to the initial question can be that (definition) cybersecurity is the 
+absence of attacks and (claim) the absence of weaknesses implies cybersecurity.
+This, however, raises another question: ``which are all the errors/weaknesses
 in (the engineering of) a system that make it possible for a cybersecurity attack
-to happen?''. More rigorusly, which scientific theory predicts all
-possible errors that causes cybersecurity attacks? 
-}
-
-With our approach, a list of weaknesses emerges
-from the mathematical formulation of the system and predicts 4 main classes
-of weaknesses. Those classes of weaknesses are used to calculate
-all the insecurity configurations of all the components of a system,
-obtaining a precise estimation of all potential cybersecurity-related risks
-in any given system. Our hypothesis can be falsified by means of experiments,
-testing if all the predicted weaknesses are present in the system under
-test, or testing if other (not predicted by our hypothesis) weaknesses are
-present.
+to happen?''. More rigorously, which scientific theory predicts all
+possible weaknesses that causes cybersecurity attacks?
+
+With our approach, a list of weaknesses emerges from the mathematical
+formulation of a system in a framework called $\abftheory$-framework and
+predicts 4 main classes of weaknesses. Those classes of weaknesses are used to
+calculate all the insecurity configurations of all the components of a system,
+obtaining a precise estimation of all potential cybersecurity-related risks in
+any given system. Our hypothesis can be falsified by means of experiments,
+testing if all the predicted weaknesses are present in the system under test,
+or testing if other (not predicted by our hypothesis) weaknesses are present.
+In fact, if any security weaknesses were to be found in a system and not
+predicted by our hypothesis, the hypothesis could be declared incomplete.  If a
+security weakness would be predicted by our hypothesis but found to be impossible
+to realize, our hypothesis could be declared as wrong.
 
@@
 security.  We formulate a security hypothesis in Section~\ref{sec:hypothesis};
-which we use to propose a theory on system security in
-Section~\ref{sec:theory}. In Section~\ref{sec:secra}, we apply our theory for
-the tool-assisted security risk assessment of a Cyber-Physical System (CPS), with an ad-hoc example,
-based on the SWaT testbed \autocite{Mathur2016swat}.  This application shows
-how our theory can be used to predict all of the possible security weaknesses
-of a system, allowing the falsification of our theory.  In fact, if any
-security weaknesses were to be found in a system and not predicted by our
-theory, the theory could be declared incomplete.  Similarly, if a security
-weakness would be predicted by our theory but found to be impossible to
-realize, our theory could be declared as wrong.
+which we use to propose a mathematical formulation of our hypothesis on system security in
+Section~\ref{sec:theory}. In Section~\ref{sec:secra}, we apply our hypothesis for
+the tool-assisted security risk assessment of a Cyber-Physical System (CPS), with an ad-hoc example.
+%based on the SWaT testbed \autocite{Mathur2016swat}. 
+This application shows
+how our hypothesis can be used to predict all of the possible security weaknesses
+of a system, allowing the falsification of our hypothesis.
 
@@
 a network of databases of weaknesses (e.g.
-CWE~\autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE~\autocite{CVE},
-NVD~\autocite{NIST2020NVD}), and attacks (e.g. CAPEC~\autocite{MITRE2020CAPEC},
-ATT\&CK~\autocite{MITRE2020ATTACK}).  Those insecurity issues can be related to
+CWE \autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE \autocite{CVE},
+NVD \autocite{NIST2020NVD}), and attacks (e.g. CAPEC \autocite{MITRE2020CAPEC},
+ATT\&CK \autocite{MITRE2020ATTACK}).  Those insecurity issues can be related to
 the violation of one or more requirements (explicit or implicit) in the
@@
 non-repudiation, or introducing new ones such as ``responsibility'' in the
-RITE approach (see~\autocite{Samonas2014cia} for an overview of the evolution of the CIA triad).
+RITE approach (see \autocite{Samonas2014cia} for an overview of the evolution of the CIA triad).
 The link between security requirements and vulnerabilities is reported 
-in the NVD databases by the CVSS\autocite{Mell2007CVSS}) scoring system.
+in the NVD databases by the CVSS \autocite{Mell2007CVSS} scoring system.
 The CVSS evaluates of the severity of a vulnerability
@@
 which safety requirements are defined (e.g. in standards such as the IEC 61508
-or 61511\autocite{IEC201761511}), are based on empirical tests and measurements.
+or 61511 \autocite{IEC201761511}), are based on empirical tests and measurements.
 While reasoning by induction
@@
 		animals, and other features of the world develop of their own
-		accord'' (e.g. the physical universe)~\autocite{wiki-nature}. 
+		accord'' (e.g. the physical universe) \autocite{wiki-nature}. 
 		\begin{itemize}
@@
 		definition, so that no other causes (e.g.  other sources) but
-		the ones in the definition are responsible for a vulnerability
+		the ones in the definition are responsible for a vulnerability.
 		%\footnote{``For this view, that
@@
 		Furthermore, the term ``threat sources'' used in the definition
-		in~\autocite{cnssi20104009} may be identified with both Nature
+		in \autocite{cnssi20104009} may be identified with both Nature
 		and Mankind, not differentiating between safety and security.
@@
 channel that allows the violation of the security policy
-[\ldots]''~\autocite{cnssi20104009}. For example, a software or procedure that takes
+[\ldots]'' \autocite{cnssi20104009}. For example, a software or procedure that takes
 advantage of the vulnerability causing an \emph{attack} to the system may
@@
 	\item \emph{Causality} refers to the causality principle; defined
-		in~\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
+		in \autocite{Spirkin1983Dialectical} as ``Causality is a genetic
 		connection of phenomena through which one thing (the cause)
@@
 		occur on computer software, hardware, or something electronic
-		(usually computerized).''~\autocite{wiki-exploit}.
+		(usually computerized).'' \autocite{wiki-exploit}.
 	\item An \emph{Attack}, as defined by the International Standard
@@
 		In fact, we consider the term \emph{hack} as
-		non-malicious (as, e.g. in~\autocite{Stallman2002hacker}).
-	\item A \emph{Threat}, as defined in~\autocite{cnssi20104009}, is ``Any
+		non-malicious (as, e.g. in \autocite{Stallman2002hacker}).
+	\item A \emph{Threat}, as defined in \autocite{cnssi20104009}, is ``Any
 		circumstance or event with the potential to adversely impact
@@
 	\item \emph{Defect}: ``anything that renders the product not reasonably
-		safe''~\autocite{Robinson2019writing} (i.e. a characteristic of
+		safe'' \autocite{Robinson2019writing} (i.e. a characteristic of
 		an object which hinders its proper usability).
-	\item \emph{Failure}, as defined in\autocite{Merriam2020failure}, is ``a state of
+	\item \emph{Failure}, as defined in \autocite{Merriam2020failure}, is ``a state of
 		inability to perform a normal function''. The term is
@@
 	\item \emph{Hazard}, ``a potential source of
-		harm''~\autocite{iet2017glossary}.
+		harm'' \autocite{iet2017glossary}.
 \end{itemize}
@@
 that huge repository of dishonest behaviors should be kept as definition, such
-as CAPEC~\autocite{MITRE2020CAPEC}; or that dishonesty of an agent with respect
+as CAPEC \autocite{MITRE2020CAPEC}; or that dishonesty of an agent with respect
 to a security protocol should be defined as a number of predefined actions as
@@
 say that the presence of vulnerabilities is a necessary condition in order to
-cause an Attack in the system.  Those vulnerabilities are, in turn, caused by
+cause an attack in the system.  Those vulnerabilities are, in turn, caused by
 the presence of weaknsesses in the system.  Weaknesses are errors in the design
@@
 
-\section{Security Hypothesis -- ABF Theory of Agents}\label{sec:hypothesis}
+\section{A Security Hypothesis in the $\abftheory$-framework}\label{sec:hypothesis}
 In order to address the problem raised by Herley, we shall define how to
@@
 Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent.
-However, the same ideas revolved around for thousands of years.
-Some relevant examples for our objective are the following\footnote{
+However, the same ideas revolved around for thousands of years\footnote{
 It is interesting to notice that 
@@
 }.
+Some relevant examples for our objective are the following.
 \begin{itemize}
@@
 variable (which admits the two alternatives True/False) its information is
-defined as Believed to be True/False and not Believed to be the opposite.  Due
+defined as believed to be True/False and not believed to be the opposite.  Due
 to the definition of information and then with its relation to the
@@
 		\footnote{``The logical structure of information is one of the
-		most basic and one of the most basic and one of the simplest
+		most basic and one of the simplest
 		thing in the wide and wonderful world of logical analysis. This
@@
 (assertions, beliefs, and facts), based on a standard definition of mereology,
-i.e. based on the definition of parthood relation between \emph{Parts}.  Due to
+i.e. based on the definition of parthood relation between \emph{parts}.  Due to
 the necessity of considering different relations between parts (as we'll show afterwards)
@@
 considering the relations in Table~\ref{tab:rcc358}.  For the sake of
-readability, we use the term \emph{region} both to refer to a mereological Part
+readability, we use the term \emph{region} both to refer to a mereological part
 and to a topological region.  Our aim is to create a meronomy (hierarchy of part-whole
@@
 of the number of insecurity configurations to emerge from the mathematical
-formulation of our cybersecurity theory.
+formulation of our cybersecurity hypothesis.
 
@@
 $\agentuniverse$ (see in Table~\ref{tab:rcc358}).  We use the Region Connection
-Calculus (RCC), as defined in~\autocite{bennettLogics,improvingRCC}, to provide
+Calculus (RCC), as defined in \autocite{bennettLogics,improvingRCC}, to provide
 an axiomatization of the mereo-topological concepts. In its broader definition,
@@
 
-As already presented in \autocite{Santaca2016abf} it follows that, defining
+As already presented in \autocite{Santaca2016abf}, it follows that, defining
 a system with a fixed number of regions, there exist
@@
 While a quantitative analysis reveals how many possible configurations of an
-agent (i.e. a system) exist w.r.t. the $\abf$ theory (i.e. $54/125$ in RCC5), a
+agent (i.e. a system) exist w.r.t. the $\abftheory$-framework (i.e. $54/125$ in RCC5), a
 qualitative analysis of the different configurations describe the
-configurations allowed by the $\abf$ theory, and how those configurations can
+configurations allowed by the $\abftheory$-framework, and how those configurations can
 be categorized.  In Table~\ref{tab:5com}, we provide the generic composition
@@
 input and output assertions.  Given an agent $a$ and a collection of asserted
-predicates $\Phi$, the Input assertions are those received by $a$ from an agent
+predicates $\Phi$, the input assertions are those received by $a$ from an agent
 $s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
@@
 quality.'' --  Ellis J. McTaggart, The Unreality of
-Time\autocite{Mctaggart1908unreality}}: 
+Time \autocite{Mctaggart1908unreality}}: 
 \begin{itemize}
 	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\beliefRegion$},
-		where the relation between Input-assertions and beliefs
+		where the relation between input-assertions and beliefs
 		describes the soundness of the execution of the functional
@@
 true.'' Sextus Empiricus, Outlines of Pyrrhonism,
-II-83\autocite{Empiricus1990Pyrrhonism}}.  The relation between beliefs and
+II-83 \autocite{Empiricus1990Pyrrhonism}}.  The relation between beliefs and
 facts determines the competence (on the subjects defined by the facts) of an
@@
 We note that our SE define the properties of a secure system, and correlated
-weaknesses can be found in the CWE dataset.  For example, SE-$1$ can be sees as
+weaknesses can be found in the CWE dataset.  For example, SE-$1$ can be seen as
 correlated to the weakness class of ``Improper Interaction Between Multiple
@@
 CWE--$682$. All those CWE are in the top ``view'' of the ``research concepts''
-in\autocite{MITRE2020CWEresearch}, while the other classes of weaknesses do not
-have a direct counterpart in our theory; we believe they can be sees as sub-classes,
+in \autocite{MITRE2020CWEresearch}, while the other classes of weaknesses do not
+have a direct counterpart in our hypothesis; we believe they can be sees as sub-classes,
 but a full comparison with the CWE is out of the scope of this article.
@@
 the $\abftheory$-framework) and, based on that definition, what the security risk is and
-how to quantify it in a risk matrix.
+how to quantify it in a risk matrix. The following definition holds for 
+abstract systems defined in the $\abftheory$-framework but will be refined
+for CPS afterwards in the paper.
 \begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
@@
 \begin{definition}{\bf Risk --}
-A risk is the whole space of potential designs of a specification with respect to the
+A risk is determined by the uncertainty related to the whole space of potential designs of a specification with respect to the
 	$\abftheory$-framework.
@@
 if there's a connection between the weakness and the asset that
-an attacker wants to impact. Therefore, a risk matrix should
+an attacker wants to impact. When the whole system is considered an asset,
+all weaknesses are equally likely to be exploited.
+Therefore, a risk matrix should
 capture the number of insecure configurations of a system, rather
@@
 a channel and a functional architecture. One, however, may argue that a similar
-connect is needed between a channel and the port itself. While this is not 
-excluded by the $\abftheory$-framework, would obviously lead to an infinite
+connection is needed between a channel and the port itself. While this is not 
+excluded by the $\abftheory$-framework, it would obviously lead to an infinite
 nested structure of ports. In order to avoid this, we assume
@@
 fact, syntactic sugar to express the relation between assertions and beliefs.
-We only define secure input port but the definition of a secure output port is
-symmetrical.  Moreover, we note that the definition of an input/output port can
+We note that the definition of an input/output port can
 be considered ``secure'', meaning that we implicitly formulated the requirement
@@
 
-\paragraph{Port Weaknesses}
+\paragraph{Port Weaknesses} We now apply the $\abftheory$-framework to
+list all possible cybersecurity weaknesses of a port.
 \begin{theorem}{Port Weaknesses.}
 From our definition of ports, it follows that there exist only the following \emph{six}
-types of weaknesses, generating six types of insecure port in RCC5 (the
-notation is reported for input-ports, i.e. on the left-hand side of the arrow):
+types of weaknesses, generating six types of insecure port in RCC5:
+%(the notation is reported for input-ports, i.e. on the left-hand side of the arrow):
 \begin{enumerate}[start=1, label={W\arabic*)}]
@@
 		become belief of the agent)
-	\item \emph{insertion port}, where some new
-		information is believed by $a$ as incoming from the port but
-		$s$ didn't send it
-	\item \emph{injection port}, where information
-		coming from $s$ is substituted with new information which
-		becomes believed by $a$
+	\item \emph{insertion port}, where some new information is transferred along with 
+		the information incoming from a channel, and then sent to
+		the intended recipient (agent)
+	\item \emph{injection port}, where part of the incoming information
+		is substituted with new information and transferred to the intended recipient
 	\item \emph{selective port}, where some information passes the port and
@@
 	\begin{enumerate}[start=1, label={W5.\arabic*)}]
-		\item drop, 
-		\item drop+insert.
+		\item dropped, 
+		\item replaced.
 	\end{enumerate}
@@
 			$\beliefRegion$ contains and extend $\assertionRegion$
-			which we call \emph{injection}
+			which we call \emph{insertion}
 		\item if $\ppi{\assertionRegion}{\beliefRegion})$ then
 			$\assertionRegion$ contains and extend $\beliefRegion$
-			which we call \emph{selective drop}
+			which we call \emph{drop} (or selective drop to stress the difference with the full drop)
 		\item if $\po{\assertionRegion}{\beliefRegion})$ then a part of
@@
 			$\beliefRegion$ which is a combination of
-			\emph{selective drop and generation}
+			\emph{selective drop and generation} which we call \emph{injection}.
 	\end{enumerate}
@@
 mono-directional channel (channel from now on) and an agent, as we did for the
-ports, since the $\abftheory$-framework is a theory of agents.  In fact, if a channel
+ports, since the $\abftheory$-framework is a logical theory of agents.  In fact, if a channel
 were considered an agent (channel-agent) then the question would be how an
@@
 %$\Rcc{\assertionRegion_2}{\beliefRegion_3}$)
-The same result can be obtained
-by analyzing the relation between the outputs of a functional block and the
-inputs of another functional block, %($\Rcc{\beliefRegion_o}{\beliefRegion_i}$),
-where functional blocks are constituents of the functional architecture as
-described afterwards\fixnote{mv}{Questo è difficile da seguire. Non si può spostare a dopo aver definito i functional blocks e il modo in cui li interpretate nel vostro MAS?}.  
 As depicted in Figure~\ref{fig:atom}, so far we have
@@
 to channel and channel to output-port.  In fact, the weaknesses of a channel
-are defined in terms of the weaknesses of ports.  Io order to define a
-functional block without encountering an infinitely recursive definition, we
-must reach the same conclusions as for the channel. Therefore, describing the
-information as flowing over a channel or in a functional block is purely
-syntactic sugar.
+are defined in terms of the weaknesses of ports.  
+The same result can be obtained by analyzing the relation between the outputs
+of a functional block and the inputs of another functional block, where
+functional blocks are constituents of the functional architecture as described
+afterwards.  In order to define a functional block without encountering an
+infinitely recursive definition, we must reach the same conclusions as for the
+channel. Therefore, describing the information as flowing over a channel or in
+a functional block is purely syntactic sugar.
 
 We can summarize these results by saying that the relations between assertions
-and beliefs, Output assertions of an agent and Input assertions of another
-agent, or Output beliefs of a functional block and Input beliefs of another
+and beliefs, output assertions of an agent and input assertions of another
+agent, or output beliefs of a functional block and input beliefs of another
 block can \emph{only} be affected by the following weaknesses: replace, drop,
@@
 \begin{definition}{\bf Functional Block and Architecture --}\label{def:funblock}
-	A functional block of an agent 
-	takes a region \fixnote{mv}{Perché qui specificate region mentre nelle definizioni di port e channel non lo facevate?} of input-beliefs and outputs a region of output-beliefs. 
-	A functional architecture is an interconnected system of functional blocks.
+	A functional block of an agent takes beliefs as  inputs (input-beliefs) and
+	returns output-beliefs.  A functional architecture is an
+	interconnected system of functional blocks.
 \end{definition}
-It is evident that the quality of a functional block cannot be determined
+The quality of a functional block cannot be determined
 by the difference between its inputs and outputs (as we did for
-ports and channels). The behavior of a functional block
+ports and channels), because the behavior of a functional block
 cannot be determined in general; since any functional block will have 
 its own purpose based on functional requirements. 
-Therefore, while a port semantics is determined by the relation 
+Therefore, while the semantics of a port is determined by the relation 
 between assertions and beliefs, the semantics of a functional block 
 is determined by the relation between facts/requirements and I/O beliefs.
-
 In other words, a functional block is a generic agent with no pre-defined general
@@
 any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
-Figure~\ref{fig:soundness}, facts are strategic rules that define how
+Figure~\ref{fig:soundness}, facts are definitory rules that define how
 the system shall behave (by specification), while reality may be shown to
@@
 without going into the design of the behavior of agents. We can give an example
-but research is needed to apply our theory
+but research is needed to apply our hypothesis
 to the behavior of agents.
@@
 
-We summarize our results by categorizing the weaknesses
-predicted by our theory into: data-flow-related and functionality-related
-weaknesses; as in Table~\ref{tab:weaknesses}.
+We summarize our results by categorizing the weaknesses predicted by our hypothesis
+into: data-flow-related and functionality-related weaknesses; as in
+Table~\ref{tab:weaknesses}. Functionality-related weaknesses can be seen as a
+general formulation of our hypothesis while data-flow-related weaknesses as an
+application of our hypothesis to components with defined behavior/requirements.
 \begin{table*}[t]
@@
 	DR & Drops all the inputs and inserts new malicious data & The component never performs/carries the expected behavior/information\\[.1cm]
-	PP & Selectively drops inputs & Part of the expected outputs are not generated in response to the correct inputs \\[.1cm]
+	PP & Selectively drops inputs & part of the expected outputs are not generated in response to the correct inputs \\[.1cm]
 	PPi& Forwards all the inputs but crafts and inserts new malicious data & The components correctly performs/carries the expected behavior/information when the correct inputs are provided but is subject to input injections \\[.1cm]
@@
 flows) between functional blocks. We define $R\subseteq V\times F$, where $F$
-is the set of all the requirements of $S$, and extend $G$ as $G'=\langle
-E'\times V'\rangle$ with $E'= E\cup R$ and $V'=V\cup F$. Let $\pi:
+is the set of all the requirements of $S$, and extend $G$ as $G'=\langle V',E'\rangle$ with $E'= E\cup R$ and $V'=V\cup F$. Let $\pi:
 E'\rightarrow RCC$ (where $RCC$ is the set of relations in the RCC) be the
@@
 \subsection{Cybersecurity Risk Assessment}\label{sec:secra}
-In order to test our theory we implemented a tool-chain (open-source under the
+In order to test our hypothesis we implemented a tool-chain (open-source under the
 AGPLv3 license and available at \autocite{v-research2020cybersecurity}) for the
 identification of weaknesses and the precise calculation of potential insecure
-configurations. The engineering of the $\abftheory$ is summarized in the UML
+configurations. The engineering of the $\abftheory$-framework for CPS is summarized in the UML
 Class diagram in Figure~\ref{fig:secraclassdiagram} (in appendix).
@@
 \begin{itemize}
-	\item a \emph{Deployment Diagram} describes the \emph{Physical
-		Architecture}. Each agent is defined as a Node with (physical)
-		ports, and agent's ports are connected via information Flow
+	\item a \emph{deployment diagram} describes the \emph{physical
+		architecture}. Each agent is defined as an UML node with (physical)
+		ports, and agent's ports are connected via UML information flow
 		connectors, representing the physical channel.
-	\item a \emph{Functional Architecture} is linked to each agent in the
-		deployment diagram and is defined by an \emph{Object Diagram}.
-		The Object Diagram is composed by Instances of functional
-		blocks, connected via information Flow connectors.
+	\item a \emph{functional architecture} is linked to each agent in the
+		deployment diagram and is defined by an \emph{object diagram}.
+		The object diagram is composed by instances of functional
+		blocks, connected via information flow connectors.
 	\item the connection between the two diagrams is implemented by
@@
 The tool generates a graph-like internal structure which represents
-the specification (called ABF graph). The ABF graph defines the system as
+the specification (called $\abftheory$-graph). The $\abftheory$-graph defines the system as
 a number of regions of assertions, beliefs, and facts. Those regions
@@
 in the Excel file.
-
 In our approach, security requirements are not imposed by the
@@
 of the specified system.
-
-The risk matrix is often defined as a function of 
-likelihood and impact of attacks. We suggest
-that security weaknesses are all equally likely to be exploited
-if there's a connection between the weakness and the asset that
-an attacker wants to impact.
 
@@
 \end{figure*}
-The results, summarized in Figure~\ref{fig:results} reports more than 16 millions
+The results, summarized in Figure~\ref{fig:results}, reports more than 16 millions
 possible scenarios in which at least one component diverge from the specification.
@@
 \section{Conclusion and Future Works}
-\fixnote{mv}{Se ho capito bene, detto a parole mie, quello che il vostro
-approccio permette di fare, dati requisiti e architettura del sistema, è di
-elencare in maniera sistematica tutti i "punti" in cui potrebbero sorgere dei
-problemi. Spiegherei un po' meglio, qui nelle conclusioni o nella sezione
-precedente, come questo approccio dovrebbe essere usato nella pratica... Voglio
-dire che non mi sembra necessariamente in opposizione ad altri approcci ma
-magari può integrarsi a questi.  Inoltre quando parlate dei passi successivi,
-ad esempio il considerare la generazione di test cases, se avete già qualche
-idea su come procedere, potreste discuterne brevemente.}
-We proposed a foundational theory on security, arguing that
+%\fixnote{mv}{Se ho capito bene, detto a parole mie, quello che il vostro
+%approccio permette di fare, dati requisiti e architettura del sistema, è di
+%elencare in maniera sistematica tutti i "punti" in cui potrebbero sorgere dei
+%problemi. Spiegherei un po' meglio, qui nelle conclusioni o nella sezione
+%precedente, come questo approccio dovrebbe essere usato nella pratica... Voglio
+%dire che non mi sembra necessariamente in opposizione ad altri approcci ma
+%magari può integrarsi a questi.  Inoltre quando parlate dei passi successivi,
+%ad esempio il considerare la generazione di test cases, se avete già qualche
+%idea su come procedere, potreste discuterne brevemente.}
+We proposed an hypothesis for a foundational theory on security, arguing that
 security-related issues are not related to the maliciousness of an agent but to
@@
 of proper security risk assessment processes.
-Those design, again, lack of security verification processes based on a solid 
+Those designs, again, lack of security verification processes based on a solid 
 security foundational theory and then permit the generation of insecure implementation.
 The security verification and test-case generation will be the focus of our next steps.
-
-We conclude that the problem of security is a problem related to the 
-many possible design and, in turn, implementation given a specification.
-The problem is related to the epistemological search for truth,
-where the challenge is to relate information and belief to human knowledge.
-Scientifically, generalizing the human to an agent, the problem is
-to relate assertions and beliefs, to facts. From an engineering standpoint,
-by generalizing the concept of agents as architectural subsystems, the
-problem is to link channels (and ports) and functional architectures to 
-requirements.
+%We conclude that the problem of security is a problem related to the 
+%many possible design and, in turn, implementation given a specification.
+%The problem is related to the epistemological search for truth,
+%where the challenge is to relate information and belief to human knowledge.
+%Scientifically, generalizing the human to an agent, the problem is
+%to relate assertions and beliefs, to facts. From an engineering standpoint,
+%by generalizing the concept of agents as architectural subsystems, the
+%problem is to link channels (and ports) and functional architectures to 
+%requirements.
 
@@
 	\includegraphics[width=\textwidth]{secra_classDiagram.png}
-	\caption{Security Risk Assessment -- Class Diagram}
+	\caption{$\abftheory$-framework for CPS Design -- Class Diagram}
 	\label{fig:secraclassdiagram}

### 2020-08-03 reports/paper_0/main.tex

@@
 	development of a theory that determines (all and only) the possible
-	insecurity and security configurations of any abstract system.  Our
+	insecure and secure configurations of any abstract system.  Our
 	hypothesis is that cybersecurity weaknesses (i.e. errors) are at the
@@
 support the non-falseness of a scientific theory. 
-
 If we consider a natural (real-world) phenomenon, such as gravity, we can think
@@
 observation that allows us to declare an arbitrary system or technique
-secure''. With security, Herley (as in the reminder of this paper) only 
-focuses on cybersecurity and his intuition is that there is no scientific theory that can predict
-the security of a system, nor a theory that can predict all possible
-insecurities of a system (which, by negation may be used as a theory of
-security).  Herley then uses this argument to show that ``claims that any
-measure is necessary for security are empirically unfalsifiable''. An example
-is whether or not password masking makes a system secure. Password masking
-prevents a complete leakage of the password and a system in which password
-masking is implemented is believed to be secure because an attacker won't be
-seeing the password by ``shoulder surging''.  However, this reasoning is
-circular and it is always true that a system in which password masking is
+secure''. With security, Herley only focuses on cybersecurity (we also use
+security and insecurity, in this paper, only to refer to cyberinsecurity and
+cybersecurity) and his intuition is that there is no scientific theory that can
+predict the cybersecurity of a system, nor a theory that can predict all
+possible insecurities of a system (which, by negation may be used as a theory
+of cybersecurity).  Herley then uses this argument to show that ``claims that
+any measure is necessary for security are empirically unfalsifiable''. An
+example is whether or not password masking makes a system secure. Password
+masking prevents a complete leakage of the password and a system, in which
+password masking is implemented, is believed to be secure because an attacker
+won't be seeing the password by ``shoulder surfing''.  However, this reasoning
+is circular and it is always true that a system in which password masking is
 implemented is a system in which a password is masked. No experiment can ever
@@
 unfalsifiability is a fallacy of a theory), Herley concludes that there is no
-scientific theory of cybersecurity; which means that cybersecurity lays in the
-realm of pseudo-sciences \autocite{Herley2016usenixvideo}. Herley, e.g.  in
+scientific theory supporting this type of cybersecurity measures; generalizing,
+cybersecurity seems to lay in the realm of pseudo-sciences
+\autocite{Herley2016usenixvideo}. Herley, e.g.  in
 \autocite{Herley2017justifying}, discusses the implications of a nonscientific
-approach to cybersecurity, and highlights the tremendous impact on all the
+approach to cybersecurity, and highlights the tremendous impact on the
 scientific research and engineering of systems. While the criticism is
@@
 theory.  We consider the problem raised by Herley not confined to ``computer
-security'' but rather we reason on any abstract system (so that our hypothesis may
+security'' but rather we reason on any abstract system (so that our scientific
+hypothesis\footnote{In the reminder of this paper, we will use the word
+hypothesis to refer to ``scientific hypothesis'' as a proposed scientific
+theory that has not gone through an extensive series of tests. Instead, we
+use ``logical theory'' to refer to a set of formal logical axioms.} may
 be tested in any sound implementation such as networks, mechanical, cyber, or
@@
 a set of abstract properties in many formal approaches to the investigation of
-the security of systems.  Let's, for the sake of simplicity, consider the case
-of a system implementing a security protocol (e.g. an authentication protocol).
-The proof of security obtained from the formal verification of the design of
+the cybersecurity of systems.  Let's, for the sake of simplicity, consider the case
+of a system implementing a cybersecurity protocol (e.g. an authentication protocol).
+The proof of cybersecurity obtained from the formal verification of the design of
 the protocol (e.g. with respect to a formalization of the authentication
-property) is indeed falsifiable against the security properties verified.  For
+property) is indeed falsifiable against the cybersecurity properties verified.  For
 example, in the protocol verification community, cybersecurity is often defined
@@
 process (of systems) without considering many aspects related to the design or
-implementation of a protocol that may generate security vulnerabilities later
+implementation of a protocol that may generate cybersecurity vulnerabilities later
 in the development process.  An infamous example is the KRAK attack on the WPA2
@@
 4-way handshake'' but ``the problem is that the proofs do not model key
-installation.'' The theories underlying the verification of security protocols
-are based on assumptions which non-evidently apply to a general security
+installation.'' The theories underlying the verification of cybersecurity protocols
+are based on assumptions which non-evidently apply to a general cybersecurity
 theory; otherwise the assumptions would clearly state the necessary element
@@
 cryptography assumption (i.e. assuming that the cryptographic primitives have
-no security flaw), and assuming that no violation to any security property
+no cybersecurity flaw), and assuming that no violation to any cybersecurity property
 occurs after message I), the freedom of choosing the scope determines that the
@@
 be considered in the verification process.  This choice has a tremendous impact
-on the focus and findings of the verification of the security of the protocol.
+on the focus and findings of the verification of the cybersecurity of the protocol.
 While this may seem to turn upon minutiae and foreseeable, this highlights the
 false sense of security that may derive from a non-falsifiable theory of system
-security. 
+cybersecurity. 
 In other words, if (as Herley mentions in \autocite{Herley2016usenixvideo}) 
@@
 system is secure'' the formula becomes tautological. In fact, $P$ and $Secure$ can 
-be considered as semantically equivalent (security is confidentiality and vice-versa).
+be considered as semantically equivalent (cybersecurity is confidentiality and vice-versa).
 Instead of starting from reasoning on what makes a system secure or insecure,
@@
 We focus on insecurities only caused by the exploitation of
-cybersecurity attacks, and we assume that achieving security means preventing all those attacks 
+cybersecurity attacks, and we assume that achieving cybersecurity means preventing all those attacks 
 from being exploitable or exploited.
@@
 or testing if other (not predicted by our hypothesis) weaknesses are present.
-In fact, if any security weaknesses were to be found in a system and not
+In fact, if any cybersecurity weaknesses were to be found in a system and not
 predicted by our hypothesis, the hypothesis could be declared incomplete.  If a
-security weakness would be predicted by our hypothesis but found to be impossible
+cybersecurity weakness would be predicted by our hypothesis but found to be impossible
 to realize, our hypothesis could be declared as wrong.
@@
 reporting a literature review on the main concepts and definitions related to
-security.  We formulate a security hypothesis in Section~\ref{sec:hypothesis};
-which we use to propose a mathematical formulation of our hypothesis on system security in
+cybersecurity.  We formulate a cybersecurity hypothesis in Section~\ref{sec:hypothesis};
+which we use to propose a mathematical formulation of our hypothesis on system cybersecurity in
 Section~\ref{sec:theory}. In Section~\ref{sec:secra}, we apply our hypothesis for
-the tool-assisted security risk assessment of a Cyber-Physical System (CPS), with an ad-hoc example.
+the tool-assisted cybersecurity risk assessment of a Cyber-Physical System (CPS), with an ad-hoc example.
 %based on the SWaT testbed \autocite{Mathur2016swat}. 
 This application shows
-how our hypothesis can be used to predict all of the possible security weaknesses
+how our hypothesis can be used to predict all of the possible cybersecurity weaknesses
 of a system, allowing the falsification of our hypothesis.
@@
 	\includegraphics[width=.8\textwidth]{safety-security_3.pdf}
-	\caption{Overview of keywords related to security and safety}
+	\caption{Overview of Keywords Related to Cybersecurity and Safety}
 	\label{fig:safety-security}
@@
 security) are expressed by the same word, e.g. sicurezza in Italian.  A
-semantic distinction between safety and security is correlated to a
+semantic distinction between safety and (cyber)security is correlated to a
 belief\footnote{A belief has to be intended as a proposition which is supposed
@@
 natural environment (e.g. natural events such as wearing of hardware
-components), whereas security deals with \emph{incidents} posed by mankind (e.g.
+components), whereas cybersecurity deals with \emph{incidents} posed by mankind (e.g.
 attackers and bugs). The fundamental difference between nature and mankind (and,
@@
 specification, design or implementation of a system. The correlation 
-between insecurity flaws and security requirements has been used to define
+between insecurity flaws and cybersecurity requirements has been used to define
 standards such as the IEC 62443-1-3 (the Industrial communication networks -
@@
 security levels) which defines requirements as ``confidentiality of information
-in transit/at-rest''. More generally, the idea of defining security requirements
+in transit/at-rest''. More generally, the idea of defining cybersecurity requirements
 as properties of a system was initially defined in 1970s with the 
@@
 RITE approach (see \autocite{Samonas2014cia} for an overview of the evolution of the CIA triad).
-The link between security requirements and vulnerabilities is reported 
+The link between cybersecurity requirements and vulnerabilities is reported 
 in the NVD databases by the CVSS \autocite{Mell2007CVSS} scoring system.
@@
 and quantitatively evaluates the impact on the CIA triad.
-While security requirements, weaknesses, vulnerabilities, and attacks
+While cybersecurity requirements, weaknesses, vulnerabilities, and attacks
 have been extensively studied and implemented both in academia and industry
@@
 no scientific falsifiable theory correlate
-security requirements to necessary and sufficient conditions (e.g. mitigations) 
+cybersecurity requirements to necessary and sufficient conditions (e.g. mitigations) 
 to declare a system secure \autocite{Herley2016unfalsifiability}.
@@
 as CAPEC \autocite{MITRE2020CAPEC}; or that dishonesty of an agent with respect
-to a security protocol should be defined as a number of predefined actions as
+to a cybersecurity protocol should be defined as a number of predefined actions as
 in
@@
 
-\section{A Security Hypothesis in the $\abftheory$-framework}\label{sec:hypothesis}
+\section{A Cybersecurity Hypothesis in the $\abftheory$-framework}\label{sec:hypothesis}
 In order to address the problem raised by Herley, we shall define how to
@@
 to the maliciousness of agents interacting with the system, we show that
-security doesn't seem to stem from a malicious nature but, rather, insecurity
-raises from the lack of well-defined security requirements for the development
-process of a system. We argue that the high number of security vulnerability
+cybersecurity doesn't seem to stem from a malicious nature but, rather, insecurity
+raises from the lack of well-defined cybersecurity requirements for the development
+process of a system. We argue that the high number of cybersecurity vulnerability
 reported today are simply the realization of potential system configurations,
@@
 As a reference, we consider the following steps of the engineering process
-and our objective is to be able to predict the security requirements
+and our objective is to be able to predict the cybersecurity requirements
 instead of axiomatically assume them.
@@
 	\item \emph{Architecture Design}, where the specification is structured into \emph{functional and physical architectures}
-	\item \emph{Security Risk Assessment}, where the potential \emph{weaknesses} (errors)
-		are identified and, consequently, \emph{security requirements} are
+	\item \emph{Cybersecurity Risk Assessment}, where the potential \emph{weaknesses} (errors)
+		are identified and, consequently, \emph{cybersecurity requirements} are
 		captured.
@@
 	%	the functional blocks (behaviors), the data flow, and communication logic.
-	%\item \emph{Verification of Security Requirements}, where the data flow, behaviors, and communication protocols are verified against the security requirements.
+	%\item \emph{Verification of Security Requirements}, where the data flow, behaviors, and communication protocols are verified against the cybersecurity requirements.
 \end{enumerate}
@@
 CVSS) to assign a quantitative evaluation 
-of the security of each entry. Instead, we want a precise calculation
+of the cybersecurity of each entry. Instead, we want a precise calculation
 of the number of insecurity configurations to emerge from the mathematical
@@
 state space for a single agent. The color coding of the table represents the
-security risk related to a generic agent, the risk is highest on the top left
+cybersecurity risk related to a generic agent, the risk is highest on the top left
 corner of the matrix, lowest on the bottom right corner.
@@
 
-\subsubsection{$\abf$ Security Enumeration (SE)} The following security
+\subsubsection{$\abf$ CyberSecurity Enumeration (CSE)} The following cybersecurity
 requirement for a CPS specification can be summarized:
 \begin{enumerate}
-	\item[SE-$1$] Proper interaction between correctly-behaving agents is
+	\item[CSE-$1$] Proper interaction between correctly-behaving agents is
 		defined as $\eq{\assertionRegion_a}{\beliefRegion_a}$ for an
@@
 	\begin{enumerate}
-		\item[SE-$1.1$] The equality relation
+		\item[CSE-$1.1$] The equality relation
 			$\eq{\assertionRegion_{s\rightarrow
@@
 			inputs and expected inputs.
-		\item[SE-$1.2$] Similarly, the equality relation
+		\item[CSE-$1.2$] Similarly, the equality relation
 			$\eq{\assertionRegion_{a\rightarrow
@@
 	\end{enumerate}
-\item[SE-$2$] The proper adherence of the data transmitted between agents with respect to
+\item[CSE-$2$] The proper adherence of the data transmitted between agents with respect to
 	requirements, is defined as $\eq{\assertionRegion}{\factRegion}$. 
-\item[SE-$3$] The proper adherence of the behavior (in terms of input and output beliefs) with respect to
+\item[CSE-$3$] The proper adherence of the behavior (in terms of input and output beliefs) with respect to
 		requirements is defined as $\eq{\beliefRegion}{\factRegion}$.
@@
 
-We note that our SE define the properties of a secure system, and correlated
-weaknesses can be found in the CWE dataset.  For example, SE-$1$ can be seen as
+We note that our CSE define the properties of a secure system, and correlated
+weaknesses can be found in the CWE dataset.  For example, CSE-$1$ can be seen as
 correlated to the weakness class of ``Improper Interaction Between Multiple
-Correctly-Behaving Entities'' defined by the CWE--$435$, SE-$2$ with the
+Correctly-Behaving Entities'' defined by the CWE--$435$, CSE-$2$ with the
 ``Insufficient Control Flow Management'' defined by MITRE in the CWE--$691$,
-and SE-$3$ with the ``Improper Calculation'' defined by MITRE in the
+and CSE-$3$ with the ``Improper Calculation'' defined by MITRE in the
 CWE--$682$. All those CWE are in the top ``view'' of the ``research concepts''
@@
 We are now in the position to define what a secure system is (with respect to
-the $\abftheory$-framework) and, based on that definition, what the security risk is and
+the $\abftheory$-framework) and, based on that definition, what the cybersecurity risk is and
 how to quantify it in a risk matrix. The following definition holds for 
@@
 for CPS afterwards in the paper.
-\begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
-	A secure system is a system where SE-$1$, SE-$2$, and SE-$3$ holds for
+\begin{definition}{\bf Cybersecurity of a System or an Agent --}\label{def:security}
+	A secure system is a system where CSE-$1$, CSE-$2$, and CSE-$3$ holds for
 	each agent composing the system.
@@
 and what is the magnitude of the incidents produced by this exploitation), we suggest
-that security weaknesses are all equally likely to be exploited
+that cybersecurity weaknesses are all equally likely to be exploited
 if there's a connection between the weakness and the asset that
@@
 
-\section{Prediction of Security Weaknesses}\label{sec:theory}
+\section{Prediction of Cybersecurity Weaknesses}\label{sec:theory}
 Several standards mandate a secure-by-design approach in which cybersecurity
@@
 \end{enumerate}
-However, standards do not describe in detail how to perform a security risk 
+However, standards do not describe in detail how to perform a cybersecurity risk 
 assessment and only vaguely define the overall objective, which
-can be summarized as to provide an understanding of the potential security risks.
-Roughly speaking, a security risk assessment (e.g. CORAS \autocite{Lund2010model}) 
+can be summarized as to provide an understanding of the potential cybersecurity risks.
+Roughly speaking, a cybersecurity risk assessment (e.g. CORAS \autocite{Lund2010model}) 
 methodology starts from the (i) identification of assets, (ii) determines
@@
 As Herley puts it \autocite{Herley2016usenixvideo}, ``we don't have a test for security and
-we don't have a metric for security'' while standards require a security risk assessment
-which, in turn, mandates a metric to evaluate the security risk. 
+we don't have a metric for security'' while standards require a cybersecurity risk assessment
+which, in turn, mandates a metric to evaluate the cybersecurity risk. 
 In contrast, in this section, we define how to specify a CPS in the $\abftheory$-framework
-and we identify a security metric.
+and we identify a cybersecurity metric.
 
@@
 to MAS. We now map MAS to CPS and show how this mapping allows us
-to predict security weaknesses in a system.
+to predict cybersecurity weaknesses in a system.
 
@@
 
-\paragraph{Security Weaknesses -- The RIDI-Hypothesis} 
+\paragraph{Cybersecurity Weaknesses -- The RIDI-Hypothesis} 
 It is important to note that all the results of the application of the
@@
 four main categories of weaknesses: Replace, Insertion, Drop, Injection. We can,
-then, deduce the following security properties to mitigate security weaknesses
+then, deduce the following cybersecurity properties to mitigate cybersecurity weaknesses
 of a port or a channel (between ports, functional blocks, or both).
@@
 be extended to capture the semantics (i.e. the logic) of the communication and
-security protocols, with the relation between assertions and facts we would
+cybersecurity protocols, with the relation between assertions and facts we would
 compare protocol logics with the requirements. We won't consider the verification of
@@
 \end{tabular}
-\caption{Weaknesses Categorized~\label{tab:weaknesses}}
+\caption{Weaknesses Categorization~\label{tab:weaknesses}}
 \end{table*}
@@
 	\includegraphics[width=.9\columnwidth]{v-secra.pdf}
-	\caption{Security Risk Assessment Tool}
+	\caption{Cybersecurity Risk Assessment Tool}
 	\label{fig:secra}
 \end{figure}
-As depicted in Figure~\ref{fig:secra}, the security risk assessment process
+As depicted in Figure~\ref{fig:secra}, the cybersecurity risk assessment process
 starts with the definition of the use cases and architectural requirements; the
@@
 The Excel file also reports the total number of configurations as indicating
-the security risk associated to the specification. A user can change the status
+the cybersecurity risk associated to the specification. A user can change the status
 of each weakness in the Excel file, from the initial (default) status (open)
@@
 in the Excel file.
-In our approach, security requirements are not imposed by the
+In our approach, cybersecurity requirements are not imposed by the
 specification but are automatically extracted by our tool as potential weaknesses.
@@
 	\centering
-	\includegraphics[width=.6\columnwidth]{eng_cs1.png}
-	\caption{Water Level Reader Deployment Diagram}
+	\includegraphics[width=.5\columnwidth]{eng_cs1.png}
+	\caption{Water level reader deployment diagram}
 	\label{fig:eng_cs1}
@@
 We proposed an hypothesis for a foundational theory on security, arguing that
-security-related issues are not related to the maliciousness of an agent but to
-the vagueness of the security controls on the engineering processes.
+cybersecurity-related issues are not related to the maliciousness of an agent but to
+the vagueness of the cybersecurity controls on the engineering processes.
 The current state-of-the-art processes allow, given a specification,
-an engineer to design the system in such a way that security issues arise due to the lack
-of proper security risk assessment processes.
-Those designs, again, lack of security verification processes based on a solid 
-security foundational theory and then permit the generation of insecure implementation.
-The security verification and test-case generation will be the focus of our next steps.
+an engineer to design the system in such a way that cybersecurity issues arise due to the lack
+of proper cybersecurity risk assessment processes.
+Those designs, again, lack of cybersecurity verification processes based on a solid 
+cybersecurity foundational theory and then permit the generation of insecure implementation.
+The cybersecurity verification and test-case generation will be the focus of our next steps.
 %We conclude that the problem of security is a problem related to the 
@@
 
-\section*{Acknowledgment}
-We thank Katia Santac\`a for the help in the development of the initial ideas, and for
-the helpful discussions that allowed us to make this step forward.
+%\section*{Acknowledgment}
+%We thank Katia Santac\`a and Luca Vigan\`o for their help in the development of
+%the initial ideas, and for the helpful discussions that allowed us to make this
+%step forward. We also thank Marco Volpe for his valuable feedback.
 

### 2020-08-03 reports/paper_0/main.tex
review abstract

@@
 \begin{abstract}
-	The objective of this research is lay the foundation for the
+	The objective of this research is to lay the foundations for the
 	development of a theory that determines (all and only) the possible
-	insecure and secure configurations of any abstract system.  Our
-	hypothesis is that cybersecurity weaknesses (i.e. errors) are at the
-	bottom of the causality chain that leads to cybersecurity attacks.
-	Based on this hypothesis, we formulate a scientific (falsifiable by
-	experiments) hypothesis that predicts all the weaknesses that lead to
-	the presence of cybersecurity weaknesses in the architectural design of
-	an abstract system. The mathematical formulation of our hypothesis is
+	insecure and secure configurations of any abstract system. We claim
+	that cybersecurity weaknesses (i.e. errors) are at the beginning of the
+	causality chain that leads to cybersecurity attacks.  We, then,
+	formulate a scientific (falsifiable by experiments) hypothesis that we
+	use to predict all the weaknesses in the architectural design of an
+	abstract system. The mathematical formulation of our hypothesis is
 	based on a correlation between the epistemological concepts of
@@
 	mathematical formula.  We implemented a prototype cybersecurity risk
-	assessment tool that, based on our hypothesis, predicts weaknesses in a UML
-	model of a (cyber-physical) system.
+	assessment tool that, based on our hypothesis, predicts weaknesses in a
+	UML model of a (cyber-physical) system.
 \end{abstract}

### 2020-08-03 reports/paper_0/main.tex
intro review

@@
 be considered acceptable as long as no empirical experiment falsify it. For
-example, the theory of gravity, in its mathematical setting, can predict how
+example, a theory of gravity, in its mathematical setting, can predict how
 objects attract each other, and those predictions can be tested by empirical
-experiment. If the theory correctly abstracts the phenomenon, no empirical
+experiment. If the theory correctly represents the phenomenon, no empirical
 experiment should falsify the predictions of that theory.
@@
 \end{figure}
-As an illustrative example, in Figure~\ref{fig:protocol-example}, under the perfect
-cryptography assumption (i.e. assuming that the cryptographic primitives have
-no cybersecurity flaw), and assuming that no violation to any cybersecurity property
-occurs after message I), the freedom of choosing the scope determines that the
-flaws related to the dishonest impersonification of the Server may or may not
-be considered in the verification process.  This choice has a tremendous impact
-on the focus and findings of the verification of the cybersecurity of the protocol.
-While this may seem to turn upon minutiae and foreseeable, this highlights the
-false sense of security that may derive from a non-falsifiable theory of system
-cybersecurity. 
-In other words, if (as Herley mentions in \autocite{Herley2016usenixvideo}) 
-the problem is in the way we reason on cybersecurity, we should
-see it in the way we reply to simple cybersecurity (foundamental) questions.
-Let us consider the foundamental question ``what is cybersecurity?'',
-as in \autocite{hintikka2013inquiry}, we can translate the wh-question into an equivalent
-existentially quantified formula. 
-But if the formula is $\exists P. P(s)\rightarrow Secure(s)$,
-(where we ask if there exists a predicate $P$ such that, given any arbitrary 
-system $s$, when $P$ holds for $s$ then $s$ can be considered secure),
-when the answer is (for example) ``if the system is confidential, the
-system is secure'' the formula becomes tautological. In fact, $P$ and $Secure$ can 
-be considered as semantically equivalent (cybersecurity is confidentiality and vice-versa).
+As an illustrative example, in Figure~\ref{fig:protocol-example}, under the
+perfect cryptography assumption (i.e. assuming that the cryptographic
+primitives have no cybersecurity flaw), and assuming that no violation to any
+cybersecurity property occurs after message I), the freedom of choosing the
+scope determines that the flaws related to the dishonest impersonification of
+the Server may or may not be considered in the verification process.  This
+choice has a tremendous impact on the focus and findings of the verification of
+the cybersecurity of the protocol.  While this may seem to turn upon minutiae
+and foreseeable, this highlights the false sense of security that may derive
+from a non-falsifiable theory of system cybersecurity.  In other words, as
+Herley mentions in \autocite{Herley2016usenixvideo}, the problem is in the way
+we reason on cybersecurity, and we can see it in the way we reply to simple
+cybersecurity (foundamental) questions.  Let us consider the general question
+``what is cybersecurity?''; as in \autocite{hintikka2013inquiry}, we can
+translate the wh-question into an equivalent existentially quantified formula.
+But if the formula is $\exists P. P(s)\rightarrow Secure(s)$, (where we ask if
+there exists a predicate $P$ such that, given any arbitrary system $s$, when
+$P$ holds for $s$ then $s$ can be considered secure), when the answer is (for
+example) ``if the system is confidential, the system is secure'' the formula
+becomes tautological. In fact, $P$ and $Secure$ can be considered as
+semantically equivalent (cybersecurity is confidentiality and vice-versa).
 Instead of starting from reasoning on what makes a system secure or insecure,
-we reason on what causes insecurities.
-We focus on insecurities only caused by the exploitation of
-cybersecurity attacks, and we assume that achieving cybersecurity means preventing all those attacks 
-from being exploitable or exploited.
-Our hypothesis is that cybersecurity attacks are only caused by 
-the presence of errors in the design or implementation of a system (i.e. cybersecurity weaknesses).
-For example, a design or implementation weakness in a sanitization function
-may allow an attacker to perform SQL-injections. If weaknesses are necessary
-to enable cybersecurity attacks, a theory that predicts all potential 
-weaknesses can be effectively used to predict the
-(in)security of a system.
-Therefore, 
-a possible answer to the initial question can be that (definition) cybersecurity is the 
-absence of attacks and (claim) the absence of weaknesses implies cybersecurity.
-This, however, raises another question: ``which are all the errors/weaknesses
-in (the engineering of) a system that make it possible for a cybersecurity attack
-to happen?''. More rigorously, which scientific theory predicts all
-possible weaknesses that causes cybersecurity attacks?
+we reason on what causes insecurities.  We focus on insecurities only caused by
+the exploitation of cybersecurity attacks, and we assume that achieving
+cybersecurity means preventing all those attacks from being exploitable or
+exploited.  Our hypothesis is that \emph{cybersecurity attacks are only caused by the
+presence of errors in the design or implementation of a system} (i.e.
+cybersecurity weaknesses).  For example, a design or implementation weakness in
+a sanitization function may allow an attacker to perform SQL-injections. If
+weaknesses are necessary to enable cybersecurity attacks, a theory that
+predicts all potential weaknesses can be effectively used to predict the
+(in)security of a system.  Therefore, a possible answer to the initial question
+can be that (definition) cybersecurity is the absence of attacks and (claim)
+the absence of weaknesses implies cybersecurity.  This, however, raises another
+question: ``which are all the errors/weaknesses in (the engineering of) a
+system that make it possible for a cybersecurity attack to happen?''. More
+rigorously, which scientific theory predicts all possible weaknesses that
+causes cybersecurity attacks?
 

### 2020-08-25 reports/paper_0/main.tex
minor changes in abstract

@@
 	The objective of this research is to lay the foundations for the
-	development of a theory that determines (all and only) the possible
+	development of a scientific theory that determines (all and only) the possible
 	insecure and secure configurations of any abstract system. We claim

### 2020-09-01 reports/paper_0/main.tex
full review

@@
 	mathematical formula.  We implemented a prototype cybersecurity risk
-	assessment tool that, based on our hypothesis, predicts weaknesses in a
+	assessment tool that, based on our hypothesis, predicts the weaknesses in a
 	UML model of a (cyber-physical) system.
@@
 secure''. With security, Herley only focuses on cybersecurity (we also use
-security and insecurity, in this paper, only to refer to cyberinsecurity and
+security and insecurity, in this paper, only to refer to cyber-insecurity and
 cybersecurity) and his intuition is that there is no scientific theory that can
@@
 example is whether or not password masking makes a system secure. Password
-masking prevents a complete leakage of the password and a system, in which
+masking prevents a complete leakage of the password, and a system, in which
 password masking is implemented, is believed to be secure because an attacker
-won't be seeing the password by ``shoulder surfing''.  However, this reasoning
+won't be seeing the password by ``shoulder surfing''.  This reasoning
 is circular and it is always true that a system in which password masking is
@@
 cyber-physical system, or even a single computer or a single device such as a
-hard-drive).  There is also an apparent inconsistency in
-\autocite{Herley2016unfalsifiability} that we seek to clarify before following
-(as we agree) the scientific path drawn by Herley: cybersecurity is defined as
-a set of abstract properties in many formal approaches to the investigation of
-the cybersecurity of systems.  Let's, for the sake of simplicity, consider the case
-of a system implementing a cybersecurity protocol (e.g. an authentication protocol).
-The proof of cybersecurity obtained from the formal verification of the design of
-the protocol (e.g. with respect to a formalization of the authentication
-property) is indeed falsifiable against the cybersecurity properties verified.  For
-example, in the protocol verification community, cybersecurity is often defined
-as a formalization of the high-level properties confidentiality, integrity, and
-authentication. The problem in such approaches is not the definition of what
-cybersecurity is but the generality of the results, since they tackle a
-specific step\footnote{In fact, engineering models such as the V-model start
-with the requirements definition, and physical and functional architecture
-design, before the logic of protocols is even considered.} of the engineering
-process (of systems) without considering many aspects related to the design or
-implementation of a protocol that may generate cybersecurity vulnerabilities later
-in the development process.  An infamous example is the KRAK attack on the WPA2
-4-way handshake protocol where the author wrote \autocite{krak2020web}: ``our
-attacks do not violate the security properties proven in formal analysis of the
-4-way handshake'' but ``the problem is that the proofs do not model key
-installation.'' The theories underlying the verification of cybersecurity protocols
-are based on assumptions which non-evidently apply to a general cybersecurity
-theory; otherwise the assumptions would clearly state the necessary element
-that needs to be considered (e.g. modeled) in order to obtain a secure system.
-As an example, the so called Dolev-Yao attacker model\footnote{The Dolev-Yao
-attacker model can be considered as abstract model of an attacker that can be
-used for the analysis of protocol specifications.} \autocite{Dolev1983security}
-only applies to specific instances (often called scenarios) and abstraction of
-the protocol.  This, in turn, creates a false sense of security since it
-requires non-justifiable assumptions on the abstraction of the system under
-verification.
+hard-drive).  
+%There is also an apparent inconsistency in
+%\autocite{Herley2016unfalsifiability} that we seek to clarify before following
+%(as we agree) the scientific path drawn by Herley: cybersecurity is defined as
+%a set of abstract properties in many formal approaches to the investigation of
+%the cybersecurity of systems.  Let's, for the sake of simplicity, consider the case
+%of a system implementing a cybersecurity protocol (e.g. an authentication protocol).
+%The proof of cybersecurity obtained from the formal verification of the design of
+%the protocol (e.g. with respect to a formalization of the authentication
+%property) is indeed falsifiable against the cybersecurity properties verified.  For
+%example, in the protocol verification community, cybersecurity is often defined
+%as a formalization of the high-level properties confidentiality, integrity, and
+%authentication. The problem in such approaches is not the definition of what
+%cybersecurity is but the generality of the results, since they tackle a
+%specific step\footnote{In fact, engineering models such as the V-model start
+%with the requirements definition, and physical and functional architecture
+%design, before the logic of protocols is even considered.} of the engineering
+%process (of systems) without considering many aspects related to the design or
+%implementation of a protocol that may generate cybersecurity vulnerabilities later
+%in the development process.  An infamous example is the KRAK attack on the WPA2
+%4-way handshake protocol where the author wrote \autocite{krak2020web}: ``our
+%attacks do not violate the security properties proven in formal analysis of the
+%4-way handshake'' but ``the problem is that the proofs do not model key
+%installation.'' The theories underlying the verification of cybersecurity protocols
+%are based on assumptions which non-evidently apply to a general cybersecurity
+%theory; otherwise the assumptions would clearly state the necessary element
+%that needs to be considered (e.g. modeled) in order to obtain a secure system.
+%As an example, the so called Dolev-Yao attacker model\footnote{The Dolev-Yao
+%attacker model can be considered as abstract model of an attacker that can be
+%used for the analysis of protocol specifications.} \autocite{Dolev1983security}
+%only applies to specific instances (often called scenarios) and abstraction of
+%the protocol.  This, in turn, creates a false sense of security since it
+%requires non-justifiable assumptions on the abstraction of the system under
+%verification.
 %More specifically, for the formal security verification of a system in which
@@
 %which is the only important one of this discussion of protocols} 
-More importantly, some of the choices made by the
-modelers/engineers using those formal approaches may completely change the
-results of the formal verification of the system (an interesting example on how
-difficult it is to even just compare the approaches is given in
-\autocite{Cremers2009comparing}). 
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.7\columnwidth]{protocol-example.pdf}
-	\caption{Abstraction of an ad-hoc esemplificative protocol execution}
-	\label{fig:protocol-example}
-\end{figure}
-As an illustrative example, in Figure~\ref{fig:protocol-example}, under the
-perfect cryptography assumption (i.e. assuming that the cryptographic
-primitives have no cybersecurity flaw), and assuming that no violation to any
-cybersecurity property occurs after message I), the freedom of choosing the
-scope determines that the flaws related to the dishonest impersonification of
-the Server may or may not be considered in the verification process.  This
-choice has a tremendous impact on the focus and findings of the verification of
-the cybersecurity of the protocol.  While this may seem to turn upon minutiae
-and foreseeable, this highlights the false sense of security that may derive
-from a non-falsifiable theory of system cybersecurity.  In other words, as
+%More importantly, some of the choices made by the
+%modelers/engineers using those formal approaches may completely change the
+%results of the formal verification of the system (an interesting example on how
+%difficult it is to even just compare the approaches is given in
+%\autocite{Cremers2009comparing}). 
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=.7\columnwidth]{protocol-example.pdf}
+%	\caption{Abstraction of an ad-hoc esemplificative protocol execution}
+%	\label{fig:protocol-example}
+%\end{figure}
+%As an illustrative example, in Figure~\ref{fig:protocol-example}, under the
+%perfect cryptography assumption (i.e. assuming that the cryptographic
+%primitives have no cybersecurity flaw), and assuming that no violation to any
+%cybersecurity property occurs after message I), the freedom of choosing the
+%scope determines that the flaws related to the dishonest impersonification of
+%the Server may or may not be considered in the verification process.  This
+%choice has a tremendous impact on the focus and findings of the verification of
+%the cybersecurity of the protocol.  While this may seem to turn upon minutiae
+%and foreseeable, this highlights the false sense of security that may derive
+%from a non-falsifiable theory of system cybersecurity.  In other words, 
+As
 Herley mentions in \autocite{Herley2016usenixvideo}, the problem is in the way
@@
 semantically equivalent (cybersecurity is confidentiality and vice-versa).
+
 Instead of starting from reasoning on what makes a system secure or insecure,
-we reason on what causes insecurities.  We focus on insecurities only caused by
+we reason on what causes insecurities. 
+We focus on insecurities only caused by
 the exploitation of cybersecurity attacks, and we assume that achieving
@@
 Nonetheless, the extensive body of literature have scientific foundations,
-for example, providing either providing necessary or sufficient conditions
+for example, providing either necessary or sufficient conditions
 for cybersecurity.
@@
 		Common vulnerabilities and Exposures (CVE) List, is a mistake
-		that can be directly used by a hacker to gain access to a
+		that can be directly used by an attacker to gain access to a
 		system or network.  The definition is circular if we interpret
@@
 		unauthorized use of an asset''; where an \emph{Asset} is
-		``anything that has value to the organization''. We note that for
+		``anything that has value to the organization''. Furthermore, we note that for
 		the purpose of this article, we do not want to focus on a specific
@@
 called adversary or attacker). In fact, weaknesses become vulnerability if an
-attacker exploits them in an attack.  This, however, just moves the problem of
+attacker exploits them in an attack.  This, however, just shifts the problem of
 defining what cybersecurity is as the problem of defining what a dishonest
@@
 entity of the system or using the system (e.g. a device, a software, or a human
-being) and dishonesty is not necessarily related to malicious motivation but
-also to incompetence or lack of skills. As in the Dolev-Yao theory, we may
+being), and dishonesty is not necessarily related to malicious motivation but
+also to incompetence or lack of skills. As in the Dolev-Yao attacker model\footnote{The Dolev-Yao
+attacker model can be considered as abstract model of an attacker that can be
+used for the analysis of protocol specifications.} \autocite{Dolev1983security}, we may
 correlate ``being dishonest'' to ``not following the intended behavior/rules''.
@@
 system.  For example, a software can be considered an agent of the system and
-whenever it has a bug, it can be exploited causing an Incident. However, this
+whenever it has a bug, it can be exploited causing an incident. This
 may lead to the conclusion that the dishonest behaviors of agents cannot be
@@
 		engineering is the one that builds (input) or describes
-		(output) the behavior of an agent (strategy rules
-		\footnote{``The logical structure of information is one of the
-		most basic and one of the simplest
-		thing in the wide and wonderful world of logical analysis. This
-		point can be put in a deeper perspective. A distinction
-		[\ldots] ought to be made [--] between two kinds of rules (or
-		principles) in any strategic activity like knowledge seeking.
-		On the one hand you have the rules that define the game, e.g.
-		how chessmen are moved on a board. The can be called
-		\emph{definitory} rules.  They must be distinguished from rules
-		[\ldots] that deal with what is better and what is worse in the
-		game in question.  Definitory rules do not say anything about
-		this subject. Rules which do can be called \emph{strategic
-		rules}'' -- Hintikka in \autocite{Hintikka1993Information}}),
+		(output) the behavior of an agent 
+		%(definitory rules
+		%\footnote{``The logical structure of information is one of the
+		%most basic and one of the simplest
+		%thing in the wide and wonderful world of logical analysis. This
+		%point can be put in a deeper perspective. A distinction
+		%[\ldots] ought to be made [--] between two kinds of rules (or
+		%principles) in any strategic activity like knowledge seeking.
+		%On the one hand you have the rules that define the game, e.g.
+		%how chessmen are moved on a board. The can be called
+		%\emph{definitory} rules.  They must be distinguished from rules
+		%[\ldots] that deal with what is better and what is worse in the
+		%game in question.  Definitory rules do not say anything about
+		%this subject. Rules which do can be called \emph{strategic
+		%rules}'' -- Hintikka in \autocite{Hintikka1993Information}},
 		and
-	\item We consider a set of axiomatic \emph{facts} (definitory
+	\item We consider a set of axiomatic \emph{facts} 
+		(definitory
 		rules\footnote{Facts are definitory rule as they don't
@@
 		how it should be, trough a series of requirements. Those requirements 
-		may not hold, through insecurity, in the implemented system.})
+		may not hold, through insecurity, in the implemented system.}\autocite{Hintikka1993Information})
 		instead of considering the more general epistemic definition of
@@
 of the cybersecurity of each entry. Instead, we want a precise calculation
-of the number of insecurity configurations to emerge from the mathematical
+of the number of insecure configurations to emerge from the mathematical
 formulation of our cybersecurity hypothesis.
@@
 A mereotopology, as defined e.g. in \autocite{Rachavelpula2017mereotopology},
-is an ordered mathematical structure where the basic relation between regions
+is a mathematical structure where the basic relation between regions
 is the reflexive and symmetric relation \emph{Connects With} $\subseteq$,
@@
 are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
-two pairs of regions\footnote{``I am not asserting, as Lotze did, that a
-relation between $X$ and $Y$ consists of a quality in $X$ and a quality in $Y$
--- a view which I regard as quite indefensible. -- I assert that a relation $Z$
-between $X$ and $Y$ involves the existence in $X$ of the quality ``having the
-relation $Z$ to $Y$'' so that a difference of relations always involves a
-difference in quality, and a change of relations always involves a change of
-quality.'' --  Ellis J. McTaggart, The Unreality of
-Time \autocite{Mctaggart1908unreality}}: 
+two pairs of regions:%\footnote{``I am not asserting, as Lotze did, that a
+%relation between $X$ and $Y$ consists of a quality in $X$ and a quality in $Y$
+%-- a view which I regard as quite indefensible. -- I assert that a relation $Z$
+%between $X$ and $Y$ involves the existence in $X$ of the quality ``having the
+%relation $Z$ to $Y$'' so that a difference of relations always involves a
+%difference in quality, and a change of relations always involves a change of
+%quality.'' --  Ellis J. McTaggart, The Unreality of
+%Time \autocite{Mctaggart1908unreality}}
 \begin{itemize}
@@
 incompetent agent. However, we consider honesty related to sharing
-truths\footnote{``[\ldots] truth exists only in the good man, but the true in
-the bad man as well; for it is possible for the bad man to utter something
-true.'' Sextus Empiricus, Outlines of Pyrrhonism,
-II-83 \autocite{Empiricus1990Pyrrhonism}}.  The relation between beliefs and
+truths.
+%\footnote{``[\ldots] truth exists only in the good man, but the true in
+%the bad man as well; for it is possible for the bad man to utter something
+%true.'' Sextus Empiricus, Outlines of Pyrrhonism,
+%II-83 \autocite{Empiricus1990Pyrrhonism}}
+The relation between beliefs and
 facts determines the competence (on the subjects defined by the facts) of an
@@
 	\item The concept of knowledge is related to the \emph{requirements}
-		that, in a factorial way, describe how the CPS shall behave and
+		that describe how the CPS shall behave and
 		its \emph{physical architecture}.
@@
 agent.  However, we considered a port as a special type of agent to avoid
-an \emph{infinite regress}. While a channel transfers information between agents and
+an \emph{infinite regress}, as describe in the following. While a channel transfers information between agents, and
 a functional architecture elaborate information, a port is simply a connector between
@@
 excluded by the $\abftheory$-framework, it would obviously lead to an infinite
-nested structure of ports. In order to avoid this, we assume
+nested structure of ports. In order to avoid this infinite structure, we assume
 that a port doesn't require any other means to transfer information from/to a channel
@@
 $\Rcc{\rassert{s}{a}}{\beliefRegion_a}$ for the input-port or
-$\Rcc{\rassert{r}{*}}{\beliefRegion_a}$ for the output-port.  A port is, in
+$\Rcc{\rassert{r}{*}}{\beliefRegion_a}$ for the output-port (where $*$ is a wildcard for an agent).  A port is, in
 fact, syntactic sugar to express the relation between assertions and beliefs.
@@
 represents the set of functional blocks and ports of $S$, and $E\subseteq
-V\times E$ is the set of pairs representing the channels and connections (data
+V\times V$ is the set of pairs representing the channels and connections (data
 flows) between functional blocks. We define $R\subseteq V\times F$, where $F$
@@
 In our approach, cybersecurity requirements are not imposed by the
-specification but are automatically extracted by our tool as potential weaknesses.
+specification but are automatically extracted by our tool as mitigations to potential weaknesses.
 Those weaknesses are related to the different insecure configurations 
@@
 As in Figure~\ref{fig:eng_cs1}, we defined 2 agents: sensorInTank and sensorBoard,
-which represent the physical reader that needs to be placed in a tank of water, and 
+which represent the physical reader that needs to be placed in a tank, and 
 the board that interprets the readings and outputs them as digital signals.
@@
 \end{figure}
-In Figure~\ref{fig:int_cs1} we report the functional architecture that
+In Figure~\ref{fig:int_cs1}, we report the functional architecture that
 receives the incoming communications from the sensor in the tank, and 
@@
 the vagueness of the cybersecurity controls on the engineering processes.
-The current state-of-the-art processes allow, given a specification,
-an engineer to design the system in such a way that cybersecurity issues arise due to the lack
-of proper cybersecurity risk assessment processes.
-Those designs, again, lack of cybersecurity verification processes based on a solid 
-cybersecurity foundational theory and then permit the generation of insecure implementation.
+We also provided a prototype tool for the quantitative estimation of the cybersecurity
+risk, based on a UML model of the system under assessment.
+%The current state-of-the-art processes allow, given a specification,
+%an engineer to design the system in such a way that cybersecurity issues arise due to the lack
+%of proper cybersecurity risk assessment processes.
+%Those designs, again, lack of cybersecurity verification processes based on a solid 
+%cybersecurity foundational theory and then permit the generation of insecure implementation.
 The cybersecurity verification and test-case generation will be the focus of our next steps.

### 2020-09-01 reports/paper_0/main.tex
anonymize

@@
 \usepackage{rotating}
+
+% Space hacks, ...yes, from outer space!
+%\addtolength{\textfloatsep}{-6mm}
+%\addtolength{\abovecaptionskip}{-3mm}
+%\addtolength{\columnsep}{-1mm}
+%\setlength{\evensidemargin}{0.1in}
+%\setlength{\oddsidemargin}{0.1in}
+%\setlength{\textwidth}{6.47in}
+%\setlength{\textwidth}{5.1in}
 
@@
 
-\author{\IEEEauthorblockN{Marco Rocchetto}
-\IEEEauthorblockA{V-Research\\
-Verona, Italy\\
-Email: marco@v-research.it}
-\and
-\IEEEauthorblockN{Francesco Beltramini}
-\IEEEauthorblockA{V-Research\\
-Verona, Italy\\
-Email: francesco@v-research.it}
+\author{
+%	\IEEEauthorblockN{Marco Rocchetto}
+%\IEEEauthorblockA{V-Research\\
+%Verona, Italy\\
+%Email: marco@v-research.it}
+%\and
+%\IEEEauthorblockN{Francesco Beltramini}
+%\IEEEauthorblockA{V-Research\\
+%Verona, Italy\\
+%Email: francesco@v-research.it}
 }
@@
 Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent.
-However, the same ideas revolved around for thousands of years\footnote{
+However, the same ideas revolved around for thousands of years
+\footnote{
 It is interesting to notice that 
@@
 \end{enumerate}
-However, standards do not describe in detail how to perform a cybersecurity risk 
+Standards do not describe in detail how to perform a cybersecurity risk 
 assessment and only vaguely define the overall objective, which
@@
 	\centering
-	\includegraphics[width=.8\textwidth]{mashyp.pdf}
+	\includegraphics[width=.7\textwidth]{mashyp.pdf}
 	\caption{Mapping Epistemological Concepts to (Cyber-Physical) Systems Engineering}

### 2020-09-01 reports/paper_0/main.tex
remove figure 9 to fit space

@@
 \subsection{From Multi-Agent to Cyber-Physical Systems}
-\begin{figure*}[t]
+\begin{figure}[t]
 	\centering
-	\includegraphics[width=.7\textwidth]{mashyp.pdf}
+	\includegraphics[width=\columnwidth]{abftheory.pdf} %{mashyp.pdf}
 	\caption{Mapping Epistemological Concepts to (Cyber-Physical) Systems Engineering}
 	\label{fig:mashyp}
-\end{figure*}
+\end{figure}
 
@@
 %$\Rcc{\assertionRegion_2}{\beliefRegion_3}$)
-As depicted in Figure~\ref{fig:atom}, so far we have
+%As depicted in Figure~\ref{fig:atom}, 
+So far we have
 considered information generated by a port $P_I$ and then sent through a
@@
 
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.9\columnwidth]{engineering_relations.pdf}
-	\caption{Relations between Ports and Channel}
-	\label{fig:atom}
-\end{figure}
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=.9\columnwidth]{engineering_relations.pdf}
+%	\caption{Relations between Ports and Channel}
+%	\label{fig:atom}
+%\end{figure}
 
@@
 \end{figure*}
-The results, summarized in Figure~\ref{fig:results}, reports more than 16 millions
+The results, summarized in Figure~\ref{fig:results}, reports 16777216
 possible scenarios in which at least one component diverge from the specification.

### 2020-09-02 reports/paper_0/main.tex
remove images (ports & behavior)

@@
 		accord'' (e.g. the physical universe) \autocite{wiki-nature}.
-		\begin{itemize}
-			\item There exist several terms to refer to an
-				\emph{attacker}, e.g. threat agent or threat source,
-				but, from now on, we consider the
-				Causality principle to be the \emph{threat
-				source}, Nature or Mankind to be the
-				\emph{threat agents} and an \emph{attacker} as
-				a specific malicious threat agent which materializes a
-				threat.
-		\end{itemize}
-	\item \emph{Vulnerability}\footnote{The term vulnerability is not
-		present in the Encyclopedia of Cryptography and Security, while
-		it is used in 12 entries (such as in the definition of
-		``penetration testing'' \autocite{caddy2005pentest})
-		highlighting how commonly this word is used without a proper
-		supporting semantics.}, as defined in \autocite{cnssi20104009}
+		There exist several terms to refer to an \emph{attacker}, e.g.
+		threat agent or threat source, but, from now on, we consider
+		the Causality principle to be the \emph{threat source}, Nature
+		or Mankind to be the \emph{threat agents} and an
+		\emph{attacker} as a specific malicious threat agent which
+		materializes a threat.
+	\item \emph{Vulnerability},
+		%\footnote{The term vulnerability is not
+		%present in the Encyclopedia of Cryptography and Security, while
+		%it is used in 12 entries (such as in the definition of
+		%``penetration testing'' \autocite{caddy2005pentest})
+		%highlighting how commonly this word is used without a proper
+		%supporting semantics.}, 
+		as defined in \autocite{cnssi20104009}
 		(and adopted in \autocite{nist2013800-53}), is a ``weakness in an
@@
 at the very early stages of the engineering process.
-
 As a reference, we consider the following steps of the engineering process
@@
 Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent.
-However, the same ideas revolved around for thousands of years
-\footnote{
+However, the same ideas revolved around for thousands of years\footnote{
 It is interesting to notice that 
-in \autocite{Empiricus1990Pyrrhonism}, the author states: ``The
- logical criterion also may be used in three senses -- of the
- agent, or the instrument, or the ``according to what''; the
- agent, for instance, may be a man, the instrument either sense
- perception or intelligence, and the ``according to what'' the
- application of the impression ``according to'' which the man
- proceeds to judge by means of one of the aforesaid
- instruments.''. The author than proceeds showing that
- even in this framework, knowledge still is non-apprehensible.
-}.
+in \autocite{Empiricus1990Pyrrhonism}, the author states: ``The logical
+criterion also may be used in three senses -- of the agent, or the instrument,
+or the ``according to what''; the agent, for instance, may be a man, the
+instrument either sense perception or intelligence, and the ``according to
+what'' the application of the impression ``according to'' which the man
+proceeds to judge by means of one of the aforesaid instruments''.}.
 Some relevant examples for our objective are the following.
@@
 		instead of considering the more general epistemic definition of
-		knowledge. Specifically, facts describes:
-		\begin{itemize}
-			\item The functional architecture of each agent
-			\item The physical/structural (HW/SW) architecture of each
-				subsystem of agents and agent within a
+		knowledge. Specifically, facts describes: the functional
+		architecture of each agent, and the physical/structural (HW/SW)
+		architecture of each subsystem of agents and agent within a
 		subsystem
-		\end{itemize}
 \end{enumerate}
@@
 \paragraph{Input and Output Ports}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\columnwidth]{IOports.pdf}
-	\caption{Input and Output Port of an agent}
-	\label{fig:ioports}
-\end{figure}
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=\columnwidth]{IOports.pdf}
+%	\caption{Input and Output Port of an agent}
+%	\label{fig:ioports}
+%\end{figure}
 Since the $\abftheory$-framework is a theory of agents, we could consider ports as
@@
 \paragraph{Communication Channels}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\columnwidth]{channel.pdf}
-	\caption{Communication over a Mono-directional Channel}
-	\label{fig:channel}
-\end{figure}
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=\columnwidth]{channel.pdf}
+%	\caption{Communication over a Mono-directional Channel}
+%	\label{fig:channel}
+%\end{figure}
 In this work, we only consider

### 2020-09-02 reports/paper_0/main.tex
anonymize ref to github repo

@@
 
-As already presented in \autocite{Santaca2016abf}, it follows that, defining
+As in \autocite{Santaca2016abf}, it follows that, defining
 a system with a fixed number of regions, there exist
@@
 In order to test our hypothesis we implemented a tool-chain (open-source under the
-AGPLv3 license and available at \autocite{v-research2020cybersecurity}) for the
+AGPLv3 license and available at \autocite{v-research2020cybersecurity-anonymous}) for the
 identification of weaknesses and the precise calculation of potential insecure

### 2020-09-02 reports/paper_0/main.tex
review up to IV-A-e

@@
 Nonetheless, the extensive body of literature have scientific foundations,
-for example, providing either necessary or sufficient conditions
+for example, providing formal frameworks for the verification of 
+properties %either necessary or sufficient conditions
 for cybersecurity.
@@
 There is no agreement between the research communities (e.g.
-Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent.
-However, the same ideas revolved around for thousands of years\footnote{
+Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent
+but the same ideas revolved around for thousands of years\footnote{
 It is interesting to notice that 
@@
 of the scope of this paper. 
-
-We now define those concepts abstractly, inspired by the
+Instead, we now define those concepts abstractly, inspired by the
 $\abftheory$-framework  defined in \autocite{Santaca2016abf} and, in
@@
 		that information is from a sender to 
-		recipient is defined; and we call it \emph{assertion}  
+		recipient; and we call it \emph{assertion}  
 	\item Similarly, the portion of \emph{beliefs} we consider for system
@@
 		%rules}'' -- Hintikka in \autocite{Hintikka1993Information}},
-		and
 	\item We consider a set of axiomatic \emph{facts} 
@@
 For completeness, we report in the next paragraph 
-the calculation done in \autocite{Santaca2016abf}.
+the calculation done by the authors in \autocite{Santaca2016abf}.
 
@@
 \end{definition}
-While a risk matrix is often defined as a function of the 
-likelihood and impact of attacks (based on quantitative ad-hoc estimation
-of how likely it is that an attacker will exploit one or more vulnerabilities
-and what is the magnitude of the incidents produced by this exploitation), we suggest
-that cybersecurity weaknesses are all equally likely to be exploited
-if there's a connection between the weakness and the asset that
-an attacker wants to impact. When the whole system is considered an asset,
-all weaknesses are equally likely to be exploited.
-Therefore, a risk matrix should
-capture the number of insecure configurations of a system, rather
-than predicating over likelihoods of weaknesses exploitation.
+While a risk matrix is often defined as a function of the likelihood and impact
+of attacks (based on quantitative ad-hoc estimation of how likely it is that an
+attacker will exploit one or more vulnerabilities and what is the magnitude of
+the incidents produced by this exploitation), we suggest that cybersecurity
+weaknesses are all equally likely to be exploited if there's a connection
+between the weakness and the asset that an attacker wants to impact. When the
+whole system is considered an asset, all weaknesses are equally likely to be
+exploited.  Therefore, a risk matrix should capture the number of insecure
+configurations of a system, rather than predicating over likelihoods of
+weaknesses exploitation.
 
@@
 		granularity of the design. For example, a modeler can model a
-		specific device, as in Figure~\ref{fig:mashyp}. In this case,
-		the system is not decomposed into sub-systems or devices and
-		the device is considered as an agent.
+		specific device as a system not decomposed into sub-systems or devices (and
+		the device is considered as an agent).
 	\item Agents reason over beliefs (i.e. transform beliefs into other
@@
 	\begin{enumerate}
-		\item if $\eq{\assertionRegion}{\beliefRegion})$ then either
+		\item if $\eq{\assertionRegion}{\beliefRegion}$ then either
 			$\assertionRegion=\beliefRegion=\emptyset$ (no
@@
 			communication)
-		\item if $\dr{\assertionRegion}{\beliefRegion})$ then
+		\item if $\dr{\assertionRegion}{\beliefRegion}$ then
 			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
@@
 			which we call \emph{replace} (i.e. drop and insert)
-		\item if $\pp{\assertionRegion}{\beliefRegion})$ then
+		\item if $\pp{\assertionRegion}{\beliefRegion}$ then
 			$\beliefRegion$ contains and extend $\assertionRegion$
 			which we call \emph{insertion}
-		\item if $\ppi{\assertionRegion}{\beliefRegion})$ then
+		\item if $\ppi{\assertionRegion}{\beliefRegion}$ then
 			$\assertionRegion$ contains and extend $\beliefRegion$
 			which we call \emph{drop} (or selective drop to stress the difference with the full drop)
-		\item if $\po{\assertionRegion}{\beliefRegion})$ then a part of
+		\item if $\po{\assertionRegion}{\beliefRegion}$ then a part of
 			the $\assertionRegion$ is contained in the

### 2020-09-03 reports/paper_0/main.tex

@@
 	\item Delete: part of output-nonce is dropped. For example, the last bit of 1101 and 1100 is dropped, resulting in the two equal and non-fresh nonces: 110 and 110.
-	\item Inject: substitution at bit level results in non-fresh nonces. For example, 110 becomes 111 becomes equal by flipping either the last 0 of the former nonce or the last 1 of the latter.
+	\item Inject: substitution at bit level results in non-fresh nonces. For example, 110 and 111 becomes equal by flipping either the last 0 of the former or the last 1 of the latter.
 \end{itemize}
@@
 
-\subsection{Security and Insecurity of a System}
+\subsection{Security and Insecurity of a System}\label{sec:formula}
 \begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
@@
 a number of regions of assertions, beliefs, and facts. Those regions
-are connected by a generic relation which is evaluated as follows.
+are connected by a generic relation which is evaluated as follows (according to
+the formula in Section~\ref{sec:formula}).
 The graph is translated into 
@@
 which represent the physical reader that needs to be placed in a tank, and 
-the board that interprets the readings and outputs them as digital signals.
+the board that interprets the readings and outputs them as signals.
 The two components are connected by a wire.
@@
 We proposed an hypothesis for a foundational theory on security, arguing that
-cybersecurity-related issues are not related to the maliciousness of an agent but to
+cybersecurity-related issues are not linked to the maliciousness of an agent but to
 the vagueness of the cybersecurity controls on the engineering processes.

### 2020-09-03 prototypes/alpha_1/secra.py
bugfix - correct calculation of prod in excel

@@ -532,7 +532,8 @@ def write_report(path,spec_package,risk_structure,cyclic_risk_struct,components)
                 risk_sheet.write_formula(risk_id, 1, i)
                 risk_id+=1
                 relation_count+=1
-            risk_sheet.write(risk_id-1,2,"=SUM(B"+str(risk_id)+":B"+str(risk_id-(relation_count-1)))
+            sum_in_if_tmp="SUM(B"+str(risk_id-(relation_count-1))+":B"+str(risk_id)+")"
+            risk_sheet.write(risk_id-1,2,"=IF("+sum_in_if_tmp+"=0,\"\","+sum_in_if_tmp+")")
 
     cyclic_sheet_id=0
     for subgraph in cyclic_risk_struct.values():
@@ -543,7 +544,7 @@ def write_report(path,spec_package,risk_structure,cyclic_risk_struct,components)
             for rel in relations:
                 if(str(rel).lower() in relation2row.keys()):
                     #If status is mitigated 0, 1 oth.
-                    cyclic_struct_sheet.write(row,col,"=IF('"+weak_sheet_name+"'!H"+str(relation2row[str(rel).lower()])+"=\"mitigated\", 0, 1")
+                    cyclic_struct_sheet.write(row,col,"=IF('"+weak_sheet_name+"'!H"+str(relation2row[str(rel).lower()])+"=\"mitigated\", 0, 1)")
                 else:
                     cyclic_struct_sheet.write(row,col,1)
                 col+=1
@@ -558,7 +559,7 @@ def write_report(path,spec_package,risk_structure,cyclic_risk_struct,components)
     risk_id+=1
     risk_sheet.write(risk_id+1, 0, "RISK", cell_format['first_risk'])
     risk_sheet.write_formula(risk_id+1, 2, "=PRODUCT(C"+str(risk_id+1)+":C1)", cell_format['all_risk'])
-    risk_sheet.write(risk_id+2, 0, "The total risk is the total number of configurations of the system", cell_format['all_risk'])
+    risk_sheet.write(risk_id+2, 0, "The total risk is the total number of insecure configurations of the system", cell_format['all_risk'])
 
     workbook.close()
 

### 2020-09-03 reports/paper_0/main.tex
Implement Francesco+Livia' feedback

@@
 	requirements, functional architectures, and channels.  Ultimately, our
-	hypothesis allows the definition of the cybersecurity of a system as a
+	hypothesis allows for the definition of the cybersecurity of a system as a
 	mathematical formula.  We implemented a prototype cybersecurity risk
@@
 
-Currently, the most complete understanding of insecurity issues, is stored into
+Currently, the most complete understanding of insecurity issues is stored into
 a network of databases of weaknesses (e.g.
@@
 to provide tools for the testing or verification of systems, 
-no scientific falsifiable theory correlate
+no scientific falsifiable theory correlates
 cybersecurity requirements to necessary and sufficient conditions (e.g. mitigations) 
 to declare a system secure \autocite{Herley2016unfalsifiability}.
-Nonetheless, the extensive body of literature have scientific foundations,
+Nonetheless, the extensive body of literature has scientific foundations,
 for example, providing formal frameworks for the verification of 
@@
 defined in general (e.g. due to the heterogeneity of agents and systems) and
-that huge repository of dishonest behaviors should be kept as definition, such
+that a huge repository of dishonest behaviors should be kept as definition, such
 as CAPEC \autocite{MITRE2020CAPEC}; or that dishonesty of an agent with respect
@@
 \begin{enumerate}
-	\item \emph{System Specification}, where the \emph{functional and physical requirements} are defined
-	\item \emph{Architecture Design}, where the specification is structured into \emph{functional and physical architectures}
+	\item \emph{System Specification}, where the \emph{functional and physical requirements} are defined.
+	\item \emph{Architecture Design}, where the specification is structured into \emph{functional and physical architectures}.
 	\item \emph{Cybersecurity Risk Assessment}, where the potential \emph{weaknesses} (errors)
@@
 There is no agreement between the research communities (e.g.
-Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent
+Multi-Agent-System, Epistemic Logic) on what the constituent of an agent are,
 but the same ideas revolved around for thousands of years\footnote{
@@
 knowledge is defined, as in \autocite{Steup2020epistemology}, as a set of
-proposition known by an agent, such that: (i) knowledge requires belief, (ii)
-knowledge require truth, (iii) knowledge must be \emph{properly justified}, and
+propositions known by an agent, such that: (i) knowledge requires belief, (ii)
+knowledge requires truth, (iii) knowledge must be \emph{properly justified}, and
 the only objective of information is to exchange beliefs between agents.  As
@@
 probabilistic correlation to truth/reality, we consider information in relation
-with an agent's beliefs. Similarly, we consider beliefs to define the actual
+to an agent's beliefs. Similarly, we consider beliefs to define the actual
 behavior of an agent or a system; i.e. we consider behaviors as belief revision
@@
 		that information is from a sender to 
-		recipient; and we call it \emph{assertion}  
+		recipient; and we call it \emph{assertion}.
 	\item Similarly, the portion of \emph{beliefs} we consider for system
 		engineering is the one that builds (input) or describes
-		(output) the behavior of an agent 
+		(output) the behavior of an agent.
 		%(definitory rules
@@
 		instead of considering the more general epistemic definition of
-		knowledge. Specifically, facts describes: the functional
+		knowledge. Specifically, facts describe: the functional
 		architecture of each agent, and the physical/structural (HW/SW)
 		architecture of each subsystem of agents and agent within a
-		subsystem
+		subsystem.
 \end{enumerate}
@@
 Similarly to \autocite{Santaca2016abf}, we define agents as a meronomy (an
-hierarchy of Part-Whole relations) over the constituent that we previously defined
+hierarchy of Part-Whole relations) over the constituents that we previously defined
 (assertions, beliefs, and facts), based on a standard definition of mereology,
@@
 $X$ and $Y$. We note that all the RCC relations are symmetric with the exception
-of those that has an explicit (related) inverse.
+of those that have an explicit (related) inverse.
 
@@
 As in \autocite{Santaca2016abf}, it follows that, defining
-a system with a fixed number of regions, there exist
+a system with a fixed number of regions, there exists
 an upper-bound to the number of possible configuration of a system, defined by
@@
 \begin{itemize}%[nosep]
-\item \emph{RCC3} --- theoretical: $3^3=27$,  correct: 15 
-\item \emph{RCC5} --- theoretical: $5^3=125$, correct: 54
-\item \emph{RCC8} --- theoretical: $8^3=512$, correct: 193
+\item \emph{RCC3} --- theoretical: $3^3=27$,  correct: 15.
+\item \emph{RCC5} --- theoretical: $5^3=125$, correct: 54.
+\item \emph{RCC8} --- theoretical: $8^3=512$, correct: 193.
 \end{itemize}
@@
 implement or diverge from the intended nominal behavior. We note 
-that, the numbers reported here do not consider the details of the
+that the numbers reported here do not consider the details of the
 engineering process and should be considered a limit of an abstract 
@@
 and beliefs with facts determines the quality with respect to the nominal
-(specified) system.  Given that facts defines what needs to be true in the system,
-the relation of assertions an facts determines the degree of quality between
+(specified) system.  Given that facts define what needs to be true in the system,
+the relation of assertions and facts determines the degree of quality between
 the real information circulating in a system (or within an agent) and
@@
 	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
-		Control System (ICS) Security''
+		Control System (ICS) Security''.
 	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
@@
 		from concept phase through production, operation, service, and
-		decommissioning''
+		decommissioning''.
 \end{enumerate}
@@
 \begin{enumerate}[start=1, label={W\arabic*)}]
-	\item \emph{replace port}, where assertions reach
+	\item \emph{Replace port}, where assertions reach
 		the port but is replaced with different and un-related information
-		before passing the boundary 
-	\item \emph{drop port}, where assertions reach the
+		before passing the boundary.
+	\item \emph{Drop port}, where assertions reach the
 		port but do not pass the boundary of the agent (i.e. do not
-		become belief of the agent)
-	\item \emph{insertion port}, where some new information is transferred along with 
+		become belief of the agent).
+	\item \emph{Insertion port}, where some new information is transferred along with 
 		the information incoming from a channel, and then sent to
-		the intended recipient (agent)
-	\item \emph{injection port}, where part of the incoming information
-		is substituted with new information and transferred to the intended recipient
-	\item \emph{selective port}, where some information passes the port and
+		the intended recipient (agent).
+	\item \emph{Injection port}, where part of the incoming information
+		is substituted with new information and transferred to the intended recipient.
+	\item \emph{Selective port}, where some information passes the port and
 		part is either:
 	\begin{enumerate}[start=1, label={W5.\arabic*)}]
-		\item dropped, 
-		\item replaced.
+		\item Dropped, or 
+		\item Replaced.
 	\end{enumerate}
@@
 	\begin{enumerate}
-		\item if $\eq{\assertionRegion}{\beliefRegion}$ then either
+		\item If $\eq{\assertionRegion}{\beliefRegion}$ then either
 			$\assertionRegion=\beliefRegion=\emptyset$ (no
@@
 			$\assertionRegion=\beliefRegion\neq\emptyset$ (forward
-			communication)
-		\item if $\dr{\assertionRegion}{\beliefRegion}$ then
+			communication).
+		\item If $\dr{\assertionRegion}{\beliefRegion}$ then
 			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
@@
 			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
-			which we call \emph{replace} (i.e. drop and insert)
-		\item if $\pp{\assertionRegion}{\beliefRegion}$ then
+			which we call \emph{replace} (i.e. drop and insert).
+		\item If $\pp{\assertionRegion}{\beliefRegion}$ then
 			$\beliefRegion$ contains and extend $\assertionRegion$
-			which we call \emph{insertion}
-		\item if $\ppi{\assertionRegion}{\beliefRegion}$ then
+			which we call \emph{insertion}.
+		\item If $\ppi{\assertionRegion}{\beliefRegion}$ then
 			$\assertionRegion$ contains and extend $\beliefRegion$
-			which we call \emph{drop} (or selective drop to stress the difference with the full drop)
-		\item if $\po{\assertionRegion}{\beliefRegion}$ then a part of
+			which we call \emph{drop} (or selective drop to stress the difference with the full drop).
+		\item If $\po{\assertionRegion}{\beliefRegion}$ then a part of
 			the $\assertionRegion$ is contained in the
@@
 \begin{enumerate}[start=6, label={W\arabic*)}]
-	\item \emph{secure output port} and \emph{input drop port}
-	\item \emph{secure output port} and \emph{input insertion port}
-	\item \emph{output drop port} and \emph{input drop port}
-	\item \emph{output drop port} and \emph{input insertion port}
-	\item \emph{output drop port} and \emph{input secure port}
-	\item \emph{output injection port} and \emph{input secure port}
+	\item \emph{Secure output port} and \emph{input drop port}.
+	\item \emph{Secure output port} and \emph{input insertion port}.
+	\item \emph{Output drop port} and \emph{input drop port}.
+	\item \emph{Output drop port} and \emph{input insertion port}.
+	\item \emph{Output drop port} and \emph{input secure port}.
+	\item \emph{Output injection port} and \emph{input secure port}.
 \end{enumerate}
@@
 \begin{itemize}
-	\item \emph{Order-preserving} -- it shall be known if information is \emph{replaced}
-	\item \emph{Availability} -- it shall be known if information is \emph{dropped} or \emph{selectively dropped}
-	\item \emph{Integrity} -- it shall be known if information is \emph{injected}
-	\item \emph{Authentication} -- it shall be known if information is \emph{inserted}
+	\item \emph{Order-preserving} -- it shall be known if information is \emph{replaced}.
+	\item \emph{Availability} -- it shall be known if information is \emph{dropped} or \emph{selectively dropped}.
+	\item \emph{Integrity} -- it shall be known if information is \emph{injected}.
+	\item \emph{Authentication} -- it shall be known if information is \emph{inserted}.
 \end{itemize}

### 2020-09-04 prototypes/alpha_1/secra.py
minor changes

@@ -376,7 +376,7 @@ def write_report(path,spec_package,risk_structure,cyclic_risk_struct,components)
     #weak_semantics['channel']={ 'po':{'weakness':"selectively drops inputs and inserts new malicious data",'mitigation':"m1"}, 'pp':{'weakness':"forwards all the inputs but crafts and inserts new malicious data",'mitigation':"m2"}, 'ppi':{'weakness':"selectively drops inputs",'mitigation':"m3"}, 'dr':{'weakness':"drops all the inputs and inserts new malicious data",'mitigation':"m4"}, 'ppb0':{'weakness':"generates new outputs even when there's no incoming data from the socket",'mitigation':"m5"}, 'ppia0':{'weakness':"drops all the incoming data",'mitigation':"m6"} }
     weak_semantics['weak_port']={ 'po':{'weakness':"selectively drops inputs and inserts new malicious data",'mitigation':"m1"}, 'pp':{'weakness':"forwards all the inputs but crafts and inserts new malicious data",'mitigation':"m2"}, 'ppi':{'weakness':"selectively drops inputs",'mitigation':"m3"}, 'dr':{'weakness':"drops all the inputs and inserts new malicious data",'mitigation':"m4"} }
     weak_semantics['weak_channel']={ 'po':{'weakness':"selectively drops inputs and inserts new malicious data",'mitigation':"m1"}, 'pp':{'weakness':"forwards all the inputs but crafts and inserts new malicious data",'mitigation':"m2"}, 'ppi':{'weakness':"selectively drops inputs",'mitigation':"m3"}, 'dr':{'weakness':"drops all the inputs and inserts new malicious data",'mitigation':"m4"} }
-    weak_semantics['weak_out_block']={ 'po':{'weakness':"the component has a Byzantine behavior where occasionally outputs the expected output given the correct inputs. Not all the inputs are handled properly, nor all the expected outputs are always generated when correct inputs are given.",'mitigation':"m5"}, 'pp':{'weakness':"part of the expected outputs are not generated in response to the correct inputs",'mitigation':"m6"}, 'ppi':{'weakness':"the components correctly performs the expected behavior when the correct inputs are provided but is subject to input injections",'mitigation':"m7"}, 'dr':{'weakness':"the component never performs the expected behavior (e.g. physical damage)",'mitigation':"m8"} }
[…]

### 2020-09-04 reports/paper_0/main.tex

@@
 Since the $\abftheory$-framework is a theory of agents, we could consider ports as
-agents that allows the exchange of information between a channel and another
+agents that allow the exchange of information between a channel and another
 agent.  However, we considered a port as a special type of agent to avoid
-an \emph{infinite regress}, as describe in the following. While a channel transfers information between agents, and
+an \emph{infinite regress}, as described in the following. While a channel transfers information between agents, and
 a functional architecture elaborate information, a port is simply a connector between
@@
 	\item \emph{Replace port}, where assertions reach
-		the port but is replaced with different and un-related information
+		the port but are replaced with different and un-related information
 		before passing the boundary.
@@
 A functional architecture takes information as input-beliefs and transforms the
-information into output-beliefs. Those transformations occurs within the
-functional architecture, where functional blocks transforms beliefs into other
+information into output-beliefs. Those transformations occur within the
+functional architecture, where functional blocks transform beliefs into other
 beliefs. Similarly to channels, we could consider a functional block as a
@@
 	\item $\ppi{\beliefRegion}{\factRegion}$ the components
-	        correctly performs the expected behavior when the correct
+	        correctly perform the expected behavior when the correct
 	        inputs are provided but is subject to input
@@
 
-As an example, consider a functional block that performs the summation of two
+As an example, considering a functional block that performs the summation of two
 inputs, as in Figure~\ref{fig:sum}, defined by the requirement $r := b_3 = b_1
@@
 	specification over the physical and functional architectures, that
-	uniquely, i.e.  tested against the range of design possible in the
+	uniquely, i.e.  tested against the range of designs possible in the
 	$\abftheory$-framework, defines the design built on top of those
@@
 As depicted in Figure~\ref{fig:secra}, the cybersecurity risk assessment process
-starts with the definition of the use cases and architectural requirements; the
-specification of a system.  In our process, the specification is manually
+starts with the definition of the use cases and architectural requirements.  In our process, the specification is manually
 translated into a UML design where:
@@
 		architecture}. Each agent is defined as an UML node with (physical)
-		ports, and agent's ports are connected via UML information flow
+		ports, and agent' ports are connected via UML information flow
 		connectors, representing the physical channel.
@@
 We proposed an hypothesis for a foundational theory on security, arguing that
-cybersecurity-related issues are not linked to the maliciousness of an agent but to
-the vagueness of the cybersecurity controls on the engineering processes.
-We also provided a prototype tool for the quantitative estimation of the cybersecurity
-risk, based on a UML model of the system under assessment.
+cybersecurity-related issues are not linked to the maliciousness of an agent
+but to the vagueness in the design processes.  We also provided a prototype
+tool for the quantitative estimation of the cybersecurity risk, based on a UML
+model of the system under assessment.
 %The current state-of-the-art processes allow, given a specification,

### 2021-01-18 reports/paper_0/main.tex
add names to paper
shorten intro

@@
 \author{
-%	\IEEEauthorblockN{Marco Rocchetto}
-%\IEEEauthorblockA{V-Research\\
-%Verona, Italy\\
-%Email: marco@v-research.it}
-%\and
-%\IEEEauthorblockN{Francesco Beltramini}
-%\IEEEauthorblockA{V-Research\\
-%Verona, Italy\\
-%Email: francesco@v-research.it}
+	\IEEEauthorblockN{Marco Rocchetto}
+\IEEEauthorblockA{V-Research\\
+Verona, Italy\\
+Email: marco@v-research.it}
+\and
+\IEEEauthorblockN{Francesco Beltramini}
+\IEEEauthorblockA{V-Research\\
+Verona, Italy\\
+Email: francesco@v-research.it}
 }
@@
 %from a non-falsifiable theory of system cybersecurity.  In other words, 
-As
-Herley mentions in \autocite{Herley2016usenixvideo}, the problem is in the way
-we reason on cybersecurity, and we can see it in the way we reply to simple
-cybersecurity (foundamental) questions.  Let us consider the general question
-``what is cybersecurity?''; as in \autocite{hintikka2013inquiry}, we can
-translate the wh-question into an equivalent existentially quantified formula.
-But if the formula is $\exists P. P(s)\rightarrow Secure(s)$, (where we ask if
-there exists a predicate $P$ such that, given any arbitrary system $s$, when
-$P$ holds for $s$ then $s$ can be considered secure), when the answer is (for
-example) ``if the system is confidential, the system is secure'' the formula
-becomes tautological. In fact, $P$ and $Secure$ can be considered as
-semantically equivalent (cybersecurity is confidentiality and vice-versa).
+%As
+%Herley mentions in \autocite{Herley2016usenixvideo}, the problem is in the way
+%we reason on cybersecurity, and we can see it in the way we reply to simple
+%cybersecurity (foundamental) questions.  Let us consider the general question
+%``what is cybersecurity?''; 
+%%as in \autocite{hintikka2013inquiry}, we can
+%%translate the wh-question into an equivalent existentially quantified formula.
+%%But if the formula is $\exists P. P(s)\rightarrow Secure(s)$, 
+%%(where we ask if
+%if we detail the question as:
+%there exists a predicate $P$ such that, given any arbitrary system $s$, when
+%$P$ holds for $s$ then $s$ can be considered secure?
+%%), 
+%When the answer is (for
+%example) ``if the system is confidential, the system is secure'' the reasoning %formula
+%becomes tautological. In fact, $P$ and $Secure$ can be considered as
+%semantically equivalent (cybersecurity is confidentiality and vice-versa).
 

### 2021-03-01 reports/paper_0/main.tex
change authors and email in paper_0
fix #23 in paper_0

@@
 \usepackage{rotating}
-
-% Space hacks, ...yes, from outer space!
-%\addtolength{\textfloatsep}{-6mm}
-%\addtolength{\abovecaptionskip}{-3mm}
-%\addtolength{\columnsep}{-1mm}
-%\setlength{\evensidemargin}{0.1in}
-%\setlength{\oddsidemargin}{0.1in}
-%\setlength{\textwidth}{6.47in}
-%\setlength{\textwidth}{5.1in}
 
@@
 \author{
-	\IEEEauthorblockN{Marco Rocchetto}
+	\IEEEauthorblockN{The Science Club}
 \IEEEauthorblockA{V-Research\\
 Verona, Italy\\
-Email: marco@v-research.it}
-\and
-\IEEEauthorblockN{Francesco Beltramini}
-\IEEEauthorblockA{V-Research\\
-Verona, Italy\\
-Email: francesco@v-research.it}
+Email: knowledgezero@v-research.it}
 }
@@
 The internal structures of the tool
-can be viewed as PDF and the results are reported into an Excel file.
-The Excel file also reports the total number of configurations as indicating
+can be viewed as PDF and the results are reported into an spreadsheet file.
+The spreadsheet file also reports the total number of configurations as indicating
 the cybersecurity risk associated to the specification. A user can change the status
-of each weakness in the Excel file, from the initial (default) status (open)
+of each weakness in the spreadsheet file, from the initial (default) status (open)
 to ``mitigated''. As a consequence, the risk is re-calculated on-the-fly, i.e.
 without the need of running the tool again, based on annotations and formulas 
-in the Excel file.
+in the spreadsheet file.
 In our approach, cybersecurity requirements are not imposed by the
@@
 	\includegraphics[width=.7\textwidth]{results_excel.pdf}
-	\caption{Partial View of the results in the Excel file}
+	\caption{Partial View of the results in the spreadsheet file}
 	\label{fig:results}

### 2021-03-01 reports/paper_0/main.tex
manual merge with oliviero's branch

@@
 
-\newcommand{\fixn}[2]{{\bf FIX}\footnote{#1: #2}}
-
 \newtheorem{definition}{Definition}%[section]
@@
 	The objective of this research is to lay the foundations for the
-	development of a scientific theory that determines (all and only) the possible
-	insecure and secure configurations of any abstract system. We claim
-	that cybersecurity weaknesses (i.e. errors) are at the beginning of the
-	causality chain that leads to cybersecurity attacks.  We, then,
+	development of a scientific theory that determines (all and only) the
+	possible insecure and secure configurations of any abstract system. We
+	claim that cybersecurity weaknesses (i.e. errors) are at the beginning
+	of the causality chain that leads to cybersecurity attacks.  We, then,
 	formulate a scientific (falsifiable by experiments) hypothesis that we
@@
 	based on a correlation between the epistemological concepts of
-	knowledge, beliefs, and information with the engineering concepts of
+	facts, beliefs, and assertions with the engineering concepts of
 	requirements, functional architectures, and channels.  Ultimately, our
-	hypothesis allows for the definition of the cybersecurity of a system as a
-	mathematical formula.  We implemented a prototype cybersecurity risk
-	assessment tool that, based on our hypothesis, predicts the weaknesses in a
-	UML model of a (cyber-physical) system.
+	hypothesis allows for the definition of a mathematical formula which
+	describes the cybersecurity of a system.  We implemented a prototype
+	cybersecurity risk assessment tool that, based on our hypothesis,
+	predicts the weaknesses in a UML model of a (cyber-physical) system.
 \end{abstract}
@@
 \IEEEpeerreviewmaketitle
+
+\fixnote{Oliviero}{Two general remarks.
+First, I'd maybe specify more what concretely A, B and F "look like", if that makes sense.
+How do we concretely specify a fact in the ABF framework?
+If I am not mistaken, RCC defines axiomatically how in general a mereotopology should be, and a *concrete* mereotopology corresponds to a particular system specification.
+Is this correct?
+If so, make explicit how one would encode this, if this make sense.
+At the very end you talk about a particular configuration being SAT w.r.t.
+the algebraic strucutre; is this algebraic structure, indeed, the system specification?
+You also sometimes speak about logic formulas; what is the connection between these logical formulas and regions?
+Is a region a "set" of logical formulas, or is it that a region is logically defined?
+
+Secondly, and this is probably due to my inexperience with systems design (and ABF), it is not totally clear to me exactly what a belief is; more concretely, what is the difference between assertions and beliefs.
+Beliefs define the inputs and the outputs; assertions are what is exchanged between agents.
+Wouldn't the assertions already define the inputs and outputs?
+or are the beliefs the "expected" I/Os, while assertions are the "actual" I/Os?
+Indeed, a functional block is defined by its inputs/outputs.
+Then, if EQ(F,A) hold, aren't we already happy?
+Because, if all the inputs and outputs are compliant with your requirements, your system is functionally "correct", and behaves as you want.
+Am I mistaken?
+Perhaps there are some other aspects that I am not considering.}
+
+\fixnote{Oliviero}{Question: the region F encodes the requirements.
+Then, what if we encode a requirement wrongly?
+or, what if we do not realise that some implict requirements are missing?
+Indeed, one might not see why integrity is a requirement until a breach happens.
+Is this within the scope of the theory?
+Finally; are we sure that all reasonable requirements can be encoded in the region F?}
 
@@
 Similarly, the results of the application of a scientific theory (i.e. its
-predictions) can be verified in the abstract mathematical structure but how can
+predictions) can be verified in the abstract mathematical structure, but how can
 we be sure that the theory correctly characterizes the phenomenon it describes? 
@@
 predict the cybersecurity of a system, nor a theory that can predict all
-possible insecurities of a system (which, by negation may be used as a theory
+possible insecurities of a system (which, by negation, may be used as a theory
 of cybersecurity).  Herley then uses this argument to show that ``claims that
@@
 implemented is a system in which a password is masked. No experiment can ever
-falsify this statement.  Given that any theory that is not falsifiable by an
+falsify this statement\footnote{In other words, one cannot design an experiment to show that
+password masking is not secure or unnecessary, and a correlated problem is that
+this may lead to a scenario where more measures are added but there is no good
+way of saying which measures are superflous.}. Given that any theory that is not falsifiable by an
 empirical experiment is well known\footnote{``A theory which is not refutable
@@
 With our approach, a list of weaknesses emerges from the mathematical
-formulation of a system in a framework called $\abftheory$-framework and
+formulation of a system in a framework called $\abftheory$-framework, and
 predicts 4 main classes of weaknesses. Those classes of weaknesses are used to
@@
 and security are not syntactically differentiated and both terms (safety and
-security) are expressed by the same word, e.g. sicurezza in Italian.  A
+security) are expressed by the same word, e.g. ``sicurezza'' in Italian.  A
 semantic distinction between safety and (cyber)security is correlated to a
 belief\footnote{A belief has to be intended as a proposition which is supposed
-to be true by the majority of people in our society, without a scientific
-underlying theory but based on partial empirical evidences or inductive reasoning based on partial empirical evidences.} that
+to be true but no conclusive evidence makes it known to be true.} that
 safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
@@
 reported today are simply the realization of potential system configurations,
-which deviate from the nominal behavior because the \emph{intended (or
-nominal) behavior} of the system is not precisely defined in the specification
+which deviate from the  \emph{intended (or
+nominal) behavior} because the nominal behavior of the system is not precisely defined in the specification
 at the very early stages of the engineering process.
@@
 	\item In \autocite{Hintikka1962knowledge}, Hintikka describes the
-		difference between knowledge and belief (as epistemological
-		concepts), and the whole Doxastic logic defines in details how
+		belief (as epistemological
+		concepts) and their relation with knowledge,
+		and the whole Doxastic logic defines in details how
 		beliefs can be formalized.
 	\item In \autocite{Hintikka1993Information}, Hintikka describes the concept
-		of information and the difference with knowledge and belief.
+		of information (and the difference with knowledge and belief).
 	\item In \autocite{Santaca2016abf}, the authors defines an agent as a
-		tuple of assertions, beliefs, and facts.
+		tuple of assertions, beliefs, and facts. A detailed discussion
+		on facts and assertions can be found in \autocite{facts2007stanford} and
+		\autocite{assertion2007stanford} respectively.
 \end{itemize}
 For our argument, as in \autocite{Santaca2016abf}, an agent is composed by its
-knowledge, beliefs, and the information or assertions it provides; where
-knowledge is defined, as in \autocite{Steup2020epistemology}, as a set of
-propositions known by an agent, such that: (i) knowledge requires belief, (ii)
-knowledge requires truth, (iii) knowledge must be \emph{properly justified}, and
-the only objective of information is to exchange beliefs between agents.  As
-defined in \autocite{Hintikka1993Information}, ``information is specified by
-specifying which alternatives concerning the reality it admits and which
-alternatives excludes''. This means that if we consider a propositional
-variable (which admits the two alternatives True/False) its information is
-defined as believed to be True/False and not believed to be the opposite.  Due
-to the definition of information and then with its relation to the
-probabilistic correlation to truth/reality, we consider information in relation
-to an agent's beliefs. Similarly, we consider beliefs to define the actual
-behavior of an agent or a system; i.e. we consider behaviors as belief revision
-systems.
-The difference between knowledge and information as used by Hintikka, and facts
-and assertions as used in \autocite{Santaca2016abf}, are not precisely discussed
-in \autocite{Santaca2016abf}.  Furthermore, those epistemological concepts are
-difficult to define
-\autocite{Gettier2012knowledge}.
+beliefs, and the information or assertions it provides.
+%knowledge, beliefs, and the information or assertions it provides; where
+%knowledge is defined, as in \autocite{Steup2020epistemology}, as a set of
+%propositions known by an agent, such that: (i) knowledge requires belief, (ii)
+%knowledge requires truth, (iii) knowledge must be \emph{properly justified}, and
+%the only objective of information is to exchange beliefs between agents.
+As in \autocite{Steup2020epistemology}, the only objective of information is to
+exchange beliefs between agents and, as defined in \autocite{Hintikka1993Information},
+``information is specified by specifying which alternatives concerning the
+reality it admits and which alternatives excludes''. This means that if we
+consider a propositional variable (which admits the two alternatives
+True/False) its information is defined as believed to be True/False and not
+believed to be the opposite.  Due to the definition of information and then
+with its relation to the probabilistic correlation to truth/reality, we
+consider information in relation to an agent's beliefs. Similarly, we consider
+beliefs to define the actual behavior of an agent or a system; i.e. we consider
+behaviors as belief revision systems.
+%The difference between knowledge and information as used by Hintikka, and facts
+The concepts of facts and assertions as used in \autocite{Santaca2016abf}, are
+not precisely discussed in \autocite{Santaca2016abf}.  Furthermore, those
+epistemological concepts are difficult to define
+\autocite{Gettier2012knowledge} in general.
 %\footnote{``\emph{Theaetetus}: [\ldots] He said
@@
 of the scope of this paper. 
-Instead, we now define those concepts abstractly, inspired by the
+Instead, we now define those concepts informally, inspired by the
 $\abftheory$-framework  defined in \autocite{Santaca2016abf} and, in
-Section~\ref{sec:theory} we detail them as engineering concepts.
+Section~\ref{sec:theory} we detail them, and properly define them, as engineering concepts.
+\fixnote{Oliviero}{Is that actually what is happening?
+During this section III, we refer multiple times to engineering concepts such as system, functional specification and what not.
+We sometimes refer to the relationship between functional architecture and beliefs but that is not made precise.
+Maybe we should use only abstract terms, until the connection with those engineering aspects are introduced thoroughly?
+{\bf mr}: what I wanted to say here (and I probably failed) is that we define A, B, and F informally (and I changed abstractly with informally in the text) but we'll define them more precisely (and as engineering concept) later on.
+I modified the sentence to make this clearer but it seems to me to me that you're saying that the text is unnecessarily complicated by the mixed used of generic terms as beliefs and functional blocks/architecture.
+If that is the case we should definitely make the section clearer.}
 \begin{enumerate}
@@
 \end{enumerate}
+
+\fixnote{Oliviero}{There are some parts I am perplexed about about this section from "There is no agreement..." up to here).
+It is unclear to me why we present and define the general concept of knowledge rather then the concept of fact, since we use mostly "facts" and as defined in [41] an agent is cmposed by assertions, beliefs and facts (and not knowledge).
+Indeed, it seems from your last bullet point that a fact is a particular case of knowledge (as you say, "more general"), but the exact definition of fact is not really clear from here, at least to me.
+Further; why are facts requirements?
+Intuitively, I would define a fact to be something that is true; requirements are something that we *want* to be true.
+I understood why we do so later on in the paper; we are later on interested, by RCC, in whether the beliefs of our agents are consistent (EQ) with the facts, that is, we want our system to behave consistently with the requirements.
+I would maybe make this explicit: we interpret requirements as facts, since we consider them to be the truth, towards which our agents must be compliant with (rather than: facts are requirements and not the actual truth).
+Another thing I found to be unclear is your definition of belief.
+You start by saying that we will define these things abstractly but you refer here to a very system-oriented definition of belief: inputs and outputs.
+Indeed, you say that beliefs define the behaviour, but the exact definition of belief was not super clear to me.
+{\bf mr}: I think you're right and the key points are (i) what's a fact?
+(ii) why stressing knowledge if abf is based on facts?
+(iii) what's the relation between facts and requirements?
+A fact is a statement which is taken as true.
+it's not necessarily true in general but the discussion is much complex (unfortunately).
+I wanted to add it to the paper but someone said that it may be tricky since those concepts have been used for 2k years at least.
+But, as an example, ``Socrates is a man'' is a fact (i.e.
+a true statement or, better, a statement taken as true), right?
+do we \emph{know} it?
+how?
+did we induce it from empirical evidences by testing every single man ever existed?
+I guess you see what I mean.
+So, in my limited view, facts are statement taken as true, knowledge can be expressed by true statements, and requirements are statement taken as true (by developers and everyone else).
+The link between those concepts seems rather strong but I think you are suggesting to simplify the text changing the focus on facts or detailing the relation between knowledge and facts.
+I chose the former and changed the text accordingly.}
+
 We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}.
@@
 
+\fixnote{Oliviero}{Further: what is a region?
+Is it a "prime" concept, that is, a primitive concept?
+Or is it a set?
+if it is a primitive concept, then defining connectedness in terms of set operations such as $\subseteq$ is a little counterintuitive; if a region is a set (or some other well-known mathematical concept), then state so explicitly somewhere, imho (as I said at the beginning, explicitly describing how these ABF regions look like would make this much clearer, I think).
+{\bf mr} you're right.
+it's a primitive concept, I changed it in the tab and in the text.}
+
 \subsection{Mereo-topological Reasoning}
@@
 \hline
-&&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& $\mathit{X}\subseteq \mathit{Y}$ \\
-&&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& $\mathit{X}\not\subseteq \mathit{Y}$\\
+&&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& reflexive and symmetric \\
+&&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& irreflexive or antisymmetric\\
 &&&Part of				& $\mathit{P}(\mathit{X},\mathit{Y})$		& $\forall \mathit{Z} ~\mathit{C}(\mathit{Z},\mathit{X}) \rightarrow \mathit{C}(\mathit{Z},\mathit{Y})$\\
 &&&Overlaps			& $\mathit{O}(\mathit{X},\mathit{Y})$		& $\exists \mathit{Z} ~\mathit{P}(\mathit{Z},\mathit{X})\wedge \mathit{P}(\mathit{Z},\mathit{Y})$\\
+\Tdot&\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
 \Tdot&&&  \textbf{Overlaps Not Equal} 	& $\mathit{ONE}(\mathit{X},\mathit{Y})$		& $\mathit{O}(\mathit{X},\mathit{Y}) \land \neg \mathit{EQ}(\mathit{X},\mathit{Y})$ \\
-\Tdot&\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
 \Tdot&\Tdot&\Tdot& \textbf{DiscRete from} 		& $\mathit{DR}(\mathit{X},\mathit{Y})$		& $\neg \mathit{O}(\mathit{X},\mathit{Y})$\\
@@
 of the cybersecurity of each entry. Instead, we want a precise calculation
-of the number of insecure configurations to emerge from the mathematical
+of the number of insecure configurations
+\fixnote{Oliviero}{is this the same as a state?
+If so, maybe mention this in brackets in the definition.
+{\bf mr}: not necessarily but it will be in the next prototype and paper.
+For now, I'd keep it easy and general (plus, we'd need to introduce the concept of state oth.)} of a system to emerge from the mathematical
 formulation of our cybersecurity hypothesis.
@@
 The general formula to calculate the number of different types of agents is
-$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
+$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$\fixnote{Oliviero}{What if our calculus uses relations of different arities?
+I guess we sum this up for all arities?
+{\bf mr}: I think so, and I see your point (we could generalize the formula) maybe it's not that important?
+not sure},
 between $n$ different regions. Therefore, $r^{\binom{n}{k}}$
@@
 
-We categorize agents by first analyzing the relations between each pair of
-regions defining an agent (i.e.  $\abf$), and then we categorize the different
+We first analyze the relations between each pair of
+regions (i.e.  $\abf$), and then we categorize the different
 agents as tuple of the three regions.  For the sake of simplicity, soundness is
@@
 consider that, by definition, assertions are defined as transfer of information
-between two agents, e.g., $a$ and $b$.  Therefore, as depicted in
+between two agents, e.g., between two agents.  Therefore, as depicted in
 Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
 input and output assertions.  Given an agent $a$ and a collection of asserted
-predicates $\Phi$, the input assertions are those received by $a$ from an agent
+predicates $\Phi$
+\fixnote{Oliviero}{If I am not mistaken, this is the first time we refer to assertions as concrete formulas/predicates.
+Perhaps we should make explicit before the "form" of such assertions?
+{\bf mr}: I'm not sure I got this point.
+Do you mean we should refer to a logic (e.g.
+FOL)?
+If so, we could but it has a cost (in terms of simplicity) and I'm not sure how to reuse that in the paper to justify that ``cost''},
+the input assertions are those received by $a$ from an agent
 $s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
-are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
+are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. With a slight abuse
+of notation, in the text we simply write $\assertionRegion_{s\rightarrow a}$ when we abstract
+away from the content of the assertion (i.e. $\Phi$).
+We shall consider
 two pairs of regions:%\footnote{``I am not asserting, as Lotze did, that a
@@
 		defined as $\eq{\assertionRegion_a}{\beliefRegion_a}$ for an
-		agent $a$ while can be detailed as follows when multiple agents
+		agent $a$, and can be detailed as follows when multiple agents
 		are considered.
@@
 \begin{definition}{\bf Risk --}
-A risk is determined by the uncertainty related to the whole space of potential designs of a specification with respect to the
-	$\abftheory$-framework.
-\end{definition}
+	\fix{mr}{A risk is determined by the uncertainty related to the whole space of potential designs of a specification with respect to the
+	$\abftheory$-framework.}
+\end{definition}\fixnote{Oliviero}{iirc, this is the first time we speak about uncertainity.
+What do you mean exactly?
+Do you mean that we are uncertain about which state our CPS is in?
+Is it not knowing that we are in the secure state?}
 The definition of Risk leads to the risk matrix in
@@
 	(RCC5 in our case).
-\end{definition}
+\end{definition}\fixnote{Oliviero}{So, the granularity is just uniformly distributed between all possible configurations, right?
+Or is it the actual numeric value unimportant, as long as it "ordered" in the right way?
+{\bf mr}: right, the former}
 While a risk matrix is often defined as a function of the likelihood and impact
@@
 to MAS. We now map MAS to CPS and show how this mapping allows us
-to predict cybersecurity weaknesses in a system.
+to predict cybersecurity weaknesses in a system.\fixnote{Oliviero}{again, to me we actually referred quite a lot to systems and architectures.}
 
@@
 		or device) is composed by a \emph{functional architecture} that
-		transforms input-beliefs into output-beliefs.
+		transforms input-beliefs into output-beliefs.\fixnote{Oliviero}{We often times used functional architectures before.
+Mabybe we should make this connection precise before, then?
+Or maybe avoid using this CPS-oriented terminology before, and keep it more abstract?}
 	\item Components of a CPS have \emph{ports} to exchange information
@@
 		defined by \emph{channels}.
-	\item The concept of knowledge is related to the \emph{requirements}
+	\item The concept of facts is related to the \emph{requirements}
 		that describe how the CPS shall behave and
@@
 	transforms assertions from a sender $s$ ($\rassert{s}{a}$) to beliefs
-	of an agent $a$ ($\beliefRegion_a$), while an output-port transforms belief
-	into assertion.
+	of an agent $a$ ($\beliefRegion_a$), while an output-port transforms beliefs
+	into assertions.
 \end{definition}
@@
 
-%\begin{definition}{\bf Secure Input Port \fixnote{mv}{chiarirei meglio intuitivamente, magari fuori della definizione, quale sia la differenza tra port e secure port.
-La differenza è in quell'always?}--}\label{def:secport}
-%	A secure input port always allows information as incoming assertions to
-%	flow from a sender $s$ to the beliefs (generated by the behavior) of
-%	the recipient $a$ (agent).  
-%	%If $p_I$ is a secure input port,
-%	%$\world\models\varphi$, $\varphi\in\rassert{s}{a}$, and $a$ has the
-%	%only input port $p_I$, $\world'\models\beliefRegion_a$ and
-%	%$\varphi\in\beliefRegion_a$ for any $\world\modalrelation\world'$.
-%\end{definition}
-
 \paragraph{Port Weaknesses} We now apply the $\abftheory$-framework to
@@
 types of weaknesses, generating six types of insecure port in RCC5:
-%(the notation is reported for input-ports, i.e. on the left-hand side of the arrow):
 \begin{enumerate}[start=1, label={W\arabic*)}]
@@
 	$X$; where a base is a collection of regions in a (mereo)topology such that
-	every open region can be written as union of elements of that base. 
-	We distinguish between regions that contain information and regions that
-	don't by writing the latter as $\emptyset$.
+	%every open region\fixnote{Oliviero}{First time we refer to open regions; perhaps briefly introduce them?} can be written as union of elements of that base.
+	every region can be written as union of elements of that base. 
+	We distinguish between regions that are related to information and regions that
+	don't by writing the latter as $\emptyset$ (e.g. a region corresponding to
+	an assertion is written $A\neq\emptyset$, while a region corresponding to an empty assertion
+	as $A=\emptyset$).
 	\begin{enumerate}
@@
 output-port, there is no insecure behavior but only the combination of the
-weaknesses of the input and output port; therefore there exist $(7^2)-2=47$
+weaknesses of the input and output port; therefore there exist $(7^2)-1=48$
 theoretical configurations ($7^2$ because there are $6$ insecure types of port,
 plus $1$ secure type, on both input and output side; and we exclude the
-configuration with $2$ secure types as input and output, $-2$); where only 43
+configuration with $2$ secure types as input and output, $-1$); where only 44
 are possible. For the sake of readability, we report 6 examples in the
@@
 	interconnected system of functional blocks.
-\end{definition}
+\end{definition}\fixnote{Oliviero}{We referred to this notion many times in the esposition, and define it precisely only here. I'd perhaps speak more abstractly until this is introduced.}
 The quality of a functional block cannot be determined
@@
 	\item Replace: the output-nonces of the block are replaced with
-		non-fresh numbers
+		non-fresh numbers.
 	\item Insert: the output-nonces are concatenated with other numbers,
@@
 		end respectively, resulting in the two identical (and the
-		non-fresh) nonces 1110 and 1110
+		non-fresh) nonces 1110 and 1110.
 	\item Delete: part of output-nonce is dropped. For example, the last bit of 1101 and 1100 is dropped, resulting in the two equal and non-fresh nonces: 110 and 110.
@@
 
+\fixnote{Oliviero}{Ok, maybe here my inexperience with systems design is showing, but these hypothesis are not super clear to me.
+I understand the next part quite well, but the actual hypthesis, not so much.
+What are the designs possible with respect to the framework?
+Do you mean all the possible configurations of <rcc(A,B),rcc(A,F),rcc(B,F)> for each agent?
+What does "testing against" this range of possible designs?}
+
 Based on these hypothesis, we can formulate the concepts of security and
@@
 flows) between functional blocks. We define $R\subseteq V\times F$, where $F$
-is the set of all the requirements of $S$, and extend $G$ as $G'=\langle V',E'\rangle$ with $E'= E\cup R$ and $V'=V\cup F$. Let $\pi:
+is the set of all the requirements of $S$,
+\fixnote{Oliviero}{I assume that $(v,f)\in R$ iff f is a requirement over v?
+Is it possible that this is unknown, i.e., we do not know which devices a requirement predicates over?
+{\bf mr}: I guess it is possible but I believe we may not consider this case here.
+If there's a functional block $v$ there must be a requirement related to it.
+It may be wrong and we may not precisely know it, but it is there.}
+and extend $G$ as $G'=\langle V',E'\rangle$ with $E'= E\cup R$ and $V'=V\cup F$. Let $\pi:
 E'\rightarrow RCC$ (where $RCC$ is the set of relations in the RCC) be the
@@
 configuration is satisfiable with respect to the logical theory defining the
-algebraic structure (mereotopology) and constraints of the calculus RCC
+algebraic structure (mereotopology)\fixnote{Oliviero}{Which structure are we talking about, here?
+Is it the system specification?
+See my previous comments on specifying what A,B, and F actually look like.} and constraints of the calculus RCC
 (otherwise $\sigma$ returns $0$), we can define
 \begin{displaymath}
-	I=\sum_{p\in\{\Pi\setminus\pi_{eq}\}} \sigma(p)
+	I=\sum_{p\in(\Pi\setminus\pi_{eq})} \sigma(p)
 \end{displaymath}

### 2021-03-01 reports/paper_0/main.tex
fix all oliviero's fixes

@@
 \IEEEpeerreviewmaketitle
-
-\fixnote{Oliviero}{Two general remarks.
-First, I'd maybe specify more what concretely A, B and F "look like", if that makes sense.
-How do we concretely specify a fact in the ABF framework?
-If I am not mistaken, RCC defines axiomatically how in general a mereotopology should be, and a *concrete* mereotopology corresponds to a particular system specification.
-Is this correct?
-If so, make explicit how one would encode this, if this make sense.
-At the very end you talk about a particular configuration being SAT w.r.t.
-the algebraic strucutre; is this algebraic structure, indeed, the system specification?
-You also sometimes speak about logic formulas; what is the connection between these logical formulas and regions?
-Is a region a "set" of logical formulas, or is it that a region is logically defined?
-
-Secondly, and this is probably due to my inexperience with systems design (and ABF), it is not totally clear to me exactly what a belief is; more concretely, what is the difference between assertions and beliefs.
-Beliefs define the inputs and the outputs; assertions are what is exchanged between agents.
-Wouldn't the assertions already define the inputs and outputs?
-or are the beliefs the "expected" I/Os, while assertions are the "actual" I/Os?
-Indeed, a functional block is defined by its inputs/outputs.
-Then, if EQ(F,A) hold, aren't we already happy?
-Because, if all the inputs and outputs are compliant with your requirements, your system is functionally "correct", and behaves as you want.
-Am I mistaken?
-Perhaps there are some other aspects that I am not considering.}
-
-\fixnote{Oliviero}{Question: the region F encodes the requirements.
-Then, what if we encode a requirement wrongly?
-or, what if we do not realise that some implict requirements are missing?
-Indeed, one might not see why integrity is a requirement until a breach happens.
-Is this within the scope of the theory?
-Finally; are we sure that all reasonable requirements can be encoded in the region F?}
 
@@
 	\item In \autocite{Hintikka1962knowledge}, Hintikka describes the
-		belief (as epistemological
+		concept of belief (as epistemological
 		concepts) and their relation with knowledge,
@@
 For our argument, as in \autocite{Santaca2016abf}, an agent is composed by its
-beliefs, and the information or assertions it provides.
+\emph{beliefs} (internal information considered true by the agent), and
+%and the information or 
+\emph{assertions} (an exchange of information between agents). % it provides.
 %knowledge, beliefs, and the information or assertions it provides; where
@@
 %the only objective of information is to exchange beliefs between agents.
-As in \autocite{Steup2020epistemology}, the only objective of information is to
-exchange beliefs between agents and, as defined in \autocite{Hintikka1993Information},
+%As in \autocite{Steup2020epistemology}, 
+%the only objective of information is to
+%exchange beliefs between agents and, 
+As defined in \autocite{Hintikka1993Information},
 ``information is specified by specifying which alternatives concerning the
@@
 consider information in relation to an agent's beliefs. Similarly, we consider
-beliefs to define the actual behavior of an agent or a system; i.e. we consider
-behaviors as belief revision systems.
+beliefs to define the actual behavior of an agent or a system; meaning
+that an agent has its own beliefs on top of which it bases its behavior,
+the consequences of which, in turn, will update its initial beliefs.
+%; i.e. we consider
+%behaviors as belief revision systems.
 %The difference between knowledge and information as used by Hintikka, and facts
@@
 Section~\ref{sec:theory} we detail them, and properly define them, as engineering concepts.
-\fixnote{Oliviero}{Is that actually what is happening?
-During this section III, we refer multiple times to engineering concepts such as system, functional specification and what not.
-We sometimes refer to the relationship between functional architecture and beliefs but that is not made precise.
-Maybe we should use only abstract terms, until the connection with those engineering aspects are introduced thoroughly?
-{\bf mr}: what I wanted to say here (and I probably failed) is that we define A, B, and F informally (and I changed abstractly with informally in the text) but we'll define them more precisely (and as engineering concept) later on.
-I modified the sentence to make this clearer but it seems to me to me that you're saying that the text is unnecessarily complicated by the mixed used of generic terms as beliefs and functional blocks/architecture.
-If that is the case we should definitely make the section clearer.}
 \begin{enumerate}
@@
 		%rules}'' -- Hintikka in \autocite{Hintikka1993Information}},
-	\item We consider a set of axiomatic \emph{facts} 
-		(definitory
-		rules\footnote{Facts are definitory rule as they don't
-		define how a real system is finally implemented but
-		how it should be, trough a series of requirements. Those requirements 
-		may not hold, through insecurity, in the implemented system.}\autocite{Hintikka1993Information})
-		instead of considering the more general epistemic definition of
-		knowledge. Specifically, facts describe: the functional
-		architecture of each agent, and the physical/structural (HW/SW)
-		architecture of each subsystem of agents and agent within a
-		subsystem.
+	\item We consider a set of axiomatic \emph{facts} determining how the
+		system shall be. Similarly, requirements determine which facts
+		must hold when the system is implemented\footnote{We consider facts as
+		definitory rules\autocite{Hintikka1993Information} since they
+		don't define how a real system is finally implemented but how
+		it shall be, trough a series of requirements.  Those
+		requirements may not hold, through insecurity, in the
+		implemented system.}. Specifically, facts/requirements describe:
+		the functional architecture of each agent, and the
+		physical/structural (HW/SW) architecture of each subsystem of
+		agents and agent within a subsystem.
 \end{enumerate}
-
-\fixnote{Oliviero}{There are some parts I am perplexed about about this section from "There is no agreement..." up to here).
-It is unclear to me why we present and define the general concept of knowledge rather then the concept of fact, since we use mostly "facts" and as defined in [41] an agent is cmposed by assertions, beliefs and facts (and not knowledge).
-Indeed, it seems from your last bullet point that a fact is a particular case of knowledge (as you say, "more general"), but the exact definition of fact is not really clear from here, at least to me.
-Further; why are facts requirements?
-Intuitively, I would define a fact to be something that is true; requirements are something that we *want* to be true.
-I understood why we do so later on in the paper; we are later on interested, by RCC, in whether the beliefs of our agents are consistent (EQ) with the facts, that is, we want our system to behave consistently with the requirements.
-I would maybe make this explicit: we interpret requirements as facts, since we consider them to be the truth, towards which our agents must be compliant with (rather than: facts are requirements and not the actual truth).
-Another thing I found to be unclear is your definition of belief.
-You start by saying that we will define these things abstractly but you refer here to a very system-oriented definition of belief: inputs and outputs.
-Indeed, you say that beliefs define the behaviour, but the exact definition of belief was not super clear to me.
-{\bf mr}: I think you're right and the key points are (i) what's a fact?
-(ii) why stressing knowledge if abf is based on facts?
-(iii) what's the relation between facts and requirements?
-A fact is a statement which is taken as true.
-it's not necessarily true in general but the discussion is much complex (unfortunately).
-I wanted to add it to the paper but someone said that it may be tricky since those concepts have been used for 2k years at least.
-But, as an example, ``Socrates is a man'' is a fact (i.e.
-a true statement or, better, a statement taken as true), right?
-do we \emph{know} it?
-how?
-did we induce it from empirical evidences by testing every single man ever existed?
-I guess you see what I mean.
-So, in my limited view, facts are statement taken as true, knowledge can be expressed by true statements, and requirements are statement taken as true (by developers and everyone else).
-The link between those concepts seems rather strong but I think you are suggesting to simplify the text changing the focus on facts or detailing the relation between knowledge and facts.
-I chose the former and changed the text accordingly.}
 
@@
 \end{figure}
-
-\fixnote{Oliviero}{Further: what is a region?
-Is it a "prime" concept, that is, a primitive concept?
-Or is it a set?
-if it is a primitive concept, then defining connectedness in terms of set operations such as $\subseteq$ is a little counterintuitive; if a region is a set (or some other well-known mathematical concept), then state so explicitly somewhere, imho (as I said at the beginning, explicitly describing how these ABF regions look like would make this much clearer, I think).
-{\bf mr} you're right.
-it's a primitive concept, I changed it in the tab and in the text.}
 
@@
 readability, we use the term \emph{region} both to refer to a mereological part
-and to a topological region.  Our aim is to create a meronomy (hierarchy of part-whole
-relations)
-instead of the taxonomies (categorization based on discrete sets) 
-such as the one provided
-in \autocite{NIST2020NVD,MITRE2020CVE} 
-so that we don't need to rely on a scoring system (such as the 
-CVSS) to assign a quantitative evaluation 
-of the cybersecurity of each entry. Instead, we want a precise calculation
-of the number of insecure configurations
-\fixnote{Oliviero}{is this the same as a state?
-If so, maybe mention this in brackets in the definition.
-{\bf mr}: not necessarily but it will be in the next prototype and paper.
-For now, I'd keep it easy and general (plus, we'd need to introduce the concept of state oth.)} of a system to emerge from the mathematical
-formulation of our cybersecurity hypothesis.
+and to a topological region.  Our aim is to create a meronomy (hierarchy of
+part-whole relations) instead of the taxonomies (categorization based on
+discrete sets) such as the one provided in \autocite{NIST2020NVD,MITRE2020CVE}
+so that we don't need to rely on a scoring system (such as the CVSS) to assign
+a quantitative evaluation of the cybersecurity of each entry. Instead, we want
+a precise calculation of the number of insecure configurations of a system to
+emerge from the mathematical formulation of our cybersecurity hypothesis.
 
@@
 is a mathematical structure where the basic relation between regions
-is the reflexive and symmetric relation \emph{Connects With} $\subseteq$,
+is the reflexive and symmetric relation \emph{Connects With},
 that we use to order a universe of agents
@@
 The general formula to calculate the number of different types of agents is
-$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$\fixnote{Oliviero}{What if our calculus uses relations of different arities?
-I guess we sum this up for all arities?
-{\bf mr}: I think so, and I see your point (we could generalize the formula) maybe it's not that important?
-not sure},
+$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
 between $n$ different regions. Therefore, $r^{\binom{n}{k}}$
@@
 input and output assertions.  Given an agent $a$ and a collection of asserted
-predicates $\Phi$
-\fixnote{Oliviero}{If I am not mistaken, this is the first time we refer to assertions as concrete formulas/predicates.
-Perhaps we should make explicit before the "form" of such assertions?
-{\bf mr}: I'm not sure I got this point.
-Do you mean we should refer to a logic (e.g.
-FOL)?
-If so, we could but it has a cost (in terms of simplicity) and I'm not sure how to reuse that in the paper to justify that ``cost''},
+predicates $\Phi$,
 the input assertions are those received by $a$ from an agent
@@
 \begin{definition}{\bf Risk --}
-	\fix{mr}{A risk is determined by the uncertainty related to the whole space of potential designs of a specification with respect to the
-	$\abftheory$-framework.}
-\end{definition}\fixnote{Oliviero}{iirc, this is the first time we speak about uncertainity.
-What do you mean exactly?
-Do you mean that we are uncertain about which state our CPS is in?
-Is it not knowing that we are in the secure state?}
+The risk is the uncertainty related to the whole space of potential designs
+	resulting from a specification in the $\abftheory$-framework.
+\end{definition}
 The definition of Risk leads to the risk matrix in
@@
 	(RCC5 in our case).
-\end{definition}\fixnote{Oliviero}{So, the granularity is just uniformly distributed between all possible configurations, right?
-Or is it the actual numeric value unimportant, as long as it "ordered" in the right way?
-{\bf mr}: right, the former}
+\end{definition}
 While a risk matrix is often defined as a function of the likelihood and impact
@@
 to MAS. We now map MAS to CPS and show how this mapping allows us
-to predict cybersecurity weaknesses in a system.\fixnote{Oliviero}{again, to me we actually referred quite a lot to systems and architectures.}
+to predict cybersecurity weaknesses in a system.
 
@@
 		or device) is composed by a \emph{functional architecture} that
-		transforms input-beliefs into output-beliefs.\fixnote{Oliviero}{We often times used functional architectures before.
-Mabybe we should make this connection precise before, then?
-Or maybe avoid using this CPS-oriented terminology before, and keep it more abstract?}
+		transforms input-beliefs into output-beliefs.
 	\item Components of a CPS have \emph{ports} to exchange information
@@
 	$X$; where a base is a collection of regions in a (mereo)topology such that
-	%every open region\fixnote{Oliviero}{First time we refer to open regions; perhaps briefly introduce them?} can be written as union of elements of that base.
 	every region can be written as union of elements of that base. 
@@
 	interconnected system of functional blocks.
-\end{definition}\fixnote{Oliviero}{We referred to this notion many times in the esposition, and define it precisely only here. I'd perhaps speak more abstractly until this is introduced.}
+\end{definition}
 The quality of a functional block cannot be determined
@@
 \begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
-	A system security design is given by a precise system 
-	specification over the physical and functional architectures, that
-	uniquely, i.e.  tested against the range of designs possible in the
-	$\abftheory$-framework, defines the design built on top of those
+	A system security design (in the $\abftheory$-framework) is given by a
+	precise system specification over the physical and functional
+	architectures that uniquely defines the design built on top of those
 	requirements.
@@
 \begin{hypothesis}{\bf System Insecurity Design --}\label{hyp:insecurity}
-	If, given a system specification as a collection of
-	requirements on the system, there exist
-	a non-unique design with respect to those requirements, the number of
-	equivalent designs that fulfills the requirements quantitatively
-	defines the magnitude of insecurity of a system design.
+	If, given a system specification as a collection of requirements, there
+	exist a non-unique design with respect to those requirements, the
+	number of equivalent designs (that fulfills the requirements)
+	quantitatively defines the magnitude of insecurity of a system design.
 \end{hypothesis}
-
-\fixnote{Oliviero}{Ok, maybe here my inexperience with systems design is showing, but these hypothesis are not super clear to me.
-I understand the next part quite well, but the actual hypthesis, not so much.
-What are the designs possible with respect to the framework?
-Do you mean all the possible configurations of <rcc(A,B),rcc(A,F),rcc(B,F)> for each agent?
-What does "testing against" this range of possible designs?}
 
@@
 is the set of all the requirements of $S$,
-\fixnote{Oliviero}{I assume that $(v,f)\in R$ iff f is a requirement over v?
-Is it possible that this is unknown, i.e., we do not know which devices a requirement predicates over?
-{\bf mr}: I guess it is possible but I believe we may not consider this case here.
-If there's a functional block $v$ there must be a requirement related to it.
-It may be wrong and we may not precisely know it, but it is there.}
 and extend $G$ as $G'=\langle V',E'\rangle$ with $E'= E\cup R$ and $V'=V\cup F$. Let $\pi:
@@
 configuration is satisfiable with respect to the logical theory defining the
-algebraic structure (mereotopology)\fixnote{Oliviero}{Which structure are we talking about, here?
-Is it the system specification?
-See my previous comments on specifying what A,B, and F actually look like.} and constraints of the calculus RCC
+algebraic structure (mereotopology) and constraints of the calculus RCC
 (otherwise $\sigma$ returns $0$), we can define

### 2021-03-02 reports/paper_0/main.tex
remove last excel ref

@@
 (to name a few)\footnote{In formal protocol verifiers, in the DY model, the
-attacker has a set of hardcoded inference rules that defines how the attacker
+attacker has a set of hard-coded inference rules that defines how the attacker
 deduces new messages based on the observed message exchange.}.  As depicted in
@@
 cause an attack in the system.  Those vulnerabilities are, in turn, caused by
-the presence of weaknsesses in the system.  Weaknesses are errors in the design
+the presence of weaknesses in the system.  Weaknesses are errors in the design
 or implementation of a system.  Therefore, a theory on cybersecurity should
@@
 \emph{beliefs} (internal information considered true by the agent), and
-%and the information or 
 \emph{assertions} (an exchange of information between agents). % it provides.
@@
 	\cellcolor{abf-rg-4}\dr{$\assertionRegion$}{$\factRegion$}\\ % e q
-% & \cellcolor{abfred}& %dr
-%	\cellcolor{abf-rg-1}& %\pp{$\assertionRegion$}{$\factRegion$} & %po
-%	\cellcolor{abf-rg-2}  & %pp
-%	\cellcolor{abf-rg-3} & %\pp{$\assertionRegion$}{$\factRegion$} & %ppi
-%	\cellcolor{abf-rg-4} \\ %eq
 \hline %po
@@
 	\centering
-	\includegraphics[width=\columnwidth]{abftheory.pdf} %{mashyp.pdf}
+	\includegraphics[width=\columnwidth]{abftheory.pdf}
 	\caption{Mapping Epistemological Concepts to (Cyber-Physical) Systems Engineering}
@@
 between output and input assertions of an agent)
-%$\Rcc{\assertionRegion_o}{\assertionRegion_i}$) 
 lead to the same results of the
 analysis of a pair of an input and output port.
-%(i.e.
-%$\Rcc{\beliefRegion_1}{\assertionRegion_2}$ and
-%$\Rcc{\assertionRegion_2}{\beliefRegion_3}$)
-%As depicted in Figure~\ref{fig:atom}, 
 So far we have
@@
 
-%the insecurity of a system 
-%($I$) can be expressed as the following equation.
-%\begin{displaymath}
-%	I=\prod_{a\in Ag} r_i^{\binom{n}{k}}-\delta_T
-%\end{displaymath}
-%where $r_i^{\binom{n}{k}}$ represents the number of possible configurations
-%(permutations) of an agent (defined as a tuple of RCC relations as in
-%Definition~\ref{def:opsystem}) in which at least one of the relation is not EQ.
-%$r_i$ represents the RCC relations with arity $k$ between $n$ different regions
-%in the $\abftheory$ theory. $\delta_T$ represents the number of configurations
-%which are not satisfiable with respect to the logical theory defining the
-%algebraic structure (topology) and constraints of the calculus (RCC). $Ag$ is
-%the set of all the agents in the system under evaluation and the product
-%represents all the unintended, insecure, configurations.  Given that there is
-%only $1$ nominal configuration of the system where all the agents are defined
-%only with EQ relations (and then the only secure configuration), which can be
-%expressed as $sad$ it follows that security can be defined as
-%\begin{displaymath}
-%	I=\prod_{a\in Ag} r_i^{\binom{n}{k}}-\delta_T
-%\end{displaymath}
-%
-%if $I=0$ the system can be considered secure, 
-%otherwise $I$ represents the number of possible insecure configurations, where
-%$I\in\mathbb{N}$.
-
 \subsection{Cybersecurity Risk Assessment}\label{sec:secra}
@@
 	\centering
-	\includegraphics[width=.7\textwidth]{results_excel.pdf}
+	\includegraphics[width=.7\textwidth]{results.pdf}
 	\caption{Partial View of the results in the spreadsheet file}
@@
 
-%\section{Discussions}\label{sec:discussion}
-%\paragraph{Abstraction Level}
-%The level of abstraction is always a concern when it comes to the definition of
-%a system model.  In this paper, we justified our assumptions on the atomicity
-%of ports, channels, and functional blocks but nothing prevents a modeler to
-%consider other atomic element of the system specification and design.  Those
-%additional elements may detail the system and may change the syntactic sugar we
-%defined; however, it won't change the underlying (semantic) reasoning unless
-%another calculus is used (instead of RCC).  Furthermore, other Infosec
-%properties may be considered, such as authenticity or non-repudiation. We,
-%however, leave them as future development.
-%
-%\paragraph{$\abftheory$ Falsification} The theory we developed can be falsified
-%by providing an experiment that shows a weaknesses not predicted by the
-%$\abftheory$-framework, or by showing that the weaknesses predicted for a certain
-%experiment are impossible to realize.  This theory, however, needs to be
-%detailed to be applicable to a full engineering process (we only cover the
-%first steps).  We propose the following complete process as a draft for
-%potential next steps in this direction.
-%\begin{enumerate}
-%\item System Specification
-%\item Architecture Design
-%\item Security Risk Assessment and Security Requirement Extraction
-%\item Design Refinement, where the semantics of the functional behavior and
-%	communication protocols are defined 
-%\item Security Requirement Extension, based on the design refinement
-%\item Verification of Security Requirements (data flow, behaviors, \& protocol
-%	logic)
-%\item Software Prototype Implementation/Synthesis
-%\item Automated Security Testing (abstract test-suite generation \&
-%	concretization)
-%\item Hardware Prototype Implementation
-%\item Automated Security Testing
-%\item Penetration Testing
-%  \end{enumerate}
-%
-%\paragraph{Relation with DY}
-%The DY attacker model has been extensively
-%used for the identification of logical weaknesses in security protocols.
-%Our theory should predict all possible weaknesses which, in turn,
-%should enable all the attacks identified by the DY model. 
-%A proof that the two theories/models correctly predicts
-%(or lacks in predicting) weaknesses and related protocol flaws would
-%connect risk assessment and protocol verification, the first
-%two steps of the secure engineering of systems.
-
 \section{Conclusion and Future Works}
-%\fixnote{mv}{Se ho capito bene, detto a parole mie, quello che il vostro
-%approccio permette di fare, dati requisiti e architettura del sistema, è di
-%elencare in maniera sistematica tutti i "punti" in cui potrebbero sorgere dei
-%problemi. Spiegherei un po' meglio, qui nelle conclusioni o nella sezione
-%precedente, come questo approccio dovrebbe essere usato nella pratica... Voglio
-%dire che non mi sembra necessariamente in opposizione ad altri approcci ma
-%magari può integrarsi a questi.  Inoltre quando parlate dei passi successivi,
-%ad esempio il considerare la generazione di test cases, se avete già qualche
-%idea su come procedere, potreste discuterne brevemente.}
 We proposed an hypothesis for a foundational theory on security, arguing that
@@
 model of the system under assessment.
-%The current state-of-the-art processes allow, given a specification,
-%an engineer to design the system in such a way that cybersecurity issues arise due to the lack
-%of proper cybersecurity risk assessment processes.
-%Those designs, again, lack of cybersecurity verification processes based on a solid 
-%cybersecurity foundational theory and then permit the generation of insecure implementation.
 The cybersecurity verification and test-case generation will be the focus of our next steps.
-%We conclude that the problem of security is a problem related to the 
-%many possible design and, in turn, implementation given a specification.
-%The problem is related to the epistemological search for truth,
-%where the challenge is to relate information and belief to human knowledge.
-%Scientifically, generalizing the human to an agent, the problem is
-%to relate assertions and beliefs, to facts. From an engineering standpoint,
-%by generalizing the concept of agents as architectural subsystems, the
-%problem is to link channels (and ports) and functional architectures to 
-%requirements.
-
-%\section*{Acknowledgment}
-%We thank Katia Santac\`a and Luca Vigan\`o for their help in the development of
-%the initial ideas, and for the helpful discussions that allowed us to make this
-%step forward. We also thank Marco Volpe for his valuable feedback.
 
@@
 
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-%\bibliographystyle{IEEEtran}
-%\bibliography{../literature}
-
 \end{document}

### 2022-03-07 reports/cimss_2022/main.tex
draft cimsss paper

--- 2021-03-02 reports/paper_0/main.tex
@@
-\documentclass[conference]{IEEEtran}
+%\documentclass[conference]{llncs}
+\documentclass[runningheads]{llncs}
+\usepackage[T1]{fontenc}
 \usepackage[table]{xcolor}
@@
 \usepackage{tikz}
-\usepackage{amsthm}
+%\usepackage{amsthm}
 \usepackage{rotating}
@@
 \newcommand{\fix}[2]{{\color{red}{\bf tofix:} #2}}
+\newcommand{\toremove}[2]{{\color{blue}{\bf toremove:} #2}}
 % \renewcommand{\fixnote}[2]{}
@@
 
-\newtheorem{definition}{Definition}%[section]
+%\newtheorem{definition}{Definition}%[section]
 \newtheorem{hypothesis}{Hypothesis}%[section]
-\newtheorem{theorem}{Theorem}%[section]
+%\newtheorem{theorem}{Theorem}%[section]
 
@@
 
-\author{
-	\IEEEauthorblockN{The Science Club}
-\IEEEauthorblockA{V-Research\\
+\author{Michele Ambrosi\inst{1} \and
+Francesco Beltramini\inst{1} \and Federico De Meo\inst{1} \and Olivero Nardi\inst{1} \and Mattia Pacchin\inst{1} \and Marco Rocchetto}
+
+\authorrunning{M. Ambrosi et al.}
+
+\institute{V-Research\\
 Verona, Italy\\
-Email: knowledgezero@v-research.it}
-}
-
+\email{knowledgezero@v-research.it}}
 \maketitle
@@
 	predicts the weaknesses in a UML model of a (cyber-physical) system.
+  \keywords{Risk Management, Cyber-Physical Systems, Risk Assessment, Security Framework}
 \end{abstract}
 
-\IEEEpeerreviewmaketitle
-
 \section{Introduction}\label{sec:intro}
+\fix{mr}{
+\begin{enumerate}
+  \item Herley's work (1 paragraph)
+  \item Implications on Risk Assessment (no theory implies no complete set of weakness which, in turn, implies no risk assessment)
+  \item 1 page: contributions
+\end{enumerate}
+}
 \paragraph{Context and motivation} 

### 2022-03-07 reports/cimss_2022/compile.sh

@@ -1,0 +1,32 @@
+#!/bin/bash
+latexOutputDir="latex-output-dir"
+latexInputFile="main.tex"
+pdfOutputDir="${latexOutputDir}/${latexInputFile%.*}.pdf"
+
+if [ "$1" != "-short" ] && [ "$1" != "-s" ]; then
+	echo "remove"
+	rm -rf $latexOutputDir
+fi
+
+if [[ ! -d $latexOutputDir ]]; then 
+	mkdir $latexOutputDir
+fi
+
+if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
+then 
+	if [ "$1" != "-short" ] && [ "$1" != "-s" ] 
+	then 
+		if biber --nodieonerror --input-directory=$latexOutputDir --output-directory=$latexOutputDir ${latexInputFile%.*}  
+		then
+			if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
+			then 
+				pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile 
+			fi
+		fi
+	fi
+
+	if [[ -f $pdfOutputDir ]]
+	then
+		mv $pdfOutputDir `dirname $0`
+	fi
+fi

### 2022-03-08 reports/cimss_2022/main-cut.tex
first revision/cut

--- 2022-03-07 reports/cimss_2022/main.tex
@@
 \section{Introduction}\label{sec:intro}
-\fix{mr}{
-\begin{enumerate}
-  \item Herley's work (1 paragraph)
-  \item Implications on Risk Assessment (no theory implies no complete set of weakness which, in turn, implies no risk assessment)
-  \item 1 page: contributions
-\end{enumerate}
-}
 \paragraph{Context and motivation} 
@@
 support the non-falseness of a scientific theory.
-If we consider a natural (real-world) phenomenon, such as gravity, we can think
-of a scientific theory as an \emph{abstract} mathematical formula (or a set of
-formulas) that defines axiomatically a certain (natural) phenomenon in a
-specific algebraic structure. While such a theory can be proved to be correct
-within the mathematical setting it has been conceived in (e.g. with respect to
-other axioms of the algebraic structure where it has been postulated, or with
-respect to other background theories), no proof can be carried out to verify
-that the theory properly abstracts the phenomenon it describes (i.e. that the
-theory is a sound or complete approximation of the real phenomenon).
-Similarly, the results of the application of a scientific theory (i.e. its
-predictions) can be verified in the abstract mathematical structure, but how can
-we be sure that the theory correctly characterizes the phenomenon it describes? 
-Popper's famous ``demarcation principle'' draws a line between scientific and
-pseudo-scientific theories by proposing that a scientific theory can
-be considered acceptable as long as no empirical experiment falsify it. For
-example, a theory of gravity, in its mathematical setting, can predict how
-objects attract each other, and those predictions can be tested by empirical
-experiment. If the theory correctly represents the phenomenon, no empirical
-experiment should falsify the predictions of that theory.
 
@@
 of cybersecurity).  Herley then uses this argument to show that ``claims that
-any measure is necessary for security are empirically unfalsifiable''. An
-example is whether or not password masking makes a system secure. Password
-masking prevents a complete leakage of the password, and a system, in which
-password masking is implemented, is believed to be secure because an attacker
-won't be seeing the password by ``shoulder surfing''.  This reasoning
-is circular and it is always true that a system in which password masking is
-implemented is a system in which a password is masked. No experiment can ever
-falsify this statement\footnote{In other words, one cannot design an experiment to show that
-password masking is not secure or unnecessary, and a correlated problem is that
-this may lead to a scenario where more measures are added but there is no good
-way of saying which measures are superflous.}. Given that any theory that is not falsifiable by an
-empirical experiment is well known\footnote{``A theory which is not refutable
-by any conceivable event is nonscientific. Irrefutability is not a virtue of a
-theory (as people often think) but a vice.'' -- Karl Popper, Conjectures and
-Refutations \autocite{popper1962conjectures}.} to be nonscientific (i.e.
-unfalsifiability is a fallacy of a theory), Herley concludes that there is no
-scientific theory supporting this type of cybersecurity measures; generalizing,
-cybersecurity seems to lay in the realm of pseudo-sciences
-\autocite{Herley2016usenixvideo}. Herley, e.g.  in
-\autocite{Herley2017justifying}, discusses the implications of a nonscientific
-approach to cybersecurity, and highlights the tremendous impact on the
-scientific research and engineering of systems. While the criticism is
-investigated in \autocite{Herley2016unfalsifiability}, no solution is provided
-nor envisioned.
+any measure is necessary for security are empirically unfalsifiable''.
 
@@
 presence of errors in the design or implementation of a system} (i.e.
-cybersecurity weaknesses).  For example, a design or implementation weakness in
-a sanitization function may allow an attacker to perform SQL-injections. If
-weaknesses are necessary to enable cybersecurity attacks, a theory that
-predicts all potential weaknesses can be effectively used to predict the
-(in)security of a system.  Therefore, a possible answer to the initial question
-can be that (definition) cybersecurity is the absence of attacks and (claim)
-the absence of weaknesses implies cybersecurity.  This, however, raises another
-question: ``which are all the errors/weaknesses in (the engineering of) a
-system that make it possible for a cybersecurity attack to happen?''. More
-rigorously, which scientific theory predicts all possible weaknesses that
-causes cybersecurity attacks?
+cybersecurity weaknesses).
 
@@
 
-\paragraph{Structure} 
-We start, in Section~\ref{sec:literature}, by detailing the problem statement,
-reporting a literature review on the main concepts and definitions related to
-cybersecurity.  We formulate a cybersecurity hypothesis in Section~\ref{sec:hypothesis};
-which we use to propose a mathematical formulation of our hypothesis on system cybersecurity in
-Section~\ref{sec:theory}. In Section~\ref{sec:secra}, we apply our hypothesis for
-the tool-assisted cybersecurity risk assessment of a Cyber-Physical System (CPS), with an ad-hoc example.
-%based on the SWaT testbed \autocite{Mathur2016swat}. 
-This application shows
-how our hypothesis can be used to predict all of the possible cybersecurity weaknesses
-of a system, allowing the falsification of our hypothesis.
-
 \section{Literature Review}\label{sec:literature}
-\begin{figure*}[t]
-	\centering
-	\includegraphics[width=.8\textwidth]{safety-security_3.pdf}
-	\caption{Overview of Keywords Related to Cybersecurity and Safety}
-	\label{fig:safety-security}
-\end{figure*}
-
-In most of the natural languages the concepts of safety
-and security are not syntactically differentiated and both terms (safety and
-security) are expressed by the same word, e.g. ``sicurezza'' in Italian.  A
-semantic distinction between safety and (cyber)security is correlated to a
-belief\footnote{A belief has to be intended as a proposition which is supposed
-to be true but no conclusive evidence makes it known to be true.} that
-safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
-natural environment (e.g. natural events such as wearing of hardware
-components), whereas cybersecurity deals with \emph{incidents} posed by mankind (e.g.
-attackers and bugs). The fundamental difference between nature and mankind (and,
-in turn, between safety and cybersecurity) is believed to be on the different
-intents (incidents are intentional, whereas accidents are not) of the causes
-that generates a threat. Namely, nature is believed not to have malicious
-intents (but unfortunate causes-effects), whereas threats generated by mankind are
-malicious\footnote{Of course, logical flaws or bugs may be
-introduced by other means (e.g. ignorance) without explicit malicious intents,
-but the exploitation of those flaws is considered (for now, and detailed
-afterwards in the article) malicious, and thus we consider any vulnerability to
-be malicious even if it is due to the lack of skills.}.
-This conclusion seems to be against a general formulation of a cybersecurity theory;
-how can we define a theory that predicts what a human being will do?
+
 Cybersecurity attacks seem to be related to the creativity of the attacker
@@
 \subsection{Terminology}
-Most of the safety-preserving principles in the field of engineering of
-safety-critical cyber-physical systems (such as elevators and aircraft), upon
-which safety requirements are defined (e.g. in standards such as the IEC 61508
-or 61511 \autocite{IEC201761511}), are based on empirical tests and measurements.
-While reasoning by induction
-%\footnote{``So, whenever they argue ``Every man is an
-%animal and Socrates is a man; therefore Socrates is an animal,'' proposing to
-%deduce from the universal proposition ``every man is an animal'' the particular
-%proposition ``Socrates therefore is an animal,'' which in fact goes (as we have
-%mentioned) to establish by way of induction the universal proposition, the
-%fall into the error of circular reasoning, since they are establishing the
-%universal proposition inductively by means of each of the particulars and
-%deducing the particular proposition from the universal syllogistically.''
-%Sextus Empiricus, Outlines of Pyrrhonism II-195 \autocite{Empiricus1990Pyrrhonism}}
-based on the empirical observation should be avoided in general,
-since it may easily lead to false beliefs \autocite{Popper1959logic}, 
-inducing general principle by means of tests
-is often justified by the supposed impossibility of defining a theory
-that correctly predicts failures.
-A failure of a wire due to environment (e.g. due to humidity,
-dust, heat \&c) is defined from empirical evidences and processes have been
-standardized to test qualities of hardware components.  This process completely
-breaks down when a malicious environment (i.e. an attacker) is considered
-instead of the (supposedly honest and predictable) natural environment.
-Therefore, the same approach that is in use for safety, seems not to be
-applicable for cybersecurity (e.g. for cybersecurity testing).  An overview on the
-aforementioned aspects of safety and cybersecurity is depicted in
-Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
+We provide a baseline for a definition of
 the terms that structure our current understanding of cybersecurity.
@@
 \begin{itemize}
-	\item \emph{Mankind} ``refers collectively to humans''
-		\autocite{wiki-mankind}, while the concept of \emph{Nature} is
-		related ``to the intrinsic characteristics that plants,
-		animals, and other features of the world develop of their own
-		accord'' (e.g. the physical universe) \autocite{wiki-nature}.
-		There exist several terms to refer to an \emph{attacker}, e.g.
-		threat agent or threat source, but, from now on, we consider
-		the Causality principle to be the \emph{threat source}, Nature
-		or Mankind to be the \emph{threat agents} and an
-		\emph{attacker} as a specific malicious threat agent which
-		materializes a threat.
 	\item \emph{Vulnerability},
@@
 		controls, or implementation that could be exploited by a threat
-		source''. On the one hand, the definition is broad to enclose
-		as much causes (that generates a vulnerability) as possible,
-		on the other hand the term vulnerability should have a complete and sound
-		definition, so that no other causes (e.g.  other sources) but
-		the ones in the definition are responsible for a vulnerability.
-		%\footnote{``For this view, that
-		%\emph{That Which Is Not} exists, can never predominate. You
-		%must debar your thought from this way of search, nor let
-		%ordinary experience in its variety force you along this way,
-		%(namely, that of allowing) the eye, sightless as it is, and the
-		%ear, full of sound, and the tongue, to rule; but (you must)
-		%judge by means of the Reason (Logos) the much-contested proof
-		%which is expounded by me.'' -- Parmenides of Elea, On Nature
-		%(circa 500 B.C.), fragments B7.1–8.2
-		%\autocite{Hakim2016philosophy}}.
-		Furthermore, the term ``threat sources'' used in the definition
-		in \autocite{cnssi20104009} may be identified with both Nature
-		and Mankind, not differentiating between safety and security.
-\end{itemize}
-
-As depicted in Figure~\ref{fig:safety-security}, a vulnerability does not
-necessarily become a threat for the system, unless exploited ``through a
-channel that allows the violation of the security policy
-[\ldots]'' \autocite{cnssi20104009}. For example, a software or procedure that takes
-advantage of the vulnerability causing an \emph{attack} to the system may
-result in several correlated incidents and threats.  The process of
-exploitation of a defect as a vulnerability is reported in
-Figure~\ref{fig:safety-security} such that the difference between exploit and failure,
-and attack and accident, is to be found just in the maliciousness of the intents
-that causes this process (i.e. excluding the intent, the terms are just syntactic transformation from a vulnerability to defect, from
-accident to incident).
-
-\begin{itemize}
+		source''.
+
 	\item \emph{Weakness}. The definition given by the MITRE in
@@
 		that a vulnerability is caused by a weakness.
-	\item \emph{Causality} refers to the causality principle; defined
-		in \autocite{Spirkin1983Dialectical} as ``Causality is a genetic
-		connection of phenomena through which one thing (the cause)
-		under certain conditions gives rise to, causes something else
-		(the effect). The essence of causality is the generation and
-		determination of one phenomenon by another. In this respect,
-		causality differs from various other kinds of connection, for
-		example, the simple temporal sequence of phenomena, of the
-		regularities of accompanying processes''. 
-		%One can consider the
-		%causality principle to be formalized by the K Modal Logic.
+
 	\item An \emph{Exploit} ``[\ldots]
@@
 		unauthorized use of an asset''; where an \emph{Asset} is
-		``anything that has value to the organization''. Furthermore, we note that for
-		the purpose of this article, we do not want to focus on a specific
-		organization or business to define asset but, in general, on any 
-		abstract organization (e.g. a company or a society).
+		``anything that has value to the organization''.
 		We do not consider ethical hackers as attacking a system. 
@@
 
-Our literature review shows that most of the definitions relate insecurity to
-dis-honesty (also called maliciousness or adversarial) of an agent (often
-called adversary or attacker). In fact, weaknesses become vulnerability if an
-attacker exploits them in an attack.  This, however, just shifts the problem of
-defining what cybersecurity is as the problem of defining what a dishonest
-agent can or cannot do in a system, where an agent is any virtual or physical
-entity of the system or using the system (e.g. a device, a software, or a human
-being), and dishonesty is not necessarily related to malicious motivation but
-also to incompetence or lack of skills. As in the Dolev-Yao attacker model\footnote{The Dolev-Yao
-attacker model can be considered as abstract model of an attacker that can be
-used for the analysis of protocol specifications.} \autocite{Dolev1983security}, we may
-correlate ``being dishonest'' to ``not following the intended behavior/rules''.
-In the case of a generic system, a dishonest agent is, therefore, any agent
-that doesn't follow the intended behavior (or functionality or logic) of the
-system.  For example, a software can be considered an agent of the system and
-whenever it has a bug, it can be exploited causing an incident. This
-may lead to the conclusion that the dishonest behaviors of agents cannot be
-defined in general (e.g. due to the heterogeneity of agents and systems) and
-that a huge repository of dishonest behaviors should be kept as definition, such
-as CAPEC \autocite{MITRE2020CAPEC}; or that dishonesty of an agent with respect
-to a cybersecurity protocol should be defined as a number of predefined actions as
-in
-\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
-(to name a few)\footnote{In formal protocol verifiers, in the DY model, the
-attacker has a set of hard-coded inference rules that defines how the attacker
-deduces new messages based on the observed message exchange.}.  As depicted in
-Figure~\ref{fig:causality}, in order to define a theory on cybersecurity we may
+As depicted in Figure~\ref{fig:causality}, in order to define a theory on cybersecurity we may
 say that the presence of vulnerabilities is a necessary condition in order to
@@
 In order to address the problem raised by Herley, we shall define how to
-distinguish between a secure and an insecure system. While most of the
-literature have correlated the problem of in-security
-to the maliciousness of agents interacting with the system, we show that
-cybersecurity doesn't seem to stem from a malicious nature but, rather, insecurity
-raises from the lack of well-defined cybersecurity requirements for the development
-process of a system. We argue that the high number of cybersecurity vulnerability
-reported today are simply the realization of potential system configurations,
-which deviate from the  \emph{intended (or
-nominal) behavior} because the nominal behavior of the system is not precisely defined in the specification
-at the very early stages of the engineering process.
+distinguish between a secure and an insecure system. 
 As a reference, we consider the following steps of the engineering process
@@
 
-\subsection{Epistemology and Multi-Agent Systems}\label{sec:system}
-The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
-of system as ``A combination of interacting elements organized to achieve one
-or more stated purposes.''\autocite{ISO201515288}. If a system is structured
-into multiple sub-systems, a sub-system can be defined in the same way as a
-system is defined (i.e. as a composition of interacting elements). A hierarchy
-of sub-systems can be limited by a ``bottom-subsystem'' often called device (or
-element in the aforementioned definition) which is not itself composed by other
-elements but is considered atomic.         For the sake of simplicity, we
-consider a device as a system with only one element: itself.  We then first define a
-system as composed by a single element and then extend the definition to a
-``combination of interacting'' elements. Given that we use concepts from the
-multi-agent system research field, instead of using the word element we will
-use the word agent.
-
-There is no agreement between the research communities (e.g.
-Multi-Agent-System, Epistemic Logic) on what the constituent of an agent are,
-but the same ideas revolved around for thousands of years\footnote{
-It is interesting to notice that 
-in \autocite{Empiricus1990Pyrrhonism}, the author states: ``The logical
-criterion also may be used in three senses -- of the agent, or the instrument,
-or the ``according to what''; the agent, for instance, may be a man, the
-instrument either sense perception or intelligence, and the ``according to
-what'' the application of the impression ``according to'' which the man
-proceeds to judge by means of one of the aforesaid instruments''.}.
-Some relevant examples for our objective are the following.
-\begin{itemize}
-	\item In \autocite{Hintikka1962knowledge}, Hintikka describes the
-		concept of belief (as epistemological
-		concepts) and their relation with knowledge,
-		and the whole Doxastic logic defines in details how
-		beliefs can be formalized.
-	\item In \autocite{Hintikka1993Information}, Hintikka describes the concept
-		of information (and the difference with knowledge and belief).
-	\item In \autocite{Santaca2016abf}, the authors defines an agent as a
-		tuple of assertions, beliefs, and facts. A detailed discussion
-		on facts and assertions can be found in \autocite{facts2007stanford} and
-		\autocite{assertion2007stanford} respectively.
-\end{itemize}
-For our argument, as in \autocite{Santaca2016abf}, an agent is composed by its
-\emph{beliefs} (internal information considered true by the agent), and
-\emph{assertions} (an exchange of information between agents). % it provides.
-%knowledge, beliefs, and the information or assertions it provides; where
-%knowledge is defined, as in \autocite{Steup2020epistemology}, as a set of
-%propositions known by an agent, such that: (i) knowledge requires belief, (ii)
-%knowledge requires truth, (iii) knowledge must be \emph{properly justified}, and
-%the only objective of information is to exchange beliefs between agents.
-%As in \autocite{Steup2020epistemology}, 
-%the only objective of information is to
-%exchange beliefs between agents and, 
-As defined in \autocite{Hintikka1993Information},
-``information is specified by specifying which alternatives concerning the
-reality it admits and which alternatives excludes''. This means that if we
-consider a propositional variable (which admits the two alternatives
-True/False) its information is defined as believed to be True/False and not
-believed to be the opposite.  Due to the definition of information and then
-with its relation to the probabilistic correlation to truth/reality, we
-consider information in relation to an agent's beliefs. Similarly, we consider
-beliefs to define the actual behavior of an agent or a system; meaning
-that an agent has its own beliefs on top of which it bases its behavior,
-the consequences of which, in turn, will update its initial beliefs.
-%; i.e. we consider
-%behaviors as belief revision systems.
-%The difference between knowledge and information as used by Hintikka, and facts
-The concepts of facts and assertions as used in \autocite{Santaca2016abf}, are
-not precisely discussed in \autocite{Santaca2016abf}.  Furthermore, those
-epistemological concepts are difficult to define
-\autocite{Gettier2012knowledge} in general.
-%\footnote{``\emph{Theaetetus}: [\ldots] He said
-%that knowledge was true opinion accompanied by reason, but that unreasoning
-%true opinion was outside of the sphere of knowledge; and matters of which there
-%is not a rational explanation are unknowable -- yes, that is what he called
-%them -- and those of which there is are knowable. [\ldots] \emph{Socrates}:
-%[\ldots] the primary elements of which we and all else are composed admit of no
-%rational explanation; for each alone by itself can only be named, and no
-%qualification can be added, neither that it is nor that it is not, for that
-%would at once be adding to it existence or non-existence, whereas we must add
-%nothing to it, if we are to speak of that itself alone.  [\ldots]'' Plato --
-%Theaetetus 201 \autocite{Plato1914Plato}}.
-For example, Hintikka in
-\autocite{Hintikka1993Information} states that ``A purely logical definition of
-information is impossible''. However, a detailed formalization of those concepts is out
-of the scope of this paper. 
-Instead, we now define those concepts informally, inspired by the
-$\abftheory$-framework  defined in \autocite{Santaca2016abf} and, in
-Section~\ref{sec:theory} we detail them, and properly define them, as engineering concepts.
-\begin{enumerate}
-	\item We consider information only when the intention of exchanging 
-		that information is from a sender to 
-		recipient; and we call it \emph{assertion}.
-	\item Similarly, the portion of \emph{beliefs} we consider for system
-		engineering is the one that builds (input) or describes
-		(output) the behavior of an agent.
-		%(definitory rules
-		%\footnote{``The logical structure of information is one of the
-		%most basic and one of the simplest
-		%thing in the wide and wonderful world of logical analysis. This
-		%point can be put in a deeper perspective. A distinction
-		%[\ldots] ought to be made [--] between two kinds of rules (or
-		%principles) in any strategic activity like knowledge seeking.
-		%On the one hand you have the rules that define the game, e.g.
-		%how chessmen are moved on a board. The can be called
-		%\emph{definitory} rules.  They must be distinguished from rules
-		%[\ldots] that deal with what is better and what is worse in the
-		%game in question.  Definitory rules do not say anything about
-		%this subject. Rules which do can be called \emph{strategic
-		%rules}'' -- Hintikka in \autocite{Hintikka1993Information}},
-	\item We consider a set of axiomatic \emph{facts} determining how the
-		system shall be. Similarly, requirements determine which facts
-		must hold when the system is implemented\footnote{We consider facts as
-		definitory rules\autocite{Hintikka1993Information} since they
-		don't define how a real system is finally implemented but how
-		it shall be, trough a series of requirements.  Those
-		requirements may not hold, through insecurity, in the
-		implemented system.}. Specifically, facts/requirements describe:
-		the functional architecture of each agent, and the
-		physical/structural (HW/SW) architecture of each subsystem of
-		agents and agent within a subsystem.
-\end{enumerate}
-
-We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}.
-The notation will be introduced in the next section but, intuitively,
-Figure~\ref{fig:system-agent} shows a system composed by two agents. The system
-has two ports denoted by ``I'' (with incoming assertions) and ``O'' (with outgoing assertions).
-The bottom part of Figure~\ref{fig:system-agent} details the structure of a single
-agent which is not decomposed into other (sub-)agents. In this case, ports translates
-incoming assertions into beliefs that feeds the behavior block, and outgoing beliefs
-from the behavior block are then translated back into (outgoing) assertions.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.9\columnwidth]{system-agent.pdf}
-	\caption{Example of system and agent structure}
-	\label{fig:system-agent}
-\end{figure}
-
 \subsection{Mereo-topological Reasoning}
+\fix{mr}{To be reviewed by Marco}
 \begin{table*}[t]
@@
 consider that, by definition, assertions are defined as transfer of information
-between two agents, e.g., between two agents.  Therefore, as depicted in
+between two agents.  Therefore, as depicted in
 Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
@@
 		describes the soundness of the execution of the functional
-		architecture w.r.t. input elicitation. With more details, the
-		ideal specification of the functional architecture, along with
-		the expected inputs, defines the functional behavior of an
-		ideal system.  If all the inputs (assertions) are correctly
+		architecture w.r.t. input elicitation.
+		If all the inputs (assertions) are correctly
 		handled in the functional specification (beliefs) the
@@
 		completeness of the behavior defined in the specification
-		w.r.t. the input elicitation.  With more details, if all the
+		w.r.t. the input elicitation. If all the
 		outputs (assertions) of the functional architecture can be
@@
 through assertions generates beliefs, a dishonest agent may circulate false
-information, generating false beliefs.  We note that it is often implied that
-the intention behind circulating false information discerns a dishonest and an
-incompetent agent. However, we consider honesty related to sharing
-truths.
+information, generating false beliefs.
 %\footnote{``[\ldots] truth exists only in the good man, but the true in
@@
 			compliant with the expected inputs of the functional
-			architecture}. For example, the inputs of the user of a
-			SW must be sanitized to exclude deviations w.r.t. the
-			expected inputs of the functions implemented in the SW.
-			Another example is the type checking between allowed
-			inputs and expected inputs.
+			architecture}.
 		\item[CSE-$1.2$] Similarly, the equality relation
@@
 Several standards mandate a secure-by-design approach in which cybersecurity
-shall be considered at the very early stages of the design process.  For
-example,
-\begin{enumerate}[noitemsep]
-	\item DO-326A -- ``Airworthiness Security Process Specification''
-		requires a cybersecurity risk assessment of the design and
-		``are the only Acceptable Means of Compliance (AMC) by FAA
-		\& EASA for aviation cyber-security airworthiness certification,
-		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
-	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
-		Control System (ICS) Security''.
-	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
-		for Cyber-Physical Vehicle Systems'' defines ``set of
-		high-level guiding principles for Cybersecurity as it relates
-		to cyber-physical vehicle systems'' and states that
-		``incorporate Cybersecurity into cyber-physical vehicle systems
-		from concept phase through production, operation, service, and
-		decommissioning''.
-\end{enumerate}
+shall be considered at the very early stages of the design process.
 Standards do not describe in detail how to perform a cybersecurity risk 
@@
 can be summarized as to provide an understanding of the potential cybersecurity risks.
-Roughly speaking, a cybersecurity risk assessment (e.g. CORAS \autocite{Lund2010model}) 
-methodology starts from the (i) identification of assets, (ii) determines
-the threat scenarios (correlating vulnerabilities, threats and incidents) e.g. 
-threat diagrams in the CORAS
-approach, (iii) calculates the risk as a function
-of impact and likelihood of threat scenarios, and (iv) finally provides
-treatments (mandating a partial re-design) and residual risks.
 All the methodologies and tools we reviewed (e.g. Threatmodeler \autocite{Threatmodeler},
@@
 the identification of threats and for the quantitative estimation of risks.
-As Herley puts it \autocite{Herley2016usenixvideo}, ``we don't have a test for security and
-we don't have a metric for security'' while standards require a cybersecurity risk assessment
-which, in turn, mandates a metric to evaluate the cybersecurity risk. 
 In contrast, in this section, we define how to specify a CPS in the $\abftheory$-framework
@@
 
-So far, we have reasoned on systems in the abstract, considering
-generic Multi-Agent System (MAS); mapping Epistemological concepts
-to MAS. We now map MAS to CPS and show how this mapping allows us
+We now map MAS to CPS and show how this mapping allows us
 to predict cybersecurity weaknesses in a system.
@@
 assuming any RCC relation, between inputs and outputs of a port, which is not
-an equality can be considered as generating a weaknesses, as we describe in
-more details in the next paragraph.
+an equality can be considered as generating a weaknesses.
 
@@
 focus on the architecture specification step of the engineering process
-without going into the design of the behavior of agents. We can give an example
-but research is needed to apply our hypothesis
-to the behavior of agents.
-
-If we consider a functional block that generates nonces (i.e. number used
-only once), the semantics of this block must define a procedure such that 
-a freshness requirement holds. By the RIDI-Hypothesis we have that
-the following weaknesses may occur:
-\begin{itemize}
-	\item Replace: the output-nonces of the block are replaced with
-		non-fresh numbers.
-	\item Insert: the output-nonces are concatenated with other numbers,
-		resulting in non-fresh nonces. For example, the nonces 110 and
-		111 are modified by inserting a 1 and a 0 at the beginning and
-		end respectively, resulting in the two identical (and the
-		non-fresh) nonces 1110 and 1110.
-	\item Delete: part of output-nonce is dropped. For example, the last bit of 1101 and 1100 is dropped, resulting in the two equal and non-fresh nonces: 110 and 110.
-	\item Inject: substitution at bit level results in non-fresh nonces. For example, 110 and 111 becomes equal by flipping either the last 0 of the former or the last 1 of the latter.
-\end{itemize}
-
+without going into the design of the behavior of agents.
 We summarize our results by categorizing the weaknesses predicted by our hypothesis

### 2022-03-08 reports/cimss_2022/compile-cut.sh

@@ -1,0 +1,32 @@
+#!/bin/bash
+latexOutputDir="latex-output-dir"
+latexInputFile="main-cut.tex"
+pdfOutputDir="${latexOutputDir}/${latexInputFile%.*}.pdf"
+
+if [ "$1" != "-short" ] && [ "$1" != "-s" ]; then
+	echo "remove"
+	rm -rf $latexOutputDir
+fi
+
+if [[ ! -d $latexOutputDir ]]; then 
+	mkdir $latexOutputDir
+fi
+
+if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
+then 
+	if [ "$1" != "-short" ] && [ "$1" != "-s" ] 
+	then 
+		if biber --nodieonerror --input-directory=$latexOutputDir --output-directory=$latexOutputDir ${latexInputFile%.*}  
+		then
+			if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
+			then 
+				pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile 
+			fi
+		fi
+	fi
+
+	if [[ -f $pdfOutputDir ]]
+	then
+		mv $pdfOutputDir `dirname $0`
+	fi
+fi

### 2022-03-14 reports/cimss_2022/main-cut.tex
review 14 March 2022 (marco & michele)

@@
 \documentclass[runningheads]{llncs}
+%\usepackage{minipage}
 \usepackage[T1]{fontenc}
@@
 	development of a scientific theory that determines (all and only) the
-	possible insecure and secure configurations of any abstract system. We
+	possible insecure and secure configurations of any abstract system to be used for the risk assessment of systems. We
 	claim that cybersecurity weaknesses (i.e. errors) are at the beginning
-	of the causality chain that leads to cybersecurity attacks.  We, then,
-	formulate a scientific (falsifiable by experiments) hypothesis that we
-	use to predict all the weaknesses in the architectural design of an
-	abstract system. The mathematical formulation of our hypothesis is
-	based on a correlation between the epistemological concepts of
-	facts, beliefs, and assertions with the engineering concepts of
-	requirements, functional architectures, and channels.  Ultimately, our
+	of the causality chain that leads to cybersecurity attacks.  We
+	formulate an hypothesis that we
+	use to predict the weaknesses in the architectural design of an
+	system. Our
 	hypothesis allows for the definition of a mathematical formula which
@@
 \section{Introduction}\label{sec:intro}
-\paragraph{Context and motivation} 
+\paragraph{Context and motivation.}
 A \emph{scientific theory} is an explanation of a phenomenon such that the
@@
 support the non-falseness of a scientific theory. 
-
 In \autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
@@
 
-\paragraph{Contributions} 
+\paragraph{Contributions.} 
 The goal of this paper is to lay the foundations of a scientific cybersecurity
@@
 %\end{figure}
-%As an illustrative example, in Figure~\ref{fig:protocol-example}, under the
+%As an illustrative example, in Fig.~\ref{fig:protocol-example}, under the
 %perfect cryptography assumption (i.e. assuming that the cryptographic
@@
 %semantically equivalent (cybersecurity is confidentiality and vice-versa).
-
 Instead of starting from reasoning on what makes a system secure or insecure,
@@
 cybersecurity weaknesses).
-
 With our approach, a list of weaknesses emerges from the mathematical
@@
 and thus unpredictable. 
-
 Currently, the most complete understanding of insecurity issues is stored into
@@
 
-\begin{itemize}
-	\item \emph{Vulnerability},
-		%\footnote{The term vulnerability is not
-		%present in the Encyclopedia of Cryptography and Security, while
-		%it is used in 12 entries (such as in the definition of
-		%``penetration testing'' \autocite{caddy2005pentest})
-		%highlighting how commonly this word is used without a proper
-		%supporting semantics.}, 
-		as defined in \autocite{cnssi20104009}
+\emph{Vulnerability}.
+	As defined in \autocite{cnssi20104009}
 	(and adopted in \autocite{nist2013800-53}), is a ``weakness in an
@@
 
-	\item \emph{Weakness}. The definition given by the MITRE in
+\emph{Weakness}. The definition given by the MITRE in
 	\autocite{MITRE2020CWEweakness} of weakness is: `` a type of
@@
 
-	\item An \emph{Exploit} ``[\ldots]
+\emph{Exploit} ``[\ldots]
 	(from the English verb to exploit, meaning to use something to
@@
 	(usually computerized).'' \autocite{wiki-exploit}.
-	\item An \emph{Attack}, as defined by the International Standard
+
+\emph{Attack}. As defined by the International Standard
 	ISO/IEC 27000, is an ``attempt to destroy, expose, alter,
 	disable, steal or gain unauthorized access to or make
-		unauthorized use of an asset''; where an \emph{Asset} is
+	unauthorized use of an asset''; where an{Asset} is
 	``anything that has value to the organization''.
 	We do not consider ethical hackers as attacking a system. 
-		In fact, we consider the term \emph{hack} as
+	In fact, we consider the term{hack} as
 	non-malicious (as, e.g. in \autocite{Stallman2002hacker}).
-	\item A \emph{Threat}, as defined in \autocite{cnssi20104009}, is ``Any
+
+\emph{Threat}. As defined in \autocite{cnssi20104009}, is ``Any
 	circumstance or event with the potential to adversely impact
@@
 	information, and/or denial of service''.
-	\item \emph{Defect}: ``anything that renders the product not reasonably
-		safe'' \autocite{Robinson2019writing} (i.e. a characteristic of
-		an object which hinders its proper usability).
-	\item \emph{Failure}, as defined in \autocite{Merriam2020failure}, is ``a state of
-		inability to perform a normal function''. The term is
-		structured and detailed in
-		\autocite{cnssi20104009,iet2017glossary}, but relying on an
-		abstract notion of failure without a specific definition.
-	\item \emph{Hazard}, ``a potential source of
-		harm'' \autocite{iet2017glossary}.
-\end{itemize}
 
@@
 	\centering
-	\includegraphics[width=.6\columnwidth]{causality.pdf}
+	\includegraphics[width=.35\columnwidth]{causality.pdf}
 	\caption{Etiology of cybersecurity}
@@
 
-As depicted in Figure~\ref{fig:causality}, in order to define a theory on cybersecurity we may
+As depicted in Fig.~\ref{fig:causality}, in order to define a theory on cybersecurity we may
 say that the presence of vulnerabilities is a necessary condition in order to
@@
 \section{A Cybersecurity Hypothesis in the $\abftheory$-framework}\label{sec:hypothesis}
-In order to address the problem raised by Herley, we shall define how to
-distinguish between a secure and an insecure system. 
-As a reference, we consider the following steps of the engineering process
-and our objective is to be able to predict the cybersecurity requirements
-instead of axiomatically assume them.
+To address the problem raised by Herley, we define how to
+distinguish between a secure and an insecure system in the following steps 
+of the engineering process:
 \begin{enumerate}
-	\item \emph{System Specification}, where the \emph{functional and physical requirements} are defined.
-	\item \emph{Architecture Design}, where the specification is structured into \emph{functional and physical architectures}.
-	\item \emph{Cybersecurity Risk Assessment}, where the potential \emph{weaknesses} (errors)
-		are identified and, consequently, \emph{cybersecurity requirements} are
-		captured.
+	\item \emph{System Specification}: the \emph{functional and physical requirements} are defined.
+	\item \emph{Architecture Design}: the specification is structured into \emph{functional and physical architectures}.
+	\item \emph{Cybersecurity Risk Assessment}: potential \emph{weaknesses} (errors) and
+		\emph{cybersecurity requirements} are identified.	
     %\item \emph{Design Refinement}, where the physical and functional
@@
 \end{enumerate}
-Given that, in our investigation, we changed the focus from the attacker to the
-potential designs of a system, we start by defining a general framework for the definition
-of a system. On top of this framework we will identify system weaknesses as potential
-design errors.
-
-\subsection{Mereo-topological Reasoning}
-\fix{mr}{To be reviewed by Marco}
+We changed the focus from the attacker as the source of insecurity to the
+potential design errors of a system. We now define a framework for the
+definition of a system that we use to identify weaknesses as potential design
+errors.
+
+\subsection{Mereo-topological Reasoning (2.5)}
 \begin{table*}[t]
 \centering
-\begin{tabular}{ccclll} 
-\rota{\textbf{RCC3}}&\rota{\textbf{RCC5}}&\rota{\textbf{RCC8}}&\textbf{Terminology} & \textbf{Notation} & \textbf{Definition} \\
+\begin{tabular}{cclll} 
+\rota{\textbf{RCC3}}&\rota{\textbf{RCC5}}&\textbf{Terminology} & \textbf{Notation} & \textbf{Definition} \\
 \hline
-&&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& reflexive and symmetric \\
-&&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& irreflexive or antisymmetric\\
-&&&Part of				& $\mathit{P}(\mathit{X},\mathit{Y})$		& $\forall \mathit{Z} ~\mathit{C}(\mathit{Z},\mathit{X}) \rightarrow \mathit{C}(\mathit{Z},\mathit{Y})$\\
-&&&Overlaps			& $\mathit{O}(\mathit{X},\mathit{Y})$		& $\exists \mathit{Z} ~\mathit{P}(\mathit{Z},\mathit{X})\wedge \mathit{P}(\mathit{Z},\mathit{Y})$\\
-\Tdot&\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
-\Tdot&&&  \textbf{Overlaps Not Equal} 	& $\mathit{ONE}(\mathit{X},\mathit{Y})$		& $\mathit{O}(\mathit{X},\mathit{Y}) \land \neg \mathit{EQ}(\mathit{X},\mathit{Y})$ \\
-\Tdot&\Tdot&\Tdot& \textbf{DiscRete from} 		& $\mathit{DR}(\mathit{X},\mathit{Y})$		& $\neg \mathit{O}(\mathit{X},\mathit{Y})$\\
-&\Tdot&\Tdot&\textbf{Partial-Overlap}	& $\mathit{PO}(\mathit{X},\mathit{Y})$ 		& $\mathit{O}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{X},\mathit{Y}) \wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\
-&\Tdot&&\textbf{Proper-Part-of} 	& $\mathit{PP}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\ 
-	&\Tdot&&\textbf{Proper-Part-of-\textit{\textbf{i}}nverse} & $\mathit{PPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{Y},\mathit{X}) \wedge \neg \mathit{P}(\mathit{X},\mathit{Y})$\\
-&&\Tdot&\textbf{Externally Connected} 	& $\mathit{EC}(\mathit{X},\mathit{Y})$ 		& $\mathit{C}(\mathit{X},\mathit{Y}) \wedge \neg\mathit{O}(\mathit{X},\mathit{Y})$\\ 
-&&\Tdot&\textbf{Tangential PP} 	& $\mathit{TPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Tangential PPi} 	& $\mathit{TPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{TPP}(\mathit{Y},\mathit{X})$\\ 
-&&\Tdot&\textbf{Non-Tangential PP} 	& $\mathit{NTPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\neg\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
-&&\Tdot&\textbf{Non-Tangential PPi} 	& $\mathit{NTPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{NTPP}(\mathit{Y},\mathit{X})$\\ 
+&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& reflexive and symmetric \\
+&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& irreflexive or antisymmetric\\
+&&Part of				& $\mathit{P}(\mathit{X},\mathit{Y})$		& $\forall \mathit{Z} ~\mathit{C}(\mathit{Z},\mathit{X}) \rightarrow \mathit{C}(\mathit{Z},\mathit{Y})$\\
+&&Overlaps			& $\mathit{O}(\mathit{X},\mathit{Y})$		& $\exists \mathit{Z} ~\mathit{P}(\mathit{Z},\mathit{X})\wedge \mathit{P}(\mathit{Z},\mathit{Y})$\\
+\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
+\Tdot&&  \textbf{Overlaps Not Equal} 	& $\mathit{ONE}(\mathit{X},\mathit{Y})$		& $\mathit{O}(\mathit{X},\mathit{Y}) \land \neg \mathit{EQ}(\mathit{X},\mathit{Y})$ \\
+\Tdot&\Tdot& \textbf{DiscRete from} 		& $\mathit{DR}(\mathit{X},\mathit{Y})$		& $\neg \mathit{O}(\mathit{X},\mathit{Y})$\\
+&\Tdot&\textbf{Partial-Overlap}	& $\mathit{PO}(\mathit{X},\mathit{Y})$ 		& $\mathit{O}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{X},\mathit{Y}) \wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\
+&\Tdot&\textbf{Proper-Part-of} 	& $\mathit{PP}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\ 
+	&\Tdot&\textbf{Proper-Part-of-\textit{\textbf{i}}nverse} & $\mathit{PPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{Y},\mathit{X}) \wedge \neg \mathit{P}(\mathit{X},\mathit{Y})$\\
+%&&\Tdot&\textbf{Externally Connected} 	& $\mathit{EC}(\mathit{X},\mathit{Y})$ 		& $\mathit{C}(\mathit{X},\mathit{Y}) \wedge \neg\mathit{O}(\mathit{X},\mathit{Y})$\\ 
+%&&\Tdot&\textbf{Tangential PP} 	& $\mathit{TPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
+%&&\Tdot&\textbf{Tangential PPi} 	& $\mathit{TPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{TPP}(\mathit{Y},\mathit{X})$\\ 
+%&&\Tdot&\textbf{Non-Tangential PP} 	& $\mathit{NTPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\neg\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\
+%&&\Tdot&\textbf{Non-Tangential PPi} 	& $\mathit{NTPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{NTPP}(\mathit{Y},\mathit{X})$\\ 
 \end{tabular}
-\caption{RCC3, RCC5, and RCC8 relations between regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
+\caption{RCC3 and RCC5 relations between regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
 \end{table*}
 
-Similarly to \autocite{Santaca2016abf}, we define agents as a meronomy (an
+As in \autocite{Santaca2016abf}, we define agents as a meronomy (an
 hierarchy of Part-Whole relations) over the constituents that we previously defined
@@
 mereo-topology \autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
-considering the relations in Table~\ref{tab:rcc358}.  For the sake of
+considering the relations in Tab.~\ref{tab:rcc358}.  For the sake of
 readability, we use the term \emph{region} both to refer to a mereological part
@@
 that we use to order a universe of agents
-$\agentuniverse$ (see in Table~\ref{tab:rcc358}).  We use the Region Connection
+$\agentuniverse$ (see in Tab.~\ref{tab:rcc358}).  We use the Region Connection
 Calculus (RCC), as defined in \autocite{bennettLogics,improvingRCC}, to provide
 an axiomatization of the mereo-topological concepts. In its broader definition,
-the RCC theory is composed by eight axioms, and is known as RCC8. In the text,
-for brevity, we will often focus only on RCC5 without loss of generality. Using
+the RCC theory is composed by eight axioms, and is known as RCC8 \autocite{Grutter2008rcc}. Using
 RCC5 instead of RCC8 prevents us from considering tangential connections
@@
 RCC5.
-
-In Table~\ref{tab:rcc}, we summarize the axioms of the RCC (see, e.g., \autocite{Grutter2008rcc}).  We can now define a system
+In Tab.~\ref{tab:rcc}, we summarize the axioms of the RCC (see, e.g., \autocite{Grutter2008rcc}).  We can now define a system
 over the mereotopology using the RCC calculus, as follows, where $\rcc(X,Y)$ on
@@
 
-\paragraph{Number of different configurations of a system}
+\paragraph{Number of different configurations of a system.}
 The general formula to calculate the number of different types of agents is
@@
 \item \emph{RCC5} --- theoretical: $5^3=125$, correct: 54.
-\item \emph{RCC8} --- theoretical: $8^3=512$, correct: 193.
+%\item \emph{RCC8} --- theoretical: $8^3=512$, correct: 193.
 \end{itemize}
 
-Hence, even if considering a different number of regions than the three
-$\assertionRegion$, $\beliefRegion$, and $\factRegion$ exponentially affects
-the number of theoretical agents, the application of RCC downscales that number
-of a factor that ranges from 1.8 to 2.5. In addition, using RCC5 we consider
-3.6 times more (different) types of agents than RCC3, but using RCC8 would
-allow us to consider 3.5 times more different agents.
-In the quantitative evaluation of a single agent, depicted in Figure~\ref{fig:quantitative},
+%Hence, even if considering a different number of regions than the three
+%$\assertionRegion$, $\beliefRegion$, and $\factRegion$ exponentially affects
+%the number of theoretical agents, the application of RCC downscales that number
+%of a factor that ranges from 1.8 to 2.5. In addition, using RCC5 we consider
+%3.6 times more (different) types of agents than RCC3, but using RCC8 would
+%allow us to consider 3.5 times more different agents.
+In the quantitative evaluation of a single agent, depicted in Fig.~\ref{fig:quantitative},
 we argue that only 1 configuration represents the nominal (expected) behavior 
@@
 	\centering
-	\includegraphics[width=\columnwidth]{quantitative.pdf}
+	\includegraphics[width=.8\columnwidth]{quantitative.pdf}
 	\caption{Cybersecurity risk for a single agent}
@@
 
-\subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}
+\subsection{Qualitative Evaluation of Agent Space in $\abf$ (3)}\label{sec:agentspace}
 \begin{table}[t]
 \centering
-\begin{adjustbox}{width=\columnwidth}
+\begin{adjustbox}{width=.7\columnwidth}
 \begin{tabular}{r||c|c|c|c|c} 
@@
 \end{adjustbox}
-\caption{RCC5 composition table over 3 regions. The results show that there
-exist 54 possible relations and the coloring anticipates the ideal risk matrix
-(green the secure state with low risk, red the high risk state, and a gradient
+\caption{RCC5 composition table over 3 regions generates the ideal risk matrix
+(green the low risk, red the high risk state, and a gradient
 of medium risk states). 
@@
 configurations allowed by the $\abftheory$-framework, and how those configurations can
-be categorized.  In Table~\ref{tab:5com}, we provide the generic composition
+be categorized.  In Tab.~\ref{tab:5com}, we provide the generic composition
 table of RCC5 over 3 regions instantiated over $\abf$, which shows the whole
@@
 corner of the matrix, lowest on the bottom right corner. 
-
-In Figure~\ref{fig:soundness}, the relation between facts, and assertions and
+In Fig.~\ref{fig:soundness}, the relation between facts, and assertions and
 beliefs (as inputs and outputs of the behavior of an agent) is illustrated.
@@
 to determine the soundness of the design with respect to the specification.
-
 We first analyze the relations between each pair of
@@
 	\centering
-	\includegraphics[width=.9\columnwidth]{soundness.pdf}
+	\includegraphics[width=.5\columnwidth]{soundness.pdf}
 	\caption{Example relation between facts, and assertions and beliefs}
@@
 
-\paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration} In order
+\paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration.} In order
 to reason on the relation between assertions and beliefs we first need to
@@
 between two agents.  Therefore, as depicted in
-Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
+Fig.~\ref{fig:system-agent}, an agent has two main categories of assertions,
 input and output assertions.  Given an agent $a$ and a collection of asserted
@@
 \paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and
-\Rcc{$\beliefRegion$}{$\factRegion$} -- Honesty and Competence} 
+\Rcc{$\beliefRegion$}{$\factRegion$} -- Honesty and Competence.} 
 The relation of assertions
@@
 
-We are now in the position to define what a secure system is (with respect to
+We can now define what a secure system is (with respect to
 the $\abftheory$-framework) and, based on that definition, what the cybersecurity risk is and
@@
 	A secure system is a system where CSE-$1$, CSE-$2$, and CSE-$3$ holds for
-	each agent composing the system.
+	each agent of the system.
 \end{definition}
@@
 The definition of Risk leads to the risk matrix in
-Figure~\ref{fig:quantitative}, defined as follows.
+Fig.~\ref{fig:quantitative}, defined as follows.
 \begin{definition}{\bf Risk Matrix --}
-	The risk matrix, as summarized in Table~\ref{tab:5com}, is a function of the three relations
+	The risk matrix, as summarized in Tab.~\ref{tab:5com}, is a function of the three relations
 	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
@@
 
-We now map MAS to CPS and show how this mapping allows us
-to predict cybersecurity weaknesses in a system.
-
 \subsection{From Multi-Agent to Cyber-Physical Systems}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{abftheory.pdf}
+	\includegraphics[width=.7\columnwidth]{abftheory.pdf}
 	\caption{Mapping Epistemological Concepts to (Cyber-Physical) Systems Engineering}
@@
 
-As depicted in Figure~\ref{fig:mashyp}, we relate MAS and CPS as follows.
+As depicted in Fig.~\ref{fig:mashyp}, we relate MAS and CPS as follows.
 \begin{itemize}
@@
 
-\paragraph{Input and Output Ports}
+\paragraph{Input and Output Ports.}
 %\begin{figure}[t]
@@
 
-\paragraph{Port Weaknesses} We now apply the $\abftheory$-framework to
+\paragraph{Port Weaknesses.} We now apply the $\abftheory$-framework to
 list all possible cybersecurity weaknesses of a port.
@@
 	\item \emph{Selective port}, where some information passes the port and
-		part is either:
-	\begin{enumerate}[start=1, label={W5.\arabic*)}]
-		\item Dropped, or 
-		\item Replaced.
-	\end{enumerate}
+    part is either: (W5.1) Dropped or (WP5.2) Replaced.
+	%\begin{enumerate}[start=1, label={W5.\arabic*)}]
+	%	\item Dropped, or 
+	%	\item Replaced.
+	%\end{enumerate}
 \end{enumerate}
@@
 			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
-			(we call \emph{insert} the former and \emph{full drop}
+			(we call \emph{insert} the former, \emph{full drop}
 			the latter case), or
 			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
-			which we call \emph{replace} (i.e. drop and insert).
+			called \emph{replace} (i.e. drop and insert).
 		\item If $\pp{\assertionRegion}{\beliefRegion}$ then
@@
 
-\paragraph{Communication Channels}
-%\begin{figure}[t]
-%	\centering
-%	\includegraphics[width=\columnwidth]{channel.pdf}
-%	\caption{Communication over a Mono-directional Channel}
-%	\label{fig:channel}
-%\end{figure}
+\paragraph{Communication Channels.}
 In this work, we only consider
@@
 the channel). 
-
-We start by considering the difference between a (communication)
+We consider first the difference between a (communication)
 mono-directional channel (channel from now on) and an agent, as we did for the
@@
 The \emph{quality of a mono-directional} channel is defined as the $\rcc$
-relation between the assertions made by the sender and the ones received by the
+relation between the assertions of the sender and the ones received by the
 receiver, i.e. $\Rcc{\assertionRegion_s}{\assertionRegion_r}$.
 
-\paragraph{Channel Weaknesses}
+\paragraph{Channel Weaknesses.}
 Given that a mono-directional bottom-channel is assumed to be perfectly
@@
 
-\paragraph{Cybersecurity Weaknesses -- The RIDI-Hypothesis} 
-It is important to note that all the results of the application of the
+\paragraph{Cybersecurity Weaknesses -- The RIDI-Hypothesis.} 
+All the results of the application of the
 $\abftheory$-framework to channels (the analysis of the RCC relations
@@
 
-\paragraph{Functional Architecture}
+\paragraph{Functional Architecture.}
 A functional architecture takes information as input-beliefs and transforms the
@@
 
-\paragraph{Requirements as Facts}
+\paragraph{Requirements as Facts.}
 During the specification phase, for any agent, channel, port, functional block
@@
 any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
-Figure~\ref{fig:soundness}, facts are definitory rules that define how
+Fig.~\ref{fig:soundness}, facts are definitory rules that define how
 the system shall behave (by specification), while reality may be shown to
 be insecure (i.e. diverging from the expected behavior).
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.3\columnwidth]{sum.pdf}
-	\caption{Example of a Functional Block}
-	\label{fig:sum}
-\end{figure}
-
 As an example, considering a functional block that performs the summation of two
-inputs, as in Figure~\ref{fig:sum}, defined by the requirement $r := b_3 = b_1
+inputs defined by the requirement $r := b_3 = b_1
 + b_2$ for any $b_1$ and $b_2$. The possible relations between the beliefs generated by the behavior of the functional block and
@@
 
-\paragraph{Assertions and Facts}
+\paragraph{Assertions and Facts.}
 The whole reasoning on the relation between beliefs and facts can be duplicated
@@
 into: data-flow-related and functionality-related weaknesses; as in
-Table~\ref{tab:weaknesses}. Functionality-related weaknesses can be seen as a
+Tab.~\ref{tab:weaknesses}. Functionality-related weaknesses can be seen as a
 general formulation of our hypothesis while data-flow-related weaknesses as an
@@
 \centering
-	\begin{tabular}{c C{.35\textwidth} C{.5\textwidth}} 
+	\begin{tabular}{c C{.3\textwidth} C{.55\textwidth}} 
 		\textbf{RCC5} & \textbf{Quantity: Data Flow} & \textbf{Quality: Requirements Adherence}\\
@@
 	PPi& Forwards all the inputs but crafts and inserts new malicious data & The components correctly performs/carries the expected behavior/information when the correct inputs are provided but is subject to input injections \\[.1cm]
-	PO & Selectively drops inputs and inserts new data & Byzantine behavior - occasionally outputs the expected output given the correct inputs.
-Not all the inputs are handled properly, nor all the expected outputs are always generated when correct inputs are given
+	PO & Selectively drops inputs and inserts new data & Byzantine behavior.
+Occasionally outputs the expected output given the correct inputs.
+Not all inputs are handled properly, nor all the expected outputs are always generated when correct inputs are given
 \end{tabular}
@@
 configurations. The engineering of the $\abftheory$-framework for CPS is summarized in the UML
-Class diagram in Figure~\ref{fig:secraclassdiagram} (in appendix).
+Class diagram in Fig.~\ref{fig:secraclassdiagram} (in appendix).
 
@@
 	\centering
-	\includegraphics[width=.9\columnwidth]{v-secra.pdf}
+	\includegraphics[width=.6\columnwidth]{v-secra.pdf}
 	\caption{Cybersecurity Risk Assessment Tool}
@@
 \end{figure}
-As depicted in Figure~\ref{fig:secra}, the cybersecurity risk assessment process
+As depicted in Fig.~\ref{fig:secra}, the cybersecurity risk assessment process
 starts with the definition of the use cases and architectural requirements.  In our process, the specification is manually
@@
 
-\paragraph{Case Studies}
+\paragraph{Case Studies.}
 We report the results of the evaluation of a water level reader (sensor ad-hoc example).
+
 \begin{figure}
-	\centering
-	\includegraphics[width=.5\columnwidth]{eng_cs1.png}
-	\caption{Water level reader deployment diagram}
+%	\centering
+\begin{minipage}[t]{0.3\textwidth}
+\includegraphics[width=\columnwidth]{eng_cs1.png}
+\end{minipage}
+\begin{minipage}[t]{0.7\textwidth}
+\includegraphics[width=\columnwidth]{internal_cs1.png}
+\end{minipage}
+  \caption{Sensor board object (left) and water level reader deployment (right) diagrams.}
+%	\centering
+\label{fig:int_cs1}
 \label{fig:eng_cs1}
 \end{figure}
-As in Figure~\ref{fig:eng_cs1}, we defined 2 agents: sensorInTank and sensorBoard,
+
+As in Fig.~\ref{fig:eng_cs1} (left), we defined 2 agents: sensorInTank and sensorBoard,
 which represent the physical reader that needs to be placed in a tank, and 
@@
 The two components are connected by a wire.
-\begin{figure}
-	\centering
-	\includegraphics[width=.9\columnwidth]{internal_cs1.png}
-	\caption{Sensor Board Object Diagram}
-	\label{fig:int_cs1}
-\end{figure}
-In Figure~\ref{fig:int_cs1}, we report the functional architecture that
+In Fig.~\ref{fig:int_cs1} (right), we report the functional architecture that
 receives the incoming communications from the sensor in the tank, and 
 communicates them encrypted.
-\begin{figure*}
-	\centering
-	\includegraphics[width=.7\textwidth]{results.pdf}
-	\caption{Partial View of the results in the spreadsheet file}
-	\label{fig:results}
-\end{figure*}
-The results, summarized in Figure~\ref{fig:results}, reports 16777216
+The results, summarized in Fig.~\ref{fig:results}, reports 16777216
 possible scenarios in which at least one component diverge from the specification.
@@
 
+\printbibliography
+
 \appendix
+\section{Class Diagram for $\abftheory$-framework}
 The Class Diagram for the Engineering of the $\abftheory$-framework is reported in
-Figure~\ref{fig:secraclassdiagram}. A \emph{specification} of a CPS is viewed
+Fig.~\ref{fig:secraclassdiagram}. A \emph{specification} of a CPS is viewed
 as an aggregation of \emph{architectures} which can describe the functional or
@@
 \end{sidewaysfigure*}
-\printbibliography
 

### 2022-03-18 reports/cimss_2022/main-cut.tex
cimss reduced to 18 pages

@@
 \usepackage[utf8]{inputenc}
-\usepackage[style=numeric]{biblatex}
+\usepackage[style=numeric-comp]{biblatex}
 \usepackage{amsmath}
 \usepackage{amsfonts}
-\addbibresource{../literature.bib}% Syntax for version >= 1.2
+\addbibresource{./literature.bib}% Syntax for version >= 1.2
 \usepackage{graphicx}
@@
 theory should not be conceived by using the principle of
-induction\footnote{Albert  Einstein wrote to Popper: ``[...] and I think (like
+induction\footnote{Einstein to Popper: ``[...] and I think (like
 you, by the way) that theory cannot be fabricated out of the results of
@@
 hypothesis to refer to ``scientific hypothesis'' as a proposed scientific
-theory that has not gone through an extensive series of tests. Instead, we
+theory that has not gone through an extensive series of tests. We
 use ``logical theory'' to refer to a set of formal logical axioms.} may
@@
 hard-drive).  
-%There is also an apparent inconsistency in
-%\autocite{Herley2016unfalsifiability} that we seek to clarify before following
-%(as we agree) the scientific path drawn by Herley: cybersecurity is defined as
-%a set of abstract properties in many formal approaches to the investigation of
-%the cybersecurity of systems.  Let's, for the sake of simplicity, consider the case
-%of a system implementing a cybersecurity protocol (e.g. an authentication protocol).
-%The proof of cybersecurity obtained from the formal verification of the design of
-%the protocol (e.g. with respect to a formalization of the authentication
-%property) is indeed falsifiable against the cybersecurity properties verified.  For
-%example, in the protocol verification community, cybersecurity is often defined
-%as a formalization of the high-level properties confidentiality, integrity, and
-%authentication. The problem in such approaches is not the definition of what
-%cybersecurity is but the generality of the results, since they tackle a
-%specific step\footnote{In fact, engineering models such as the V-model start
-%with the requirements definition, and physical and functional architecture
-%design, before the logic of protocols is even considered.} of the engineering
-%process (of systems) without considering many aspects related to the design or
-%implementation of a protocol that may generate cybersecurity vulnerabilities later
-%in the development process.  An infamous example is the KRAK attack on the WPA2
-%4-way handshake protocol where the author wrote \autocite{krak2020web}: ``our
-%attacks do not violate the security properties proven in formal analysis of the
-%4-way handshake'' but ``the problem is that the proofs do not model key
-%installation.'' The theories underlying the verification of cybersecurity protocols
-%are based on assumptions which non-evidently apply to a general cybersecurity
-%theory; otherwise the assumptions would clearly state the necessary element
-%that needs to be considered (e.g. modeled) in order to obtain a secure system.
-%As an example, the so called Dolev-Yao attacker model\footnote{The Dolev-Yao
-%attacker model can be considered as abstract model of an attacker that can be
-%used for the analysis of protocol specifications.} \autocite{Dolev1983security}
-%only applies to specific instances (often called scenarios) and abstraction of
-%the protocol.  This, in turn, creates a false sense of security since it
-%requires non-justifiable assumptions on the abstraction of the system under
-%verification.
-%More specifically, for the formal security verification of a system in which
-%two or more agents are communicating, a formalized scenario needs to be
-%defined by a modeler who chooses (among others): (i) a scope of the
-%formalization (e.g.  excluding the server that distributes the public key is
-%often done when verifying the security of authentication protocols), (ii) the
-%number of sessions (even tough some approaches do reason on an infinite number
-%of sessions such as \autocite{Escobar2007maudenpa}),\fixnote{Luca}{Nowadays, most
-%tools work for infinite sessions: Maude-NPA, Tamarin, ProVerif, OFMC.} (iii)
-%honesty/dishonesty of the peers (e.g.  in the ASLan++ language
-%\autocite{Oheimb2010aslan++}),\fixnote{Luca}{Nowadays, most tools allow for easy
-%switch of the resulting sessions} and (iv) the abstraction of the
-%cryptographic primitives (e.g.  ProVerif vs CryptoVerif
-%\autocite{Blanchet2017symbolic}).\fixnote{Luca}{This is the only claim that, more
-%or less, still holds, even though more and more tools do cryptoanalysis. The
-%good news is that, I believe, you do not need (i)-(iv) at all, but you can
-%simply say that by the nature of modeling a specification, you are abstracting
-%away certain things (e.g., cryptography, actual message content, PKI, etc.),
-%and the vulnerabilities might be hidden there} While many tools and theories
-%improved those issues, there is no agreement on which should be the definitive
-%approach (if any).\fixnote{Luca}{I don't think that this is important. In many
-%fields of research you don't have a single definitive approach and it is good
-%that it is so. I would remove this sentence, but only focus on the next point,
-%which is the only important one of this discussion of protocols} 
-%More importantly, some of the choices made by the
-%modelers/engineers using those formal approaches may completely change the
-%results of the formal verification of the system (an interesting example on how
-%difficult it is to even just compare the approaches is given in
-%\autocite{Cremers2009comparing}). 
-%\begin{figure}[t]
-%	\centering
-%	\includegraphics[width=.7\columnwidth]{protocol-example.pdf}
-%	\caption{Abstraction of an ad-hoc esemplificative protocol execution}
-%	\label{fig:protocol-example}
-%\end{figure}
-%As an illustrative example, in Fig.~\ref{fig:protocol-example}, under the
-%perfect cryptography assumption (i.e. assuming that the cryptographic
-%primitives have no cybersecurity flaw), and assuming that no violation to any
-%cybersecurity property occurs after message I), the freedom of choosing the
-%scope determines that the flaws related to the dishonest impersonification of
-%the Server may or may not be considered in the verification process.  This
-%choice has a tremendous impact on the focus and findings of the verification of
-%the cybersecurity of the protocol.  While this may seem to turn upon minutiae
-%and foreseeable, this highlights the false sense of security that may derive
-%from a non-falsifiable theory of system cybersecurity.  In other words, 
-%As
-%Herley mentions in \autocite{Herley2016usenixvideo}, the problem is in the way
-%we reason on cybersecurity, and we can see it in the way we reply to simple
-%cybersecurity (foundamental) questions.  Let us consider the general question
-%``what is cybersecurity?''; 
-%%as in \autocite{hintikka2013inquiry}, we can
-%%translate the wh-question into an equivalent existentially quantified formula.
-%%But if the formula is $\exists P. P(s)\rightarrow Secure(s)$, 
-%%(where we ask if
-%if we detail the question as:
-%there exists a predicate $P$ such that, given any arbitrary system $s$, when
-%$P$ holds for $s$ then $s$ can be considered secure?
-%%), 
-%When the answer is (for
-%example) ``if the system is confidential, the system is secure'' the reasoning %formula
-%becomes tautological. In fact, $P$ and $Secure$ can be considered as
-%semantically equivalent (cybersecurity is confidentiality and vice-versa).
 Instead of starting from reasoning on what makes a system secure or insecure,
@@
 \begin{figure}[t]
-	\centering
-	\includegraphics[width=.35\columnwidth]{causality.pdf}
-	\caption{Etiology of cybersecurity}
+\begin{minipage}[t]{0.4\textwidth}
+	\includegraphics[width=\columnwidth]{causality.pdf}
+\end{minipage}
+\begin{minipage}[t]{0.6\textwidth}
+	\includegraphics[width=.9\columnwidth]{abftheory.pdf}
+\end{minipage}
+  \caption{Etiology of cybersecurity (left). Mapping Epistemological Concepts to (Cyber-Physical) Systems Engineering (right).}
 	\label{fig:causality}
+	\label{fig:mashyp}
 \end{figure}
 
-As depicted in Fig.~\ref{fig:causality}, in order to define a theory on cybersecurity we may
-say that the presence of vulnerabilities is a necessary condition in order to
+As in Fig.~\ref{fig:causality}, order to define a theory on cybersecurity we may
+say that the presence of vulnerabilities is a necessary condition to
 cause an attack in the system.  Those vulnerabilities are, in turn, caused by
 the presence of weaknesses in the system.  Weaknesses are errors in the design
-or implementation of a system.  Therefore, a theory on cybersecurity should
+or implementation of a system and a theory on cybersecurity should
 first predict the errors in a system design.
@@
 		\emph{cybersecurity requirements} are identified.	
-    %\item \emph{Design Refinement}, where the physical and functional
-	%	architectures are extended with \emph{mitigations} for the
-	%	weaknesses. Mitigations at this step details the semantics of
-	%	the functional blocks (behaviors), the data flow, and communication logic.
-	%\item \emph{Verification of Security Requirements}, where the data flow, behaviors, and communication protocols are verified against the cybersecurity requirements.
 \end{enumerate}
@@
 part-whole relations) instead of the taxonomies (categorization based on
-discrete sets) such as the one provided in \autocite{NIST2020NVD,MITRE2020CVE}
+discrete sets) such as the one provided in \autocite{NIST2020NVD,CVE}
 so that we don't need to rely on a scoring system (such as the CVSS) to assign
@@
 emerge from the mathematical formulation of our cybersecurity hypothesis.
-
-A mereotopology, as defined e.g. in \autocite{Rachavelpula2017mereotopology},
+A mereotopology, as defined in \autocite{Rachavelpula2017mereotopology},
 is a mathematical structure where the basic relation between regions
@@
 the possible relations between the different regions.
-For completeness, we report in the next paragraph 
-the calculation done by the authors in \autocite{Santaca2016abf}.
-
-\paragraph{Number of different configurations of a system.}
+For completeness, we report in Appendix~\ref{app:agcalc}
+some details on the calculation done by the authors in \autocite{Santaca2016abf}.
 The general formula to calculate the number of different types of agents is
 $r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
-between $n$ different regions. Therefore, $r^{\binom{n}{k}}$
-expresses the number of permutation of $r$
-relations over ${\binom{n}{k}}$ elements with repetitions, 
-with ${\binom{n}{k}}$ being the number of
-$k$-ary combinations of $n$ regions.
-In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
+between $n$ different regions. In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
 ($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
-considered in the RCC are binary.  Hence, using RCC5 (with five different
-spatial relations) over three regions, we can theoretically define up to 125
-different type of agents. However, only 54 of the 125 (as showed in
-\autocite{improvingRCC}) combinations are topologically correct with respect to
-the definition of the relations of RCC5. Generalizing to all the RCCs:
+considered in the RCC are binary.  Hence, we have up to 125
+different type of agents but only 54 of the 125 (as showed in
+\autocite{improvingRCC}) combinations are topologically correct. 
+Generalizing to all the RCCs:
 
@@
 \item \emph{RCC5} --- theoretical: $5^3=125$, correct: 54.
-%\item \emph{RCC8} --- theoretical: $8^3=512$, correct: 193.
 \end{itemize}
@@
 %allow us to consider 3.5 times more different agents.
-In the quantitative evaluation of a single agent, depicted in Fig.~\ref{fig:quantitative},
+In the quantitative evaluation of a single agent, as in Fig.~\ref{fig:quantitative},
 we argue that only 1 configuration represents the nominal (expected) behavior 
@@
 representation of the system.
+
 \begin{figure}[t]
-	\centering
-	\includegraphics[width=.8\columnwidth]{quantitative.pdf}
-	\caption{Cybersecurity risk for a single agent}
+\begin{minipage}[t]{0.5\textwidth}
+\includegraphics[width=.9\columnwidth]{quantitative.pdf}
+\end{minipage}
+\begin{minipage}[t]{0.5\textwidth}
+\includegraphics[width=.9\columnwidth]{soundness.pdf}
+\end{minipage}
+  \caption{Cybersecurity risk for a single agent (left). Example relation between facts, and assertions and beliefs.}
 \label{fig:quantitative}
+\label{fig:soundness}
 \end{figure}
@@
 \centering
-\begin{adjustbox}{width=.7\columnwidth}
+\begin{adjustbox}{width=.6\columnwidth}
 \begin{tabular}{r||c|c|c|c|c} 
@@
 degrees of non-soundness.  A similar argument can be done for completeness.
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.5\columnwidth]{soundness.pdf}
-	\caption{Example relation between facts, and assertions and beliefs}
-	\label{fig:soundness}
-\end{figure}
-
-\paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration.} In order
-to reason on the relation between assertions and beliefs we first need to
-consider that, by definition, assertions are defined as transfer of information
-between two agents.  Therefore, as depicted in
-Fig.~\ref{fig:system-agent}, an agent has two main categories of assertions,
+
+\paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration.} 
+By definition, assertions are defined as transfer of information
+between two agents. An agent has two main categories of assertions,
 input and output assertions.  Given an agent $a$ and a collection of asserted
@@
 We shall consider
-two pairs of regions:%\footnote{``I am not asserting, as Lotze did, that a
-%relation between $X$ and $Y$ consists of a quality in $X$ and a quality in $Y$
-%-- a view which I regard as quite indefensible. -- I assert that a relation $Z$
-%between $X$ and $Y$ involves the existence in $X$ of the quality ``having the
-%relation $Z$ to $Y$'' so that a difference of relations always involves a
-%difference in quality, and a change of relations always involves a change of
-%quality.'' --  Ellis J. McTaggart, The Unreality of
-%Time \autocite{Mctaggart1908unreality}}
+two pairs of regions:
 \begin{itemize}
@@
 but a full comparison with the CWE is out of the scope of this article.
-
 We can now define what a secure system is (with respect to
@@
 \subsection{From Multi-Agent to Cyber-Physical Systems}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.7\columnwidth]{abftheory.pdf}
-	\caption{Mapping Epistemological Concepts to (Cyber-Physical) Systems Engineering}
-	\label{fig:mashyp}
-\end{figure}
-
-As depicted in Fig.~\ref{fig:mashyp}, we relate MAS and CPS as follows.
+As in Fig.~\ref{fig:mashyp}, we relate MAS and CPS as follows.
 \begin{itemize}
@@
 \paragraph{Input and Output Ports.}
-%\begin{figure}[t]
-%	\centering
-%	\includegraphics[width=\columnwidth]{IOports.pdf}
-%	\caption{Input and Output Port of an agent}
-%	\label{fig:ioports}
-%\end{figure}
 Since the $\abftheory$-framework is a theory of agents, we could consider ports as
@@
 excluded by the $\abftheory$-framework, it would obviously lead to an infinite
-nested structure of ports. In order to avoid this infinite structure, we assume
+nested structure of ports. To avoid this infinite structure, we assume
 that a port doesn't require any other means to transfer information from/to a channel
@@
 assertions received or sent and the belief, i.e.
-$\Rcc{\rassert{s}{a}}{\beliefRegion_a}$ for the input-port or
-$\Rcc{\rassert{r}{*}}{\beliefRegion_a}$ for the output-port (where $*$ is a wildcard for an agent).  A port is, in
+$\Rcc{\rassert{s}{a}}{\beliefRegion_a}$ for input-port or
+$\Rcc{\rassert{r}{*}}{\beliefRegion_a}$, for output-port (where $*$ is a wildcard for an agent).  A port is, in
 fact, syntactic sugar to express the relation between assertions and beliefs.
@@
 that a port always forwards the information without modifying it in any way.
-This is assumed because we define a port as if the RCC equality relation holds
+This is assumed since we defined a port as if the RCC equality relation holds
 between the input/output assertions and the input/output beliefs. In contrast,
-assuming any RCC relation, between inputs and outputs of a port, which is not
-an equality can be considered as generating a weaknesses.
+assuming any \emph{other} RCC relation between inputs and outputs of a port can be considered as generating a weaknesses.
 
 \paragraph{Port Weaknesses.} We now apply the $\abftheory$-framework to
-list all possible cybersecurity weaknesses of a port.
+list all possible cybersecurity weaknesses of a port. From our definition of ports, it follows that
 \begin{theorem}{Port Weaknesses.}
-From our definition of ports, it follows that there exist only the following \emph{six}
+There exist only the following \emph{six}
 types of weaknesses, generating six types of insecure port in RCC5:
@@
 		become belief of the agent).
-	\item \emph{Insertion port}, where some new information is transferred along with 
+	\item \emph{Insertion port}, where new information is transferred along with 
 		the information incoming from a channel, and then sent to
-		the intended recipient (agent).
+		the recipient (agent).
 	\item \emph{Injection port}, where part of the incoming information
@@
 functional blocks are constituents of the functional architecture as described
-afterwards.  In order to define a functional block without encountering an
+afterwards.  To define a functional block without encountering an
 infinitely recursive definition, we must reach the same conclusions as for the
-channel. Therefore, describing the information as flowing over a channel or in
+channel. So, describing the information as flowing over a channel or in
 a functional block is purely syntactic sugar.
-
 We can summarize these results by saying that the relations between assertions
@@
 \end{itemize}
-%We can also generalize the weaknesses related to the data flow (i.e. the flow
-%of information as assertions or beliefs) as the relation between output-input
-%beliefs between functional blocks, output-input assertions of a channel,
-%inputs-outputs of a port, as follows.
-%\begin{itemize}
-%	\item \emph{equal}, expected nominal behavior
-%	\item \emph{discrete-from}, drops all the inputs and inserts new malicious data
-%	\item \emph{proper part}, selectively drops inputs
-%	\item \emph{proper part inverse}, forwards all the inputs but crafts and
-%	      inserts new malicious data
-%	\item \emph{partial overlap}, selectively drops inputs and inserts new data.
-%\end{itemize}
-
-%\begin{figure}[t]
-%	\centering
-%	\includegraphics[width=.9\columnwidth]{engineering_relations.pdf}
-%	\caption{Relations between Ports and Channel}
-%	\label{fig:atom}
-%\end{figure}
 
@@
 	\item $\po{\beliefRegion}{\factRegion}$ the component has a Byzantine
-		behavior where occasionally outputs the expected output given
-		the correct inputs. Not all the inputs are handled properly,
+		behavior. Not all the inputs are handled properly,
 		nor all the expected outputs are always generated when correct
 		inputs are given.
-	\item $\pp{\beliefRegion}{\factRegion}$ part of the expected outputs are not
-	        generated in response to the correct
+	\item $\pp{\beliefRegion}{\factRegion}$ some expected outputs are not
+	        generated with the correct
 	        inputs
@@
 In other words, any requirement is defined as a fact since they must be true in
-any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
+any design or implementation. As in Section~\ref{sec:theory} and depicted in
 Fig.~\ref{fig:soundness}, facts are definitory rules that define how
@@
 		where $\beliefRegion_3$ represents the region of the
-		outputs of the functional block $sum$ while the region
-		$\beliefRegion_{1+2}$ represents the expected
+		outputs of $sum$ while
+		$\beliefRegion_{1+2}$ the expected
 		outputs of an ideal implementation of the requirement $r$.
-		Therefore, the functional block correctly implements the
+		The functional block correctly implements the
 		requirements.
@@
 		$\dr{\beliefRegion}{\factRegion}=\dr{\beliefRegion_3}{\beliefRegion_{1+2}}$,
-		the function block never implements the requirements.
+		the functional block do not implement the requirements.
 	\item
-		$\pp{\beliefRegion}{\factRegion}=\pp{\beliefRegion_3}{\beliefRegion_{1+2}}$, there exist some inputs for which
-		the output is incorrect.
-	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_{1+2}}$, there exist some outputs that are not the result of the summation of the inputs but given the correct inputs the function always outputs correctly.
+		$\pp{\beliefRegion}{\factRegion}=\pp{\beliefRegion_3}{\beliefRegion_{1+2}}$, some inputs produce incorrect output.
+	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_{1+2}}$, some outputs are not the result of the summation of the inputs but with the correct inputs the function outputs correctly.
 	\item
 		$\po{\beliefRegion}{\factRegion}=\po{\beliefRegion_3}{\beliefRegion_{1+2}}$,
-		the component has a Byzantine behavior where occasionally
-		outputs the expected output given the correct inputs. Not all
+		Byzantine behavior where occasionally
+		outputs are produced with the correct inputs. Not all
 		the inputs are handled properly, nor all the expected outputs
-		are always generated when correct inputs are given.
+		are generated when correct inputs are given.
 \end{itemize}
@@
 into: data-flow-related and functionality-related weaknesses; as in
-Tab.~\ref{tab:weaknesses}. Functionality-related weaknesses can be seen as a
-general formulation of our hypothesis while data-flow-related weaknesses as an
+Tab.~\ref{tab:weaknesses}. Functional weaknesses can be seen as a
+general formulation of our hypothesis, while data-flow weaknesses as an
 application of our hypothesis to components with defined behavior/requirements.
@@
 \centering
-	\begin{tabular}{c C{.3\textwidth} C{.55\textwidth}} 
+	\begin{tabular}{c C{.3\textwidth} C{.6\textwidth}} 
 		\textbf{RCC5} & \textbf{Quantity: Data Flow} & \textbf{Quality: Requirements Adherence}\\
@@
 		EQ & Expected/Nominal & Expected/Nominal\\[.1cm]
-	DR & Drops all the inputs and inserts new malicious data & The component never performs/carries the expected behavior/information\\[.1cm]
+	DR & Drops all inputs and inserts new data & The component never performs/carries the expected behavior/information\\[.1cm]
 	PP & Selectively drops inputs & part of the expected outputs are not generated in response to the correct inputs \\[.1cm]
@@
 Occasionally outputs the expected output given the correct inputs.
-Not all inputs are handled properly, nor all the expected outputs are always generated when correct inputs are given
+Not all inputs are handled properly, nor all expected outputs always generated on correct inputs
 \end{tabular}
@@
 \subsection{Cybersecurity Risk Assessment}\label{sec:secra}
-In order to test our hypothesis we implemented a tool-chain (open-source under the
-AGPLv3 license and available at \autocite{v-research2020cybersecurity-anonymous}) for the
-identification of weaknesses and the precise calculation of potential insecure
-configurations. The engineering of the $\abftheory$-framework for CPS is summarized in the UML
-Class diagram in Fig.~\ref{fig:secraclassdiagram} (in appendix).
-
-\begin{figure}
+\begin{figure}[t]
 	\centering
-	\includegraphics[width=.6\columnwidth]{v-secra.pdf}
+	\includegraphics[width=.5\columnwidth]{v-secra.pdf}
 	\caption{Cybersecurity Risk Assessment Tool}
@@
 \end{figure}
-As depicted in Fig.~\ref{fig:secra}, the cybersecurity risk assessment process
+To test our hypothesis we implemented a tool-chain (open-source with
+AGPLv3 license, available at \autocite{v-research2020cybersecurity}) for the
+identification of weaknesses and the calculation of potential insecure
+configurations. The engineering of the $\abftheory$-framework for CPS is summarized in the UML
+Class diagram in Fig.~\ref{fig:secraclassdiagram} (in appendix). 
+As in Fig.~\ref{fig:secra}, the cybersecurity risk assessment process
 starts with the definition of the use cases and architectural requirements.  In our process, the specification is manually
@@
 
-The tool generates a graph-like internal structure which represents
-the specification (called $\abftheory$-graph). The $\abftheory$-graph defines the system as
+The tool generates a graph-like structure which represents
+the specification ($\abftheory$-graph). The $\abftheory$-graph defines the system as
 a number of regions of assertions, beliefs, and facts. Those regions
@@
 identifies all potential weaknesses. 
-The internal structures of the tool
-can be viewed as PDF and the results are reported into an spreadsheet file.
+The $\abftheory$-graph can be viewed as PDF and the results are reported into an spreadsheet file.
 The spreadsheet file also reports the total number of configurations as indicating
 the cybersecurity risk associated to the specification. A user can change the status
-of each weakness in the spreadsheet file, from the initial (default) status (open)
-to ``mitigated''. As a consequence, the risk is re-calculated on-the-fly, i.e.
+of each weakness in the spreadsheet file from the default status (open)
+to ``mitigated'' and the risk is re-calculated on-the-fly, i.e.
 without the need of running the tool again, based on annotations and formulas 
@@
 In our approach, cybersecurity requirements are not imposed by the
-specification but are automatically extracted by our tool as mitigations to potential weaknesses.
-Those weaknesses are related to the different insecure configurations 
+specification but are automatically extracted by our tool as mitigations to potential weaknesses,
+which are related to the insecure configurations 
 of the specified system.
 
-\paragraph{Case Studies.}
-We report the results of the evaluation of a water level reader (sensor ad-hoc example).
-
-\begin{figure}
-%	\centering
+\paragraph{Case Study.}
+\begin{figure}[t]
 \begin{minipage}[t]{0.3\textwidth}
@@
   \caption{Sensor board object (left) and water level reader deployment (right) diagrams.}
-%	\centering
 \label{fig:int_cs1}
@@
 \end{figure}
-
+We report here the results of the evaluation of a water level reader (sensor ad-hoc example). 
 As in Fig.~\ref{fig:eng_cs1} (left), we defined 2 agents: sensorInTank and sensorBoard,
-which represent the physical reader that needs to be placed in a tank, and 
+as the physical reader that needs to be placed in a tank, and 
 the board that interprets the readings and outputs them as signals.
@@
 In Fig.~\ref{fig:int_cs1} (right), we report the functional architecture that
-receives the incoming communications from the sensor in the tank, and 
+receives the incoming communications from the sensor in the tank and 
 communicates them encrypted.
-The results, summarized in Fig.~\ref{fig:results}, reports 16777216
-possible scenarios in which at least one component diverge from the specification.
+The tool (App.~\ref{app:results}) reports 16777216
+scenarios in which at least one component diverge from the specification.
 
@@
 cybersecurity-related issues are not linked to the maliciousness of an agent
-but to the vagueness in the design processes.  We also provided a prototype
-tool for the quantitative estimation of the cybersecurity risk, based on a UML
-model of the system under assessment.
-The cybersecurity verification and test-case generation will be the focus of our next steps.
+but to the vagueness in the design processes.  We provided a prototype
+tool for the quantitative estimation of the cybersecurity risk based on a UML
+model of a system.
+The verification and test-case generation will be our next steps.
 
@@
 
+\section{Number of different configurations of a system.}\label{agcalc}
+The general formula to calculate the number of different types of agents is
+$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
+between $n$ different regions. Therefore, $r^{\binom{n}{k}}$
+expresses the number of permutation of $r$
+relations over ${\binom{n}{k}}$ elements with repetitions, 
+with ${\binom{n}{k}}$ being the number of
+$k$-ary combinations of $n$ regions.
+In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
+($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
+considered in the RCC are binary.  Hence, using RCC5 (with five different
+spatial relations) over three regions, we can theoretically define up to 125
+different type of agents. However, only 54 of the 125 (as showed in
+\autocite{improvingRCC}) combinations are topologically correct with respect to
+the definition of the relations of RCC5. Generalizing to all the RCCs:
+
+\begin{itemize}%[nosep]
+\item \emph{RCC3} --- theoretical: $3^3=27$,  correct: 15.
+\item \emph{RCC5} --- theoretical: $5^3=125$, correct: 54.
+\end{itemize}
+
+\section{Overview of the Results of the Tool}\label{app:results}
+In Fig.~\ref{fig:results} we show a screenshot of the 
+results reported by our tool.
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{results.pdf}
+	\caption{Partial View of the results in the spreadsheet file}
+	\label{fig:results}
+\end{figure}
+
 \end{document}

### 2022-03-18 reports/cimss_2022/main-cut.tex
anonymize authors in cimss submission

@@
 
-\author{Michele Ambrosi\inst{1} \and
-Francesco Beltramini\inst{1} \and Federico De Meo\inst{1} \and Olivero Nardi\inst{1} \and Mattia Pacchin\inst{1} \and Marco Rocchetto}
-
-\authorrunning{M. Ambrosi et al.}
-
-\institute{V-Research\\
-Verona, Italy\\
-\email{knowledgezero@v-research.it}}
+\author{}
+
+\authorrunning{}
+
+\institute{}
 \maketitle
@@
 To test our hypothesis we implemented a tool-chain (open-source with
-AGPLv3 license, available at \autocite{v-research2020cybersecurity}) for the
+AGPLv3 license, available at \autocite{v-research2020cybersecurity-anonymous}) for the
 identification of weaknesses and the calculation of potential insecure

### 2022-03-21 reports/cimss_2022/main-cut.tex
implement fed's feedback

@@
 	disable, steal or gain unauthorized access to or make
-	unauthorized use of an asset''; where an{Asset} is
+	unauthorized use of an asset''; where an \emph{Asset} is
 	``anything that has value to the organization''.
 	We do not consider ethical hackers as attacking a system. 
-	In fact, we consider the term{hack} as
+	In fact, we consider the term \emph{hack} as
 	non-malicious (as, e.g. in \autocite{Stallman2002hacker}).
@@
 
-As in Fig.~\ref{fig:causality}, order to define a theory on cybersecurity we may
+As in Fig.~\ref{fig:causality}, in order to define a theory on cybersecurity we may
 say that the presence of vulnerabilities is a necessary condition to
@@
 \end{minipage}
-  \caption{Cybersecurity risk for a single agent (left). Example relation between facts, and assertions and beliefs.}
+  \caption{Cybersecurity risk for a single agent (left). Example relation between facts, and assertions and beliefs (right).}
 \label{fig:quantitative}
@@
 in \autocite{MITRE2020CWEresearch}, while the other classes of weaknesses do not
-have a direct counterpart in our hypothesis; we believe they can be sees as sub-classes,
+have a direct counterpart in our hypothesis; we believe they can be seen as sub-classes,
 but a full comparison with the CWE is out of the scope of this article.
@@
 The definition of Risk leads to the risk matrix in
-Fig.~\ref{fig:quantitative}, defined as follows.
+Tab.~\ref{fig:quantitative}, defined as follows.
 \begin{definition}{\bf Risk Matrix --}
@@
 an \emph{infinite regress}, as described in the following. While a channel transfers information between agents, and
-a functional architecture elaborate information, a port is simply a connector between
+a functional architecture processes information, a port is simply a connector between
 a channel and a functional architecture. One, however, may argue that a similar
@@
 	DR & Drops all inputs and inserts new data & The component never performs/carries the expected behavior/information\\[.1cm]
-	PP & Selectively drops inputs & part of the expected outputs are not generated in response to the correct inputs \\[.1cm]
+	PP & Selectively drops inputs & Part of the expected outputs are not generated in response to the correct inputs \\[.1cm]
 	PPi& Forwards all the inputs but crafts and inserts new malicious data & The components correctly performs/carries the expected behavior/information when the correct inputs are provided but is subject to input injections \\[.1cm]

### 2022-04-24 reports/cimss_2022/main-cut.tex
Oliviero's feedback

@@
 
+\newcommand{\ON}[1]{\footnote{\textcolor{blue}{[[Oliviero: #1]]}}}
+
 \begin{document}
@@
 	of the causality chain that leads to cybersecurity attacks.  We
-	formulate an hypothesis that we
+	formulate a hypothesis that we
 	use to predict the weaknesses in the architectural design of an
@@
 theory should not be conceived by using the principle of
-induction\footnote{Einstein to Popper: ``[...] and I think (like
+induction,\footnote{Einstein to Popper: ``[...] and I think (like
 you, by the way) that theory cannot be fabricated out of the results of
-observation, but that it can only be invented.'' \autocite{Popper1959logic}},
+observation, but that it can only be invented.'' \autocite{Popper1959logic}}
 and that empirical experiments should be considered as the only evidence to
@@
 security'' but rather we reason on any abstract system (so that our scientific
-hypothesis\footnote{In the reminder of this paper, we will use the word
+hypothesis\footnote{In the remainder of this paper, we will use the word
 hypothesis to refer to ``scientific hypothesis'' as a proposed scientific
@@
 use ``logical theory'' to refer to a set of formal logical axioms.} may
-be tested in any sound implementation such as networks, mechanical, cyber, or
+be tested in any sound implementation such as a network, a mechanical, cyber, or
 cyber-physical system, or even a single computer or a single device such as a
 hard-drive).  
-Instead of starting from reasoning on what makes a system secure or insecure,
-we reason on what causes insecurities. 
+Instead of starting from reasoning about what makes a system secure or insecure,
+we reason about what causes insecurities. 
 We focus on insecurities only caused by
@@
 With our approach, a list of weaknesses emerges from the mathematical
-formulation of a system in a framework called $\abftheory$-framework, and
-predicts 4 main classes of weaknesses. Those classes of weaknesses are used to
+formulation of a system in a framework called $\abftheory$-framework\ON{Qui, direi o "framework called ABF" o "framework called the ABF-framework"}, and
+predicts 4 main classes of weaknesses. Those classes of weaknesses\ON{rimuoverei ripetizione di "of weaknesses"} are used to
 calculate all the insecurity configurations of all the components of a system,
@@
 any given system. Our hypothesis can be falsified by means of experiments,
-testing if all the predicted weaknesses are present in the system under test,
+testing if all the predicted weaknesses are present in the system under consideration,
 or testing if other (not predicted by our hypothesis) weaknesses are present.
@@
 \begin{enumerate}
-	\item \emph{System Specification}: the \emph{functional and physical requirements} are defined.
-	\item \emph{Architecture Design}: the specification is structured into \emph{functional and physical architectures}.
+	\item \emph{System Specification}: the \emph{functional and physical requirements} \ON{proporrei di togliere l'and dal corsivo} are defined.
+	\item \emph{Architecture Design}: the specification is structured into \emph{functional and physical architectures}\ON{same as the above}.
 	\item \emph{Cybersecurity Risk Assessment}: potential \emph{weaknesses} (errors) and
@@
 
-\subsection{Mereo-topological Reasoning (2.5)}
+\subsection{Mereo-topological Reasoning}
 \begin{table*}[t]
@@
 
-As in \autocite{Santaca2016abf}, we define agents as a meronomy (an
+As in \autocite{Santaca2016abf}, we define agents\ON{so far, we spoke about systems, from now we focus on agents.
+Shall we briefly motivate why we think of systems in term of agents?
+Or, at least, say that we use the two terms intercambiabilmente?
+Later on in the paper, we do say "agent (i.e.
+system)", but that comes much after this, and so the text inbetween might be puzzling to some.}\ON{Further note: In section 4.1 we explicitly state this.
+Perhaps shall we here just say "we consider a CPS as a hierarchy of agents, as we will detail later in Section 4.1"?
+(or something similar)} as a meronomy (an
 hierarchy of Part-Whole relations) over the constituents that we previously defined
@@
 i.e. based on the definition of parthood relation between \emph{parts}.  Due to
-the necessity of considering different relations between parts (as we'll show afterwards)
+the necessity of considering different relations between parts (as we will show afterwards)
 we extend the mereology to a
@@
 \begin{definition}{\bf System State --}\label{def:opsystem}
-	A CPS system (or a sub-system) state is defined as the tuple
-	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
-	where $\assertionRegion,\beliefRegion,\factRegion$, are regions of
-	assertions, beliefs (i.e. the beliefs generated by the behavior), and
-	facts expressed as requirements respectively.
+	A CPS system (or a sub-system) state is defined as the\ON{the->a} tuple
+	$s=\langle\rcc(\factRegion,\beliefRegion),\,\rcc(\factRegion,\assertionRegion),\,\rcc(\beliefRegion,\assertionRegion)\rangle$,
+	where $\assertionRegion$,$\beliefRegion$, and $\factRegion$ are regions of
+	assertions, beliefs (i.e. the beliefs generated by the behavior)\ON{is "generated" the right word here? Anyhow, the bracketed part is not super clear to me.}, and
+	facts respectively, expressed as requirements.
 \end{definition}
 
-As in \autocite{Santaca2016abf}, it follows that, defining
-a system with a fixed number of regions, there exists
+\ON{Shouldn't we introduce these terms here? A, B, F? I mean even with short sentences: assertions are the exchanged data, facts requirements, beliefs are the behaviour...}
+
+As in \autocite{Santaca2016abf}, it follows that, by defining
+a system with \ON{with->as?} a fixed number of regions, there exists
 an upper-bound to the number of possible configuration of a system, defined by
 the possible relations between the different regions.
-For completeness, we report in Appendix~\ref{app:agcalc}
+For completeness, we report in Appendix~\ref{app:agcalc}\ON{here}
 some details on the calculation done by the authors in \autocite{Santaca2016abf}.
@@
 between $n$ different regions. In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
-($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
+($\assertionRegion$, $\beliefRegion$, and $\factRegion$), and all the relations
 considered in the RCC are binary.  Hence, we have up to 125
-different type of agents but only 54 of the 125 (as showed in
+different types of agents but only 54 of the 125 (as showed in
 \autocite{improvingRCC}) combinations are topologically correct. 
@@
 of the agent while the other configurations are either impossible to 
-implement or diverge from the intended nominal behavior. We note 
+implement or diverge from the intended nominal behavior.\ON{Here, I would explicitly say that what we mean is that, only when the assertiosn, beliefs and facts are equal, then the system is behaving as excepted.
+That is, I would explicitly say is that the only nominal state is EQ,EQ,EQ.
+I mean, it is in the table, but a bit implicit to me.} We note
 that the numbers reported here do not consider the details of the
@@
 
-\subsection{Qualitative Evaluation of Agent Space in $\abf$ (3)}\label{sec:agentspace}
+\subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}
 \begin{table}[t]
@@
 \caption{RCC5 composition table over 3 regions generates the ideal risk matrix
-(green the low risk, red the high risk state, and a gradient
-of medium risk states). 
+(green the low-risk state, red the high-risk state, and a gradient
+of intermediate risk states). 
 \all{$\assertionRegion$}{$\factRegion$} = \{\dr{$\assertionRegion$}{$\factRegion$}, \po{$\assertionRegion$}{$\factRegion$}, \pp{$\assertionRegion$}{$\factRegion$}, \ppi{$\assertionRegion$}{$\factRegion$}, \eq{$\assertionRegion$}{$\factRegion$}\}
@@
 While a quantitative analysis reveals how many possible configurations of an
-agent (i.e. a system) exist w.r.t. the $\abftheory$-framework (i.e. $54/125$ in RCC5), a
+agent (i.e. a system) exist w.r.t. the $\abftheory$-framework (i.e.\ON{e.g.?} $54/125$ in RCC5), a
 qualitative analysis of the different configurations describe the
@@
 We first analyze the relations between each pair of
-regions (i.e.  $\abf$), and then we categorize the different
-agents as tuple of the three regions.  For the sake of simplicity, soundness is
-opposed to non-soundness in the following, however, with the RCC one should
+regions (i.e.  $\abf$), and then categorize different
+agents as tuples of the three regions.\ON{but we don't actually do this, right?
+Maybe I missed it.
+If not, we should remove this sentence.
+If yes, then we should clarify this.
+(At least to me :-))}  For the sake of simplicity, soundness is
+opposed to non-soundness in the following, however, with the RCC, one should\ON{should->could?}
 consider different ``degrees'' of non-soundness.  For example, in RCC5, if we
 consider EQ between two regions as representing soundness, DR over the same
-regions represents non-soundness; while PP, PO, PPi represents the different
+regions represents ``total'' non-soundness; while PP, PO, and PPi each represent different
 degrees of non-soundness.  A similar argument can be done for completeness.
@@
 By definition, assertions are defined as transfer of information
-between two agents. An agent has two main categories of assertions,
+between two agents.\ON{first time we say this!} An agent has two main categories of assertions,
 input and output assertions.  Given an agent $a$ and a collection of asserted
@@
 are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. With a slight abuse
-of notation, in the text we simply write $\assertionRegion_{s\rightarrow a}$ when we abstract
+of notation,\ON{alternative: in the text we drop the $\Phi$ when the content of the assertions is not relevant.} in the text we simply write $\assertionRegion_{s\rightarrow a}$ when we abstract
 away from the content of the assertion (i.e. $\Phi$).
@@
 		handled in the functional specification (beliefs) the
-		specification is complete.
+		specification is complete.\ON{Do we perhaps mean "sound"? We used sound above}
 
@@
 $\Rcc{\rassert{s}{a}}{\beliefRegion_a}$ for input-port or
-$\Rcc{\rassert{r}{*}}{\beliefRegion_a}$, for output-port (where $*$ is a wildcard for an agent).  A port is, in
+$\Rcc{\rassert{r}{*}}{\beliefRegion_a}$, for output-port (where $*$ is a wildcard for an agent)\ON{should this $r$ be an $a$?}.  A port is, in
 fact, syntactic sugar to express the relation between assertions and beliefs.
@@
 \paragraph{Port Weaknesses.} We now apply the $\abftheory$-framework to
-list all possible cybersecurity weaknesses of a port. From our definition of ports, it follows that
-\begin{theorem}{Port Weaknesses.}
+list all possible cybersecurity weaknesses of a port. From our definition of ports, the following holds.
+\begin{theorem}[Port Weaknesses]
 There exist only the following \emph{six}
@@
 	an assertion is written $A\neq\emptyset$, while a region corresponding to an empty assertion
-	as $A=\emptyset$).
+	as $A=\emptyset$).\ON{Do we use the weight/$|A|$ thing in the proof? I don't see it... If not, should we get rid of it?}
 	\begin{enumerate}
@@
 defined as agents with the pre-defined behavior (i.e. defined in an axiomatic
-way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
+way) of forwarding any input-assertion as output-assertion, without modifying it.\footnote{Nothing
 prevents us from introducing additional constraints to the channel as storing
 assertions that are transferred over the channel, or filter out some
-input-assertions.}.
+input-assertions.}
 
@@
 	A mono-directional channel between two agents ($s \rightarrow r$) is
-	defined as an agent whose behavior is (dogmatically) defined as: to
+	defined as\ON{remove defined as} an agent whose behavior is (dogmatically) defined as: to
 	forward any assertion received from $s$ over an input-port, to  the
@@
 weaknesses of the input and output port; therefore there exist $(7^2)-1=48$
-theoretical configurations ($7^2$ because there are $6$ insecure types of port,
+theoretical configurations ($7^2$ because there are $6$ insecure types of port,\ON{refer to Theorem 1?}
 plus $1$ secure type, on both input and output side; and we exclude the
-configuration with $2$ secure types as input and output, $-1$); where only 44
+configuration with $2$ secure types as input and output, hence the $-1$); where only 44
 are possible. For the sake of readability, we report 6 examples in the
@@
 
+\ON{why from 50?}
+
 \begin{enumerate}[start=50, label={W\arabic*)}]
@@
 	        generated with the correct
-	        inputs
+	        inputs.
 	\item $\ppi{\beliefRegion}{\factRegion}$ the components
@@
 	        inputs are provided but is subject to input
-	        injections
+	        injections.
 	\item $\dr{\beliefRegion}{\factRegion}$ the component
 		never performs the expected behavior (e.g. physical
-		damage)
+		damage).
 \end{enumerate}
@@
 		$\dr{\beliefRegion}{\factRegion}=\dr{\beliefRegion_3}{\beliefRegion_{1+2}}$,
-		the functional block do not implement the requirements.
+		the functional block does not implement the requirements.
 	\item
-		$\pp{\beliefRegion}{\factRegion}=\pp{\beliefRegion_3}{\beliefRegion_{1+2}}$, some inputs produce incorrect output.
-	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_{1+2}}$, some outputs are not the result of the summation of the inputs but with the correct inputs the function outputs correctly.
+		$\pp{\beliefRegion}{\factRegion}=\pp{\beliefRegion_3}{\beliefRegion_{1+2}}$, the block produces incorrect outputs for some inputs.
+	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_{1+2}}$, some outputs are not the result of the summation of the inputs but with the correct inputs the function outputs correctly.\ON{I didn't understand this}
 	\item
@@
 \subsection{Security and Insecurity of a System}\label{sec:formula}
+
+\ON{Maybe start with "We are ready to state our main hypotheses."}
+
 \begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
-	A system security design (in the $\abftheory$-framework) is given by a
+	A system security design\ON{A secure system design?} (in the $\abftheory$-framework) is given by a
 	precise system specification over the physical and functional
-	architectures that uniquely defines the design built on top of those
+	architectures that uniquely defines the design built\ON{maybe "to be built"?} on top of those
 	requirements.
@@
 	If, given a system specification as a collection of requirements, there
-	exist a non-unique design with respect to those requirements, the
-	number of equivalent designs (that fulfills the requirements)
-	quantitatively defines the magnitude of insecurity of a system design.
+	exist a non-unique design with respect to those requirements\ON{I'd change in: there exist more than one design satisfying these requirements}, the
+	number of equivalent\ON{equivalent in what sense?
+In the sense that they equivalently satisfy the requirements?
+If so, I would change this word with "possible".} designs (that fulfill the requirements)
+	quantitatively defines \ON{quantitatively defines->quantifies} the magnitude of insecurity of a system design\ON{I am not sure about this definition.
+But this maybe because of my lack of experience with security.
+My problem stems from the uses of the word "design".
+In someway, we seem to use "design" as "collection of requirements", but also as something to be built in accordance to such requirements.
+So what is insecure: the collection of requirements, of the design built on top of them?
+Right now, my intuition says that it is the collection of requirements that is insecure, but the use of the word "design" suggests otherwise.
+Maybe we need to discuss this on the science club...}.
 \end{hypothesis}
@@
 is the set of all the requirements of $S$,
-and extend $G$ as $G'=\langle V',E'\rangle$ with $E'= E\cup R$ and $V'=V\cup F$. Let $\pi:
+and extend $G$ as $G'=\langle V',E'\rangle$ with $V'=V\cup F$ and $E'= E\cup R$. Let $\pi:
 E'\rightarrow RCC$ (where $RCC$ is the set of relations in the RCC) be the
@@
 algebraic structure (mereotopology) and constraints of the calculus RCC
-(otherwise $\sigma$ returns $0$), we can define
+(otherwise $\sigma$ returns $0$),\ON{I'd add a footnote here, saying: "In other words, $\sigma$ returns $1$ if and only if a configuration is possible with the respect to the axioms of the RCC." This might be redunant, but I think it miht help a bit.} we define:
 \begin{displaymath}
@@
 \end{displaymath}
-where $\pi_{eq}\in\Pi$ is the output of the function $\pi$ that associates only EQ relations
+Here, $\pi_{eq}\in\Pi$ is the output of the function $\pi$ that associates only EQ relations
 to any $e\in E'$, and $I\in\mathbb{N}$ represents all the insecurity configurations
@@
 \begin{itemize}
-	\item a \emph{deployment diagram} describes the \emph{physical
+	\item A \emph{deployment diagram} describes the \emph{physical
 		architecture}. Each agent is defined as an UML node with (physical)
-		ports, and agent' ports are connected via UML information flow
+		ports, and agent's ports are connected via UML information flow
 		connectors, representing the physical channel.
-	\item a \emph{functional architecture} is linked to each agent in the
+	\item A \emph{functional architecture} is linked to each agent in the
 		deployment diagram and is defined by an \emph{object diagram}.
@@
 		blocks, connected via information flow connectors.
-	\item the connection between the two diagrams is implemented by
+	\item The connection between the two diagrams is implemented by
 		``sockets'', functional blocks connected to a 
@@
 
-\section{Conclusion and Future Works}
-We proposed an hypothesis for a foundational theory on security, arguing that
+\section{Conclusion and Future Work}
+We proposed a hypothesis for a foundational theory on security, arguing that
 cybersecurity-related issues are not linked to the maliciousness of an agent

### 2022-04-24 reports/cimss_2022/main-cut.tex
Oliviero's feedback, pt. 2

@@
 \section{Introduction}\label{sec:intro}
-\paragraph{Context and motivation.}
+\paragraph{Context and motivation.}\ON{General: let us remind ourselves to address (at least discuss between us) the concerns of the reviewers.}
 A \emph{scientific theory} is an explanation of a phenomenon such that the

### 2022-04-26 reports/cimss_2022/main-cut.tex
fix oliviero's feedback

--- 2022-04-24 reports/cimss_2022/main-cut.tex
@@
 With our approach, a list of weaknesses emerges from the mathematical
-formulation of a system in a framework called $\abftheory$-framework\ON{Qui, direi o "framework called ABF" o "framework called the ABF-framework"}, and
-predicts 4 main classes of weaknesses. Those classes of weaknesses\ON{rimuoverei ripetizione di "of weaknesses"} are used to
+formulation of a system in a framework called $\abftheory$ (as in Assertions,
+Beliefs, Facts in \autocite{Santaca2016abf})\footnote{Intuitively, as we will
+see later, assertions correspond to an exchange of
+information between agents, beliefs to internal
+information considered true by the agent, and facts to
+requirements.} predicts 4 main classes of weaknesses. Those classes are used to
 calculate all the insecurity configurations of all the components of a system,
@@
 \begin{enumerate}
-	\item \emph{System Specification}: the \emph{functional and physical requirements} \ON{proporrei di togliere l'and dal corsivo} are defined.
-	\item \emph{Architecture Design}: the specification is structured into \emph{functional and physical architectures}\ON{same as the above}.
+  \item \emph{System Specification}: the \emph{functional} and \emph{physical requirements} are defined.
+  \item \emph{Architecture Design}: the specification is structured into \emph{functional} and \emph{physical architectures}.
 	\item \emph{Cybersecurity Risk Assessment}: potential \emph{weaknesses} (errors) and
@@
 
-As in \autocite{Santaca2016abf}, we define agents\ON{so far, we spoke about systems, from now we focus on agents.
-Shall we briefly motivate why we think of systems in term of agents?
-Or, at least, say that we use the two terms intercambiabilmente?
-Later on in the paper, we do say "agent (i.e.
-system)", but that comes much after this, and so the text inbetween might be puzzling to some.}\ON{Further note: In section 4.1 we explicitly state this.
-Perhaps shall we here just say "we consider a CPS as a hierarchy of agents, as we will detail later in Section 4.1"?
-(or something similar)} as a meronomy (an
-hierarchy of Part-Whole relations) over the constituents that we previously defined
-(assertions, beliefs, and facts), based on a standard definition of mereology,
-i.e. based on the definition of parthood relation between \emph{parts}.  Due to
+Following \autocite{Santaca2016abf}, we define a system as a hierarchy of agents. Furthermore, we model an agent as a meronomy (an
+hierarchy of Part-Whole relations) over the aforementioned constituents 
+(assertions, beliefs, and facts), based on a standard definition of mereology
+(i.e. based on the definition of parthood relation between \emph{parts}).  Due to
 the necessity of considering different relations between parts (as we will show afterwards)
@@
 \begin{definition}{\bf System State --}\label{def:opsystem}
-	A CPS system (or a sub-system) state is defined as the\ON{the->a} tuple
+	A CPS system (or a sub-system) state is defined as a tuple
 	$s=\langle\rcc(\factRegion,\beliefRegion),\,\rcc(\factRegion,\assertionRegion),\,\rcc(\beliefRegion,\assertionRegion)\rangle$,
 	where $\assertionRegion$,$\beliefRegion$, and $\factRegion$ are regions of
-	assertions, beliefs (i.e. the beliefs generated by the behavior)\ON{is "generated" the right word here? Anyhow, the bracketed part is not super clear to me.}, and
+	assertions, beliefs (i.e. the beliefs generated by the behavior), and
 	facts respectively, expressed as requirements.
@@
 
-\ON{Shouldn't we introduce these terms here? A, B, F? I mean even with short sentences: assertions are the exchanged data, facts requirements, beliefs are the behaviour...}
-
 As in \autocite{Santaca2016abf}, it follows that, by defining
-a system with \ON{with->as?} a fixed number of regions, there exists
+a system as a fixed number of regions, there exists
 an upper-bound to the number of possible configuration of a system, defined by
 the possible relations between the different regions.
-For completeness, we report in Appendix~\ref{app:agcalc}\ON{here}
+For completeness, we report in Appendix~\ref{app:agcalc}
 some details on the calculation done by the authors in \autocite{Santaca2016abf}.
@@
 of the agent while the other configurations are either impossible to 
-implement or diverge from the intended nominal behavior.\ON{Here, I would explicitly say that what we mean is that, only when the assertiosn, beliefs and facts are equal, then the system is behaving as excepted.
-That is, I would explicitly say is that the only nominal state is EQ,EQ,EQ.
-I mean, it is in the table, but a bit implicit to me.} We note
+implement or diverge from the intended nominal behavior. We note 
 that the numbers reported here do not consider the details of the
@@
 While a quantitative analysis reveals how many possible configurations of an
-agent (i.e. a system) exist w.r.t. the $\abftheory$-framework (i.e.\ON{e.g.?} $54/125$ in RCC5), a
+agent (i.e. a system) exist w.r.t. the $\abftheory$-framework (e.g., $54/125$ in RCC5), a
 qualitative analysis of the different configurations describe the
@@
 to determine the soundness of the design with respect to the specification.
-We first analyze the relations between each pair of
-regions (i.e.  $\abf$), and then categorize different
-agents as tuples of the three regions.\ON{but we don't actually do this, right?
-Maybe I missed it.
-If not, we should remove this sentence.
-If yes, then we should clarify this.
-(At least to me :-))}  For the sake of simplicity, soundness is
-opposed to non-soundness in the following, however, with the RCC, one should\ON{should->could?}
+We now analyze the relations between each pair of
+regions (i.e., $\abf$). For the sake of simplicity, soundness is
+opposed to non-soundness in the following, however, with the RCC, one could
 consider different ``degrees'' of non-soundness.  For example, in RCC5, if we
@@
 By definition, assertions are defined as transfer of information
-between two agents.\ON{first time we say this!} An agent has two main categories of assertions,
+between two agents. An agent has two main categories of assertions,
 input and output assertions.  Given an agent $a$ and a collection of asserted
@@
 are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. With a slight abuse
-of notation,\ON{alternative: in the text we drop the $\Phi$ when the content of the assertions is not relevant.} in the text we simply write $\assertionRegion_{s\rightarrow a}$ when we abstract
-away from the content of the assertion (i.e. $\Phi$).
+of notation, in the text we drop the $\Phi$ when the content of the assertions is not relevant.
 We shall consider
@@
 		handled in the functional specification (beliefs) the
-		specification is complete.\ON{Do we perhaps mean "sound"? We used sound above}
+		specification is complete.
 
@@
 $\Rcc{\rassert{s}{a}}{\beliefRegion_a}$ for input-port or
-$\Rcc{\rassert{r}{*}}{\beliefRegion_a}$, for output-port (where $*$ is a wildcard for an agent)\ON{should this $r$ be an $a$?}.  A port is, in
+$\Rcc{\rassert{a}{r}}{\beliefRegion_a}$, for output-port.  A port is, in
 fact, syntactic sugar to express the relation between assertions and beliefs.
@@
 list all possible cybersecurity weaknesses of a port. From our definition of ports, the following holds.
-\begin{theorem}[Port Weaknesses]
+\begin{theorem}[Port Weaknesses]\label{thm:pw}
 There exist only the following \emph{six}
@@
 	We distinguish between regions that are related to information and regions that
-	don't by writing the latter as $\emptyset$ (e.g. a region corresponding to
-	an assertion is written $A\neq\emptyset$, while a region corresponding to an empty assertion
-	as $A=\emptyset$).\ON{Do we use the weight/$|A|$ thing in the proof? I don't see it... If not, should we get rid of it?}
+  are not (i.e., regions $A$ such that $|A|=0$) by writing the latter as $\emptyset$.
+
   \begin{enumerate}
@@
 	A mono-directional channel between two agents ($s \rightarrow r$) is
-	defined as\ON{remove defined as} an agent whose behavior is (dogmatically) defined as: to
+	an agent whose behavior is defined as: to
 	forward any assertion received from $s$ over an input-port, to  the
@@
 weaknesses of the input and output port; therefore there exist $(7^2)-1=48$
-theoretical configurations ($7^2$ because there are $6$ insecure types of port,\ON{refer to Theorem 1?}
-plus $1$ secure type, on both input and output side; and we exclude the
+theoretical configurations ($7^2$ because there are $6$ insecure types of port -- see Theorem \ref{thm:pw} -- plus $1$ secure type, on both input and output side; and we exclude the
 configuration with $2$ secure types as input and output, hence the $-1$); where only 44
@@
 
-\section{Number of different configurations of a system.}\label{agcalc}
+\section{Number of different configurations of a system.}\label{app:agcalc}
 The general formula to calculate the number of different types of agents is

### 2022-04-26 reports/cimss_2022/main-cut.tex
final draft

@@
 
-\author{}
+\author{Michele Ambrosi, Francesco Beltramini, Federico De Meo,
+ Oliviero Nardi, Mattia Pacchin, and Marco Rocchetto}
 
@@
 
-\institute{}
+\institute{V-Research, Verona, Italy\\email: knowledgezero@v-research.it}
 \maketitle
@@
 Following \autocite{Santaca2016abf}, we define a system as a hierarchy of agents. Furthermore, we model an agent as a meronomy (an
-hierarchy of Part-Whole relations) over the aforementioned constituents 
+hierarchy of part-whole relations) over the aforementioned constituents 
 (assertions, beliefs, and facts), based on a standard definition of mereology
@@
 \begin{definition}{\bf System State --}\label{def:opsystem}
-	A CPS system (or a sub-system) state is defined as a tuple
+  A Cyber-Physical System (CPS), or a sub-system, state is defined as a tuple
 	$s=\langle\rcc(\factRegion,\beliefRegion),\,\rcc(\factRegion,\assertionRegion),\,\rcc(\beliefRegion,\assertionRegion)\rangle$,
@@
 the possible relations between the different regions.
-For completeness, we report in Appendix~\ref{app:agcalc}
-some details on the calculation done by the authors in \autocite{Santaca2016abf}.
 The general formula to calculate the number of different types of agents is
@@
 \autocite{improvingRCC}) combinations are topologically correct. 
-Generalizing to all the RCCs:
-
-\begin{itemize}%[nosep]
-\item \emph{RCC3} --- theoretical: $3^3=27$,  correct: 15.
-\item \emph{RCC5} --- theoretical: $5^3=125$, correct: 54.
-\end{itemize}
-
-%Hence, even if considering a different number of regions than the three
-%$\assertionRegion$, $\beliefRegion$, and $\factRegion$ exponentially affects
-%the number of theoretical agents, the application of RCC downscales that number
-%of a factor that ranges from 1.8 to 2.5. In addition, using RCC5 we consider
-%3.6 times more (different) types of agents than RCC3, but using RCC8 would
-%allow us to consider 3.5 times more different agents.
+For RCC5 there are $5^3=125$ theoretical combinations but
+only 54 are correct with respect to the axioms.
+
 In the quantitative evaluation of a single agent, as in Fig.~\ref{fig:quantitative},
@@
 \subsection{From Multi-Agent to Cyber-Physical Systems}
-As in Fig.~\ref{fig:mashyp}, we relate MAS and CPS as follows.
+As in Fig.~\ref{fig:mashyp}, we relate MAS (Multi-Agent System) and CPS as follows.
 \begin{itemize}
@@
 configuration with $2$ secure types as input and output, hence the $-1$); where only 44
-are possible. For the sake of readability, we report 6 examples in the
-following but the proof by exhaustion over all the possible cases is
-straightforward.
-
+are possible. 
 \begin{enumerate}[start=6, label={W\arabic*)}]
@@
 \end{enumerate}
+
+For the sake of readability, we reported 6 examples
+but the proof by exhaustion (up to W49) over all the possible cases is
+straightforward.
 
@@
 
-\ON{why from 50?}
-
 \begin{enumerate}[start=50, label={W\arabic*)}]
@@
 		$\pp{\beliefRegion}{\factRegion}=\pp{\beliefRegion_3}{\beliefRegion_{1+2}}$, the block produces incorrect outputs for some inputs.
-	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_{1+2}}$, some outputs are not the result of the summation of the inputs but with the correct inputs the function outputs correctly.\ON{I didn't understand this}
+  \item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_{1+2}}$, not all outputs result from a summation of two inputs (but with the expected inputs the function outputs correctly.)
 	\item
@@
 	DR & Drops all inputs and inserts new data & The component never performs/carries the expected behavior/information\\[.1cm]
-	PP & Selectively drops inputs & Part of the expected outputs are not generated in response to the correct inputs \\[.1cm]
+	PP & Selectively drops inputs & part of the expected outputs are not generated in response to the correct inputs \\[.1cm]
 	PPi& Forwards all the inputs but crafts and inserts new malicious data & The components correctly performs/carries the expected behavior/information when the correct inputs are provided but is subject to input injections \\[.1cm]
@@
 \subsection{Security and Insecurity of a System}\label{sec:formula}
-
-\ON{Maybe start with "We are ready to state our main hypotheses."}
+We are ready to state our main hypotheses.
 
 \begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
-	A system security design\ON{A secure system design?} (in the $\abftheory$-framework) is given by a
+	A system security design (in the $\abftheory$-framework) is given by a
 	precise system specification over the physical and functional
-	architectures that uniquely defines the design built\ON{maybe "to be built"?} on top of those
+	architectures that uniquely defines the design to be built on top of those
 	requirements.
@@
 	If, given a system specification as a collection of requirements, there
-	exist a non-unique design with respect to those requirements\ON{I'd change in: there exist more than one design satisfying these requirements}, the
-	number of equivalent\ON{equivalent in what sense?
-In the sense that they equivalently satisfy the requirements?
-If so, I would change this word with "possible".} designs (that fulfill the requirements)
-	quantitatively defines \ON{quantitatively defines->quantifies} the magnitude of insecurity of a system design\ON{I am not sure about this definition.
-But this maybe because of my lack of experience with security.
-My problem stems from the uses of the word "design".
-In someway, we seem to use "design" as "collection of requirements", but also as something to be built in accordance to such requirements.
-So what is insecure: the collection of requirements, of the design built on top of them?
-Right now, my intuition says that it is the collection of requirements that is insecure, but the use of the word "design" suggests otherwise.
-Maybe we need to discuss this on the science club...}.
+	exist a non-unique design with respect to those requirements, the
+	number of possible designs that fulfill the requirements
+	quantitatively defines the magnitude of insecurity of a system design with respect to the specification.
 \end{hypothesis}
@@
 algebraic structure (mereotopology) and constraints of the calculus RCC
-(otherwise $\sigma$ returns $0$),\ON{I'd add a footnote here, saying: "In other words, $\sigma$ returns $1$ if and only if a configuration is possible with the respect to the axioms of the RCC." This might be redunant, but I think it miht help a bit.} we define:
+(otherwise $\sigma$ returns $0$),\footnote{In other words, $\sigma$ returns $1$
+if and only if a configuration is satisfiable with the respect to the axioms of
+the RCC.} we define:
 \begin{displaymath}
@@
 
-\section{Number of different configurations of a system.}\label{app:agcalc}
-The general formula to calculate the number of different types of agents is
-$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
-between $n$ different regions. Therefore, $r^{\binom{n}{k}}$
-expresses the number of permutation of $r$
-relations over ${\binom{n}{k}}$ elements with repetitions, 
-with ${\binom{n}{k}}$ being the number of
-$k$-ary combinations of $n$ regions.
-In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
-($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
-considered in the RCC are binary.  Hence, using RCC5 (with five different
-spatial relations) over three regions, we can theoretically define up to 125
-different type of agents. However, only 54 of the 125 (as showed in
-\autocite{improvingRCC}) combinations are topologically correct with respect to
-the definition of the relations of RCC5. Generalizing to all the RCCs:
-
-\begin{itemize}%[nosep]
-\item \emph{RCC3} --- theoretical: $3^3=27$,  correct: 15.
-\item \emph{RCC5} --- theoretical: $5^3=125$, correct: 54.
-\end{itemize}
-
 \section{Overview of the Results of the Tool}\label{app:results}

### 2022-04-26 reports/cimss_2022/main-cut.tex
last fixes

@@
 
-\authorrunning{}
+\authorrunning{Ambrosi et al.}
 
@@
 \section{Introduction}\label{sec:intro}
-\paragraph{Context and motivation.}
 A \emph{scientific theory} is an explanation of a phenomenon such that the
@@
 any measure is necessary for security are empirically unfalsifiable''. 
-
-\paragraph{Contributions.} 
-The goal of this paper is to lay the foundations of a scientific cybersecurity
+The goal of this paper is to address this issue and to lay the foundations of a scientific cybersecurity
 theory.  We consider the problem raised by Herley not confined to ``computer
@@
 To test our hypothesis we implemented a tool-chain (open-source with
-AGPLv3 license, available at \autocite{v-research2020cybersecurity-anonymous}) for the
+AGPLv3 license, available at \autocite{v-research2020cybersecurity}) for the
 identification of weaknesses and the calculation of potential insecure
@@
 
+\section{Overview of the Results of the Tool}\label{app:results}
+In Fig.~\ref{fig:results} we show a screenshot of the 
+results reported by our tool.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{results.pdf}
+	\caption{Partial View of the results in the spreadsheet file}
+	\label{fig:results}
+\end{figure}
+
 \begin{sidewaysfigure*}
@@
 
-\section{Overview of the Results of the Tool}\label{app:results}
-In Fig.~\ref{fig:results} we show a screenshot of the 
-results reported by our tool.
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{results.pdf}
-	\caption{Partial View of the results in the spreadsheet file}
-	\label{fig:results}
-\end{figure}
-
 \end{document}

### 2022-04-29 reports/cimss_2022/compile.sh
removed old main

@@ -1,32 +1,0 @@
-#!/bin/bash
-latexOutputDir="latex-output-dir"
-latexInputFile="main.tex"
-pdfOutputDir="${latexOutputDir}/${latexInputFile%.*}.pdf"
-
-if [ "$1" != "-short" ] && [ "$1" != "-s" ]; then
-	echo "remove"
-	rm -rf $latexOutputDir
-fi
-
-if [[ ! -d $latexOutputDir ]]; then 
-	mkdir $latexOutputDir
-fi
-
-if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
-then 
-	if [ "$1" != "-short" ] && [ "$1" != "-s" ] 
-	then 
-		if biber --nodieonerror --input-directory=$latexOutputDir --output-directory=$latexOutputDir ${latexInputFile%.*}  
-		then
-			if pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile
-			then 
-				pdflatex -halt-on-error -output-directory=$latexOutputDir $latexInputFile 
-			fi
-		fi
-	fi
-
-	if [[ -f $pdfOutputDir ]]
-	then
-		mv $pdfOutputDir `dirname $0`
-	fi
-fi

### 2022-11-24 reports/cimss_2022/main-cut.tex

@@
 $\agentuniverse$ (see in Tab.~\ref{tab:rcc358}).  We use the Region Connection
-Calculus (RCC), as defined in \autocite{bennettLogics,improvingRCC}, to provide
+Calculus (RCC), as defined in \autocite{bennettLogics,Grutter2008rcc}, to provide
 an axiomatization of the mereo-topological concepts. In its broader definition,
@@
 different types of agents but only 54 of the 125 (as showed in
-\autocite{improvingRCC}) combinations are topologically correct. 
+\autocite{Grutter2008rcc}) combinations are topologically correct. 
 For RCC5 there are $5^3=125$ theoretical combinations but
